### 2018-12-20 draft_p2.tex
- *TODO* Introduction - I Description of the test case    + experimental setup    + LBM    + Drag force    + basic pheno. of drag flucts.    + The drag as a random process - Direct sampling of extreme fluctuations   + Extracting extreme fluctuations from a timeseries   + Instant. drag     ++ Contribution of the foreboy pressure     ++ Dynamical aspects     ++ *TODO* conclusion   + *TODO* AVG drag - *TODO* GKTL - *TODO* Conclusion

% \include{header_p2.tex}

\documentclass[pre,aps,floatfix,10pt,superscriptaddress]{revtex4-1}
\pdfoutput=1
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage[svgnames]{xcolor}
\usepackage{graphicx}
\usepackage{acro}
\usepackage{subfig}
\usepackage{caption}
\graphicspath{{../figures/}}

\captionsetup[figure]{justification=raggedright}

\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}

\DeclareAcronym{pdf}{
  short=PDF,
  long=Probability Density Function,
}
\DeclareAcronym{iid}{
  short=i.i.d.,
  long=Independent Identically Distributed
}
\DeclareAcronym{ou}{
  short=OU,
  long=Ornstein-Uhlenbeck,
}
\DeclareAcronym{gktl}{
  short=GKTL,
  long=Giardina-Kurchan-Tailleur-Lecomte,
}
\DeclareAcronym{ams}{
  short=AMS,
  long=Adaptive Multilevel Splitting,
}
\DeclareAcronym{tams}{
  short=TAMS,
  long=Trajectory Adaptive Multilevel Splitting,
}
\DeclareAcronym{scgf}{
  short=SCGF,
  long=Scaled Cumulant Generating Function,
}
\DeclareAcronym{lbm}{
  short=LBM,
  long=Lattice Boltzmann Method,
}
\DeclareAcronym{lbe}{
  short=LBE,
  long=Lattice Boltzmann Equation,
}
\DeclareAcronym{lgca}{
  short=LGCA,
  long=Lattice Gas Cellular Automata,
}
\DeclareAcronym{lbgk}{
  short=LBGK,
  long=Lattice Bhatnagar-Gross-Krook
}
\DeclareAcronym{OU}{
  short=OU,
  long=Ornstein--Ulhenbeck
}
\DeclareAcronym{dns}{
  short=DNS,
  long=Direct Numerical Simulation,
}
\DeclareAcronym{md}{
  short=MD,
  long=Molecular Dynamics,
}
\DeclareAcronym{cfd}{
  short=CFD,
  long=Computational Fluid Dynamics,
}

\begin{document}
\title{A statistical mechanics approach to the simulation of extreme fluctuations of the time-averaged drag acting on an obstacle immersed in a turbulent flow}

\author{Thibault Lestang}
\email{thibault.lestang@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\author{Freddy Bouchet}
\email{freddy.bouchet@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\author{Emmanuel Lévêque}
\email{emmanuel.leveque@ec-lyon.fr}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\maketitle
\section{Introduction}

\section{Description of the test case for extreme drag fluctuations sampling}
\label{sec:test_flow}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
  \caption{Typical vorticity field in the computational domain. The square obstacle is represented by the gray area in the center of the channel.}
\end{figure}
%% Description de la géométrie
We consider the interaction of a two dimensional turbulent channel flow with a square cylinder, illustrated in figure~\ref{fig:computational_domain}.
At the inlet of the channel, a constant parabolic velocity profile is imposed.
We note $u_0$ the imposed fluid velocity at the inlet, at the center of the channel.
Shortly after the inlet, the flow interacts with a grid, thus leading to turbulence.
Generated eddies are advected downstream by the mean-flow and grow in size along the way---a consequence of the inverse energy cascade taking place in two-dimensional turbulent flows.
A square obstacle is mounted in the channel.
Its distance from the grid is such that the typical size of the eddies it interacts with is close to the siye of the obstacle itself.
Fluid structures are then advected further downstream where they are damped in a region of increasing viscosity, a so-called \textit{sponge layer}~\cite{ref1,ref2}.
The outlet is modelled as an open boundary, where a zero momentum flux along the mean-flow direction in imposed.

%% Méthode Boltzmann sur réseau
The flow dynamics are computed by means of the \ac{lbm}~\cite{ref1,ref2}.
[[Paragraphe LBM]]

%% Descritpion d'une fluctuations de trainée typique
The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts exterted on the obstacle.
In particular, the \textit{drag force} is defined as the result of both pressure and viscous efforts, projected along the mean-flow direction, that is
\begin{equation}
  \label{eq:drag_definition}
  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ the identiy matrix, $\boldsymbol{\tau}$ the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.
to the formation of boundary layers at the top and bottom boundaries of the obstacle.
Figure~\ref{fig:typical_drag_signal} illustrates the time evoluction of the drag acting on the square obstacle.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
  \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}. The correlation time $\tau_c$ is defined further in this text and correspond to the typical fluctuation timescale of the drag.}
  \label{fig:typical_drag_signal}
\end{figure}
The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients fields in the vicinity of the obstacle.
Furthermore, we observed that the contribution of viscous stresses to the overall drag is three order of magnitudes smaller than the contrbution of pressure.
This a consequence of the blunt shape of the ostacle, that results in an important pressure difference between the forebody (the upstream boundary) and the base (the downstream boundary).
As a result, the drag acting on the square cylinder is very well approximated by
\begin{equation}
  \label{eq:drag_approx}
  f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
where $p_{fb}(t)$ amd $p_{base}(t)$ denote the pressure integrated over the forebody and the base, respectively.

The typical timescale of variation of the drag is therefore corresponds to the typical timescale of variation of the pressure in the vicinity of the obstale.
It can be estimated by the \textit{turnover time}
\begin{equation}
  \label{eq:turnover_time}
  \tau_0 = \frac{R}{U}
\end{equation}
where $R$ is the diamenter of the cylinder and $U$ the mean-flow velocity.

%% The phenomenology of typical drag fluctuations

Figure~\ref{fig:typical_drag_fluctuation} illustrate the evolution of the vorticity field in the vicinity of the obstacle.
Over a few turnover times, boundary layers form along the top and bottom boundaries.
The generated vorticity turns into eddies that are advected downstream by the mean flow.
As the result, the vorticity generated along the top and bottom boundary layers have a very small impact
on the the pressure field in the vicinity of the base of the obstacle.
As result, typical fluctuations of the drag are mostly due to the fluctating pressure in the vicinity of the upstream side of the obstacle.
Such fluctuating pressure directly orginates from the incoming turbulent flow.
In the following section~\ref{sec:direct_sampling}, we illustrate very different dynamics for \textit{extreme} fluctuations of the drag acting on the obstacle.
In particular, we show that extreme fluctuations primarily result from fluctuations of the base pressure, whereas forebody pressure fluctuations play very little role.

\subsection{The drag random process: Probability Density Function, time correlations}
\label{sec:pdfs}

\begin{figure}
  \centering
    \subfloat[\ac{pdf} for the drag fluctuations $f'_d$]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{../figures/pdf_drag_SQ_asymetry_expo}}
    \subfloat[Autocorrelation function of the the instantaneous drag $f_d$]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{../figures/correlation_with_log}}  
  \caption{\textbf{(a)} \ac{pdf} describing the statistics of the instantaneous drag fluctuations $f'_d = (f_d - \bar{f_d})/\sigma$. This \ac{pdf} has been estimated on the basis of a timeseries spanning $T_{tot} = 10^6 \tau_c$. $\bar{f_d}$ and $\sigma$ denote the average and the standard deviation computed over the whole timeseries, respectively. The linear fits highlight the exponential behaviour of the the \ac{pdf} for rare events. \textbf{(b)} Autocorrelation function for the instantaneous drag $f_d$, defined as $(\mathbb{E}[f_d(t+\tau)f_d(t)]-\mathbb{E}[f_d]^2)/\sigma^2$ and estimated from a timeseries spanning $T_{tot} = 10^6 \tau_c$. This figure shows the exponential decay of the correlations over time.}
  \label{fig:pdf_drag}
\end{figure}

In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of a very long timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ with $T_{tot} \approx 4\times 10^{6}\tau_0$.
Two comments are in order:
\begin{itemize}
\item The probabilty is skewed towards positive fluctuations. As an indication, the skewness coefficient is ??.
\item The tails of the distribution are well described by an exponential \ac{pdf}, \textit{i.e.} $\mathbb{P}(f_d=f) \propto e^{-\alpha f}\mathrm{d}f$.
\end{itemize}
The interaction of the flow is responsible for this particular shape of the \ac{pdf}.
This can be seen by considering the flow \textit{without} obstacle, and computing the \ac{pdf} of the drag acting on a control surface corresponding to the surface of the obstacle.
This experiment is depicted in figure~\ref{fig:no_square}.
One can see in figure~\ref{fig:pdf_drag} that the \ac{pdf} of the drag process exhibits a very different shape when no obstacle interacts with the flow.
In particular, it is symmetric and does not display exponential tails.
The exponential behaviour of the \ac{pdf} describing rare events plays an important role in the description of the phenomenology of extreme fluctuations of the \textit{time-averaged} drag.
This topic is further discussed in section~\ref{sec:avg_drag}.

Figure~\ref{fig:correlation_function} displays the autocorrelation function for the drag process, defined as 
\begin{equation}
  \label{eq:correlation_def}
  C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
\end{equation}
Time correlations decay exponentially.
As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.
From figure~\ref{fig:correlation_function}, we find that $\tau_c \approx 5\tau_0$.
From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.

\section{Direct sampling of extreme fluctuations of the drag acting on a square cylinder in two-dimensional turbulence}
\label{sec:direct_sampling}

\paragraph*{Paragraphe d'introduction: les motivations d'un étude directe ?}
Rare event algorithms relies on the simulation of an ensemble of trajectories.
Members of the ensemble can be duplicated or removed from the ensemble, based on the evaluation of a \textit{score function}.
A score function is an obervable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achived by the trajectory.
In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observalbe over the trajectory.
However, for high-dimensional dynamics with a very complex phase space, the simple distance to the target event is likely to be a very poor criterion for selecting rare trajectories.
\textbf{[...]}

\subsection{Extracting extreme drag fluctuations from a very long timeseries}
\label{sec:extreme_extraction}

We performed the simulation of the flow dynmaics described in section~\ref{sec:test_flow} over $T_{tot}=10^6\tau_c$.
The current value of the drag $f_d(t)$ was recorded at each time step, thus yielding a timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
As a post-processing step, we could identify the particular drag fluctuations with an amplitude greater than a fixed threshold amplitude $a$.
The threshold $a$ was then chosen so that the typical timescale of occurence of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
This typical timescale of occurence is referred to as the \textit{return time} and can be computed as a function of the fluctuation amplitude on the basis of the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$~\cite{lestang2018}.
Figure~\ref{return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard devaition of the drag process.
We denote by $(t^{\star}, f_d^{\star})$ an fluctation of amplitude $f_d^{\star}$ occuring at time $t^{\star}$.
On the basis of the drag timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we extracted the set of extreme fluctations
\begin{equation}
  \{(t^{\star}, f_d^{\star}) | f_d^{\star} \geq a\} \quad \mbox{with} \quad a=7.6\sigma.
\end{equation}
As a result, we extracted $104$ independent extreme drag fluctuations with a return time greater than $10^4\tau_c$.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{return_time_drag}
  \caption{Drag fluctuation amplitude as a function the correponding return time. The return time $r(a)$ is the typical timescale of occurence of a fluctuation of amplitude larger than $a$.}
  \label{fig:return_time_instant}
\end{figure}
\subsection{Instantaneous drag}
\label{sec:instantaneous_drag}

\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
\label{sec:forebody_and_base_contribution}
It was illustrated in section~\ref{sec:test_flow} that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations that are caused by the updtream turbulent flow.
In the following we investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuatons, in the case of the $104$ extreme drag events ${(t^{\star}, f_d^{\star})_n}_{1\leq n \leq 104}$ extracted from the timeseries.
For a given event , we consider the effective fluctuation, \textit{i.e.} the difference from the average value $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
it can be further decomposed into 
\begin{equation}
  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
\end{equation}
where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denotes the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.  
\begin{figure}
  \centering
  \label{fig:contribution_forebody_and_base_pressure}
\end{figure}
It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the drag fluctuation.
By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinderare dominated by the variation of the pressure in the vicinity of the base of the obstacle.
In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation relatively to the foreboy pressure fluctuation.

\subsubsection{Dynamical apsects of extreme drag fluctations}
\label{sec:dynamical_aspects}

\begin{figure}
  \centering
  \begin{minipage}{0.85\linewidth}
    \begin{tabular}{c|c}
    \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
    \hline
    \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
    \end{tabular}
  \end{minipage}
\begin{minipage}[h]{0.1\linewidth}
  \centering
  \includegraphics[width=\linewidth]{scale_vorticity.png}
\end{minipage}
\caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries. The corresponding drag timeseries can be found in figure~\ref{fig:extremes_timeseries}.
  The high value for the drag results from the formation of a strong negative (red) vortex inducing a pressure drop at the base of the obstacle. The formation of such a structure is aided by important vorticity production at the bottom boundary of the obstacle coupled with the influence of positive vorticity in the vicinity of the base.}
\end{figure}

% \begin{figure}
%   \centering
%   \begin{minipage}[h]{0.85\linewidth}
%     \begin{tabular}[h]{c|c}
%     \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_streamlines_a}\includegraphics[width=0.4\linewidth]{event_95_streamlines.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_streamlines.png}} \\
%     \hline
%     \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_streamlines.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_streamlines.png}} 
%   \end{tabular}
% \end{minipage}
% \begin{minipage}[h]{0.1\linewidth}
%   \centering
%   \includegraphics[width=\linewidth]{scale_density.png}
% \end{minipage}
% \caption{\label{fig:top_4_events_streamlines} Density field at $t=t^{\star}$ for the four highest drag fluctuations in the control simulation for \tf{2}. Recall that the density $\rho$ is proportional to the pressure $p$, following the ideal gas state equation $p=c_s^2\rho$. See chapter~\ref{chap:test_flows} and appendix~\ref{app:lbm}. Additionally, velocity streamlines are depicted, representing advection by the flow at $t=t^{\star}$, see note~\ref{note:streamlines}. Blue areas indicate areas of lower pressure while red regions indicate regions of higher pressure. The formation of a vortex very close to the base boundary leads to a strong pressure drop. The downstream vortices originating from the top boundary layer separation are clearly visible and constrain the formation of the base vortex to a region very close to the base boundary.}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.1\linewidth]{scale_vorticity.png}
%   \begin{tabular}[h]{cccc}
%     \subfloat[$t = t^{\star}-0.54\tau_c$]{\label{frame_a}\includegraphics[width=.2\linewidth]{animation_event_95_0035}} & \subfloat[$t = t^{\star}-0.5\tau_c$]{\label{frame_b}\includegraphics[width=.2\linewidth]{animation_event_95_0036}} &     \subfloat[$t = t^{\star}-0.46\tau_c$]{\label{frame_c}\includegraphics[width=.2\linewidth]{animation_event_95_0037}} &     \subfloat[$t = t^{\star}-0.44\tau_c$]{\label{frame_d}\includegraphics[width=.2\linewidth]{animation_event_95_0038}} \\
%     \subfloat[$t = t^{\star}-0.4\tau_c$]{\label{frame_e}\includegraphics[width=.2\linewidth]{animation_event_95_0039}} &     \subfloat[$t = t^{\star}-0.36\tau_c$]{\label{frame_f}\includegraphics[width=.2\linewidth]{animation_event_95_0040}} &     \subfloat[$t = t^{\star}-0.34\tau_c$]{\label{frame_g}\includegraphics[width=.2\linewidth]{animation_event_95_0041}} &     \subfloat[$t = t^{\star}-0.3\tau_c$]{\label{frame_h}\includegraphics[width=.2\linewidth]{animation_event_95_0042}} \\
%     \subfloat[$t = t^{\star}-0.26\tau_c$]{\label{frame_i}\includegraphics[width=.2\linewidth]{animation_event_95_0043}} &     \subfloat[$t = t^{\star}-0.24\tau_c$]{\label{frame_j}\includegraphics[width=.2\linewidth]{animation_event_95_0044}} &     \subfloat[$t = t^{\star}-0.2\tau_c$]{\label{frame_k}\includegraphics[width=.2\linewidth]{animation_event_95_0045}} &     \subfloat[$t = t^{\star}-0.16\tau_c$]{\label{frame_l}\includegraphics[width=.2\linewidth]{animation_event_95_0046}} \\
%     \subfloat[$t = t^{\star}-0.12\tau_c$]{\label{frame_m}\includegraphics[width=.2\linewidth]{animation_event_95_0047}} &     \subfloat[$t = t^{\star}-0.08\tau_c$]{\label{frame_n}\includegraphics[width=.2\linewidth]{animation_event_95_0048}} &     \subfloat[$t = t^{\star}-0.04\tau_c$]{\label{frame_o}\includegraphics[width=.2\linewidth]{animation_event_95_0049}} &     \subfloat[$t = t^{\star}$]{\label{frame_p}\includegraphics[width=.2\linewidth]{animation_event_95_0050}} \\
%   \end{tabular}
%   \caption{\label{fig:type_1_event}Flow dynamics corresponding to a particular type 1 event, which drag timeseries and flow fields at $t=t^{\star}$ are displayed in figures~\ref{fig:extremes_timeseries}, \ref{fig:top_4_events_vorticity} and~\ref{fig:top_4_events_streamlines}, respectively. In each frame the vorticity field is displayed, as well as the velocity streamlines, representing advection by the flow at each instant. See note~\ref{note:streamlines} for a discussion of streamlines.
%     The maximum of the drag is attained for $t=t^{\star}$ and corresponds to frame~\ref{frame_p}.  The sequence starts with a shear boundary layer forming over the top boundary of the obstacle (frames~\ref{frame_a} to~\ref{frame_d}). This boundary layer separates (frames~\ref{frame_e} to~\ref{frame_j}) and results in the formation of a large positive eddy in the vicinity of the base of the obstacle (frames~\ref{frame_k} and~\ref{frame_l}). In the meantime, frames~\ref{frame_j} to~\ref{frame_m} depict the formation of an attached strong shear layer at the bottom boundary. Eventually, frames~\ref{frame_n} to~\ref{frame_p} illustrate that this results in vorticity production very close to the base, aided by the large positive eddy originating from the top boundary layer separation.}
% \end{figure}

We now describe the dynamics associated with some of the extreme drag fluctuations sampled in the contrl timeseries.
Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events with the larger fluctuation in the whole timeseries.
Remarkably, all four events display very similar flow configurations.
In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels can reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
Such vorticity is produced through viscous shear along a boundary of the obstacle tangential to the mean flow direction.
By contrast with typical configurations, we observe that this vorticity is not directly advected by the mean flow, but instead constrained in the vicinity of the base of the obstacle.
Indeed, in each of the four examples displayed in figure~\ref{fig:top_four_events}, a large vortical region is located at roughly one obstacle length of the base.
As illustrated by the streamlines displayed in figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.

Figure~\ref{fig:dynamics} illustrates the dynamics leading to the extreme drag fluctuation corresponding to figure~\ref{fig:top_four_events_a}.
Figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} show that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy, advected by the flow above the obstacle.
The interaction of this eddy with the top boundary layer eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
This can be seen see in figure~\ref{fig:dynamics_k}.
Although this vorticity is too far from the base to directly affect the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle, as can be observed in figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.

Figure~\ref{fig:top_four_events_timeseries} shows the drag timeseries corresponding to the four events featured in figure~\ref{fig:top_four_events}.
The timeseries are centered on the extreme event.
It illustrates that these fluctuations are short-lived, as their temporel extension is rouhly one correlation time unit.
This is the time it takes for the block eddy to form and be advected away by the mean flow, thus frreeing the vorticity concentrated in the vicinity of the base of the obstacle.

Among the $104$ sampled events, $80\%$ can be related to very similar dynamics as the one described in the previous paragraph. In order to identify such events, we monitored the averaged shear along the top or bottom booundary:
\begin{equation}
  \label{eq:avg_shear_def}
  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x}
\end{equation}
where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary, depending on the event.
Figure~\ref{fig:type_1_and_type_2_a} displays the label of each sampled events, sorted according to the amplitude of the corresponding fluctuation.
It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the hghest fluctuation amplitude.
Figure~\ref{fig:type_1_and_type_2_a} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of these events.
For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
As illustrated in figure~\ref{fig:extremes_timeseries}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
Correspondingly, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
Interestingly, these excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
This confirms that the increase of $\overline{\gamma}$ acts as a precursor for extreme drag fluctuations.

The remaining $20\%$ of the sampled events correspond to slightly different dynmaics.
In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
This viscous shear is induced by a large vortex detached from the obstacle.

\subsubsection{conclusion}
On the basis

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-12-23 draft_p2.tex
Ajoute la section sur la trainée intégée

@@
 
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress]{revtex4-1}
+\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage]{revtex4-1}
 \pdfoutput=1
@@
 [[Paragraphe LBM]]
+
+\subsection{The drag force}
+\label{sec:drag_force}
 
@@
 
-\subsubsection{conclusion}
-On the basis
+\subsection{Extreme fluctuations of the time-averaged drag }
+\label{sec:time_avg}
+
+% Motivation
+In the previous section, we observed that rare events corresponding to extremely high values of the drag acting on the square obstacle have a lifetime of about $2\tau_c$.
+In many applications, this duration is much smaller than the timescale of interest.
+This is for instance the case when considering the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
+
+In such cases, a relevant observable is the \textit{time-averaged} drag
+\begin{equation}
+  \label{eq:def_time_averaged_drag}
+  F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
+\end{equation}
+where $f_d$ denotes the instantaneous drag and $T\leq \tau_c$ a timescale of interest.
+In the following we consider the case where $T \gg \tau_c$, typically $T=10\tau_c$.
+In this context, a fluctuation of the average $F_T(t)$ is relatex to roughly $T / 2\tau_c$ independant fluctuations of $f_d$ over the time interval $[t;t+T]$.
+What is the phenomenology leading to extreme values of $F_T(t)$ ?
+Do an exceptionally large value of the average drag result from a single excpetionally large value of the instantaneous drag (case (1)) ?
+Or from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
+From the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$ we deduced the timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ and identified extreme positive fluctuations for which the drag exceeds a fixed threshold $a$ in the same way as described in section~\ref{sec:extreme_extraction}.
+Setting $a=5.2\sigma _T$ leads to $84$ independant events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.
+
+Figure~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t^{\star})\}_{t^{\star} \leq t \leq t^{\star}+T}$ for four of the highest fluctuations in the timeseries.
+It illustrates that the phenomenology of the extreme fluctuations of the time-averaged drag cannot be reduced to the two above scenarios. 
+Indeed, both case (1) and case (2) are featured in figure~\ref{fig:extreme_avg}, as well as intermediate cases where the very large value of the drag results from both a very large fluctuation and a large number of positive, rather typical, fluctuations.
+
+This marginal phenomenology can be connected to the form of the tail of the \ac{pdf} describing extreme positive drag fluctations, illustrated in figure~\ref{fig:pdf_drag}.
+Let $X$ be a random variable whose \ac{pdf} is denoted $\mathbb{P}$ and standard deviation $\sigma_X$.
+Considering an extreme positive value $a$ of $S_N=\sum_{n=1}{N}X_n$, the probability $p_1$(resp. $p_2$) of case (1) (resp. case (2)) writes:   
+\begin{equation}
+  \label{eq:indep}
+  p_{1}(\sum_{1}^{N} X_n=a) \approx \mathcal{P}\left(\frac{a}{N}\right)^{N} \,\,\, \text{and} \,\,\, p_{2}(\sum_{1}^{N} X_n=a)\approx \mathcal{P}(a)  
+\end{equation}
+If $\mathbb{P}$ has an exponential posititve tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
+\begin{equation}
+  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\sim} C\left(e^{-\alpha \frac{a}{N}}\right)^{N} e^{-\alpha a} = 1.
+  \label{eq:ratio_exp}
+\end{equation}
+
+Past this duration, the blocking eddy responsible for the accumulation of vortity in the base of the obstacle is advected downstream.
+The high local vorticity is turn advected and the base pressure goes back to typical values.
 

### 2018-12-24 draft_p2.tex
Intro sur les algorithmes

@@
 
-Past this duration, the blocking eddy responsible for the accumulation of vortity in the base of the obstacle is advected downstream.
-The high local vorticity is turn advected and the base pressure goes back to typical values.
+\section{Rare event algorithms}
+\label{sec:rare_events_algorithms}
+
+In the limit of very rare events, direct sampling approaches become unrealistic. As an example, consider the drag process $t \to f_d(t)$.
+In the limit of extreme positive fluctuations, the return time of a fluctuations $f_d \geq a$ verifies
+\begin{equation}
+  \label{eq:return_time_scaling}
+  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
+\end{equation}
+where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
+The computational cost required to sample events $f_d \geq a$ therefore scales like $e^{\alpha a}$.
+
+In this section we discuss the application of rare event algorithms to the numerical sampling of extreme drag forces on immersed objects, on the example
+of the flow presented in section~\ref{sec:test_flow}.
+The purpose of these algorithm is to sample rare events for a computational cost well inferior to their return time.
+Many of them rely on the simulation of an ensemble of trajectories in parallel, which can be duplicated or discarded from the ensemble depending on their relevance to
+the realisation of extreme events of interest. 
+Numerous algorithms and applications have been proposed since the original ideas of Kahn and Harris in the early 50s~\cite{kahnharris}, which can be sorted in two categories.
+On the one hand, \textit{splitting} approaches decompose rare events into a succession of less rare events that can be sampled with a higher probability.
+On the other hand, \textit{importance sampling} relies on the sampling of trajectories according to a modified distribution, that is biased towards the rare events of interest.
+In section~\ref{sec:ams}, we highlight that sampling  extreme flucutations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
+In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme flucutations of the time-averaged drag.
+As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
+
+\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
+\label{sec:ams}
+
+\subsection{Extreme time-averaged drag forces with the \acl{gktl} algorithm}
+\label{sec:gktl}
 

### 2018-12-26 draft_p2.tex
Remanie un peu l'intro de la section sur les algos. Ajoute les figures dans la sous-section AMS Début de commentaire de la figure_AMS (principe d'une itération)

@@
 The purpose of these algorithm is to sample rare events for a computational cost well inferior to their return time.
-Many of them rely on the simulation of an ensemble of trajectories in parallel, which can be duplicated or discarded from the ensemble depending on their relevance to
+Many of them involve keeping track of an ensemble of $N$ trajectories $\{x_n(t)\}_{n}$.
+Over the iterations of the algorithm, trajectories are duplicated or discarded from the ensemble according to precise selection rules that bias the sampling towards the rare events of interest.
+Such selection rules are designed in such a way that the sampling bias is known at each iteration of the algorithm.
+In other words, generated trajectories are given a weight which knowledge allow for the computation of their probability and expectation values according to the unbiased stationary distribution of the dynamics.
+
+depending on their relevance to
 the realisation of extreme events of interest. 
@@
 \label{sec:ams}
+The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event can be made easier by splitting the event
+into a sequence of events that can be sampled with a higher probabilbity.
+Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date.
+In recent years, it was applied to several problems
+
+In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
+The algorithm is sketched in figure~\ref{figure_AMS}.
+For further details about the algorithm and assoiated mathematical results, see~\cite{lestang2018} and references therein.
+In the remainder of this section, we illustrate that the application of the \ac{TAMS}, fails to easily and efficiently sample rare drag fluctuations occuring in the test flow described in section~\ref{sec:test_flow}.
+We analyse those difficulties in light of the conclusions of sections~\ref{sec:instantaneous_drag} and~\ref{sec:time_avg} on the dynamics of extreme drags.
+
+To begin with, figure~\ref{fig:comparaison_temps_retour} illustrate expected results from the \ac{tams}, based on a one-dimensional \ac{ou} process.
+In this case, it is clear that the algorithm allows for the sampling of very rare fluctuations for a computational cost that is many order of magnitude lower than their return time.
+
+We then apply the \ac{tams} to the test flow, to sample extreme drag fluctuations on the square obstacle.
+
+  \centering
+  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
+  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
+    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
+The amplitude of fluctuations for the \ac{ou} process~\eqref{eq:OUprocess_chap_ams} are represented along the Y-axis.
+Along the X-axis is represented the typical computational cost required to sample fluctuations of the corresponding amplitude in both the \ac{tams} and direct sampling.
+In the latter, the typical computational cost is simply the return time $r(a)$.
+The computational cost associated to a level $a$ is the \ac{tams} is denoted by $C(a)$.
+The orange line represents the evolution of the level $\mathcal{Q}_j^{\star}$ as a function of the cost $C(a)$.
+The blue stars represent the maximum of $x(t)$ over the resampled trajectory \emph{after} it is resampled, that is at iteration $j+1$.
+Finally, the solid black line is the analytical solution for the return time of amplitude $a$~\cite{lestang_computing_2018}.}
+  \label{fig:comparaison_temps_de_retour}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{ams/instant_256_20}
+    \caption{Maximum of the instantaneous drag along the resampled trajectories, denoted by $\tilde{f}_d^{(j)}$, as a function of the corresponding computational cost $C_j$, defined in~\eqref{eq:comp_cost_iterations}.
+In both figures, the \ac{tams} is iterated until extinction.
+    The solid black line indicates the typical fluctuation amplitude $a$ as a function of the corresponding  \textit{return time} $r(a)$.
+It is the typical timescale of occurrence of fluctuations above $a$.
+Furthermore, the square markers depict the maximum $\tilde{f}_{d,n}^{(0)},\,\, 1\leq n \leq N_c$ over the initial trajectories.
+    This figure illustrates that in both experiments the \ac{tams} does not improve the sampling of rare fluctuations.
+On can see that the maximum values among the resampled trajectories concentrates on values already sampled over the initial trajectories.
+Eventually, extinction is reached as the algorithm is unable to generate a fluctuation higher than the maximum fluctuation in the initial ensemble of trajectories.
+    This behaviour is illustrated in figure~\ref{fig:extinction_32_5} on the basis of the drag timeseries involved in the experiment of figure~\ref{fig:comparaison_ams_return_time_instant_a}.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{ams/iter_181}
+    \caption{Visualisation of the drag timeseries corresponding to the set of trajectories in the ensemble, at different iterations along the algorithm.
+In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
+Trajectories have a duration $T_a = 5\tau_c$ and their number is $N_c = 32$.
+    One can see that, as the number of iterations increases, the number of independent resampled trajectories drastically decreases until extinction at iteration $181$.
+The timeseries corresponding to the two initial trajectories exhibiting the highest fluctuations are displayed in colour.
+In all four figures, colour is inherited from the parent trajectory.
+It illustrates that the two strongest initial trajectories quickly become predominant in the ensemble, as the other are discarded.
+    As can be seen in figure~\ref{fig:subfigc}, trajectories only separate well after the maximum is reached.
+Therefore, resampling the blue trajectories based on the red trajectories can only change the overall maximum by a very small amount.
+Consequently, blue trajectories are discarded one after the other, each time resampled into another instance of the red trajectory.
+This new instance has roughly the same maximum as the others.}
+\end{figure}
 

### 2018-12-26 draft_p2.tex
Presque fini la sous section sur l'AMS. Reste à écrire un pragraphe de perspectives

@@
 \section{Introduction}
-
+The purpose of this paper is twofold.
 \section{Description of the test case for extreme drag fluctuations sampling}
@@
 \label{sec:ams}
-The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event can be made easier by splitting the event
-into a sequence of events that can be sampled with a higher probabilbity.
+% Introduction to AMS
+The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the event into a sequence of events that can be sampled with a higher probabilbity.
 Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date.
-In recent years, it was applied to several problems
+In recent years, it was applied to several problems, ranging from the computation of the dissociation time of biomolecules~\cite{teo2016} to the simulation of rare relaminarisations in a stochastic model of wall turbulence~\cite{rolland2017}.
 
+% Quick description of TAMS
 In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
+The \ac{ams}, and therefore the \ac{tams}, relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t))$.
+It is an observable of the dynamics that acts as a measure of the distance of the current state $\mathbf{x}(t)$ to the region of the state space that correspond to the rare event(s) of interest.
+The \ac{ams} algorithm is an iterative procedure, that starts from an ensemble of $N$ independant trajectories of the dynamics over a time-interval $[0;T_a]$.
+At each iteration, trajectories with the lowest maxima of the score function over $[0;T_a]$ are discarded and replaced by trajectories that achieved a higher value of the score function at some point of their
+history,
 The algorithm is sketched in figure~\ref{figure_AMS}.
 For further details about the algorithm and assoiated mathematical results, see~\cite{lestang2018} and references therein.
-In the remainder of this section, we illustrate that the application of the \ac{TAMS}, fails to easily and efficiently sample rare drag fluctuations occuring in the test flow described in section~\ref{sec:test_flow}.
-We analyse those difficulties in light of the conclusions of sections~\ref{sec:instantaneous_drag} and~\ref{sec:time_avg} on the dynamics of extreme drags.
 
-To begin with, figure~\ref{fig:comparaison_temps_retour} illustrate expected results from the \ac{tams}, based on a one-dimensional \ac{ou} process.
-In this case, it is clear that the algorithm allows for the sampling of very rare fluctuations for a computational cost that is many order of magnitude lower than their return time.
+% Example: OU process
+We first describe the result of the \ac{tams} on a simple case.
+We consider a one-dimensional \ac{ou} process
+\begin{equation}
+  \label{eq:ou}
+  \dot{x} = -x + \sqrt{2\epsilon}\eta (t)  
+\end{equation}
 
-We then apply the \ac{tams} to the test flow, to sample extreme drag fluctuations on the square obstacle.
+and apply the \ac{tams} to simulate trajectories $\{\mathbf{x}(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
+Figure~\ref{fig:compraison_temps_de_retour} illustrate the amplitude of the maximum over trajectories along the iterations of the algorithm.
+The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
+Figure~\ref{fig:compraison_temps_de_retour} illustrates that the \ac{tams} samples extreme fluctuations $x \geq a$ for a computational cost several order of magnitude lower than the return time of such fluctuations.
+Undoubtedly, an \ac{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} and \ac{tams}.
+Indeed, the state space is one-dimensional and the choice of the score function poses no questions: it is simply the observable $x$ itself.
+In addition, the noise term in~\eqref{eq:ou} has no correlations in time, which means that newly generated trajectories will quickly separate from their ancesotrs.
 
+% Application to turbulent flow past an obstacle
+Can we obtain similar results by applying the \ac{tams} to the test flow described in\ref{sec:test_flow} ?n
+The objective is to sample trajectories that display extreme fluctuatons of the drag $f_d$ acting on the square obstacle.
+However, by contrast with a simple \ac{ou} process, the phase space in higly-dimensinal and the choice of the score function $xi (\mathbf{x}(t))$ is not so clear.
+In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
+For such complex dynamics however, there is no resons for the drag itself to be a good measure of how far is a given state in phase space to regions where the drag is extreme.
+Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, whcih is precisely what we are tryping to achieve.
+
+% Commentaire figure instant_256_20
+
+Figure~\ref{fig:instant_256_20} displays similar information as figure~\ref{fig:comparaison_temps_de_retour} for $N=256$ initial trajectories over $T_a = 20 \tau_c$.
+Additionally, figure~\ref{fig:instant_256_20} shows the distribuion of the maximum of the drag among the initial trajectories.
+Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
+These maxima are depicted by the stars in figure~\ref{fig:instant_256_20}.
+However, the maximum over the resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
+
+% Commentaire de la figure iter_181
+
+Figure~\ref{fig:instant_256_20} can be better understood by visualizing the set of the last resampled trajectories. 
+One can see that, ultimately, trajectories have been resampled from the same initial trajectory.
+We explain this phenonemon as follows.
+Because the flow dynamics hae a memory, it takes a certain time before a resampled trajectory separate from its parent.
+This is clearly visible on figure~\ref{fig:instant_256_20}.
+In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting point will overlap over a timescale called
+the <<Lyapunov timesacle>>~\cite{lyapunovRef}.
+However, we observe that the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag proces.
+This means that the resampling of a trajectory based on the value of the drag at $t^{\star}$ cannot lead to larger drag values for $t^{\star} \leq t \leq \tau_L$.
+For $t > \tau_L$, the drag process has no memory of the drag fluctuations on which the resampling was based and the probability of observing an extreme fluctuation over $\tau_l \leq t \leq T_a$ is very low.
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_0$ is linked to the convective nature of the flow.
+Indeed, fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$.
+
+% perspectives
+
+\begin{itemize}
+\item more trajectories
+\item different score functino?
+\item Change algorithm
+\end{itemize}
+\begin{figure}
   \centering

### 2018-12-27 draft_p2.tex
Commencé la section sur GKTL Intro; description de l'algo; choix des parametes; reste description des resultats et perspectives

@@
 
+We now turn to the sampling of extreme fluctuations of the time-averaged drag $F_T$.
+The \ac{ams} could be used in the same way as described in section~\ref{sec:ams} for the instantaneous drag, choosing $F_T$ itself as a score function.
+However, we observed a similar behaviour: the \ac{tams} resampling procedure fails to yield trajectories with higher fluctuations of $F_T$.
+
+When it comes to time-averaged observable, one can make use of a different algorithm the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
+It was originally designed to compute large deviations of time-averaged dynamical observables.
+By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not belong to the family of splitting algorithms.
+Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
+The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
+The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.
+
+\subsubsection{The algorithm}
+\label{sec:gktl_description}
+Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a population of trajectories.
+In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
+The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is tilted in order to favour trajectories with large values of a chosen time averaged observable.
+This sort of algorithm has first been proposed by~\cite{Giardina2006} and has been used to study rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
+
+More precisely, we perform simulations of an ensemble of $N$ trajectories $\left\{ X_{n}(t)\right\}$ (with $n=1,2,...,N$) starting from random initial conditions.
+Like in section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
+We consider an observable of interest $A(X(t))$ and a resampling time $\tau$.
+At times $t_{i}=i\tau$ (with $i=1,2,...,T_{a}/\tau$) we assign to each trajectory $n$ a weight $W_{n}^{i}$ defined as
+\begin{equation}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}
+\end{equation}
+For each trajectory $X_{n}$, a random number of copies of the trajectory are generated, on average proportional to the weight $W_{n}^{i}$ and such that the total number of trajectories produced at each event is equal to $N$.
+For deterministic systems, a small perturbation is introduced for clones at times $t_i$, so that clones separate from their parent. 
+The parameter $k$ is chosen by the user in order to control the strength of the selection and thus to target a class of extreme events of interest.
+The larger the value of $k$, the more trajectories with large values of the time average observable will survive the selection.
+
+Let us denote formally $\mathbb{P}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right)$ the probability to observe a trajectory $\left\{ x(t)\right\} _{0\leq t\leq T_{a}}$ in the model, and $\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}} \right)$ the probability to observe the same trajectory with the algorithm.
+By construction of the algorithm through the weights~\eqref{eq:Weight}, we have
+\begin{align}
+\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(x(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right).\label{eq:Biased_Path_Approximation}
+\end{align}
+where the normalisation factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(X(t))dt}\right]$, denoting by $\mathbb{E}_{0}$ the expectation value with respect to $\mathbb{P}_{0}$, and $\underset{N\rightarrow\infty}{\sim}$ means that this is true only asymptotically for large $N$.
+The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
+Equation~\eqref{eq:Biased_Path_Approximation} is obtained by assuming the mean field approximation
+\begin{equation}
+R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{0}^{t_{_{1}}}A(X_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{0}^{t_{1}}A(X(t))dt}\right],\label{eq:Mean_Field_Approximation}
+\end{equation}
+which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{Giardina2006,Giardina2011}:
+\begin{equation}
+\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(X(t))dt}\right].\label{eq:Estimate_Lambda}
+\end{equation}
+The validity of the mean field approximation and the fact that the typical relative error due to this approximation is of order $1/\sqrt{N}$ has been proven~\cite{DelMoralBook,DelMoral2013} to be true for a family of rare event algorithms including the one adopted in this paper.
+
+Formula~\eqref{eq:Biased_Path_Approximation} is valid only for times $T_{a}$ that are integer multiples of the resampling time $\tau$.
+The killed trajectories have to be discarded from the statistics.
+Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
+In this way one obtains an effective ensemble of $N$ trajectories from time 0 to time $T_{a}$, distributed according to $\mathbb{P}_{k}$.
+All trajectories reconstructed in this way are real solutions of the model: we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
+
+The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$, from the distribution $\mathbb{P}_{k}$.
+This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
+For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\ref{these_lestang,nemoto_bouchet2016}
+As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
+An estimator is then given by
+\begin{equation}
+\mathbb{E}_{0}\left[O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ X_{n}(t)\right\} _{0\leq t\leq T_{a}}\right)\mbox{e}^{-k\int_{0}^{T_{a}}A(X_{n}(t))dt}\mbox{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
+\end{equation}
+where the $X_{n}$ are the $N$ backward reconstructed trajectories.
+Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
+As a result, the effective ensemble of trajectories generated by the \ac{gktl} algorithm contains a larger amount of such events, and empirical estimators such as~\eqref{eq:GK_O_estimator} have a dramatically lower statistical error.
+
+\subsubsection{Application to extreme fluctuations of the time-averaged drag}
+\label{sec:gktl_drag}
+
+We now discuss the application the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
+The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
+Consistenly with section~\ref{sec:time_avg}, we choose the averaging window $T=10\tau_c$.
+
+% Choice of parameters and perturbation
+The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C = N_c \times T_a$.
+In the following series of experiements we set $T_a = T = 10\tau_c$.
+The sampling from the \ac{gktl} alorithm will be compared to the sampling from a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
+To match this computational cost with the \ac{gktl} experiments, we set $N_c=16384$.
+Lastly, one must set the cloning period $\tau$.
+On the one hand, too small a cloning period may result in unfortunate loss of information as clones do not separate from their parent over a cloning period.
+On the other hand, choosing $\tau \gg \tau_c$ leads to too few cloning stages.
+As a consequence, we set $\tau = \tau_c /2$.
+
+% Commentaire de la figure illustration_IS
+
+Figure~\ref{fig:illustration_IS} illustrates the importance sampling performed by the \ac{gktl} algorithm.
+illustrates the amplitude of the fluctuations of $F_{T_a}$ corresponding to the sampled trajectories for each three experiments.
+More precisely, it illustrates the estimate of the biased drag \ac{pdf} $\rho_k$ in each case:
+\begin{multline}
+  \label{eq:estimate_biased_measure}
+  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)] \\
+  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F).
+\end{multline}
+Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the unbiased drag \ac{pdf} $\rho_0(F) = \mathbb{E}_{\set{P}_0}[\delta(F_{T_a} - F)]$.
+Note that this figure is analogous to figure~\ref{fig:illstration_importance_sampling} on page \pageref{fig:illstration_importance_sampling}.
+One can see that, as $k$ is increased, the biased \ac{pdf} is shifted towards extreme values.
+However, the experiment with $k=0.03$ does not seem to increase the sampling of rare event with respect to the experiment with $k=0.025$.
+Actually, the shape of the estimate of the biased \ac{pdf} shows that the reconstructed trajectories concentrate around a single value of the fluctuation $F'_{T_a} = F_{T_a}-\overline{F_{T_a}} \approx 5\sigma_{T_a}$, where $\overline{F_{T_a}}$ and $\sigma_{T_a}$ denote the average and standard deviation of $F_{T_a}$, respectively.
+This suggests a strong overlap of trajectories in the reconstructed ensemble, which we address next.
+
+\paragraph{}
+Figure~\ref{fig:nbIndep} illustrates the overlap of trajectories in the reconstructed ensemble.
+Along the X-axis, we represent the $M = T_a / \tau$ iterations of the evolution/selection procedure.
+For each iteration $m$ we plot the number of independent trajectories, \textit{i.e.} the number of trajectories that exhibit different paths $\{X(t)\}_{(m-1)\tau \leq t \leq m\tau}$ over the interval $t\in [(m-1)\tau;m\tau]$.
+In practice we trace back the history of the trajectories backward in time starting from the final time $T_a = M\tau$, in a way very similar to the \emph{Method 1} presented in section~\ref{sec:reconstruction_practice}.
+Figure~\ref{fig:nbIndep} illustrates that, at early times, the vast majority of the trajectories in the reconstructed ensemble overlap.
+This is also illustrated, in a different way, by the red trajectories in figure~\ref{fig:schema_algo_b}.
+Say $N$ trajectories of the reconstructed ensemble overlap over the interval $[0;m\tau]$, with $1\leq m \leq M$.
+Among these $N$ trajectories, only one trajectory $j$ actually originates from $t=0$.
+Indeed, the $N-1$ others have been discarded following the selection stage of a given iteration $\tilde{m}\geq m$, and replaced by a clone of trajectory $j$ over the interval $[0;\tilde{m}\tau]$.
+In the following we call \emph{fully independent trajectories} the trajectories that do not overlap at all over $[0;T_a]$, \textit{i.e.} trajectories that do not overlap over the first period $[0;\tau]$.
+From figure~\ref{fig:nbIndep} one can see that the number of fully independent trajectories is very small compared to the total number of trajectories $N_c$.
+
+\paragraph{}
+In this context, can the \ac{gktl} yield more \emph{independent} events than a direct sampling based on a long simulation ?
+If yes, how much more ?
+\begin{table}
+  \centering
+  \subfloat[$k=0.02$]{
+  \begin{tabular}{l|ccccc}
+    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+    \hline
+    0\% & 759 & 258 & 32 & 5 & 0  \\ 
+    25\% & 942 & 372 & 62 & 9 & 0  \\ 
+    50\% & 1594 & 799 & 155 & 22 & 0  \\ 
+    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
+  \end{tabular}}\\
+\subfloat[$k=0.025$]{\label{tbl:nbTraj_tbl_b}
+  \begin{tabular}{l|ccccc}
+    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+    \hline
+    0\% & 305 & 179 & 52 & 5 & 0  \\ 
+    25\% & 451 & 310 & 144 & 39 & 4  \\ 
+    50\% & 1019 & 834 & 521 & 198 & 27  \\ 
+    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
+  \end{tabular}}\\
+\subfloat[$k=0.03$]{\label{tbl:nbTraj_tbl_c}
+  \begin{tabular}{l|ccccc}
+    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+    \hline
+    0\% & 100 & 80 & 33 & 8 & 1  \\ 
+    25\% & 186 & 166 & 110 & 46 & 7  \\ 
+    50\% & 539 & 510 & 391 & 205 & 36  \\ 
+    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
+  \end{tabular}}
+\caption{Number of independent trajectories corresponding to fluctuations $F'_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the reconstructed ensemble, for the three experiments $k=0.02,0.025,0.03$.
+Parameters of the \ac{gktl} experiments are $N_c = 16384$, $T_a=10\tau_c$, $\tau = \tau_c /2$ and $\epsilon = 0.02$.
+In each case we count as independent trajectories with an overlap of $0\%$ (zero overlap), $25\%$ and $50\%$, respectively.
+In addition, $N_{N_c}^{(gauss)}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ realisations obtained from a direct sampling approach.
+This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.
+It is defined in~\eqref{eq:gaussian_number}.}
+\label{tbl:nbTraj_tbl}
+\end{table}
+Table~\ref{tbl:nbTraj_tbl} displays the number of independent trajectories that correspond to fluctuations $F'_{T_a} \geq a$ with $a=\sigma_{T_a}, 2\sigma_{T_a},3\sigma_{T_a}...$, for each of the three experiments.
+In each case, we count the number of trajectories according to the allowed degree of overlap.
+For instance, allowing a $25\%$ overlap, we consider trajectories that overlap over $[0;T_a/4]$ as independent.
+In order to compare with a direct sampling approach, we could perform a very long simulation of computational cost $N_c\times T_a$, divide it into $N_c$ realisations $F_{T_a}$ and count the number of fluctuations $F'_{T_a} \geq a$ above a given threshold value $a$.
+However, we found that the \ac{pdf} describing the statistics of fluctuations $F'_{T_a}$---with $T_a = 10\tau_c$--- is very close to a Gaussian \ac{pdf}.
+As a result, the expected number of fluctuations $F'_{T_a} \geq a$ among $N_c$ realisations is
+\begin{equation}
+  N^{(gauss)}_{N_c} \approx N_c \times \set{P}(F'_{T_a} \geq a) = N_c \times \frac{1}{\sqrt{2\pi \sigma_{T_a}}}\int_{a}^{\infty}e^{-x^2/2\sigma^2_{T_a}}\mathrm{d}x.
+  \label{eq:gaussian_number}
+\end{equation}
+Considering small fluctuations $F'_{T_a} \geq a$ with $a = \sigma$ and $a = 2\sigma$, one can see in table~\ref{tbl:nbTraj_tbl} that the number of sampled independent trajectories actually decreases as $k$ is increased.
+This results from the fact that the overall number of independent trajectories decreases as $k$ is increased, as illustrated in figure~\ref{fig:nbIndep}.
+More importantly, considering fluctuations above $3\sigma$, the \ac{gktl} experiments for $k=0.025$ and $k=0.03$ provide a larger number of fluctuations than a direct sampling approach with similar computational cost.
+For instance, table~\ref{tbl:nbTraj_tbl_b} shows that the experiment with $k=0.025$ provides $5$ fully independent trajectories corresponding to fluctuations $F'_{T_a} \geq 3\sigma$.
+This is ten times more than what can be expected from a direct sampling.
+Notice however how drastic is the reduction of the number of independent trajectories as the allowed degree of overlap is diminished.
+Allowing for a $50\%$ overlap, the reconstructed ensemble for the experiment $k=0.025$ contains roughly $200$ independent trajectories corresponding to fluctuations $F'_{T_a} \geq 3\sigma$.
+Generally speaking, we observe in all three experiments that the number of independent trajectories is very small compared to the total number of trajectories $N_c = 16384$.
+For instance, table~\ref{tbl:nbTraj_tbl_c} shows that, allowing a $50\%$ overlap between independent trajectories, $521$ trajectories can be used to study the dynamics leading to fluctuations $F'_{T_a} \geq 3\sigma$.
+This represents $521 / N_c \approx 3\%$ of the overall computational cost of the experiment.
+We stress that, in spite of this rather poor yield, the \ac{gktl} still results in a larger amount of extreme fluctuations, with respect to a direct sampling approach.
+However, there is clearly room from improvement.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{illustration_IS}
+  \caption{\textbf{(a)}: Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 
+  The \ac{gktl} algorithm is used with $N_c = 16384$ copies and $T_a = T= 10\tau_c$ for three increasing values of the bias $k$.
+Additionally, the cloning period is $\tau = \tau_c/2$, resulting in $M=20$ iterations of the evolution/selection procedure.
+Note that in this particular example the last cloning stage at $t_{19}$ is not performed.
+  As $k$ is increased, the sampled \ac{pdf} is shifted towards extreme values.
+  Along the X-axis we display the fluctuation $F'_{T} = F_{T} - \overline{F_{T}}$, where the average $\overline{F_{T}}$ and standard deviation $\sigma_{T}$ have been computed over a timeseries spanning $10^6\tau_c$.
+  One can see that the distribution for $k=0.03$ is peaked for a particular value of the fluctuation.
+This indicates that a significant part of the sampled trajectories actually overlap over the major part if their history.
+  A a result, the average drag only differ by a small amount.}
+\end{figure}
 \end{document}

### 2018-12-29 draft_p2.tex
Section sur GKTL. Maque conclusion de la section

@@
 The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_0$ is linked to the convective nature of the flow.
-Indeed, fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$.
+Indeed, fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for
+trajectories to separate.
 
@@
 
-Figure~\ref{fig:illustration_IS} illustrates the importance sampling performed by the \ac{gktl} algorithm.
-illustrates the amplitude of the fluctuations of $F_{T_a}$ corresponding to the sampled trajectories for each three experiments.
-More precisely, it illustrates the estimate of the biased drag \ac{pdf} $\rho_k$ in each case:
-\begin{multline}
+We performed three experiments, corresponding to three different values of the parameter $k$.
+This parameter sets the amplitude of the typical fluctuations within the biased ensemble sampled by the algorithm.
+In other words, a higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.
+In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
+In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled flucutations through the Gartner-Ellis theorem~\cite{GEref}.
+In the large time limit $T_a\to \infty$, the typical level of fluctuations $\mathbb{E}_{\mathbb{P}_k}[ F_{T_a} ]$ within the biased ensemble of trajectories
+can be related to the \textit{Scaled Cumulant Generating Function} $\lambda(k)$~\cite{ref}.
+
+Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved with the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
+For each three experiment, it displays the estimate of the biased \ac{pdf}
+\begin{equation}
   \label{eq:estimate_biased_measure}
-  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)] \\
-  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F).
-\end{multline}
-Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the unbiased drag \ac{pdf} $\rho_0(F) = \mathbb{E}_{\set{P}_0}[\delta(F_{T_a} - F)]$.
-Note that this figure is analogous to figure~\ref{fig:illstration_importance_sampling} on page \pageref{fig:illstration_importance_sampling}.
-One can see that, as $k$ is increased, the biased \ac{pdf} is shifted towards extreme values.
-However, the experiment with $k=0.03$ does not seem to increase the sampling of rare event with respect to the experiment with $k=0.025$.
-Actually, the shape of the estimate of the biased \ac{pdf} shows that the reconstructed trajectories concentrate around a single value of the fluctuation $F'_{T_a} = F_{T_a}-\overline{F_{T_a}} \approx 5\sigma_{T_a}$, where $\overline{F_{T_a}}$ and $\sigma_{T_a}$ denote the average and standard deviation of $F_{T_a}$, respectively.
-This suggests a strong overlap of trajectories in the reconstructed ensemble, which we address next.
+  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
+  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
+\end{equation}
+computed over the set of $N=16384$ trajectories sampled by the algorithm.
+Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} describing the time-averaged drag process.
+It corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.
 
-\paragraph{}
-Figure~\ref{fig:nbIndep} illustrates the overlap of trajectories in the reconstructed ensemble.
-Along the X-axis, we represent the $M = T_a / \tau$ iterations of the evolution/selection procedure.
-For each iteration $m$ we plot the number of independent trajectories, \textit{i.e.} the number of trajectories that exhibit different paths $\{X(t)\}_{(m-1)\tau \leq t \leq m\tau}$ over the interval $t\in [(m-1)\tau;m\tau]$.
-In practice we trace back the history of the trajectories backward in time starting from the final time $T_a = M\tau$, in a way very similar to the \emph{Method 1} presented in section~\ref{sec:reconstruction_practice}.
-Figure~\ref{fig:nbIndep} illustrates that, at early times, the vast majority of the trajectories in the reconstructed ensemble overlap.
-This is also illustrated, in a different way, by the red trajectories in figure~\ref{fig:schema_algo_b}.
-Say $N$ trajectories of the reconstructed ensemble overlap over the interval $[0;m\tau]$, with $1\leq m \leq M$.
-Among these $N$ trajectories, only one trajectory $j$ actually originates from $t=0$.
-Indeed, the $N-1$ others have been discarded following the selection stage of a given iteration $\tilde{m}\geq m$, and replaced by a clone of trajectory $j$ over the interval $[0;\tilde{m}\tau]$.
-In the following we call \emph{fully independent trajectories} the trajectories that do not overlap at all over $[0;T_a]$, \textit{i.e.} trajectories that do not overlap over the first period $[0;\tau]$.
-From figure~\ref{fig:nbIndep} one can see that the number of fully independent trajectories is very small compared to the total number of trajectories $N_c$.
+% Limitations
 
-\paragraph{}
-In this context, can the \ac{gktl} yield more \emph{independent} events than a direct sampling based on a long simulation ?
-If yes, how much more ?
+For a fixed number of trajectories $N$, there is a limited range of $k$ values for which the algorithm is able to yield an accurate estimate of $\rho_k$.
+Setting $k$ to a high value causes finite population effects: trajectories overlap over most of their duration.
+This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
+This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, therefore leading to the same value for the averaged drag $F_T$.
+See~\cite{ref1,ref2,ref3} for further details about the scaling of the required population size as $k$ is increased.
+
+% Malgré les limitations, c'est toujours mieux
+% qu'un échantillonnage direct
+Unsurprisingly, sampling larger fluctuations thus reauire a larger number of populations, and therefore a larger computational cost.
+Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
+not be sampled through a direct sampling with equal computational cost.
+Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
+As a comparison, it also displays the average number of such fluctuations in a long timeseries of duration $T_{tot}=10^6\tau_c$, divided in $T_{tot} / T_a$ events.
+This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian, motivated by figure~\ref{fig:pdf_averared}.
+While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
+
 \begin{table}
   \centering
-  \subfloat[$k=0.02$]{
   \begin{tabular}{l|ccccc}
-    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+     & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
     \hline
-    0\% & 759 & 258 & 32 & 5 & 0  \\ 
-    25\% & 942 & 372 & 62 & 9 & 0  \\ 
-    50\% & 1594 & 799 & 155 & 22 & 0  \\ 
-    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
-  \end{tabular}}\\
-\subfloat[$k=0.025$]{\label{tbl:nbTraj_tbl_b}
-  \begin{tabular}{l|ccccc}
-    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-    \hline
-    0\% & 305 & 179 & 52 & 5 & 0  \\ 
-    25\% & 451 & 310 & 144 & 39 & 4  \\ 
-    50\% & 1019 & 834 & 521 & 198 & 27  \\ 
-    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
-  \end{tabular}}\\
-\subfloat[$k=0.03$]{\label{tbl:nbTraj_tbl_c}
-  \begin{tabular}{l|ccccc}
-    Overlap & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-    \hline
-    0\% & 100 & 80 & 33 & 8 & 1  \\ 
-    25\% & 186 & 166 & 110 & 46 & 7  \\ 
-    50\% & 539 & 510 & 391 & 205 & 36  \\ 
-    $N_{N_c}^{(gauss)}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
-  \end{tabular}}
-\caption{Number of independent trajectories corresponding to fluctuations $F'_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the reconstructed ensemble, for the three experiments $k=0.02,0.025,0.03$.
-Parameters of the \ac{gktl} experiments are $N_c = 16384$, $T_a=10\tau_c$, $\tau = \tau_c /2$ and $\epsilon = 0.02$.
-In each case we count as independent trajectories with an overlap of $0\%$ (zero overlap), $25\%$ and $50\%$, respectively.
-In addition, $N_{N_c}^{(gauss)}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ realisations obtained from a direct sampling approach.
-This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.
-It is defined in~\eqref{eq:gaussian_number}.}
+    $k=0.02$ & 1594 & 799 & 155 & 22 & 0  \\ 
+    $k=0.025$ 1019 & 834 & 521 & 198 & 27  \\ 
+    $k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
+    $\mathcal{N}_{direct}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
+  \end{tabular}
+  \caption{Number of independent corresponding to fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$.
+Parameters of the \ac{gktl} experiments are $N_c = 16384$, $T_a=10\tau_c$, $\tau = \tau_c /2$.
+Within the biased ensemble, trajectories are considered distinct if they overlap for less than $50\%$ of their duration.
+    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ independant realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.
+    This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.}
 \label{tbl:nbTraj_tbl}
 \end{table}
-Table~\ref{tbl:nbTraj_tbl} displays the number of independent trajectories that correspond to fluctuations $F'_{T_a} \geq a$ with $a=\sigma_{T_a}, 2\sigma_{T_a},3\sigma_{T_a}...$, for each of the three experiments.
-In each case, we count the number of trajectories according to the allowed degree of overlap.
-For instance, allowing a $25\%$ overlap, we consider trajectories that overlap over $[0;T_a/4]$ as independent.
-In order to compare with a direct sampling approach, we could perform a very long simulation of computational cost $N_c\times T_a$, divide it into $N_c$ realisations $F_{T_a}$ and count the number of fluctuations $F'_{T_a} \geq a$ above a given threshold value $a$.
-However, we found that the \ac{pdf} describing the statistics of fluctuations $F'_{T_a}$---with $T_a = 10\tau_c$--- is very close to a Gaussian \ac{pdf}.
-As a result, the expected number of fluctuations $F'_{T_a} \geq a$ among $N_c$ realisations is
-\begin{equation}
-  N^{(gauss)}_{N_c} \approx N_c \times \set{P}(F'_{T_a} \geq a) = N_c \times \frac{1}{\sqrt{2\pi \sigma_{T_a}}}\int_{a}^{\infty}e^{-x^2/2\sigma^2_{T_a}}\mathrm{d}x.
-  \label{eq:gaussian_number}
-\end{equation}
-Considering small fluctuations $F'_{T_a} \geq a$ with $a = \sigma$ and $a = 2\sigma$, one can see in table~\ref{tbl:nbTraj_tbl} that the number of sampled independent trajectories actually decreases as $k$ is increased.
-This results from the fact that the overall number of independent trajectories decreases as $k$ is increased, as illustrated in figure~\ref{fig:nbIndep}.
-More importantly, considering fluctuations above $3\sigma$, the \ac{gktl} experiments for $k=0.025$ and $k=0.03$ provide a larger number of fluctuations than a direct sampling approach with similar computational cost.
-For instance, table~\ref{tbl:nbTraj_tbl_b} shows that the experiment with $k=0.025$ provides $5$ fully independent trajectories corresponding to fluctuations $F'_{T_a} \geq 3\sigma$.
-This is ten times more than what can be expected from a direct sampling.
-Notice however how drastic is the reduction of the number of independent trajectories as the allowed degree of overlap is diminished.
-Allowing for a $50\%$ overlap, the reconstructed ensemble for the experiment $k=0.025$ contains roughly $200$ independent trajectories corresponding to fluctuations $F'_{T_a} \geq 3\sigma$.
-Generally speaking, we observe in all three experiments that the number of independent trajectories is very small compared to the total number of trajectories $N_c = 16384$.
-For instance, table~\ref{tbl:nbTraj_tbl_c} shows that, allowing a $50\%$ overlap between independent trajectories, $521$ trajectories can be used to study the dynamics leading to fluctuations $F'_{T_a} \geq 3\sigma$.
-This represents $521 / N_c \approx 3\%$ of the overall computational cost of the experiment.
-We stress that, in spite of this rather poor yield, the \ac{gktl} still results in a larger amount of extreme fluctuations, with respect to a direct sampling approach.
-However, there is clearly room from improvement.
 

### 2019-01-07 draft_p2.tex
Ré-itération générale du papier

@@
 We consider the interaction of a two dimensional turbulent channel flow with a square cylinder, illustrated in figure~\ref{fig:computational_domain}.
-At the inlet of the channel, a constant parabolic velocity profile is imposed.
-We note $u_0$ the imposed fluid velocity at the inlet, at the center of the channel.
-Shortly after the inlet, the flow interacts with a grid, thus leading to turbulence.
-Generated eddies are advected downstream by the mean-flow and grow in size along the way---a consequence of the inverse energy cascade taking place in two-dimensional turbulent flows.
-A square obstacle is mounted in the channel.
-Its distance from the grid is such that the typical size of the eddies it interacts with is close to the siye of the obstacle itself.
-Fluid structures are then advected further downstream where they are damped in a region of increasing viscosity, a so-called \textit{sponge layer}~\cite{ref1,ref2}.
-The outlet is modelled as an open boundary, where a zero momentum flux along the mean-flow direction in imposed.
+A constant parabolic velocity profile is imposed at the inlet of the channel and turbulence is
+generated by means of a grid.
+A square obstacle is mounted at the center of the channel.
+Eddies are created in the vicinity of the grid with a typical size that correspond to the step of the grid.
+As they are advected downstream, eddies interact with each other and grow in size, until they interact with the obstacle.
+The distance separating the obstacle from the grid is such that the typical size of the eddies it interacts with is close to the size of the obstacle itself.
+The flow in the vicinity of the outlet is regularized by means of a \textit{sponge layer}~\cite{ref1,ref2}, \textit{i.e.} a region where the viscosity is progressivaly increased.
+Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
 
@@
 %% Descritpion d'une fluctuations de trainée typique
-The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts exterted on the obstacle.
-In particular, the \textit{drag force} is defined as the result of both pressure and viscous efforts, projected along the mean-flow direction, that is
+The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts.
+In particular, the \textit{drag force} is defined as the result of both pressure and viscous constraints, projected along the mean-flow direction.
+It reads
 \begin{equation}
   \label{eq:drag_definition}
-  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x
+  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x,
 \end{equation}
-where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ the identiy matrix, $\boldsymbol{\tau}$ the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.
-to the formation of boundary layers at the top and bottom boundaries of the obstacle.
-Figure~\ref{fig:typical_drag_signal} illustrates the time evoluction of the drag acting on the square obstacle.
+where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ is the identity matrix, $\boldsymbol{\tau}$ denotes the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.
+
+%% Viscous effects are negligible
+The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients in the vicinity of the obstacle.
+Furthermore, because of the blunt shape of the obstacle, the contribution of viscous constraints to the overall drag is negligage with respect to pressure constraints.
+As a result, the drag mainly originates from the pressure difference between the \textit{forebody}---the upstream boundary---and the \textit{base}---the downstream boundary.
+Consequently, the drag acting on the square cylinder is very well approximated by
+\begin{equation}
+  \label{eq:drag_approx}
+  f_d(t) = p_{fb}(t) - p_{b}(t)
+\end{equation}
+where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base, respectively.
+
+%% The phenomenology of typical drag fluctuations
+The typical timescale of variation of the drag therefore corresponds to the timescale of pressure fluctuations in the vicinity of the obstacle.
+An estimate of such timescale is the \textit{turnover time}
+\begin{equation}
+  \label{eq:turnover_time}
+  \tau_0 = \frac{R}{U},
+\end{equation}
+where $R$ is the diameter of the cylinder and $U$ the mean-flow velocity.
+Figure~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle over a few thousands turnover times.
+The evolution over a few turnover times of the vorticity field around the obstacle is displayed in figure~\ref{fig:typical_vorticity}.
+One can see that the interaction of the incoming flow with the square obstacle primarily results in boundary layers along the top and bottom boundaries.
+However, the vorticity corresponding to these boundary layers only have a small impact on the the pressure field in the vicinity of the obstacle, because it is rapidly advected dowmstream by the mean-flow.
+As a consequence, the base pressure varies much less than the forebody pressure, and typical drag fluctuations mostly result from the fluctuating pressure at the front of the obstacle.
+
+In the following section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
+In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a smaller role.
+
 \begin{figure}
@@
 \end{figure}
-The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients fields in the vicinity of the obstacle.
-Furthermore, we observed that the contribution of viscous stresses to the overall drag is three order of magnitudes smaller than the contrbution of pressure.
-This a consequence of the blunt shape of the ostacle, that results in an important pressure difference between the forebody (the upstream boundary) and the base (the downstream boundary).
-As a result, the drag acting on the square cylinder is very well approximated by
-\begin{equation}
-  \label{eq:drag_approx}
-  f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-where $p_{fb}(t)$ amd $p_{base}(t)$ denote the pressure integrated over the forebody and the base, respectively.
-
-The typical timescale of variation of the drag is therefore corresponds to the typical timescale of variation of the pressure in the vicinity of the obstale.
-It can be estimated by the \textit{turnover time}
-\begin{equation}
-  \label{eq:turnover_time}
-  \tau_0 = \frac{R}{U}
-\end{equation}
-where $R$ is the diamenter of the cylinder and $U$ the mean-flow velocity.
-
-%% The phenomenology of typical drag fluctuations
-
-Figure~\ref{fig:typical_drag_fluctuation} illustrate the evolution of the vorticity field in the vicinity of the obstacle.
-Over a few turnover times, boundary layers form along the top and bottom boundaries.
-The generated vorticity turns into eddies that are advected downstream by the mean flow.
-As the result, the vorticity generated along the top and bottom boundary layers have a very small impact
-on the the pressure field in the vicinity of the base of the obstacle.
-As result, typical fluctuations of the drag are mostly due to the fluctating pressure in the vicinity of the upstream side of the obstacle.
-Such fluctuating pressure directly orginates from the incoming turbulent flow.
-In the following section~\ref{sec:direct_sampling}, we illustrate very different dynamics for \textit{extreme} fluctuations of the drag acting on the obstacle.
-In particular, we show that extreme fluctuations primarily result from fluctuations of the base pressure, whereas forebody pressure fluctuations play very little role.
 
@@
 In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
-Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of a very long timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ with $T_{tot} \approx 4\times 10^{6}\tau_0$.
-Two comments are in order:
-\begin{itemize}
-\item The probabilty is skewed towards positive fluctuations. As an indication, the skewness coefficient is ??.
-\item The tails of the distribution are well described by an exponential \ac{pdf}, \textit{i.e.} $\mathbb{P}(f_d=f) \propto e^{-\alpha f}\mathrm{d}f$.
-\end{itemize}
-The interaction of the flow is responsible for this particular shape of the \ac{pdf}.
-This can be seen by considering the flow \textit{without} obstacle, and computing the \ac{pdf} of the drag acting on a control surface corresponding to the surface of the obstacle.
-This experiment is depicted in figure~\ref{fig:no_square}.
-One can see in figure~\ref{fig:pdf_drag} that the \ac{pdf} of the drag process exhibits a very different shape when no obstacle interacts with the flow.
-In particular, it is symmetric and does not display exponential tails.
-The exponential behaviour of the \ac{pdf} describing rare events plays an important role in the description of the phenomenology of extreme fluctuations of the \textit{time-averaged} drag.
-This topic is further discussed in section~\ref{sec:avg_drag}.
+In order to estimate its statistical properties, we performed a simulation of the dynamics over a duration $T_{tot} = 4\times 10^6\tau_c$.
+In the following of this paper we refer to this long simulation as the \textit{control run}.
+Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of the control timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ obtained from the control run.
+One can see that the tails of this distribution, that describe the probability of very rare events, can be very well modelled by an exponential \ac{pdf} of the form $p(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
+In addition, figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} of the process describing the drag acting on a control surface equivalent to the surface of  the obstacle, in a flow \textit{without obstacle}.
+This different setup is illustrated in figure~\ref{fig:no_square}.
+In this case, the corresponding \ac{pdf} estimate for the drag process is quasi-symmetric and, more importantly, does not display exponential tails.
+We conclude that the interation of the flow with the square obstacle is responsible for the exponential tails of the \ac{pdf} of the drag process.
+In section~\ref{sec:time_avg}, we relate this very particular shape of the \ac{pdf} describing extreme drag fluctuations to the phenomenology of extreme fluctuations of the time-average of the drag process.
 
-Figure~\ref{fig:correlation_function} displays the autocorrelation function for the drag process, defined as 
-\begin{equation}
-  \label{eq:correlation_def}
-  C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
-\end{equation}
-Time correlations decay exponentially.
-As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.
-From figure~\ref{fig:correlation_function}, we find that $\tau_c \approx 5\tau_0$.
-From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
-This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
+Lastly, figure~\ref{fig:correlation_function} illustrates that the drag process has exponentially decaying time correlations.
+In the following, we call \textit{correlation time}, the typical time it takes for the drag process to decorrelate.
+We note it $\tau_c$.
+From figure~\ref{fig:correlation_function}, one can see that $\tau_c \approx 4\tau_0$.
+This means that the drag decorrelates after only a few number of eddies interacted with the square obstacle.
+
+% \begin{equation}
+%   \label{eq:correlation_def}
+%   C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
+% \end{equation}
+
+% As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.
+
+% From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
+% This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
 
@@
 \paragraph*{Paragraphe d'introduction: les motivations d'un étude directe ?}
+%% TODO: Introduire la notion de temps de retour
 Rare event algorithms relies on the simulation of an ensemble of trajectories.
 Members of the ensemble can be duplicated or removed from the ensemble, based on the evaluation of a \textit{score function}.
-A score function is an obervable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
-For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achived by the trajectory.
-In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observalbe over the trajectory.
+A score function is an observable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
+For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achieved by the trajectory.
+In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observable over the trajectory.
 However, for high-dimensional dynamics with a very complex phase space, the simple distance to the target event is likely to be a very poor criterion for selecting rare trajectories.
@@
 
-We performed the simulation of the flow dynmaics described in section~\ref{sec:test_flow} over $T_{tot}=10^6\tau_c$.
-The current value of the drag $f_d(t)$ was recorded at each time step, thus yielding a timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-As a post-processing step, we could identify the particular drag fluctuations with an amplitude greater than a fixed threshold amplitude $a$.
-The threshold $a$ was then chosen so that the typical timescale of occurence of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
-This typical timescale of occurence is referred to as the \textit{return time} and can be computed as a function of the fluctuation amplitude on the basis of the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$~\cite{lestang2018}.
+On the basis of the control timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we tracked every drag fluctuation with an amplitude greater than a fixed threshold amplitude $a$.
+The value of this threshold was chosen so that the return time of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
 Figure~\ref{return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
-According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard devaition of the drag process.
-We denote by $(t^{\star}, f_d^{\star})$ an fluctation of amplitude $f_d^{\star}$ occuring at time $t^{\star}$.
-On the basis of the drag timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we extracted the set of extreme fluctations
-\begin{equation}
-  \{(t^{\star}, f_d^{\star}) | f_d^{\star} \geq a\} \quad \mbox{with} \quad a=7.6\sigma.
-\end{equation}
-As a result, we extracted $104$ independent extreme drag fluctuations with a return time greater than $10^4\tau_c$.
+According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard deviation of the drag process.
+We denote by $(t^{\star}, f_d^{\star})$ a fluctuation of peak amplitude $f_d^{\star}$ occurring at time $t^{\star}$, as illustrated in figure~\ref{fig:extraction}.
+From the control timeseries, we identified $104$ independent extreme drag fluctuations whith an amplitude greater than $7.6\sigma$, \textit{i.e.} with a return time greater than $10^4\tau_c$.
+In the following sections~\ref{sec:forebody_and_base_contribution} and~\ref{sec:instantaneous_drag}, we describe the phenomenology of extreme drag fluctuations on the basis of this set of
+events.
 
@@
   \includegraphics[width=\linewidth]{return_time_drag}
-  \caption{Drag fluctuation amplitude as a function the correponding return time. The return time $r(a)$ is the typical timescale of occurence of a fluctuation of amplitude larger than $a$.}
+  \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
   \label{fig:return_time_instant}
@@
 \label{sec:forebody_and_base_contribution}
-It was illustrated in section~\ref{sec:test_flow} that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations that are caused by the updtream turbulent flow.
-In the following we investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuatons, in the case of the $104$ extreme drag events ${(t^{\star}, f_d^{\star})_n}_{1\leq n \leq 104}$ extracted from the timeseries.
-For a given event , we consider the effective fluctuation, \textit{i.e.} the difference from the average value $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
-it can be further decomposed into 
+In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
+We now investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuations, in the case of the sampled \textit{extreme} drag fluctuations.
+Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
+In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
+It can be further decomposed into 
 \begin{equation}
@@
 \end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denotes the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
+where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
 Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
@@
 \end{figure}
-It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the drag fluctuation.
-By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinderare dominated by the variation of the pressure in the vicinity of the base of the obstacle.
-In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation relatively to the foreboy pressure fluctuation.
+It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the overall drag fluctuation.
+By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinder are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
+In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation, relatively to the forebody pressure fluctuation.
 
-\subsubsection{Dynamical apsects of extreme drag fluctations}
+\subsubsection{Dynamical aspects of extreme drag fluctuations}
 \label{sec:dynamical_aspects}
@@
 
-We now describe the dynamics associated with some of the extreme drag fluctuations sampled in the contrl timeseries.
-Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events with the larger fluctuation in the whole timeseries.
-Remarkably, all four events display very similar flow configurations.
-In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels can reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
-Such vorticity is produced through viscous shear along a boundary of the obstacle tangential to the mean flow direction.
+We now turn to a description of the flow dynamics for some of the extreme drag fluctuations, sampled in the control timeseries.
+Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events corresponding to the largest fluctuation amplitudes.
+One can see that they correspond to very similar flow configurations.
+In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
+Such strong vorticity is produced through viscous shear along either the bottom or top boundary of the obstacle.
+
 By contrast with typical configurations, we observe that this vorticity is not directly advected by the mean flow, but instead constrained in the vicinity of the base of the obstacle.
 Indeed, in each of the four examples displayed in figure~\ref{fig:top_four_events}, a large vortical region is located at roughly one obstacle length of the base.
-As illustrated by the streamlines displayed in figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.
+As illustrated by figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.
 
-Figure~\ref{fig:dynamics} illustrates the dynamics leading to the extreme drag fluctuation corresponding to figure~\ref{fig:top_four_events_a}.
-Figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} show that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy, advected by the flow above the obstacle.
-The interaction of this eddy with the top boundary layer eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
+Figure~\ref{fig:top_four_events_timeseries} displays the drag timeseries corresponding to the four events illustrated in figure~\ref{fig:top_four_events}.
+It illustrates that the lifetime of such fluctuations temporal extension is roughly one correlation time unit $\tau_c$.
+It is the time it takes for the block eddy to form and be advected away by the mean flow, thus freeing the vorticity concentrated in the vicinity of the base of the obstacle.
+This can be seen in figure~\ref{fig:dynamics}, which illustrate the vorticty dynamics from $t^{\star}-\tau_c/2$ to $t^{\star}$ on a particular example of an extreme drag fluctuation. 
+On can see in figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy.
+This interaction eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
 This can be seen see in figure~\ref{fig:dynamics_k}.
-Although this vorticity is too far from the base to directly affect the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle, as can be observed in figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.
+Although this vorticity is too far from the base to directly affect the base pressure, and thus the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle---see figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.
 
-Figure~\ref{fig:top_four_events_timeseries} shows the drag timeseries corresponding to the four events featured in figure~\ref{fig:top_four_events}.
-The timeseries are centered on the extreme event.
-It illustrates that these fluctuations are short-lived, as their temporel extension is rouhly one correlation time unit.
-This is the time it takes for the block eddy to form and be advected away by the mean flow, thus frreeing the vorticity concentrated in the vicinity of the base of the obstacle.
-
-Among the $104$ sampled events, $80\%$ can be related to very similar dynamics as the one described in the previous paragraph.
-In order to identify such events, we monitored the averaged shear along the top or bottom booundary:
+We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
+In order to identify such events, we monitored the averaged shear along the top or bottom boundary
 \begin{equation}
   \label{eq:avg_shear_def}
-  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x}
+  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
 \end{equation}
-where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary, depending on the event.
-Figure~\ref{fig:type_1_and_type_2_a} displays the label of each sampled events, sorted according to the amplitude of the corresponding fluctuation.
-It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the hghest fluctuation amplitude.
-Figure~\ref{fig:type_1_and_type_2_a} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of these events.
+where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary.
+% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
+% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
+Figure~\ref{fig:gamma_asof_fd} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
 For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
 As illustrated in figure~\ref{fig:extremes_timeseries}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
-Correspondingly, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-Interestingly, these excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
-This confirms that the increase of $\overline{\gamma}$ acts as a precursor for extreme drag fluctuations.
+As a consequence, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
+These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
+This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
 
-The remaining $20\%$ of the sampled events correspond to slightly different dynmaics.
+The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
 In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
 This viscous shear is induced by a large vortex detached from the obstacle.
+For a more detailed description of such dynamics, see the chapter 3 of ~\ref{ref:these_lestang}.
 
@@
 % Motivation
-In the previous section, we observed that rare events corresponding to extremely high values of the drag acting on the square obstacle have a lifetime of about $2\tau_c$.
-In many applications, this duration is much smaller than the timescale of interest.
-This is for instance the case when considering the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
+In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
+In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $tau_c$.
+However, In many applications, this duration is much smaller than the timescale of interest.
+Consider for insatnce the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
 
+%% Definition de la trainée intégrée
 In such cases, a relevant observable is the \textit{time-averaged} drag
@@
 In the following we consider the case where $T \gg \tau_c$, typically $T=10\tau_c$.
-In this context, a fluctuation of the average $F_T(t)$ is relatex to roughly $T / 2\tau_c$ independant fluctuations of $f_d$ over the time interval $[t;t+T]$.
+In this context, a fluctuation of the average $F_T(t)$ is related to roughly $T / \tau_c$ independent fluctuations of the \textit{instantaneous} drag $f_d$, over the time interval $[t;t+T]$.
+
+%% Quelle est la phénoménologie de la trainée intégrée ?
 What is the phenomenology leading to extreme values of $F_T(t)$ ?
-Do an exceptionally large value of the average drag result from a single excpetionally large value of the instantaneous drag (case (1)) ?
-Or from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
-From the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$ we deduced the timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ and identified extreme positive fluctuations for which the drag exceeds a fixed threshold $a$ in the same way as described in section~\ref{sec:extreme_extraction}.
-Setting $a=5.2\sigma _T$ leads to $84$ independant events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.
+Do an exceptionally large value of the average drag result from a single exceptionally large value of the instantaneous drag (case (1)) ?
+Or instead from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
+From the control timeseries, see section~\ref{sec:pdfs}, one can compute a control timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$.
+It is then possible to identify extreme positive fluctuations for which the time-averaged $F_T$ drag exceeds a fixed threshold $a$, in the very same way as described in section~\ref{sec:extreme_extraction}.
+Setting $a=5.2\sigma _T$ leads to $84$ independent events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.
 
-Figure~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t^{\star})\}_{t^{\star} \leq t \leq t^{\star}+T}$ for four of the highest fluctuations in the timeseries.
-It illustrates that the phenomenology of the extreme fluctuations of the time-averaged drag cannot be reduced to the two above scenarios. 
-Indeed, both case (1) and case (2) are featured in figure~\ref{fig:extreme_avg}, as well as intermediate cases where the very large value of the drag results from both a very large fluctuation and a large number of positive, rather typical, fluctuations.
+Figure~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t^{\star})\}_{t^{\star} \leq t \leq t^{\star}+T}$ for four of the sampled extreme fluctuations sampled in the control timeseries for $F_T$.
+It illustrates that the phenomenology of the extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
+Indeed, both cases are featured in figure~\ref{fig:extreme_avg}, along with intermediate cases where the very large value of the drag results from both a very large fluctuation and a large number of typical, however all positive, fluctuations of the instantaneous drag $f_d$.
 
-This marginal phenomenology can be connected to the form of the tail of the \ac{pdf} describing extreme positive drag fluctations, illustrated in figure~\ref{fig:pdf_drag}.
+This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{papier_JSTAT}.
+An estimate of this \ac{pdf} is displayed in figure~\ref{fig:pdf_drag}.
 Let $X$ be a random variable whose \ac{pdf} is denoted $\mathbb{P}$ and standard deviation $\sigma_X$.
@@
 \end{equation}
-If $\mathbb{P}$ has an exponential posititve tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
+If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
 \begin{equation}
@@
 
-In the limit of very rare events, direct sampling approaches become unrealistic. As an example, consider the drag process $t \to f_d(t)$.
-In the limit of extreme positive fluctuations, the return time of a fluctuations $f_d \geq a$ verifies
+%% Limites de l'approche par échantillonnage direct
+In the limit of very rare events and/or very complex dynamics such as turbulent flows in industrial contexts, a direct sampling approach is unrealistic.
+For instance, in the limit of extreme positive fluctuations, the return time of a fluctuations $f_d \geq a$ verifies~\ref{sec:lestang2018}
 \begin{equation}
@@
 where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
-The computational cost required to sample events $f_d \geq a$ therefore scales like $e^{\alpha a}$.
+As a result, the computational cost required to sample events $f_d \geq a$ scales like $e^{\alpha a}$.
 
-In this section we discuss the application of rare event algorithms to the numerical sampling of extreme drag forces on immersed objects, on the example
+%% Objectif des algo.
+In this section we discuss the application of \textbf{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
 of the flow presented in section~\ref{sec:test_flow}.
-The purpose of these algorithm is to sample rare events for a computational cost well inferior to their return time.
-Many of them involve keeping track of an ensemble of $N$ trajectories $\{x_n(t)\}_{n}$.
+The purpose of rare events algorithms is to make it possible to sample rare events for a computational cost well inferior to their return time.
+In sections~\ref{sec:ams} and~\ref{sec:gktl} we consider the \acl{ams} and \acl{gktl} algorithms, respectively.
+Both algorithms rely on the simulation of an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$, where $\{\mathbf{x}(t)\}_{0\leq t T_a}$ is a formal notation for a trajectory
+of duration $T_a$ in the phase space describing the dynamical system.
+In this case the dynamical system is the flow described in section~\ref{sec:test_flow}.
 Over the iterations of the algorithm, trajectories are duplicated or discarded from the ensemble according to precise selection rules that bias the sampling towards the rare events of interest.
 Such selection rules are designed in such a way that the sampling bias is known at each iteration of the algorithm.
-In other words, generated trajectories are given a weight which knowledge allow for the computation of their probability and expectation values according to the unbiased stationary distribution of the dynamics.
+In other words, generated trajectories are given a statistical weight, which knowledge allows for the computation of both their probability and expectation values of any trajectory observables.
 
-depending on their relevance to
-the realisation of extreme events of interest. 
-Numerous algorithms and applications have been proposed since the original ideas of Kahn and Harris in the early 50s~\cite{kahnharris}, which can be sorted in two categories.
-On the one hand, \textit{splitting} approaches decompose rare events into a succession of less rare events that can be sampled with a higher probability.
-On the other hand, \textit{importance sampling} relies on the sampling of trajectories according to a modified distribution, that is biased towards the rare events of interest.
-In section~\ref{sec:ams}, we highlight that sampling  extreme flucutations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme flucutations of the time-averaged drag.
-As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
+% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
+% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
+% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
 
@@
 % Introduction to AMS
-The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the event into a sequence of events that can be sampled with a higher probabilbity.
-Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date.
-In recent years, it was applied to several problems, ranging from the computation of the dissociation time of biomolecules~\cite{teo2016} to the simulation of rare relaminarisations in a stochastic model of wall turbulence~\cite{rolland2017}.
+The \acl{ams}~\cite{amspaper} inherits from the basic ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman,rolland_simonnet}.
+Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date (\textbf{Sans doute un peu <<too-much>>.
+Avec cette phrase je voudrais justifier que l'on s'intéresse à l'AMS et pas à un autre algo.}).
+In recent years, it was applied to several problems, for instance the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
 
@@
 In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
-The \ac{ams}, and therefore the \ac{tams}, relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t))$.
-It is an observable of the dynamics that acts as a measure of the distance of the current state $\mathbf{x}(t)$ to the region of the state space that correspond to the rare event(s) of interest.
-The \ac{ams} algorithm is an iterative procedure, that starts from an ensemble of $N$ independant trajectories of the dynamics over a time-interval $[0;T_a]$.
-At each iteration, trajectories with the lowest maxima of the score function over $[0;T_a]$ are discarded and replaced by trajectories that achieved a higher value of the score function at some point of their
-history,
-The algorithm is sketched in figure~\ref{figure_AMS}.
-For further details about the algorithm and assoiated mathematical results, see~\cite{lestang2018} and references therein.
+The \ac{tams} relies on the definition of a time-dependant \textit{score function} $\xi (\mathbf{x}(t),t)$, as well as an iterative procedure starting from an ensemble $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$ of independant trajectories.
+Each iteration consists in discarding the trajectories with the lowest maxima of the score function $\xi$ over the whole trajectory of duration $[0;T_a]$.
+Discarded trajectories are resampled based on member of the ensemble of trajectories that achieved a higher value of the score function at some point of their history.
+As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{figure_AMS}.
+For further details about the algorithm and corresponding mathematical results, see~\cite{lestang2018} and references therein.
 
 % Example: OU process
-We first describe the result of the \ac{tams} on a simple case.
-We consider a one-dimensional \ac{ou} process
+We first describe results of the \ac{tams} on a simple case.
+We temporarily let fluid dynamics aside and consider a one-dimensional \ac{ou} process
 \begin{equation}
@@
 \end{equation}
+We apply the \ac{tams} to simulate trajectories $\{x(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
+For the sake of simplicity, a single trajectory is resampled at each iteration of the \ac{tams}.
+We denote by $a_j$ the maximum of the resampled trajectory at iteration $j$.
+The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
+Figure~\ref{fig:compraison_temps_de_retour} illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
+More importantly, these very rare trajectories are sampled for a computational cost that several order of magnitude lower than the return time of the corresponding fluctuations.
 
-and apply the \ac{tams} to simulate trajectories $\{\mathbf{x}(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
-Figure~\ref{fig:compraison_temps_de_retour} illustrate the amplitude of the maximum over trajectories along the iterations of the algorithm.
-The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
-Figure~\ref{fig:compraison_temps_de_retour} illustrates that the \ac{tams} samples extreme fluctuations $x \geq a$ for a computational cost several order of magnitude lower than the return time of such fluctuations.
 Undoubtedly, an \ac{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} and \ac{tams}.
 Indeed, the state space is one-dimensional and the choice of the score function poses no questions: it is simply the observable $x$ itself.
-In addition, the noise term in~\eqref{eq:ou} has no correlations in time, which means that newly generated trajectories will quickly separate from their ancesotrs.
+In addition, the noise term in~\eqref{eq:ou} has no correlations in time, which means that newly generated trajectories will quickly separate from their ancestors.
 
 % Application to turbulent flow past an obstacle
-Can we obtain similar results by applying the \ac{tams} to the test flow described in\ref{sec:test_flow} ?n
-The objective is to sample trajectories that display extreme fluctuatons of the drag $f_d$ acting on the square obstacle.
-However, by contrast with a simple \ac{ou} process, the phase space in higly-dimensinal and the choice of the score function $xi (\mathbf{x}(t))$ is not so clear.
+Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
+To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
+Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
+However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
+As a consequence the choice of the score function $xi (\mathbf{x}(t),t)$ is not so clear.
 In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
-For such complex dynamics however, there is no resons for the drag itself to be a good measure of how far is a given state in phase space to regions where the drag is extreme.
-Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, whcih is precisely what we are tryping to achieve.
+For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
+Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.
 
@@
 
-Figure~\ref{fig:instant_256_20} displays similar information as figure~\ref{fig:comparaison_temps_de_retour} for $N=256$ initial trajectories over $T_a = 20 \tau_c$.
-Additionally, figure~\ref{fig:instant_256_20} shows the distribuion of the maximum of the drag among the initial trajectories.
+We apply the \ac{tams} with $N=256$ initial trajectories, with a duration $T_a = 20 \tau_c$.
+Similarly to figure~\ref{fig:comparaison_temps_de_retour}, figure~\ref{fig:instant_256_20} displays the maximum drag achieved by resampled trajectories.
+Figure~\ref{fig:instant_256_20} also shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
 Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
 These maxima are depicted by the stars in figure~\ref{fig:instant_256_20}.
-However, the maximum over the resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
+However, one can see that resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
 
@@
 
-Figure~\ref{fig:instant_256_20} can be better understood by visualizing the set of the last resampled trajectories. 
-One can see that, ultimately, trajectories have been resampled from the same initial trajectory.
-We explain this phenonemon as follows.
-Because the flow dynamics hae a memory, it takes a certain time before a resampled trajectory separate from its parent.
-This is clearly visible on figure~\ref{fig:instant_256_20}.
-In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting point will overlap over a timescale called
-the <<Lyapunov timesacle>>~\cite{lyapunovRef}.
-However, we observe that the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag proces.
+Figure~\ref{fig:instant_256_20} can be better understood by visualising the drag signals corresponding to the last resampled trajectories.
+Figure~\ref{fig:instant_256_20_b} displays the ensemble of trajectories after $181$ iterations of the \ac{tams} resampling procedure.
+Comparison with figure~\ref{fig:instant_256_20_b} suggests that, utimately, trajectories are resampled from the same initial trajectory.
+This phenomenon can be explained as follows.
+Because the flow dynamics have a memory, it takes a certain time before a resampled trajectory separate from its parent.
+In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting points will overlap over a timescale called
+the <<Lyapunov timescale>>~\cite{lyapunovRef}.
+However, we observe that, in our particular case, the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag process.
 This means that the resampling of a trajectory based on the value of the drag at $t^{\star}$ cannot lead to larger drag values for $t^{\star} \leq t \leq \tau_L$.
-For $t > \tau_L$, the drag process has no memory of the drag fluctuations on which the resampling was based and the probability of observing an extreme fluctuation over $\tau_l \leq t \leq T_a$ is very low.
+For $t > \tau_L$, the drag process has no memory of the drag fluctuations on which the resampling was based.
+As a consequence, the probability of observing an extreme fluctuation over $\tau_l \leq t \leq T_a$ is very low.
 The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_0$ is linked to the convective nature of the flow.
-Indeed, fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for
-trajectories to separate.
+As described in section~\ref{sec:instantaneous_drag}, extreme fluctuations of the instantaneous drag $f_d$ have a lifetime of roughly $\tau_c$.
+Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for
+trajectories to separate, \textit{i.e} the Lyapunov timescale.
 
 % perspectives
+This result suggest that the \ac{tams}--- and \ac{ams}---cannot be expected to efficiently sample rare trajectories for dynamics with such a phenomenology.
+% More trajectories ?
+Note that, because of the complexity of the dynamics, we were limited to a rather low number of trajectories, relatively to previous use cases of these algorithms~\cite{refs}.
+Targeting events with a fixed return time, one can show that the computational cost of a \ac{tams} experiment roughly grows linearly with the number of trajectories $N$~\cite{lestang2018}.
+As a result, multiplying the number of trajectories by a factor 10 or 100 would require important computational resources as well as an efficient parallel implementation of the \ac{tams}.
+Despite the relative simplicity of the \ac{tams} algorithm, the latter implementation is difficult to achieve~\cite{ref}.
+In any cases, from the observations made in this section, we do not expect that increasing the number of trajectories is able to better the efficiency of the sampling for such dynamics.
+% Another score function ?
+%%
+% The choice of the simple score function was already discusses above...
+% Maybe not necessary to RE-discuss it here.
 
-\begin{itemize}
-\item more trajectories
-\item different score functino?
-\item Change algorithm
-\end{itemize}
+% Modify the TAMS to fit the phenomenology ?
+A promising perspective is to adapt the \ac{tams} to dynamics for which the Lyapunov time is larger than the typical duration of extreme events, such as the convective flow involved in this work.
+In the original \ac{tams}, at later stages of the algorithm, new trajectories are resampled starting from points very close to the maximum of the parent trajectory.
+Resampled trajectories have no chance to display a significantly higher maximum as they overlap their parent trajectory over the Lyapunov timescale.
+As an example, the resampling procedure could be modified so that new trajectories are resampled roughly a Lyapunov time \textit{before} the maximum is attained.
+In this way, it gives the opportunity for new events, associated to a higher fluctuations, to occur.
+However, such modification(s) of the resampling procedure must be implemented with care, so as to preserve the mathematical properties of the \ac{ams} and \ac{tams}.
 \begin{figure}
@@
 We now turn to the sampling of extreme fluctuations of the time-averaged drag $F_T$.
-The \ac{ams} could be used in the same way as described in section~\ref{sec:ams} for the instantaneous drag, choosing $F_T$ itself as a score function.
-However, we observed a similar behaviour: the \ac{tams} resampling procedure fails to yield trajectories with higher fluctuations of $F_T$.
+The \ac{tams} algorithm could be used in the same way as described in section~\ref{sec:ams} for the instantaneous drag, choosing $F_T$ itself as a score function.
+However, this leads to the same problem: the resampling fails to efficiently generate new, higher fluctuations of $F_T$.
 
-When it comes to time-averaged observable, one can make use of a different algorithm the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
-It was originally designed to compute large deviations of time-averaged dynamical observables.
-By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not belong to the family of splitting algorithms.
+For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
+It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
+averaging time~\cite{ref_large_dev}.
+By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
 Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
@@
 
-\subsubsection{The algorithm}
+\subsubsection{The \ac{gktl} algorithm}
 \label{sec:gktl_description}
-Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a population of trajectories.
+Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
 In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
-The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is tilted in order to favour trajectories with large values of a chosen time averaged observable.
-This sort of algorithm has first been proposed by~\cite{Giardina2006} and has been used to study rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
+The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
+The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous_varizani,grassberger}.
+Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{delMoral_Feynman_Kac_book}.
+The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{Giardina2006}.
+Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
 
-More precisely, we perform simulations of an ensemble of $N$ trajectories $\left\{ X_{n}(t)\right\}$ (with $n=1,2,...,N$) starting from random initial conditions.
-Like in section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
-We consider an observable of interest $A(X(t))$ and a resampling time $\tau$.
+The \ac{gktl} algorithm consists in simulation an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{1 \leq n \leq N}$ starting from independant random initial conditions.
+Similarly to section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
+We consider an observable of interest $A(\mathbf{x}(t))$ and a cloning period $\tau$.
 At times $t_{i}=i\tau$ (with $i=1,2,...,T_{a}/\tau$) we assign to each trajectory $n$ a weight $W_{n}^{i}$ defined as
 \begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}.
 \end{equation}
-For each trajectory $X_{n}$, a random number of copies of the trajectory are generated, on average proportional to the weight $W_{n}^{i}$ and such that the total number of trajectories produced at each event is equal to $N$.
+For each trajectory $\{\mathbf{x}_{n}\}_{0 \leq t_i}$, a random number of copies of the trajectory are generated.
+This number is on average proportional to the weight $W_{n}^{i}$, such that the total number of trajectories produced at each event is equal to $N$.
 For deterministic systems, a small perturbation is introduced for clones at times $t_i$, so that clones separate from their parent. 
-The parameter $k$ is chosen by the user in order to control the strength of the selection and thus to target a class of extreme events of interest.
-The larger the value of $k$, the more trajectories with large values of the time average observable will survive the selection.
+The parameter  is chosen by the user in order to control the strength of the selection.
+The free parameter $k$ sets the amplitude of the typical fluctuations within the biased ensemble of trajectory sampled by the algorithm.
+A higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.
 
@@
 
-We now discuss the application the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
+We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
 The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
-Consistenly with section~\ref{sec:time_avg}, we choose the averaging window $T=10\tau_c$.
+Consistently with section~\ref{sec:time_avg}, we choose the averaging window $T=10\tau_c$.
 
 % Choice of parameters and perturbation
-The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C = N_c \times T_a$.
-In the following series of experiements we set $T_a = T = 10\tau_c$.
-The sampling from the \ac{gktl} alorithm will be compared to the sampling from a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
-To match this computational cost with the \ac{gktl} experiments, we set $N_c=16384$.
+The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
+In the following series of experiments we set $T_a = T = 10\tau_c$.
+The sampling from the \ac{gktl} algorithm will be compared to the sampling from the control run, \textit{i.e.} a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
+We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+(\textbf{Plus vraiment clair pour moi pourquoi je n'ai pas utilisé 10X plus de clones, pour matcher le coût numérique de la simulation de contrôle...})
 Lastly, one must set the cloning period $\tau$.
-On the one hand, too small a cloning period may result in unfortunate loss of information as clones do not separate from their parent over a cloning period.
-On the other hand, choosing $\tau \gg \tau_c$ leads to too few cloning stages.
-As a consequence, we set $\tau = \tau_c /2$.
+On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
+On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
+A rule of thumb is to set $\tau \approx \tau_c$.
+In the following, we set $\tau = \tau_c /2$.
 
@@
 We performed three experiments, corresponding to three different values of the parameter $k$.
-This parameter sets the amplitude of the typical fluctuations within the biased ensemble sampled by the algorithm.
-In other words, a higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.
 In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
-In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled flucutations through the Gartner-Ellis theorem~\cite{GEref}.
-In the large time limit $T_a\to \infty$, the typical level of fluctuations $\mathbb{E}_{\mathbb{P}_k}[ F_{T_a} ]$ within the biased ensemble of trajectories
-can be related to the \textit{Scaled Cumulant Generating Function} $\lambda(k)$~\cite{ref}.
+In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
+% In the large time limit $T_a\to \infty$, the typical level of fluctuations $\mathbb{E}_{\mathbb{P}_k}[ F_{T_a} ]$ within the biased ensemble of trajectories
+% can be related to the \textit{Scaled Cumulant Generating Function} $\lambda(k)$~\cite{ref}.
 
-Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved with the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
-For each three experiment, it displays the estimate of the biased \ac{pdf}
+Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
+For each of the three experiments, it displays the estimate of the biased \ac{pdf}
 \begin{equation}
@@
 computed over the set of $N=16384$ trajectories sampled by the algorithm.
-Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} describing the time-averaged drag process.
-It corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.
+Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
+Note that it corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.
 
@@
 For a fixed number of trajectories $N$, there is a limited range of $k$ values for which the algorithm is able to yield an accurate estimate of $\rho_k$.
-Setting $k$ to a high value causes finite population effects: trajectories overlap over most of their duration.
+Setting $k$ to too high a value causes harmful effects resulting from the finiteness of the ensemble of trajectories: trajectories overlap over most of their duration.
 This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
@@
 % qu'un échantillonnage direct
-Unsurprisingly, sampling larger fluctuations thus reauire a larger number of populations, and therefore a larger computational cost.
+Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
 Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
@@
 Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
-As a comparison, it also displays the average number of such fluctuations in a long timeseries of duration $T_{tot}=10^6\tau_c$, divided in $T_{tot} / T_a$ events.
-This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian, motivated by figure~\ref{fig:pdf_averared}.
+As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
+This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
+This hypothesis is motivated by figure~\ref{fig:pdf_averared} (\textbf{figure manquante}).
 While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
 
+In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+In the latter case, each fluctuation typically lasts a correlation time $\tau_c$.
+In this context, the \ac{gktl} algorithm is particularly suited to sample trajectories that display a succession of these fluctuations.
+(\textbf{conclusion à développer.})
 \begin{table}
@@
 Within the biased ensemble, trajectories are considered distinct if they overlap for less than $50\%$ of their duration.
-    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ independant realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.
+    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.
     This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.}
@@
 \end{figure}
+
+\section{Conclusion}
+\label{conlusion}
+
 \end{document}

### 2019-01-07 draft_p2.tex
Ajoute un paragraphe d'intro pour la section sur l'échantillonnage direct des fluctuations de trainée

@@
 \begin{document}
-\title{A statistical mechanics approach to the simulation of extreme fluctuations of the time-averaged drag acting on an obstacle immersed in a turbulent flow}
+\title{Brouillon PoF,PRF..}
 
@@
 
-\paragraph*{Paragraphe d'introduction: les motivations d'un étude directe ?}
-%% TODO: Introduire la notion de temps de retour
-Rare event algorithms relies on the simulation of an ensemble of trajectories.
-Members of the ensemble can be duplicated or removed from the ensemble, based on the evaluation of a \textit{score function}.
-A score function is an observable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
-For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achieved by the trajectory.
-In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observable over the trajectory.
-However, for high-dimensional dynamics with a very complex phase space, the simple distance to the target event is likely to be a very poor criterion for selecting rare trajectories.
-\textbf{[...]}
+In this section we describe the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
 
+How rare is a fluctuation $f_d \geq a$ can be determined by computing its \textit{return time} $r(a)$.
+It is defined as the average time between two consecutive occurrences of a fluctuation $f_d \geq a$.
+Extreme drag fluctuations are \textit{rare events}, that is fluctuations for which $r(a) \gg \tau_c$.
+In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang2018}
+\begin{equation}
+  \label{eq:return_time}
+  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
+\end{equation}
+where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
+Figure~\ref{fig:return_time_instant} illustrates the fluctuation amplitude $a$ as a function of the return time $r(a)$, for the drag process $f_d(t)$.
+Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang2018}.
+
+In the following we discuss extreme fluctuations of both the instantaneous drag $f_d$ and time-averaged drag with return times larger than $10^4\tau_c$.
+In this section extreme drag fluctuations are simply sampled by simulating the flow over a very long duration.
+This approach is referred to as \textit{direct sampling}, as opposed to approaches involving the use of rare events algorithms, discussed in section~\ref{sec:rare_events_algorithms}.
+Such approach is made possible by the relative simplicity of the flow described in section~\ref{sec:test_flow}.
+The control run mentioned in the previous section, spanning $T_{tot} = 10^6 \tau_c$, could be achieved in a wall-clock time of a few weeks and gives access to extreme fluctuations $f_d \geq a$ with a return time $r(a) \leq 10^6\tau_c$.
 \subsection{Extracting extreme drag fluctuations from a very long timeseries}
@@
   \centering
-  \includegraphics[width=\linewidth]{return_time_drag}
+  \includegraphics[width=.6\linewidth]{return_time_drag}
   \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
@@
 In the limit of very rare events and/or very complex dynamics such as turbulent flows in industrial contexts, a direct sampling approach is unrealistic.
-For instance, in the limit of extreme positive fluctuations, the return time of a fluctuations $f_d \geq a$ verifies~\ref{sec:lestang2018}
-\begin{equation}
-  \label{eq:return_time_scaling}
-  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
-\end{equation}
-where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
+Indeed, according to equation~\eqref{eq:return_time_scaling}, the return time of a fluctuation $f_d \geq a$ verifies $r(a) \propto e^{\alpha a}$.
 As a result, the computational cost required to sample events $f_d \geq a$ scales like $e^{\alpha a}$.

### 2019-02-07 draft_p2.tex
Adds new figures to draft

@@
 \usepackage{caption}
-\graphicspath{{../figures/}}
+\graphicspath{{./figures/}}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
+  \includegraphics[width=\linewidth]{illustr_vort/illustration_vorticity_tf2}
   \caption{Typical vorticity field in the computational domain. The square obstacle is represented by the gray area in the center of the channel.}
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
+  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
   \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}.
@@
   \centering
-    \subfloat[\ac{pdf} for the drag fluctuations $f'_d$]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{../figures/pdf_drag_SQ_asymetry_expo}}
-    \subfloat[Autocorrelation function of the the instantaneous drag $f_d$]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{../figures/correlation_with_log}}  
+    \subfloat[\ac{pdf} for the drag fluctuations $f'_d$]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{./PDF_drag/PDF_drag.png}}
+    \subfloat[Autocorrelation function of the the instantaneous drag $f_d$]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{autocorr_drag/autocorr_drag.png}}  
   \caption{\textbf{(a)} \ac{pdf} describing the statistics of the instantaneous drag fluctuations $f'_d = (f_d - \bar{f_d})/\sigma$.
@@
   \centering
-  \includegraphics[width=.6\linewidth]{return_time_drag}
+  \includegraphics[width=.6\linewidth]{return_time/return_time.png}
   \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
@@
   \centering
-  \label{fig:contribution_forebody_and_base_pressure}
+  \includegraphics{pressure_ratio/pressure_ratio.png}
+  \label{fig:pressure_ratio}
 \end{figure}
@@
 By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinder are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
-In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation, relatively to the forebody pressure fluctuation.
+In addition, figure~\ref{fig:pressure_ratio} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation, relatively to the forebody pressure fluctuation.
 
@@
   \centering
-  \begin{minipage}{0.85\linewidth}
-    \begin{tabular}{c|c}
-    \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
-    \hline
-    \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
-    \end{tabular}
-  \end{minipage}
-\begin{minipage}[h]{0.1\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{scale_vorticity.png}
-\end{minipage}
+  \includegraphics[width=.8\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
   \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\linewidth]{timeseries_extremes/timeseries_extremes.png}
+\end{figure}
+
+% \begin{figure}
+%   \centering
+%   \begin{minipage}{0.85\linewidth}
+%     \begin{tabular}{c|c}
+%     \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
+%     \hline
+%     \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
+%     \end{tabular}
+%   \end{minipage}
+% \begin{minipage}[h]{0.1\linewidth}
+%   \centering
+%   \includegraphics[width=\linewidth]{scale_vorticity.png}
+% \end{minipage}
+% \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
+The corresponding drag timeseries can be found in figure~\ref{fig:extremes_timeseries}.
+%   The high value for the drag results from the formation of a strong negative (red) vortex inducing a pressure drop at the base of the obstacle.
+The formation of such a structure is aided by important vorticity production at the bottom boundary of the obstacle coupled with the influence of positive vorticity in the vicinity of the base.}
+% \end{figure}
 
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
+  \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
   \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
+  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
     \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/instant_256_20}
+  \includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
     \caption{Maximum of the instantaneous drag along the resampled trajectories, denoted by $\tilde{f}_d^{(j)}$, as a function of the corresponding computational cost $C_j$, defined in~\eqref{eq:comp_cost_iterations}.
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/iter_181}
+  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
     \caption{Visualisation of the drag timeseries corresponding to the set of trajectories in the ensemble, at different iterations along the algorithm.
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{illustration_IS}
+%  \includegraphics[width=0.7\linewidth]{illustration_IS}
   \caption{\textbf{(a)}: Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 

### 2019-02-07 figures/illustr_extrms_vorticity/make_plot_vorticity.py
Updates script and figure for illustration of vorticity fields for extreme events

@@ -85,9 +85,8 @@ for i in range(0,Nevents):
 cbar=fig.colorbar(im, ax=ax_list)
 cbar.ax.tick_params(labelsize=16)
 
-
-
-plt.show()
 fname = 'illustr_extrms_vorticity.png'
 plt.savefig(fname)
+
+plt.show()
             

### 2019-02-07 draft_p2.tex
rewrites the description of the flow fields

@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
+\end{figure}
+
 % \begin{figure}
@@
 
-We now turn to a description of the flow dynamics for some of the extreme drag fluctuations, sampled in the control timeseries.
-Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events corresponding to the largest fluctuation amplitudes.
-One can see that they correspond to very similar flow configurations.
-In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
-Such strong vorticity is produced through viscous shear along either the bottom or top boundary of the obstacle.
+We now turn to a description of the flow dynamics that correspond to extremely high values of the drag acting on the square obstacle.
 
-By contrast with typical configurations, we observe that this vorticity is not directly advected by the mean flow, but instead constrained in the vicinity of the base of the obstacle.
-Indeed, in each of the four examples displayed in figure~\ref{fig:top_four_events}, a large vortical region is located at roughly one obstacle length of the base.
-As illustrated by figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.
+Figure~\ref{fig:timeseries_extremes} displays the average of the drag timeseries corresponding to the 104 extreme events sampled in the control run, in the vicinity of the fluctuation.
+It illustrates that extreme drag fluctuations are very local in time, with a lifetime of roughly one correlation time unit.
+Starting from typical values, extreme drags are typically reached in less than a correlation time.
+Moreover, such very high drag levels do not persist over time.
+To understand why it is so, it is useful to visualize the corresponding flow fields.
 
-Figure~\ref{fig:top_four_events_timeseries} displays the drag timeseries corresponding to the four events illustrated in figure~\ref{fig:top_four_events}.
-It illustrates that the lifetime of such fluctuations temporal extension is roughly one correlation time unit $\tau_c$.
-It is the time it takes for the block eddy to form and be advected away by the mean flow, thus freeing the vorticity concentrated in the vicinity of the base of the obstacle.
-This can be seen in figure~\ref{fig:dynamics}, which illustrate the vorticty dynamics from $t^{\star}-\tau_c/2$ to $t^{\star}$ on a particular example of an extreme drag fluctuation. 
-On can see in figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy.
-This interaction eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
-This can be seen see in figure~\ref{fig:dynamics_k}.
-Although this vorticity is too far from the base to directly affect the base pressure, and thus the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle---see figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.
+To begin with, figure~\ref{fig:top_four_events} displays the vorticity field corresponding to the peak drag value, for a subset of the extreme events sampled in the control simulation.
+We observe that, in each case, extreme drag flcutuations correspond to very similar flow configurations.
+More precisely, a region of strong vorticity is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
+This vorticity results from a boundary layer, along either the top or bottom side of the obstacle.
+Such extreme vorticity is responsible for a significant pressure drop in the vicinity of the base of the obstacle, which in turn leads to an extreme value of the drag, see figure~\ref{fig:pressure_ratio}.
+
+How are such vorticity reached ?
+In contrast with the typical flow dynamics, illustrated in figure~\ref{vorticity_typical}, the vorticity generated within the boundary layers is not directly advectd by the means flow.
+Figure~\ref{fig:illustr_density_streamlines} shows that, instead, this vorticity is constrained in the vicinity of the base of the obstacle.
+Such constraint results from a large vortical region located roughly one obstacle length of the base, that prevents vorticity to be advected downstream.
+
+Figure~\ref{fig:dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:illustr_density_streamlines}.
+From $t^{\star}-0.4\tau_c$ to $t^{\star}-0.16\tau_c$, a strong negative eddy is generated along the bottom boundary layer of the obstacle and advected downstream by the mean flow.
+This leads to a region of negative vorticity at roughly one obstacle length of the base of the obstacle, constraining the positive vorticity generated along the top boundary.
+However, it can be seen in figure~\ref{fig:dynamics} that the positive vorticity only accumulates for a short duration.
+The blocking negative vortex is in turn advected downstream, thus freeing the trapped positive vorticity.
+
+The lifetime of an extreme drag fluctuation therefore corresponds to the duration over which the blocking vortex remains in the vicinity of the base.
+Consistently with figure~\ref{fig:timeseries_extremes}, this duration is roughly one correlation time unit $\tau_c$.
 
@@
 % It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-Figure~\ref{fig:gamma_asof_fd} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
+Figure~\ref{fig:shear_asof_drag} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
 For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
@@
 For a more detailed description of such dynamics, see the chapter 3 of ~\ref{ref:these_lestang}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
+  \caption{\label{fig:shear_asof_drag}}
+\end{figure}
 

### 2019-02-07 draft_p2.tex
Adds first modifications from Emmanuel

--- 2019-01-07 draft_p2.tex
@@
-% \include{header_p2.tex}
-
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage]{revtex4-1}
+\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage,preprint]{revtex4-1}
 \pdfoutput=1
@@
 \usepackage{caption}
-\graphicspath{{../figures/}}
+\graphicspath{{figures/}}
 
@@
 \begin{document}
-\title{Brouillon PoF,PRF..}
+
+\title{Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
 
@@
 \email{thibault.lestang@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
+\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
 \author{Freddy Bouchet}
 \email{freddy.bouchet@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
+\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
 \author{Emmanuel Lévêque}
 \email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+
+\begin{abstract}
+This study evaluates the relevance of importance sampling techniques to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+The main idea behind importance sampling is here to replace a long simulation by a set of shorter simulations running in parallel with dynamics that are replicated or abandoned in order to sample large events more frequently.
+%
+Such techniques have already proved their efficiency for simulating extreme events in... with significant run-time savings.
+Application to turbulent fluid-mechanical systems remains mainly open. 
+%
+The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two important sampling algorithms, namely ....
+Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort...
+However, limitations...
+pitfalls...
+Importantly, such techniques also provide access to the flow scenario yielding extreme events. In the present case, it is found that... blocking effect
+\end{abstract}
+
 \maketitle
+
 \section{Introduction}
-The purpose of this paper is twofold.
-\section{Description of the test case for extreme drag fluctuations sampling}
+
+// Recall the previous success of importance sampling techniques with references. //
+Explain why interesting to test application for turbulent systems. // To what extent a turbulent system is different from previous case studies? // 
+\\
+
+Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic fluid motions associated with enormous internal forces.
+In the turbulence literature, ``extreme'' can refer to fluctuations that deviate from the mean value by $O(10)$ standard deviations.
+// true ?
+//
+The nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
+\\
+
+From the viewpoint of dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium.
+The use of perturbative methods in identifying resonant interactions (among degrees of freedom)  leading to extreme fluctuations is not successful.
+In this context, simulation offers a practical means to gain insight into the extreme events. However, their numerical sampling requires long simulation since these events are rare as well. 
+\\
+
+// Our motivations and expectations. //
+
+\section{Description of the case study}
 \label{sec:test_flow}
@@
   \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
-  \caption{Typical vorticity field in the computational domain. The square obstacle is represented by the gray area in the center of the channel.}
+  \caption{Our case study is a grid-generated turbulent flow impinging a squared obstacle located at the center of a channel (in two dimensions).
+The vorticity is displayed with an arbitrary colormap from blue (negative values) to red (positive values).
+The Reynolds number based on the grid spacing is $Re=...$.
+// no color bar, show the grid //}
+  \label{fig:computational_domain}
 \end{figure}
-%% Description de la géométrie
-We consider the interaction of a two dimensional turbulent channel flow with a square cylinder, illustrated in figure~\ref{fig:computational_domain}.
-A constant parabolic velocity profile is imposed at the inlet of the channel and turbulence is
-generated by means of a grid.
-A square obstacle is mounted at the center of the channel.
-Eddies are created in the vicinity of the grid with a typical size that correspond to the step of the grid.
-As they are advected downstream, eddies interact with each other and grow in size, until they interact with the obstacle.
-The distance separating the obstacle from the grid is such that the typical size of the eddies it interacts with is close to the size of the obstacle itself.
-The flow in the vicinity of the outlet is regularized by means of a \textit{sponge layer}~\cite{ref1,ref2}, \textit{i.e.} a region where the viscosity is progressivaly increased.
-Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
 
-%% Méthode Boltzmann sur réseau
-The flow dynamics are computed by means of the \ac{lbm}~\cite{ref1,ref2}.
-[[Paragraphe LBM]]
+The drag exerted by a grid-generated turbulent flow onto a squared obstacle is considered as our representative case study.
+The flow is contained in a long channel as illustrated in Fig.~\ref{fig:computational_domain}.
+Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and make possible a systematic study.
+Nevertheless, this fluid-mechanical system exhibits the characteristic features of turbulent fluid-structure interaction.
+// true even if 2d?
+// In the present situation, the obstacle does not deform or move.
+Through  this specific setting, our main motivation is to evaluate the operability of importance sampling techniques to capture extreme events with a significant run-time savings.
+\\
+
+A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid. 
+At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{ref1,ref2} where the viscosity is artificially enhanced. 
+% Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
+\\
+
+Turbulent eddies are generated in the near-wake of the grid and carried downstream. // grid is usually characterized by the spacing and the mesh size, are they equal here? //
+They interact with each other and grow in size as expected in two-dimensional turbulence. 
+The size of the eddies that eventually hit the square (at the center of the channel) is comparable to the size of this obstacle, which results in strong fluctuations of the drag acting on the square. 
+\\
+
+The flow dynamics are computed by the lattice Boltzmann (LB) method ~\cite{ref1,ref2}. 
+While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
+Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
+Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
+This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
+This is, in short, the rationale behind the LB approach.
+In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
+%
+More precisely, the fluid is viewed as populations of particles
+that collide, redistribute and propagate along the different links of a discrete lattice. 
+In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
+Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
+\[
+\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
+\]
+for the mass density and the fluid momentum respectively.
+The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{D2Q9}
+	\caption{Sketch of the D2Q9 lattice.
+Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
+By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
+	\label{fig:D2Q9}
+\end{figure}
+
+The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
+Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+Namely, an \emph{exact} streaming step
+\[
+f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
+\]
+during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
+\[
+f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
+\]
+which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
+\[
+\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
+\]
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
+\[
+f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
+\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
+\] 
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
+Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
+$\Delta x \to 0$.
+
+// algorithm for boundary conditions // computation of the stress with LBM // grid resolution, number of grid points in each direction, Reynolds, Mach numbers, etc.
+// low Mach number approaches the nearly-incompressible limit
+\\
+
+As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
+\[
+\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\]
+so that the total stress expresses as
+\begin{equation}\label{eq:def_stress}
+\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\end{equation}
 
@@
 
-%% Descritpion d'une fluctuations de trainée typique
-The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts.
-In particular, the \textit{drag force} is defined as the result of both pressure and viscous constraints, projected along the mean-flow direction.
-It reads
+The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle. The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and reads 
 \begin{equation}
   \label{eq:drag_definition}
-  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x,
+  f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x})  
 \end{equation}
-where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ is the identity matrix, $\boldsymbol{\tau}$ denotes the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.
-
-%% Viscous effects are negligible
-The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients in the vicinity of the obstacle.
-Furthermore, because of the blunt shape of the obstacle, the contribution of viscous constraints to the overall drag is negligage with respect to pressure constraints.
-As a result, the drag mainly originates from the pressure difference between the \textit{forebody}---the upstream boundary---and the \textit{base}---the downstream boundary.
-Consequently, the drag acting on the square cylinder is very well approximated by
+where $\mathcal{S}$ is the surface of the obstacle. 
+In our situation, the viscous stress makes a negligible contribution to the drag.
+This latter therefore results mostly from pressure forces, which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
+%
+Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
+As a consequence, the drag may be viewed as the pressure difference 
 \begin{equation}
@@
 \end{equation}
-where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base, respectively.
+where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base of the obstacle, respectively.
+
+\vskip 5cm
+END CORRECTIONS EMMANUEL
+\vskip 5cm
 
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
+  % \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
   \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}.
@@
   \centering
-  \includegraphics[width=.6\linewidth]{return_time_drag}
+ % \includegraphics[width=.6\linewidth]{return_time_drag}
   \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
@@
     \begin{tabular}{c|c}
-    \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
+%    \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
     \hline
-    \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
+%    \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
     \end{tabular}
@@
   \centering
-  \includegraphics[width=\linewidth]{scale_vorticity.png}
+%  \includegraphics[width=\linewidth]{scale_vorticity.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
+%  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
   \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
+%  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
     \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/instant_256_20}
+%  \includegraphics[width=.7\linewidth]{ams/instant_256_20}
     \caption{Maximum of the instantaneous drag along the resampled trajectories, denoted by $\tilde{f}_d^{(j)}$, as a function of the corresponding computational cost $C_j$, defined in~\eqref{eq:comp_cost_iterations}.
@@
   \centering
-  \includegraphics[width=.7\linewidth]{ams/iter_181}
+%  \includegraphics[width=.7\linewidth]{ams/iter_181}
     \caption{Visualisation of the drag timeseries corresponding to the set of trajectories in the ensemble, at different iterations along the algorithm.

### 2019-02-07 draft_p2.tex

@@
   \centering
-  \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
+  \includegraphics[width=\linewidth]{illustr_vort/illustration_vorticity_tf2.png}
   \caption{Our case study is a grid-generated turbulent flow impinging a squared obstacle located at the center of a channel (in two dimensions).
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{D2Q9}
+	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
 	\caption{Sketch of the D2Q9 lattice.
@@
 
-\vskip 5cm
-END CORRECTIONS EMMANUEL
-\vskip 5cm
-
 %% The phenomenology of typical drag fluctuations
@@
   \centering
-  % \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
+  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
   \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}.
@@
   \centering
- % \includegraphics[width=.6\linewidth]{return_time_drag}
+  \includegraphics[width=.6\linewidth]{return_time/return_time.png}
   \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
@@
   \centering
-  \begin{minipage}{0.85\linewidth}
-    \begin{tabular}{c|c}
-%    \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
-    \hline
-%    \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
-    \end{tabular}
-  \end{minipage}
-\begin{minipage}[h]{0.1\linewidth}
-  \centering
-%  \includegraphics[width=\linewidth]{scale_vorticity.png}
-\end{minipage}
+  \includegraphics[width=.8\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
   \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\linewidth]{timeseries_extremes/timeseries_extremes.png}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
+\end{figure}
+
+% \begin{figure}
+%   \centering
+%   \begin{minipage}{0.85\linewidth}
+%     \begin{tabular}{c|c}
+%     \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
+%     \hline
+%     \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
+%     \end{tabular}
+%   \end{minipage}
+% \begin{minipage}[h]{0.1\linewidth}
+%   \centering
+%   \includegraphics[width=\linewidth]{scale_vorticity.png}
+% \end{minipage}
+% \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
+The corresponding drag timeseries can be found in figure~\ref{fig:extremes_timeseries}.
+%   The high value for the drag results from the formation of a strong negative (red) vortex inducing a pressure drop at the base of the obstacle.
+The formation of such a structure is aided by important vorticity production at the bottom boundary of the obstacle coupled with the influence of positive vorticity in the vicinity of the base.}
+% \end{figure}
 
@@
   \centering
-%  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
+  \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
   \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
@@
   \centering
-%  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
+  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
     \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
@@
   \centering
-%  \includegraphics[width=.7\linewidth]{ams/instant_256_20}
+  \includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
     \caption{Maximum of the instantaneous drag along the resampled trajectories, denoted by $\tilde{f}_d^{(j)}$, as a function of the corresponding computational cost $C_j$, defined in~\eqref{eq:comp_cost_iterations}.
@@
   \centering
-%  \includegraphics[width=.7\linewidth]{ams/iter_181}
+  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
     \caption{Visualisation of the drag timeseries corresponding to the set of trajectories in the ensemble, at different iterations along the algorithm.

### 2019-02-17 figures/PDF_AVG/plot_pdf_AVG.py
Adds figures ecoulement_typique and PDF_AVG

@@ -1,0 +1,41 @@
+import numpy as np
+import matplotlib.style
+import matplotlib.pyplot as plt
+
+plt.style.use('ggplot')
+
+def ret_gaussian(x):
+    return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
+
+musigma = np.fromfile("musigma_AVG_10.dat", float, -1, "");
+
+buf = np.fromfile("pdfData_AVG.dat", float, -1, "");
+n=len(buf)/2
+f = buf[0:n]*musigma[1]
+xi = (buf[n:]-musigma[0])/musigma[1]
+
+fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
+
+lineSQ, = ax.plot(xi,f)
+lineSQ.set_label('PDF of $F_T$')
+
+
+
+# buf = np.fromfile("pdfData_NOSQ.dat", float, -1, "");
+# n=len(buf)/2
+# f = buf[0:n]
+# xi = buf[n:]
+lineNOSQ, = ax.plot(xi,ret_gaussian(xi), label='Gaussian PDF')
+
+
+ax.set_yscale('log')
+ax.set_xticks([-5,-3,-1,0,1,3,5])
+ax.set_xticklabels(['$-5\sigma_T$','$-3\sigma_T$','$-\sigma_T$','$0$','$\sigma_T$','$3\sigma_T$','$5\sigma_T$'], fontsize=16)
+ax.set_xlabel('$F_T$',fontsize=22)
+plt.ylabel('$\log(PDF)$',fontsize=22)
+ax.legend(loc='best',fontsize=22)
+
+fname = 'PDF_AVG.png'
+plt.savefig(fname)
+
+plt.show()

### 2019-02-17 figures/ecoulement_typique/compute_events.cpp

@@ -1,0 +1,136 @@
+#include <iostream>
+#include <fstream>
+#include <string>
+#include "libpipeLBM.h"
+
+#define _PATH_TO_INIT_ "/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/run_for_pops/populations/pops_1.datout"
+
+using namespace std;
+
+int main(int argc, char* argv[])
+{
+  int Dx = 513; int Dy = 129;
+  int N = 2000;
+  pipeLBM *myLB = new pipeLBM(Dx, Dy);
+  myLB->setInletBC("poiseuille");
+  myLB->setOutletBC("open");
+  
+  int spongeStart = (int) 3*(Dx-1)/4+1;
+  int R = (Dy-1)/16; int x0 = (Dx-1)/16;
+  int L = 16; int xsq = (Dx-1)/2; int ysq = (Dy-1)/2 - L/2;
+  double U0 = 0.05; double tau = 0.501; double F0 = 0.0;
+				    
+  myLB->setSpongeLayer(spongeStart, Dx);
+  myLB->setParameters(tau, U0, F0);
+
+  Obstacle* obs[2];
+  squareObstacle mySquare(xsq, ysq, L);
+  obs[0] = new Grid(x0, R, Dy);
+  //  obs[1] = new squareObstacle(xsq, ysq, L);
+  obs[1] = &mySquare;
+  myLB->setObstacles(obs, 2);
+  string path_to_file=_PATH_TO_INIT_;
+  bool errFlag;
+  errFlag = myLB->initFromFile(path_to_file);
+  if(errFlag){cout << "PB with init" << endl;}
+
+  double *fd = new double[N];
+  // double *viscousRear = new double[N];
+  // double *viscousTop = new double[N];
+  // double *viscousFront = new double[N];
+  // double *viscousBot = new double[N];
+  // double *pRear = new double[N];
+  // double *pFront = new double[N];
+  // double *shearBotVec = new double[N];
+  // double *shearTopVec = new double[N];
+  double *meanTransVelocityVec = new double[N];
+  double shearTop = 0, shearBot=0, meanTransVelocity=0;
+  int k = 0;
+  int kk=0;
+  for (int t=0;t<N;t++)
+    {
+      myLB->displayPercentage(t,N);
+      myLB->computeStress(obs[1]);
+      fd[t] = obs[1]->getDrag();
+      if(t%400==0)
+      	{
[…]

### 2019-02-17 figures/ecoulement_typique/plot_ecoulement_typique.py

@@ -1,0 +1,105 @@
+# Produces a figure illustrating the vorticity field at the moment of peak drag for 7 events.
+# The figure consists of 7 axes. One axes is four times bigger as the others, and contains 
+# annotations
+import numpy as np
+import csv
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
+
+Dx = 513
+Dy = 129
+N = Dx*Dy
+F = np.zeros(104)
+idx = np.zeros(104)
+
+prefix = './'
+
+fig = plt.figure(figsize=(16,7), constrained_layout=True)
+# Declares a 2x5 grid.
+
+grid = plt.GridSpec(2, 5, figure=fig)
+# the first row is divided into five snapshots
+for i in range(0,5):
+    plt.subplot(grid[0,i])
+#the second row is for the drag signal
+plt.subplot(grid[1,:])
+
+ax_list = fig.axes
+nplots = len(ax_list)
+
+path_to_file_ux = prefix+'ux.dat'
+path_to_file_uy = prefix+'uy.dat'
+fx = open(path_to_file_ux, "rb")  # reopen the file
+fy = open(path_to_file_uy, "rb")  # reopen the file
+bbox = dict(boxstyle="round", fc="0.9")
+
+for i in range(0,nplots-1):
+        # trick to advance in time with row major ordering
+        
+        fx.seek(i*N*8, 0)  
+        fy.seek(i*N*8, 0)  
+
+        # Get velocity fields
+        ux = np.fromfile(fx, float, N, "")
+        ux = np.reshape(ux, (129,513), 'F')
+        uy = np.fromfile(fy, float, N, "")
+        uy = np.reshape(uy, (129,513), 'F')
+    
+        ymin=32;ymax=96
+        xmin=236;xmax=310
+        # Compute vorticity using finite-differences
+        dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
+        dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
+        vort = dxy - dyx
+
+        # Print time as title
+        timeFromBeg=0.1+i*0.2
+        tit='$t={:2.1f}\\tau_c$'.format(timeFromBeg)
+
[…]

### 2019-02-17 figures/typical_drag_signal/plot_drag_timeseries.py
Update typical_drag_signl: bigger labels, x units is tau_c

@@ -15,8 +15,12 @@ ax.set_xlim(0, 1000)
 line, = ax.plot(tspan, f)
 line.set_linewidth(.7)
 
-ax.set_xlabel(r'$t / \tau_0$',fontsize=16)
-plt.ylabel('$f_d(t)$',fontsize=16)
+ax.set_xlabel(r'$t / \tau_c$',fontsize=22)
+plt.ylabel('$f_d(t)$',fontsize=22)
+ax.set_yticks(np.arange(-0.10,0.30,0.10))
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+
 
 fname= 'typical_drag_signal.png'
 plt.savefig(fname)

### 2019-02-17 draft_p2.tex
Adds figures ecoulement_typique and PDF_AVG
Updates caption for ecoulement typique

@@
 The correlation time $\tau_c$ is defined further in this text and correspond to the typical fluctuation timescale of the drag.}
+  \label{fig:typical_drag_signal}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
+  \caption{Vorticity dynamics corresponding to typical drag fluctuations over one correlation time unit $\tau_c$.
+Negative vorticity is generated along the bottom boundary, and advected away from the the obstacle by the mean flow.}
   \label{fig:typical_drag_signal}
@@
 This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
-This hypothesis is motivated by figure~\ref{fig:pdf_averared} (\textbf{figure manquante}).
+This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
 While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+  \caption{PDF of average drag over $10\tau_c$}
+  \label{fig:PDF_AVG}
+\end{figure}
 

### 2019-02-17 draft_p2.tex
updates figures and legend for autocorr and pdf instant

@@
   \centering
-    \subfloat[\ac{pdf} for the drag fluctuations $f'_d$]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[Autocorrelation function of the the instantaneous drag $f_d$]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{autocorr_drag/autocorr_drag.png}}  
-  \caption{\textbf{(a)} \ac{pdf} describing the statistics of the instantaneous drag fluctuations $f'_d = (f_d - \bar{f_d})/\sigma$.
-This \ac{pdf} has been estimated on the basis of a timeseries spanning $T_{tot} = 10^6 \tau_c$.
-$\bar{f_d}$ and $\sigma$ denote the average and the standard deviation computed over the whole timeseries, respectively.
-The linear fits highlight the exponential behaviour of the the \ac{pdf} for rare events.
-\textbf{(b)} Autocorrelation function for the instantaneous drag $f_d$, defined as $(\mathbb{E}[f_d(t+\tau)f_d(t)]-\mathbb{E}[f_d]^2)/\sigma^2$ and estimated from a timeseries spanning $T_{tot} = 10^6 \tau_c$.
-This figure shows the exponential decay of the correlations over time.}
+    \subfloat[]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{./PDF_drag/PDF_drag.png}}
+    \subfloat[]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{autocorr_drag/autocorr_drag.png}}  
+  \caption{\textbf{(a)} \ac{pdf} describing the statistics of the instantaneous drag fluctuations, computed on the basis of the control timeseries.
+\textbf{(b)} Autocorrelation function for the instantaneous drag $f_d$, defined as $(\mathbb{E}[f_d(t+\tau)f_d(t)]-\mathbb{E}[f_d]^2)/\sigma^2$, computed over the control run.
+It shows the exponential decay of the correlations over time.}
   \label{fig:pdf_drag}
@@
 In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
-In order to estimate its statistical properties, we performed a simulation of the dynamics over a duration $T_{tot} = 4\times 10^6\tau_c$.
+In order to estimate its statistical properties, we performed a simulation of the dynamics over a duration $T_{tot} = 4\times 10^6\tau_0$.
 In the following of this paper we refer to this long simulation as the \textit{control run}.
+Furthermore, we denote by $\bar{f}_d$ and $\sigma$ the average and variance computed over the control run, respectively.
 Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of the control timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ obtained from the control run.
@@
 Lastly, figure~\ref{fig:correlation_function} illustrates that the drag process has exponentially decaying time correlations.
-In the following, we call \textit{correlation time}, the typical time it takes for the drag process to decorrelate.
-We note it $\tau_c$.
-From figure~\ref{fig:correlation_function}, one can see that $\tau_c \approx 4\tau_0$.
+In the following we call \textit{correlation time}, denoted by $\tau_c$, the typical time it takes for the drag process to decorrelate.
+We observe that $\tau_c \approx \tau_0$.
 This means that the drag decorrelates after only a few number of eddies interacted with the square obstacle.
@@
 
-In this section we describe the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
+We now turn to a description of the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
 

### 2019-02-17 figures/PDF_drag/plot_pdf.py

@@ -10,7 +10,7 @@ n=len(buf)/2
 f = buf[0:n]
 xi = buf[n:]
 
-fig,ax = plt.subplots(1,1)
+fig,ax = plt.subplots(1,1, figsize=(7.5,5), constrained_layout=True)
 lineSQ, = ax.plot(xi,f)
 lineSQ.set_label('With obstacle')
 
@@ -21,9 +21,11 @@ xi = buf[n:]
 lineNOSQ, = ax.plot(xi,f, label='No obstacle')
 
 ax.set_yscale('log')
-ax.set_xlabel('$f_d$',fontsize=16)
-plt.ylabel('$\log(PDF)$',fontsize=16)
-ax.legend(loc='upper right')
+ax.set_xlabel('$f_d - \\bar{f}_d$',fontsize=22)
+plt.ylabel('$\log(PDF)$',fontsize=22)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+ax.legend(loc='upper right', fontsize=16)
 
 fname = 'PDF_drag.png'
 plt.savefig(fname)

### 2019-02-17 figures/autocorr_drag/plot_correlation.py

@@ -10,15 +10,17 @@ n=len(buf)/2
 tDomain = buf[0:n]
 c = buf[n:]
 
-fig,ax = plt.subplots(1,1)
+fig,ax = plt.subplots(1,1, figsize=(7.5,5), constrained_layout=True)
 ax.set_xlim(0, 3.5)
 line0, =ax.plot(np.array([0, 3.5]), np.array([0, 0]))
 lineC, = ax.plot(tDomain/2000,c)
 line0.set_linestyle('--')
 line0.set_linewidth(0.7)
 
-ax.set_xlabel(r'$\tau$',fontsize=16)
-plt.ylabel(r'$C(\tau)$',fontsize=16)
+ax.set_xlabel(r'$\tau$',fontsize=22)
+plt.ylabel(r'$C(\tau)$',fontsize=22)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
 
 fname = 'autocorr_drag.png'
 plt.savefig(fname)

### 2019-02-17 figures/return_time/plot_return_time.py
Updates return_time legend and figure

@@ -17,17 +17,19 @@ mu = 0.0252;
 sig = 0.0412;
 avec = (avec-mu)/sig;
 
-fig,ax = plt.subplots(1,1,figsize=(6,4),constrained_layout=True)
+fig,ax = plt.subplots(1,1,figsize=(7,3),constrained_layout=True)
 ax.set_xlim(10, 1e5)
 ax.set_ylim(2.7, 9)
 line, = ax.plot(r, avec)
 ax.set_xscale('log')
 #line.set_linewidth(.7)
 
-ax.set_xlabel(r'$r(f_d \geq a) / \tau_c$',fontsize=16)
-ax.set_ylabel('$a$',fontsize=16)
+ax.set_xlabel(r'$r(f_d \geq a) / \tau_c$',fontsize=22)
+ax.set_ylabel('$a$',fontsize=22)
 ax.set_yticks([3,5,7,9])
 ax.set_yticklabels((r'$3\sigma$',r'$5\sigma$',r'$7\sigma$',r'$9\sigma$'))
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
 
 fname = 'return_time.png'
 plt.savefig(fname)

### 2019-02-17 draft_p2.tex
Updates return_time legend and figure
Update plot and legend for pressure_ratio

@@
   \includegraphics[width=.6\linewidth]{return_time/return_time.png}
-  \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
+  \caption{Drag fluctuation amplitude as a function the corresponding return time. Here $f_d$ denotes the centered and reduced drag with $f_d=0$ and $\sigma = 1$.}
   \label{fig:return_time_instant}
@@
 \label{sec:forebody_and_base_contribution}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
+  \label{fig:pressure_ratio}
+  \caption{Contribution of the forebody and base pressure fluctuation to the overall drag fluctuation for the 104 sampled extreme drag fluctuations in the control run.
+To each of these fluctuation correspond a unique value of the drag $f^{\star}_d$, and a unique pair ($\tilde{p}^{\star}_{base}$,$\tilde{p}^{\star}_{fb}$).
+The dashed lines are first order least square fits of the data.
+They highlight that, as the value of the fluctuation increases, so is the relative contribution of the forebody pressure fluctuation.
+}
+\end{figure}
+
 In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
@@
 Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
-In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
+In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \bar{f}_d$.
 It can be further decomposed into 
@@
 \end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
+where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \bar{p}_{base}$ .
 Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
-\begin{figure}
-  \centering
-  \includegraphics{pressure_ratio/pressure_ratio.png}
-  \label{fig:pressure_ratio}
-\end{figure}
+
 It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the overall drag fluctuation.

### 2019-02-17 figures/pressure_ratio/plot_pressure_ratio.py

@@ -62,8 +62,11 @@ fitRear.set_linestyle('--')
 fitRear.set_linewidth(0.7)
 
 
-ax.set_xlabel(r'$\tilde{f}_d$',fontsize=16)
-ax.legend(loc='best', fontsize=16)
+ax.set_xlabel(r'$\tilde{f}^{\star}_d / \sigma$',fontsize=22)
+ax.legend(loc='best', fontsize=22)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+ax.set_xlim((7.5,12))
 
 fname = 'pressure_ratio.png'
 plt.savefig(fname)

### 2019-02-17 draft_p2.tex
Updates legend for illustr_extrms_vorticity, illustr_density_streamlines and timeseries_extremes

@@
   \includegraphics[width=.8\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-  \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
-The corresponding drag timeseries can be found in figure~\ref{fig:extremes_timeseries}.
-    The high value for the drag results from the formation of a strong negative (red) vortex inducing a pressure drop at the base of the obstacle.
-The formation of such a structure is aided by important vorticity production at the bottom boundary of the obstacle coupled with the influence of positive vorticity in the vicinity of the base.}
+  \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for some of the highest drag fluctuations in the control timeseries.
+    The extreme value for the drag results from the concentration of vorticity in the vicinity of the base of the obstacle, causing a pressure drop.}
 \end{figure}
@@
   \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
+  \label{density+streamlines}
+  \caption{Pressure field and velocity streamlines in the vicinity of the obstacle at $t=t^{\star}$.
+Becasue of the blocking vortex, vorticity produced along the top boundary of the obsatcle has no other choice but to accumulate at the base of the obstacle.}
 \end{figure}
@@
   \centering
-  \includegraphics[width=.5\linewidth]{timeseries_extremes/timeseries_extremes.png}
+  \includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
+  \caption{Average drag timeseries in the vicinity of an extreme drag fluctuation at $t=t^{\star}$.
+The blue continuous line results from an average over the 104 extreme events sampled in the control run.
+This illustrates that extreme drag variations are very local in time, with a typical lifetime of one correlation time unit $\tau_c$.}
 \end{figure}

### 2019-02-17 figures/timeseries_extremes/plot_force.py
Updates timeseries_extremes plotx

@@ -31,9 +31,11 @@ ax.set_yticklabels(['${:2.1f}\\sigma$'.format(y) for y in ax.get_yticks()], font
 ax.set_xlim((-2,2))
 plt.annotate('', xy=(-0.5,-0.35), xytext=(0.5,-0.35), arrowprops=dict(arrowstyle='<->',lw=2))
 plt.text(-0.1, -0.05, r'$\tau_c$', fontsize=18)
-ax.legend(loc='best',fontsize=12)
-ax.set_ylabel('$f_d(t)$',fontsize=16)
-ax.set_xlabel('$t$',fontsize=16)
+ax.legend(loc='best',fontsize=16)
+ax.set_ylabel('$f_d(t)$',fontsize=22)
+ax.set_xlabel('$t$',fontsize=22)
+plt.xticks(fontsize=18)
+plt.yticks(fontsize=18)
 
 fname = 'timeseries_extremes.png'
 plt.savefig(fname)

### 2019-02-17 draft_p2.tex
Removes dead code
Updates legend for dynamics_extremes and shear_asof_drag

@@
   \includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
+  \label{fig:timeseries_extremes}
   \caption{Average drag timeseries in the vicinity of an extreme drag fluctuation at $t=t^{\star}$.
@@
   \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
+  \label{fig:vorticity_dynamics}
+  \caption{Vorticity dynamics associated to an extreme fluctuation of the drag acting on the square obstacle at $t=t^{\star}$.}
 \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \begin{minipage}{0.85\linewidth}
-%     \begin{tabular}{c|c}
-%     \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_vorticity_a}\includegraphics[width=0.4\linewidth]{event_95_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_vort.png}} \\
-%     \hline
-%     \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_vort.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_vort.png}}
-%     \end{tabular}
-%   \end{minipage}
-% \begin{minipage}[h]{0.1\linewidth}
-%   \centering
-%   \includegraphics[width=\linewidth]{scale_vorticity.png}
-% \end{minipage}
-% \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
-The corresponding drag timeseries can be found in figure~\ref{fig:extremes_timeseries}.
-%   The high value for the drag results from the formation of a strong negative (red) vortex inducing a pressure drop at the base of the obstacle.
-The formation of such a structure is aided by important vorticity production at the bottom boundary of the obstacle coupled with the influence of positive vorticity in the vicinity of the base.}
-% \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \begin{minipage}[h]{0.85\linewidth}
-%     \begin{tabular}[h]{c|c}
-%     \subfloat[$f'_d = 11.6\sigma$]{\label{fig:top_4_events_streamlines_a}\includegraphics[width=0.4\linewidth]{event_95_streamlines.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 11.2\sigma$]{\includegraphics[width=0.4\linewidth]{event_69_streamlines.png}} \\
-%     \hline
-%     \subfloat[$f'_d = 10.6\sigma$]{\includegraphics[width=0.4\linewidth]{event_53_streamlines.png}} \hspace{0.5cm} & \hspace{0.5cm} \subfloat[$f'_d = 10\sigma$]{\includegraphics[width=0.4\linewidth]{event_20_streamlines.png}}
-%   \end{tabular}
-% \end{minipage}
-% \begin{minipage}[h]{0.1\linewidth}
-%   \centering
-%   \includegraphics[width=\linewidth]{scale_density.png}
-% \end{minipage}
-% \caption{\label{fig:top_4_events_streamlines} Density field at $t=t^{\star}$ for the four highest drag fluctuations in the control simulation for \tf{2}.
-Recall that the density $\rho$ is proportional to the pressure $p$, following the ideal gas state equation $p=c_s^2\rho$.
-See chapter~\ref{chap:test_flows} and appendix~\ref{app:lbm}.
-Additionally, velocity streamlines are depicted, representing advection by the flow at $t=t^{\star}$, see note~\ref{note:streamlines}.
-Blue areas indicate areas of lower pressure while red regions indicate regions of higher pressure.
-The formation of a vortex very close to the base boundary leads to a strong pressure drop.
-The downstream vortices originating from the top boundary layer separation are clearly visible and constrain the formation of the base vortex to a region very close to the base boundary.}
-% \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.1\linewidth]{scale_vorticity.png}
-%   \begin{tabular}[h]{cccc}
-%     \subfloat[$t = t^{\star}-0.54\tau_c$]{\label{frame_a}\includegraphics[width=.2\linewidth]{animation_event_95_0035}} & \subfloat[$t = t^{\star}-0.5\tau_c$]{\label{frame_b}\includegraphics[width=.2\linewidth]{animation_event_95_0036}} &     \subfloat[$t = t^{\star}-0.46\tau_c$]{\label{frame_c}\includegraphics[width=.2\linewidth]{animation_event_95_0037}} &     \subfloat[$t = t^{\star}-0.44\tau_c$]{\label{frame_d}\includegraphics[width=.2\linewidth]{animation_event_95_0038}} \\
-%     \subfloat[$t = t^{\star}-0.4\tau_c$]{\label{frame_e}\includegraphics[width=.2\linewidth]{animation_event_95_0039}} &     \subfloat[$t = t^{\star}-0.36\tau_c$]{\label{frame_f}\includegraphics[width=.2\linewidth]{animation_event_95_0040}} &     \subfloat[$t = t^{\star}-0.34\tau_c$]{\label{frame_g}\includegraphics[width=.2\linewidth]{animation_event_95_0041}} &     \subfloat[$t = t^{\star}-0.3\tau_c$]{\label{frame_h}\includegraphics[width=.2\linewidth]{animation_event_95_0042}} \\
-%     \subfloat[$t = t^{\star}-0.26\tau_c$]{\label{frame_i}\includegraphics[width=.2\linewidth]{animation_event_95_0043}} &     \subfloat[$t = t^{\star}-0.24\tau_c$]{\label{frame_j}\includegraphics[width=.2\linewidth]{animation_event_95_0044}} &     \subfloat[$t = t^{\star}-0.2\tau_c$]{\label{frame_k}\includegraphics[width=.2\linewidth]{animation_event_95_0045}} &     \subfloat[$t = t^{\star}-0.16\tau_c$]{\label{frame_l}\includegraphics[width=.2\linewidth]{animation_event_95_0046}} \\
-%     \subfloat[$t = t^{\star}-0.12\tau_c$]{\label{frame_m}\includegraphics[width=.2\linewidth]{animation_event_95_0047}} &     \subfloat[$t = t^{\star}-0.08\tau_c$]{\label{frame_n}\includegraphics[width=.2\linewidth]{animation_event_95_0048}} &     \subfloat[$t = t^{\star}-0.04\tau_c$]{\label{frame_o}\includegraphics[width=.2\linewidth]{animation_event_95_0049}} &     \subfloat[$t = t^{\star}$]{\label{frame_p}\includegraphics[width=.2\linewidth]{animation_event_95_0050}} \\
-%   \end{tabular}
-%   \caption{\label{fig:type_1_event}Flow dynamics corresponding to a particular type 1 event, which drag timeseries and flow fields at $t=t^{\star}$ are displayed in figures~\ref{fig:extremes_timeseries}, \ref{fig:top_4_events_vorticity} and~\ref{fig:top_4_events_streamlines}, respectively.
-In each frame the vorticity field is displayed, as well as the velocity streamlines, representing advection by the flow at each instant.
-See note~\ref{note:streamlines} for a discussion of streamlines.
-%     The maximum of the drag is attained for $t=t^{\star}$ and corresponds to frame~\ref{frame_p}.
-The sequence starts with a shear boundary layer forming over the top boundary of the obstacle (frames~\ref{frame_a} to~\ref{frame_d}).
-This boundary layer separates (frames~\ref{frame_e} to~\ref{frame_j}) and results in the formation of a large positive eddy in the vicinity of the base of the obstacle (frames~\ref{frame_k} and~\ref{frame_l}).
-In the meantime, frames~\ref{frame_j} to~\ref{frame_m} depict the formation of an attached strong shear layer at the bottom boundary.
-Eventually, frames~\ref{frame_n} to~\ref{frame_p} illustrate that this results in vorticity production very close to the base, aided by the large positive eddy originating from the top boundary layer separation.}
-% \end{figure}
 
@@
 
-Figure~\ref{fig:dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:illustr_density_streamlines}.
+Figure~\ref{fig:vorticity_dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:illustr_density_streamlines}.
 From $t^{\star}-0.4\tau_c$ to $t^{\star}-0.16\tau_c$, a strong negative eddy is generated along the bottom boundary layer of the obstacle and advected downstream by the mean flow.
 This leads to a region of negative vorticity at roughly one obstacle length of the base of the obstacle, constraining the positive vorticity generated along the top boundary.
-However, it can be seen in figure~\ref{fig:dynamics} that the positive vorticity only accumulates for a short duration.
+However, it can be seen in figure~\ref{fig:vorticity_dynamics} that the positive vorticity only accumulates for a short duration.
 The blocking negative vortex is in turn advected downstream, thus freeing the trapped positive vorticity.
@@
   \includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-  \caption{\label{fig:shear_asof_drag}}
+  \caption{\label{fig:shear_asof_drag} Evolution of the shear alongised the top or bottom boundary of the obsatcle as a function of the drag, for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
+To each oneof the grey lines correspond a unique extreme event.
+The blue line is the average over the 104 extreme event sampled in the control run.
+}
 \end{figure}

### 2019-02-17 draft_p2.tex
Update legend for shear_asof_drag and illustr_AMS

@@
 To each oneof the grey lines correspond a unique extreme event.
-The blue line is the average over the 104 extreme event sampled in the control run.
-}
+The blue line is the average over the 104 extreme event sampled in the control run.}
 \end{figure}
@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
+  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
+The initial ensemble of blue trajectories is first generated.
+These trajectories are independant and can be simulated in parallel.
+On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded from the ensemble (dashed blue line).
+Among the two remaining trajectories, trajectory 3 is chosen at random and copied until $Q_1$.
+It is then freely simulated from the branching poin to $T_a$.
+In case of deterministic dynamics, as samll perturbation is introduced at the branching.
+This procedure is typically iterated $J$ times, or until all trajectories go beyond a fixed threshold $Q$.}
 \end{figure}

### 2019-02-17 draft_p2.tex
Update position of plots illustr_AMS and AMS_OU and updates legend of AMS_OU as  well as explanation in main text

@@
 \label{sec:ams}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
+  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
+The initial ensemble of blue trajectories is first generated.
+These trajectories are independant and can be simulated in parallel.
+On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded from the ensemble (dashed blue line).
+Among the two remaining trajectories, trajectory 3 is chosen at random and copied until $Q_1$.
+It is then freely simulated from the branching poin to $T_a$.
+In case of deterministic dynamics, as samll perturbation is introduced at the branching.
+This procedure is typically iterated $J$ times, or until all trajectories go beyond a fixed threshold $Q$.}
+\end{figure}
+
 % Introduction to AMS
@@
 % Example: OU process
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
+The orange line represents the evolution of the maximum over resampled trajectories as a function of the computational cost $C_{TAMS}$ (iterations of the the selection-mutation step).
+The solid blue line is the analytical solution for the return time of amplitude $a$ for the \ac{ou} process~\cite{lestang_computing_2018}.}
+  \label{fig:comparaison_temps_de_retour}
+\end{figure}
 We first describe results of the \ac{tams} on a simple case.
@@
 The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
-Figure~\ref{fig:compraison_temps_de_retour} illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
+Figure~\ref{fig:compraison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach.
+In the latter, the typical computational cost is simply the return time $r(a)$.
+illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
 More importantly, these very rare trajectories are sampled for a computational cost that several order of magnitude lower than the return time of the corresponding fluctuations.
@@
 However, such modification(s) of the resampling procedure must be implemented with care, so as to preserve the mathematical properties of the \ac{ams} and \ac{tams}.
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
-The initial ensemble of blue trajectories is first generated.
-These trajectories are independant and can be simulated in parallel.
-On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded from the ensemble (dashed blue line).
-Among the two remaining trajectories, trajectory 3 is chosen at random and copied until $Q_1$.
-It is then freely simulated from the branching poin to $T_a$.
-In case of deterministic dynamics, as samll perturbation is introduced at the branching.
-This procedure is typically iterated $J$ times, or until all trajectories go beyond a fixed threshold $Q$.}
-\end{figure}
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
-The amplitude of fluctuations for the \ac{ou} process~\eqref{eq:OUprocess_chap_ams} are represented along the Y-axis.
-Along the X-axis is represented the typical computational cost required to sample fluctuations of the corresponding amplitude in both the \ac{tams} and direct sampling.
-In the latter, the typical computational cost is simply the return time $r(a)$.
-The computational cost associated to a level $a$ is the \ac{tams} is denoted by $C(a)$.
-The orange line represents the evolution of the level $\mathcal{Q}_j^{\star}$ as a function of the cost $C(a)$.
-The blue stars represent the maximum of $x(t)$ over the resampled trajectory \emph{after} it is resampled, that is at iteration $j+1$.
-Finally, the solid black line is the analytical solution for the return time of amplitude $a$~\cite{lestang_computing_2018}.}
-  \label{fig:comparaison_temps_de_retour}
-\end{figure}
 

### 2019-02-17 figures/AMS_OU/make_plot_validation_OU.py
Updates plot AMS_OU

@@ -19,20 +19,22 @@ n = len(buf)/2
 avec = buf[0:n]
 rtheo = buf[n:]
 
-fig,ax = plt.subplots(figsize=(6,4),constrained_layout=True)
+fig,ax = plt.subplots(figsize=(8,5),constrained_layout=True)
 #ax.plot(xx,amax/sig,marker='o',markersize=1,linestyle='None');
 lineLvl, = ax.plot(xx,Z/sig,linestyle='-',label='TAMS');
-ax.plot(rtheo,avec/sig, linestyle='-',label='Direct sampling')
+ax.plot(rtheo,avec/sig, linestyle='-',label='Direct sampling (return time)')
 plt.annotate('', xy=(12584,10), xytext=(1.7e19,10), arrowprops=dict(arrowstyle='<->',lw=2,color=lineLvl.get_color()))
-plt.text(1e7, 10.2, r'$C(a) = 10^{-17}r(a), a=10\sigma$', fontsize=12)
+plt.text(1e7, 10.2, r'$C(a) = 10^{-17}r(a), a=10\sigma$', fontsize=18)
 ax.set_xscale('log')
-ax.set_xlabel(r'$r(f_d \geq a), C_{TAMS}(f_d \leq a)$',fontsize=16)
-ax.set_ylabel('$a$',fontsize=16)
+ax.set_xlabel(r'$r(f_d \geq a), C_{TAMS}(f_d \geq a)$',fontsize=22)
+ax.set_ylabel('$a$',fontsize=22)
 ax.set_yticks([0,2,4,6,8,10])
 ax.set_xticks([1e2,1e8,1e14,1e20])
 ax.set_xticklabels((r'$10^2\tau_c$',r'$10^8\tau_c$',r'$10^{14}\tau_c$',r'$10^{20}\tau_c$'))
 ax.set_yticklabels((r'$0$',r'$2\sigma$',r'$4\sigma$',r'$6\sigma$',r'$8\sigma$',r'$10\sigma$'))
-ax.legend(loc='best')
+ax.legend(loc='best', fontsize=18)
+plt.xticks(fontsize=18)
+plt.yticks(fontsize=18)
 
 fname = 'AMS_OU.png'
 plt.savefig(fname)

### 2019-02-17 draft_p2.tex
Updates plot and legends for AMS_drag_trajectories and AMS_drag_resampling

@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
+  \label{fig:illustr_AMS}
   \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
@@
 % Commentaire figure instant_256_20
-
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
+    \caption{Maximum of the instantaneous drag along the resampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
+For these particular dynamics, the \ac{tams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation in the initial ensemble.}
+\end{figure}
 We apply the \ac{tams} with $N=256$ initial trajectories, with a duration $T_a = 20 \tau_c$.
 Similarly to figure~\ref{fig:comparaison_temps_de_retour}, figure~\ref{fig:instant_256_20} displays the maximum drag achieved by resampled trajectories.
-Figure~\ref{fig:instant_256_20} also shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
+In addition, figure~\ref{fig:instant_256_20} shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
 Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
@@
 % Commentaire de la figure iter_181
-
-Figure~\ref{fig:instant_256_20} can be better understood by visualising the drag signals corresponding to the last resampled trajectories.
-Figure~\ref{fig:instant_256_20_b} displays the ensemble of trajectories after $181$ iterations of the \ac{tams} resampling procedure.
-Comparison with figure~\ref{fig:instant_256_20_b} suggests that, utimately, trajectories are resampled from the same initial trajectory.
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
+  \label{fig:AMS_drag_trajectories}
+  \caption{Ensemble of trajectories after $181$ iterations of the selection-mutation procedure.In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
+Trajectories have a duration $T_a = 5\tau_c$ and their number is $N_c = 32$.
+    The \ac{tams} fails to efficiently sample rare trajectories, because all trajectories are ultimately resampled from the trajectory displaying the highest maximum in the initial ensemble.}
+\end{figure}
+Figure~\ref{fig:AMS_drag_trajectories} can be better understood by visualising the drag signals corresponding to the last resampled trajectories.
+Figure~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after $181$ iterations of the \ac{tams} resampling procedure (see figure~\ref{fig:illustr_AMS}).
+One can see that, utimately, trajectories are resampled from one unique initial trajectory.
 This phenomenon can be explained as follows.
@@
 However, such modification(s) of the resampling procedure must be implemented with care, so as to preserve the mathematical properties of the \ac{ams} and \ac{tams}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-    \caption{Maximum of the instantaneous drag along the resampled trajectories, denoted by $\tilde{f}_d^{(j)}$, as a function of the corresponding computational cost $C_j$, defined in~\eqref{eq:comp_cost_iterations}.
-In both figures, the \ac{tams} is iterated until extinction.
-    The solid black line indicates the typical fluctuation amplitude $a$ as a function of the corresponding  \textit{return time} $r(a)$.
-It is the typical timescale of occurrence of fluctuations above $a$.
-Furthermore, the square markers depict the maximum $\tilde{f}_{d,n}^{(0)},\,\, 1\leq n \leq N_c$ over the initial trajectories.
-    This figure illustrates that in both experiments the \ac{tams} does not improve the sampling of rare fluctuations.
-On can see that the maximum values among the resampled trajectories concentrates on values already sampled over the initial trajectories.
-Eventually, extinction is reached as the algorithm is unable to generate a fluctuation higher than the maximum fluctuation in the initial ensemble of trajectories.
-    This behaviour is illustrated in figure~\ref{fig:extinction_32_5} on the basis of the drag timeseries involved in the experiment of figure~\ref{fig:comparaison_ams_return_time_instant_a}.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-    \caption{Visualisation of the drag timeseries corresponding to the set of trajectories in the ensemble, at different iterations along the algorithm.
-In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
-Trajectories have a duration $T_a = 5\tau_c$ and their number is $N_c = 32$.
-    One can see that, as the number of iterations increases, the number of independent resampled trajectories drastically decreases until extinction at iteration $181$.
-The timeseries corresponding to the two initial trajectories exhibiting the highest fluctuations are displayed in colour.
-In all four figures, colour is inherited from the parent trajectory.
-It illustrates that the two strongest initial trajectories quickly become predominant in the ensemble, as the other are discarded.
-    As can be seen in figure~\ref{fig:subfigc}, trajectories only separate well after the maximum is reached.
-Therefore, resampling the blue trajectories based on the red trajectories can only change the overall maximum by a very small amount.
-Consequently, blue trajectories are discarded one after the other, each time resampled into another instance of the red trajectory.
-This new instance has roughly the same maximum as the others.}
-\end{figure}
 

### 2019-02-17 figures/AMS_drag_resampling/make_plot_256.py

@@ -32,7 +32,7 @@ Nresamp = len(amax)
 Cvec = Nc*Ta*np.ones(Nresamp) + Ta*np.arange(1,Nresamp+1)
 
 # Init. figure
-fig,ax = plt.subplots(figsize=(9,5), constrained_layout=True)
+fig,ax = plt.subplots(figsize=(8,5), constrained_layout=True)
 
 # Gathers the maxima over the initial ensemble
 maxArray = np.zeros(Nc)
@@ -58,12 +58,12 @@ resamp.set_markeredgecolor(resamp.get_markerfacecolor())
 resamp.set_label('Maxima over resampled trajectories')
 
 ax.set_xscale('log')
-ax.set_xlabel(r'$C_{TAMS} / \tau_c$',fontsize=16)
-ax.set_ylabel(r'$f_d^{\star} / \sigma$',fontsize=18)
-ax.tick_params(axis="x", labelsize=16)
-ax.tick_params(axis="y", labelsize=16)
+ax.set_xlabel(r'$C_{TAMS} / \tau_c$',fontsize=22)
+ax.set_ylabel(r'$f_d / \sigma$',fontsize=22)
+ax.tick_params(axis="x", labelsize=18)
+ax.tick_params(axis="y", labelsize=18)
 
-ax.legend(loc=7, fontsize=12)
+ax.legend(loc='best', fontsize=18)
 
 fname = 'AMS_drag_resampling.png'
 plt.savefig(fname)

### 2019-02-17 figures/AMS_drag_trajectories/plot_pre_extinct.py

@@ -1,5 +1,5 @@
 import numpy as np
-1;4601;0cimport matplotlib.pyplot as plt
+import matplotlib.pyplot as plt
 
 plt.style.use('ggplot')
 
@@ -26,7 +26,7 @@ Ta = 5;
 tspan = np.linspace(0,5,5000)
 maxArray=np.zeros(Nc)
 # Init. figure
-fig,ax = plt.subplots(figsize=(9,5), constrained_layout=True)
+fig,ax = plt.subplots(figsize=(8,5), constrained_layout=True)
 
 for i in range(0,Nc):
     x = np.fromfile(prefix+'traj_{:d}_costFunc_iter_0.dat'.format(i))
@@ -50,12 +50,12 @@ for t in range(1,182):
         x=(x-m)/sig
         ax.lines[ktIdx].set_ydata(x)
 
-ax.set_xlabel(r'$t / \tau_c$',fontsize=16)
-ax.set_ylabel(r'$f_d(t) / \sigma$',fontsize=18)
-ax.tick_params(axis="x", labelsize=16)
-ax.tick_params(axis="y", labelsize=16)
+ax.set_xlabel(r'$t / \tau_c$',fontsize=22)
+ax.set_ylabel(r'$f_d(t) / \sigma$',fontsize=22)
+ax.tick_params(axis="x", labelsize=18)
+ax.tick_params(axis="y", labelsize=18)
 
-ax.legend(loc='best', fontsize=12)
+ax.legend(loc='best', fontsize=18)
 ax.set_xlim((0,5))
 
 fname = 'AMS_drag_trajectories.png'

### 2019-03-20 draft_p2.tex
Updates discussion of results for GKTL

@@
 
-\subsubsection{Application to extreme fluctuations of the time-averaged drag}
+\subsubsection{Application of the \ac{GKTL} algorithm to extreme fluctuations of the time-averaged drag}
 \label{sec:gktl_drag}
@@
 
-% Choice of parameters and perturbation
-The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
+% Computational cost and subsequent choice of parameters
+The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N$: $C_{gktl} = N_c \times T_a$.
 In the following series of experiments we set $T_a = T = 10\tau_c$.
-The sampling from the \ac{gktl} algorithm will be compared to the sampling from the control run, \textit{i.e.} a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
-We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-(\textbf{Plus vraiment clair pour moi pourquoi je n'ai pas utilisé 10X plus de clones, pour matcher le coût numérique de la simulation de contrôle...})
-Lastly, one must set the cloning period $\tau$.
+The number of the clones is set to $N=16384$.
+As a consequence, the computational cost is $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+Last but not least, one must set the cloning period $\tau$.
 On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
@@
 Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
-For each of the three experiments, it displays the estimate of the biased \ac{pdf}
+For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
 \begin{equation}
   \label{eq:estimate_biased_measure}
-  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
-  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
+  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
+  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
 \end{equation}
@@
 Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
-Note that it corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.
+This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\\ref{sec:pdfs}).
+Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
 
@@
 
-For a fixed number of trajectories $N$, there is a limited range of $k$ values for which the algorithm is able to yield an accurate estimate of $\rho_k$.
-Setting $k$ to too high a value causes harmful effects resulting from the finiteness of the ensemble of trajectories: trajectories overlap over most of their duration.
+For a fixed number of trajectories $N$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite
+size effects harm the efficiency and accuracy of the selection procedure.
+Typically, at each selection step, the majority of the clones are resampled, with the expection of only one or two.
 This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
-This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, therefore leading to the same value for the averaged drag $F_T$.
-See~\cite{ref1,ref2,ref3} for further details about the scaling of the required population size as $k$ is increased.
+This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, because the resampling of
+discarded trajectories is based on a very small subset of the ensemble of $N$ trajectories.
+Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the
+bias $k$ is increased.
+An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_max$ as
+a function of the ensemble size $N$. 
+See~\cite{ref1,ref2,ref3} for further details.
+
+For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the
+diversity of the trajectories in the biased ensemble.
+This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.
 
@@
 % qu'un échantillonnage direct
-Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
-Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
+Sampling larger fluctuations therefore requires a larger population size, and a larger computational cost.
+Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to extreme fluctuations of the time-averaged drag that could
 not be sampled through a direct sampling with equal computational cost.
-Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
+For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$.
+Note that trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.} are counted as one.  
 As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
-This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
+To estimate this number we assumed that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
 This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
-While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
+Even though such a Gaussian assumption does not hold for fluctuations $F_T \geq 4\sigma_T$,
+it may still provide numbers relevant to the comparison with the algorithm.
 \begin{figure}

### 2019-04-13 draft_p2.tex
Overwrite intro of part IV.B. (GKTL) by intro of part IV.B.1. (was redundant)

--- 2019-02-17 draft_p2.tex
@@
 For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
-It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
-averaging time~\cite{ref_large_dev}.
-By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
-Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
-The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
-The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.
-
-\subsubsection{The \ac{gktl} algorithm}
-\label{sec:gktl_description}
 Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
@@
 Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
+% It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
+% averaging time~\cite{ref_large_dev}.
+% By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
+% Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
+% The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
+% The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.
 
+\subsubsection{The \ac{gktl} algorithm}
+\label{sec:gktl_description}
 The \ac{gktl} algorithm consists in simulation an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{1 \leq n \leq N}$ starting from independant random initial conditions.

### 2019-04-14 draft_p2.tex
Fusionne la version d Emmanuel apres la discussion du 21/03

@@
 
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic fluid motions associated with enormous internal forces.
+Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces.
 In the turbulence literature, ``extreme'' can refer to fluctuations that deviate from the mean value by $O(10)$ standard deviations.
@@
 //
-The nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
+In an engineering context, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
 \\
@@
 The use of perturbative methods in identifying resonant interactions (among degrees of freedom) leading to extreme fluctuations is not successful.
-In this context, simulation offers a practical means to gain insight into the extreme events. However, their numerical sampling requires long simulation since these events are rare as well. 
+In this context, simulation offers a practical means to gain physical insight into these events and estimate their frequency of occurrence, or return times.
+Nevertheless, this requires very long simulation since these events are also very rare.
+It is here our motivation to evaluate the practicality and efficiency of importance sampling techniques to sample extreme events in developed turbulent flows.
 \\
@@
 
-%% The phenomenology of typical drag fluctuations
-The typical timescale of variation of the drag therefore corresponds to the timescale of pressure fluctuations in the vicinity of the obstacle.
-An estimate of such timescale is the \textit{turnover time}
+%% Definition du turnover time
+The typical timescale of drag fluctuations may be estimated from dimensional analysis as
 \begin{equation}
@@
 where $R$ is the diameter of the cylinder and $U$ the mean-flow velocity.
-Figure~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle over a few thousands turnover times.
-The evolution over a few turnover times of the vorticity field around the obstacle is displayed in figure~\ref{fig:typical_vorticity}.
-One can see that the interaction of the incoming flow with the square obstacle primarily results in boundary layers along the top and bottom boundaries.
-However, the vorticity corresponding to these boundary layers only have a small impact on the the pressure field in the vicinity of the obstacle, because it is rapidly advected dowmstream by the mean-flow.
+Let us mention that the viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+
+% Description du signal de trainee typique
+Figure~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle over a one thousands turnover times.
+The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+
+% Description du champ de vort. correspondant a une fluctuation typique
+Figure~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle over a few turnover times.
+The vorticity generated along the boundaries of the obstacle is advected downstream by the mean flow.
 As a consequence, the base pressure varies much less than the forebody pressure, and typical drag fluctuations mostly result from the fluctuating pressure at the front of the obstacle.
 
-In the following section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
-In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a smaller role.
+In the section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
+In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a lesser role.
 
@@
   \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}.
-The correlation time $\tau_c$ is defined further in this text and correspond to the typical fluctuation timescale of the drag.}
+  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
+The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
   \label{fig:typical_drag_signal}
 \end{figure}
+
 \begin{figure}
@@
   \includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-  \caption{Vorticity dynamics corresponding to typical drag fluctuations over one correlation time unit $\tau_c$.
+  \caption{Vorticity dynamics corresponding to typical drag fluctuations over four turnover times~\eqref{eq:turnover_time}.
 Negative vorticity is generated along the bottom boundary, and advected away from the the obstacle by the mean flow.}
@@
   \centering
-    \subfloat[]{\label{fig:pdf_drag_a}\includegraphics[width=.5\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[]{\label{fig:pdf_drag_b}\includegraphics[width=.5\linewidth]{autocorr_drag/autocorr_drag.png}}  
-  \caption{\textbf{(a)} \ac{pdf} describing the statistics of the instantaneous drag fluctuations, computed on the basis of the control timeseries.
-\textbf{(b)} Autocorrelation function for the instantaneous drag $f_d$, defined as $(\mathbb{E}[f_d(t+\tau)f_d(t)]-\mathbb{E}[f_d]^2)/\sigma^2$, computed over the control run.
-It shows the exponential decay of the correlations over time.}
+    \subfloat[\ac{pdf} of drag fluctuations around mean.]
+    {\label{fig:pdf_drag_a}
+    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
+    \subfloat[Autocorrelation of drag.]
+    {\label{fig:pdf_drag_b}
+    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
+      \caption{\textbf{(a)} Estimate of the zero-mean \ac{pdf} for drag fluctuations $f'_d = (f_d - \bar{f_d})$ where $\bar{f_d}$ denotes the averaged value.
+The drag is evaluated both in the presence and absence of the obstacle.
+Note that both \acsp{pdf} have different variance, and are here not normalized.
+  \textbf{(b)} Autocorrelation function of the drag defined as $\left( \overline{f_d(t+\tau)f_d(t)}-\overline{f_d}^2 \right) /\sigma^2$.
+The lag $\tau$ is normalized by the turnover-time~\eqref{eq:turnover_time}.}
   \label{fig:pdf_drag}
@@
 
-In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
-In order to estimate its statistical properties, we performed a simulation of the dynamics over a duration $T_{tot} = 4\times 10^6\tau_0$.
-In the following of this paper we refer to this long simulation as the \textit{control run}.
-Furthermore, we denote by $\bar{f}_d$ and $\sigma$ the average and variance computed over the control run, respectively.
-Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of the control timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ obtained from the control run.
-One can see that the tails of this distribution, that describe the probability of very rare events, can be very well modelled by an exponential \ac{pdf} of the form $p(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
-In addition, figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} of the process describing the drag acting on a control surface equivalent to the surface of  the obstacle, in a flow \textit{without obstacle}.
-This different setup is illustrated in figure~\ref{fig:no_square}.
-In this case, the corresponding \ac{pdf} estimate for the drag process is quasi-symmetric and, more importantly, does not display exponential tails.
-We conclude that the interation of the flow with the square obstacle is responsible for the exponential tails of the \ac{pdf} of the drag process.
-In section~\ref{sec:time_avg}, we relate this very particular shape of the \ac{pdf} describing extreme drag fluctuations to the phenomenology of extreme fluctuations of the time-average of the drag process.
+Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
+The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+It is therefore natural to model the drag as a scalar random process $f_d(t)$.
 
-Lastly, figure~\ref{fig:correlation_function} illustrates that the drag process has exponentially decaying time correlations.
-In the following we call \textit{correlation time}, denoted by $\tau_c$, the typical time it takes for the drag process to decorrelate.
-We observe that $\tau_c \approx \tau_0$.
-This means that the drag decorrelates after only a few number of eddies interacted with the square obstacle.
+Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
+The \ac{pdf} of drag fluctuations deviates from a normal law and shows an exponential tail for large positive fluctuations, i.e.  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
+%
+Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
+%
+Lastly, the autocorrelation function of the drag is shown in Fig.~\ref{fig:pdf_drag_b}.
+It is found drag fluctuations are correlated over a time scale of the order of $\tau_0$, suggesting that the drag signal looses its memory typically after the sweeping of a few eddies past the obstacle.
+%
+This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
+%
+In the following we define the \textit{correlation time} $\tau_c = 4\tau_0$.
+Note that the ratio $tau_0 / tau_c$ may be viewed as a \textit{Strouhal number} $St$.
+A value $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
 
-% \begin{equation}
-%   \label{eq:correlation_def}
-%   C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
-% \end{equation}
-
-% As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.
-
-% From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
-% This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
-
-\section{Direct sampling of extreme fluctuations of the drag acting on a square cylinder in two-dimensional turbulence}
+\section{Extreme fluctuations of the drag by means of direct sampling}
 \label{sec:direct_sampling}
 
-We now turn to a description of the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
+In this section we describe the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
 
@@
   \includegraphics[width=.6\linewidth]{return_time/return_time.png}
-  \caption{Drag fluctuation amplitude as a function the corresponding return time. Here $f_d$ denotes the centered and reduced drag with $f_d=0$ and $\sigma = 1$.}
+  \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
   \label{fig:return_time_instant}
 \end{figure}
+
 \subsection{Instantaneous drag}
@@
 \label{sec:forebody_and_base_contribution}
+In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
+We now investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuations, in the case of the sampled \textit{extreme} drag fluctuations.
+Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
+In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
+It can be further decomposed into 
+\begin{equation}
+  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
+\end{equation}
+where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
 
+Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
 \begin{figure}
@@
 \end{figure}
-
-In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
-We now investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuations, in the case of the sampled \textit{extreme} drag fluctuations.
-Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
-In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \bar{f}_d$.
-It can be further decomposed into 
-\begin{equation}
-  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
-\end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \bar{p}_{base}$ .
-Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
-
 It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the overall drag fluctuation.
@@
   \includegraphics[width=.8\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-  \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for some of the highest drag fluctuations in the control timeseries.
-    The extreme value for the drag results from the concentration of vorticity in the vicinity of the base of the obstacle, causing a pressure drop.}
+  \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
+    The high value for the drag results from the generation of an intense vortex inducing a pressure drop at the base of the obstacle.
+The formation of such a structure is aided by important negative (positive) vorticity production at the bottom (top) boundary of the obstacle coupled with the influence of positive (negative) vorticity in the vicinity of the base.}
 \end{figure}
@@
   \caption{Pressure field and velocity streamlines in the vicinity of the obstacle at $t=t^{\star}$.
-Becasue of the blocking vortex, vorticity produced along the top boundary of the obsatcle has no other choice but to accumulate at the base of the obstacle.}
+Because of the blocking vortex, vorticity produced along the top boundary of the obsatcle has no other choice but to accumulate at the base of the obstacle.}
 \end{figure}

### 2019-04-14 draft_p2.tex
Moves figure typical_drag_signal and description to part II.B. The Drag as a Random Process

@@
 
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle over a one thousands turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-
 % Description du champ de vort. correspondant a une fluctuation typique
@@
 In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a lesser role.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
-The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
-  \label{fig:typical_drag_signal}
-\end{figure}
 
@@
 
+% Description du signal de trainee typique
+Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
+The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+It is therefore natural to model the drag as a scalar random process $f_d(t)$.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
+  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
+The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
+  \label{fig:typical_drag_signal}
+\end{figure}
+
+Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
+The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
+%
+Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
 \begin{figure}
@@
 \end{figure}
-
-Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-It is therefore natural to model the drag as a scalar random process $f_d(t)$.
-
-Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
-The \ac{pdf} of drag fluctuations deviates from a normal law and shows an exponential tail for large positive fluctuations, i.e.  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
 %

### 2019-04-15 draft_p2.tex
Fix typos

@@
 In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $tau_c$.
+In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $\tau_c$.
 However, In many applications, this duration is much smaller than the timescale of interest.
-Consider for insatnce the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
+Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
 
@@
 %% Objectif des algo.
-In this section we discuss the application of \textbf{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
+In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
 of the flow presented in section~\ref{sec:test_flow}.
@@
 However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
-As a consequence the choice of the score function $xi (\mathbf{x}(t),t)$ is not so clear.
+As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not so clear.
 In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.

### 2019-04-15 draft_p2.tex
Minor modifications in part IV.B.2 results GKTL

@@
 The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
-Consistently with section~\ref{sec:time_avg}, we choose the averaging window $T=10\tau_c$.
+Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.
 
@@
 The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
-In the following series of experiments we set $T_a = T = 10\tau_c$.
-The sampling from the \ac{gktl} algorithm will be compared to the sampling from the control run, \textit{i.e.} a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
+In the following we set $T_a = T = 10\tau_c$.
 We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-(\textbf{Plus vraiment clair pour moi pourquoi je n'ai pas utilisé 10X plus de clones, pour matcher le coût numérique de la simulation de contrôle...})
 Lastly, one must set the cloning period $\tau$.
@@
 On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
-A rule of thumb is to set $\tau \approx \tau_c$.
-In the following, we set $\tau = \tau_c /2$.
-
-% Commentaire de la figure illustration_IS
+As a consequence, a rule of thumb is to set $\tau \approx \tau_c$.
+In the following, $\tau = \tau_c /2$.
 
@@
 In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
-% In the large time limit $T_a\to \infty$, the typical level of fluctuations $\mathbb{E}_{\mathbb{P}_k}[ F_{T_a} ]$ within the biased ensemble of trajectories
-% can be related to the \textit{Scaled Cumulant Generating Function} $\lambda(k)$~\cite{ref}.
+Each experiment was run using $512$ Intel processors, each processor being assigned the computation of $N_c / 512 = 32$ clones.
+The walltime for one experiment was roughly two days.
 
+% Commentaire de la figure IS_GKTL
 Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
@@
 Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
-Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
+Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
 not be sampled through a direct sampling with equal computational cost.

### 2019-04-15 draft_p2.tex

@@
 % qu'un échantillonnage direct
-Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
-Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
-not be sampled through a direct sampling with equal computational cost.
-Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
+Sampling larger fluctuations therefore requires a larger population size, and a larger computational cost.
+Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to extreme fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
+For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$.
+Note that trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.} are counted as one.  
 As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
-This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
+To estimate this number we assumed that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
 This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
-While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
+Even though such a Gaussian assumption does not hold for fluctuations $F_T \geq 4\sigma_T$,
+it may still provide numbers relevant to the comparison with the algorithm.
 \begin{figure}

### 2019-04-15 draft_p2.tex
Fine grained updates for part IV.B.2 Results GKTL

@@
 
-\subsubsection{Application of the \ac{GKTL} algorithm to extreme fluctuations of the time-averaged drag}
+\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
 \label{sec:gktl_drag}
@@
 In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
-Each experiment was run using $512$ Intel processors, each processor being assigned the computation of $N_c / 512 = 32$ clones.
-The walltime for one experiment was roughly two days.
 
 % Commentaire de la figure IS_GKTL
-Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
+Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
 For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
@@
 Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
-This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\\ref{sec:pdfs}).
+This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\ref{sec:pdfs}).
 Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
@@
 
-For a fixed number of trajectories $N$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite
-size effects harm the efficiency and accuracy of the selection procedure.
-Typically, at each selection step, the majority of the clones are resampled, with the expection of only one or two.
-This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
-This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, because the resampling of
-discarded trajectories is based on a very small subset of the ensemble of $N$ trajectories.
-Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the
-bias $k$ is increased.
-An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_max$ as
-a function of the ensemble size $N$. 
+For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
+For $k \gtrapprox k_{max}$ the majority of the clones are resampled, with the expection of only one or two.
+Because the resampling is based on a very small subset of the ensemble of $N_c$ trajectories, most of the trajectories in the biased ensemble overlap.
+This effect is visible in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
+Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
+An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
 See~\cite{ref1,ref2,ref3} for further details.
 
-For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the
-diversity of the trajectories in the biased ensemble.
+For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
 This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

### 2019-04-16 draft_p2.tex
Finished figure timeseries_extrms_AVG and added it to the paper

@@
 Indeed, both cases are featured in figure~\ref{fig:extreme_avg}, along with intermediate cases where the very large value of the drag results from both a very large fluctuation and a large number of typical, however all positive, fluctuations of the instantaneous drag $f_d$.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+  \caption{Instantaneous drag timeseries corresponding to the highest fluctuations of the time-averaged drag $F_T$ in the control timeseries. The averaging window is $T=10\tau_c$.
+  Here $f_d$ denotes the zero-mean, unit variance instantaneous drag. The time-average is normalized by the standard deviation for the time-averaged drag $\sigma_T$.}
+  \label{fig:extreme_avg}
+\end{figure}
 

### 2019-04-16 figures/timeseries_extrms_AVG/sort_extremes.py

@@ -27,20 +27,7 @@ for i in range(0,Nevents):
     fileName=prefix+'event_{:d}_AVG/'.format(i+1)+'data_force.datout'
     # Reads instant. drag
     f = np.fromfile(fileName, float, -1, "")
-
-    # Substract mean value
-    f = f-m
-    # Computes average
     F[i] = np.mean(f)
-    # Records min and max over signal, normalized to std dev
-    maxArray[i]=np.amax(f)/sig
-    minArray[i]=np.amin(f)/sig
-    
-globalDragMaximum = np.amax(maxArray)
-globalDragMinimum = np.amax(minArray)
-
-print globalDragMaximum
-print globalDragMinimum
 
 np.savetxt('list.txt', np.argsort(-F), '%0.2d')
 

### 2019-04-16 figures/timeseries_extrms_AVG/timeseries_extrms_AVG.py

@@ -1,35 +1,68 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
+plt.style.use('ggplot')
 
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/' \
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/' \
          + 'averaged_drag/'
 
-F = np.loadtxt('list.txt', dtype=int)
+# list.txt is generated with ./sort_extremes.py
+# contains the sorted list of extremes according to the
+# time-averaged drag (highest value first)
+sortedI = np.loadtxt('list.txt', dtype=int)
+
 sig=0.0412
 m=0.0252
-
-Nevents = len(F)
-
+# Gets the value of mu and sigma for time-averaged drag
 musigma = np.fromfile("../PDF_AVG/musigma_AVG_10.dat", float, -1, "");
 
-fig, axes = plt.subplots(nrows=2,ncols=3, figsize=(18,6))
+T = 10 # Physical duration of instant. trajectories
 
-for i in range(0,6):
-    fileName=prefix+'event_{:d}_AVG/'.format(F[i]+1)+'data_force.datout'
+Nsubplots = 6
+minArray=np.zeros(Nsubplots) # containers for max and min over instant. drag
+maxArray=np.zeros(Nsubplots) # timeseries
+fig, axes = plt.subplots(nrows=2,ncols=3, figsize=(18,6), constrained_layout=True)
+for i in range(0,Nsubplots):
+    fileName=prefix+'event_{:d}_AVG/'.format(sortedI[i]+1)+'data_force.datout'
     f = np.fromfile(fileName, float, -1, "")
-    f = f-m
-    amp = np.sum(f)/(len(f)-1)/musigma[1]
-    f = f/sig
+    # amplitude of fluctuation for time-average
+    F_avg = np.sum(f)/(len(f)-1)
+    F_avg = (F_avg-musigma[0])/musigma[1]
+    # Compute centered and reduced drag
+    # important to do this *after* computing F_avg
+    f = (f-m)/sig
     N = len(f)
+    # row and col are computed so that plots go from left
+    # to right, in descending order of F_avg
     row = i // 3
     col = i % 3
-    axes[row,col].plot(np.linspace(0,10,N),f)
-    axes[row, col].set_xlim((0, 10))
[…]

### 2019-04-18 figures/PDF_AVG/plot_pdf_AVG.py
Minor modif to PDF_AVG figure

@@ -32,7 +32,8 @@ line.set_label('PDF of $F_T$')
 lineGauss, = ax.plot(xi,ret_gaussian(xi), label='Gaussian PDF')
 
 # Plots the local least-square fit for the positive tail
-lineFit, = ax.plot(xiTrunc,np.exp(p(xiTrunc)))
+lineFit, = ax.plot(xiTrunc,np.exp(p(xiTrunc)), linewidth=2, \
+                   color='k')
 
 # Set figure parameters
 ax.set_yscale('log')

### 2019-04-18 figures/table_nb_traj/table_nb_traj_fit.py
Add script to generate estimate of number of events in GKTL runs and according to direct approach with fit of PDF

@@ -1,0 +1,122 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy import special,integrate
+
+def ret_gaussian(x):
+    """ Returns zero-mean, unit variance at point
+    """
+    return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
+
+def integrate_composite_pdf(borneInf, borneSup, pol, norm_factor=1):
+    """Return value of integral of the piecewise function
+         f(x) = f_gauss(x) for x <=3
+                f_fit(x)   for x > 3
+       where f_gauss is the zero-mean, unit variance Gaussian PDF and 
+       f_fit is the exponential of a polynomial pol
+
+    Parameters
+    ----------
+    borneInf : float
+    borneSup : float
+    pol : poly1d
+    norm_factor : float, optional
+
+    Returns
+    -------
+    integral : float
+    """
+    xvec = np.linspace(borneInf,borneSup,1000)
+    dx = xvec[1]-xvec[0]
+    gaussian_part = ret_gaussian(xvec[xvec<=3])
+    non_gaussian_part = np.exp(p(xvec[xvec>3]))
+    full_pdf = np.append(gaussian_part,non_gaussian_part)
+
+    return integrate.trapz(full_pdf,xvec,dx)/norm_factor
+
+Nc = 16384 # Nb of total traj
+NN = 20 # Nb of cloning steps
+sig = 0.0087714 # Std dev of time-averaged drag (T=10\tau_c)
+
+# List containing the names of the result directories
+directories = ['k_0.02_2018_03_26_10_01_46/', 'k_0.025_2018_03_25_20_13_40/',
+	   'k_0.03_2018_03_25_20_13_40/']
+prefix = '/home/tlestang/transition_hdd/thibault/These/tailleur_lecomte/lbm/2_sq/' \
+         + 'grid/illustration_IS/'
+
+# Open a text file
+fo = open("foo.txt", "w")
+for directory in directories:
+    # Read labels and average drag
+    # labels is a Nc*NN vector containing the index of ancestors
+    # for the next evolution step
+    # Ex: Ancestor for traj. 1 after third step is
+    # labels[2*Nc + 1]
+    #          |
+    #    Ancestors for first 2 steps run from 0 to 2*Nc-1   
+    full_path = prefix+directory+'labels.datout'
+    labels = np.fromfile(full_path, np.int32, -1, "")
+    full_path = prefix+directory+'A.datout'
[…]

### 2019-04-19 draft_p2.tex
Update table I in draft_p2.tex with new values from the least square fit of PDF_AVG
Add figure IS_GKTL and add it to .tex file

@@
     $k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-    $\mathcal{N}_{direct}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005  \\ 
+    $\mathcal{N}_{direct}$ & 2601.258& 374.862& 24.268& 0.958& 0.000\\ 
   \end{tabular}
@@
   \centering
-%  \includegraphics[width=0.7\linewidth]{illustration_IS}
-  \caption{\textbf{(a)}: Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 
+  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
+  \caption{Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 
   The \ac{gktl} algorithm is used with $N_c = 16384$ copies and $T_a = T= 10\tau_c$ for three increasing values of the bias $k$.
-Additionally, the cloning period is $\tau = \tau_c/2$, resulting in $M=20$ iterations of the evolution/selection procedure.
-Note that in this particular example the last cloning stage at $t_{19}$ is not performed.
-  As $k$ is increased, the sampled \ac{pdf} is shifted towards extreme values.
-  Along the X-axis we display the fluctuation $F'_{T} = F_{T} - \overline{F_{T}}$, where the average $\overline{F_{T}}$ and standard deviation $\sigma_{T}$ have been computed over a timeseries spanning $10^6\tau_c$.
-  One can see that the distribution for $k=0.03$ is peaked for a particular value of the fluctuation.
-This indicates that a significant part of the sampled trajectories actually overlap over the major part if their history.
-  A a result, the average drag only differ by a small amount.}
+Additionally, the cloning period is $\tau = \tau_c/2$, resulting in $M=20$ iterations of the evolution/selection procedure.}
 \end{figure}

### 2019-04-19 figures/IS_GKTL/IS_GKTL.py

@@ -1,0 +1,49 @@
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+def ret_gaussian(x):
+    """ Returns zero-mean, unit variance at point
+    """
+    return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
+
+prefix = '/home/tlestang/transition_hdd/thibault/These/tailleur_lecomte/lbm/2_sq/' \
+         + 'grid/illustration_IS/'
+
+directories = ['k_0.02_2018_03_26_10_01_46/', 'k_0.025_2018_03_25_20_13_40/',
+           'k_0.03_2018_03_25_20_13_40/'];
+k_value = [0.02, 0.015, 0.03]
+
+sig = 0.0087714;
+
+fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
+
+i=0
+for directory in directories:
+    A = np.fromfile(prefix+directory+'A.datout', float, -1, "")
+    A = A/sig # A is already zero-mean
+
+    # https://seaborn.pydata.org/generated/seaborn.kdeplot.html
+    sns.kdeplot(A, shade=True, ax=ax, label='$k={:1.2f}$'.format(k_value[i]))
+    i=i+1
+    
+xvec = np.linspace(-2,5,100)
+# Plot the origina PDF estimate (gaussian approx)
+ax.plot(xvec,ret_gaussian(xvec),linestyle='--', label='$k=0$')
+ax.set_xlim((-2,7))
+ax.set_ylim((0,0.7))
+
+ax.set_xticks([-2,0,2,4,6])
+ax.set_yticks(ax.get_yticks()[::2]) # Plot only half of ticks
+ax.set_xticklabels(['$-2\\sigma_T$','$0$','$2\\sigma_T$','$4\\sigma_T$','$6\\sigma_T$'])
+ax.tick_params(axis='both', which='major', labelsize=16)
+
+ax.set_xlabel('$F$', fontsize=22)
+ax.set_ylabel('$\\rho_{k}(F_T = F)$', fontsize=22)
+ax.legend(loc='best', fontsize=16)
+
+# Prints figure
+fname = 'IS_GKTL.png'
+plt.savefig(fname)
+
+plt.show()

### 2019-04-19 draft_p2.tex
Update text in part IV.B.2 Application of GKTL

@@
 We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
-The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
+The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag} acting on the square.
 Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.
@@
   \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
-  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
+  \approx \frac{1}{N}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
 \end{equation}
-computed over the set of $N=16384$ trajectories sampled by the algorithm.
-Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
-This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\ref{sec:pdfs}).
+computed over the set of $N_c=16384$ trajectories sampled by the algorithm.
+Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the unbiased \ac{pdf} describing the time-averaged drag process.
+This estimate was computed on the basis the control run, see section~\ref{sec:pdfs}.
 Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
@@
 For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
-For $k \gtrapprox k_{max}$ the majority of the clones are resampled, with the expection of only one or two.
-Because the resampling is based on a very small subset of the ensemble of $N_c$ trajectories, most of the trajectories in the biased ensemble overlap.
-This effect is visible in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
+For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
+This effect is visible in figure~\ref{fig:illustration_IS}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
 Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
@@
 See~\cite{ref1,ref2,ref3} for further details.
-
+% how to assess the value of k_max ?
 For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
@@
 % qu'un échantillonnage direct
-Sampling larger fluctuations therefore requires a larger population size, and a larger computational cost.
-Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to extreme fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
-For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$.
-Note that trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.} are counted as one.  
-As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
-To estimate this number we assumed that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
+Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
+For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T} \geq a$.
+Note that we count as one trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.}.
+As a comparison, table~\ref{tbl:nbTraj_tbl} indicates the average number of fluctuations one can expect from simulating $N_c$ independant trajectories of duration $T_a$ $\{f_d(t)\}_{0\leq t \leq T_a}$, this with the same computational cost $C_{gktl}$.
+This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
 This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
-Even though such a Gaussian assumption does not hold for fluctuations $F_T \geq 4\sigma_T$,
-it may still provide numbers relevant to the comparison with the algorithm.
+
 \begin{figure}
@@
   \end{tabular}
-  \caption{Number of independent corresponding to fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$.
-Parameters of the \ac{gktl} experiments are $N_c = 16384$, $T_a=10\tau_c$, $\tau = \tau_c /2$.
-Within the biased ensemble, trajectories are considered distinct if they overlap for less than $50\%$ of their duration.
-    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.
-    This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.}
+  \caption{Number of fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$.
+    Within the biased ensemble, trajectories are counted as one if they overlap over more than $50\%$ of their duration.
+    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.}
 \label{tbl:nbTraj_tbl}
@@
   \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-  \caption{Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 
-  The \ac{gktl} algorithm is used with $N_c = 16384$ copies and $T_a = T= 10\tau_c$ for three increasing values of the bias $k$.
-Additionally, the cloning period is $\tau = \tau_c/2$, resulting in $M=20$ iterations of the evolution/selection procedure.}
+  \caption{Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$. The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
+The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
 \end{figure}

### 2019-04-19 figures/PDF_AVG/plot_pdf_AVG.py
Takes the least-square fit out of PDF_AVG

@@ -32,8 +32,8 @@ line.set_label('PDF of $F_T$')
 lineGauss, = ax.plot(xi,ret_gaussian(xi), label='Gaussian PDF')
 
 # Plots the local least-square fit for the positive tail
-lineFit, = ax.plot(xiTrunc,np.exp(p(xiTrunc)), linewidth=2, \
-                   color='k')
+#lineFit, = ax.plot(xiTrunc,np.exp(p(xiTrunc)), linewidth=2, \
+#                   color='k')
 
 # Set figure parameters
 ax.set_yscale('log')

### 2019-04-20 draft_p2.tex
Takes the least-square fit out of PDF_AVG
Add conclusion and new placement of figures for part IV.B.2 GKTL results

@@
 % Commentaire de la figure IS_GKTL
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
+  \caption{Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$. The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
+The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
+\end{figure}
 Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
@@
 % qu'un échantillonnage direct
-Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
+We now assess the efficiency of the \ac{gktl} algorithm, with respect to a direct sampling approach. 
 For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T} \geq a$.
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-  \caption{PDF of average drag over $10\tau_c$}
-  \label{fig:PDF_AVG}
-\end{figure}
-
-In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-In the latter case, each fluctuation typically lasts a correlation time $\tau_c$.
-In this context, the \ac{gktl} algorithm is particularly suited to sample trajectories that display a succession of these fluctuations.
-(\textbf{conclusion à développer.})
 \begin{table}
@@
     $k=0.02$ & 1594 & 799 & 155 & 22 & 0  \\ 
-    $k=0.025$ 1019 & 834 & 521 & 198 & 27  \\ 
+    $k=0.025$ & 1019 & 834 & 521 & 198 & 27  \\ 
     $k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-    $\mathcal{N}_{direct}$ & 2601.258& 374.862& 24.268& 0.958& 0.000\\ 
+    $\mathcal{N}_{direct}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005 
   \end{tabular}
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-  \caption{Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$. The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
-The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
+  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+  \caption{PDF of average drag over $10\tau_c$}
+  \label{fig:PDF_AVG}
 \end{figure}
+
+% Conclusion de cette partie
+Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a direct sampling approach of similar computational cost.
+In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
+However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two vert large fluctuations of the instantaneous drag.
+Further works must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
 

### 2019-04-20 draft_p2.tex
Divide AMS subsection into two subsubsections

@@
 
+\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+\label{sec:ams_ou}
 % Example: OU process
@@
 \end{figure}
-We first describe results of the \ac{tams} on a simple case.
 We temporarily let fluid dynamics aside and consider a one-dimensional \ac{ou} process
@@
 
+\subsubsection{The \ac{ams} for extreme drag fluctuations}`
+\label{sec:ams_drag}
 % Application to turbulent flow past an obstacle

### 2019-04-20 figures/AMS_drag_resampling/make_plot_256.py
Updates figure AMS_drag_resampling

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 plt.style.use('ggplot')
-
+color_list = plt.rcParams['axes.prop_cycle'].by_key()['color']
 # This scripts produces a figure that illustrates how the maximum over the resamplied trajectories
 # saturates. It shows that that the TAMS does not lead to higher fluctuations, with respect to the
 # initial ensemble of trajectories.
@@ -11,7 +11,7 @@ plt.style.use('ggplot')
 # transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/courbe_efficacite_instant
 # Original MATLAB script is make_plot_256.m
 
-prefix = '/home/thibault/transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/' \
+prefix = '/home/tlestang/transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/' \
         +'data/instant_256_20/'
 
 # Defines average and std of drag process, compute over the control run
@@ -50,10 +50,10 @@ init.set_color(line.get_color())
 init.set_label('Maxima over initial trajectories')
 
 # Plots maxima over resampled trajectories
-resamp, =ax.plot(Cvec, amax, linestyle='None')
+resamp, = ax.plot(Cvec, amax, linestyle='None')
 resamp.set_marker('o')
 resamp.set_markersize(3)
-resamp.set_markerfacecolor('#6666ff');
+resamp.set_markerfacecolor(color_list[1]);
 resamp.set_markeredgecolor(resamp.get_markerfacecolor())
 resamp.set_label('Maxima over resampled trajectories')
 

### 2019-04-20 draft_p2.tex
Fix sentence about AMS

@@
 The \acl{ams}~\cite{amspaper} inherits from the basic ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman,rolland_simonnet}.
-Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date (\textbf{Sans doute un peu <<too-much>>.
-Avec cette phrase je voudrais justifier que l'on s'intéresse à l'AMS et pas à un autre algo.}).
-In recent years, it was applied to several problems, for instance the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
+Over the past ten years the \ac{AMS} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
 

### 2019-04-20 figures/typical_drag_signal/plot_drag_timeseries.py
Fix unit of time in figure typical_drag_signal

@@ -4,18 +4,20 @@ import matplotlib.pyplot as plt
 
 plt.style.use('ggplot')
 
-path_to_file = "/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/control_run/control_1" \
+path_to_file = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/control_run/control_1" \
                + "/data_force.datout"
-tau_c = 1000;
-f = np.fromfile(path_to_file, float, 250*tau_c, "");
-tspan = np.linspace(0,1000,len(f))
+tau_0 = 500; # Turnover time in units of LBM timesteps
+             # The correlation time tau_c is 2000 timesteps, and the
+             # tau_c is 4tau_0
+f = np.fromfile(path_to_file, float, 500*tau_0, "");
+tspan = np.linspace(0,tau_0,len(f))
 
 fig,ax = plt.subplots(1,1,figsize=(8,3), constrained_layout=True)
-ax.set_xlim(0, 1000)
+ax.set_xlim(0, tau_0)
 line, = ax.plot(tspan, f)
 line.set_linewidth(.7)
 
-ax.set_xlabel(r'$t / \tau_c$',fontsize=22)
+ax.set_xlabel(r'$t / \tau_0$',fontsize=22)
 plt.ylabel('$f_d(t)$',fontsize=22)
 ax.set_yticks(np.arange(-0.10,0.30,0.10))
 plt.xticks(fontsize=16)

### 2019-04-20 draft_p2.tex
Change \overline to \langle \rangle in caption for autocorr_drag
Updates figure illustr_ecoulement and justification for 2D in main text

@@
   \centering
-  \includegraphics[width=\linewidth]{illustr_vort/illustration_vorticity_tf2.png}
+  \includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
   \caption{Our case study is a grid-generated turbulent flow impinging a squared obstacle located at the center of a channel (in two dimensions).
 The vorticity is displayed with an arbitrary colormap from blue (negative values) to red (positive values).
-The Reynolds number based on the grid spacing is $Re=...$.
-// no color bar, show the grid //}
-  \label{fig:computational_domain}
+The Reynolds number based on the grid spacing is $Re=...$.}
+  \label{fig:illustr_ecoulement}
 \end{figure}
@@
 The drag exerted by a grid-generated turbulent flow onto a squared obstacle is considered as our representative case study.
-The flow is contained in a long channel as illustrated in Fig.~\ref{fig:computational_domain}.
+The flow is contained in a long channel as illustrated in Fig.~\ref{fig:illustr_ecoulement}.
 Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and make possible a systematic study.
-Nevertheless, this fluid-mechanical system exhibits the characteristic features of turbulent fluid-structure interaction.
-// true even if 2d?
-// In the present situation, the obstacle does not deform or move.
+Nevertheless, this fluid-mechanical system embeds the key features that makes the application of rare event algorithms both relevant and challenging for turbulent fluid-structure interaction problems.
+In particular, spatio-temporal chaos that results in the unpredictable emergence of coherent fluid structures, in turn leading to  strong fluctuations of the mechanical efforts on the structure.
+In the present situation, the obstacle does not deform or move.
 Through  this specific setting, our main motivation is to evaluate the operability of importance sampling techniques to capture extreme events with a significant run-time savings.  
@@
 Note that both \acsp{pdf} have different variance, and are here not normalized.
-  \textbf{(b)} Autocorrelation function of the drag defined as $\left( \overline{f_d(t+\tau)f_d(t)}-\overline{f_d}^2 \right) /\sigma^2$.
+  \textbf{(b)} Autocorrelation function of the drag defined as $\left( \langle f_d(t+\tau)f_d(t) \rangle-\langle f_d \rangle ^2 \right) /\sigma^2$.
 The lag $\tau$ is normalized by the turnover-time~\eqref{eq:turnover_time}.}
@@
 The \acl{ams}~\cite{amspaper} inherits from the basic ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman,rolland_simonnet}.
-Over the past ten years the \ac{AMS} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
+Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
 

### 2019-04-20 figures/illustr_ecoulement/plot_ecoulement.py

@@ -1,0 +1,74 @@
+# Produces a figure illustrating the vorticity field in the whole comp. domain
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
+
+Dx = 513
+Dy = 129
+N = Dx*Dy
+F = np.zeros(104)
+idx = np.zeros(104)
+
+# Make use of flow data generated for figure ecoulement typique
+prefix = '../ecoulement_typique/'
+
+fig, ax = plt.subplots(figsize=(16,7), constrained_layout=True)
+# Declares a 2x5 grid.
+
+path_to_file_ux = prefix+'ux.dat'
+path_to_file_uy = prefix+'uy.dat'
+fx = open(path_to_file_ux, "rb")  # reopen the file
+fy = open(path_to_file_uy, "rb")  # reopen the file
+bbox = dict(boxstyle="round", fc="0.9")
+
+# Get velocity fields
+ux = np.fromfile(fx, float, N, "")
+ux = np.reshape(ux, (129,513), 'F')
+uy = np.fromfile(fy, float, N, "")
+uy = np.reshape(uy, (129,513), 'F')
+    
+ymin=1;ymax=128
+xmin=1;xmax=512
+# Compute vorticity using finite-differences
+dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
+dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
+vort = dxy - dyx
+
+im=ax.imshow(vort, interpolation='spline36', cmap='BrBG', vmin=-0.07, vmax=0.07)
+
+# Adds a filled Rectangle to mark the square obstacle
+rect = ax.add_artist(Rectangle((256,55), 14, 14))
+rect.set_color('#545454')
+# Show the grid
+R = 8
+ny,nx  = ux.shape
+n_lamina = ((ny-1)/R)/2
+x0 = (nx-1)/16
+for i in range(0,int(n_lamina)):
+    y_start = 2*i*R + R/2
+    y_end = y_start + R
+    grid = ax.add_artist(Rectangle((x0-1,y_start-1), 1, R-1))
+    grid.set_color('#545454')
+
+# Write ticks and labels
+xticks = ax.get_xticks()
+xticks[:] = [x / 17 for x in xticks]
[…]

### 2019-04-20 draft_p2.tex
Adds into and merge with Emmanuel's initial draft

@@
 In the turbulence literature, ``extreme'' can refer to fluctuations that deviate from the mean value by $O(10)$ standard deviations.
-// true ?
-//
 In an engineering context, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
@@
 \\
-
-// Our motivations and expectations. //
+Rare event sampling for turbulence has traditionally been approached from simplified dynamics, such as Burger's turbulence with a stochastic forcing~\cite{bec_burgers_2007}.
+In this context it is possible to describe the dynamics through an action analoguous to a Gibbs measure for an equilibrium sytem.
+This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}.
+Even though Monte Carlo simulations are not expected to be competitive with traditional approaches such as finite difference or pseudo-spectral methods, they provide a simulation framework in which constraints can be applied so as to certain events preferentially.
+A different approach is the Instanton method~\cite{grafke_instanton_2015}, also relying on the knowledge of an action for the dynamics.
+For stochastically driven systms, in the limit of weak noise, it can be shown that the probability of extreme events is dominated by the action minima~\cite{freidlin2013random}.
+The trajectory that that minimizes the action is referred to as the \textit{instanton}~\cite{gurarie_instantons_1996}.
+The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.
+Such approaches are however limited to stochastically driven dynamics, simple enough so that an action can be derived and sampled, or minimized.
+This is for instance the case of the stochastically driven 1D Burger's equation.
+\\
+Whilst Burgers turbulence provids an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.
+In this paper, we discuss a thrid approach to the numerical simulation of rare events in turbuent flows, that aims at flows for which an action cannot be derived.
+It is based on the application of \textit{rare event algorithms}, coupled with a numerical simulation of the flow dynamics.
+Even though the original ideads can be traced back to the early 1950's, rare event algorithms have been of ever-growing interest voer the last 20 years.
+Successful application of such algorithms can be found in various domains such as chemistry~\cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016},biophysics~\cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear phsics~\cite{louvin2017}, nonlinear dynamical systems~\cite{tailleur_probing_2007}, communication networks simulation~\cite{villen-altamirano_restart:_1994}.
+\\
+However, a large body of algorithms are limited to systems at equilibrium, for which the statioany ry measure can be sampled by means of Monte Carlo methods.
+For systems far from equilibrium, such as turbulent flows, the sampling of rare events must be done at the level of the trajectories.
+In the present paper with discuss two different algorithms suitable to out-of-equilibrum dynamics.
+The Adaptive Multilevel Algorithm (2006)~\cite{cerou_adaptive_2007} inherits from older ideas about splitting approaches~\cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999}.
+In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland_computing_2016} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.
+Over the last decade, the main theoritical framework for the study of rare events in statistical physics has been the one of large deviation theory~\cite{touchette_large_2009}.
+Alongside, numerical methods have been developed to sample rare events based on general ideas and algorithms~\cite{moral_feynman-kac_2004}.
+Among them, the \ac{gktl} algorithm~\cite{giardina_direct_2006} is particularly suited to rare events in dynamical systems~\cite{giardina_simulating_2011,Laffargue_2013}.
+\\
+The applicability of such algorithms to flows relevant to envorinmental or industrial applications remains open.
+Recently, the \cite{gktl} allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}.
+Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.
+In the present paper, we consider a different context of application: turbulent flows around objects.
 

### 2019-04-21 draft_p2.tex
Fix figure references

@@
 Negative vorticity is generated along the bottom boundary, and advected away from the the obstacle by the mean flow.}
-  \label{fig:typical_drag_signal}
+  \label{fig:typical_vorticity}
 \end{figure}
@@
 In the following we define the \textit{correlation time} $\tau_c = 4\tau_0$.
-Note that the ratio $tau_0 / tau_c$ may be viewed as a \textit{Strouhal number} $St$.
+Note that the ratio $\tau_0 / \tau_c$ may be viewed as a \textit{Strouhal number} $St$.
 A value $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
@@
 The value of this threshold was chosen so that the return time of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
-Figure~\ref{return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
-According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard deviation of the drag process.
-We denote by $(t^{\star}, f_d^{\star})$ a fluctuation of peak amplitude $f_d^{\star}$ occurring at time $t^{\star}$, as illustrated in figure~\ref{fig:extraction}.
+Figure~\ref{fig:return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
+According to figure~\ref{fig:return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard deviation of the drag process.
+We denote by $(t^{\star}, f_d^{\star})$ a fluctuation of peak amplitude $f_d^{\star}$ occurring at time $t^{\star}$.
 From the control timeseries, we identified $104$ independent extreme drag fluctuations whith an amplitude greater than $7.6\sigma$, \textit{i.e.} with a return time greater than $10^4\tau_c$.
@@
 
-Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
+Figure~\ref{fig:pressure_ratio} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
 \begin{figure}
@@
   \includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-  \label{fig:pressure_ratio}
-  \caption{Contribution of the forebody and base pressure fluctuation to the overall drag fluctuation for the 104 sampled extreme drag fluctuations in the control run.
+  \caption{\label{fig:pressure_ratio} Contribution of the forebody and base pressure fluctuation to the overall drag fluctuation for the 104 sampled extreme drag fluctuations in the control run.
 To each of these fluctuation correspond a unique value of the drag $f^{\star}_d$, and a unique pair ($\tilde{p}^{\star}_{base}$,$\tilde{p}^{\star}_{fb}$).
@@
   \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-  \label{density+streamlines}
-  \caption{Pressure field and velocity streamlines in the vicinity of the obstacle at $t=t^{\star}$.
+  \caption{\label{fig:density+streamlines} Pressure field and velocity streamlines in the vicinity of the obstacle at $t=t^{\star}$.
 Because of the blocking vortex, vorticity produced along the top boundary of the obsatcle has no other choice but to accumulate at the base of the obstacle.}
@@
   \includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-  \label{fig:timeseries_extremes}
-  \caption{Average drag timeseries in the vicinity of an extreme drag fluctuation at $t=t^{\star}$.
+  \caption{\label{fig:timeseries_extremes} Average drag timeseries in the vicinity of an extreme drag fluctuation at $t=t^{\star}$.
 The blue continuous line results from an average over the 104 extreme events sampled in the control run.
@@
   \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-  \label{fig:vorticity_dynamics}
-  \caption{Vorticity dynamics associated to an extreme fluctuation of the drag acting on the square obstacle at $t=t^{\star}$.}
+  \caption{\label{fig:vorticity_dynamics} Vorticity dynamics associated to an extreme fluctuation of the drag acting on the square obstacle at $t=t^{\star}$.}
 \end{figure}
@@
 
-To begin with, figure~\ref{fig:top_four_events} displays the vorticity field corresponding to the peak drag value, for a subset of the extreme events sampled in the control simulation.
+To begin with, figure~\ref{fig:top_4_events_vorticity} displays the vorticity field corresponding to the peak drag value, for a subset of the extreme events sampled in the control simulation.
 We observe that, in each case, extreme drag flcutuations correspond to very similar flow configurations.
-More precisely, a region of strong vorticity is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
+More precisely, a region of strong vorticity is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_vorticity}.
 This vorticity results from a boundary layer, along either the top or bottom side of the obstacle.
@@
 How are such vorticity reached ?
-In contrast with the typical flow dynamics, illustrated in figure~\ref{vorticity_typical}, the vorticity generated within the boundary layers is not directly advectd by the means flow.
-Figure~\ref{fig:illustr_density_streamlines} shows that, instead, this vorticity is constrained in the vicinity of the base of the obstacle.
+In contrast with the typical flow dynamics, illustrated in figure~\ref{fig:typical_vorticity}, the vorticity generated within the boundary layers is not directly advectd by the means flow.
+Figure~\ref{fig:density+streamlines} shows that, instead, this vorticity is constrained in the vicinity of the base of the obstacle.
 Such constraint results from a large vortical region located roughly one obstacle length of the base, that prevents vorticity to be advected downstream.
 
-Figure~\ref{fig:vorticity_dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:illustr_density_streamlines}.
+Figure~\ref{fig:vorticity_dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:density+streamlines}.
 From $t^{\star}-0.4\tau_c$ to $t^{\star}-0.16\tau_c$, a strong negative eddy is generated along the bottom boundary layer of the obstacle and advected downstream by the mean flow.
@@
 % It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
+  \caption{\label{fig:shear_asof_drag} Evolution of the shear alongised the top or bottom boundary of the obsatcle as a function of the drag, for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
+To each oneof the grey lines correspond a unique extreme event.
+The blue line is the average over the 104 extreme event sampled in the control run.}
+\end{figure}
 Figure~\ref{fig:shear_asof_drag} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
 For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
-As illustrated in figure~\ref{fig:extremes_timeseries}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
+As illustrated in figure~\ref{fig:timeseries_extremes}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
 As a consequence, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
@@
 This viscous shear is induced by a large vortex detached from the obstacle.
-For a more detailed description of such dynamics, see the chapter 3 of ~\ref{ref:these_lestang}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-  \caption{\label{fig:shear_asof_drag} Evolution of the shear alongised the top or bottom boundary of the obsatcle as a function of the drag, for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-To each oneof the grey lines correspond a unique extreme event.
-The blue line is the average over the 104 extreme event sampled in the control run.}
-\end{figure}
+For a more detailed description of such dynamics, see the chapter 3 of ~\cite{ref:these_lestang}.
 
@@
 In the limit of very rare events and/or very complex dynamics such as turbulent flows in industrial contexts, a direct sampling approach is unrealistic.
-Indeed, according to equation~\eqref{eq:return_time_scaling}, the return time of a fluctuation $f_d \geq a$ verifies $r(a) \propto e^{\alpha a}$.
+Indeed, according to equation~\eqref{eq:return_time}, the return time of a fluctuation $f_d \geq a$ verifies $r(a) \propto e^{\alpha a}$.
 As a result, the computational cost required to sample events $f_d \geq a$ scales like $e^{\alpha a}$.
@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \label{fig:illustr_AMS}
-  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
+  \caption{\label{fig:illustr_AMS} Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
 The initial ensemble of blue trajectories is first generated.
@@
 Discarded trajectories are resampled based on member of the ensemble of trajectories that achieved a higher value of the score function at some point of their history.
-As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{figure_AMS}.
+As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS}.
 For further details about the algorithm and corresponding mathematical results, see~\cite{lestang2018} and references therein.
@@
 The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
-Figure~\ref{fig:compraison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach.
+Figure~\ref{fig:comparaison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach.
 In the latter, the typical computational cost is simply the return time $r(a)$.
@@
   \includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-    \caption{Maximum of the instantaneous drag along the resampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
+  \caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag along the resampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
 For these particular dynamics, the \ac{tams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation in the initial ensemble.}
@@
 We apply the \ac{tams} with $N=256$ initial trajectories, with a duration $T_a = 20 \tau_c$.
-Similarly to figure~\ref{fig:comparaison_temps_de_retour}, figure~\ref{fig:instant_256_20} displays the maximum drag achieved by resampled trajectories.
-In addition, figure~\ref{fig:instant_256_20} shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
+Similarly to figure~\ref{fig:comparaison_temps_de_retour}, figure~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by resampled trajectories.
+In addition, figure~\ref{fig:AMS_drag_resampling} shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
 Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
-These maxima are depicted by the stars in figure~\ref{fig:instant_256_20}.
+These maxima are depicted by the stars in figure~\ref{fig:AMS_drag_resampling}.
 However, one can see that resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
@@
   \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-  \label{fig:AMS_drag_trajectories}
-  \caption{Ensemble of trajectories after $181$ iterations of the selection-mutation procedure.In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
+  \caption{\label{fig:AMS_drag_trajectories} Ensemble of trajectories after $181$ iterations of the selection-mutation procedure.In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
 Trajectories have a duration $T_a = 5\tau_c$ and their number is $N_c = 32$.
@@
 In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting points will overlap over a timescale called
-the <<Lyapunov timescale>>~\cite{lyapunovRef}.
+the \textit{Lyapunov timescale}~\cite{lyapunovRef}.
 However, we observe that, in our particular case, the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag process.
@@
 This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
-For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\ref{these_lestang,nemoto_bouchet2016}
+For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{these_lestang,nemoto_bouchet2016}.
 As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
@@
   \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-  \caption{Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$. The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
+  \caption{\label{fig:IS_GKTL} Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$.
+The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
 The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
 \end{figure}
-Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
+Figure~\ref{fig:IS_GKTL} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
 For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
@@
 computed over the set of $N_c=16384$ trajectories sampled by the algorithm.
-Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the unbiased \ac{pdf} describing the time-averaged drag process.
+Moreover, figure~\ref{fig:IS_GKTL} displays an estimate of the unbiased \ac{pdf} describing the time-averaged drag process.
 This estimate was computed on the basis the control run, see section~\ref{sec:pdfs}.
@@
 For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
-This effect is visible in figure~\ref{fig:illustration_IS}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
+This effect is visible in figure~\ref{fig:IS_GKTL}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
 Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.

### 2019-04-21 draft_p2.tex
Add conclusion of paper

@@
 \label{conlusion}
+We assessedthe application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
+\\
+In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
+By means of very long simulations---without rare event algorithm---we observed that such extreme events are caused by the concentration of vorticity very close to the base of the square.
+Because of the sweeping of the fluid structure responsible for such a situation, extreme drag levels do not persist over time.
+The lifetime of extreme drag fluctuations is of the order of the turnover time, and the corresponding drag signal is very peaked around the extreme value.
+Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are very well described by an exponential \ac{pdf}.
+This property can be linked to the phenomenology of extremes of the time-averaged drag.
+Indeed, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations or an exceptional succession of typical fluctuations that
+add up forming an extremely large value of the average.
+Such a phenomenology can be described by a stochastic process with an exponential \ac{pdf}.
+\\
+On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
+In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling approach.
+This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which resampled trajectories separate from their parent following the
+addition of a small perturbation in the initial conditions.
+The \ac{gktl} has been applied to the sampling of trajectories displaying extreme fluctuations of the time averaged drag.
+In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a direct approach, allowing the simulation of trajectories that are out of reach without algorithm.
+\\
+The application of rare event algorithms relies on the definition of a score function, on which the selection of trajectories is based upon.
+Although the dynamics itself is a black-box from the point of view of the algorithm, the efficiency of the sampling depends on the phenomenology.
+For complex dynamics, including turbulent flows, the choice of the score function is made difficult by the very little knowledge one has of the rare events of interest.
+This is, after all, our motivation for the simulation of these events.
+In such situations, it is then tempting to settle for a straightforward choice for the score function.
+This is the approach we followed in this work, setting the score function to the drag itself.
+However, for dynamics as complex as the turbulent flow around an obstacle, it is unlikely that selecting trajectories based on the value of the drag favours the occurrence of extreme drag fluctuations
+in the future.
+Under such circumstances, the \ac{tams} fails drastically.
+Successful usage of the \ac{tams} and similar splitting algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that good score functions
+are very difficult to identify, if even possible.
+A popular direction in current research involves taking advantage of the recent advances in learning methods to compute score functions beforehand.
+For the time-averaged drag, we showed that the \ac{gktl} algorithm is able to greatly outperform a direct sampling approach.
+This is because the sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
 

### 2019-04-22 draft_p2.tex
Fix citations

@@
 In this context it is possible to describe the dynamics through an action analoguous to a Gibbs measure for an equilibrium sytem.
-This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}.
+This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice}.
 Even though Monte Carlo simulations are not expected to be competitive with traditional approaches such as finite difference or pseudo-spectral methods, they provide a simulation framework in which constraints can be applied so as to certain events preferentially.
-A different approach is the Instanton method~\cite{grafke_instanton_2015}, also relying on the knowledge of an action for the dynamics.
+A different approach is the Instanton method~\cite{grafke2015instanton}, also relying on the knowledge of an action for the dynamics.
 For stochastically driven systms, in the limit of weak noise, it can be shown that the probability of extreme events is dominated by the action minima~\cite{freidlin2013random}.
 The trajectory that that minimizes the action is referred to as the \textit{instanton}~\cite{gurarie_instantons_1996}.
-The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.
+The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{grafke2015instanton,chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017}.
 Such approaches are however limited to stochastically driven dynamics, simple enough so that an action can be derived and sampled, or minimized.
@@
 \\
-Whilst Burgers turbulence provids an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.
+Although Burgers turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.
 In this paper, we discuss a thrid approach to the numerical simulation of rare events in turbuent flows, that aims at flows for which an action cannot be derived.
@@
 The Adaptive Multilevel Algorithm (2006)~\cite{cerou_adaptive_2007} inherits from older ideas about splitting approaches~\cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999}.
-In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland_computing_2016} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.
+In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland2016computing} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.
 Over the last decade, the main theoritical framework for the study of rare events in statistical physics has been the one of large deviation theory~\cite{touchette_large_2009}.
@@
 The applicability of such algorithms to flows relevant to envorinmental or industrial applications remains open.
-Recently, the \cite{gktl} allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}.
+Recently, the \ac{gktl} algorithm allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}.
 Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.
@@
 A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid. 
-At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{ref1,ref2} where the viscosity is artificially enhanced. 
+At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
 % Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
@@
 
-The flow dynamics are computed by the lattice Boltzmann (LB) method ~\cite{ref1,ref2}. 
+The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}. 
 While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
@@
 Extreme drag fluctuations are \textit{rare events}, that is fluctuations for which $r(a) \gg \tau_c$.
-In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang2018}
+In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang_computing_2018}
 \begin{equation}
@@
 Figure~\ref{fig:return_time_instant} illustrates the fluctuation amplitude $a$ as a function of the return time $r(a)$, for the drag process $f_d(t)$.
-Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang2018}.
+Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang_computing_2018}.
 
@@
 This viscous shear is induced by a large vortex detached from the obstacle.
-For a more detailed description of such dynamics, see the chapter 3 of ~\cite{ref:these_lestang}.
+For a more detailed description of such dynamics, see chapter 3 of~\cite{lestang:tel-01974316}.
 
@@
 
-This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{papier_JSTAT}.
+This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_computing_2018}.
 An estimate of this \ac{pdf} is displayed in figure~\ref{fig:pdf_drag}.
@@
 % Introduction to AMS
-The \acl{ams}~\cite{amspaper} inherits from the basic ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman,rolland_simonnet}.
-Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland2017}.
+The \acl{ams}~\cite{cerou_adaptive_2007} inherits from the basic ideas of \textit{splitting} algorithms~\cite{KahnHarris1951}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
 
@@
 As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS}.
-For further details about the algorithm and corresponding mathematical results, see~\cite{lestang2018} and references therein.
+For further details about the algorithm and corresponding mathematical results, see~\cite{lestang_computing_2018} and references therein.
 
@@
 
-For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
+For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
 Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
@@
 The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
-The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous_varizani,grassberger}.
-Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{delMoral_Feynman_Kac_book}.
-The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{Giardina2006}.
-Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
+The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous1994go,grassberger2002go}.
+Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
+The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{giardina_direct_2006}.
+Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems~\cite{giardina_direct_2006,tailleur_probing_2007}.
 % It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
@@
 \end{equation}
-which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{Giardina2006,Giardina2011}:
+which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{giardina_direct_2006,giardina_simulating_2011}:
 \begin{equation}
@@
 This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
-For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{these_lestang,nemoto_bouchet2016}.
+For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{lestang:tel-01974316,brewer2018efficient}.
 As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
@@
 We performed three experiments, corresponding to three different values of the parameter $k$.
-In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
-In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
+In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{touchette_large_2009}.
+In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{touchette_large_2009}.
 
@@
 An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
-See~\cite{ref1,ref2,ref3} for further details.
+See~\cite{nemoto2017finite,Guevara_Hidalgo_2018} for further details.
 % how to assess the value of k_max ?
@@
 \label{conlusion}
-We assessedthe application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
+We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
 \\
@@
 
+\bibliographystyle{unsrt}
+\bibliography{draft_p2}
+
 \end{document}

### 2019-04-22 draft_p2.tex
Update part I description of the case study

@@
 The vorticity is displayed with an arbitrary colormap from blue (negative values) to red (positive values).
-The Reynolds number based on the grid spacing is $Re=...$.}
+The Reynolds number based on the grid spacing is $Re=1200$.}
   \label{fig:illustr_ecoulement}
@@
 In the present situation, the obstacle does not deform or move.
-Through  this specific setting, our main motivation is to evaluate the operability of importance sampling techniques to capture extreme events with a significant run-time savings.  
+Through  this specific setting, our main motivation is to evaluate the operability of rare event sampling techniques to capture extreme events with a significant run-time savings.  
+\\
+A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid.
+Turbulent eddies are generated in the near-wake of the grid and carried downstream.
+The size of the eddies that eventually interact with the square obstacle, mounted at the center of the channel, is comparable to its size, thus resulting in strong fluctuations of the drag acting on the square.
+At the outlet of the channel, an open boundary boundary is prescribed.
+In addition, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
 \\
 
-A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid. 
-At the outlet of the channel, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
-% Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.
+%% The LBM %%
+
+The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}, using a non-orthogonal central moment relaxation scheme~\ref{De_Rosis_2016}.
+The \ac{lbm} solves the Navier-Stokes equations, in a weakly compressible limit, by simulating the dynamics of fictitious particles at a mesoscopic level.
+The dynamics are discretized on an isotropic lattice of spacing $\Delta x$.
+The \ac{lbm} is further described in appendix~\ref{sec:lbm} and references therein.
+
+%% LB parameters  %
+
+The lattice spacing is set to $\Delta x = R/16$, where $R$ denotes the side of the square obstacle.
+The channel has a dimension $L_x \times L_y$ with $L_y = 8R$ and $L_x = 4L_y = 32R$.
+The lattice therefore has a resolution of $N_y = 129$ lattice nodes in the $y$-direction and $N_x = 513$ nodes in the $x$-direction.
+The grid spacing is set to $g = 8\Delta x$.
+The characteristic velocity at the inlet is set to $U_0 = 0.05 \frac{\Delta x}{\Delta t}$, where $\Delta t$ is the lattice timestep.
+\textcolor{red}{@todo@Emmanuel: Briefly explain why the low Mach number and the nearly incompressible limit.}
+Finally, the viscosity is set so that the Reynolds number is $Re_{grid} = U_0g/\nu = 1200$.
+The Reynolds number based on the mean flow velocity and the size of the obstacle is $Re_{obs} \approx 1700$.
+
+% Boundary conditions %
+
+The channel walls and the surface of the obstacle are decribed by the halfway bounce-back scheme~\cite{ref_BB}\textcolor{red}{@todo@Emmanuel: Provide citation for halfway bounce-back boundary condition}.
+The open boundary at the outlet is based on a second order extrapolation of the velocity along the $x$-direction.
+The extrapolated velocity is then enforced using a regularization based on a finite difference estimation of the local stress tensor~\cite{latt2008straight}.
+The parabolic profile at the inlet is imposed by setting populations at the corresponding equilibrium, see appendix~\ref{sec:lbm}.
 \\
-
-Turbulent eddies are generated in the near-wake of the grid and carried downstream. // grid is usually characterized by the spacing and the mesh size, are they equal here? //
-They interact with each other and grow in size as expected in two-dimensional turbulence. 
-The size of the eddies that eventually hit the square (at the center of the channel) is comparable to the size of this obstacle, which results in strong fluctuations of the drag acting on the square. 
-\\
-
-The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}. 
-While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
-Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
-Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
-This is, in short, the rationale behind the LB approach.
-In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
-%
-More precisely, the fluid is viewed as populations of particles
-that collide, redistribute and propagate along the different links of a discrete lattice. 
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
-
-// algorithm for boundary conditions // computation of the stress with LBM // grid resolution, number of grid points in each direction, Reynolds, Mach numbers, etc.
-// low Mach number approaches the nearly-incompressible limit
-\\
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
 
@@
 In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting points will overlap over a timescale called
-the \textit{Lyapunov timescale}~\cite{lyapunovRef}.
+the \textit{Lyapunov timescale}~\cite{lyapunovRef}. \textcolor{red}{@todo@Freddy: Provide citation for Lyapunov timescale}
 However, we observe that, in our particular case, the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag process.
@@
 
+\appendix*
+\section{The \acl{lbm}}
+\label{sec:lbm}
+While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
+Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
+Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
+This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
+This is, in short, the rationale behind the LB approach.
+In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
+\\
+More precisely, the fluid is viewed as populations of particles
+that collide, redistribute and propagate along the different links of a discrete lattice. 
+In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
+Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
+\[
+\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
+\]
+for the mass density and the fluid momentum respectively.
+The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
+	\caption{Sketch of the D2Q9 lattice.
+Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
+By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
+	\label{fig:D2Q9}
+\end{figure}
+
+The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
+Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+Namely, an \emph{exact} streaming step
+\[
+f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
+\]
+during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
+\[
+f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
+\]
+which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
+\[
+\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
+\]
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
+\[
+f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
+\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
+\] 
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
+Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
+$\Delta x \to 0$.
+\\
+As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
+\[
+\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\]
+so that the total stress expresses as
+\begin{equation}\label{eq:def_stress}
+\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\end{equation}
+
 \bibliographystyle{unsrt}

### 2019-04-22 draft_p2.tex
Update title of paper and perspectives on AMS

@@
 
-\title{Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+\title{Rare event algorithms for the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
 
@@
 However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
-As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not so clear.
+As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not straightforward.
 In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
-For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
-Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.
 
@@
 % perspectives
-This result suggest that the \ac{tams}--- and \ac{ams}---cannot be expected to efficiently sample rare trajectories for dynamics with such a phenomenology.
-% More trajectories ?
-Note that, because of the complexity of the dynamics, we were limited to a rather low number of trajectories, relatively to previous use cases of these algorithms~\cite{refs}.
-Targeting events with a fixed return time, one can show that the computational cost of a \ac{tams} experiment roughly grows linearly with the number of trajectories $N$~\cite{lestang2018}.
-As a result, multiplying the number of trajectories by a factor 10 or 100 would require important computational resources as well as an efficient parallel implementation of the \ac{tams}.
-Despite the relative simplicity of the \ac{tams} algorithm, the latter implementation is difficult to achieve~\cite{ref}.
-In any cases, from the observations made in this section, we do not expect that increasing the number of trajectories is able to better the efficiency of the sampling for such dynamics.
-% Another score function ?
-%%
-% The choice of the simple score function was already discusses above...
-% Maybe not necessary to RE-discuss it here.
-
-% Modify the TAMS to fit the phenomenology ?
-A promising perspective is to adapt the \ac{tams} to dynamics for which the Lyapunov time is larger than the typical duration of extreme events, such as the convective flow involved in this work.
-In the original \ac{tams}, at later stages of the algorithm, new trajectories are resampled starting from points very close to the maximum of the parent trajectory.
-Resampled trajectories have no chance to display a significantly higher maximum as they overlap their parent trajectory over the Lyapunov timescale.
+A direct application of the \ac{tams}, choosing the drag itself as the score function,  cannot be expected to efficiently sample rare trajectories for flows with such a phenomenology.
+Indeed, there is \textit{a priori} no reason for the drag itself to be a good measure of how likely future drag fluctuations are.
+Finding a better score function requires an intricate understanding of the dynamics leading to extreme drag fluctuations, beyond the phenomenological description of section~\ref{sec:dynamical_aspects}.
+Furthermore, such fine-tuning is problem dependent, and the study should be done each time the algorithm is applied to a similar, however different, system.
+A current line of research is to adapt the algorithm itself to the phenomenology.
 As an example, the resampling procedure could be modified so that new trajectories are resampled roughly a Lyapunov time \textit{before} the maximum is attained.
-In this way, it gives the opportunity for new events, associated to a higher fluctuations, to occur.
+In this way new events, associated to higher fluctuations, have the opportunity to occur.
 However, such modification(s) of the resampling procedure must be implemented with care, so as to preserve the mathematical properties of the \ac{ams} and \ac{tams}.
+%
 

### 2019-05-14 draft_p2.tex
Modifications emmanuel 3/05/2019

@@
 
+\newcommand{\todo}[1]{{\color{red}{#1}}}
+
 \begin{document}
 
-\title{Rare event algorithms for the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
 
@@
 \affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
 \author{Freddy Bouchet}
@@
 \email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
 
 \begin{abstract}
-This study evaluates the relevance of importance sampling techniques to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea behind importance sampling is here to replace a long simulation by a set of shorter simulations running in parallel with dynamics that are replicated or abandoned in order to sample large events more frequently.
+This study evaluates the relevance of  rare-event sampling to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+The main idea is here to replace a long simulation by a set of much shorter simulations (running in parallel) with dynamics that are replicated or abandoned in order to sample large-amplitude events more frequently.
 %
-Such techniques have already proved their efficiency for simulating extreme events in... with significant run-time savings.
-Application to turbulent fluid-mechanical systems remains mainly open. 
+Such technique has already proved its efficiency for simulating extreme events in {<...>} with significant run-time savings.
+Application to fluid-structure interaction problems remains mainly open. 
 %
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two important sampling algorithms, namely ....
-Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort...
-However, limitations...
-pitfalls...
-Importantly, such techniques also provide access to the flow scenario yielding extreme events. In the present case, it is found that... blocking effect
+The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely {<...>}.
+Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort.
+However, {<limitations and pitfalls>}.
+% Importantly, such techniques also provide direct access to the flow scenario yielding extreme events. In the present case, it is found that {<blocking effect>}.
 \end{abstract}
@@
 
-// Recall the previous success of importance sampling techniques with references. //
-Explain why interesting to test application for turbulent systems. // To what extent a turbulent system is different from previous case studies? // 
-\\
+% general comments on physical problem %
+%
+Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
+``Extreme'' refers here to fluctuations that can deviate from mean value by ${\cal{O}}(10)$ standard deviations.
+In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
+Such anomalous efforts can threaten the structural integrity of embedded structures.
+%
+From the viewpoint of (chaotic) dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
+The use of perturbative methods in identifying resonant interactions (among degrees of freedom) possibly responsible for extreme fluctuations is not successful.
+Simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
+Nevertheless, this requires very long simulation since these events are also very rare.
+It is here our motivation to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost.
+%
 
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces.
-In the turbulence literature, ``extreme'' can refer to fluctuations that deviate from the mean value by $O(10)$ standard deviations.
-In an engineering context, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
-\\
+% general def on rare-event sampling %
+%
+Rare-event sampling is an umbrella term for numerical methods intended to sample preferentially regions of the phase space which would otherwise be rarely visited  through brute-force simulation.
+%
 
-From the viewpoint of dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium.
-The use of perturbative methods in identifying resonant interactions (among degrees of freedom) leading to extreme fluctuations is not successful.
-In this context, simulation offers a practical means to gain physical insight into these events and estimate their frequency of occurrence, or return times.
-Nevertheless, this requires very long simulation since these events are also very rare.
-It is here our motivation to evaluate the practicality and efficiency of importance sampling techniques to sample extreme events in developed turbulent flows.
-\\
-Rare event sampling for turbulence has traditionally been approached from simplified dynamics, such as Burger's turbulence with a stochastic forcing~\cite{bec_burgers_2007}.
-In this context it is possible to describe the dynamics through an action analoguous to a Gibbs measure for an equilibrium sytem.
-This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice}.
-Even though Monte Carlo simulations are not expected to be competitive with traditional approaches such as finite difference or pseudo-spectral methods, they provide a simulation framework in which constraints can be applied so as to certain events preferentially.
-A different approach is the Instanton method~\cite{grafke2015instanton}, also relying on the knowledge of an action for the dynamics.
-For stochastically driven systms, in the limit of weak noise, it can be shown that the probability of extreme events is dominated by the action minima~\cite{freidlin2013random}.
-The trajectory that that minimizes the action is referred to as the \textit{instanton}~\cite{gurarie_instantons_1996}.
-The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{grafke2015instanton,chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017}.
-Such approaches are however limited to stochastically driven dynamics, simple enough so that an action can be derived and sampled, or minimized.
-This is for instance the case of the stochastically driven 1D Burger's equation.
-\\
-Although Burgers turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.
-In this paper, we discuss a thrid approach to the numerical simulation of rare events in turbuent flows, that aims at flows for which an action cannot be derived.
-It is based on the application of \textit{rare event algorithms}, coupled with a numerical simulation of the flow dynamics.
-Even though the original ideads can be traced back to the early 1950's, rare event algorithms have been of ever-growing interest voer the last 20 years.
-Successful application of such algorithms can be found in various domains such as chemistry~\cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016},biophysics~\cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear phsics~\cite{louvin2017}, nonlinear dynamical systems~\cite{tailleur_probing_2007}, communication networks simulation~\cite{villen-altamirano_restart:_1994}.
-\\
-However, a large body of algorithms are limited to systems at equilibrium, for which the statioany ry measure can be sampled by means of Monte Carlo methods.
-For systems far from equilibrium, such as turbulent flows, the sampling of rare events must be done at the level of the trajectories.
-In the present paper with discuss two different algorithms suitable to out-of-equilibrum dynamics.
-The Adaptive Multilevel Algorithm (2006)~\cite{cerou_adaptive_2007} inherits from older ideas about splitting approaches~\cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999}.
-In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland2016computing} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.
-Over the last decade, the main theoritical framework for the study of rare events in statistical physics has been the one of large deviation theory~\cite{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events based on general ideas and algorithms~\cite{moral_feynman-kac_2004}.
-Among them, the \ac{gktl} algorithm~\cite{giardina_direct_2006} is particularly suited to rare events in dynamical systems~\cite{giardina_simulating_2011,Laffargue_2013}.
-\\
-The applicability of such algorithms to flows relevant to envorinmental or industrial applications remains open.
-Recently, the \ac{gktl} algorithm allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}.
-Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.
-In the present paper, we consider a different context of application: turbulent flows around objects.
+%
+% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
+%
+% this paragraph would certainly need to be reshaped!
+%
+In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
+In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium.
+This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}, which provides a framework for rare-event sampling.
+%
+An alternative approach is based on \emph{instanton} \cite{grafke_instanton_2015}.
+For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}.
+Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.
+%
+An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimized.
+%
+%
+%Whilst Burgers' turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or environmental flows.
+%
+In the present paper, a more straightforward approach is investigated.
+It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.
+% dedicated to dynamical systems for which an action cannot be easily derived or minimized,
+%    
+%
+Even though such ideas trace back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
+%
+% already mentioned in the beginning? This sentence is misleading.
+%
+%However, a large body of algorithms are limited to systems at equilibrium, for which the stationary measure can be sampled by means of Monte Carlo methods. 
+%
+An original contribution of our study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \todo{TRUE?}
+% 
+Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
+%
 
-\section{Description of the case study}
+The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a more detailed description will be given later in the paper.
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \cite{aristoff_adaptive_2015}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
+
+Over the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
+Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
+Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
+%
+Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
+This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
+Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.
+
+\section{Description of the numerical case study}
 \label{sec:test_flow}
@@
 \includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-  \caption{Our case study is a grid-generated turbulent flow impinging a squared obstacle located at the center of a channel (in two dimensions).
-The vorticity is displayed with an arbitrary colormap from blue (negative values) to red (positive values).
-The Reynolds number based on the grid spacing is $Re=1200$.}
+\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the center of a channel in two dimensions.
+The flow is artificially damped near the end of the channel.
+In the developed flow, turbulent eddies have typically the size of the square.
+The vorticity is displayed with an arbitrary color map from blue (negative values) to red (positive values).}
 \label{fig:illustr_ecoulement}
@@
 
-The drag exerted by a grid-generated turbulent flow onto a squared obstacle is considered as our representative case study.
-The flow is contained in a long channel as illustrated in Fig.~\ref{fig:illustr_ecoulement}.
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and make possible a systematic study.
-Nevertheless, this fluid-mechanical system embeds the key features that makes the application of rare event algorithms both relevant and challenging for turbulent fluid-structure interaction problems.
-In particular, spatio-temporal chaos that results in the unpredictable emergence of coherent fluid structures, in turn leading to  strong fluctuations of the mechanical efforts on the structure.
-In the present situation, the obstacle does not deform or move.
-Through  this specific setting, our main motivation is to evaluate the operability of rare event sampling techniques to capture extreme events with a significant run-time savings.  
-\\
-A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid.
-Turbulent eddies are generated in the near-wake of the grid and carried downstream.
-The size of the eddies that eventually interact with the square obstacle, mounted at the center of the channel, is comparable to its size, thus resulting in strong fluctuations of the drag acting on the square.
-At the outlet of the channel, an open boundary boundary is prescribed.
-In addition, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
-\\
+% introduce the flow 
+%
+The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
+% why this flow?
+%
+Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
+% 
+We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
+Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
+%
+%
+% rapid description of the flow
+Turbulent eddies generated in the near-wake of the grid are carried downstream. They interact with each other and grow in size as expected in two-dimensional turbulence. 
+The dimension of the grid is such that the size of the eddies that eventually hit the square is comparable to the size of this latter, which results in strong fluctuations of the drag acting on the square.
+%
+%Finally, the obstacle does not deform or move either. 
+%
+%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
+%
 
-%% The LBM %%
+% lattice Boltzmann method 
+%
+The flow dynamics is integrated by the lattice Boltzmann (LB) method in our numerical simulations. 
+While traditional methods in \emph{computational fluid dynamics} rely on a discretization of the Navier-Stokes equations, the LB method considers the fluid at a mesoscopic level.
+Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
+This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
+%namely the Navier-Stokes dynamics in the weakly compressible limit
+Further details are given in Appendix and references therein.
+In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
 
-The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}, using a non-orthogonal central moment relaxation scheme~\ref{De_Rosis_2016}.
-The \ac{lbm} solves the Navier-Stokes equations, in a weakly compressible limit, by simulating the dynamics of fictitious particles at a mesoscopic level.
-The dynamics are discretized on an isotropic lattice of spacing $\Delta x$.
-The \ac{lbm} is further described in appendix~\ref{sec:lbm} and references therein.
-
-%% LB parameters  %
-
-The lattice spacing is set to $\Delta x = R/16$, where $R$ denotes the side of the square obstacle.
-The channel has a dimension $L_x \times L_y$ with $L_y = 8R$ and $L_x = 4L_y = 32R$.
-The lattice therefore has a resolution of $N_y = 129$ lattice nodes in the $y$-direction and $N_x = 513$ nodes in the $x$-direction.
-The grid spacing is set to $g = 8\Delta x$.
-The characteristic velocity at the inlet is set to $U_0 = 0.05 \frac{\Delta x}{\Delta t}$, where $\Delta t$ is the lattice timestep.
-\textcolor{red}{@todo@Emmanuel: Briefly explain why the low Mach number and the nearly incompressible limit.}
-Finally, the viscosity is set so that the Reynolds number is $Re_{grid} = U_0g/\nu = 1200$.
-The Reynolds number based on the mean flow velocity and the size of the obstacle is $Re_{obs} \approx 1700$.
-
-% Boundary conditions %
-
-The channel walls and the surface of the obstacle are decribed by the halfway bounce-back scheme~\cite{ref_BB}\textcolor{red}{@todo@Emmanuel: Provide citation for halfway bounce-back boundary condition}.
-The open boundary at the outlet is based on a second order extrapolation of the velocity along the $x$-direction.
-The extrapolated velocity is then enforced using a regularization based on a finite difference estimation of the local stress tensor~\cite{latt2008straight}.
-The parabolic profile at the inlet is imposed by setting populations at the corresponding equilibrium, see appendix~\ref{sec:lbm}.
-\\
+% geometry
+%
+The simulated flow develops in a long plane channel of dimension $512 \times 128$ mesh points.
+The squared obstacle has size $R=16$ (in mesh unit) and is located at the center of the channel.
+The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
+% boundary conditions
+%
+No-slip boundary conditions are enforced on top and bottom walls of the channel by using an halfway bounce-back procedure \cite{lbm_book}.  
+%
+Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
+The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $(\Delta x/\Delta t)$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The distributions are imposed at equilibrium (see Appendix).
+In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$.
+The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method.
+Near the end of the channel, the flow is progressively damped within a sponge layer where the viscosity is artificially enhanced.
+Finally, the oulet boundary condition relies on a second-order extrapolation of the velocity and mass density.
+Then, the extrapolated distributions are evaluated through a regularization procedure (development onto the first two tensor Hermite polynomials) relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
 
@@
 
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle. The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and reads 
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
+	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
+	The vorticity is given in lattice units. \todo{THE COLORMAP SEEMS TO BE REVERSED IN THE WHOLE PAPER: BLUE SHOULD BE NEGATIVE AND RED POSITIVE: $\omega = \partial_x u_y - \partial_y u_x$}}
+	\label{fig:typical_vorticity}
+\end{figure}
+
+% drag signal
+%
+The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
+The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and expresses formally as
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{S}$ is the surface of the obstacle. 
-In our situation, the viscous stress makes a negligible contribution to the drag.
-This latter therefore results mostly from pressure forces, which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
+where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
+Here, the viscous stress makes a negligible contribution to the drag, which therefore results mostly from pressure forces.
+% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
 %
 Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag may be viewed as the pressure difference 
+As a consequence, the drag can be expressed as the pressure difference 
 \begin{equation}
 \label{eq:drag_approx}
-  f_d(t) = p_{fb}(t) - p_{b}(t)
+f_d(t) = p_{fb}(t) - p_{base}(t)
 \end{equation}
-where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base of the obstacle, respectively.
+between the pressure integrated over the forebody ($p_{fb}(t)$) and the base of the obstacle ($p_{base}(t)$).
 
-%% Definition du turnover time
-The typical timescale of drag fluctuations may be estimated from dimensional analysis as
+% Definition of turnover time
+%
+The typical timescale (turnover time) of drag fluctuations may be estimated from dimensional analysis as
 \begin{equation}
 \label{eq:turnover_time}
-  \tau_0 = \frac{R}{U},
+\tau_0 = \frac{R}{U}
 \end{equation}
-where $R$ is the diameter of the cylinder and $U$ the mean-flow velocity.
-Let us mention that the viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+where $R$ is the size of the square and $U$ the averaged velocity. 
+The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+%
+% typical fluctuation of the drag
+%
+Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
+It is observed that some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
+Therefore, the base pressure varies much less than the forebody pressure, and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
+In section~\ref{sec:direct_sampling}, we shall highlight that the situations leading to  {extreme} fluctuations of the drag are different.
+In particular, extreme fluctuations mainly result from large variations of the base pressure, whereas forebody pressure variations play a lesser role.
 
-% Description du champ de vort. correspondant a une fluctuation typique
-Figure~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle over a few turnover times.
-The vorticity generated along the boundaries of the obstacle is advected downstream by the mean flow.
-As a consequence, the base pressure varies much less than the forebody pressure, and typical drag fluctuations mostly result from the fluctuating pressure at the front of the obstacle.
-
-In the section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
-In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a lesser role.
+\subsection{The drag as a random process: Probability Density Function, correlation in time}
+\label{sec:pdfs}
 
@@
 	\centering
-  \includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-  \caption{Vorticity dynamics corresponding to typical drag fluctuations over four turnover times~\eqref{eq:turnover_time}.
-Negative vorticity is generated along the bottom boundary, and advected away from the the obstacle by the mean flow.}
-  \label{fig:typical_vorticity}
-\end{figure}
-
-\subsection{The drag random process: Probability Density Function, time correlations}
-\label{sec:pdfs}
-
-% Description du signal de trainee typique
-Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-It is therefore natural to model the drag as a scalar random process $f_d(t)$.
-\begin{figure}
-  \centering
 	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
+	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
 The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
@@
 
-Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
-The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
+% drag signal
 %
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
+Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square over five hundred turnover times.
+The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+It is therefore natural to model the drag as a (scalar) random process $f_d(t)$.
+
 \begin{figure}
   \centering
-    \subfloat[\ac{pdf} of drag fluctuations around mean.]
+    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
     {\label{fig:pdf_drag_a}
     	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[Autocorrelation of drag.]
+    \subfloat[Autocorrelation of drag]
     {\label{fig:pdf_drag_b}
     	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
-      \caption{\textbf{(a)} Estimate of the zero-mean \ac{pdf} for drag fluctuations $f'_d = (f_d - \bar{f_d})$ where $\bar{f_d}$ denotes the averaged value.
-The drag is evaluated both in the presence and absence of the obstacle.
-Note that both \acsp{pdf} have different variance, and are here not normalized.
-  \textbf{(b)} Autocorrelation function of the drag defined as $\left( \langle f_d(t+\tau)f_d(t) \rangle-\langle f_d \rangle ^2 \right) /\sigma^2$.
-The lag $\tau$ is normalized by the turnover-time~\eqref{eq:turnover_time}.}
+      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$, where $\bar{f_d}$ denotes the time averaged value.
+The drag is evaluated both in the presence (red) and in the absence of the obstacle (blue).
+Note that the \acsp{pdf} not been normalized.
+  \textbf{(b)} Autocorrelation function of the drag defined as $\overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
+The time lag has been normalized by $\tau_c \simeq 4\tau_0$ so that $C(1)=0$: $\tau_c$ can be considered as the correlation time of the drag signal.}
   \label{fig:pdf_drag}
 \end{figure}
+
+% drag statistics
+% PDF and correlation time
+%
+Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$.
+This very long simulation will be referred to as the \textit{control run} in the remainder.
+It has been made possible by the relative simplicity of the investigated two-dimensional flow; its execution time is about a few weeks (wall-clock time).
+The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
+It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(\tilde f_d) \propto e^{-\lambda \tilde f_d}$.
+%
+Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a (virtual) control surface along the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
+This suggests that the asymmetry of the \ac{pdf} and the positive exponential tail are closely related to the no-slip condition on the obstacle.
+%
 %
 Lastly, the autocorrelation function of the drag is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found drag fluctuations are correlated over a time scale of the order of $\tau_0$, suggesting that the drag signal looses its memory typically after the sweeping of a few eddies past the obstacle.
+It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
+The drag therefore looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
 %
-This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
+This observation has important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
 %
-In the following we define the \textit{correlation time} $\tau_c = 4\tau_0$.
-Note that the ratio $\tau_0 / \tau_c$ may be viewed as a \textit{Strouhal number} $St$.
-A value $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
+In the remainder, $\tau_c$ will be referred to as the \textit{correlation time} of the drag.
+The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
 
@@
 
-In this section we describe the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.
+\begin{figure}
+	\centering
+	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
+	\caption{Amplitude of drag fluctuation (in standard-deviation unit) as a function the corresponding return time.
+The return time $r(a)$ is defined as the typical period of occurrence of a fluctuation of amplitude larger than $a$.}
+	\label{fig:return_time_instant}
+\end{figure}
 
-How rare is a fluctuation $f_d \geq a$ can be determined by computing its \textit{return time} $r(a)$.
-It is defined as the average time between two consecutive occurrences of a fluctuation $f_d \geq a$.
-Extreme drag fluctuations are \textit{rare events}, that is fluctuations for which $r(a) \gg \tau_c$.
-In this context, the occurrence of the fluctuations follows a Poisson process and~\cite{lestang_computing_2018}
+% return time given from direct sampling
+%
+As a first step, the phenomenology of extreme fluctuations of the drag is investigated in this section by means of \textit{direct sampling} of the control run.
+This brute-force approach is here used as opposed to optimized approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
+Nevertheless, this direct approach will provide a trustworthy baseline for the validation of these elaborate algorithms.
+
+How rare is a fluctuation $f_d \geq a$ can be quantified by its so-called \textit{return time} $r(a)$.
+This latter is defined as the average time between two consecutive occurrences of events of amplitude $f_d \geq a$.
+\todo{suffisamment clair?
+un schema pour fixer les choses?}
+Extreme drag fluctuations are \textit{rare events} in the sense that $r(a) \gg \tau_c$ (correlation time).
+In that case, the occurrence of the fluctuations follows a Poisson process \cite{lestang_computing_2018} and
 \begin{equation}
   \label{eq:return_time}
-  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\alpha a}
+  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\lambda a}
 \end{equation}
-where $\alpha$ is the rate describing the positive tail of the drag \ac{pdf}, see figure~\ref{fig:pdf_drag}.
-Figure~\ref{fig:return_time_instant} illustrates the fluctuation amplitude $a$ as a function of the return time $r(a)$, for the drag process $f_d(t)$.
-Such plot can be obtained from a sample timeseries following a simple procedure~\cite{lestang_computing_2018}.
+where $\lambda$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
+%
+Fig.~\ref{fig:return_time_instant} illustrates the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$ computed from the \textit{direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
+It is found that the return time indeed diverges  exponentially with the amplitude of fluctuation.
+Let us also remark that under-sampling effects explain the deviation from the exponential law at the largest amplitudes.
+% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
 
-In the following we discuss extreme fluctuations of both the instantaneous drag $f_d$ and time-averaged drag with return times larger than $10^4\tau_c$.
-In this section extreme drag fluctuations are simply sampled by simulating the flow over a very long duration.
-This approach is referred to as \textit{direct sampling}, as opposed to approaches involving the use of rare events algorithms, discussed in section~\ref{sec:rare_events_algorithms}.
-Such approach is made possible by the relative simplicity of the flow described in section~\ref{sec:test_flow}.
-The control run mentioned in the previous section, spanning $T_{tot} = 10^6 \tau_c$, could be achieved in a wall-clock time of a few weeks and gives access to extreme fluctuations $f_d \geq a$ with a return time $r(a) \leq 10^6\tau_c$.
 \subsection{Extracting extreme drag fluctuations from a very long timeseries}
@@
 
-On the basis of the control timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we tracked every drag fluctuation with an amplitude greater than a fixed threshold amplitude $a$.
-The value of this threshold was chosen so that the return time of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
-Figure~\ref{fig:return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
-According to figure~\ref{fig:return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard deviation of the drag process.
-We denote by $(t^{\star}, f_d^{\star})$ a fluctuation of peak amplitude $f_d^{\star}$ occurring at time $t^{\star}$.
-From the control timeseries, we identified $104$ independent extreme drag fluctuations whith an amplitude greater than $7.6\sigma$, \textit{i.e.} with a return time greater than $10^4\tau_c$.
-In the following sections~\ref{sec:forebody_and_base_contribution} and~\ref{sec:instantaneous_drag}, we describe the phenomenology of extreme drag fluctuations on the basis of this set of
-events.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.6\linewidth]{return_time/return_time.png}
-  \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
-  \label{fig:return_time_instant}
-\end{figure}
+% intro
+%
+We have extracted from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$ the fluctuations of the drag with a return time $r(a)$ greater than  $10^4~\tau_c$.
+This set will be considered as representative of \emph{extreme events} in the upcoming study.
+This choice has been driven by the need to collect enough large-amplitude events and possibly identify generic features.
+%
+According to Fig.~\ref{fig:return_time_instant}, the related threshold amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
+Therefore, 104 independent excursions with $f_d(t) > 7.6~\sigma$ have been identified.
+For convenience, each excursion will be characterized by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
+%
+The phenomenology of extreme drag fluctuations is now examined on the basis of this set of events.
 
@@
 \label{sec:forebody_and_base_contribution}
-In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
-We now investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuations, in the case of the sampled \textit{extreme} drag fluctuations.
-Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
-In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
-It can be further decomposed into 
-\begin{equation}
-  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
-\end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
 
-Figure~\ref{fig:pressure_ratio} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.
 \begin{figure}
@@
 	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-  \caption{\label{fig:pressure_ratio} Contribution of the forebody and base pressure fluctuation to the overall drag fluctuation for the 104 sampled extreme drag fluctuations in the control run.
-To each of these fluctuation correspond a unique value of the drag $f^{\star}_d$, and a unique pair ($\tilde{p}^{\star}_{base}$,$\tilde{p}^{\star}_{fb}$).
-The dashed lines are first order least square fits of the data.
-They highlight that, as the value of the fluctuation increases, so is the relative contribution of the forebody pressure fluctuation.
-}
+	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
+An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
+	As the amplitude of the drag increases, so does the relative variation of the base pressure.
+\todo{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
 \end{figure}
-It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the overall drag fluctuation.
-By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinder are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
-In addition, figure~\ref{fig:pressure_ratio} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation, relatively to the forebody pressure fluctuation.
 
-\subsubsection{Dynamical aspects of extreme drag fluctuations}
+% teasing
+%
+In section~\ref{sec:test_flow}, it has been pointed out that typical drag fluctuations  originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
+We shall see that the situation is different in the case of {extreme} drag fluctuations.
+
+% relative contribution of forebody and base pressure
+%
+Let $(t^{\star}, f_d^{\star})$ refer to an extreme drag event.
+The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
+\begin{equation}
+  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star}
+\end{equation}
+where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the variations of the forebody pressure and base pressure, respectively.
+%
+Fig.~\ref{fig:pressure_ratio} displays the relative contributions
+$\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
+\todo{please confirm signs! I have added a minus sign: $-\tilde{p}_{base}^{\star}$}  
+%
+It is found that the base pressure variation contributes typically $80\%$ to the overall drag fluctuation.
+Therefore, extreme amplitude of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
+Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the contribution of the base pressure relatively to the forebody pressure.
+
+\subsubsection{Fluid dynamics related to extreme drag fluctuations}
 \label{sec:dynamical_aspects}
@@
 	\centering
-  \includegraphics[width=.8\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-  \caption{\label{fig:top_4_events_vorticity} Vorticity field at $t=t^{\star}$ for the four highest drag fluctuations in the timeseries.
-    The high value for the drag results from the generation of an intense vortex inducing a pressure drop at the base of the obstacle.
-The formation of such a structure is aided by important negative (positive) vorticity production at the bottom (top) boundary of the obstacle coupled with the influence of positive (negative) vorticity in the vicinity of the base.}
+	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
+	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centered around extreme fluctuations occurring  at $t=t^{\star}$.
+The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
+Extreme drag events exhibit a typical lifetime of one correlation time $\tau_c$.
+The profile is slightly skewed indicating that the step up is slower than the return to typical values.
+\todo{ $\tilde f_d$ rather than $f_d$ in figure?}}
+\end{figure}
+
+% mean profile around burst
+%
+The focus is now on the flow scenarii that yield extreme values of the drag. At first,
+Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
+A peaked profile is observed with a width corresponding roughly to one correlation time $\tau_c$.
+This implies in particular that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
+%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
+Interestingly, the profile is also slightly skewed indicating that the step up of the drag  is slower than the return to typical values past the peak value.
+%Moreover, such very high drag levels do not persist over time.
+To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
+  \caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
+  }
 \end{figure}
@@
   \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-  \caption{\label{fig:density+streamlines} Pressure field and velocity streamlines in the vicinity of the obstacle at $t=t^{\star}$.
-Because of the blocking vortex, vorticity produced along the top boundary of the obsatcle has no other choice but to accumulate at the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-  \caption{\label{fig:timeseries_extremes} Average drag timeseries in the vicinity of an extreme drag fluctuation at $t=t^{\star}$.
-The blue continuous line results from an average over the 104 extreme events sampled in the control run.
-This illustrates that extreme drag variations are very local in time, with a typical lifetime of one correlation time unit $\tau_c$.}
+  \caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
+In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.
+\todo{$f_d$ or $\tilde f_d$?}}
 \end{figure}
@@
   \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-  \caption{\label{fig:vorticity_dynamics} Vorticity dynamics associated to an extreme fluctuation of the drag acting on the square obstacle at $t=t^{\star}$.}
+  \caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
 \end{figure}
 
-We now turn to a description of the flow dynamics that correspond to extremely high values of the drag acting on the square obstacle.
+% flow scenario
+%
+Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
+%
+In each case, an intense vortical structure is visible near the base of the obstacle.
+The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
+%
+The formation of this vortex originates from an intense negative (positive) vorticity layer at the top (bottom) boundary of the obstacle. 
+%
+This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strongly enhanced drag.
+In contrast, nothing special happens near the forebody of the obstacle.
 
-Figure~\ref{fig:timeseries_extremes} displays the average of the drag timeseries corresponding to the 104 extreme events sampled in the control run, in the vicinity of the fluctuation.
-It illustrates that extreme drag fluctuations are very local in time, with a lifetime of roughly one correlation time unit.
-Starting from typical values, extreme drags are typically reached in less than a correlation time.
-Moreover, such very high drag levels do not persist over time.
-To understand why it is so, it is useful to visualize the corresponding flow fields.
+The high pressure drop on the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
+As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents him to be swept away downstream.
+% 
+This scenario is better evidenced in Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
+%
+Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
+This positive vorticity then prevent the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
+When the blocking vortex is in turn advected downstream, the vortex against the base is released.
+%
+Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past a region of size comparable to the size of the obstacle, and is therefore of the order of $\tau_c$.
+This scenario is generic and has been observed for most extreme events sampled in the control run.
 
-To begin with, figure~\ref{fig:top_4_events_vorticity} displays the vorticity field corresponding to the peak drag value, for a subset of the extreme events sampled in the control simulation.
-We observe that, in each case, extreme drag flcutuations correspond to very similar flow configurations.
-More precisely, a region of strong vorticity is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_vorticity}.
-This vorticity results from a boundary layer, along either the top or bottom side of the obstacle.
-Such extreme vorticity is responsible for a significant pressure drop in the vicinity of the base of the obstacle, which in turn leads to an extreme value of the drag, see figure~\ref{fig:pressure_ratio}.
+%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
 
-How are such vorticity reached ?
-In contrast with the typical flow dynamics, illustrated in figure~\ref{fig:typical_vorticity}, the vorticity generated within the boundary layers is not directly advectd by the means flow.
-Figure~\ref{fig:density+streamlines} shows that, instead, this vorticity is constrained in the vicinity of the base of the obstacle.
-Such constraint results from a large vortical region located roughly one obstacle length of the base, that prevents vorticity to be advected downstream.
-
-Figure~\ref{fig:vorticity_dynamics} illustrate the time evolution of the vorticity field, in the vicinity of the drag fluctuation, for the event depicted in figure~\ref{fig:density+streamlines}.
-From $t^{\star}-0.4\tau_c$ to $t^{\star}-0.16\tau_c$, a strong negative eddy is generated along the bottom boundary layer of the obstacle and advected downstream by the mean flow.
-This leads to a region of negative vorticity at roughly one obstacle length of the base of the obstacle, constraining the positive vorticity generated along the top boundary.
-However, it can be seen in figure~\ref{fig:vorticity_dynamics} that the positive vorticity only accumulates for a short duration.
-The blocking negative vortex is in turn advected downstream, thus freeing the trapped positive vorticity.
-
-The lifetime of an extreme drag fluctuation therefore corresponds to the duration over which the blocking vortex remains in the vicinity of the base.
-Consistently with figure~\ref{fig:timeseries_extremes}, this duration is roughly one correlation time unit $\tau_c$.
-
-We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-In order to identify such events, we monitored the averaged shear along the top or bottom boundary
-\begin{equation}
-  \label{eq:avg_shear_def}
-  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary.
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
 \begin{figure}
@@
 	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-  \caption{\label{fig:shear_asof_drag} Evolution of the shear alongised the top or bottom boundary of the obsatcle as a function of the drag, for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-To each oneof the grey lines correspond a unique extreme event.
-The blue line is the average over the 104 extreme event sampled in the control run.}
+	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
+Each trajectory corresponds to an event.
+The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
 \end{figure}
-Figure~\ref{fig:shear_asof_drag} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
-For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
-As illustrated in figure~\ref{fig:timeseries_extremes}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
-As a consequence, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
+
+Since the occurrence of large drag amplitudes originates from the production of vorticity along the top or bottom side of the square, it is now proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square.
+Precisely,
+\begin{equation}
+  \label{eq:avg_shear_def}
+  \overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x}
+\end{equation}
+where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
+%
+% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
+% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
+%
+Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the sampled extreme events.
+Before and after the extremal fluctuation, \textit{i.e.}  $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in the region related to typical values of both $\overline{\gamma}$ and $f_d$.
+On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
+%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
+This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations, as also described in the previous flow scenarii.
 
-The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
+%The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
 In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
@@
 For a more detailed description of such dynamics, see chapter 3 of~\cite{lestang:tel-01974316}.
+
+\newpage
 
@@
   \centering
-  \includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+  %\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
   \caption{Instantaneous drag timeseries corresponding to the highest fluctuations of the time-averaged drag $F_T$ in the control timeseries. The averaging window is $T=10\tau_c$.
@@
 \label{sec:lbm}
-While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
-Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
-Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
-This is, in short, the rationale behind the LB approach.
-In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
-\\
-More precisely, the fluid is viewed as populations of particles
-that collide, redistribute and propagate along the different links of a discrete lattice. 
+
+% details about LBM
+In the LB method, the fluid is viewed as populations of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
@@
 Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
+By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
 	\label{fig:D2Q9}
@@
 
+% algo
+%
 The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+The LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
 Namely, an \emph{exact} streaming step
@@
 \]
-during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
+during which particles move with their own velocity to a neighboring node, and an instantaneous collision step
 \[
@@
 \]
-which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
+which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
+The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
 \[
@@
 \]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas.
+%
+The equilibrium function is given  by
 \[
@@
 \] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. Einstein summation convention is assumed.
+Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
 $\Delta x \to 0$.
-\\
+
 As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
@@
 \end{equation}
+Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \todo{REF ADR}.
 
 \bibliographystyle{unsrt}
-\bibliography{draft_p2}
+\bibliography{biblio}
 

### 2019-05-14 figures/dynamics_extremes/plot_dynamics.py
Reverse vorticity field

@@ -12,7 +12,7 @@ N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
 
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+prefix = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/"
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is lowest fluctuation
@@ -60,6 +60,9 @@ for i in range(0,nplots):
         dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
         dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
         vort = dxy - dyx
+        # The sign of the vorticity is wrong if using a right-handed coord. system.
+        # The velocity field should probably be flipped upside-down
+        vort = - vort
 
         # Print fluctuation amplitude as title
         timeFromPeak=(offset+1-50)*0.04

### 2019-05-14 figures/ecoulement_typique/plot_ecoulement_typique.py

@@ -12,7 +12,8 @@ N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
 
-prefix = './'
+prefix = '/home/tlestang/transition_hdd/thibault/These/' \
+         + 'articles/data/ecoulement_typique/'
 
 fig = plt.figure(figsize=(16,7), constrained_layout=True)
 # Declares a 2x5 grid.
@@ -51,6 +52,9 @@ for i in range(0,nplots-1):
         dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
         dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
         vort = dxy - dyx
+        # The sign of the vorticity is wrong if using a right-handed coord. system.
+        # The velocity field should probably be flipped upside-down
+        vort = - vort
 
         # Print time as title
         timeFromBeg=0.1+i*0.2

### 2019-05-14 figures/illustr_ecoulement/plot_ecoulement.py

@@ -1,0 +1,77 @@
+# Produces a figure illustrating the vorticity field in the whole comp. domain
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
+
+Dx = 513
+Dy = 129
+N = Dx*Dy
+F = np.zeros(104)
+idx = np.zeros(104)
+# Make use of flow data generated for figure ecoulement typique
+prefix = '/home/tlestang/transition_hdd/thibault/These/' \
+         + 'articles/data/ecoulement_typique/'
+
+fig, ax = plt.subplots(figsize=(16,7), constrained_layout=True)
+# Declares a 2x5 grid.
+
+path_to_file_ux = prefix+'ux.dat'
+path_to_file_uy = prefix+'uy.dat'
+fx = open(path_to_file_ux, "rb")  # reopen the file
+fy = open(path_to_file_uy, "rb")  # reopen the file
+bbox = dict(boxstyle="round", fc="0.9")
+
+# Get velocity fields
+ux = np.fromfile(fx, float, N, "")
+ux = np.reshape(ux, (129,513), 'F')
+uy = np.fromfile(fy, float, N, "")
+uy = np.reshape(uy, (129,513), 'F')
+    
+ymin=1;ymax=128
+xmin=1;xmax=512
+# Compute vorticity using finite-differences
+dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
+dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
+vort = dxy - dyx
+# The sign of the vorticity is wrong if using a right-handed coord. system.
+# The velocity field should probably be flipped upside-down
+vort = - vort
+
+im=ax.imshow(vort, interpolation='spline36', cmap='BrBG', vmin=-0.07, vmax=0.07)
+
+# Adds a filled Rectangle to mark the square obstacle
+rect = ax.add_artist(Rectangle((256,55), 14, 14))
+rect.set_color('#545454')
+# Show the grid
+R = 8
+ny,nx  = ux.shape
+n_lamina = ((ny-1)/R)/2
+x0 = (nx-1)/16
+for i in range(0,int(n_lamina)):
+    y_start = 2*i*R + R/2
+    y_end = y_start + R
[…]

### 2019-05-14 figures/illustr_extrms_vorticity/make_plot_vorticity.py

@@ -11,7 +11,7 @@ Dy = 129
 N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is lowest fluctuation
@@ -59,6 +59,9 @@ for i in range(0,Nevents):
         dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
         dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
         vort = dxy - dyx
+        # The sign of the vorticity is wrong if using a right-handed coord. system.
+        # The velocity field should probably be flipped upside-down
+        vort = - vort
 
         # Print fluctuation amplitude as title
         tit='{:2.1f}$\\sigma$'.format(F[Nevents-i-1])

### 2019-05-15 figures/illustrate_return_time/illustrate_return_time.py
Update illustrate_return time script

@@ -42,8 +42,6 @@ for i, bundle in enumerate(list_of_bundles):
     point_array[i] = tspan[np.argmax(f[bundle])+bundle[0]]
     value_array[i] = np.amax(f[bundle])
 
-
-
 fig,ax = plt.subplots(1,1,figsize=(12,4), constrained_layout=True)
 ax.set_xlim(0, T)
 line, = ax.plot(tspan, f)
@@ -51,6 +49,16 @@ line2, = ax.plot(point_array, value_array, 'o', linestyle='None', markersize=4)
 line3, = ax.plot([tspan[0], tspan[-1]], [a, a], linestyle='--', linewidth=.7)
 line3.set_color(line2.get_color())
 line.set_linewidth(.8)
+
+y0 = 0.20
+eps = 10
+props=dict(arrowstyle="->", color=line2.get_color(),
+                length_includes_head=True, width=0.002,
+                head_width=0.01, head_length=5)
+
+for x_point, x_point_fwd in zip(point_array, point_array[1:]):
+    plt.annotate('', xy=(x_point+eps, y0), xytext=(x_point_fwd-eps, y0),
+                 arrowprops=props)
 
 ax.set_xlabel(r'$t / \tau_0$',fontsize=22)
 plt.ylabel('$f_d(t)$',fontsize=22)

### 2019-05-15 draft_p2.tex
Deuxieme partie de la relecture d'Emmanuel, 15/05

@@
 \DeclareAcronym{OU}{
-	short=OU,
+	short=O-U,
 	long=Ornstein--Ulhenbeck
@@
 
-\newcommand{\todo}[1]{{\color{red}{#1}}}
+\newcommand{\EL}[1]{{\color{red}{#1}}}
+\newcommand{\TL}[1]{{\color{green}{#1}}}
 
@@
 \begin{abstract}
+	\EL{ ABSTRACT TO BE REVIEWED}
 This study evaluates the relevance of  rare-event sampling to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
@@
 %
-Such technique has already proved its efficiency for simulating extreme events in {<...>} with significant run-time savings.
+Such technique has already proved its efficiency for simulating extreme events in {...} with significant run-time savings.
 Application to fluid-structure interaction problems remains mainly open. 
 %
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely {<...>}.
+The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely {...}
 Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort.
-However, {<limitations and pitfalls>}.
+However, {limitations and pitfalls}.
 % Importantly, such techniques also provide direct access to the flow scenario yielding extreme events. In the present case, it is found that {<blocking effect>}.
@@
 The use of perturbative methods in identifying resonant interactions (among degrees of freedom) possibly responsible for extreme fluctuations is not successful.
-Simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
+Therefore, simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
 Nevertheless, this requires very long simulation since these events are also very rare.
@@
 In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium.
-This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}, which provides a framework for rare-event sampling.
+This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013} that provides a framework for rare-event sampling.
 %
@@
 %
-In the present paper, a more straightforward approach is investigated.
+In this paper, a more general approach is investigated.
 It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.
@@
 %
-An original contribution of our study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \todo{TRUE?}
+An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \EL{is it TRUE?}
 % 
@@
 
-The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a more detailed description will be given later in the paper.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \cite{aristoff_adaptive_2015}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
-
-Over the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
+The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
+In recent years, this algorithm has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
+
+Over the last decade, the main theoretical framework for the study of rare events (in statistical physics) has been the theory of large deviations \cite{touchette_large_2009}.
 Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
@@
 The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square.
+In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
 The vorticity is displayed with an arbitrary color map from blue (negative values) to red (positive values).}
@@
 We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
-Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
+%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
 %
@@
 Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-%namely the Navier-Stokes dynamics in the weakly compressible limit
 Further details are given in Appendix and references therein.
@@
 %
-No-slip boundary conditions are enforced on top and bottom walls of the channel by using an halfway bounce-back procedure \cite{lbm_book}.  
+No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure \cite{lbm_book}.  
 %
 Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
-The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $(\Delta x/\Delta t)$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
 The distributions are imposed at equilibrium (see Appendix).
@@
 The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method.
-Near the end of the channel, the flow is progressively damped within a sponge layer where the viscosity is artificially enhanced.
+Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
 Finally, the oulet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-Then, the extrapolated distributions are evaluated through a regularization procedure (development onto the first two tensor Hermite polynomials) relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
 
@@
 	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
-	The vorticity is given in lattice units. \todo{THE COLORMAP SEEMS TO BE REVERSED IN THE WHOLE PAPER: BLUE SHOULD BE NEGATIVE AND RED POSITIVE: $\omega = \partial_x u_y - \partial_y u_x$}}
+	The vorticity is given in lattice units. \EL{THE COLORMAP IS REVERSED IN THE WHOLE PAPER: BLUE SHOULD BE NEGATIVE AND RED POSITIVE}}
 	\label{fig:typical_vorticity}
@@
 \end{equation}
-between the pressure integrated over the forebody ($p_{fb}(t)$) and the base of the obstacle ($p_{base}(t)$).
+between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.
 
@@
 Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-It is observed that some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
-Therefore, the base pressure varies much less than the forebody pressure, and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
-In section~\ref{sec:direct_sampling}, we shall highlight that the situations leading to  {extreme} fluctuations of the drag are different.
-In particular, extreme fluctuations mainly result from large variations of the base pressure, whereas forebody pressure variations play a lesser role.
+It is observed that, typically, some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
+Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
+In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.
+%In particular, extreme fluctuations mainly result from large variations of the base pressure, whereas forebody pressure variations play a lesser role.
 
@@
 	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
+The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
 	\label{fig:typical_drag_signal}
@@
 %
-Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square over five hundred turnover times.
+Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
 The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-It is therefore natural to model the drag as a (scalar) random process $f_d(t)$.
+It is therefore natural to consider the drag as a random process.
 
@@
     	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[Autocorrelation of drag]
+    \subfloat[Autocorrelation of drag fluctuations]
     {\label{fig:pdf_drag_b}
     	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
-      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$, where $\bar{f_d}$ denotes the time averaged value.
-The drag is evaluated both in the presence (red) and in the absence of the obstacle (blue).
-Note that the \acsp{pdf} not been normalized.
-  \textbf{(b)} Autocorrelation function of the drag defined as $\overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The time lag has been normalized by $\tau_c \simeq 4\tau_0$ so that $C(1)=0$: $\tau_c$ can be considered as the correlation time of the drag signal.}
+      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
+The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
+%Note that the amplitudes have not been normalized.
+  \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
+The correlation time $\tau_c$ is given by $C(\tau_c)=0$.
+\EL{mettre $\tau/\tau_c$ en abscisse de l'autocorrelation (b)}
+  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
+}
   \label{fig:pdf_drag}
@@
 %
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$.
+Drag fluctuations have been recorder during a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$.
 This very long simulation will be referred to as the \textit{control run} in the remainder.
-It has been made possible by the relative simplicity of the investigated two-dimensional flow; its execution time is about a few weeks (wall-clock time).
+It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method.
+%The related execution time is about a few weeks (wall-clock time).
 The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
@@
 %
-Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a (virtual) control surface along the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
+Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a control surface corresponding the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that latter case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
 This suggests that the asymmetry of the \ac{pdf} and the positive exponential tail are closely related to the no-slip condition on the obstacle.
@@
 %
-Lastly, the autocorrelation function of the drag is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-The drag therefore looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation has important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the remainder, $\tau_c$ will be referred to as the \textit{correlation time} of the drag.
+Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
+It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$ for which $C(\tau_c)=0$.
+Therefore, the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+%
+This observation will have important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
+%
+In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag.
 The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
@@
 	\caption{Amplitude of drag fluctuation (in standard-deviation unit) as a function the corresponding return time.
-The return time $r(a)$ is defined as the typical period of occurrence of a fluctuation of amplitude larger than $a$.}
+The return time $r(a)$ is defined as the typical period of occurrence of a fluctuation of amplitude $a$.}
 	\label{fig:return_time_instant}
@@
 %
-As a first step, the phenomenology of extreme fluctuations of the drag is investigated in this section by means of \textit{direct sampling} of the control run.
+At first, the phenomenology of extreme fluctuations of the drag is investigated by means of \textit{direct sampling} of the control run.
 This brute-force approach is here used as opposed to optimized approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-Nevertheless, this direct approach will provide a trustworthy baseline for the validation of these elaborate algorithms.
+The direct approach will provide a trustworthy baseline for the validation of rare-events algorithms.
 
 How rare is a fluctuation $f_d \geq a$ can be quantified by its so-called \textit{return time} $r(a)$.
-This latter is defined as the average time between two consecutive occurrences of events of amplitude $f_d \geq a$.
-\todo{suffisamment clair?
+This latter is defined as the average time between two consecutive occurrences of events of amplitudes $f_d \geq a$.
+\EL{suffisamment clair?
 un schema pour fixer les choses?}
-Extreme drag fluctuations are \textit{rare events} in the sense that $r(a) \gg \tau_c$ (correlation time).
-In that case, the occurrence of the fluctuations follows a Poisson process \cite{lestang_computing_2018} and
+Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time: $r(a) \gg \tau_c$.
+In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and
 \begin{equation}
@@
 %
-We have extracted from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$ the fluctuations of the drag with a return time $r(a)$ greater than  $10^4~\tau_c$.
+We have extracted  the fluctuations of the drag with a return time $r(a)$ greater than  $10^4~\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
 This set will be considered as representative of \emph{extreme events} in the upcoming study.
-This choice has been driven by the need to collect enough large-amplitude events and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related threshold amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
-Therefore, 104 independent excursions with $f_d(t) > 7.6~\sigma$ have been identified.
-For convenience, each excursion will be characterized by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
-%
-The phenomenology of extreme drag fluctuations is now examined on the basis of this set of events.
+The choice of this particular threshold has been driven by the need to collect enough large-amplitude events and possibly identify generic features.
+%
+According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
+Precisely, 104 independent excursions \EL{excursion, terminologie correcte?} with $f_d(t) \geq 7.6~\sigma$ have been identified.
+For convenience, each excursion will be characterized in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
+%
+The phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
 
@@
 An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-	As the amplitude of the drag increases, so does the relative variation of the base pressure.
-\todo{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
+	As the amplitude of the drag increases, the relative variation of the base pressure also increases.
+\EL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
 \end{figure}
@@
 $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-\todo{please confirm signs! I have added a minus sign: $-\tilde{p}_{base}^{\star}$}  
+\EL{please confirm signs! I have added a minus sign: $-\tilde{p}_{base}^{\star}$}  
 %
 It is found that the base pressure variation contributes typically $80\%$ to the overall drag fluctuation.
-Therefore, extreme amplitude of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the contribution of the base pressure relatively to the forebody pressure.
+Therefore, extreme amplitude of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream the obstacle.
+Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure. 
+%relatively to the forebody pressure.
 
@@
 The profile is slightly skewed indicating that the step up is slower than the return to typical values.
-\todo{ $\tilde f_d$ rather than $f_d$ in figure?}}
+\EL{ $\tilde f_d$ rather than $f_d$ in figure?}}
 \end{figure}
@@
 %
-The focus is now on the flow scenarii that yield extreme values of the drag. At first,
+The focus is now on the flow scenarii that yield extreme values of the drag. 
 Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
 A peaked profile is observed with a width corresponding roughly to one correlation time $\tau_c$.
-This implies in particular that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
+This implies that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
 %Starting from typical values, extreme drags are typically reached in less than a correlation time. 
@@
 In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.
-\todo{$f_d$ or $\tilde f_d$?}}
+\EL{$f_d$ or $\tilde f_d$?}}
 \end{figure}
@@
 Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity then prevent the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
+This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
 When the blocking vortex is in turn advected downstream, the vortex against the base is released.
 %
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past a region of size comparable to the size of the obstacle, and is therefore of the order of $\tau_c$.
+Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$. 
 This scenario is generic and has been observed for most extreme events sampled in the control run.
@@
 
-Since the occurrence of large drag amplitudes originates from the production of vorticity along the top or bottom side of the square, it is now proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square.
+Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is now proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square.
 Precisely,
@@
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations, as also described in the previous flow scenarii.
+This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations in agreement with the previous description of flow scenarii.
 
@@
 
-\newpage
-
 \subsection{Extreme fluctuations of the time-averaged drag }
@@
 
-% Motivation
-In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $\tau_c$.
-However, In many applications, this duration is much smaller than the timescale of interest.
-Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-
-%% Definition de la trainée intégrée
-In such cases, a relevant observable is the \textit{time-averaged} drag
+\begin{figure}
+	\centering
+	%\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the time-averaged drag $F_T$ with averaging period $T=10~\tau_c$ and $\tau_c$ being the correlation time of the drag.
+$\sigma_T$ denotes the standard deviation of $F_T$.
+\EL{FIGURE IS MISSING}}
+	\label{fig:extreme_avg}
+\end{figure}
+
+% motivation
+%
+%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
+In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
+We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event was of the order of the sweeping time of the flow past the obstacle, or $\tau_c$.
+In practical applications, this duration may be much smaller than the characteristic response time of the material structure subject to these fluctuations.
+% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
+%
+% definition
+%
+Therefore, a relevant observable is the \textit{time-averaged} drag
 \begin{equation}
@@
 \end{equation}
-where $f_d$ denotes the instantaneous drag and $T\leq \tau_c$ a timescale of interest.
-In the following we consider the case where $T \gg \tau_c$, typically $T=10\tau_c$.
-In this context, a fluctuation of the average $F_T(t)$ is related to roughly $T / \tau_c$ independent fluctuations of the \textit{instantaneous} drag $f_d$, over the time interval $[t;t+T]$.
-
-%% Quelle est la phénoménologie de la trainée intégrée ?
-What is the phenomenology leading to extreme values of $F_T(t)$ ?
-Do an exceptionally large value of the average drag result from a single exceptionally large value of the instantaneous drag (case (1)) ?
-Or instead from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
-From the control timeseries, see section~\ref{sec:pdfs}, one can compute a control timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$.
-It is then possible to identify extreme positive fluctuations for which the time-averaged $F_T$ drag exceeds a fixed threshold $a$, in the very same way as described in section~\ref{sec:extreme_extraction}.
-Setting $a=5.2\sigma _T$ leads to $84$ independent events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.
-
-Figure~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t^{\star})\}_{t^{\star} \leq t \leq t^{\star}+T}$ for four of the sampled extreme fluctuations sampled in the control timeseries for $F_T$.
-It illustrates that the phenomenology of the extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are featured in figure~\ref{fig:extreme_avg}, along with intermediate cases where the very large value of the drag results from both a very large fluctuation and a large number of typical, however all positive, fluctuations of the instantaneous drag $f_d$.
-\begin{figure}
-  \centering
-  %\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-  \caption{Instantaneous drag timeseries corresponding to the highest fluctuations of the time-averaged drag $F_T$ in the control timeseries. The averaging window is $T=10\tau_c$.
-  Here $f_d$ denotes the zero-mean, unit variance instantaneous drag. The time-average is normalized by the standard deviation for the time-averaged drag $\sigma_T$.}
-  \label{fig:extreme_avg}
-\end{figure}
-
-This marginal phenomenology can be connected to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_computing_2018}.
-An estimate of this \ac{pdf} is displayed in figure~\ref{fig:pdf_drag}.
-Let $X$ be a random variable whose \ac{pdf} is denoted $\mathbb{P}$ and standard deviation $\sigma_X$.
-Considering an extreme positive value $a$ of $S_N=\sum_{n=1}{N}X_n$, the probability $p_1$(resp. $p_2$) of case (1) (resp. case (2)) writes:   
+where $f_d(t)$ denotes the instantaneous drag and $T$ is the timescale (response time) of interest. We shall consider $T=10~\tau_c$ in the following.
+
+During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be viewed as the overall contribution of roughly $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
+% phenomenology
+%
+% What is the phenomenology leading to extreme values of $F_T(t)$ ?
+It is thus legitimate to ask whether a large value of the averaged drag results from (case (1)) a single outstanding fluctuation of the instantaneous drag,  or from (case (2)) a series of moderate positive fluctuations.
+%
+%
+The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained straightforwardly by filtering in time the instantaneous drag signal $f_d(t)$. 
+In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$ and sample a set of extreme events.
+By taking $a=5.2~\sigma _T$  with $\sigma_T$ being the standard deviation of the fluctuations of $F_T$, $84$ independent extreme events have been selected.
+
+% neither case (1) nor case (2)
+%
+Fig.~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
+It is established that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
+Indeed, both cases are featured in Fig.~\ref{fig:extreme_avg} along with mixed situations where very large value of the averaged-drag results from both a very large fluctuation and a significant succession of moderate (positive) fluctuations of the instantaneous drag.
+%
+% theoretical explanation
+%
+This observation may be explained by the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_computing_2018} (see Fig.~\ref{fig:pdf_drag}). 
+Indeed, let $X$ be a random variable with \ac{pdf} $\mathbb{P}(X)$ and standard deviation $\sigma_X$.
+Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
 \begin{equation}
   \label{eq:indep}
-  p_{1}(\sum_{1}^{N} X_n=a) \approx \mathcal{P}\left(\frac{a}{N}\right)^{N} \,\,\, \text{and} \,\,\, p_{2}(\sum_{1}^{N} X_n=a)\approx \mathcal{P}(a)  
-\end{equation}
-If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \ll \sigma_X}{\propto} e^{-\alpha x}$, then both cases (1) and (2) are equiprobable, provided that the average $a=S_N/N$ is very large:
-\begin{equation}
-  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\sim} C\left(e^{-\alpha \frac{a}{N}}\right)^{N} e^{-\alpha a} = 1.
+  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left(\frac{S_N}{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
+\end{equation}
+If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
+\begin{equation}
+  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
   \label{eq:ratio_exp}
@@
 
-\section{Rare event algorithms}
+\section{Rare-event algorithms}
 \label{sec:rare_events_algorithms}
 
-%% Limites de l'approche par échantillonnage direct
-In the limit of very rare events and/or very complex dynamics such as turbulent flows in industrial contexts, a direct sampling approach is unrealistic.
-Indeed, according to equation~\eqref{eq:return_time}, the return time of a fluctuation $f_d \geq a$ verifies $r(a) \propto e^{\alpha a}$.
-As a result, the computational cost required to sample events $f_d \geq a$ scales like $e^{\alpha a}$.
-
-%% Objectif des algo.
-In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
-of the flow presented in section~\ref{sec:test_flow}.
-The purpose of rare events algorithms is to make it possible to sample rare events for a computational cost well inferior to their return time.
-In sections~\ref{sec:ams} and~\ref{sec:gktl} we consider the \acl{ams} and \acl{gktl} algorithms, respectively.
-Both algorithms rely on the simulation of an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$, where $\{\mathbf{x}(t)\}_{0\leq t T_a}$ is a formal notation for a trajectory
-of duration $T_a$ in the phase space describing the dynamical system.
-In this case the dynamical system is the flow described in section~\ref{sec:test_flow}.
-Over the iterations of the algorithm, trajectories are duplicated or discarded from the ensemble according to precise selection rules that bias the sampling towards the rare events of interest.
-Such selection rules are designed in such a way that the sampling bias is known at each iteration of the algorithm.
-In other words, generated trajectories are given a statistical weight, which knowledge allows for the computation of both their probability and expectation values of any trajectory observables.
+% limitation of direct sampling
+%
+In the limit of very rare events and complex dynamics (such as turbulent flows in industrial or environmental domains) the computational cost of  direct sampling becomes prohibitive.
+According to Eq.~\eqref{eq:return_time}, the return time of a fluctuation $f_d \geq a$ scales as $r(a) \propto e^{\lambda a}$.
+As a consequence, the computational cost required to sample (directly) events of amplitude $f_d \geq a$ diverges exponentially as $e^{\lambda a}$.
+%% objective
+% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
+The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than that related to return times.
+%
+For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
+%
+%
+Both algorithms consider an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in the phase space.
+In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
+In brief, the trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} allowing us to sample more frequently rare events of some observables of interest, \textit{e.g.} the drag.
+These selection rules are designed so that the introduced statistical bias is known at each iteration of the algorithm.
+The generated trajectories receive a weight whose knowledge makes it possible to calculate the probability and expectation values of the observables.
+% These two algorithms will be detailed in the following.
 
@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{\label{fig:illustr_AMS} Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.
-The initial ensemble of blue trajectories is first generated.
-These trajectories are independant and can be simulated in parallel.
-On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded from the ensemble (dashed blue line).
-Among the two remaining trajectories, trajectory 3 is chosen at random and copied until $Q_1$.
-It is then freely simulated from the branching poin to $T_a$.
-In case of deterministic dynamics, as samll perturbation is introduced at the branching.
-This procedure is typically iterated $J$ times, or until all trajectories go beyond a fixed threshold $Q$.}
-\end{figure}
-
-% Introduction to AMS
-The \acl{ams}~\cite{cerou_adaptive_2007} inherits from the basic ideas of \textit{splitting} algorithms~\cite{KahnHarris1951}: the sampling of a rare event is made easier by splitting the corresponding dynamical path into a sequence of events that can be sampled with a higher probability~\cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-
+  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} \EL{ou TAMS?} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
+These trajectories are independent and can be simulated in parallel.
+On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed blue line).
+Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
+It is then freely simulated from the branching point to the final time $T_a$.
+In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
+This procedure is iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.}
+\end{figure}
+
+% Introduction of AMS
+%
+The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits the basic ideas of {splitting algorithms} \cite{KahnHarris1951}, in which the sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
+%a sequence of events that can be sampled with a higher probability  
+a sequence of ``short-cuts'' that lead more rapidly to the realization of this event \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+The related selection-mutation step of this algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
+%
+% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
+%
+% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
+%
 % Quick description of TAMS
-In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
-The \ac{tams} relies on the definition of a time-dependant \textit{score function} $\xi (\mathbf{x}(t),t)$, as well as an iterative procedure starting from an ensemble $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$ of independant trajectories.
-Each iteration consists in discarding the trajectories with the lowest maxima of the score function $\xi$ over the whole trajectory of duration $[0;T_a]$.
-Discarded trajectories are resampled based on member of the ensemble of trajectories that achieved a higher value of the score function at some point of their history.
-As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS}.
-For further details about the algorithm and corresponding mathematical results, see~\cite{lestang_computing_2018} and references therein.
-
-\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+%
+We consider here a variant \EL{generalization?} of \ac{ams} called \ac{tams}, in which the branching procedure relies on a time-dependent \textit{score function} $\xi (\mathbf{x}(t),t)$.
+Each iteration therefore consists in discarding the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$.
+Discarded trajectories are re-sampled according to the remaining trajectories.
+% As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS} .
+Further details about the algorithm and corresponding mathematical results are provided in \cite{lestang_computing_2018} and references therein.
+\EL{What is the difference between AMS and TAMS? why do we need TAMS rather than AMS, This is not mentioned in text.}
+
+\subsubsection{Application of the \ac{ams} \EL{OR TAMS?} to a simple case: the \acl{ou} process}
 \label{sec:ams_ou}
+
 % Example: OU process
@@
   \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling.
-The orange line represents the evolution of the maximum over resampled trajectories as a function of the computational cost $C_{TAMS}$ (iterations of the the selection-mutation step).
-The solid blue line is the analytical solution for the return time of amplitude $a$ for the \ac{ou} process~\cite{lestang_computing_2018}.}
+    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
+The orange line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
+The solid blue line is the analytical solution for the return time of amplitude $a$.
+\EL{how many trajectories in the ensemble?
+REPLACE $f_d$ with $x$ in figure} }
   \label{fig:comparaison_temps_de_retour}
 \end{figure}
-We temporarily let fluid dynamics aside and consider a one-dimensional \ac{ou} process
+
+Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered. This should allow us to highlight differences with fluid dynamics.
 \begin{equation}
@@
 \end{equation}
-We apply the \ac{tams} to simulate trajectories $\{x(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
-For the sake of simplicity, a single trajectory is resampled at each iteration of the \ac{tams}.
-We denote by $a_j$ the maximum of the resampled trajectory at iteration $j$.
-The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
-Figure~\ref{fig:comparaison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach.
+
+The \ac{tams} is applied  to a set of \EL{$N$} trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with the objective to sample fluctuations of $x\geq a$ with $a$ being very large compared to typical values of $x$.
+The score function is here $x(t)$ and only one single trajectory is re-sampled at each iteration.
+%
+%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+%
+The computational cost of the algorithm after $J$ iterations is related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{tams} algorithm with that of a direct sampling. 
 In the latter, the typical computational cost is simply the return time $r(a)$.
-illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
-More importantly, these very rare trajectories are sampled for a computational cost that several order of magnitude lower than the return time of the corresponding fluctuations.
-
-Undoubtedly, an \ac{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} and \ac{tams}.
-Indeed, the state space is one-dimensional and the choice of the score function poses no questions: it is simply the observable $x$ itself.
-In addition, the noise term in~\eqref{eq:ou} has no correlations in time, which means that newly generated trajectories will quickly separate from their ancestors.
-
-\subsubsection{The \ac{ams} for extreme drag fluctuations}`
+Obviously, the successive re-samplings of the \ac{tams} algorithm lead very rapidly to trajectories exhibiting extreme fluctuations.
+The related computational cost is many orders of magnitude lower than that of direct sampling.
+
+Undoubtedly, the \acl{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} or \ac{tams} algorithm.
+Indeed, the state space is one-dimensional and the choice of the score function is straightforward. It is the process $x(t)$ itself.
+In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+
+\subsubsection{The \ac{ams} for extreme drag fluctuations}
 \label{sec:ams_drag}
-% Application to turbulent flow past an obstacle
-Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
-As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not straightforward.
-In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
-
-% Commentaire figure instant_256_20
+
 \begin{figure}
@@
 	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-  \caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag along the resampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
-For these particular dynamics, the \ac{tams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation in the initial ensemble.}
-\end{figure}
-We apply the \ac{tams} with $N=256$ initial trajectories, with a duration $T_a = 20 \tau_c$.
-Similarly to figure~\ref{fig:comparaison_temps_de_retour}, figure~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by resampled trajectories.
-In addition, figure~\ref{fig:AMS_drag_resampling} shows the distribution of the maximum achieved drag for the $N$ \textit{initial} trajectories.
-Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
-These maxima are depicted by the stars in figure~\ref{fig:AMS_drag_resampling}.
-However, one can see that resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
-
-% Commentaire de la figure iter_181
+	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
+The \ac{tams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
+\end{figure}
+
+% Application to instantaneous drag
+%
+% Can the \ac{tams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
+% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
+%
+Our aim is now to sample flow evolution that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
+In contrast with the \acl{ou} process, the phase space is here higlhy-dimensional with intrinsically chaotic dynamics.
+In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
+However, a natural simple choice for which we will opt in this study is certainly $\xi (\mathbf{x}(t),t) = f_d(t)$. 
+\EL{say something about the perturbation applied at branching?}
+
+% application
+%
+The \ac{tams} is applied with $N=256$ trajectories of duration $T_a = 20\tau_c$.
+%
+Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost, similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
+%
+%
+The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
+After a few iterations, the trajectories with the lowest maxima of the score function are discarded, and new trajectories with higher maxima are re-sampled.
+These maxima are depicted by the stars in Fig~\ref{fig:AMS_drag_resampling}.
+Importantly, one can see that re-sampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories. This in contrast with the case of the \acl{ou} process.
+
 \begin{figure}
@@
   \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-  \caption{\label{fig:AMS_drag_trajectories} Ensemble of trajectories after $181$ iterations of the selection-mutation procedure.In this experiment, the \ac{tams} is used with the instantaneous drag $f_d$ as a score function.
-Trajectories have a duration $T_a = 5\tau_c$ and their number is $N_c = 32$.
-    The \ac{tams} fails to efficiently sample rare trajectories, because all trajectories are ultimately resampled from the trajectory displaying the highest maximum in the initial ensemble.}
-\end{figure}
-Figure~\ref{fig:AMS_drag_trajectories} can be better understood by visualising the drag signals corresponding to the last resampled trajectories.
-Figure~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after $181$ iterations of the \ac{tams} resampling procedure (see figure~\ref{fig:illustr_AMS}).
-One can see that, utimately, trajectories are resampled from one unique initial trajectory.
-This phenomenon can be explained as follows.
-Because the flow dynamics have a memory, it takes a certain time before a resampled trajectory separate from its parent.
-In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting points will overlap over a timescale called
-the \textit{Lyapunov timescale}~\cite{lyapunovRef}. \textcolor{red}{@todo@Freddy: Provide citation for Lyapunov timescale}
-However, we observe that, in our particular case, the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag process.
-This means that the resampling of a trajectory based on the value of the drag at $t^{\star}$ cannot lead to larger drag values for $t^{\star} \leq t \leq \tau_L$.
-For $t > \tau_L$, the drag process has no memory of the drag fluctuations on which the resampling was based.
-As a consequence, the probability of observing an extreme fluctuation over $\tau_l \leq t \leq T_a$ is very low.
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_0$ is linked to the convective nature of the flow.
-As described in section~\ref{sec:instantaneous_drag}, extreme fluctuations of the instantaneous drag $f_d$ have a lifetime of roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for
-trajectories to separate, \textit{i.e} the Lyapunov timescale.
+  \caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
+In this experiment, the \ac{tams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
+Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
+  The \ac{tams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the trajectory displaying the highest maximum in the initial ensemble.}
+\end{figure}
+
+%pitfall
+%
+Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{tams} re-sampling procedure. 
+%
+Let us mention that for the sake of simplicity, the trajectory of the system is  here abusively associated with the time evolution of the observed drag.  \EL{useful to mention?}
+%
+Interestingly, it is obtained that all trajectories ultimately originate from one unique initial trajectory. 
+%
+This feature may be explained as follows.
+Because the dynamics has a memory, it takes a certain time before a re-sampled trajectory separates from its parent.
+In our situation, this memory originates from the deterministic nature of fluid dynamics.
+Two trajectories (of the drag signal) starting from infinitesimally close points typically overlap over a timescale $\tau_L$ called the \textit{Lyapunov timescale}~\cite{lyapunovRef}.
+\EL{Freddy: Provide citation for Lyapunov timescale}
+%
+In our case, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale of the drag process.
+Therefore, the re-sampling of a trajectory branched at $t=t^{\star}$ corresponding to the maximal drag amplitude cannot lead to larger values for $t^{\star} \leq t \leq t^{\star}+\tau_L$.
+For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based.
+Therefore, the probability of observing a new extreme fluctuation is also very low.
+%
+
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the ``convective nature'' of the flow.
+%
+As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and therefore to the mean velocity $U$.
+On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
+The ratio $u_{rms}/U < 1$ in our case of grid-generated turbulence implies that $\tau_L > \tau_c$.
+\EL{does it make sense to you?}
+
+%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
+Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
 
 % perspectives
-A direct application of the \ac{tams}, choosing the drag itself as the score function,  cannot be expected to efficiently sample rare trajectories for flows with such a phenomenology.
-Indeed, there is \textit{a priori} no reason for the drag itself to be a good measure of how likely future drag fluctuations are.
-Finding a better score function requires an intricate understanding of the dynamics leading to extreme drag fluctuations, beyond the phenomenological description of section~\ref{sec:dynamical_aspects}.
-Furthermore, such fine-tuning is problem dependent, and the study should be done each time the algorithm is applied to a similar, however different, system.
-A current line of research is to adapt the algorithm itself to the phenomenology.
-As an example, the resampling procedure could be modified so that new trajectories are resampled roughly a Lyapunov time \textit{before} the maximum is attained.
-In this way new events, associated to higher fluctuations, have the opportunity to occur.
-However, such modification(s) of the resampling procedure must be implemented with care, so as to preserve the mathematical properties of the \ac{ams} and \ac{tams}.
+%
+In summary, it appears that a straightforward application of the \ac{tams} by taking the drag as score function does not allow us to efficiently sample extreme fluctuations.
+%
+Indeed, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
+On the other hand, finding a better score function would require an intricate understanding of the flow scenario leading to extreme events, certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
+%
+%Furthermore, such ``tuning'' is problem dependent.
+%
+A line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time \textit{before} the maximum is reached.
+That way, new events associated to higher fluctuations may have the opportunity to develop.
+However, such modification of the procedure must be implemented with care, so as to preserve the mathematical properties (evaluation of the statistical bias) of the algorithm.
 %
@@
 
-We now turn to the sampling of extreme fluctuations of the time-averaged drag $F_T$.
-The \ac{tams} algorithm could be used in the same way as described in section~\ref{sec:ams} for the instantaneous drag, choosing $F_T$ itself as a score function.
-However, this leads to the same problem: the resampling fails to efficiently generate new, higher fluctuations of $F_T$.
-
-For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
-In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
-The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
-The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous1994go,grassberger2002go}.
-Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
-The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{giardina_direct_2006}.
-Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems~\cite{giardina_direct_2006,tailleur_probing_2007}.
+The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
+The \ac{tams} algorithm could be used in the same way as before by choosing $F_T$ as score function. However, this would lead to the same pitfall.
+% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
+%
+For a time-averaged observable,  an alternative approach is provided by the \acf{gktl} algorithm \cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
+%
+Similarly to the \ac{tams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
+In this case, however, the trajectories interact dynamically. 
+%
+At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the replica.
+%
+The weights are chosen so that, after several iterations of the algorithm, the trajectories generated are distributed according to a biased probability distribution in favor of trajectories with large values of the (time-averaged) observable of interest.
+%
+The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
+Similar ideas have been applied in a wide range of fields over the past 50 years, under different names \EL{donner des exemples}, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
+The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
+Over the past ten years, it has successfully been applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
 % It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
@@
 \label{sec:gktl_description}
-The \ac{gktl} algorithm consists in simulation an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{1 \leq n \leq N}$ starting from independant random initial conditions.
-Similarly to section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
-We consider an observable of interest $A(\mathbf{x}(t))$ and a cloning period $\tau$.
-At times $t_{i}=i\tau$ (with $i=1,2,...,T_{a}/\tau$) we assign to each trajectory $n$ a weight $W_{n}^{i}$ defined as
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}.
-\end{equation}
-For each trajectory $\{\mathbf{x}_{n}\}_{0 \leq t_i}$, a random number of copies of the trajectory are generated.
+As with the \ac{tams} algorithm, the \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%
+Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$ and introduce a cloning period $\tau$.
+%
+At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
+%
+\begin{equation}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalization factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+\label{eq:Weight}
+\end{equation}
+so that $\sum_{n=1}^N W_n^i = N$.
+%
+%
+Therefore, the weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine (on average \EL{vague?}) how many copies of each trajectory are produced at time $t=t_i$. 
+% For each t $\{\mathbf{x}_{n}\}_{0 \leq t \leq  t_i}$, a random number of copies of the trajectory are generated.
 This number is on average proportional to the weight $W_{n}^{i}$, such that the total number of trajectories produced at each event is equal to $N$.
-For deterministic systems, a small perturbation is introduced for clones at times $t_i$, so that clones separate from their parent. 
-The parameter  is chosen by the user in order to control the strength of the selection.
-The free parameter $k$ sets the amplitude of the typical fluctuations within the biased ensemble of trajectory sampled by the algorithm.
-A higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.
-
-Let us denote formally $\mathbb{P}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right)$ the probability to observe a trajectory $\left\{ x(t)\right\} _{0\leq t\leq T_{a}}$ in the model, and $\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}} = \left\{ x(t)\right\} _{0\leq t\leq T_{a}} \right)$ the probability to observe the same trajectory with the algorithm.
-By construction of the algorithm through the weights~\eqref{eq:Weight}, we have
+%
+For deterministic systems, a small perturbation is applied to the clones so that they can rapidly separate from their parents. 
+%
+% The parameter  is chosen by the user in order to control the strength of the selection.
+%
+The free parameter $k$ drives the amplitude of fluctuations targeted by the biased ensemble of trajectories.
+A higher value of $k$ gives a higher weight to higher fluctuations and thus drives the sampling to more extreme events.
+
+Let 
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
+refer formally to the probability of the trajectory 
+$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ of our dynamical system.
+%
+% $\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}} \right)$ 
+%
+By construction, the probability to obtain the same trajectory by resorting to the previous cloning procedure is given by
+%
 \begin{align}
-\mathbb{P}_{k}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(x(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}=\left\{ x(t)\right\} _{0\leq t\leq T_{a}}\right).\label{eq:Biased_Path_Approximation}
+\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)
+\label{eq:Biased_Path_Approximation}
 \end{align}
-where the normalisation factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(X(t))dt}\right]$, denoting by $\mathbb{E}_{0}$ the expectation value with respect to $\mathbb{P}_{0}$, and $\underset{N\rightarrow\infty}{\sim}$ means that this is true only asymptotically for large $N$.
-The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
-Equation~\eqref{eq:Biased_Path_Approximation} is obtained by assuming the mean field approximation
-\begin{equation}
-R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{0}^{t_{_{1}}}A(X_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{0}^{t_{1}}A(X(t))dt}\right],\label{eq:Mean_Field_Approximation}
-\end{equation}
-which, by induction, and using a formula similar to~\eqref{eq:Mean_Field_Approximation} at each step of the induction, leads to~\cite{giardina_direct_2006,giardina_simulating_2011}:
-\begin{equation}
-\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(X(t))dt}\right].\label{eq:Estimate_Lambda}
-\end{equation}
-The validity of the mean field approximation and the fact that the typical relative error due to this approximation is of order $1/\sqrt{N}$ has been proven~\cite{DelMoralBook,DelMoral2013} to be true for a family of rare event algorithms including the one adopted in this paper.
-
-Formula~\eqref{eq:Biased_Path_Approximation} is valid only for times $T_{a}$ that are integer multiples of the resampling time $\tau$.
-The killed trajectories have to be discarded from the statistics.
-Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
-In this way one obtains an effective ensemble of $N$ trajectories from time 0 to time $T_{a}$, distributed according to $\mathbb{P}_{k}$.
-All trajectories reconstructed in this way are real solutions of the model: we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
-
-The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$, from the distribution $\mathbb{P}_{k}$.
-This is done using the backward reconstructed trajectories and inverting formula~\eqref{eq:Biased_Path_Approximation}.
-For more details concerning the implementation of the algorithm and the corresponding data analysis, see~\cite{lestang:tel-01974316,brewer2018efficient}.
-As an example, say one wants to estimate the expectation value of an observable $O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)$.
-An estimator is then given by
-\begin{equation}
-\mathbb{E}_{0}\left[O\left(\left\{ X(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ X_{n}(t)\right\} _{0\leq t\leq T_{a}}\right)\mbox{e}^{-k\int_{0}^{T_{a}}A(X_{n}(t))dt}\mbox{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
-\end{equation}
-where the $X_{n}$ are the $N$ backward reconstructed trajectories.
-Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
+where the normalization factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]$ with $\mathbb{E}_{0}$ being the expectation value with respect to $\mathbb{P}_{0}$.
+%
+% obvious? 
+%and $\underset{N\rightarrow\infty}{\sim}$ meaning that this is true only asymptotically for large $N$.
+%
+%
+%The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
+%
+Eq.~\eqref{eq:Biased_Path_Approximation} relies on the mean-field approximation
+\begin{equation}
+R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}(t))dt}\right]
+\label{eq:Mean_Field_Approximation}
+\end{equation}
+which by induction (and assuming similar relations for the successive $R_i$'s) eventually leads  to~\cite{giardina_direct_2006,giardina_simulating_2011}
+\begin{equation}
+\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(\mathbf{X}(t))dt}\right].\label{eq:Estimate_Lambda}
+\end{equation}
+\EL{IS IT TRUE?}
+The validity of this mean-field approximation and the typical relative error related to this approximation being of order $1/\sqrt{N}$ have been proven rigorously \cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
+%
+%
+The killed trajectories, \emph{i.e.} with a weight smaller than $1$ \EL{true?}, are discarded from the statistics.
+%
+% no because the trajectories are cloned
+%
+% Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
+%
+Eventually, an effective ensemble of $N$ trajectories for $0 \leq t \leq T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+All trajectories are solutions of the dynamical system since we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
+\EL{quid the small perturbation to separated the clones? comment required}
+%
+The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+%
+More details concerning the implementation of the algorithm and the corresponding data analysis are provided in \cite{lestang:tel-01974316,brewer2018efficient}.
+
+%
+%\EL{next paragraph needs rephrasing or suppress}
+%
+%For instance, the expectation value of the observable $O\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}\right)$ is given by
+%\begin{equation}
+%\mathbb{E}_{0}\left[O\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ \mathbf{X}_{n}(t)\right\} _{0\leq t\leq T_{a}}\right){e}^{-k\int_{0}^{T_{a}}A(\mathbf{X}_{n}(t))dt}{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
+%\end{equation}
+%where the $X_{n}$ are the $N$ backward reconstructed trajectories.
+%Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
 As a result, the effective ensemble of trajectories generated by the \ac{gktl} algorithm contains a larger amount of such events, and empirical estimators such as~\eqref{eq:GK_O_estimator} have a dramatically lower statistical error.
@@
 
-We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
-The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag} acting on the square.
-Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
+	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10~\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
+The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
+	The dashed line denotes the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
+\EL{$F_T-\overline{F_T}$ in the figure!}}
+\end{figure}
+
+The application of the \ac{gktl} algorithm is now studied for the flow dynamics introduced in section~\ref{sec:test_flow} with the aim of sampling trajectories with extreme fluctuations of the \textit{time-averaged} drag.
 
 % Choice of parameters and perturbation
-The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
-In the following we set $T_a = T = 10\tau_c$.
-We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-Lastly, one must set the cloning period $\tau$.
-On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
-On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
-As a consequence, a rule of thumb is to set $\tau \approx \tau_c$.
-In the following, $\tau = \tau_c /2$.
-
-We performed three experiments, corresponding to three different values of the parameter $k$.
-In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{touchette_large_2009}.
-In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{touchette_large_2009}.
-
-% Commentaire de la figure IS_GKTL
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-  \caption{\label{fig:IS_GKTL} Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$.
-The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
-The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
-\end{figure}
-Figure~\ref{fig:IS_GKTL} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
-For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
+The computational cost $C_{gktlS}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+In practice, $N=16384$ and $T_a = T = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+A cloning period $\tau$ that is too short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
+As a result, a rule of thumb is to set $\tau \approx \tau_c$. We took $\tau = \tau_c /2$.
+%
+Three numerical experiments corresponding to three different values of the parameter $k$ have been performed.
+%
+%
+%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
+%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
+%
+%
+As a proof of realizability, Fig.~\ref{fig:IS_GKTL} displays the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag, $\tilde F_T \equiv F_T - \overline{F_T}$:
 \begin{equation}
   \label{eq:estimate_biased_measure}
-  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
-  \approx \frac{1}{N}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
-\end{equation}
-computed over the set of $N_c=16384$ trajectories sampled by the algorithm.
-Moreover, figure~\ref{fig:IS_GKTL} displays an estimate of the unbiased \ac{pdf} describing the time-averaged drag process.
-This estimate was computed on the basis the control run, see section~\ref{sec:pdfs}.
-Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-
-% Limitations
-
-For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
-For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
-This effect is visible in figure~\ref{fig:IS_GKTL}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
-Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
-An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
-See~\cite{nemoto2017finite,Guevara_Hidalgo_2018} for further details.
+  \rho_k(F) 
+  %= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
+  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
+\end{equation}
+in addition to the unbiased \ac{pdf} based on the control run.
+%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
+
+% limitations
+%
+For a fixed number of trajectories $N$, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
+%
+indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is over-peaked.
+%
+Such finite-size effect is expected to gradually increase as the bias amplitude  $k$ is increased.
+%
+An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
+This point has been very recently examined in \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
 % how to assess the value of k_max ?
-For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
-This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.
-
-% Malgré les limitations, c'est toujours mieux
-% qu'un échantillonnage direct
-We now assess the efficiency of the \ac{gktl} algorithm, with respect to a direct sampling approach. 
-For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T} \geq a$.
-Note that we count as one trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.}.
-As a comparison, table~\ref{tbl:nbTraj_tbl} indicates the average number of fluctuations one can expect from simulating $N_c$ independant trajectories of duration $T_a$ $\{f_d(t)\}_{0\leq t \leq T_a}$, this with the same computational cost $C_{gktl}$.
-This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
-This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
+In this study, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
+This diversity may be monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
 
@@
 	\centering
-  \begin{tabular}{l|ccccc}
+	\begin{tabular}{|l|c|c|c|c|c|}
+		\hline
 		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
 		\hline
-    $k=0.02$ & 1594 & 799 & 155 & 22 & 0  \\ 
-    $k=0.025$ & 1019 & 834 & 521 & 198 & 27  \\ 
+		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
+		\hline
+		$~k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
+		\hline
 		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-    $\mathcal{N}_{direct}$ & 2599.408 & 372.738 & 22.117 & 0.519 & 0.005 
+		\hline
+		${N}_{direct}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
+		\hline
 	\end{tabular}
-  \caption{Number of fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$.
-    Within the biased ensemble, trajectories are counted as one if they overlap over more than $50\%$ of their duration.
-    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.}
+	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,~...$ captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
+		Note that trajectories that overlap for more than half of their duration are counted as one.
+		${N}_{direct}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$ of the same duration.}
 	\label{tbl:nbTraj_tbl}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-  \caption{PDF of average drag over $10\tau_c$}
-  \label{fig:PDF_AVG}
-\end{figure}
-
-% Conclusion de cette partie
+% Comparisons with direct sampling
+%
+The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
+Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged $F_{T} \geq a$ for the different values of the bias parameter $k$.
+Note that trajectories that overlap for more than half of their duration are counted as one.
+%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
+Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations one can expect from simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$, \emph{i.e.} with the same computational cost $C_{gktl}$.
+%
+%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+%  \caption{PDF of average drag over $10\tau_c$}
+%  \label{fig:PDF_AVG}
+%\end{figure}
+
+% Conclusion
+%
 Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a direct sampling approach of similar computational cost.
@@
 
+\EL{rajouter le calcul des temps de retour par gktl pour etoffer un peu la partie du papier sur gktl -- il faudrait montrer un resultat qui montre que l'on obtient la bonne statistique}
+
 \section{Conclusion}
 \label{conlusion}
+
+\EL{CONCLUSION TO BE REVIEWED}
+
 We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
@@
 \end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \todo{REF ADR}.
+Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \cite{De_Rosis_2016}.
 

### 2019-05-15 draft_p2.tex
Add figure illustrate_return_time and add it to main text

@@
 How rare is a fluctuation $f_d \geq a$ can be quantified by its so-called \textit{return time} $r(a)$.
-This latter is defined as the average time between two consecutive occurrences of events of amplitudes $f_d \geq a$.
-\EL{suffisamment clair?
-un schema pour fixer les choses?}
+This latter is defined as the average time between two consecutive occurrences of events of amplitudes $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
+  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurence of the event: $r(a) = \mathbb{E}[\tau]$.
+Note that we assume $r(a) >> \tau_0$, and therefore consider that such events are points in time.}}
+\end{figure}
 Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time: $r(a) \gg \tau_c$.

### 2019-05-15 figures/illustrate_return_time/illustrate_return_time.py

@@ -3,6 +3,19 @@ import matplotlib.style
 import matplotlib.pyplot as plt
 
 def bundle_nearest_points(I, dist):
+    """
+       Given a list of number I (int or floats), return a list of lists, each sublist containing
+       subsequent elements distant from less than dist of the next one.
+
+       Parameters
+       -----------
+       I: list
+       dist: int or float
+
+       Returns
+       --------
+       list of lists
+    """
     bundle = []
     list_of_bundles = []
     bundle.append(I[0])
@@ -29,10 +42,11 @@ T = 1000
 f = np.fromfile(path_to_file, float, T*tau_0, "");
 tspan = np.linspace(0,T,len(f))
 
-a = 0.15
-I = np.argwhere(f>a)
+a = 0.17 # threshold
+I = np.argwhere(f>a) # find all indices where f>a
 
-list_of_bundles = bundle_nearest_points(I, 2000)
+tau_c = 4*tau_0
+list_of_bundles = bundle_nearest_points(I, tau_c)
 
 point_array = np.zeros(len(list_of_bundles))
 value_array = np.zeros(len(list_of_bundles))                    
@@ -50,15 +64,22 @@ line3, = ax.plot([tspan[0], tspan[-1]], [a, a], linestyle='--', linewidth=.7)
 line3.set_color(line2.get_color())
 line.set_linewidth(.8)
 
-y0 = 0.20
-eps = 10
-props=dict(arrowstyle="->", color=line2.get_color(),
-                length_includes_head=True, width=0.002,
-                head_width=0.01, head_length=5)
+y0 = 0.22 # height of the arrows
+dy0 = 0.017 # How higher is the tau letters wrt arrows
+dx0 = 10 # Shift tau letters to the left
+eps = 1 # Shorten the arrows on both sides
+props=dict(arrowstyle="<->", color=line2.get_color(), linewidth=1.5)
 
-for x_point, x_point_fwd in zip(point_array, point_array[1:]):
+# Plot the arrows and text
+for i, tup in enumerate(zip(point_array, point_array[1:])):
+    x_point, x_point_fwd = tup
     plt.annotate('', xy=(x_point+eps, y0), xytext=(x_point_fwd-eps, y0),
                  arrowprops=props)
[…]

### 2019-05-15 figures/autocorr_drag/plot_correlation.py
Update x label of autocorrelation function

@@ -6,7 +6,7 @@ plt.style.use('ggplot')
 
 
 buf = np.fromfile("correlationData.dat", float, -1, "");
-n=len(buf)/2
+n=int(len(buf)/2)
 tDomain = buf[0:n]
 c = buf[n:]
 
@@ -17,7 +17,7 @@ lineC, = ax.plot(tDomain/2000,c)
 line0.set_linestyle('--')
 line0.set_linewidth(0.7)
 
-ax.set_xlabel(r'$\tau$',fontsize=22)
+ax.set_xlabel(r'$\tau / \tau_c$',fontsize=22)
 plt.ylabel(r'$C(\tau)$',fontsize=22)
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)

### 2019-05-15 draft_p2.tex
Update x label of autocorrelation function
excusrion --> fluctuation in Extracting extreme drag fluct. from timeseries (p10)

@@
 The correlation time $\tau_c$ is given by $C(\tau_c)=0$.
-\EL{mettre $\tau/\tau_c$ en abscisse de l'autocorrelation (b)}
   %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
@@
 According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
-Precisely, 104 independent excursions \EL{excursion, terminologie correcte?} with $f_d(t) \geq 7.6~\sigma$ have been identified.
-For convenience, each excursion will be characterized in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
+Precisely, 104 independent fluctuations \EL{excursion, terminologie correcte?} \TL{peut etre, mais fluctuation est correct et plus precis?.}with $f_d(t) \geq 7.6~\sigma$ have been identified.
+For convenience, each fluctuation will be characterized in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
 %

### 2019-05-15 figures/pressure_ratio/plot_pressure_ratio.py
Fix signs for figure pressure_ratio

@@ -13,7 +13,7 @@ meanDrag = 0.0252;
 M = (meanPf + meanPr)/2;
 sig = 0.0412;
 mu = 0.0252;
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/';
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/';
 maxDrag = np.zeros(NbEvents)
 frontContrib = np.zeros(NbEvents)
 rearContrib = np.zeros(NbEvents)
@@ -38,10 +38,10 @@ maxDrag = (maxDrag-mu) / sig
 fig,ax = plt.subplots(figsize=(9,5), constrained_layout=True)
 
 
-lineRear, =ax.plot(maxDrag, rearContrib, marker='o', label=r'$\tilde{p}^{\star}_{base}$')
+lineRear, =ax.plot(maxDrag, rearContrib, marker='o', label=r'$-\tilde{p}^{\star}_{base}/\tilde{f}^{\star}_d$')
 lineRear.set_linestyle('None')
 
-lineFront, =ax.plot(maxDrag, frontContrib, marker='^', label=r'$\tilde{p}^{\star}_{fb}$')
+lineFront, =ax.plot(maxDrag, frontContrib, marker='^', label=r'$\tilde{p}^{\star}_{fb}/\tilde{f}^{\star}_d$')
 lineFront.set_linestyle('None')
 
 coefFront = np.polyfit(maxDrag, frontContrib, 1);

### 2019-05-15 figures/timeseries_extremes/plot_force.py
\tilde{f}_d instead of f_d for figure 8 (timeseries_extremes)

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
 
 sig=0.0412
 m=0.0252
@@ -32,7 +32,7 @@ ax.set_xlim((-2,2))
 plt.annotate('', xy=(-0.5,-0.35), xytext=(0.5,-0.35), arrowprops=dict(arrowstyle='<->',lw=2))
 plt.text(-0.1, -0.05, r'$\tau_c$', fontsize=18)
 ax.legend(loc='best',fontsize=16)
-ax.set_ylabel('$f_d(t)$',fontsize=22)
+ax.set_ylabel('$\tilde{f}_d(t)$',fontsize=22)
 ax.set_xlabel('$t$',fontsize=22)
 plt.xticks(fontsize=18)
 plt.yticks(fontsize=18)

### 2019-05-15 draft_p2.tex
\tilde{f}_d instead of f_d for figure 8 (timeseries_extremes)
Fix f tilde in PDF and corresponding text

@@
 The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(\tilde f_d) \propto e^{-\lambda \tilde f_d}$.
+It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
 %
@@
 	As the amplitude of the drag increases, the relative variation of the base pressure also increases.
-\EL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
+\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
 \end{figure}
@@
 $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-\EL{please confirm signs! I have added a minus sign: $-\tilde{p}_{base}^{\star}$}  
 %
@@
 The profile is slightly skewed indicating that the step up is slower than the return to typical values.
-\EL{ $\tilde f_d$ rather than $f_d$ in figure?}}
+\TL{ $\tilde f_d$ rather than $f_d$ in figure.}}
 \end{figure}

### 2019-05-15 figures/PDF_drag/plot_pdf.py

@@ -6,7 +6,7 @@ plt.style.use('ggplot')
 
 
 buf = np.fromfile("pdfData_SQ_centered.dat", float, -1, "");
-n=len(buf)/2
+n=int(len(buf)/2)
 f = buf[0:n]
 xi = buf[n:]
 
@@ -15,13 +15,13 @@ lineSQ, = ax.plot(xi,f)
 lineSQ.set_label('With obstacle')
 
 buf = np.fromfile("pdfData_NOSQ.dat", float, -1, "");
-n=len(buf)/2
+n=int(len(buf)/2)
 f = buf[0:n]
 xi = buf[n:]
 lineNOSQ, = ax.plot(xi,f, label='No obstacle')
 
 ax.set_yscale('log')
-ax.set_xlabel('$f_d - \\bar{f}_d$',fontsize=22)
+ax.set_xlabel('$\\tilde{f}_d$',fontsize=22)
 plt.ylabel('$\log(PDF)$',fontsize=22)
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)

### 2019-05-15 figures/return_time/plot_return_time.py
Set \tilde{f}_d instead of f_d in figures return_tim and fix it in timeseries_extemes

@@ -4,11 +4,11 @@ import matplotlib.pyplot as plt
 
 plt.style.use('ggplot')
 
-path_to_file = "/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/scripts/" \
+path_to_file = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/scripts/" \
                + "courbe_temps_de_retour_drag_instant/return_times_drag_instant.dat"
 
 buf = np.fromfile(path_to_file, float, -1, "");
-n=len(buf)/2
+n=int(len(buf)/2)
 r = buf[0:n]
 avec = buf[n:]
 
@@ -24,7 +24,7 @@ line, = ax.plot(r, avec)
 ax.set_xscale('log')
 #line.set_linewidth(.7)
 
-ax.set_xlabel(r'$r(f_d \geq a) / \tau_c$',fontsize=22)
+ax.set_xlabel(r'$r(\tilde{f}_d \geq a) / \tau_c$',fontsize=22)
 ax.set_ylabel('$a$',fontsize=22)
 ax.set_yticks([3,5,7,9])
 ax.set_yticklabels((r'$3\sigma$',r'$5\sigma$',r'$7\sigma$',r'$9\sigma$'))

### 2019-05-15 figures/timeseries_extremes/plot_force.py

@@ -32,7 +32,7 @@ ax.set_xlim((-2,2))
 plt.annotate('', xy=(-0.5,-0.35), xytext=(0.5,-0.35), arrowprops=dict(arrowstyle='<->',lw=2))
 plt.text(-0.1, -0.05, r'$\tau_c$', fontsize=18)
 ax.legend(loc='best',fontsize=16)
-ax.set_ylabel('$\tilde{f}_d(t)$',fontsize=22)
+ax.set_ylabel('$\\tilde{f}_d(t)$',fontsize=22)
 ax.set_xlabel('$t$',fontsize=22)
 plt.xticks(fontsize=18)
 plt.yticks(fontsize=18)

### 2019-05-15 figures/illustr_density_streamlines/make_plot_density_2.py
Change title of figure density_streamlines

@@ -13,7 +13,7 @@ y = np.asarray(range(0,64))
 N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
-prefix = '/home/thibault/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is lowest fluctuation
@@ -67,7 +67,7 @@ rhoCropped = rho[ymin:ymax,xmin:xmax]
 im = ax_list.imshow(rhoCropped, interpolation='spline36', cmap='RdBu', vmin=0.85, vmax=1.15)
     
 # Print fluctuation amplitude as title
-tit='$f_d = {:2.1f}\\sigma$'.format(F[0])
+tit='$\\tilde{f}'+'_d = {:2.1f}\\sigma$'.format(F[0])
 ax_list.set_title(tit,fontsize=22)
 
 # Adds a filled Rectangle to mark the square obstacle

### 2019-05-15 draft_p2.tex
Change title of figure density_streamlines
Minor modies following E's remarks

@@
 In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.
-\EL{$f_d$ or $\tilde f_d$?}}
+\TL{changed to $\tilde f_d$.}}
 \end{figure}
@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} \EL{ou TAMS?} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
+  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
 These trajectories are independent and can be simulated in parallel.
@@
 %
-We consider here a variant \EL{generalization?} of \ac{ams} called \ac{tams}, in which the branching procedure relies on a time-dependent \textit{score function} $\xi (\mathbf{x}(t),t)$.
-Each iteration therefore consists in discarding the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$.
+\TL{The selection of trajectories rely on a \textit{score function} $\xi (\mathbf{x}(t),t)$.}
+Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$.
 Discarded trajectories are re-sampled according to the remaining trajectories.
-% As an illustration, a particular iteration of the \ac{tams} algorithm is sketched in figure~\ref{fig:illustr_AMS} .
-Further details about the algorithm and corresponding mathematical results are provided in \cite{lestang_computing_2018} and references therein.
-\EL{What is the difference between AMS and TAMS? why do we need TAMS rather than AMS, This is not mentioned in text.}
-
-\subsubsection{Application of the \ac{ams} \EL{OR TAMS?} to a simple case: the \acl{ou} process}
+\TL{For a recent review of the \ac{ams}, its history and applications, see~\cite{cerou_2019}.}
+\TL{Note that we use a specific form of the the \ac{ams} in which trajectories have a fixed duration $T_a$.
+It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing} and the following results are not limited to this particular variant.}
+
+\subsubsection{Application of the \ac{ams} to a simple case: the \acl{ou} process}
 \label{sec:ams_ou}
@@
 The solid blue line is the analytical solution for the return time of amplitude $a$.
-\EL{how many trajectories in the ensemble?
-REPLACE $f_d$ with $x$ in figure} }
+\TL{replaced $f_d$ with $x$ in figure} }
   \label{fig:comparaison_temps_de_retour}
@@
 
-The \ac{tams} is applied  to a set of \EL{$N$} trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with the objective to sample fluctuations of $x\geq a$ with $a$ being very large compared to typical values of $x$.
+\TL{The \ac{tams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5$.Note that, for the process as define in~\eqref{eq:ou}, the correlation time is $\tau_c = 1$.}
+Our objective to sample fluctuations of $x\geq a$ with $a$ being very large compared to typical values of $x$.
 The score function is here $x(t)$ and only one single trajectory is re-sampled at each iteration.
@@
 However, a natural simple choice for which we will opt in this study is certainly $\xi (\mathbf{x}(t),t) = f_d(t)$. 
-\EL{say something about the perturbation applied at branching?}
+\TL{Let us stress here that the flow dynamics are deterministic.
+As a result, resampled trajectories do not differ from the parent trajectory their are resampled
+from.
+A workaround is to introduce a slight perturbation in the state of the flow at the branching point (see figure~\ref{fig:illustr_AMS}).
+Because of the flow chaoticity, resampled trajectories will separate from their parent over the \textit{Lyapunov timescale} $\tau_L$.}\EL{Freddy: provide reference for Lyapunov timescale.}
 
@@
 %
-Let us mention that for the sake of simplicity, the trajectory of the system is  here abusively associated with the time evolution of the observed drag.  \EL{useful to mention?}
+Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  \EL{useful to mention?}\TL{Yes, good point!}
 %
@@
 This feature may be explained as follows.
-Because the dynamics has a memory, it takes a certain time before a re-sampled trajectory separates from its parent.
+Because the dynamics has a memory, it takes a certain time $\tau_L$ before a re-sampled trajectory separates from its parent.
 In our situation, this memory originates from the deterministic nature of fluid dynamics.
-Two trajectories (of the drag signal) starting from infinitesimally close points typically overlap over a timescale $\tau_L$ called the \textit{Lyapunov timescale}~\cite{lyapunovRef}.
-\EL{Freddy: Provide citation for Lyapunov timescale}
-%
-In our case, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale of the drag process.
+%
+\TL{In our case, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale of the drag process.}
 Therefore, the re-sampling of a trajectory branched at $t=t^{\star}$ corresponding to the maximal drag amplitude cannot lead to larger values for $t^{\star} \leq t \leq t^{\star}+\tau_L$.
@@
 %
-Therefore, the weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine (on average \EL{vague?}) how many copies of each trajectory are produced at time $t=t_i$. 
+\TL{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are produced at time $t=t_i$.
+See~\cite{brewer_efficient_2018, nemoto_population-dynamics_2016} for information on the practical implementation of the algorithm.}
 % For each t $\{\mathbf{x}_{n}\}_{0 \leq t \leq  t_i}$, a random number of copies of the trajectory are generated.

### 2019-05-22 draft_p2.tex
Reprise partielle de la presentation de GKTL

@@
 \end{equation}
+The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
 which by induction (and assuming similar relations for the successive $R_i$'s) eventually leads  to~\cite{giardina_direct_2006,giardina_simulating_2011}
@@
 \EL{IS IT TRUE?}
-The validity of this mean-field approximation and the typical relative error related to this approximation being of order $1/\sqrt{N}$ have been proven rigorously \cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
+The validity of this mean-field approximation and the 
 %

### 2019-05-22 draft_p2.tex
Add references to issues to Emmanuel's comments

@@
 \begin{abstract}
-	\EL{ ABSTRACT TO BE REVIEWED}
+	\EL{ ABSTRACT TO BE REVIEWED \#8}
 This study evaluates the relevance of  rare-event sampling to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
@@
 An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \EL{is it TRUE?}
+Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \EL{is it TRUE? \#11}
 % 
@@
 	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
-	The vorticity is given in lattice units. \EL{THE COLORMAP IS REVERSED IN THE WHOLE PAPER: BLUE SHOULD BE NEGATIVE AND RED POSITIVE}}
+	The vorticity is given in lattice units.}
 	\label{fig:typical_vorticity}
@@
 $\sigma_T$ denotes the standard deviation of $F_T$.
-\EL{FIGURE IS MISSING}}
+\EL{FIGURE IS MISSING} \TL{Addressed in issue \#9}}
 	\label{fig:extreme_avg}
@@
   \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
+  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?
+\#1} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
 These trajectories are independent and can be simulated in parallel.
@@
 A workaround is to introduce a slight perturbation in the state of the flow at the branching point (see figure~\ref{fig:illustr_AMS}).
-Because of the flow chaoticity, resampled trajectories will separate from their parent over the \textit{Lyapunov timescale} $\tau_L$.}\EL{Freddy: provide reference for Lyapunov timescale.}
+Because of the flow chaoticity, resampled trajectories will separate from their parent over the \textit{Lyapunov timescale} $\tau_L$.}\EL{Freddy: provide reference for Lyapunov timescale. \#12}
 
@@
 The ratio $u_{rms}/U < 1$ in our case of grid-generated turbulence implies that $\tau_L > \tau_c$.
-\EL{does it make sense to you?}
+\EL{does it make sense to you?} \TL{A priori oui.
+Mais j'aimerai prendre 10-15min pour bien y réféchir \#2}
 
@@
 The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-Similar ideas have been applied in a wide range of fields over the past 50 years, under different names \EL{donner des exemples}, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
+Similar ideas have been applied in a wide range of fields over the past 50 years, under different names \EL{donner des exemples \#3}, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
 The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
@@
 \end{equation}
-\EL{IS IT TRUE?}
+\EL{IS IT TRUE?} \TL{Oui. Il faudrait un peu reécrire ce paragraphe pour que ca soit clair, voir issue \#5}
 The validity of this mean-field approximation and the 
@@
 All trajectories are solutions of the dynamical system since we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
-\EL{quid the small perturbation to separated the clones? comment required}
+\EL{quid the small perturbation to separated the clones? comment required \#6}
 %
@@
 	The dashed line denotes the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-\EL{$F_T-\overline{F_T}$ in the figure!}}
+\EL{$F_T-\overline{F_T}$ in the figure! \#4}}
 \end{figure}
@@
 
-\EL{rajouter le calcul des temps de retour par gktl pour etoffer un peu la partie du papier sur gktl -- il faudrait montrer un resultat qui montre que l'on obtient la bonne statistique}
+\EL{rajouter le calcul des temps de retour par gktl pour etoffer un peu la partie du papier sur gktl -- il faudrait montrer un resultat qui montre que l'on obtient la bonne statistique} \TL{Demande pas mal de travail en comparaison des autres points.
+Voir issue \#10.}
 
@@
 
-\EL{CONCLUSION TO BE REVIEWED}
+\EL{CONCLUSION TO BE REVIEWED \#7}
 

### 2019-05-22 draft_p2.tex
Update biblio #13

@@
 In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium.
-This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on \emph{instanton} \cite{grafke_instanton_2015}.
+This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
+%
+An alternative approach is based on \emph{instanton} \cite{grafke2015instanton}.
 For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}.
-Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.
+Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
 %
@@
 Discarded trajectories are re-sampled according to the remaining trajectories.
-\TL{For a recent review of the \ac{ams}, its history and applications, see~\cite{cerou_2019}.}
+\TL{For a recent review of the \ac{ams}, its history and applications, see~\cite{cerou2019adaptive}.}
 \TL{Note that we use a specific form of the the \ac{ams} in which trajectories have a fixed duration $T_a$.
-It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing} and the following results are not limited to this particular variant.}
+It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing_2018} and the following results are not limited to this particular variant.}
 
@@
 \TL{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are produced at time $t=t_i$.
-See~\cite{brewer_efficient_2018, nemoto_population-dynamics_2016} for information on the practical implementation of the algorithm.}
+See~\cite{brewer2018efficient, nemoto2016population} for information on the practical implementation of the algorithm.}
 % For each t $\{\mathbf{x}_{n}\}_{0 \leq t \leq  t_i}$, a random number of copies of the trajectory are generated.

### 2019-05-25 draft_p2.tex
Re-ecriture de la description de GKTL (#5); Commentaire sur l'influence de la perturbation des clones pour AMS et GKTL (#6); References pour algo. similaires a GKTL (#3)

@@
 \TL{Let us stress here that the flow dynamics are deterministic.
-As a result, resampled trajectories do not differ from the parent trajectory their are resampled
-from.
-A workaround is to introduce a slight perturbation in the state of the flow at the branching point (see figure~\ref{fig:illustr_AMS}).
-Because of the flow chaoticity, resampled trajectories will separate from their parent over the \textit{Lyapunov timescale} $\tau_L$.}\EL{Freddy: provide reference for Lyapunov timescale. \#12}
+Randomness must therefore be artificially introduced for the resampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
+In practice we introduce a small perturbation in the state of the flow at the branching points.
+Due to the flow chaoticity, resampled trajectories will then separate from their parent, as shown in figure~\ref{fig:illustr_AMS}.
+The timescale over which two trajectories separate due to an infinitesimal perturbation of the initial state is kown as the \textit{Lyapunov timescale}.
+It is denoted $\tau_L$ in the remainder of this paper.
+We make the important assumption that such a perturbation at branching times has a negligeabale impact on the properties of the sampled rare events.
+The validity of this assumption can be verified by performing a brute force simulation of the dynamics over a long duration, regurlary perturbing the state of the flow.
+On can eventually check that the statistics of the drag are consistent with the statistics computed from the control simulation shown in section~\ref{sec:pdfs}.}
+\EL{Freddy: provide reference for Lyapunov timescale. \#12}
 
@@
 The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-Similar ideas have been applied in a wide range of fields over the past 50 years, under different names \EL{donner des exemples \#3}, depending of the specific application domain~\cite{moral_feynman-kac_2004}.
+\TL{Similar ideas have been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
 The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
@@
 \TL{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are produced at time $t=t_i$.
-See~\cite{brewer2018efficient, nemoto2016population} for information on the practical implementation of the algorithm.}
-% For each t $\{\mathbf{x}_{n}\}_{0 \leq t \leq  t_i}$, a random number of copies of the trajectory are generated.
-This number is on average proportional to the weight $W_{n}^{i}$, such that the total number of trajectories produced at each event is equal to $N$.
-%
-For deterministic systems, a small perturbation is applied to the clones so that they can rapidly separate from their parents. 
-%
-% The parameter  is chosen by the user in order to control the strength of the selection.
-%
-The free parameter $k$ drives the amplitude of fluctuations targeted by the biased ensemble of trajectories.
-A higher value of $k$ gives a higher weight to higher fluctuations and thus drives the sampling to more extreme events.
-
-Let 
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refer formally to the probability of the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ of our dynamical system.
-%
-% $\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}} \right)$ 
-%
-By construction, the probability to obtain the same trajectory by resorting to the previous cloning procedure is given by
+See~\cite{brewer2018efficient, nemoto2016population} for information on the practical implementation of the algorithm.
+The application of this resampling at each step $t_i$ eventually leads to a biased sampling of trajectory space, for which trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+The sampled biased distribution writes
 %
@@
 \end{align}
-where the normalization factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]$ with $\mathbb{E}_{0}$ being the expectation value with respect to $\mathbb{P}_{0}$.
-%
-% obvious? 
-%and $\underset{N\rightarrow\infty}{\sim}$ meaning that this is true only asymptotically for large $N$.
-%
-%
-%The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
-%
-Eq.~\eqref{eq:Biased_Path_Approximation} relies on the mean-field approximation
-\begin{equation}
-R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}(t))dt}\right]
+where
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
+refer formally to the probability of observing the trajectory 
+$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
+The normalization factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+Note that
+\begin{equation}
+  \label{eq:mean_field}
+  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]
+\end{equation}
+with $\mathbb{E}_{0}$ being the expectation value with respect to the
+distribution $\mathbb{P}_{0}$.
+This result relies on a \textit{mean-field approximation}
+\begin{equation}
+R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right].
 \label{eq:Mean_Field_Approximation}
 \end{equation}
+where $\mathbb{E}_{i}[.]$ denote the expectation value with prespect to the biased distribtuion $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
 The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
-which by induction (and assuming similar relations for the successive $R_i$'s) eventually leads  to~\cite{giardina_direct_2006,giardina_simulating_2011}
-\begin{equation}
-\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(\mathbf{X}(t))dt}\right].\label{eq:Estimate_Lambda}
-\end{equation}
-\EL{IS IT TRUE?} \TL{Oui. Il faudrait un peu reécrire ce paragraphe pour que ca soit clair, voir issue \#5}
-The validity of this mean-field approximation and the 
-%
-%
-The killed trajectories, \emph{i.e.} with a weight smaller than $1$ \EL{true?}, are discarded from the statistics.
-%
-% no because the trajectories are cloned
-%
-% Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
-%
+%
+Pruned trajectories, are discarded from the statistics.
 Eventually, an effective ensemble of $N$ trajectories for $0 \leq t \leq T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-All trajectories are solutions of the dynamical system since we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
-\EL{quid the small perturbation to separated the clones? comment required \#6}
+A key feature of the \ac{gktl} algorithm is that the sampling procedure do not involve any alteration of the dynamics, and therefore all trajectories in the sampled ensemble are solutions of the dynamical system.
+This is true for stochastic dynamics.
+However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for determinisitc dynamics and we proceed in the same way as as described in section~\ref{sec:ams_drag}.}
+
 %
@@
 More details concerning the implementation of the algorithm and the corresponding data analysis are provided in \cite{lestang:tel-01974316,brewer2018efficient}.
-
-%
-%\EL{next paragraph needs rephrasing or suppress}
-%
-%For instance, the expectation value of the observable $O\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}\right)$ is given by
-%\begin{equation}
-%\mathbb{E}_{0}\left[O\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}\right)\right]\underset{N\rightarrow\infty}{\sim}\frac{1}{N}\sum_{n=1}^{N}O\left(\left\{ \mathbf{X}_{n}(t)\right\} _{0\leq t\leq T_{a}}\right){e}^{-k\int_{0}^{T_{a}}A(\mathbf{X}_{n}(t))dt}{e}^{T_{a}\lambda(k,T_{a})},\label{eq:GK_O_estimator}
-%\end{equation}
-%where the $X_{n}$ are the $N$ backward reconstructed trajectories.
-%Events that can be considered rare with respect to $\mathbb{\mathbb{P}}_{0}$ are oversampled following $\mathbb{P}_{k}$.
-As a result, the effective ensemble of trajectories generated by the \ac{gktl} algorithm contains a larger amount of such events, and empirical estimators such as~\eqref{eq:GK_O_estimator} have a dramatically lower statistical error.
 

### 2019-05-26 figures/illustr_AMS/make_figure_AMS.py
Update figure illustr_AMS #1

@@ -1,0 +1,78 @@
+import os
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+prefix = '/home/tlestang/transition_hdd/thibault/These/AMS/AMS_OU_AVG/run_for_figure_AMS/'
+with open(os.path.join(prefix, 'input.in'), 'r') as f:
+    lines = [line.rstrip() for line in f]
+Nc = int(lines[0])
+Ta = int(lines[1])
+T = int(lines[2])
+dt = float(lines[3])
+
+ndtAVG = int(np.floor(T/dt))
+
+idx = int(np.fromfile(os.path.join(prefix, 'killed_copie.datout'), np.int32, -1, ''))
+
+Nt = int(np.floor(Ta/dt))
+tspan = np.linspace(0,Ta-T,Nt-ndtAVG+1)
+lblidx = [1, 2, 3]
+
+filename = os.path.join(prefix, 'clone_{}_iter_1.traj'.format(idx))
+x = np.fromfile(filename, float, -1, '')
+
+series = pd.Series(x).rolling(ndtAVG).mean()
+xAVG = series.values[ndtAVG-1:]
+
+fig,ax = plt.subplots(1,1, figsize=(18,9), constrained_layout=True)
+ax.plot(tspan, xAVG, linewidth=2.5)
+
+amax = [0,0,0]
+linestyle = ['-']*3
+linestyle[idx]='--'
+for j in range(0,3):
+    filename = os.path.join(prefix, 'clone_{}_init.traj'.format(j))
+    x = np.fromfile(filename, float, -1, '')
+    series = pd.Series(x).rolling(ndtAVG).mean()
+    xAVG = series.values[ndtAVG-1:]
+
+    amax[j] = np.amax(xAVG)
+    ii = np.argmax(xAVG)
+    ax.plot(tspan, xAVG, linestyle=linestyle[j], color='k', linewidth=2.5)
+
+    ax.plot(ii*dt, amax[j], linestyle=None, marker='s', markersize=7, color='k')
+    plt.plot(tspan, amax[j]*np.ones(len(xAVG)), color='k', linestyle='--')
+
+brch_pt, = ax.plot(1.75,amax[idx], 'o', linestyle=None, markersize=10, color='r')
+
[…]

### 2019-05-26 draft_p2.tex
Update figure illustr_AMS #1

@@
   \centering
-  \includegraphics[width=.7\linewidth]{illustr_AMS/figure_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm to compute the probability that a real-valued observable $A(\mathbf{x}(t))$ reaches values larger than $Q$ throughout a trajectory of duration $T_a$.\EL{QUID of the score function mentioned in the text: is the observable?
-\#1} The initial ensemble of blue trajectories (associated with the time evolution of the observable $A$) is first generated.
-These trajectories are independent and can be simulated in parallel.
-On the basis of the respective maxima $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed blue line).
+  \includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
+  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
+Lines 1, 2 and 3 respresent the evolution of the score function for the current ensemble of trajectories.
+On the basis of their respective maximum $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
 Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then freely simulated from the branching point to the final time $T_a$.
+It is then simulated from the branching point to the final time $T_a$.
 In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This procedure is iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.}
+Depending on the application, this resampling procedure can iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.
+\TL{MAJ figure et ordonnee \#1}}
 \end{figure}

### 2019-05-26 draft_p2.tex

@@
 %
-An alternative approach is based on \emph{instanton} \cite{grafke2015instanton}.
+An alternative approach is based on \emph{instantons} \cite{grafke2015instanton}.
 For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}.
@@
 
+%% LB parameters  %
+
 % drag signal
@@
 In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.
-%In particular, extreme fluctuations mainly result from large variations of the base pressure, whereas forebody pressure variations play a lesser role.
 
@@
 In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and
+
 \begin{equation}
@@
 \end{figure}
-
 % flow scenario
@@
 Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
+
 \begin{equation}
@@
 \label{sec:rare_events_algorithms}
-
 % limitation of direct sampling
@@
 % These two algorithms will be detailed in the following.
-
 % In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
@@
 
-\subsubsection{Application of the \ac{ams} to a simple case: the \acl{ou} process}
+\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
 \label{sec:ams_ou}
-
 % Example: OU process
@@
 \end{figure}
-
 Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered. This should allow us to highlight differences with fluid dynamics.
+
 \begin{equation}
@@
 \end{equation}
-
 \TL{The \ac{tams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5$.Note that, for the process as define in~\eqref{eq:ou}, the correlation time is $\tau_c = 1$.}
@@
 \label{sec:gktl}
-
 The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
@@
 This result relies on a \textit{mean-field approximation}
+
 \begin{equation}
@@
 
+% Conclusion de cette partie
+Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a direct sampling approach of similar computational cost.
+In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
+However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two vert large fluctuations of the instantaneous drag.
+Further works must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
+
 \section{Conclusion}
 \label{conlusion}
-
-\EL{CONCLUSION TO BE REVIEWED \#7}
-
 We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
@@
 \label{sec:lbm}
+While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
+Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
+Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
+This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
+This is, in short, the rationale behind the LB approach.
+In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
+\\
+More precisely, the fluid is viewed as populations of particles
+that collide, redistribute and propagate along the different links of a discrete lattice. 
+In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
+Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
+\[
+\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
+\]
+for the mass density and the fluid momentum respectively.
+The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
+	\caption{Sketch of the D2Q9 lattice.
+Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
+By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
+	\label{fig:D2Q9}
+\end{figure}
+
+The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
+Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+Namely, an \emph{exact} streaming step
+\[
+f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
+\]
+during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
+\[
+f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
+\]
+which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
+\[
+\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
+\]
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
+\[
+f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
+\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
+\] 
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
+Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
+$\Delta x \to 0$.
+\\
+As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
+\[
+\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\]
+so that the total stress expresses as
+\begin{equation}\label{eq:def_stress}
+\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\end{equation}
+
+\EL{CONCLUSION TO BE REVIEWED \#7}
+
+We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
+\\
+In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
+By means of very long simulations---without rare event algorithm---we observed that such extreme events are caused by the concentration of vorticity very close to the base of the square.
+Because of the sweeping of the fluid structure responsible for such a situation, extreme drag levels do not persist over time.
+The lifetime of extreme drag fluctuations is of the order of the turnover time, and the corresponding drag signal is very peaked around the extreme value.
+Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are very well described by an exponential \ac{pdf}.
+This property can be linked to the phenomenology of extremes of the time-averaged drag.
+Indeed, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations or an exceptional succession of typical fluctuations that
+add up forming an extremely large value of the average.
+Such a phenomenology can be described by a stochastic process with an exponential \ac{pdf}.
+\\
+On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
+In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling approach.
+This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which resampled trajectories separate from their parent following the
+addition of a small perturbation in the initial conditions.
+The \ac{gktl} has been applied to the sampling of trajectories displaying extreme fluctuations of the time averaged drag.
+In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a direct approach, allowing the simulation of trajectories that are out of reach without algorithm.
+\\
+The application of rare event algorithms relies on the definition of a score function, on which the selection of trajectories is based upon.
+Although the dynamics itself is a black-box from the point of view of the algorithm, the efficiency of the sampling depends on the phenomenology.
+For complex dynamics, including turbulent flows, the choice of the score function is made difficult by the very little knowledge one has of the rare events of interest.
+This is, after all, our motivation for the simulation of these events.
+In such situations, it is then tempting to settle for a straightforward choice for the score function.
+This is the approach we followed in this work, setting the score function to the drag itself.
+However, for dynamics as complex as the turbulent flow around an obstacle, it is unlikely that selecting trajectories based on the value of the drag favours the occurrence of extreme drag fluctuations
+in the future.
+Under such circumstances, the \ac{tams} fails drastically.
+Successful usage of the \ac{tams} and similar splitting algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that good score functions
+are very difficult to identify, if even possible.
+A popular direction in current research involves taking advantage of the recent advances in learning methods to compute score functions beforehand.
+For the time-averaged drag, we showed that the \ac{gktl} algorithm is able to greatly outperform a direct sampling approach.
+This is because the sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
+
+\appendix*
+\section{The \acl{lbm}}
+\label{sec:lbm}
 

### 2019-05-26 draft_p2.tex
Add figur timeseries_extrms_AVG

@@
 	\centering
-	%\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the time-averaged drag $F_T$ with averaging period $T=10~\tau_c$ and $\tau_c$ being the correlation time of the drag.
-$\sigma_T$ denotes the standard deviation of $F_T$.
-\EL{FIGURE IS MISSING} \TL{Addressed in issue \#9}}
+	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+	\caption{\TL{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the time-averaged drag $F_T$ observed in a control timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ with $T_{Tot} = 10^6\tau_c$ and $T = 10 \tau_c$.
+Here $\tilde{f}_d$ denotes the zero-mean, unit variance instantaneous drag.}}
 	\label{fig:extreme_avg}
@@
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$ and sample a set of extreme events.
-By taking $a=5.2~\sigma _T$  with $\sigma_T$ being the standard deviation of the fluctuations of $F_T$, $84$ independent extreme events have been selected.
+By taking $a=5.2~\sigma_T$  with $\sigma_T$ being the standard deviation of the fluctuations of $F_T$, $84$ independent extreme events have been selected.
 

### 2019-05-26 figures/timeseries_extrms_AVG/timeseries_extrms_AVG.py

@@ -14,7 +14,7 @@ sortedI = np.loadtxt('list.txt', dtype=int)
 sig=0.0412
 m=0.0252
 # Gets the value of mu and sigma for time-averaged drag
-musigma = np.fromfile("../PDF_AVG/musigma_AVG_10.dat", float, -1, "");
+musigma = np.fromfile("musigma_AVG_10.dat", float, -1, "");
 
 T = 10 # Physical duration of instant. trajectories
 
@@ -38,7 +38,7 @@ for i in range(0,Nsubplots):
     col = i % 3
     drag_signal, = axes[row,col].plot(np.linspace(0,T,N),f)
     axes[row,col].set_xlim((0, T))
-    axes[row,col].set_title(r"$F = {:2.1f}\sigma _T$".format(F_avg))
+    axes[row,col].set_title(r"$F_T = {:2.1f}\sigma _T$".format(F_avg), fontsize=26)
     # Compute min and max for instant drag signal to
     # impose common ylim to all subplots (see below)
     maxArray[i]=np.amax(f)
@@ -55,12 +55,12 @@ for i in range(0,6):
     row = i // 3
     col = i % 3
     axes[row,col].set_ylim((globalDragMinimum,globalDragMaximum))
-    axes[row,col].set_xlabel(r'$t/ \tau_c$',fontsize=18)
-    axes[row,col].set_ylabel(r'$f_d(t)$',fontsize=18)
+    axes[row,col].set_xlabel(r'$t/ \tau_c$',fontsize=28)
+    axes[row,col].set_ylabel(r'$f_d(t)$',fontsize=28)
     axes[row,col].set_yticks([0,2,4,6])
     axes[row,col].set_yticklabels(['${:2.0f}\\sigma$'.format(x) \
                                    for x in axes[row,col].get_xticks()])
-    axes[row,col].tick_params(axis='both', which='major', labelsize=16)
+    axes[row,col].tick_params(axis='both', which='major', labelsize=22)
 
 fname = 'timeseries_extrms_AVG.png'
 plt.savefig(fname)

### 2019-05-26 figures/IS_GKTL/IS_GKTL.py
Fix x label for figure IS_GKTL #4

@@ -38,8 +38,8 @@ ax.set_yticks(ax.get_yticks()[::2]) # Plot only half of ticks
 ax.set_xticklabels(['$-2\\sigma_T$','$0$','$2\\sigma_T$','$4\\sigma_T$','$6\\sigma_T$'])
 ax.tick_params(axis='both', which='major', labelsize=16)
 
-ax.set_xlabel('$F$', fontsize=22)
-ax.set_ylabel('$\\rho_{k}(F_T = F)$', fontsize=22)
+ax.set_xlabel('$\\tilde{F}_T$', fontsize=22)
+ax.set_ylabel('$\\rho_{k}(\\tilde{F}_T = F)$', fontsize=22)
 ax.legend(loc='best', fontsize=16)
 
 # Prints figure

### 2019-05-27 draft_p2.tex
Fix x label for figure IS_GKTL #4
Make colors for commands TL and EL less aggressive
Iteration de l'abstract par @tlestang #8

@@
 
-\newcommand{\EL}[1]{{\color{red}{#1}}}
-\newcommand{\TL}[1]{{\color{green}{#1}}}
+\definecolor{myred}{HTML}{b30000}
+\definecolor{mygreen}{HTML}{008000}
+\newcommand{\EL}[1]{{\color{myred}{#1}}}
+\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
 
@@
 \begin{abstract}
-	\EL{ ABSTRACT TO BE REVIEWED \#8}
-This study evaluates the relevance of  rare-event sampling to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is here to replace a long simulation by a set of much shorter simulations (running in parallel) with dynamics that are replicated or abandoned in order to sample large-amplitude events more frequently.
-%
-Such technique has already proved its efficiency for simulating extreme events in {...} with significant run-time savings.
+\TL{  This study evaluates the relevance of rare-event sampling type approaches to make accessible the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, wich corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently.
+%
+  Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of events far from reach from brute force sampling approaches.
   Application to fluid-structure interaction problems remains mainly open. 
 %
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely {...}
-Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort.
-However, {limitations and pitfalls}.
-% Importantly, such techniques also provide direct access to the flow scenario yielding extreme events. In the present case, it is found that {<blocking effect>}.
+  The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
+  Practical evidence is given that the fast sweeping time of fluid structures has a drastic influence on the efficiency of the sampling with both algorithms.
+  While is is shown that the \ac{ams} resampling does not lead to rare trajectories with siginificant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
 \end{abstract}
@@
 The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-	The dashed line denotes the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-\EL{$F_T-\overline{F_T}$ in the figure! \#4}}
+	The dashed line denotes the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.}
 \end{figure}

### 2019-05-27 draft_p2.tex
Relecture generale #11 #7 #2 #13 #6

@@
 \TL{  This study evaluates the relevance of rare-event sampling type approaches to make accessible the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, wich corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently.
+  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, which corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently.
 %
@@
   Practical evidence is given that the fast sweeping time of fluid structures has a drastic influence on the efficiency of the sampling with both algorithms.
-  While is is shown that the \ac{ams} resampling does not lead to rare trajectories with siginificant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
+  While is is shown that the \ac{ams} re-sampling does not lead to rare trajectories with significant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
 \end{abstract}
@@
 Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from mean value by ${\cal{O}}(10)$ standard deviations.
+``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
 In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Such anomalous efforts can threaten the structural integrity of embedded structures.
-%
-From the viewpoint of (chaotic) dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions (among degrees of freedom) possibly responsible for extreme fluctuations is not successful.
-Therefore, simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
-Nevertheless, this requires very long simulation since these events are also very rare.
-It is here our motivation to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost.
+Indeed, such anomalous constraints can threaten the structural integrity of embedded structures.
+%
+From the viewpoint of (chaotic) dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
+The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
+Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
+Nevertheless, this requires very long simulations since these events are also very rare.
+Our motivation is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
 %
@@
 %
-Rare-event sampling is an umbrella term for numerical methods intended to sample preferentially regions of the phase space which would otherwise be rarely visited through brute-force simulation.
+\TL{Rare-event sampling refer to a large body of methods that aim at sampling preferentially regions of phase space that correspond to rare events, that would otherwise be accessed with a very low probability through a brute-force sampling.}
 %
@@
 For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}.
-Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
-%
-An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimized.
+Therefore, suitable numerical schemes can  be used to minimise the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
+%
+An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimised.
 %
@@
 In this paper, a more general approach is investigated.
-It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.
+\TL{Je ne comprends pas cette phrase}\EL{It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.}
 % dedicated to dynamical systems for which an action cannot be easily derived or minimized,
@@
 %
-Even though such ideas trace back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
+Even though such ideas date back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
 %
@@
 An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \EL{is it TRUE? \#11}
+%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
 % 
@@
 %
-Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
+\EL{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
   This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
 Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.
+}
+\TL{Super!}
 
@@
 \includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the center of a channel in two dimensions.
+\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
 The flow is artificially damped near the end of the channel.
 In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary color map from blue (negative values) to red (positive values).}
+The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
 \label{fig:illustr_ecoulement}
@@
 % rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream. They interact with each other and grow in size as expected in two-dimensional turbulence. 
-The dimension of the grid is such that the size of the eddies that eventually hit the square is comparable to the size of this latter, which results in strong fluctuations of the drag acting on the square.
+Turbulent eddies generated in the near-wake of the grid are carried downstream.
+They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
+The dimension of the grid is such that the size of the eddies that hit the square is comparable to its side, resulting in strong fluctuations of the drag acting on the square. 
 %
@@
 %
-The simulated flow develops in a long plane channel of dimension $512 \times 128$ mesh points.
-The squared obstacle has size $R=16$ (in mesh unit) and is located at the center of the channel.
+The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
+The squared obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
 The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
@@
 Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
-The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalised by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
 The distributions are imposed at equilibrium (see Appendix).
@@
 Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the oulet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
+The extrapolated distributions are evaluated through a regularisation procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
 
@@
 The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and expresses formally as 
+The \textit{drag} is defined as the resulting force in the streamwise $x$-direction, formally 
 \begin{equation}
 \label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x})  
+f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
 \end{equation}
@@
 %
-The typical timescale (turnover time) of drag fluctuations may be estimated from dimensional analysis as
+The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
 \begin{equation}
 \label{eq:turnover_time}
-\tau_0 = \frac{R}{U}
+\tau_0 = \frac{R}{U},
 \end{equation}
 where $R$ is the size of the square and $U$ the averaged velocity. 
-The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+\EL{The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.}\TL{Redondant avec le commentaire de (1)?}
 %
@@
 Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-It is observed that, typically, some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
-Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
+It is observed that, typically, some vorticity that is generated near the front part of the obstacle rapidly detaches from it and is advected downstream by the flow.
+Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations mostly result from pressure variation at the front of the obstacle.
 In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.
@@
   \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
+The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
   \label{fig:typical_drag_signal}
@@
 Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+Fluctuations appear unpredictable and the signal exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
 It is therefore natural to model the drag as a scalar random process $f_d(t)$.
@@
   \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
-The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
+The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
   \label{fig:typical_drag_signal}
@@
 
+% drag statistics
+% pdf
 Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
+It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method. 
 The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
@@
 %
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle surface.
 \begin{figure}
@@
 % drag statistics
-% PDF and correlation time
-%
-Drag fluctuations have been recorder during a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$.
-This very long simulation will be referred to as the \textit{control run} in the remainder.
-It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method.
-%The related execution time is about a few weeks (wall-clock time).
-The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
-%
-Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a control surface corresponding the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that latter case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This suggests that the asymmetry of the \ac{pdf} and the positive exponential tail are closely related to the no-slip condition on the obstacle.
-%
-%
+% correlation time
 Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
@@
 %
-At first, the phenomenology of extreme fluctuations of the drag is investigated by means of \textit{direct sampling} of the control run.
-This brute-force approach is here used as opposed to optimized approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
+The phenomenology of extreme fluctuations of the drag is first investigated through brute-force sampling, using the control run.
+Brute-force sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
 The direct approach will provide a trustworthy baseline for the validation of rare-events algorithms. 
-
-How rare is a fluctuation $f_d \geq a$ can be quantified by its so-called \textit{return time} $r(a)$.
-This latter is defined as the average time between two consecutive occurrences of events of amplitudes $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
+How rare is a fluctuation $f_d \geq a$ can be quantified by its \textit{return time} $r(a)$.
+This latter is defined as the average time between two consecutive occurrences of fluctuations with amplitude $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
 \begin{figure}
@@
   \includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurence of the event: $r(a) = \mathbb{E}[\tau]$.
+  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurrence of the event: $r(a) = \mathbb{E}[\tau]$.
 Note that we assume $r(a) >> \tau_0$, and therefore consider that such events are points in time.}}
@@
 According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
-Precisely, 104 independent fluctuations \EL{excursion, terminologie correcte?} \TL{peut etre, mais fluctuation est correct et plus precis?.}with $f_d(t) \geq 7.6~\sigma$ have been identified.
-For convenience, each fluctuation will be characterized in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
+Precisely, 104 independent fluctuations \EL{excursion, terminologie correcte?} \TL{peut etre, mais fluctuation est correct et plus precis?.}with $f_d(t) \geq 7.6\sigma$ have been identified.
+For convenience, each fluctuation will be characterised in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
 %
@@
 %
-It is found that the base pressure variation contributes typically $80\%$ to the overall drag fluctuation.
-Therefore, extreme amplitude of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream the obstacle.
+It is found that the base pressure variation contributes typically as $80\%$ of the overall drag fluctuation.
+Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
 Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure. 
@@
 	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centered around extreme fluctuations occurring  at $t=t^{\star}$.
+	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
 The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
@@
 %
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strongly enhanced drag.
-In contrast, nothing special happens near the forebody of the obstacle.
+This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strong drag. In contrast, nothing special happens near the forebody of the obstacle.
 
 The high pressure drop on the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents him to be swept away downstream.
+As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it to be swept away.
 % 
@@
 
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is now proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square.
-Precisely,
+Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is now proposed to characterise the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square,
 \begin{equation}
   \label{eq:avg_shear_def}
-  \overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x}
+  \overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
 \end{equation}
@@
 In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event was of the order of the sweeping time of the flow past the obstacle, or $\tau_c$.
+We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
 In practical applications, this duration may be much smaller than the characteristic response time of the material structure subject to these fluctuations.
@@
 %
-The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained straightforwardly by filtering in time the instantaneous drag signal $f_d(t)$. 
+The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$ and sample a set of extreme events.
@@
 %
-This observation may be explained by the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_computing_2018} (see Fig.~\ref{fig:pdf_drag}). 
+This observation may be explained by the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_article_p1}\TL{reference vers l'article JSTAT} (see Fig.~\ref{fig:pdf_drag}).
 Indeed, let $X$ be a random variable with \ac{pdf} $\mathbb{P}(X)$ and standard deviation $\sigma_X$.
@@
 According to Eq.~\eqref{eq:return_time}, the return time of a fluctuation $f_d \geq a$ scales as $r(a) \propto e^{\lambda a}$.
-As a consequence, the computational cost required to sample (directly) events of amplitude $f_d \geq a$ diverges exponentially as $e^{\lambda a}$.
+As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially as $e^{\lambda a}$.
 %% objective
 % In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than that related to return times.
+The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than the related return time.
 %
@@
 %
-Both algorithms consider an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in the phase space.
+Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
 In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-In brief, the trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} allowing us to sample more frequently rare events of some observables of interest, \textit{e.g.} the drag.
-These selection rules are designed so that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories receive a weight whose knowledge makes it possible to calculate the probability and expectation values of the observables.
+In a nutshell, trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable of interest---\textit{e.g.} the drag.---with a higher frequency.
+These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
+The generated trajectories are associated a weight, from which one is able to compute the probability and expectation values of observables. 
 % These two algorithms will be detailed in the following.
@@
   \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
-Lines 1, 2 and 3 respresent the evolution of the score function for the current ensemble of trajectories.
+Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
 On the basis of their respective maximum $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
@@
 In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this resampling procedure can iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.
+Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.
 \TL{MAJ figure et ordonnee \#1}}
@@
 %a sequence of events that can be sampled with a higher probability  
-a sequence of ``short-cuts'' that lead more rapidly to the realization of this event \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
+a \TL{sequence of short excursions that are easier to sample}~\cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+The related selection-mutation step of this algorithm \TL{relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$} and is sketched in Fig.~\ref{fig:illustr_AMS}.
 %
@@
 %
-\TL{The selection of trajectories rely on a \textit{score function} $\xi (\mathbf{x}(t),t)$.}
+
 Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$.
@@
 \TL{Note that we use a specific form of the the \ac{ams} in which trajectories have a fixed duration $T_a$.
-It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing_2018} and the following results are not limited to this particular variant.}
+It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing_2018} and the following results are therefore not limited to this particular variant.}
 
@@
 \end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered. This should allow us to highlight differences with fluid dynamics.
+Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered.
+This allow us to highlight differences with fluid dynamics.
 
@@
 \end{equation}
-\TL{The \ac{tams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5$.Note that, for the process as define in~\eqref{eq:ou}, the correlation time is $\tau_c = 1$.}
-Our objective to sample fluctuations of $x\geq a$ with $a$ being very large compared to typical values of $x$.
-The score function is here $x(t)$ and only one single trajectory is re-sampled at each iteration.
+\TL{The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+Note that, for the process defined in~\eqref{eq:ou}, the correlation time is $\tau_c = 1$.}
+Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to typical values of $x$.
+The score function is simply $x(t)$ and only one single trajectory is re-sampled at each iteration.
 %
@@
 The computational cost of the algorithm after $J$ iterations is related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{tams} algorithm with that of a direct sampling. 
+Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
 In the latter, the typical computational cost is simply the return time $r(a)$.
-Obviously, the successive re-samplings of the \ac{tams} algorithm lead very rapidly to trajectories exhibiting extreme fluctuations.
+Obviously, the successive re-samplings of the \ac{ams} algorithm lead very rapidly to trajectories exhibiting extreme fluctuations.
 The related computational cost is many orders of magnitude lower than that of direct sampling.
 
-Undoubtedly, the \acl{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} or \ac{tams} algorithm.
-Indeed, the state space is one-dimensional and the choice of the score function is straightforward. It is the process $x(t)$ itself.
+Undoubtedly, the \acl{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+Indeed, the state space is one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
 In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
@@
 	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
-The \ac{tams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
+The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
 \end{figure}
@@
 %
-% Can the \ac{tams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
+% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
 % To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
@@
 Our aim is now to sample flow evolution that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-In contrast with the \acl{ou} process, the phase space is here higlhy-dimensional with intrinsically chaotic dynamics.
+In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
 In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However, a natural simple choice for which we will opt in this study is certainly $\xi (\mathbf{x}(t),t) = f_d(t)$. 
+However, a natural simple choice for which we will opt for in this study is $\xi (\mathbf{x}(t),t) = f_d(t)$. 
 \TL{Let us stress here that the flow dynamics are deterministic.
-Randomness must therefore be artificially introduced for the resampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice we introduce a small perturbation in the state of the flow at the branching points.
-Due to the flow chaoticity, resampled trajectories will then separate from their parent, as shown in figure~\ref{fig:illustr_AMS}.
-The timescale over which two trajectories separate due to an infinitesimal perturbation of the initial state is kown as the \textit{Lyapunov timescale}.
+Randomness must therefore be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
+In practice we introduce a small perturbation in the state of the flow at the branching points (see figure~\ref{fig:illustr_AMS}).
+Due to the chaoticity of the flow, re-sampled trajectories will then separate from their parent.
+The timescale over which two trajectories separate due to an infinitesimal perturbation of the initial state is known as the \textit{Lyapunov timescale}.
 It is denoted $\tau_L$ in the remainder of this paper.
-We make the important assumption that such a perturbation at branching times has a negligeabale impact on the properties of the sampled rare events.
-The validity of this assumption can be verified by performing a brute force simulation of the dynamics over a long duration, regurlary perturbing the state of the flow.
+We make the important assumption that such a perturbation at branching times has a negligible impact on the properties of the sampled rare events.
+The validity of this assumption can be verified by performing a brute force simulation of the dynamics over a long duration, regularly perturbing the state of the flow.
 On can eventually check that the statistics of the drag are consistent with the statistics computed from the control simulation shown in section~\ref{sec:pdfs}.}
@@
 %
-The \ac{tams} is applied with $N=256$ trajectories of duration $T_a = 20\tau_c$.
+The \ac{ams} is applied with $N=256$ trajectories of duration $T_a = 20\tau_c$.
 %
@@
 These maxima are depicted by the stars in Fig~\ref{fig:AMS_drag_resampling}.
-Importantly, one can see that re-sampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories. This in contrast with the case of the \acl{ou} process.
+Importantly, one can see that re-sampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
+This contrasts with the case of the \acl{ou} process (Fig.~\ref{fig:comparaison_temps_de_retour}).
 
@@
   \caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
-In this experiment, the \ac{tams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
+In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
 Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
-  The \ac{tams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the trajectory displaying the highest maximum in the initial ensemble.}
+  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the trajectory displaying the highest maximum in the initial ensemble.}
 \end{figure}
@@
 %
-Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{tams} re-sampling procedure. 
+Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{ams} re-sampling procedure. 
 %
@@
 %
-Interestingly, it is obtained that all trajectories ultimately originate from one unique initial trajectory. 
+Interestingly, all trajectories ultimately originate from one unique initial trajectory. 
 %
 This feature may be explained as follows.
-Because the dynamics has a memory, it takes a certain time $\tau_L$ before a re-sampled trajectory separates from its parent.
+Because the dynamic process has a memory, it takes a certain time $\tau_L$ before a re-sampled trajectory separates from its parent.
 In our situation, this memory originates from the deterministic nature of fluid dynamics.
 %
-\TL{In our case, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale of the drag process.}
-Therefore, the re-sampling of a trajectory branched at $t=t^{\star}$ corresponding to the maximal drag amplitude cannot lead to larger values for $t^{\star} \leq t \leq t^{\star}+\tau_L$.
+\TL{Importantly, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale $\tau_L$ of the drag process.}
+As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$---the maximum drag amplitude---cannot lead to larger values for $t^{\star} \leq t \leq t^{\star}+\tau_L$.
 For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based.
@@
 The ratio $u_{rms}/U < 1$ in our case of grid-generated turbulence implies that $\tau_L > \tau_c$.
-\EL{does it make sense to you?} \TL{A priori oui.
-Mais j'aimerai prendre 10-15min pour bien y réféchir \#2}
 
@@
 %
-In summary, it appears that a straightforward application of the \ac{tams} by taking the drag as score function does not allow us to efficiently sample extreme fluctuations.
+In summary, it appears that a straightforward application, defining the score function as the drag itself, does not allow us to efficiently sample extreme fluctuations.
 %
@@
 
-\subsection{Extreme time-averaged drag forces with the \acl{gktl} algorithm}
+\subsection{Importance sampling of extreme time-averaged drag forces with the \acl{gktl} algorithm}
 \label{sec:gktl}
 The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{tams} algorithm could be used in the same way as before by choosing $F_T$ as score function. However, this would lead to the same pitfall.
+The \ac{ams} algorithm could be used in the same way as in section~\ref{sec:ams_drag}, \textit{i.e.} by setting $F_T$ as the score function $\xi$.
+However, this leads to similar unsatisfactory conclusions.
 % the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
@@
 %
-Similarly to the \ac{tams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
+Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
 In this case, however, the trajectories interact dynamically. 
@@
 %
-The weights are chosen so that, after several iterations of the algorithm, the trajectories generated are distributed according to a biased probability distribution in favor of trajectories with large values of the (time-averaged) observable of interest.
+The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favours trajectories related to large values of the (time-averaged) observable of interest.
 %
@@
 % averaging time~\cite{ref_large_dev}.
-% By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
+% By contrast with the \ac{ams} and \ac{ams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
 % Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
@@
 \label{sec:gktl_description}
-As with the \ac{tams} algorithm, the \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
 %
@@
 \begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalization factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
 \label{eq:Weight}
@@
 \TL{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are produced at time $t=t_i$.
-See~\cite{brewer2018efficient, nemoto2016population} for information on the practical implementation of the algorithm.
-The application of this resampling at each step $t_i$ eventually leads to a biased sampling of trajectory space, for which trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+See~\cite{brewer2018efficient, lestang:tel-01974316} for more information on the practical implementation of the algorithm.
+The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling of trajectory space, for which trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
 The sampled biased distribution writes
@@
 \begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)
+\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
 \label{eq:Biased_Path_Approximation}
@@
 where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refer formally to the probability of observing the trajectory 
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),$ 
+refers formally to the probability of observing the trajectory 
 $\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalization factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
 Note that
@@
   \label{eq:mean_field}
-  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]
+  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
 \end{equation}
@@
 \begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right].
+R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
 \label{eq:Mean_Field_Approximation}
 \end{equation}
-where $\mathbb{E}_{i}[.]$ denote the expectation value with prespect to the biased distribtuion $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
-%
-Pruned trajectories, are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories for $0 \leq t \leq T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+where $\mathbb{E}_{i}[.]$ denote the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$~\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
+%
+Pruned trajectories are discarded from the statistics.
+Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
 A key feature of the \ac{gktl} algorithm is that the sampling procedure do not involve any alteration of the dynamics, and therefore all trajectories in the sampled ensemble are solutions of the dynamical system.
 This is true for stochastic dynamics.
-However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for determinisitc dynamics and we proceed in the same way as as described in section~\ref{sec:ams_drag}.}
+However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics and we proceed in the same way as as described in section~\ref{sec:ams_drag}.}
 
@@
 %
-More details concerning the implementation of the algorithm and the corresponding data analysis are provided in \cite{lestang:tel-01974316,brewer2018efficient}.
 
@@
 % Choice of parameters and perturbation
-The computational cost $C_{gktlS}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-In practice, $N=16384$ and $T_a = T = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+The computational cost $C_{gktl}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+In practice we set $N=16384$ and $T_a = T = 10\tau_c$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
 A cloning period $\tau$ that is too short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
 On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a rule of thumb is to set $\tau \approx \tau_c$. We took $\tau = \tau_c /2$.
-%
+As a result, a rule of thumb is to set $\tau \approx \tau_c$.
+In the following experiments we set $\tau = \tau_c /2$.
+%
+
 Three numerical experiments corresponding to three different values of the parameter $k$ have been performed.
@@
 %
-As a proof of realizability, Fig.~\ref{fig:IS_GKTL} displays the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag, $\tilde F_T \equiv F_T - \overline{F_T}$:
+Fig.~\ref{fig:IS_GKTL} displays the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag, $\tilde F_T \equiv F_T - \overline{F_T}$:
 \begin{equation}
@@
 %
-indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is over-peaked.
-%
-Such finite-size effect is expected to gradually increase as the bias amplitude  $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-This point has been very recently examined in \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
+Indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is over-peaked\TL{Anglais?}.
+%
+Such a finite-size effect is expected to gradually increase as the bias amplitude  $k$ is increased.
+%
+An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$, addressed in recent studies~\cite{nemoto2017finite,Guevara_Hidalgo_2018}.
+
 % how to assess the value of k_max ?
-In this study, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
+In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
 This diversity may be monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
@@
 		\hline
-		$~k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
+		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
 		\hline
@@
 		\hline
-		${N}_{direct}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
+		${N}_{brute-force}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
 		\hline
@@
 		Note that trajectories that overlap for more than half of their duration are counted as one.
-		${N}_{direct}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$ of the same duration.}
+		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realisations of $F_{T}$ of the same duration.}
 	\label{tbl:nbTraj_tbl}
@@
 %
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
-Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged $F_{T} \geq a$ for the different values of the bias parameter $k$.
+The efficiency of the \ac{gktl} algorithm is now assessed with respect to brute-force sampling. 
+Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
 Note that trajectories that overlap for more than half of their duration are counted as one.
@@
 %
-Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a direct sampling approach of similar computational cost.
+Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a brute-force sampling approach of similar computational cost.
 In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
 By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two vert large fluctuations of the instantaneous drag.
-Further works must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
+However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
+Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
 
@@
 Voir issue \#10.}
-
-% Conclusion de cette partie
-Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a direct sampling approach of similar computational cost.
-In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two vert large fluctuations of the instantaneous drag.
-Further works must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
 
@@
 add up forming an extremely large value of the average.
-Such a phenomenology can be described by a stochastic process with an exponential \ac{pdf}.
-\\
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling approach.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which resampled trajectories separate from their parent following the
-addition of a small perturbation in the initial conditions.
-The \ac{gktl} has been applied to the sampling of trajectories displaying extreme fluctuations of the time averaged drag.
-In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a direct approach, allowing the simulation of trajectories that are out of reach without algorithm.
+Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
 \\
@@
 This is, after all, our motivation for the simulation of these events.
-In such situations, it is then tempting to settle for a straightforward choice for the score function.
-This is the approach we followed in this work, setting the score function to the drag itself.
-However, for dynamics as complex as the turbulent flow around an obstacle, it is unlikely that selecting trajectories based on the value of the drag favours the occurrence of extreme drag fluctuations
-in the future.
-Under such circumstances, the \ac{tams} fails drastically.
-Successful usage of the \ac{tams} and similar splitting algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that good score functions
-are very difficult to identify, if even possible.
-A popular direction in current research involves taking advantage of the recent advances in learning methods to compute score functions beforehand.
-For the time-averaged drag, we showed that the \ac{gktl} algorithm is able to greatly outperform a direct sampling approach.
-This is because the sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
-
-\appendix*
-\section{The \acl{lbm}}
-\label{sec:lbm}
-While standard numerical methods rely on the discretization of the Navier-Stokes equations, the LB method emerged in the late 80s as a conceptually different approach.
-Fluid motions are simulated at a mesoscopic level that is intermediate between the microscopic and the macroscopic.
-Capturing the kinetic behavior of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, kinetic equations simpler than the original Boltzmann equation may be designed by retaining only the basic features that pertain at the macroscopic level.
-This is, in short, the rationale behind the LB approach.
-In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
-\\
-More precisely, the fluid is viewed as populations of particles
-that collide, redistribute and propagate along the different links of a discrete lattice. 
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$, where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-Therefore, the LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighboring node, is consecutive to an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which can be viewed as the relaxation of local densities towards an absolute equilibrium at the macroscopic level; the time-scale $\tau_\nu$ is related to the kinematic viscosity of the fluid by 
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas. The equilibrium function is given (assuming Einstein summation convention) by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations (with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$) as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
-\\
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-
-\EL{CONCLUSION TO BE REVIEWED \#7}
-
-We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
-\\
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of very long simulations---without rare event algorithm---we observed that such extreme events are caused by the concentration of vorticity very close to the base of the square.
-Because of the sweeping of the fluid structure responsible for such a situation, extreme drag levels do not persist over time.
-The lifetime of extreme drag fluctuations is of the order of the turnover time, and the corresponding drag signal is very peaked around the extreme value.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are very well described by an exponential \ac{pdf}.
-This property can be linked to the phenomenology of extremes of the time-averaged drag.
-Indeed, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations or an exceptional succession of typical fluctuations that
-add up forming an extremely large value of the average.
-Such a phenomenology can be described by a stochastic process with an exponential \ac{pdf}.
 \\
@@
 In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling approach.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which resampled trajectories separate from their parent following the
+This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent following the
 addition of a small perturbation in the initial conditions.
 The \ac{gktl} has been applied to the sampling of trajectories displaying extreme fluctuations of the time averaged drag.
-In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a direct approach, allowing the simulation of trajectories that are out of reach without algorithm.
+In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a brute-force approach, allowing the simulation and computation of statistics for trajectories that are out of reach without algorithm.
+The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
+It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
 \\
-The application of rare event algorithms relies on the definition of a score function, on which the selection of trajectories is based upon.
-Although the dynamics itself is a black-box from the point of view of the algorithm, the efficiency of the sampling depends on the phenomenology.
-For complex dynamics, including turbulent flows, the choice of the score function is made difficult by the very little knowledge one has of the rare events of interest.
-This is, after all, our motivation for the simulation of these events.
-In such situations, it is then tempting to settle for a straightforward choice for the score function.
-This is the approach we followed in this work, setting the score function to the drag itself.
-However, for dynamics as complex as the turbulent flow around an obstacle, it is unlikely that selecting trajectories based on the value of the drag favours the occurrence of extreme drag fluctuations
-in the future.
-Under such circumstances, the \ac{tams} fails drastically.
-Successful usage of the \ac{tams} and similar splitting algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that good score functions
+Successful usage of the \ac{ams}, \ac{gktl} and similar methods for complex flows relevant to industrial or environmental situations will require coping with the fact that quality score functions
 are very difficult to identify, if even possible.
 A popular direction in current research involves taking advantage of the recent advances in learning methods to compute score functions beforehand.
-For the time-averaged drag, we showed that the \ac{gktl} algorithm is able to greatly outperform a direct sampling approach.
-This is because the sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
 
@@
 \]
-during which particles move with their own velocity to a neighboring node, and an instantaneous collision step
+during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
 \[

### 2019-07-10 draft_p2.tex
Add plot and text about return times #10

@@
 %
-Indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
+Indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'', as most of the trajectories (in the biased ensemble) overlap. 
 This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is over-peaked\TL{Anglais?}.
@@
 %
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to brute-force sampling. 
 Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
+Trajectories that overlap for more than half of their duration are counted as one.
 %\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations one can expect from simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$, \emph{i.e.} with the same computational cost $C_{gktl}$.
+This number is compared to the expected number of fluctuations obtained from simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$, \emph{i.e.} with the same computational cost $C_{gktl}$.
+Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a brute-force sampling approach of similar computational cost.
 %
 %This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-
 %\begin{figure}
@@
 
-% Conclusion
-%
-Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a brute-force sampling approach of similar computational cost.
+\subsubsection{Computation of return times}
+\label{sec:return_times}
+Statistics for extreme events can be computed from the biased ensemble by inverting eq.~\eqref{eq:Biased_Path_Approximation}.
+For instance, it is possible to compute return times (see fig.~\ref{fig:illustrate_return_time}) for
+events unreachable by a direct approach.
+
+In the following we use the \ac{gktl} algorithm with $N$ trajectories of duration $T_a > T$.
+Each trajectory in the biased ensemble results in a timeseries for the time-averaged drag:
+\begin{equation}
+  \label{eq:time_averaged}
+  F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
+\end{equation}
+
+The return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
+\begin{equation}
+r(a) = - \frac{T_a - T}{\ln (1-p(F_T \geq a))}.
+\end{equation}
+The probability $p(F_T \geq a)$ can be estimated from the biased ensemble, inverting eq.~\eqref{eq:Biased_Path_Approximation}:
+\begin{equation}
+p(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a),
+\end{equation}
+with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise.
+That is, summing the weights of the timeseries which maximum is above $a$.
+
+Figure~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
+It comapres return times plots for two values of the bias $k$ and a fixed computational cost $N \times T_a$ to the return times obtained from a timeseries of duration $T_a = N \times T_a$.
+While a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows for the computation of statistics for drag fluctuations having a return time several order of magnitude above the computational cost.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
+  \caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle. $\tilde{F}_T$ denotes the time-averaged drag with zero mean.
+    The continuous lines are obtained based on the biased ensemble of trajectories generated by the \ac{gktl} algorithm using $N=1024$ and $T_a=30\tau_c$.
+The dashed line is the average return time plot obtained from a timeseries of duration $T_{tot} = N\times T_a$.}
+\end{figure}
+
+\subsubsection{Discussion}
+\label{sec:discussion_gktl}
+
+%
 In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
@@
 Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{rajouter le calcul des temps de retour par gktl pour etoffer un peu la partie du papier sur gktl -- il faudrait montrer un resultat qui montre que l'on obtient la bonne statistique} \TL{Demande pas mal de travail en comparaison des autres points.
-Voir issue \#10.}
 

### 2019-07-10 figures/return_times_GKTL/plot_return_times_GKTL.py

@@ -1,0 +1,83 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import interpolate
+import returntimes_GKTL as rtt
+
+def get_interpolated_return_times(directories, Ta, T, k, N, nb_points=10):
+    nb_directories = len(directories)
+
+    rmax = np.empty(nb_directories)
+    rmin = np.empty(nb_directories)
+    interpolants = []
+    for i, dirname in enumerate(directories):
+        a, ra = rtt.compute_return_times_no_dups(dirname, Ta,
+                                                 T, k, N)
+        rmax[i] = np.amax(ra); rmin[i] = np.amin(ra)
+        interpolants.append(interpolate.interp1d(ra, a, kind='next', fill_value='extrapolate'))
+    ra_interp = np.logspace(np.log(np.amin(rmin))/np.log(10),
+                            np.log(np.amax(rmax))/np.log(10), nb_points)
+    
+    interpolated_a = np.empty((nb_points, nb_directories))
+    for i, dirname in enumerate(directories):
+        interpolated_a[:,i] = interpolants[i](ra_interp)
+
+    return ra_interp, pd.DataFrame(data=interpolated_a)
+
+N = 1024
+Ta = 30
+nbfiles = 10
+tau_c = 2000
+T = 10
+
+mu_and_sigma_file = './data/musigma_AVG_10.dat'
+musigma = np.fromfile(mu_and_sigma_file, float, -1, "")
+
+fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
+ax.set_xscale('log')
+
+for k in [0.005, 0.01]:
+    GKTL_dir_template = './data/gktl_test_{}_k_{}_T_{}'.format(N,k,Ta)
+    directories = ['{}_{}'.format(GKTL_dir_template, i) for i in range(0,nbfiles)]
+    ra_interp, df = get_interpolated_return_times(directories, Ta, T,
+                                                  k, N, nb_points=30)
+
+    df=df.div(musigma[1])
+    error = 1.96*df.std(axis=1).values
+    mean = df.mean(axis=1).values
+    line, = plt.plot(ra_interp, mean, label='$k={}$'.format(k))
+    plt.fill_between(ra_interp, mean-error, mean+error, color=line.get_color(), alpha=0.2)
+    
+
[…]

### 2019-07-10 figures/return_times_GKTL/returntimes_GKTL.py

@@ -1,0 +1,52 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+def returntimes(amax, wghts, duration):
+        """
+        Estimate the return time of an observable using AMS sampling
+        """
+        wghtstilde = wghts[np.argsort(amax)][::-1]
+        amaxtilde = np.sort(amax)[::-1]
+        ra = np.zeros(len(amax))
+        for i, a in enumerate(amaxtilde):
+                ra[i] = -duration/np.log(1-np.sum(wghtstilde[0:i+1]))
+        I = ~np.isnan(ra)
+        return amaxtilde[I], ra[I]
+
+def compute_return_times(dirname, Ta, T, k, N, tau_c=2000):
+        filename = os.path.join(dirname, 'A.datout')
+        A = np.fromfile(filename, float, -1,'')
+
+        filename = os.path.join(dirname, 'amax.datout')
+        amax = np.fromfile(filename, float, -1,'')
+        amax = amax
+
+        filename = os.path.join(dirname, 'SCGF.datout')
+        lam = np.fromfile(filename, float, -1,'')
+
+        wghts = np.array([np.exp(Ta*tau_c*lam)*np.exp(-k*A_*Ta*tau_c)/N for A_ in A])
+        return returntimes(amax, wghts, Ta-T)
+
+def compute_return_times_no_dups(dirname, Ta, T, k, N, tau_c=2000):
+        filename = os.path.join(dirname, 'A.datout')
+        A = np.fromfile(filename, float, -1,'')
+
+        filename = os.path.join(dirname, 'amax.datout')
+        amax = np.fromfile(filename, float, -1,'')
+        amax = amax
+
+        filename = os.path.join(dirname, 'SCGF.datout')
+        lam = np.fromfile(filename, float, -1,'')
+
+        unique_amax, I = np.unique(amax, return_index=True)
+        unique_A = A[I]
+        N = len(unique_amax)
+        wghts = np.array([np.exp(Ta*lam*tau_c)*np.exp(-k*A_*Ta*tau_c)/N for A_ in unique_A])
+        return returntimes(unique_amax, wghts, Ta-T)
+
+
+
+    
+    

### 2019-08-23 draft_p2.tex
Modifs Emmanuel fin juillet

--- 2019-05-27 draft_p2.tex
@@
 \begin{abstract}
-\TL{  This study evaluates the relevance of rare-event sampling type approaches to make accessible the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, which corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-  Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of events far from reach from brute force sampling approaches.
+This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+The main idea is here to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
+%
+Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events out of reach by brute-force direct sampling.
 Application to fluid-structure interaction problems remains mainly open. 
 %
-  The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-  Practical evidence is given that the fast sweeping time of fluid structures has a drastic influence on the efficiency of the sampling with both algorithms.
-  While is is shown that the \ac{ams} re-sampling does not lead to rare trajectories with significant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
+The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is here taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
+Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
+While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
 \end{abstract}
@@
 In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Indeed, such anomalous constraints can threaten the structural integrity of embedded structures.
-%
-From the viewpoint of (chaotic) dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
+Such anomalous constraints can threaten the structural integrity of embedded structures.
+%
+
+From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
 The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
-Nevertheless, this requires very long simulations since these events are also very rare.
+Simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
+However, this requires very long simulations since these events are also very rare.
 Our motivation is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
@@
 %
-\TL{Rare-event sampling refer to a large body of methods that aim at sampling preferentially regions of phase space that correspond to rare events, that would otherwise be accessed with a very low probability through a brute-force sampling.}
-%
-
-%
+{Rare-event sampling refer to a large body of methods that aim at sampling preferentially regions of phase space that correspond to rare events, that would otherwise be accessed with a very low probability through a brute-force sampling.}
+%
+
 % next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
 %
-% this paragraph would certainly need to be reshaped!
-%
 In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium.
+In that case, dynamics may be accounted for through an action that is identified as analogous to a Gibbs measure for a system at equilibrium.
 This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
@@
 %
-In this paper, a more general approach is investigated.
-\TL{Je ne comprends pas cette phrase}\EL{It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.}
-% dedicated to dynamical systems for which an action cannot be easily derived or minimized,
-%    
-%
-Even though such ideas date back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
+In this paper, a more general approach is considered for dynamical systems for which an action cannot be easily derived or minimized. 
+It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
+%    
+%
+Even though such ideas date back to the early 1950's, they have received ever-growing interest during the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
 %
@@
 The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, this algorithm has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
-
-Over the last decade, the main theoretical framework for the study of rare events (in statistical physics) has been the theory of large deviations \cite{touchette_large_2009}.
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
+
+During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
 Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
@@
 %
-\EL{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
+{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
 This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
@@
 }
-\TL{Super!}
 
@@
 %
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
+Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost, and allow for a systematic study.
 % 
@@
 They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its side, resulting in strong fluctuations of the drag acting on the square. 
+The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
 %
@@
 Further details are given in Appendix and references therein.
-In our situation, the LB method has been chosen principally for its outstanding computational efficiency.
+In our context, the LB method has been chosen principally for its computational manageability and efficiency.
 
@@
 Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
-The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalised by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The distributions are imposed at equilibrium (see Appendix).
+The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The initial distributions are imposed at equilibrium (see Appendix).
 In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$.
@@
 Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularisation procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
 
@@
 The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction, formally 
+The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
 \begin{equation}
@@
 where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
-Here, the viscous stress makes a negligible contribution to the drag, which therefore results mostly from pressure forces.
+Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
 % which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
@@
 Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can be expressed as the pressure difference 
+As a consequence, the drag can eventually be expressed as the difference 
 \begin{equation}
@@
 \end{equation}
-where $R$ is the size of the square and $U$ the averaged velocity. 
-\EL{The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.}\TL{Redondant avec le commentaire de (1)?}
+where $R$ is the size of the square and $U$ is the averaged velocity. 
+%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
 %
@@
 Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-It is observed that, typically, some vorticity that is generated near the front part of the obstacle rapidly detaches from it and is advected downstream by the flow.
-Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations mostly result from pressure variation at the front of the obstacle.
-In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.
-
-\subsection{The drag as a random process: Probability Density Function, correlation in time}
+It is observed that, typically, some vorticity  is generated near the front part of the obstacle, rapidly detaches from it, and is advected downstream by the flow.
+Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations mostly result from pressure variations at the front of the obstacle.
+In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is very different.
+
+\subsection{The drag as a random process
+	%: Probability Density Function and correlation in time
+}
 \label{sec:pdfs}
@@
 %
+% Description du signal de trainee typique
 Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
 The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-It is therefore natural to consider the drag as a random process.
-
-% Description du signal de trainee typique
-Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
-Fluctuations appear unpredictable and the signal exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-It is therefore natural to model the drag as a scalar random process $f_d(t)$.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
-  \label{fig:typical_drag_signal}
-\end{figure}
+Therefore, it is  natural to model the drag as a (scalar) random process.
 
@@
 % pdf
-Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method. 
-The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle surface.
+Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$. This long simulation will be referred to as the \textit{control run} in the following.
+It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
+The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
+%
+Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
+This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
 \begin{figure}
@@
   \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c$ is given by $C(\tau_c)=0$.
+The correlation time $\tau_c\simeq 4 \tau_0$ is obtained by the condition $C(\tau_c)=0$.
   %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
@@
 Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$ for which $C(\tau_c)=0$.
-Therefore, the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation will have important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
+It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
+One can argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+%
+This observation is important for the application of rare-event algorithms as will be discussed in section~\ref{sec:rare_events_algorithms}.
 %
 In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.
+The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}.
+The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
 
@@
 	\centering
+	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
+	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
+	One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
 	\caption{Amplitude of drag fluctuation (in standard-deviation unit) as a function the corresponding return time.
-The return time $r(a)$ is defined as the typical period of occurrence of a fluctuation of amplitude $a$.}
+	}
 	\label{fig:return_time_instant}
@@
 %
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force sampling, using the control run.
-Brute-force sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-The direct approach will provide a trustworthy baseline for the validation of rare-events algorithms. 
+The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
+Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
+It will provide a trustworthy baseline for the validation of rare-events algorithms. 
 How rare is a fluctuation $f_d \geq a$ can be quantified by its \textit{return time} $r(a)$.
-This latter is defined as the average time between two consecutive occurrences of fluctuations with amplitude $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurrence of the event: $r(a) = \mathbb{E}[\tau]$.
-Note that we assume $r(a) >> \tau_0$, and therefore consider that such events are points in time.}}
-\end{figure}
-Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time: $r(a) \gg \tau_c$.
+This latter is defined as the averaged time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
+
+Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
 In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and
@@
 %
-Fig.~\ref{fig:return_time_instant} illustrates the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$ computed from the \textit{direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
-It is found that the return time indeed diverges  exponentially with the amplitude of fluctuation.
-Let us also remark that under-sampling effects explain the deviation from the exponential law at the largest amplitudes.
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$ computed from {direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
+It is consistently found that the return time  diverges  exponentially with the amplitude of fluctuation.
+Let us also remark that under-sampling effects explain some deviation from the exponential law at the largest amplitudes.
 % typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
@@
 %
-We have extracted  the fluctuations of the drag with a return time $r(a)$ greater than  $10^4~\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
+We have extracted  the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
 This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough large-amplitude events and possibly identify generic features.
+The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
 %
 According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
-Precisely, 104 independent fluctuations \EL{excursion, terminologie correcte?} \TL{peut etre, mais fluctuation est correct et plus precis?.}with $f_d(t) \geq 7.6\sigma$ have been identified.
-For convenience, each fluctuation will be characterised in the following by its maximal value $f_d^{\star}$ and the time $t^{\star}$ at which this maximum is reached.
-%
-The phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
+Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
+Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
+%
+In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
 
@@
 	As the amplitude of the drag increases, the relative variation of the base pressure also increases. 
-\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} }
+	%\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} 
+}
 \end{figure}
@@
 %
-In section~\ref{sec:test_flow}, it has been pointed out that typical drag fluctuations  originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
+In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations were originating mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
 We shall see that the situation is different in the case of {extreme} drag fluctuations.
@@
 \end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the variations of the forebody pressure and base pressure, respectively.
+where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
 %
@@
 %
-It is found that the base pressure variation contributes typically as $80\%$ of the overall drag fluctuation.
+It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
 Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
@@
 Extreme drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.
-\TL{ $\tilde f_d$ rather than $f_d$ in figure.}}
+The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
 \end{figure}
@@
 Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width corresponding roughly to one correlation time $\tau_c$.
-This implies that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
+A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
+This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
 %Starting from typical values, extreme drags are typically reached in less than a correlation time. 
 Interestingly, the profile is also slightly skewed indicating that the step up of the drag  is slower than the return to typical values past the peak value.
+This is reminiscent of time-irreversibility in turbulent dynamics.
 %Moreover, such very high drag levels do not persist over time.
@@
   \caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.
-\TL{changed to $\tilde f_d$.}}
+In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
 \end{figure}
@@
 %
-The formation of this vortex originates from an intense negative (positive) vorticity layer at the top (bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strong drag. In contrast, nothing special happens near the forebody of the obstacle.
+The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
+%
+This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strong drag.
+In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
 
 The high pressure drop on the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it to be swept away.
+As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
 % 
@@
 Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
+This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
 This scenario is generic and has been observed for most extreme events sampled in the control run.
@@
 
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is now proposed to characterise the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square,
+Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is  proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square,
 \begin{equation}
@@
 Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.}  $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in the region related to typical values of both $\overline{\gamma}$ and $f_d$.
+Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
 On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
@@
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations in agreement with the previous description of flow scenarii.
+This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
+In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
 
@@
 	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{\TL{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the time-averaged drag $F_T$ observed in a control timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ with $T_{Tot} = 10^6\tau_c$ and $T = 10 \tau_c$.
-Here $\tilde{f}_d$ denotes the zero-mean, unit variance instantaneous drag.}}
+	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
 	\label{fig:extreme_avg}
@@
 In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-In practical applications, this duration may be much smaller than the characteristic response time of the material structure subject to these fluctuations.
+
+%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
+We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
+%
+In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
 % Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
@@
 %
-Therefore, a relevant observable is the \textit{time-averaged} drag
+Therefore, a relevant observable is the \textit{time-averaged} drag defined as
 \begin{equation}
@@
 \end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the timescale (response time) of interest. We shall consider $T=10~\tau_c$ in the following.
-
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be viewed as the overall contribution of roughly $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
+where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time). We shall consider $T=10\tau_c$ in the following.
+
+%
+During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
 % phenomenology
@@
 % What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from (case (1)) a single outstanding fluctuation of the instantaneous drag,  or from (case (2)) a series of moderate positive fluctuations.
-%
-%
-The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$ and sample a set of extreme events.
-By taking $a=5.2~\sigma_T$  with $\sigma_T$ being the standard deviation of the fluctuations of $F_T$, $84$ independent extreme events have been selected.
+It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
+%
+%
+%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
+In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
+By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
 
@@
 %
-Fig.~\ref{fig:extreme_avg} displays the timeseries $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-It is established that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are featured in Fig.~\ref{fig:extreme_avg} along with mixed situations where very large value of the averaged-drag results from both a very large fluctuation and a significant succession of moderate (positive) fluctuations of the instantaneous drag.
+Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
+It is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
+Indeed, both cases are equally featured in Fig.~\ref{fig:extreme_avg}.
+Very large value of the averaged-drag results from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
 %
@@
 %
-This observation may be explained by the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations~\cite{lestang_article_p1}\TL{reference vers l'article JSTAT} (see Fig.~\ref{fig:pdf_drag}).
-Indeed, let $X$ be a random variable with \ac{pdf} $\mathbb{P}(X)$ and standard deviation $\sigma_X$.
+This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
+Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
 Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
@@
   \label{eq:indep}
-  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left(\frac{S_N}{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
+  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
 \end{equation}
@@
 \end{equation}
+The observation is therefore well supported theoretically.
 
@@
 %
-In the limit of very rare events and complex dynamics (such as turbulent flows in industrial or environmental domains) the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of a fluctuation $f_d \geq a$ scales as $r(a) \propto e^{\lambda a}$.
+In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
+According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\lambda a}$.
 As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially as $e^{\lambda a}$.
@@
 % In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than the related return time.
+The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than their related return times.
 %
@@
 %
-Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
+Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
+Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
 In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-In a nutshell, trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable of interest---\textit{e.g.} the drag.---with a higher frequency.
+%In a nutshell
+Briefly, trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
 These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories are associated a weight, from which one is able to compute the probability and expectation values of observables. 
+The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables. 
 % These two algorithms will be detailed in the following.
@@
 Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
-On the basis of their respective maximum $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
+On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
 Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
@@
 In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories exceed a fixed threshold $Q$.
-\TL{MAJ figure et ordonnee \#1}}
+Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
 \end{figure}
@@
 %
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits the basic ideas of {splitting algorithms} \cite{KahnHarris1951}, in which the sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
+The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits the basic ideas of {splitting algorithms} \cite{KahnHarris1951}.
+The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
 %a sequence of events that can be sampled with a higher probability  
-a \TL{sequence of short excursions that are easier to sample}~\cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm \TL{relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$} and is sketched in Fig.~\ref{fig:illustr_AMS}.
+a sequence of short excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
+The operating principle is sketched in Fig.~\ref{fig:illustr_AMS}.
 %
@@
 %
-
-Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$.
+Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function during the time interval $0\leq t \leq T_a$.
 Discarded trajectories are re-sampled according to the remaining trajectories.
-\TL{For a recent review of the \ac{ams}, its history and applications, see~\cite{cerou2019adaptive}.}
-\TL{Note that we use a specific form of the the \ac{ams} in which trajectories have a fixed duration $T_a$.
-It can however be formulated in terms of the classical \ac{ams}~\cite{lestang_computing_2018} and the following results are therefore not limited to this particular variant.}
+{A very recent review of the \ac{ams} algorithm, its history and applications are provided in~\cite{cerou2019adaptive}.}
+{Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
+It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}. Next results are therefore not limited to this particular variant.}
 
@@
     \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
-The orange line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
-The solid blue line is the analytical solution for the return time of amplitude $a$.
-\TL{replaced $f_d$ with $x$ in figure} }
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
   \label{fig:comparaison_temps_de_retour}
 \end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered.
-This allow us to highlight differences with fluid dynamics.
-
+Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
 \begin{equation}
   \label{eq:ou}
-  \dot{x} = -x + \sqrt{2\epsilon}\eta (t)  
-\end{equation}
-\TL{The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Note that, for the process defined in~\eqref{eq:ou}, the correlation time is $\tau_c = 1$.}
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to typical values of $x$.
-The score function is simply $x(t)$ and only one single trajectory is re-sampled at each iteration.
+  \dot{x} = -x + \sqrt{2\epsilon}\eta (t).
+\end{equation}
+This basic stochastic process will allow us to highlight differences with fluid dynamics.
+
+The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
 %
@@
 %
-The computational cost of the algorithm after $J$ iterations is related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
 Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
 In the latter, the typical computational cost is simply the return time $r(a)$.
-Obviously, the successive re-samplings of the \ac{ams} algorithm lead very rapidly to trajectories exhibiting extreme fluctuations.
-The related computational cost is many orders of magnitude lower than that of direct sampling.
-
-Undoubtedly, the \acl{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-Indeed, the state space is one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
+Obviously, the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+The computational cost is many orders of magnitude lower than that obtained by direct sampling.
+
+Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
 In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favorable features do no a priori persist in the case of fluid dynamics.
 
@@
 %
-Our aim is now to sample flow evolution that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
+Our aim is now to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
 In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
 In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However, a natural simple choice for which we will opt for in this study is $\xi (\mathbf{x}(t),t) = f_d(t)$. 
-\TL{Let us stress here that the flow dynamics are deterministic.
-Randomness must therefore be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice we introduce a small perturbation in the state of the flow at the branching points (see figure~\ref{fig:illustr_AMS}).
-Due to the chaoticity of the flow, re-sampled trajectories will then separate from their parent.
-The timescale over which two trajectories separate due to an infinitesimal perturbation of the initial state is known as the \textit{Lyapunov timescale}.
-It is denoted $\tau_L$ in the remainder of this paper.
-We make the important assumption that such a perturbation at branching times has a negligible impact on the properties of the sampled rare events.
-The validity of this assumption can be verified by performing a brute force simulation of the dynamics over a long duration, regularly perturbing the state of the flow.
-On can eventually check that the statistics of the drag are consistent with the statistics computed from the control simulation shown in section~\ref{sec:pdfs}.}
-\EL{Freddy: provide reference for Lyapunov timescale. \#12}
+However, a natural simple choice for which we have opted, is $\xi (\mathbf{x}(t),t) = f_d(t)$. 
+%
+{Let us also stress that the dynamics is here deterministic.
+Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
+In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS}). 
+%
+Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
+% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
+We assume that this perturbation at branching times has a negligible impact on the properties of the sampled rare events.
+The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
+It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
+%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
 
@@
 %
-The \ac{ams} is applied with $N=256$ trajectories of duration $T_a = 20\tau_c$.
+The \ac{ams} algorithm is applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
 %
@@
 The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-After a few iterations, the trajectories with the lowest maxima of the score function are discarded, and new trajectories with higher maxima are re-sampled.
+After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
 These maxima are depicted by the stars in Fig~\ref{fig:AMS_drag_resampling}.
-Importantly, one can see that re-sampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.
-This contrasts with the case of the \acl{ou} process (Fig.~\ref{fig:comparaison_temps_de_retour}).
+Importantly, one can observe that the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
+This obviously contrasts with the case of the \acl{ou} process.
+%(see Fig.~\ref{fig:comparaison_temps_de_retour}).
 
@@
 Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
-  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the trajectory displaying the highest maximum in the initial ensemble.}
+  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
 \end{figure}
@@
 %
-Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  \EL{useful to mention?}\TL{Yes, good point!}
+Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  
 %
@@
 This feature may be explained as follows.
-Because the dynamic process has a memory, it takes a certain time $\tau_L$ before a re-sampled trajectory separates from its parent.
-In our situation, this memory originates from the deterministic nature of fluid dynamics.
-%
-\TL{Importantly, we observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations is shorter than the Lyapunov timescale $\tau_L$ of the drag process.}
-As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$---the maximum drag amplitude---cannot lead to larger values for $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based.
-Therefore, the probability of observing a new extreme fluctuation is also very low.
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the ``convective nature'' of the flow.
-%
-As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and therefore to the mean velocity $U$.
+It takes a certain time $\tau_L$ (called Liapunov's timescale) before a re-sampled trajectory separates from its parent.
+In our situation, this ``memory effect'' originates from the deterministic nature of fluid dynamics.
+%
+We observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations, $\tau_c$, is much shorter than the Lyapunov's timescale $\tau_L$.
+As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
+For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, therefore, the probability of observing a new extreme fluctuation is also very low.
+%
+
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
+%
+As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
 On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U < 1$ in our case of grid-generated turbulence implies that $\tau_L > \tau_c$.
+The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L \gtrsim \tau_c$.
 
@@
 %
-In summary, it appears that a straightforward application, defining the score function as the drag itself, does not allow us to efficiently sample extreme fluctuations.
-%
-Indeed, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require an intricate understanding of the flow scenario leading to extreme events, certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
+In summary, it appears that a straightforward application of the \ac{ams} algorithm with the score function being the drag itself, does not allow us to efficiently sample extreme fluctuations.
+%
+Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
+On the other hand, finding a better score function would require an intricate understanding of the flow scenario leading to extreme events.
+%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
 %
@@
 %
-A line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time \textit{before} the maximum is reached.
+A possible line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
 That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care, so as to preserve the mathematical properties (evaluation of the statistical bias) of the algorithm.
+However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
 %
@@
 The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as in section~\ref{sec:ams_drag}, \textit{i.e.} by setting $F_T$ as the score function $\xi$.
-However, this leads to similar unsatisfactory conclusions.
+The \ac{ams} algorithm could be used in the same way as before
+%in section~\ref{sec:ams_drag} \textit{i.e.} 
+by taking the time-averaged observable itself as the score function.
+However, this would lead to similar unsatisfactory results.
 % the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
@@
 Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-In this case, however, the trajectories interact dynamically. 
+% In this case, however, the trajectories interact dynamically. 
 %
@@
 %
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favours trajectories related to large values of the (time-averaged) observable of interest.
+The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the observable.
 %
 The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-\TL{Similar ideas have been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
+{Similar ideas have already been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
 The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
@@
 \label{sec:gktl_description}
-Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$ and introduce a cloning period $\tau$.
+The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%
+Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
 %
@@
 %
-\TL{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are produced at time $t=t_i$.
-See~\cite{brewer2018efficient, lestang:tel-01974316} for more information on the practical implementation of the algorithm.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling of trajectory space, for which trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
+The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
+The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+%
 The sampled biased distribution writes
@@
 where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),$ 
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
 refers formally to the probability of observing the trajectory 
@@
 The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-Note that
+%
+One can mention that
 \begin{equation}
@@
 distribution $\mathbb{P}_{0}$.
-This result relies on a \textit{mean-field approximation}
+This result relies on the \textit{mean-field approximation}
 
@@
 where $\mathbb{E}_{i}[.]$ denote the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$~\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
-%
-Pruned trajectories are discarded from the statistics.
+The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\cite{DelMoralBook,DelMoral2013}.
+%
+Rejected trajectories are discarded from the statistics.
 Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-A key feature of the \ac{gktl} algorithm is that the sampling procedure do not involve any alteration of the dynamics, and therefore all trajectories in the sampled ensemble are solutions of the dynamical system.
-This is true for stochastic dynamics.
-However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics and we proceed in the same way as as described in section~\ref{sec:ams_drag}.}
-
-%
-The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+
+A key feature of the \ac{gktl} algorithm is that the sampling procedure do not involve any alteration of the dynamics.
+All trajectories in the resampled ensemble are solutions of the original dynamical.
+%This is true for stochastic dynamics.
+{Nevertheless, it should be noted that a (small) random perturbation is introduced in the cloning procedure to allow the trajectories of two clones to separate. 
+As for the \ac{ams} algorithm, it is assumed that this perturbation does not (notably) affect the statistics of the sampled trajectories.}
+% 
+%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
+%
+%
+Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
 %
@@
 	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10~\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
+	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
 The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-	The dashed line denotes the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is now studied for the flow dynamics introduced in section~\ref{sec:test_flow} with the aim of sampling trajectories with extreme fluctuations of the \textit{time-averaged} drag.
+	The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
+}
+\end{figure}
+
+The application of the \ac{gktl} algorithm is now considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
+The purpose is to sample trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.
+{Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.}
 
 % Choice of parameters and perturbation
-The computational cost $C_{gktl}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-In practice we set $N=16384$ and $T_a = T = 10\tau_c$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-A cloning period $\tau$ that is too short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is an issue.
+A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
 On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a rule of thumb is to set $\tau \approx \tau_c$.
-In the following experiments we set $\tau = \tau_c /2$.
-%
-
-Three numerical experiments corresponding to three different values of the parameter $k$ have been performed.
+As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
+In the following experiments, $\tau = \tau_c /2$. 
+%
+
+Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
 %
@@
 %
-Fig.~\ref{fig:IS_GKTL} displays the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag, $\tilde F_T \equiv F_T - \overline{F_T}$:
+Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
 \begin{equation}
@@
 in addition to the unbiased \ac{pdf} based on the control run.
+As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
 %Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-
+%
 % limitations
 %
-For a fixed number of trajectories $N$, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-Indeed for $k \gtrsim k_{max}$, the re-sampling relies on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is over-peaked\TL{Anglais?}.
+Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
+%
+For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
 %
@@
 %
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$, addressed in recent studies~\cite{nemoto2017finite,Guevara_Hidalgo_2018}.
-
+An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
+This issue is addressed in recent studies \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
+%
 % how to assess the value of k_max ?
 In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity may be monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
+This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
 
@@
 	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,~...$ captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as one.
-		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realisations of $F_{T}$ of the same duration.}
+	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
+		Note that trajectories that overlap for more than half of their duration are counted as only one.
+		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
 	\label{tbl:nbTraj_tbl}
@@
 %
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to brute-force sampling. 
+The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
 Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
+Note that trajectories that overlap for more than half of their duration are counted  as only one.
 %\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations one can expect from simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$, \emph{i.e.} with the same computational cost $C_{gktl}$.
+Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
 %
 %This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-
+%
 %\begin{figure}
@@
 %\end{figure}
-
+%
 % Conclusion
 %
-Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach from a brute-force sampling approach of similar computational cost.
-In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
+
+%
+% PARAGRAPHE SUIVANT INUTILE ?
+%
+%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
 By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
@@
 
-\EL{rajouter le calcul des temps de retour par gktl pour etoffer un peu la partie du papier sur gktl -- il faudrait montrer un resultat qui montre que l'on obtient la bonne statistique} \TL{Demande pas mal de travail en comparaison des autres points.
-Voir issue \#10.}
+\subsubsection{Computation of return times}
+\label{sec:return_times}
+Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times  for
+events unreachable by a direct approach.
+%(see Fig.~\ref{fig:illustrate_return_time})
+
+In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
+Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
+\begin{equation}
+\label{eq:time_averaged}
+F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
+\end{equation}
+and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
+\begin{equation}
+r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
+\end{equation}
+
+The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
+\begin{equation}
+\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
+\end{equation}
+with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
+
+Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
+It compares return times obtained for two values of the bias $k$ and a fixed computational cost $N \times T_a$ to the return times obtained by a direct approach applied to a single timeseries of duration corresponding to $T_{tot} = N \times T_a$.
+While a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+This is  obviously a major advantage of this rare-event sampling algorithm.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
+	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle. $\tilde{F}_T$ denotes the time-averaged drag with zero mean.
+		The blue and read lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
+The green line is the return times obtained from a single timeseries of duration $T_{tot} = N\times T_a$.
+Uncertainty ranges are shown.}
+\end{figure}
+
+%\subsubsection{Discussion}
+%\label{sec:discussion_gktl}
+%
+%%
+%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+%By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
+%However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
+%Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
 
@@
 \label{conlusion}
-We assessed the application of two rare event algorithms for the numerical simulation of extreme mechanical efforts on structures immersed in a turbulent flow, a situation relevant to many industrial applications.
-\\
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of very long simulations---without rare event algorithm---we observed that such extreme events are caused by the concentration of vorticity very close to the base of the square.
-Because of the sweeping of the fluid structure responsible for such a situation, extreme drag levels do not persist over time.
-The lifetime of extreme drag fluctuations is of the order of the turnover time, and the corresponding drag signal is very peaked around the extreme value.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are very well described by an exponential \ac{pdf}.
-This property can be linked to the phenomenology of extremes of the time-averaged drag.
-Indeed, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations or an exceptional succession of typical fluctuations that
-add up forming an extremely large value of the average.
+In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
+
+In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square placed in a turbulent channel flow, in two dimensions.
+By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
+Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
+%
+The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
+Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are  well described by an exponential \ac{pdf}.
+This property is linked to the phenomenology of extrema of the time-averaged drag.
+Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that
+pile up to yield an large value of the average.
 Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-\\
-The application of rare event algorithms relies on the definition of a score function, on which the selection of trajectories is based upon.
-Although the dynamics itself is a black-box from the point of view of the algorithm, the efficiency of the sampling depends on the phenomenology.
-For complex dynamics, including turbulent flows, the choice of the score function is made difficult by the very little knowledge one has of the rare events of interest.
-This is, after all, our motivation for the simulation of these events.
+
+The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
+%Although the dynamics itself is a black-box from the point of view of the algorithm, 
+The efficiency of the algorithm depends on the phenomenology of the dynamics and, therefore, on the choice of the score function.
+For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterization of rare events.
+These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
+In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
+%This is, after all, our motivation for the simulation of these events.
 \\
 On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling approach.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent following the
-addition of a small perturbation in the initial conditions.
-The \ac{gktl} has been applied to the sampling of trajectories displaying extreme fluctuations of the time averaged drag.
-In this case, we showed that using the \ac{gktl} leads to a tremendous improvement with respect to a brute-force approach, allowing the simulation and computation of statistics for trajectories that are out of reach without algorithm.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in a an extreme value for the average.
+In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
+This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent. 
+%following the addition of a small perturbation in the initial conditions.
+The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
+In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
+The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
 It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-\\
-Successful usage of the \ac{ams}, \ac{gktl} and similar methods for complex flows relevant to industrial or environmental situations will require coping with the fact that quality score functions
-are very difficult to identify, if even possible.
-A popular direction in current research involves taking advantage of the recent advances in learning methods to compute score functions beforehand.
+This issue is postponed for future investigations.
+
+Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
+are difficult to identify, if even possible.
+A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimize score functions beforehand.
 
@@
 \] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. Einstein summation convention is assumed.
-Finally, let us mention that this discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
+This discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
 $\Delta x \to 0$.

### 2019-08-23 draft_p2.tex
Modifs Freddy debut aout 2019

@@
 This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is here to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events out of reach by brute-force direct sampling.
-Application to fluid-structure interaction problems remains mainly open. 
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is here taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
+The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
+%
+Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events out of reach by direct sampling.
+This work is the first application to fluid-structure interaction problems. 
+%
+The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
 Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
@@
 The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
+The first use of the AMS algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
+More recently is has been applied to rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
 
@@
   \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is obtained by the condition $C(\tau_c)=0$.
+The correlation time $\tau_c\simeq 4 \tau_0$ is defined by the condition $C(\tau_c)=0$.
   %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
@@
 It will provide a trustworthy baseline for the validation of rare-events algorithms.
-How rare is a fluctuation $f_d \geq a$ can be quantified by its \textit{return time} $r(a)$.
-This latter is defined as the averaged time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-
-Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and
-
+
+{\bf The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
+The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
+As soon as the typical waiting times are much larger than than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential: $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$, where $r(a)=1/\lambda(a)$ is the average waiting time \cite{lestang_computing_2018}.
+$r(a)$ is called the {\it return time} of the level $a$.
+For systems without multistability, it is often the case that the mixing time $\tau_m$ is of the order of the correlation time $\tau_c$.
+
+How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
+We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
+
+The shape of the return time $r$ as a function of the return level $a$ is system dependent.
+It is often the case that  $r(a) = t(a)/\mathbb{P}(f_d\geq a)$, where the level dependent time scale $t(a)$, expected to be of order $\tau_c$, changes much more slowly with $a$ that the PDF ${\mathbb{P}(f_d\geq a)}$, for large values of $a$.
+As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag}), we might expect that
 \begin{equation}
  \label{eq:return_time}
-  r(a) \underset{a\to\infty}{\propto} \frac{1}{\mathbb{P}(f_d\geq a)} \propto e^{\lambda a}
-\end{equation}
-where $\lambda$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-%
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$ computed from {direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
-It is consistently found that the return time  diverges  exponentially with the amplitude of fluctuation.
-Let us also remark that under-sampling effects explain some deviation from the exponential law at the largest amplitudes.
+r(a) \underset{a\to\infty}{\propto} \exp(-la)
+\end{equation}
+where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
+It is consistently found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
+Let us also remark some deviation from the exponential law at the largest levels, that are probably the consequence of undersampling.
+}
 % typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
@@
 %
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the sampled extreme events.
+Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
 Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
@@
 \end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time). We shall consider $T=10\tau_c$ in the following.
+where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time). Relevant values of $T$ are application dependant. We shall consider $T=10\tau_c$ in the following.
 
@@
 Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-
 \begin{equation}
@@
 In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\lambda a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially as $e^{\lambda a}$.
+According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}$.
+As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially as $e^{l a}$.
 %% objective
@@
 Obviously, the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-The computational cost is many orders of magnitude lower than that obtained by direct sampling.
+For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
 
@@
 % The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-We assume that this perturbation at branching times has a negligible impact on the properties of the sampled rare events.
+We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
 The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
@@
 It takes a certain time $\tau_L$ (called Liapunov's timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' originates from the deterministic nature of fluid dynamics.
+In our situation, this ``memory effect'' originates from the fact that the score function is of dimension smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
 %
@@
 Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require an intricate understanding of the flow scenario leading to extreme events.
+On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
 %which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
@@
 
-A key feature of the \ac{gktl} algorithm is that the sampling procedure do not involve any alteration of the dynamics.
+A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics.
 All trajectories in the resampled ensemble are solutions of the original dynamical.
@@
 The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is an issue.
+We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practise.
 A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
@@
 %Although the dynamics itself is a black-box from the point of view of the algorithm, 
-The efficiency of the algorithm depends on the phenomenology of the dynamics and, therefore, on the choice of the score function.
+The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
 For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterization of rare events.

### 2019-08-27 draft_p2.tex
Premiere partie des commentaires de Freddy

@@
 However, this requires very long simulations since these events are also very rare.
-Our motivation is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
+
+Our motivations is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
 %
@@
 between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.
+Pressure fluctuations are related to the dynamics of the vorticity field, as regions of strong vorticity correspond to strong local pressure gradients.
+Indeed, considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
 
@@
 Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-It is observed that, typically, some vorticity  is generated near the front part of the obstacle, rapidly detaches from it, and is advected downstream by the flow.
-Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations mostly result from pressure variations at the front of the obstacle.
-In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is very different.
+Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field
+in the vicinity of the base is only slightly perturbed.
 
@@
 The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential: $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$, where $r(a)=1/\lambda(a)$ is the average waiting time \cite{lestang_computing_2018}.
+As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential: $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$, where $r(a)=1/\lambda(a)$ is the average waiting time \cite{lestang_computing_2018}.
 $r(a)$ is called the {\it return time} of the level $a$.
@@
 It is consistently found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also remark some deviation from the exponential law at the largest levels, that are probably the consequence of undersampling.
+Let us also point out some deviation from the exponential law at the largest levels, that are probably the consequence of undersampling.
 }
@@
 \begin{equation}
-  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star}
-\end{equation}
-where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
+  \tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
+\end{equation}
+where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
 %
 Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
+$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
 %
@@
 The observation is therefore well supported theoretically.
+See~\cite{lestang2019JSTAT} for further details.
 
@@
 Discarded trajectories are re-sampled according to the remaining trajectories.
-{A very recent review of the \ac{ams} algorithm, its history and applications are provided in~\cite{cerou2019adaptive}.}
+See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
 {Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
@@
   \label{eq:ou}
-  \dot{x} = -x + \sqrt{2\epsilon}\eta (t).
-\end{equation}
+  \dot{x} = -x + \eta (t),
+\end{equation}
+where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
 This basic stochastic process will allow us to highlight differences with fluid dynamics.
@@
 Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-It compares return times obtained for two values of the bias $k$ and a fixed computational cost $N \times T_a$ to the return times obtained by a direct approach applied to a single timeseries of duration corresponding to $T_{tot} = N \times T_a$.
+The figure shows two independant estimates, obtained using different values of the bias parameter $k$.
+Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
+In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
 While a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
@@
 	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle. $\tilde{F}_T$ denotes the time-averaged drag with zero mean.
+	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
+$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
 The blue and read lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration $T_{tot} = N\times T_a$.
-Uncertainty ranges are shown.}
+The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
+Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
+Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
 \end{figure}

### 2019-08-27 draft_p2.tex
Modifie l'intro et ajoute references

@@
 However, this requires very long simulations since these events are also very rare.
-
-Our motivations is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
-%
-
 % general def on rare-event sampling %
@@
 %
+Our motivation is to conduct the first computational study of extreme mechanical efforts acting on a immersed body, using both very long timeseries and rare event sampling techniques.
+%
 
@@
 In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In that case, dynamics may be accounted for through an action that is identified as analogous to a Gibbs measure for a system at equilibrium.
-This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on \emph{instantons} \cite{grafke2015instanton}.
-For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}.
-Therefore, suitable numerical schemes can  be used to minimise the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
-%
-An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimised.
-%
-%
-%Whilst Burgers' turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or environmental flows.
-%
-In this paper, a more general approach is considered for dynamical systems for which an action cannot be easily derived or minimized. 
+In that case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
+%
+An alternative approach is based on \emph{instantons} \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
+Instantons refer to the most probable trajectories (in phase space) that achieve a given rare event, in the limit of weak noise.
+Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
+An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
+%
+A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
+%
+
+%
+In this paper, a more general approach is considered for complex, potentially deterministic dynamical systems.
 It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
@@
 Even though such ideas date back to the early 1950's, they have received ever-growing interest during the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
-%
-% already mentioned in the beginning? This sentence is misleading.
-%
-%However, a large body of algorithms are limited to systems at equilibrium, for which the stationary measure can be sampled by means of Monte Carlo methods. 
+More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
+%
+
 %
@@
 
-The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
+The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
 In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the AMS algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
-More recently is has been applied to rare events of re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.
+The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
+More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.
 
@@
 %
-{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
+{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
 This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.
-}
+Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.}
+%
+
+{The first part of this paper presents a study of the extreme fluctuations of the drag force acting on a square in a two dimensional turbulent channel flow.
+  This study is based on the simulation of the flow over a very long duration, made possible by the simplicity of the flow.}
+%
+The motivation for this study is twofold.
+First, we provide a description of the statistics and dynamics related to extreme drag fluctuations.
+To the author's knowledge, such a description does not exist in the literature.
+Second, this study provides reference results that can be used to validate results obtained using
+rare event algorithms.
+
+% Annonce du plan
+Section~\ref{sec:test_flow} introduces the flow set-up and describes the dynamics related to typical
+drag fluctuations.
+The statistical properties of the drag are then discussed.
+In section~\ref{sec:direct_sampling} we discuss the phenomenology of extreme fluctuations of extreme
+drag fluctuations, based on a direct sampling approach.
+Both the instantaneous drag and time-averaged drag are considered.
+We find that sampled extreme events for the instantaneous drag display similar dynamics and we highlight that the phenomenology of extremes for the time-averaged drag can be related to the statistical properties of the instantaneous drag.
+This is supported by the study of extreme events for simple stochastic dynamics.
+Next, section~\ref{sec:rare_events_algorithms} describes the application of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, using the same flow set-up.
+In section~\ref{sec:ams} we show that the application of the \ac{ams} to such a flow is not straightforward and connect the difficulties to the phenomenology described in the previous sections.
+Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
+Its application allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
+Finally, we use the \ac{gktl} algorithm to compute the return times of extreme fluctuations of the
+time-averaged drag acting on the immersed obstacle.
 

### 2019-08-27 draft_p2.tex
Add a few references

@@
 See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
+For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
 {Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
@@
 Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS}). 
-%
-Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
+In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
+%
+Due to the chaotic behavior of the flow, re-sampled trajectories then sparate from their parents.
 % The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.

### 2019-08-27 figures/pressure_ratio/plot_pressure_ratio.py
Modifies labels for figure pressure_ratio

@@ -38,10 +38,10 @@ maxDrag = (maxDrag-mu) / sig
 fig,ax = plt.subplots(figsize=(9,5), constrained_layout=True)
 
 
-lineRear, =ax.plot(maxDrag, rearContrib, marker='o', label=r'$-\tilde{p}^{\star}_{base}/\tilde{f}^{\star}_d$')
+lineRear, =ax.plot(maxDrag, rearContrib, marker='o', label=r'$-\Delta p^{\star}_{base}/\tilde{f}^{\star}_d$')
 lineRear.set_linestyle('None')
 
-lineFront, =ax.plot(maxDrag, frontContrib, marker='^', label=r'$\tilde{p}^{\star}_{fb}/\tilde{f}^{\star}_d$')
+lineFront, =ax.plot(maxDrag, frontContrib, marker='^', label=r'$\Delta p^{\star}_{fb}/\tilde{f}^{\star}_d$')
 lineFront.set_linestyle('None')
 
 coefFront = np.polyfit(maxDrag, frontContrib, 1);

### 2019-08-27 figures/IS_GKTL/IS_GKTL.py
Fix figures IS_GKTL, illutrate_return_time and pressure_ratio #18 #19 #20

@@ -38,7 +38,7 @@ ax.set_yticks(ax.get_yticks()[::2]) # Plot only half of ticks
 ax.set_xticklabels(['$-2\\sigma_T$','$0$','$2\\sigma_T$','$4\\sigma_T$','$6\\sigma_T$'])
 ax.tick_params(axis='both', which='major', labelsize=16)
 
-ax.set_xlabel('$\\tilde{F}_T$', fontsize=22)
+ax.set_xlabel('$F$', fontsize=22)
 ax.set_ylabel('$\\rho_{k}(\\tilde{F}_T = F)$', fontsize=22)
 ax.legend(loc='best', fontsize=16)
 

### 2019-08-27 figures/illustrate_return_time/illustrate_return_time.py

@@ -75,7 +75,7 @@ for i, tup in enumerate(zip(point_array, point_array[1:])):
     x_point, x_point_fwd = tup
     plt.annotate('', xy=(x_point+eps, y0), xytext=(x_point_fwd-eps, y0),
                  arrowprops=props)
-    plt.text((x_point+x_point_fwd)/2-dx0, y0+dy0, '$\\tau_{}$'.format(i), fontsize=14)
+    plt.text((x_point+x_point_fwd)/2-dx0, y0+dy0, '$\\tau_{}$'.format(i+1), fontsize=14)
 
 # Plot the letter a at the left hand side of the plot slightly above
 # the threshold line

### 2019-08-27 draft_p2.tex
Modifies legend caption for figure 6
Modifies comment and figure AMS_drag_resampling

@@
 	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuation (in standard-deviation unit) as a function the corresponding return time.
+	\caption{Amplitude of drag fluctuations as a function the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
 	}
@@
 After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-These maxima are depicted by the stars in Fig~\ref{fig:AMS_drag_resampling}.
-Importantly, one can observe that the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-This obviously contrasts with the case of the \acl{ou} process.
+Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
+The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
+Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in  figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
 %(see Fig.~\ref{fig:comparaison_temps_de_retour}).

### 2019-08-27 figures/AMS_drag_resampling/make_plot_256.py

@@ -42,7 +42,7 @@ for j in range(0,Nc):
 # Plots a line that represents the level of the highest fluctuation in
 # the initial ensemble
 M = np.amax(maxArray)
-line, = ax.plot([Cvec[40], Cvec[Nresamp-1]], [M, M], '--')
+line, = ax.plot([Cvec[40], Cvec[Nresamp-1]], [M, M], '--', color=color_list[1])
 # Plots the maxima for Cvec = Nc*Ta (initial computational cost)
 init, = ax.plot(Nc*Ta*np.ones(Nc), maxArray, '+', linestyle='None')
 init.set_markersize(5)
@@ -53,7 +53,7 @@ init.set_label('Maxima over initial trajectories')
 resamp, = ax.plot(Cvec, amax, linestyle='None')
 resamp.set_marker('o')
 resamp.set_markersize(3)
-resamp.set_markerfacecolor(color_list[1]);
+resamp.set_markerfacecolor(color_list[0]);
 resamp.set_markeredgecolor(resamp.get_markerfacecolor())
 resamp.set_label('Maxima over resampled trajectories')
 

### 2019-08-27 draft_p2.tex
Take out bold font for Freddy's part on return times

@@
 
-{\bf The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
+The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
 The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
@@
 Let us also point out some deviation from the exponential law at the largest levels, that are probably the consequence of undersampling.
-}
+
 % typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.

### 2019-09-10 draft_p2.tex
Modifs Freddy 27/08

@@
 %
-In this paper, a more general approach is considered for complex, potentially deterministic dynamical systems.
+In this paper, a more general approach is considered.
 It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
+It is suited for any Markov dynamics, for example stochastic ones including diffusions or stochastic differential equations, but also potentially for chaotic deterministic dynamical systems, for example turbulent flow equations as studied in this work.
 %    
@@
 Even though such ideas date back to the early 1950's, they have received ever-growing interest during the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
+More importantly, these types of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
 %
@@
 %
-{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
+{Recently, the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}.
 This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.

### 2019-09-10 draft_p2.tex
Modifs Emmanuel 2/9

--- 2019-08-27 draft_p2.tex
@@
 %
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events out of reach by direct sampling.
+Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
 This work is the first application to fluid-structure interaction problems. 
@@
 While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
+%
+Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme drag events of a turbulent flow on a bluff body.
 \end{abstract}
@@
 The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
-Simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
+Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
 However, this requires very long simulations since these events are also very rare.
@@
 %
-{Rare-event sampling refer to a large body of methods that aim at sampling preferentially regions of phase space that correspond to rare events, that would otherwise be accessed with a very low probability through a brute-force sampling.}
-%
-Our motivation is to conduct the first computational study of extreme mechanical efforts acting on a immersed body, using both very long timeseries and rare event sampling techniques.
+{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+%
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
 %
@@
 An alternative approach is based on \emph{instantons} \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories (in phase space) that achieve a given rare event, in the limit of weak noise.
+Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
 Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
@@
 %
-In this paper, a more general approach is considered for complex, potentially deterministic dynamical systems.
+In this paper, a more general approach is considered for complex, possibly deterministic dynamical systems.
 It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
@@
 %
-
-%
-An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+%
+An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
 %Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
@@
 
-{The first part of this paper presents a study of the extreme fluctuations of the drag force acting on a square in a two dimensional turbulent channel flow.
-  This study is based on the simulation of the flow over a very long duration, made possible by the simplicity of the flow.}
-%
-The motivation for this study is twofold.
-First, we provide a description of the statistics and dynamics related to extreme drag fluctuations.
-To the author's knowledge, such a description does not exist in the literature.
-Second, this study provides reference results that can be used to validate results obtained using
-rare event algorithms.
+{The paper is organized in two parts.
+The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
+This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
+%
+The motivation is twofold.
+Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
+This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
+Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
+This assessment is developed in the second part of the paper.
 
 % Annonce du plan
-Section~\ref{sec:test_flow} introduces the flow set-up and describes the dynamics related to typical
-drag fluctuations.
+In next section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
+drag fluctuations is described.
 The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling} we discuss the phenomenology of extreme fluctuations of extreme
-drag fluctuations, based on a direct sampling approach.
+In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
+drag fluctuations is investigated based on a direct sampling approach.
 Both the instantaneous drag and time-averaged drag are considered.
-We find that sampled extreme events for the instantaneous drag display similar dynamics and we highlight that the phenomenology of extremes for the time-averaged drag can be related to the statistical properties of the instantaneous drag.
-This is supported by the study of extreme events for simple stochastic dynamics.
-Next, section~\ref{sec:rare_events_algorithms} describes the application of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, using the same flow set-up.
-In section~\ref{sec:ams} we show that the application of the \ac{ams} to such a flow is not straightforward and connect the difficulties to the phenomenology described in the previous sections.
+It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
+Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
+This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
+Next, section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
+In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
+This difficulty is put in perspective with the phenomenology developed in the previous sections.
 Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-Its application allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-Finally, we use the \ac{gktl} algorithm to compute the return times of extreme fluctuations of the
-time-averaged drag acting on the immersed obstacle.
+This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
+As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
+Perspectives and conclusion finalize this work.
 
@@
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details are given in Appendix and references therein.
-In our context, the LB method has been chosen principally for its computational manageability and efficiency.
+Further details about the LB method are given in Appendix and references therein.
+In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
 
@@
 Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
 The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$.
+In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
 The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method.
@@
 between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity field, as regions of strong vorticity correspond to strong local pressure gradients.
-Indeed, considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
+Pressure fluctuations are related to the dynamics of the vorticity field with regions of strong vorticity corresponding to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
+%
+% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
+%
 
@@
   \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is defined by the condition $C(\tau_c)=0$.
+The correlation time $\tau_c\simeq 4 \tau_0$ is defined by $C(\tau_c)=0$.
   %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
@@
 It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-One can argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as will be discussed in section~\ref{sec:rare_events_algorithms}.
+One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+%
+This observation is important for the application of rare-event algorithms as it  will be discussed in section~\ref{sec:rare_events_algorithms}.
 %
@@
 The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential: $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$, where $r(a)=1/\lambda(a)$ is the average waiting time \cite{lestang_computing_2018}.
-$r(a)$ is called the {\it return time} of the level $a$.
-For systems without multistability, it is often the case that the mixing time $\tau_m$ is of the order of the correlation time $\tau_c$.
+As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
+For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
 
@@
 We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-
-The shape of the return time $r$ as a function of the return level $a$ is system dependent.
-It is often the case that  $r(a) = t(a)/\mathbb{P}(f_d\geq a)$, where the level dependent time scale $t(a)$, expected to be of order $\tau_c$, changes much more slowly with $a$ that the PDF ${\mathbb{P}(f_d\geq a)}$, for large values of $a$.
-As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag}), we might expect that
+%
+%
+%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
+If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
+% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
+one might expect that 
 \begin{equation}
@@
 It is consistently found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, that are probably the consequence of undersampling.
+Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of undersampling.
 

### 2019-10-19 figures/return_times_GKTL/plot_return_times_GKTL.py
Swith from `next` to `linear` in interpolation for return times

@@ -14,7 +14,7 @@ def get_interpolated_return_times(directories, Ta, T, k, N, nb_points=10):
         a, ra = rtt.compute_return_times_no_dups(dirname, Ta,
                                                  T, k, N)
         rmax[i] = np.amax(ra); rmin[i] = np.amin(ra)
-        interpolants.append(interpolate.interp1d(ra, a, kind='next', fill_value='extrapolate'))
+        interpolants.append(interpolate.interp1d(ra, a, kind='slinear', bounds_error=False))
     ra_interp = np.logspace(np.log(np.amin(rmin))/np.log(10),
                             np.log(np.amax(rmax))/np.log(10), nb_points)
     

### 2019-10-19 draft_p2.tex
Add aknowledgements

@@
 \usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath,amssymb}
@@
 
+\section{Acknowledgements}
+The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
+T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
+Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I)
+The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
+
 \appendix*

### 2019-10-19 draft_p2.tex
Add Sarah's comments for intro and proofread whole paper

@@
 %
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme drag events of a turbulent flow on a bluff body.
+Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
 \end{abstract}
@@
 %
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
+Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
+Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
 ``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
@@
 From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
-However, this requires very long simulations since these events are also very rare.
+The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
+Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
+However, this requires very long simulations since these events are rare.
 % general def on rare-event sampling %
@@
 %
-In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In that case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on \emph{instantons} \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
+In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
+In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
+%
+An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
 Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
@@
 %
-In this paper, a more general approach is considered for complex, possibly deterministic dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950's, they have received ever-growing interest during the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
+In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
+It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
+%    
+%
+Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
 More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
@@
 
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
+The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
 In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
@@
 During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
-Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
+Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
+Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
+%
+{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
 This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.}
+Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
 %
@@
 %
-The motivation is twofold.
+Motivation for this study is twofold.
 Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
@@
 % Annonce du plan
-In next section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
+In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
 drag fluctuations is described.
@@
 This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
-Next, section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
+Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
 In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
@@
 As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion finalize this work.
+Perspectives and conclusion end this work.
 
@@
 %
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost, and allow for a systematic study.
+Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
 % 
@@
 %
-The flow dynamics is integrated by the lattice Boltzmann (LB) method in our numerical simulations. 
-While traditional methods in \emph{computational fluid dynamics} rely on a discretization of the Navier-Stokes equations, the LB method considers the fluid at a mesoscopic level.
+The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
+While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
 Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
@@
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the LB method are given in Appendix and references therein.
+Further details about the \ac{lbm} are given in Appendix and references therein.
 In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
@@
 The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The squared obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
+The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
 The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
@@
 %
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure \cite{lbm_book}.  
-%
-Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
+No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
+%
+Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
 The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
@@
 In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method.
+The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
 Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
@@
 	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
+	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
 	The vorticity is given in lattice units.}
@@
 \end{equation}
-between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity field with regions of strong vorticity corresponding to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
+between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
+Pressure fluctuations are related to the dynamics of the vorticity fiel.
+Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
 %
@@
 \end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity. 
+where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
 %The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
@@
 Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field
-in the vicinity of the base is only slightly perturbed.
+Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
 
@@
 % Description du signal de trainee typique
-Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
+Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
 The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
@@
 % pdf
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$. This long simulation will be referred to as the \textit{control run} in the following.
+Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$.
+This long simulation will be referred to as the \textit{control run} in the following.
 It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
@@
 %
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
+In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
+The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
 
@@
 where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}.
-It is consistently found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of undersampling.
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
+Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
+Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
 
@@
 %
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation.
+According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
 Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
@@
 %
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations were originating mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
+In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
 We shall see that the situation is different in the case of {extreme} drag fluctuations.
@@
 %
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme drag event.
+Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
 The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
@@
 The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme drag events exhibit a typical lifetime of one correlation time $\tau_c$.
+Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
 The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
@@
 The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%
+%+
 The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
 %
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and, therefore, a strong drag.
+This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
 In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
 
-The high pressure drop on the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
+The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
 As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
 % 
-This scenario is better evidenced in Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
+This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
 %
@@
 This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-When the blocking vortex is in turn advected downstream, the vortex against the base is released.
+As the blocking vortex is in turn advected downstream, the vortex against the base is released.
 %
@@
 	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to an event.
+Each trajectory corresponds to a single event.
 The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
@@
 
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is  proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square,
+Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
 \begin{equation}
@@
 
-%The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
-In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
-This viscous shear is induced by a large vortex detached from the obstacle.
-For a more detailed description of such dynamics, see chapter 3 of~\cite{lestang:tel-01974316}.
-
 \subsection{Extreme fluctuations of the time-averaged drag }
@@
 \end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time). Relevant values of $T$ are application dependant. We shall consider $T=10\tau_c$ in the following.
+where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
+Relevant values of $T$ are application dependant.
+We shall consider $T=10\tau_c$ in the following.
 
@@
 In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially as $e^{l a}$.
+According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like $r(a) \propto e^{l a}$.
+As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
 %% objective
@@
 %In a nutshell
-Briefly, trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
+Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
 These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
@@
 %
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits the basic ideas of {splitting algorithms} \cite{KahnHarris1951}.
+The \acl{ams} algorithm~\cite{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \cite{KahnHarris1951}.
 The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into 
 %a sequence of events that can be sampled with a higher probability  
-a sequence of short excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+a sequence of shorter excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
 The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
@@
 {Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}. Next results are therefore not limited to this particular variant.}
+  It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}.
+  The following results are not limited to this particular variant.}
 
@@
 In the latter, the typical computational cost is simply the return time $r(a)$.
-Obviously, the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
 For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
@@
 In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features do no a priori persist in the case of fluid dynamics.
+Such favorable features do no \textit{a priori} persist in the case of fluid dynamics.
 
@@
 In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However, a natural simple choice for which we have opted, is $\xi (\mathbf{x}(t),t) = f_d(t)$. 
+However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.
 %
@@
 %
-Due to the chaotic behavior of the flow, re-sampled trajectories then sparate from their parents.
+Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
 % The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
@@
 This feature may be explained as follows.
-It takes a certain time $\tau_L$ (called Liapunov's timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
+It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
+In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
 %
@@
 As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, therefore, the probability of observing a new extreme fluctuation is also very low.
+For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and therefore the probability of observing a new extreme fluctuation is also very low.
 %
@@
 %
-In summary, it appears that a straightforward application of the \ac{ams} algorithm with the score function being the drag itself, does not allow us to efficiently sample extreme fluctuations.
+In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
 %
@@
 %
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the replica.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the observable.
+At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
+%
+The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
 %
@@
 The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
-Over the past ten years, it has successfully been applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
-% It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
-% averaging time~\cite{ref_large_dev}.
-% By contrast with the \ac{ams} and \ac{ams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
-% Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
-% The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
-% The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.
+Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
 
@@
 This result relies on the \textit{mean-field approximation}
-
 \begin{equation}
@@
 \end{equation}
-where $\mathbb{E}_{i}[.]$ denote the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
 The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\cite{DelMoralBook,DelMoral2013}.
@@
 %This is true for stochastic dynamics.
-{Nevertheless, it should be noted that a (small) random perturbation is introduced in the cloning procedure to allow the trajectories of two clones to separate. 
-As for the \ac{ams} algorithm, it is assumed that this perturbation does not (notably) affect the statistics of the sampled trajectories.}
+{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to allow clones of a same trajectory to separate. 
+As for the \ac{ams} algorithm, it is assumed that this perturbation does not significantly affect the statistics of the sampled trajectories.}
 % 
@@
 The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practise.
+We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practice.
 A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
@@
 %
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories (in the biased ensemble) overlap. 
+For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
 This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
@@
 Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted  as only one.
+Note that trajectories that overlap for more than half of their duration are counted as one.
 %\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
@@
 In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-While a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
 This is  obviously a major advantage of this rare-event sampling algorithm.
@@
 $\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and read lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
+The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
 The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
@@
 
-%\subsubsection{Discussion}
-%\label{sec:discussion_gktl}
-%
-%%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-%By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-%However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-%Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
 \section{Conclusion}
@@
 
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square placed in a turbulent channel flow, in two dimensions.
+In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
 By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
@@
 Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extrema of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that
-pile up to yield an large value of the average.
+This property is linked to the phenomenology of extremes of the time-averaged drag.
+Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to yield an large value of the average.
 Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
@@
 The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-%Although the dynamics itself is a black-box from the point of view of the algorithm, 
 The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterization of rare events.
+For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
 These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
 In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-%This is, after all, our motivation for the simulation of these events.
-\\
+
 On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
@@
 are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimize score functions beforehand.
+A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
 
@@
 T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I)
+Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
 The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
@@
 % details about LBM
-In the LB method, the fluid is viewed as populations of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
@@
 The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The LB scheme advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
 Namely, an \emph{exact} streaming step
@@
 with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete LB scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
+This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
 $\Delta x \to 0$.

### 2020-01-07 jfm/draft_p2_jfm.tex
Add jfm styling

--- 2019-10-19 draft_p2.tex
@@
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage,preprint]{revtex4-1}
+\documentclass{jfm}
 \pdfoutput=1
@@
 \usepackage{caption}
-\graphicspath{{figures/}}
+\graphicspath{{../figures/}}
 
@@
 
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
+\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+
+\author{Thibault Lestang\aff{1}\aff{2}
+  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
+  Freddy Bouchet\aff{1}
+  \and Emmanuel L\'evêque\aff{2}}
+\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
+\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
 \begin{document}
 
-\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-
-\author{Thibault Lestang}
-\email{thibault.lestang@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\author{Freddy Bouchet}
-\email{freddy.bouchet@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\author{Emmanuel Lévêque}
-\email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
+\maketitle
 
@@
 
-\maketitle
-
 \section{Introduction}
@@
 % general comments on physical problem %
-%
+
 Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
+Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
 ``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
@@
 
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
+From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
 The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
@@
 %
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
+In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
+In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
+%
+An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
 Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
+Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
+An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
 %
@@
 %
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
+Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
+More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
 %
@@
 
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
-More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
-Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
+The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
+The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
+More recently is has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+
+During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
+Alongside numerical methods have been developed to sample rare events \citep{moral_feynman-kac_2004}. 
+Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
+%
+{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
 This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
@@
 %
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
+No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
 %
@@
 Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
 
@@
 The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
+As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
 For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
@@
 where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
 Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
@@
 The observation is therefore well supported theoretically.
-See~\cite{lestang2019JSTAT} for further details.
+See~\citep{lestang2019JSTAT} for further details.
 
@@
 %
-The \acl{ams} algorithm~\cite{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \cite{KahnHarris1951}.
+The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \citep{KahnHarris1951}.
 The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into 
 %a sequence of events that can be sampled with a higher probability  
-a sequence of shorter excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
+a sequence of shorter excursions \citep{glasserman_multilevel_1999,rolland_statistical_2015}.
 The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
@@
 %
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
+% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\citep{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\citep{rolland_extremely_2018}.
 %
@@
 Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
+See~\citep{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
+For recent applications of the \ac{ams} in the context of fluid turbulence, see \citep{bouchet2019rare,rolland_extremely_2018}.
 {Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-  It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}.
+  It can however be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
   The following results are not limited to this particular variant.}
@@
   \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
+    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
 The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
@@
 Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
+In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\citep{wouters2016rare}).
 %
@@
 %
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
+For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
 %
@@
 %
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
+The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
+{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
+The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \citep{giardina_direct_2006}.
+Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
 
@@
 The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
 The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
@@
 where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\cite{DelMoralBook,DelMoral2013}.
+The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
 %
@@
 %
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
+%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \citep{touchette_large_2009}.
+%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \citep{touchette_large_2009}.
 %
@@
 An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
+This issue is addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
 %
@@
 \end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
+and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
 \begin{equation}
@@
 \end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \cite{De_Rosis_2016}.
-
-\bibliographystyle{unsrt}
-\bibliography{biblio}
+Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability~\citep{De_Rosis_2016}.
+
+\bibliographystyle{jfm}
+\bibliography{../biblio}
 

### 2020-01-07 jfm/jfm-instructions.tex
Add jfm styling

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%       release v3.0, 16th July 2014
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997, 2014 Cambridge University Press

\documentclass{jfm}
\usepackage{graphicx}
\usepackage{epstopdf, epsfig}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\shorttitle{Guidelines for authors}
\shortauthor{A. N. Other, H.-C. Smith and J. Q. Public}

\title{Guidelines for authors and submission template}

\author{Alan N. Other\aff{1}
  \corresp{\email{jfm@damtp.cam.ac.uk}},
  H. - C. Smith\aff{1}
 \and J. Q.  Public\aff{2}}

\affiliation{\aff{1}Department of Chemical Engineering, University of America,
Somewhere, IN 12345, USA
\aff{2}Department of Aerospace and Mechanical Engineering, University of
Camford, Academic Street, Camford CF3 5QL, UK}

\begin{document}

\maketitle

\begin{abstract}
This file contains instructions for authors planning to submit a paper to the {\it Journal of Fluid Mechanics}. These instructions were generated in {\LaTeX} using the JFM style, so the {\LaTeX} source file can be used as a template for submissions. The present paragraph appears in the \verb}abstract} environment. All papers should feature a single-paragraph abstract of no more than 250 words, which provides a summary of the main aims and results. 
\end{abstract}

\begin{keywords}
Authors should not enter keywords on the manuscript, as these must be chosen by the author during the online submission process and will then be added during the typesetting process (see http://journals.cambridge.org/data/\linebreak[3]relatedlink/jfm-\linebreak[3]keywords.pdf for the full list)
\end{keywords}

\section{How to submit to the \textbf{\textit{Journal of Fluid Mechanics}}}
Authors must submit using the online submission and peer review system, Scholar One Manuscripts (formerly Manuscript Central) at http://mc.manuscriptcentral.com/jfm. If visiting the site for the first time, users must create a new account by clicking on `register here'. Once logged in, authors should click on the `Corresponding Author Centre', from which point a new manuscript can be submitted, with step-by-step instructions provided. Authors must at this stage specify whether the submission is a standard paper, or a {\it JFM Rapids} paper (see \S\ref{sec:filetypes} for more details). In addition, authors must specify an editor to whom the paper will be submitted from the drop-down list provided. Note that all editors exclusively deal with either standard papers or {\it JFM Rapids} (clearly indicated on the list), so please ensure that you choose an editor accordingly. Corresponding authors must provide a valid ORCID ID in order to submit a manuscript, either by linking an existing ORCID profile to your ScholarOne account or by creating a new ORCID profile. Once your submission is completed you will receive an email confirmation.  Book reviews should not be submitted via the online submission site and should instead be submitted by email to c.p.caulfield@damtp.cam.ac.uk.

\section{Rules of submission}\label{sec:rules_submission}
Submission of a paper implies a declaration by the author that the work has not previously been published, that it is not being considered for publication elsewhere and that it has not already been considered by a different editor of the Journal. Note that a report on a conference must be submitted within 3 months of the meeting.

\section{Types of paper}\label{sec:types_paper}
\subsection{Standard papers}
Regular submissions to JFM are termed `standard papers'. Note that such papers must contain original research, as review articles are not published in JFM. Papers should be written in a concise manner; though JFM has no page limit, each paper will be judged on its own merits, and those deemed excessive in length will be rejected or will require significant revision.  
\subsection{JFM Rapids}
{\it JFM Rapids} is devoted to the rapid publication of short, high-impact papers across the full range of fluid mechanics. Manuscripts submitted as {\it Rapids}  must be strictly 10 or fewer printed pages, and must be submitted in {\LaTeX} using the jfm.cls class file, so as to ensure that they meet the page limit and to expedite their production.  The principal and over-riding objective is to publish papers of the highest scientific quality.

Once a paper is submitted, referees are asked to provide reports with a short turnaround.  In order to be accepted for publication in {\it JFM Rapids}, such papers must require only minor revisions to improve clarity, usually without recourse to a second round of refereeing. In this case, and at the discretion of the editor, some additional pages may be allowed to address specific points raised by the referees, such as the addition of an extra figure or some explanatory text.

In cases where the editor, guided by the referees, judges that a paper has merit but requires substantial revision that will require significant refereeing, a decision of `revise and resubmit' will be given. On re-submission, such papers will be handled as standard JFM papers and if accepted will not subsequently appear as {\it JFM Rapids}.

{\it Rapids} will be published online within one month of final acceptance.  They will appear within a designated section on the Journal of Fluid Mechanics website.  Each {\it Rapid} will be cited and indexed as a JFM article but with a distinctive {\it Rapids} identifier, and will be assigned to a JFM volume. {\it Rapids} will not be included in the regular print versions of JFM, and accordingly they incur no colour figure charges.

\section{File types}\label{sec:filetypes}
Authors are strongly encouraged to compose their papers in {\LaTeX}, using the jfm.cls style file and supporting files provided at\\ https://www.cambridge.org/core/journals/journal-of-fluid-mechanics/information/\\ instructions-contributors, with the jfm-instructions.tex file serving as a template (note that this is mandatory for {\it JFM Rapids}). A PDF of the {\LaTeX} file should then be generated and submitted via the submission site. Please note that PDFs larger than 10MB are not acceptable for review. There is no need to submit the {\LaTeX} source file alongside the PDF, but upon provisional acceptance of the paper, the {\LaTeX} source file, along with individual figure files and a PDF of the final version, will need to be submitted for typesetting purposes. 
Authors may also compose standard papers in Word, though this will lead to the paper spending a longer period in production. If using Word, please note that equations must NOT be converted to picture format and the file must be saved with the option `make equation editable'. 
All submitted video abstract files should be formatted as MP4 (H.264). MP4 has full compatibility across commonly used browsers, whereas other video formats will only work on selected browsers. This will ensure the greatest possible dissemination of this work. 
\section{Preparing your manuscript}
Authors should write their papers clearly and concisely in English, adhering to JFM's established style for notation, as provided in \S\ref{notstyle}. We encourage the submission of online supplementary material alongside the manuscript where appropriate (see section \ref{online}). Metric units should be used throughout and all abbreviations must be defined at first use, even those deemed to be well known to the readership. British spelling must be used, and should follow the \textit{Shorter Oxford English Dictionary}.

\subsection{Figures}
All authors need to acquire the correct permissions and licences to reproduce figures. Figures should be as small as possible while displaying clearly all the information required, and with all lettering readable. Every effort should be taken to avoid figures that run over more than one page. Figures submitted in colour will appear online in colour but, with the exception of {\it JFM Rapids}, all figures will be printed in black and white unless authors specify during submission that figures should be printed in colour, for which there is a charge of \pounds200 plus VAT per figure (i.e. \pounds240) with a cap of \pounds1000 plus VAT per article (i.e. \pounds1200) (colour is free for {\it JFM Rapids}).  Note that separate figures for online and print will {\bf not} be accepted and it is the author's responsibility to ensure that if a figure is to appear in colour online only, that same figure will still be meaningful when printed in black and white (for example, do not rely upon colours to distinguish lines if those colours will just appear as similar shades of grey when printed). If using colour, authors should also ensure that some consistency is applied within the manuscript. For review purposes figures should be embedded within the manuscript. Upon final acceptance, however, individual figure files will be required for production. These should be submitted in EPS or high-resolution TIFF format (1200 dpi for lines, 300 dpi for halftone and colour in CMYK format, and 600 dpi for a mixture of lines and halftone). The minimum acceptable width of any line is 0.5pt. Each figure should be accompanied by a single caption, to appear beneath, and must be cited in the text. Figures should appear in the order in which they are first mentioned in the text and figure files must be named accordingly (`Figure 1.eps', `Figure 2a.tiff', etc) to assist the production process (and numbering of figures should continue through any appendices). The word \textit {figure} is only capitalized at the start of a sentence. For example see figures \ref{fig:ka} and \ref{fig:kd}. Failure to follow figure guidelines may result in a request for resupply and a subsequent delay in the production process. Note that {\em all} figures are relabelled by the typesetter, therefore the files supplied should have editable labelling. Please ensure all figure labels are carefully checked against your originals when you receive your proofs.

\begin{figure}
  \centerline{\includegraphics{trapped}}% Images in 100% size
  \caption{Trapped-mode wavenumbers, $kd$, plotted against $a/d$ for
    three ellipses:\protect\\
    ---$\!$---,
    $b/a=1$; $\cdots$\,$\cdots$, $b/a=1.5$.}
\label{fig:ka}
\end{figure}

\begin{figure}
  \centerline{\includegraphics{modes}}
  \caption{The features of the four possible modes corresponding to
  (\textit{a}) periodic\protect\\ and (\textit{b}) half-periodic solutions.}
\label{fig:kd}
\end{figure}

\subsection{Tables}
Tables, however small, must be numbered sequentially in the order in which they are mentioned in the text. The word \textit {table} is only capitalized at the start of a sentence. See table \ref{tab:kd} for an example.

\begin{table}
  \begin{center}
\def~{\hphantom{0}}
  \begin{tabular}{lccc}
      $a/d$  & $M=4$   &   $M=8$ & Callan \etal \\[3pt]
       0.1   & 1.56905 & ~~1.56~ & 1.56904\\
       0.3   & 1.50484 & ~~1.504 & 1.50484\\
       0.55  & 1.39128 & ~~1.391 & 1.39131\\
       0.7   & 1.32281 & ~10.322 & 1.32288\\
       0.913 & 1.34479 & 100.351 & 1.35185\\
  \end{tabular}
  \caption{Values of $kd$ at which trapped modes occur when $\rho(\theta)=a$}
  \label{tab:kd}
  \end{center}
\end{table}

\subsection{Online supplementary material}\label{online}
Relevant material which is not suitable for print production, such as movies or numerical simulations/animations, can be uploaded as part of the initial submission. Movies should be designated as `Movie' and each individual file must be accompanied by a separate caption and a suitable title (eg Movie 1). Accepted formats are .mov, .mpg, .mp4, and .avi, though they should be archived as a .zip or .tar file before uploading. Each movie should be no more than 10MB. Upon publication these materials will then be hosted online alongside the final published article. Likewise, should there be detailed mathematical relations, tables or figures which are likely to be useful only to a few specialists or take up excessive space in the printed journal, these can also be published online as supplementary material [designated as `Other supplementary material']. Note that supplementary material is published `as is', with no further production performed.

\section{Editorial decisions}
\subsection{Revision}
If a revision is requested, you should upload revised files following the same procedure as for submitting a new paper. You begin by clicking on `Manuscripts with decision' in your Corresponding Author Center, and then on `Create a revision'. (Note that if you abandon the process before completing the submission, to continue the submission, you must click on `Revised manuscripts in draft'.) There is a new first page showing the decision letter and a space for your reply to the referee's/editor's comments. You also have the opportunity at this stage to upload your reply to the comments as a separate file. All the values filled in on original submission are displayed again. The ID number of the paper will be appended `.R1'. Also note that if a manuscript is submitted as a {\it JFM Rapid}, but requires substantial revision, it will be re-designated as a standard paper, and the ID and paper type will be amended to reflect this.

\subsection{Provisional acceptance}
If the paper is accepted as suitable for publication you will be sent a provisional acceptance decision. This enables you to upload the final files required for production:
(1) the final PDF or word version of the paper, designated as a `main document';
(2) any source files (see section \ref{sec:filetypes}) which must be designated as `production (not for review)' and uploaded as a single .zip or .tar file.

\subsection{Acceptance}
On receipt of the production files you will be sent an email indicating completion of the acceptance process.

\section{Publication process}
Once a paper has been accepted for publication and the source files have been uploaded, the manuscript will be sent to Cambridge University Press for copyediting and typesetting, and will be assigned a digital object identifier (doi). When the proof is ready, authors will receive an email alert containing a link to the PDF of the proof, and instructions for its correction and return. It is imperative that authors check their proofs closely, particularly the equations and figures, which should be checked against the accepted file, as the production schedule does not allow for corrections at a later stage. Once ready for printing, papers will be published online on Cambridge Journals Online in the current `open' volume. {\it JFM Rapids} will also be assigned to the current open volume, but will receive an article number in place of traditional pagination, as they will not appear in the print version of that volume. Each volume will be kept open for approximately two weeks, at which point it will be considered `closed', and then printed and distributed. The following volume will be immediately opened. Note that the PDF published online is the Version of Record, matching the print version, and no further alterations/corrections to this document will be allowed. The corresponding author is emailed a link to the published article when it is first published online.

\section{Corrigenda}
The Journal will publish corrigenda that alter significant conclusions made in a paper.  Such corrigenda should be submitted to an associate editor, who will consider the submission similarly to a new paper and may consult referees if appropriate.  When published, corrigenda are clearly linked with the original articles to which they refer, and the articles to them.

The Journal does not normally publish corrigenda to amend typographical errors, so it is extremely important that authors make very careful checks of their manuscript at every stage, including the reading of proofs, prior to publication.

\section{Obtaining help}
Technical support for the online submission system is available by clicking on the `Get Help Now' link at the top-right corner of each page of the submission site. Any other questions relating to the submission or publication process should be directed to the JFM Editorial Assistant, Mrs Amanda Johns, at ajohns@cambridge.org.

\section{Cambridge Language Editing Service} 
We suggest that authors whose first language is not English have their manuscripts checked by a native English speaker before submission. This is optional but will help to ensure that any submissions that reach peer review can be judged exclusively on academic merit. We offer a Cambridge service which you can find out more about at https://www.cambridge.org/core/services/authors/language-services, and suggest that authors make contact as appropriate. Please note that use of language editing services is voluntary and at the author’s own expense. Use of these services does not guarantee that the manuscript will be accepted for publication nor does it restrict the author to submitting to a Cambridge-published journal

\section{Notation and style}\label{notstyle}
Generally any queries concerning notation and journal style can be answered by viewing recent pages in the Journal. However, the following guide provides the key points to note. It is expected that Journal style will be followed, and authors should take care to define all variables or entities upon first use. Also note that footnotes are not normally accepted.

\subsection{Mathematical notation}
\subsubsection{Setting variables, functions, vectors, matrices etc}
{\bf Italic font} should be used for denoting variables, with multiple-letter symbols avoided except in the case of dimensionless numbers such as $\Rey$, $\Pran$ and $\Pen$ (Reynolds, Prandtl, and P\'eclet numbers respectively, which are defined as \verb}\Rey}, \verb}\Pran} and {\verb}\Pen} in the template).\\
{\bf Upright Roman font} (or upright Greek where appropriate) should be used for:\\
Operators: sin, log, d, $\Delta$, e etc.\\
Constants: i ($\sqrt{-1}$), $\upi$ (defined as \verb}\upi}), etc.\\
Functions: $\Ai$, $\Bi$ (Airy functions, defined as \verb}\Ai} and \verb}\Bi}), $\Real$ (real part, defined as \verb}\Real}), $\Imag$ (imaginary part, defined as \verb}\Imag}), etc.\\
Physical units: cm, s, etc\\
Abbreviations: c.c. (complex conjugate), h.o.t. (higher-order terms), DNS, etc.\\
{\bf Bold italic font} (or bold sloping Greek) should be used for:\\
Vectors (with the centred dot for a scalar product also in bold): $\boldsymbol{i \cdot j}$\\
{\bf Bold sloping sans serif font}, defined by the \verb}\mathsfbi} macro, should be used for:\\
Tensors and matrices: $\mathsfbi{D}$ \\
{\bf Script font} (for example $\mathcal{G}$, $\mathcal{R}$) can be used as an alternative to italic when the same letter denotes a different quantity (use \verb}\mathcal} in \LaTeX)\\
The product symbol ($\times$) should only be used to denote multiplication where an equation is broken over more than one line, to denote a cross product, or between numbers (the $\cdot$ symbol should not be used, except to denote a scalar product specifically).\\

\subsubsection{Other symbols}
A centred point should be used only for the scalar product of vectors.
Large numbers that are not scientific powers should not include commas, but have the
form 1600 or 16 000 or 160 000.
Use \textit{O} to denote `of the order of', not the \LaTeX\ $\mathcal{O}$.

\section{Citations and references}
All papers included in the References section must be cited in the article, and vice versa. Citations should be included as, for example ``It has been shown \citep{Rogallo81} that...'' (using the {\verb}\citep} command, part of the natbib package) ``recent work by \citet{Dennis85}...'' (using {\verb}\citet}).
The natbib package can be used to generate citation variations, as shown below.\\
\verb#\citet[pp. 2-4]{Hwang70}#:\\
\citet[pp. 2-4]{Hwang70} \\
\verb#\citep[p. 6]{Worster92}#:\\
\citep[p. 6]{Worster92}\\
\verb#\citep[see][]{Koch83, Lee71, Linton92}#:\\
\citep[see][]{Koch83, Lee71, Linton92}\\
\verb#\citep[see][p. 18]{Martin80}#:\\
\citep[see][p. 18]{Martin80}\\
\verb#\citep{Brownell04,Brownell07,Ursell50,Wijngaarden68,Miller91}#:\\
\citep{Brownell04,Brownell07,Ursell50,Wijngaarden68,Miller91}\\
The References section can either be built from individual \verb#\bibitem# commands, or can be built using BibTex. The BibTex files used to generate the references in this document can be found in the zip file at http://journals.cambridge.org/\linebreak[3]data/\linebreak[3]relatedlink/\linebreak[3]jfm-ifc.zip.\\
Where there are up to ten authors, all authors' names should be given in the reference list. Where there are more than ten authors, only the first name should appear, followed by et al.\\

Acknowledgements should be included at the end of the paper, before the References section or any appendicies, and should be a separate paragraph without a heading. Several anonymous individuals are thanked for contributions to these instructions.

\appendix
\section{}\label{appA}
This appendix contains sample equations in the JFM style. Please refer to the {\LaTeX} source file for examples of how to display such equations in your manuscript.

\begin{equation}
  (\nabla^2+k^2)G_s=(\nabla^2+k^2)G_a=0
  \label{Helm}
\end{equation}

\begin{equation}
  \bnabla\bcdot\boldsymbol{v} = 0,\quad \nabla^{2}P=
    \bnabla\bcdot(\boldsymbol{v}\times \boldsymbol{w}).
\end{equation}

\begin{equation}
  G_s,G_a\sim 1 / (2\upi)\ln r
  \quad \mbox{as\ }\quad r\equiv|P-Q|\rightarrow 0,
  \label{singular}
\end{equation}

\begin{equation}
\left. \begin{array}{ll}  
\displaystyle\frac{\p G_s}{\p y}=0
  \quad \mbox{on\ }\quad y=0,\\[8pt]
\displaystyle  G_a=0
  \quad \mbox{on\ }\quad y=0,
 \end{array}\right\}
  \label{symbc}
\end{equation}

\begin{equation}
  -\frac{1}{2\upi} \int_0^{\infty} \gamma^{-1}[\mathrm exp(-k\gamma|y-\eta|)
   + \mathrm exp(-k\gamma(2d-y-\eta))] \cos k(x-\xi)t\:\mathrm{d} t,
   \qquad 0<y,\quad \eta<d,
\end{equation}

\begin{equation}
  \gamma(t) = \left\{
    \begin{array}{ll}
      -\mathrm{i}(1-t^2)^{1/2}, & t\le 1 \\[2pt]
      (t^2-1)^{1/2},         & t>1.
    \end{array} \right.
\end{equation}

\[
  -\frac{1}{2\upi}
   \pvi B(t)\frac{\cosh k\gamma(d-y)}{\gamma\sinh k\gamma d}
   \cos k(x-\xi)t\:\mathrm{d} t
\]

\begin{equation}
  G = -\frac{1}{4}\mathrm{i} (H_0(kr)+H_0(kr_1))
    - \frac{1}{\upi} \pvi\frac{\mathrm{e}^{-\kgd}}%
    {\gamma\sinh\kgd} \cosh k\gamma(d-y) \cosh k\gamma(d-\eta)
\end{equation}

Note that when equations are included in definitions, it may be suitable to render them in line, rather than in the equation environment: $\boldsymbol{n}_q=(-y^{\prime}(\theta),
x^{\prime}(\theta))/w(\theta)$.
Now $G_a=\squart Y_0(kr)+\Gat$ where
$r=\{[x(\theta)-x(\psi)]^2 + [y(\theta)-y(\psi)]^2\}^{1/2}$ and $\Gat$ is
regular as $kr\ttz$. However, any fractions displayed like this, other than $\thalf$ or $\squart$, must be written on the line, and not stacked (ie 1/3).

\begin{eqnarray}
  \ndq\left(\frac{1}{4} Y_0(kr)\right) & \sim &
    \frac{1}{4\upi w^3(\theta)}
    [x^{\prime\prime}(\theta)y^{\prime}(\theta)-
    y^{\prime\prime}(\theta)x^{\prime}(\theta)] \nonumber\\
  & = & \frac{1}{4\upi w^3(\theta)}
    [\rho^{\prime}(\theta)\rho^{\prime\prime}(\theta)
    - \rho^2(\theta)-2\rho^{\prime 2}(\theta)]
    \quad \mbox{as\ }\quad kr\ttz . \label{inteqpt}
\end{eqnarray}

\begin{equation}
  \frac{1}{2}\phi_i = \frac{\upi}{M} \sumjm\phi_j K_{ij}^a w_j,
  \qquad i=1,\,\ldots,\,M,
\end{equation}
where
\begin{equation}
  K_{ij}^a = \left\{
    \begin{array}{ll}
      \p G_a(\theta_i,\theta_j)/\p n_q, & i\neq j \\[2pt]
      \p\Gat(\theta_i,\theta_i)/\p n_q
      + [\rho_i^{\prime}\rho_i^{\prime\prime}-\rho_i^2-2\rho_i^{\prime 2}]
      / 4\upi w_i^3, & i=j.
  \end{array} \right.
\end{equation}

\refstepcounter{equation}
$$
  \rho_l = \lim_{\zeta \rightarrow Z^-_l(x)} \rho(x,\zeta), \quad
  \rho_{u} = \lim_{\zeta \rightarrow Z^{+}_u(x)} \rho(x,\zeta)
  \eqno{(\theequation{\mathit{a},\mathit{b}})}\label{eq35}
$$

\begin{equation}
  (\rho(x,\zeta),\phi_{\zeta\zeta}(x,\zeta))=(\rho_0,N_0)
  \quad \mbox{for}\quad Z_l(x) < \zeta < Z_u(x).
\end{equation}

\begin{subeqnarray}
  \tau_{ij} & = &
    (\overline{\overline{u}_i \overline{u}_j}
    - \overline{u}_i\overline{u}_j)
    + (\overline{\overline{u}_iu^{SGS}_j
    + u^{SGS}_i\overline{u}_j})
    + \overline{u^{SGS}_iu^{SGS}_j},\\[3pt]
  \tau^\theta_j & = &
    (\overline{\overline{u}_j\overline{\theta}}
    - \overline{u}_j \overline{\theta})
    + (\overline{\overline{u}_j\theta^{SGS}
    + u^{SGS}_j \overline{\theta}})
    + \overline{u^{SGS}_j\theta^{SGS}}.
\end{subeqnarray}

\begin{equation}
\setlength{\arraycolsep}{0pt}
\renewcommand{\arraystretch}{1.3}
\slsQ_C = \left[
\begin{array}{ccccc}
  -\omega^{-2}V'_w  &  -(\alpha^t\omega)^{-1}  &  0  &  0  &  0  \\
  \displaystyle
  \frac{\beta}{\alpha\omega^2}V'_w  &  0  &  0  &  0  &  \mathrm{i}\omega^{-1} \\
  \mathrm{i}\omega^{-1}  &  0  &  0  &  0  &  0  \\
  \displaystyle
  \mathrm{i} R^{-1}_{\delta}(\alpha^t+\omega^{-1}V''_w)  &  0
    & -(\mathrm{i}\alpha^tR_\delta)^{-1}  &  0  &  0  \\
  \displaystyle
  \frac{\mathrm{i}\beta}{\alpha\omega}R^{-1}_\delta V''_w  &  0  &  0
    &  0  & 0 \\
  (\mathrm{i}\alpha^t)^{-1}V'_w  &  (3R^{-1}_{\delta}+c^t(\mathrm{i}\alpha^t)^{-1})
    &  0  &  -(\alpha^t)^{-2}R^{-1}_{\delta}  &  0  \\
\end{array}  \right] .
\label{defQc}
\end{equation}

\begin{equation}
\etb^t = \skew2\hat{\etb}^t \exp [\mathrm{i} (\alpha^tx^t_1-\omega t)],
\end{equation}
where $\skew2\hat{\etb}^t=\boldsymbol{b}\exp (\mathrm{i}\gamma x^t_3)$. 
\begin{equation}
\mbox{Det}[\rho\omega^2\delta_{ps}-C^t_{pqrs}k^t_qk^t_r]=0,
\end{equation}

\begin{equation}
 \langle k^t_1,k^t_2,k^t_3\rangle = \langle
\alpha^t,0,\gamma\rangle  
\end{equation}

\begin{equation}
\boldsymbol{f}(\theta,\psi) = (g(\psi)\cos \theta,g(\psi) \sin \theta,f(\psi)).
\label{eq41}
\end{equation}

\begin{eqnarray}
f(\psi_1) = \frac{3b}{\upi[2(a+b \cos \psi_1)]^{{3}/{2}}}
  \int^{2\upi}_0 \frac{(\sin \psi_1 - \sin \psi)(a+b \cos \psi)^{1/2}}%
  {[1 - \cos (\psi_1 - \psi)](2+\alpha)^{1/2}}\mathrm{d}x,
\label{eq42}
\end{eqnarray}
\begin{eqnarray}
g(\psi_1) & = & \frac{3}{\upi[2(a+b \cos \psi_1)]^{{3}/{2}}}
  \int^{2\upi}_0 \left(\frac{a+b \cos \psi}{2+\alpha}\right)^{1/2}
  \left\{ \astrut f(\psi)[(\cos \psi_1 - b \beta_1)S + \beta_1P]
  \right. \nonumber\\
&& \mbox{}\times \frac{\sin \psi_1 - \sin \psi}{1-\cos(\psi_1 - \psi)}
  + g(\psi) \left[\left(2+\alpha - \frac{(\sin \psi_1 - \sin \psi)^2}
  {1- \cos (\psi - \psi_1)} - b^2 \gamma \right) S \right.\nonumber\\
&& \left.\left.\mbox{} + \left( b^2 \cos \psi_1\gamma -
  \frac{a}{b}\alpha \right) F(\frac{1}{2}\upi, \delta) - (2+\alpha)
  \cos\psi_1 E(\frac{1}{2}\upi, \delta)\right] \astrut\right\} \mathrm{d} \psi,
\label{eq43}
\end{eqnarray}
\begin{equation}
\alpha = \alpha(\psi,\psi_1) = \frac{b^2[1-\cos(\psi-\psi_1)]}%
  {(a+b\cos\psi) (a+b\cos\psi_1)},
  \quad
  \beta - \beta(\psi,\psi_1) = \frac{1-\cos(\psi-\psi_1)}{a+b\cos\psi}.
\end{equation}

\begin{equation}
\left. \begin{array}{l}
\displaystyle
H(0) = \frac{\epsilon \overline{C}_v}{\tilde{v}^{{1}/{2}}_T
(1- \beta)},\quad H'(0) = -1+\epsilon^{{2}/{3}} \overline{C}_u
+ \epsilon \skew5\hat{C}_u'; \\[16pt]
\displaystyle
H''(0) = \frac{\epsilon u^2_{\ast}}{\tilde{v}^{{1}/{2}}
_T u^2_P},\quad H' (\infty) = 0.
\end{array} \right\}
\end{equation}

\begin{lemma}
Let $f(z)$ be a trial \citet[][pp.~231--232]{Batchelor59} function defined on $[0,1]$.  Let $\varLambda_1$ denote
the ground-state eigenvalue for $-\mathrm{d}^2g/\mathrm{d} z^2=\varLambda g$,
where $g$ must satisfy $\pm\mathrm{d} g/\mathrm{d} z+\alpha g=0$ at $z=0,1$
for some non-negative constant~$\alpha$.  Then for any $f$ that is not
identically zero we have
\begin{equation}
\frac{\displaystyle
  \alpha(f^2(0)+f^2(1)) + \int_0^1 \left(
  \frac{\mathrm{d} f}{\mathrm{d} z} \right)^2 \mathrm{d} z}%
  {\displaystyle \int_0^1 f^2\mathrm{d} z}
\ge \varLambda_1 \ge
\left( \frac{-\alpha+(\alpha^2+8\upi^2\alpha)^{1/2}}{4\upi} \right)^2.
\end{equation}
\end{lemma}

\begin{corollary}
Any non-zero trial function $f$ which satisfies the boundary condition
$f(0)=f(1)=0$ always satisfies
\begin{equation}
  \int_0^1 \left( \frac{\mathrm{d} f}{\mathrm{d} z} \right)^2 \mathrm{d} z.
\end{equation}
\end{corollary}

\bibliographystyle{jfm}
% Note the spaces between the initials
\bibliography{jfm-instructions}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-01-07 draft_p2.tex
test_commit

@@
 \begin{abstract}
-This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+Blabla !!! This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
 The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.

### 2020-01-08 jfm/draft_p2_jfm.tex
Update Emmanuel following review

@@
 
+\definecolor{myred}{HTML}{b30000}
+\definecolor{mygreen}{HTML}{008000}
+\newcommand{\EL}[1]{{\color{myred}{#1}}}
+\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
+
 \title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
@@
 % general comments on physical problem %
-
+% 
 Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
@@
 % 
-{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
+{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+% 
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
 %
@@
 An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
+% Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
 % 
@@
 The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently is has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
 
 During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \citep{moral_feynman-kac_2004}. 
+Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}.
 Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
@@
   This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
+Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
 %
@@
 The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
-drag fluctuations is investigated based on a direct sampling approach.
+In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
+  drag fluctuations} is investigated based on a direct sampling approach.
 Both the instantaneous drag and time-averaged drag are considered.
@@
 We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems.
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
+% Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
 % 
@@
 % 
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
+% Finally, the obstacle does not deform or move either.
+% 
+% Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
 %
@@
 Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+% This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix and references therein.
+Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
 In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
@@
 Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
 
@@
   \includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
+  \caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
     The vorticity is given in lattice units.}
@@
 \end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
+where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
 Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
@@
 between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity fiel.
+Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
 Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
@@
   \label{eq:turnover_time}
-\tau_0 = \frac{R}{U},
+  \EL{T_0 = \frac{R}{U}},
 \end{equation}
 where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+% The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
 % 
@@
 \subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
+  % : Probability Density Function and correlation in time
 }
@@
   \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
+The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
   \label{fig:typical_drag_signal}
@@
 % pdf
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$.
+Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
 This long simulation will be referred to as the \textit{control run} in the following.
@@
 The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
 % 
@@
     \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is defined by $C(\tau_c)=0$.
-  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
+The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
+    % $\tau_c$ may therefore be considered as the correlation time of the drag signal.
   }
@@
 Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}.
+One can then argue that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
 % 
@@
 In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
+The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
+The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.
 
@@
   \includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a function the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
+  \caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
   }
@@
 The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
+The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
+As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
 For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
 
-How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
+How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
 We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
@@
 % 
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
+% The shape of the return time $r$ as a function of the return level $a$ is system dependent.
 If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
@@
   \label{eq:return_time}
-r(a) \underset{a\to\infty}{\propto} \exp(-la)
+  \EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
 \end{equation}
@@
 
-\subsection{Instantaneous drag}
+\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
 \label{sec:instantaneous_drag}
@@
     As the amplitude of the drag increases, the relative variation of the base pressure also increases.
-	%\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} 
+\EL{The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
   }
@@
 Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
-%relatively to the forebody pressure.
+% relatively to the forebody pressure.
 
@@
 This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
+% Starting from typical values, extreme drags are typically reached in less than a correlation time.
 Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-This is reminiscent of time-irreversibility in turbulent dynamics. 
-%Moreover, such very high drag levels do not persist over time.
+% This is reminiscent of time-irreversibility in turbulent dynamics.
+\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
+% Moreover, such very high drag levels do not persist over time.
 To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
@@
 The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
+% +
 The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
@@
 
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
+% We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
 
@@
 On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
+% Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
@@
 % 
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
+% In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
 In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
 
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
+% We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
 We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
@@
 We shall consider $T=10\tau_c$ in the following.
+\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:IS_GKTL}).}
 
@@
 % 
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
+% The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
 By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
+\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
+As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
 
@@
 The observation is therefore well supported theoretically.
-See~\citep{lestang2019JSTAT} for further details.
+See~\cite{lestang2019JSTAT} for further details.
 
@@
 In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-%In a nutshell
+% In a nutshell
 Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
@@
 In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
+Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
 \end{figure}
@@
 The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
-%a sequence of events that can be sampled with a higher probability  
+% a sequence of events that can be sampled with a higher probability
 a sequence of shorter excursions \citep{glasserman_multilevel_1999,rolland_statistical_2015}.
@@
 % 
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\citep{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\citep{rolland_extremely_2018}.
+% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
 % 
@@
 Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\citep{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \citep{bouchet2019rare,rolland_extremely_2018}.
+See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
+For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
 {Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
@@
 % 
-%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+% Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
 % 
@@
 In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features do no \textit{a priori} persist in the case of fluid dynamics.
+Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
 
@@
   Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\citep{wouters2016rare}).
+  In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
   % 
@@
   We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
-The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
+  The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow \EL{(see Appendix \ref{app:perturb_branching_time})}.
   It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
+% \EL{Freddy: provide reference for Lyapunov timescale. \#12}
 
@@
 After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
+% Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
 The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
 Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-%(see Fig.~\ref{fig:comparaison_temps_de_retour}).
+% (see Fig.~\ref{fig:comparaison_temps_de_retour}).
 
@@
 In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
+Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
     The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
@@
 
-%pitfall
+% pitfall
 % 
@@
 % 
-As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
+As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
 On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
@@
 
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
+% As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
 Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
@@
 On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-A possible line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
+% which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
+% 
+% Furthermore, such ``tuning'' is problem dependent.
+% 
+A possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
 That way, new events associated to higher fluctuations may have the opportunity to develop.
@@
 
-\subsection{Importance sampling of extreme time-averaged drag forces with the \acl{gktl} algorithm}
+\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
 \label{sec:gktl}
@@
 The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
+% in section~\ref{sec:ams_drag} \textit{i.e.}
 by taking the time-averaged observable itself as the score function.
@@
 % 
-Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
+\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
 % In this case, however, the trajectories interact dynamically.
@@
 {Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \citep{giardina_direct_2006}.
+The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
 Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
@@
 The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
 The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
@@
 
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics.
-All trajectories in the resampled ensemble are solutions of the original dynamical.
-%This is true for stochastic dynamics.
-{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to allow clones of a same trajectory to separate. 
-As for the \ac{ams} algorithm, it is assumed that this perturbation does not significantly affect the statistics of the sampled trajectories.}
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
+% This is true for stochastic dynamics.
+\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+  As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+% 
+% 
+% However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
 We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
@@
 
-\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
+\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
 \label{sec:gktl_drag}
@@
 % 
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \citep{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \citep{touchette_large_2009}.
+% Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
+% In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
 % 
@@
   \rho_k(F)
-  %= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
+  % = \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
   \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
@@
 As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
+% Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
 % 
@@
 Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
+% \footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
 Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
 % 
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
+% This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
+% 
+% \begin{figure}
 %   \centering
@@
 %   \label{fig:PDF_AVG}
-%\end{figure}
+% \end{figure}
 % 
@@
 % 
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+% In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
 By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
@@
 events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
+% (see Fig.~\ref{fig:illustrate_return_time})
 
@@
 \end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
+and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
 \begin{equation}
@@
 Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independant estimates, obtained using different values of the bias parameter $k$.
+The figure shows two independent estimates, obtained using different values of the bias parameter $k$.
 Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
+In addition, \EL{Fig.~\ref{fig:return_times_gktl}} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
 Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
@@
 This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to yield an large value of the average.
+Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
 Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
@@
 In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent. 
-%following the addition of a small perturbation in the initial conditions.
+This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent.
+% following the addition of a small perturbation in the initial conditions.
 The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
@@
 
-\appendix*
+\appendix
 \section{The \acl{lbm}}
-\label{sec:lbm}
+\label{app:lbm}
 
 % details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
@@
 \]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas.
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
 % 
@@
 This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
+$\Delta x \to 0$ \EL{\citep{succi_book}}.
 
@@
 \end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability~\citep{De_Rosis_2016}.
+Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
+
+\EL{
+  \section{Perturbation at branching time}
+  \label{app:perturb_branching_time}
+}
 
@@
 \end{document}
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2020-01-28 figures/pressure_ratio/plot_pressure_ratio_revision.py
Save new figure pressure_ratio

@@ -57,4 +57,7 @@ ax.legend(loc='best', fontsize=22)
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)
 
+fname = 'pressure_ratio.png'
+plt.savefig(fname)
+
 plt.show()

### 2020-01-28 jfm/draft_p2_jfm.tex
Modififs Emmanuel pour la revision: elagage de la partie sur les algorithmes.

@@
 \usepackage{caption}
+
+\usepackage[normalem]{ulem}
+\usepackage{soul}
+
 \graphicspath{{../figures/}}
@@
 
-\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
+
+\title{\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+\ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
+}
 
@@
 \begin{abstract}
-  This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+	\ZZ{This study investigates by means of direct numerical simulation, extreme drag events exerted by a  turbulent flow on a bluff body and examine the relevance of statistical algorithms to sample efficiently these events.
+Two representative algorithms are examined: AMS and GKTL.
+Briefly mention differences and complementarity.
+What do we learn?
+The take-home message is...}
+\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
 The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
@@
 Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
+}
 \end{abstract}
@@
 ``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Such anomalous constraints can threaten the structural integrity of embedded structures.
+In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures.
 %
@@
 From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
+The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
+Otherwise, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
 However, this requires very long simulations since these events are rare.
@@
 %
-
+%
 %
@@
 %
-An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-% Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
+\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+}
+%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
 %
@@
 %
+
+\smallskip
+
+\ZZ{The next two paragraphs may be moved to the introduction of section \ref{sec:rare_events_algorithms}?}
+\smallskip
 
@@
 % Annonce du plan
-In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described.
+In more detail,  the flow set-up is introduced and the dynamics related to typical
+drag fluctuations is described in section~\ref{sec:test_flow}.
 The statistical properties of the drag are then discussed.
@@
 Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
+\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
 Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
+In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
 This difficulty is put in perspective with the phenomenology developed in the previous sections.
@@
 This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
+As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
 Perspectives and conclusion end this work.
@@
 %
-We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems.
-% Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
+We believe that this system embeds the characteristic features that makes \sout{the application of rare-event algorithms} \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
+%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
 %
@@
 %
-% Finally, the obstacle does not deform or move either.
-% 
-% Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
+%Finally, the obstacle does not deform or move either.
+%
+%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
 %
@@
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs.
-% This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
+%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
@@
 where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
-% The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
 %
@@
 \subsection{The drag as a random process
-  % : Probability Density Function and correlation in time
+	%: Probability Density Function and correlation in time
 }
@@
 The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
-    % $\tau_c$ may therefore be considered as the correlation time of the drag signal.
+		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
 	}
@@
 %
-% The shape of the return time $r$ as a function of the return level $a$ is system dependent.
+%The shape of the return time $r$ as a function of the return level $a$ is system dependent.
 If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
@@
 Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
-% relatively to the forebody pressure.
+%relatively to the forebody pressure.
 
@@
 This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-% Starting from typical values, extreme drags are typically reached in less than a correlation time.
+%Starting from typical values, extreme drags are typically reached in less than a correlation time.
 Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-% This is reminiscent of time-irreversibility in turbulent dynamics.
+%This is reminiscent of time-irreversibility in turbulent dynamics.
 \EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-% Moreover, such very high drag levels do not persist over time.
+%Moreover, such very high drag levels do not persist over time.
 To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
@@
 The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-% +
+%+
 The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
@@
 
-% We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
+%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
 
@@
 On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-% Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
+%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
@@
 %
-% In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
+%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
 In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
 
-% We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
+%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
 We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
@@
 %
-% The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
+%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
@@
 Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-It is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
-Indeed, both cases are equally featured in Fig.~\ref{fig:extreme_avg}.
-Very large value of the averaged-drag results from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
-% 
-% theoretical explanation
-% 
-This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
-Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
-\begin{equation}
-  \label{eq:indep}
-  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
-\end{equation}
-If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-\begin{equation}
-  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-  \label{eq:ratio_exp}
-\end{equation}
-The observation is therefore well supported theoretically.
-See~\cite{lestang2019JSTAT} for further details.
+\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
+Indeed, both cases are %equally
+featured in Fig.~\ref{fig:extreme_avg};
+very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
+
+\bigskip
+
+\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
+
+\bigskip
+
+% theoretical explanation is removed because it is speculative
+
+%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
+%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
+%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
+%\begin{equation}
+%\label{eq:indep}
+%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
+%\end{equation}
+%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
+%\begin{equation}
+%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
+%\label{eq:ratio_exp}
+%\end{equation}
+%The observation is therefore well supported theoretically.
+%See~\cite{lestang2019JSTAT} for further details.
+
+\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience.
+Technical details are either removed (reference to a companion paper) or moved to Appendix}
 
@@
 \label{sec:rare_events_algorithms}
+
 % limitation of direct sampling
 %
-In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like $r(a) \propto e^{l a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
+
 %% objective
@@
 %
+
+\ZZ{Should develop here the idea behind rare event sampling.
+How it works.
+Include here the paragraphs of the introduction.
+Justify the choice of ams and gktl: what are differences and complementarity.
+AMS = instantaneous signal, GKTL= time-averaged signal.
+Why are they representative?
+What are the expectations?}
+
+%
 For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
@@
 In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-% In a nutshell
+%In a nutshell
 Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
@@
 The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables.
+
 % These two algorithms will be detailed in the following.
@@
 The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
-% a sequence of events that can be sampled with a higher probability
+%a sequence of events that can be sampled with a higher probability
 a sequence of shorter excursions \citep{glasserman_multilevel_1999,rolland_statistical_2015}.
@@
 
-\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-\label{sec:ams_ou}
-% Example: OU process
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-  \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
+\bigskip
+
+\ZZ{The Ornstein-uhlenbeck application has been moved to appendix \ref{app:AMS_on_OU}. Enough to mention in the text that AMS works successfully for OU and refer to apppendix?}
+
+\bigskip
+
+% SUBSECTION IS MOVED TO APPENDIX %
+
+%\subsubsection{}
+%\label{sec:ams_ou}
+%% Example: OU process
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
 The blue line is the analytical solution for the return time of amplitude $a$.}
-  \label{fig:comparaison_temps_de_retour}
-\end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-\begin{equation}
-  \label{eq:ou}
-  \dot{x} = -x + \eta (t),
-\end{equation}
-where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-% 
-% Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-% 
-The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-In the latter, the typical computational cost is simply the return time $r(a)$.
-One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+%	\label{fig:comparaison_temps_de_retour}
+%\end{figure}
+%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+%\begin{equation}
+%\label{eq:ou}
+%\dot{x} = -x + \eta (t),
+%\end{equation}
+%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+%This basic stochastic process will allow us to highlight differences with fluid dynamics.
+%
+%
+%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+%%
+%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+%%
+%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
+%In the latter, the typical computational cost is simply the return time $r(a)$.
+%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+%
+%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
 Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
 
-\subsubsection{The \ac{ams} for extreme drag fluctuations}
-\label{sec:ams_drag}
+%\subsubsection{The \ac{ams} for extreme drag fluctuations}
+%\label{sec:ams_drag}
 
@@
 	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-  \caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
+	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding \EL{computational cost $C_{AMS}$}.
 The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
@@
 %
-Our aim is now to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
+Our aim is here to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
 In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
@@
 It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-% \EL{Freddy: provide reference for Lyapunov timescale. \#12}
+%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
 
@@
 %
-Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost, similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
+Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
+% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
 %
@@
 After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-% Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
+%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
 The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-% (see Fig.~\ref{fig:comparaison_temps_de_retour}).
+%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
 
@@
 
-% pitfall
+%pitfall
 %
@@
 
-% As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
+%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
 Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
@@
 On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-% which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-% 
-% Furthermore, such ``tuning'' is problem dependent.
+%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
+%
+%Furthermore, such ``tuning'' is problem dependent.
 %
@@
 The \ac{ams} algorithm could be used in the same way as before
-% in section~\ref{sec:ams_drag} \textit{i.e.}
+%in section~\ref{sec:ams_drag} \textit{i.e.}
 by taking the time-averaged observable itself as the score function.
@@
 
-\subsubsection{The \ac{gktl} algorithm}
-\label{sec:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-% 
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-% 
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-% 
-\begin{equation}
-  W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-  \label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-% 
-% 
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
+\bigskip
+\ZZ{The theoretical introduction of GKTL is moved to appendix.}
+
+\bigskip
+
+%\subsubsection{The \ac{gktl} algorithm}
+%\label{sec:gktl_description}
+%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%%
+%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
+%%
+%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
+%%
+%\begin{equation}
+%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+%\label{eq:Weight}
+%\end{equation}
+%so that $\sum_{n=1}^N W_n^i = N$.
+%%
+%%
+%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
 The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
 For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-% 
-The sampled biased distribution writes
-% 
-\begin{align}
-  \mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-                                                                                                                                     \label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
-refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-% 
-One can mention that
-\begin{equation}
-  \label{eq:mean_field}
-  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-  R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-  \label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-% 
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
-% This is true for stochastic dynamics.
-\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-  As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-% 
-% 
-% However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+%%
+%The sampled biased distribution writes
+%%
+%\begin{align}
+%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
+%\label{eq:Biased_Path_Approximation}
+%\end{align}
+%where
+%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
+%refers formally to the probability of observing the trajectory
+%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
+%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+%%
+%One can mention that
+%\begin{equation}
+%\label{eq:mean_field}
+%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
+%\end{equation}
+%with $\mathbb{E}_{0}$ being the expectation value with respect to the
+%distribution $\mathbb{P}_{0}$.
+%This result relies on the \textit{mean-field approximation}
+%\begin{equation}
+%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
+%\label{eq:Mean_Field_Approximation}
+%\end{equation}
+%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
+%%
+%Rejected trajectories are discarded from the statistics.
+%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+%
+%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
+%%This is true for stochastic dynamics.
+%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+%%
+%%
+%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
 We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-% 
-% 
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-\label{sec:gktl_drag}
+%%
+%%
+%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+%%
+
+%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
+%\label{sec:gktl_drag}
 
@@
 %
-% Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-% In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
+%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
+%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
 %
@@
 \rho_k(F)
-  % = \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
+%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
 \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
@@
 As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-% Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
+%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
 %
@@
 Note that trajectories that overlap for more than half of their duration are counted as one.
-% \footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
+%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
 Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
 %
-% This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-% 
-% \begin{figure}
+%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
+%
+%\begin{figure}
 %  \centering
@@
 %  \label{fig:PDF_AVG}
-% \end{figure}
+%\end{figure}
 %
@@
 %
-% In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
 By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
@@
 \label{sec:return_times}
+
+\ZZ{Should insist on the efficiency of GKTL to compute return times. The formula may be moved to the appendix?}
+\bigskip
+
 Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
 events unreachable by a direct approach.
-% (see Fig.~\ref{fig:illustrate_return_time})
+%(see Fig.~\ref{fig:illustrate_return_time})
 
@@
 \label{conlusion}
-In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
+
+%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
 
@@
 Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
+%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
 
@@
 This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent.
-% following the addition of a small perturbation in the initial conditions.
+%following the addition of a small perturbation in the initial conditions.
 The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
@@
 
+\ZZ{
+	\section{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+	\label{app:AMS_on_OU}
+
+	\begin{figure}
+		\centering
+		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
+		\label{fig:comparaison_temps_de_retour}
+	\end{figure}
+	Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+	\begin{equation}
+	\label{eq:ou}
+	\dot{x} = -x + \eta (t),
+	\end{equation}
+	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+	This basic stochastic process will allow us to highlight differences with fluid dynamics.
+
+	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+	The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+	%
+	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+	%
+	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
+	In the latter, the typical computational cost is simply the return time $r(a)$.
+	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+
+	Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+	The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
+}
+
+\ZZ{
+\section{The \ac{gktl} algorithm}
+\label{app:gktl_description}
+The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%
+Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
+%
+At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
+%
+\begin{equation}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+\label{eq:Weight}
+\end{equation}
+so that $\sum_{n=1}^N W_n^i = N$.
+%
+%
+{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
+The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
+The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+%
+The sampled biased distribution writes
+%
+\begin{align}
+\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
+\label{eq:Biased_Path_Approximation}
+\end{align}
+where
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
+refers formally to the probability of observing the trajectory
+$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
+The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+%
+One can mention that
+\begin{equation}
+\label{eq:mean_field}
+Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
+\end{equation}
+with $\mathbb{E}_{0}$ being the expectation value with respect to the
+distribution $\mathbb{P}_{0}$.
+This result relies on the \textit{mean-field approximation}
+\begin{equation}
+R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
+\label{eq:Mean_Field_Approximation}
+\end{equation}
+where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
+%
+Rejected trajectories are discarded from the statistics.
+Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+
+A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
+%This is true for stochastic dynamics.
+\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+%
+%
+%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
+%
+%
+Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+%
+}
+
 \EL{
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2020-01-28 jfm/draft_p2_jfm.tex
Re-ecriture de la partie sur les algorithmes

@@
 Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-\smallskip
-
-\ZZ{The next two paragraphs may be moved to the introduction of section \ref{sec:rare_events_algorithms}?}
-\smallskip
-
-The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}.
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
-	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
 %
@@
 
-\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience.
-Technical details are either removed (reference to a companion paper) or moved to Appendix}
-
-\section{Rare-event algorithms}
+\section{Rare-events algorithms}
 \label{sec:rare_events_algorithms}
 
-% limitation of direct sampling
-%
-
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than their related return times.
-%
-
-\ZZ{Should develop here the idea behind rare event sampling.
-How it works.
-Include here the paragraphs of the introduction.
-Justify the choice of ams and gktl: what are differences and complementarity.
-AMS = instantaneous signal, GKTL= time-averaged signal.
-Why are they representative?
-What are the expectations?}
-
-%
-For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
-%
-%
-Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
-In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-%In a nutshell
-Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
-These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables.
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
+According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
+$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
+The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.
+
+In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
+computational cost much lower than their return time.
+Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
+greater complexity than the two-dimensional flow discussed in the previous sections.
+
+We focus on two algorithms, representatives of two popular approaches in rare event simulation.
+The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
+More recently, \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
+This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
+Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
+More recently it allowed the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
+This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
+Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.
+
+The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
+Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
+system.
+Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
+The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
+influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
+Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
+committor function. However, with the exception of rare analytical cases, this function is unknown.
+See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.
+
+Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
+However, resampling steps are performed along the simulation of the ensemble.
+Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
+according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
+This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
+Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
+\begin{align}
+\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
+\label{eq:Biased_Path_Approximation_main}
+\end{align}
+where
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory 
+$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
+See appendix~\ref{app:gktl_description} for a more detailed description of the cloning algorithm.
+
+Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
+rules to bias the sampling towards extreme events.
+In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
+events of interest.
+Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.
+
+Most previous applications of the \ac{ams} and cloning algorithm target stochastic dynamics.
+We stress that this is not the case in this work, as the target flow model is fully deterministic.
+Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
+The amplitude of this perturbation is chosen very low, so as not to impact the statistics of the dynamics.
+The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
+For further details about the addition of this perturbation, see appendix~\ref{sec:app_perturbation}.
+
+\subsection{The \acl{ams} algorithm}
 \label{sec:ams}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
-Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-\end{figure}
-
-% Introduction of AMS
-%
-The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \citep{KahnHarris1951}.
-The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into
-%a sequence of events that can be sampled with a higher probability
-a sequence of shorter excursions \citep{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
-The operating principle is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years.
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function during the time interval $0\leq t \leq T_a$.
-Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
-{Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-It can however be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-The following results are not limited to this particular variant.}
-
-\bigskip
-
-\ZZ{The Ornstein-uhlenbeck application has been moved to appendix \ref{app:AMS_on_OU}. Enough to mention in the text that AMS works successfully for OU and refer to apppendix?}
-
-\bigskip
-
-% SUBSECTION IS MOVED TO APPENDIX %
-
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding \EL{computational cost $C_{AMS}$}.
-The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-Our aim is here to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
-In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.
-%
-{Let us also stress that the dynamics is here deterministic.
-Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
-The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow \EL{(see Appendix \ref{app:perturb_branching_time})}.
-It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
-
-% application
-%
-The \ac{ams} algorithm is applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%
-Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
-In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{ams} re-sampling procedure.
-%
-Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.
-%
-Interestingly, all trajectories ultimately originate from one unique initial trajectory.
-%
-This feature may be explained as follows.
-It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
+We applied the \ac{ams} algorithm, described in appendix\ref{sec:app_ams}, to the sampling of extreme fluctuations of the drag applied on the square obstacle immersed in the flow described in section~\ref{sec:test_flow}.
+We chose to set the score function as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
+Let us stress that we do not expect this choice to be close to the optimal "committor" function.
+Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
+into the physics that lead to the extreme events of interest.
+Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
+It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
+score function, a choice that is often the only available one.
+
+In addition to the score function, the parameters of the \ac{ams} are the number of trajectories $N$ and the duration of trajectories $T_a$.
+As detailed in appendix~\ref{app_ams}, the size of the ensemble $N$ governs the statistical error affecting the estimators based on the sampled set of trajectories.
+Furthermore, the number of resamplings required to reach a final ensemble of trajectories of probability of the order of $10^{-\beta}$ can be shown to scale like $\beta N$ \citep{lestang_computing_2018}.
+In this work, the choice of $N$ was governed by the available computational resources.
+We performed two experiments with $N = 32$ and $N = 256$, having in mind that higher numbers are unlikely to be accessible in the case of more complex or three-dimensional flows.
+The duration of the trajectories was set to $T_a$ = $5\tau_c$, a value that is larger than the timescale of extreme drag fluctuations whilst maintaining an affordable computational cost.
+As pointed out in appendix~\ref{app_ams}, setting $T_a \gg \tau_c$ does not lead to a better sampling.
+
+\begin{figure}
+  \centering
+  \subfloat[\label{fig:AMS_drag_trajectories_a} Final ensemble of trajectories sampled by the \ac{ams}]{\includegraphics[width=.4\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}}
+  \subfloat[Drag maximum for resampled trajectories]{\label{fig:AMS_drag_trajectories_b}\includegraphics[width=.4\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}}
+  \caption{(a) Ensemble of $N = 32$ trajectories after $181$ \ac{ams} resamplings.
+    (b) Maximum of the instantaneous drag throughout the resampled trajectories as a function of the corresponding computational cost $C_{AMS}$, \textit{i.e.} as a function of the resamplings.
+The maximum over resampled trajectories overlap until saturating at the level of the largest maximum in the initial ensemble.
+  }
+\end{figure}
+
+Figure~\ref{fig:AMS_drag_trajectories_a} illustrates that the splitting introduced by the \ac{ams} does not enhance the sampling of extremes but rather has catastrophic consequences on the sampling of trajectories.
+Indeed, trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.
+This result is confirmed by increasing the number of initial trajectories from $N=32$ to $N=256$, as shown
+in figure~\ref{fig:AMS_drag_trajectories_b}.
+Recalling section~\ref{sec:instantaneous_drag}, extreme drag fluctuations have a lifetime of $\tau_c$, the timescale over which a vortex is blocked near the base of the obstacle.
+After $\tau_c$, the vortex is freed and swept downstream, and further large drag fluctuations can only result from the interaction of the obstacle with the incoming flow, beyond the correlation time of the drag fluctuations.
+Moreover, it takes a certain time $\tau_L$ (known as the Lyapunov timescale) before a trajectory resampled by the \ac{ams} separates from its parent under the effect of the initial perturbation.
 In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
 %
-We observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations, $\tau_c$, is much shorter than the Lyapunov's timescale $\tau_L$.
-As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and therefore the probability of observing a new extreme fluctuation is also very low.
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L \gtrsim \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-A possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.}
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically.
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-
-\bigskip
-\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-
-\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
-%refers formally to the probability of observing the trajectory
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%%
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is now considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-The purpose is to sample trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.
-{Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.}
-
-% Choice of parameters and perturbation
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-In the following experiments, $\tau = \tau_c /2$.
-%
-
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-\begin{equation}
-\label{eq:estimate_biased_measure}
-\rho_k(F)
-%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-\end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-%
-% limitations
-%
-Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap.
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
-%
-Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\
-		\hline
-		${N}_{brute-force}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling.
-Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
+We observe in Fig.~\ref{fig:AMS_drag_trajectories_a} that this separation timescale is larger than the duration of extreme drag fluctuations $\tau_c$, leading to an overlap of trajectories.
+
+We stress that this phenomenology is independent of the choice of $N$ and $T_a$.
+Increasing the size of the initial ensemble can only increase the amplitude of the the global maximum attained initially, but does not solve the issue of overlapping trajectories.
+Increasing $T_a$ allows further time for resampled trajectories to develop rare fluctuations, however after a duration $\Delta T \geq \tau_L > \tau_c$, and therefore decorrelated from the resampling.
+
+Our results indicate that the \ac{ams} cannot be straightforwardly applied to convective flows past obstacles,
+in which the timescale of extreme events (here a drag fluctuation) is smaller than the duration required
+for trajectories to separate under the influence of turbulence.
+The the author's knowledge, this result is the first evidence of the limitations of splitting approaches such as the \ac{ams} for complex systems, for which the only available choice for the score function displays a memory.
+\textbf{Do we agree on this last paragraph?}
+
+\section{The cloning algorithm}
+\label{sec:cloning}
+We applied the cloning algorithm described, in appendix\ref{sec:app_gktl}, to the two-dimensional flow described in section~\ref{sec:test_flow}.
+Unlike the case of the \ac{ams} in section~\ref{sec:ams}, the algorithm is here used to sampled trajectories that exhibit extremes values of
+the \emph{average} drag.
+
+The cloning algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
+The cloning strength is the multiplicative factor appearing the the bias term of equation~\eqref{eq:Biased_Path_Approximation_main}.
+The higher $k$, the larger the typical average drag for the trajectories in the ensemble sampled by the algorithm.
+There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.
+Similarly to the \ac{ams}, $N$ governs the error affecting estimates of probabilities and estimators based on the biased ensemble.
+In the cloning algorithm, the population of trajectories is periodically resampled as the trajectories are simulated from $t=0$ to
+$t=T_a$.
+The cloning period determines how often this resampling step is performed.
+For deterministic dynamics, a common rule of thumb is to chose $\tau \approx \tau_c$ \citep{giardina_direct_2006}.
+For further details about the cloning algorithm, see appendix~\ref{sec:app_gktl}.
+
+In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
+Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
+Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
+Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
+Finally, we set $\tau = \tau_c / 2$.
+\textbf{How do we set k?}
+
+Figure~\ref{sec:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
+It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
+The higher $k$, the stronger the bias.
+The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
+We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:pdf_avg}).
+Figure~\ref{sec:IS_GKTL} also illustrates that the sampling by the algorithm is very concentrated on some fluctuations values, and that this effect increases as $k$ increases.
+This results from the overlap of trajectories as a consequence of the cloning and pruning.
+Indeed, if a trajectory is cloned at step $i$, then the trajectory ${\mathbf{x}(t)}$ over the interval $[0;i\tau]$ will be the same for all the clones.
+For a fixed ensemble size $N$, the overlap of trajectories increases as $k$ increases as fewer
+trajectories are cloned into a higher number of duplicates.
+\textbf{How do we quantify this effect? It is probably application-dependant. What are the strategies to mitigate it?}
+
+Figure~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag timeseries for several trajectories within the ensemble sampled by the cloning algorithm.
+It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
+Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
+This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
+  \caption{Drag timeseries corresponding to the 6 highest fluctuations of the average drag in the ensemble sampled by the cloning algorithm.
+The cloning algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
+}
+  \label{fig:timeseries_extrms_AVG_GKTL}
+\end{figure}
+
+In addition, figure~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
+sampled by the cloning algorithm.
+It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
+Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
+  \caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
+  \label{fig:illustr_extrms_vorticity_GKTL}
+\end{figure}
+
+An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
+Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
+dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
+In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
+A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
 
@@
 
-\ZZ{Should insist on the efficiency of GKTL to compute return times. The formula may be moved to the appendix?}
-\bigskip
-
-Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
+Beyond dynamical aspects, the cloning algorithm provides ensemble of trajectories over which statistics of rare events can be computed by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}.
+In this section, we show that the biased sampling performed by the cloning algorithm allows for the computation of return times for events otherwise out-of-reach from a direct sampling approach.
+
+Each trajectory $j$ in the biased ensemble results in a timeseries of the time-averaged drag:
 \begin{equation}
 \label{eq:time_averaged}
-F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
+F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  ,
+\end{equation}
+and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
 \begin{equation}
@@
 
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
+The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}
+\begin{equation}
+  \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{kT_aF_T^{(j)}}s_j(a),
+% \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d^{(j)}(t)dt}s_j(a)
+\end{equation}
+with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise. This amounts to taking the sum of the weights of the timeseries which maximum is larger than $a$.
 
 Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independent estimates, obtained using different values of the bias parameter $k$.
+Two independent estimates are shown, obtained using different values of the cloning strength $k$.
 Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, \EL{Fig.~\ref{fig:return_times_gktl}} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
+Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the cloning algorithm allows for the computation of statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+This result shows that the biased ensemble sampled by cloning algorithm enables the estimation of return times for average drag fluctuations that would be out-of-reach from a direct sampling.
+Alternatively, for a fixed target return time, the use of the cloning algorithm can reduce the computational cost of the estimation by several orders of magnitude.
 This is obviously a major advantage of this rare-event sampling algorithm.
+
+More generally, the successful computation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the cloning algorithm may be used to compute statistical estimators in a very efficient manner.
+We note, however, that some estimators may be more affected by the quality of the sampling than others.
 

### 2020-01-28 jfm/draft_p2_jfm.tex
Add new figure IS_GKTL_2

@@
 We shall consider $T=10\tau_c$ in the following.
-\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:IS_GKTL}).}
+\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig}).}
 
@@
 
-Figure~\ref{sec:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{IS_GKTL/IS_GKTL_2}
+  \caption{Histogram of the values of the time-averaged drag $F_{T_a}$, with $T_a = 10\tau_c$, in the ensemble sampled by the cloning algorithm, for three values of the cloning strength $k$.
+In each case, the algorithm was applied with $N=1024$, $\tau = \tau_c / 2$.
+The grey histogram represents the average histogram obtained through a direct sampling, \textit{i.e.} with $k=0$.
+It was computed using a Gaussian approximation for the \ac{pdf} of the time-averaged drag.}
+  \label{fig:IS_GKTL}
+\end{figure}
+Figure~\ref{fig:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
 It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.

### 2020-02-04 jfm/appendix_ams.tex
Add appendix on AMS

\newtheorem{theo}{Theorem}

This appendix describes the \ac{ams} algorithm, gives a brief summary of the main
mathematical results and discusses its use in practice.
\subsection{Description of the algorithm}
We consider an ensemble of $N$ independent trajectories $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$, for a fixed duration $T_a$.
These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.

Starting from this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.

To start off with, each of these trajectories is associated a weight $w_0=1$.
Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
\begin{equation}
  \mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
\end{equation}
Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
\begin{equation}
  \label{eq:ams_threshold_def}
  \mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
\end{equation}
One might expect intuitively that $\ell_j=1$.
However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{Brehier2016a}.
The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
  \caption{Illustration of one iteration the \ac{ams} algorithm. Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$. Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1. Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
  \label{fig:AMS_schema}
\end{figure}

After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
$\tilde{J} = \sum_{j=1}^J \ell_j$.
Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
\subsection{How to compute the probability}
Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,Brehier2016a}:
\begin{equation}
  w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
\end{equation}
This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
\subsection{Summary of mathematical results}
We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
of iterations $N$.
We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{ams_threshold_def}.
The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
\begin{equation}
  \label{eq:estimator_ams}
  q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
\end{equation}
In the following we denote this estimator $\hat{q}_N(a)$.
One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{Brehier2016a} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
\begin{theo}
  For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
  \begin{equation}
    \mathbb{E}[\hat{q}_N]=q.
  \end{equation}
\end{theo}
Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{Brehier2016a,Rolland2015} for extensive numerical simulations concerning the role of the score function.

More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
\begin{equation}
  \sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
\end{equation}
with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
The minimal variance $-q^2\ln q$ is obtained when choosing
\begin{equation}
  \bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
  \label{eq:time_dependent_committor}
\end{equation}
for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.

\subsection{How to use it in practice}
In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.

Another parameter of the algorithm is the duration $T_a$.
This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
This can be explained as follows.
Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
This amounts to direct sampling.

Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
Rather, iterations were carried out until all the trajectories in the ensemble overlap.
Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
\begin{equation}
  \mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
\end{equation}
and the iterations stop.
This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End:

### 2020-02-04 jfm/draft_p2_jfm.tex
Add pdf
Add appendix on AMS

@@
 Finally, we set $\tau = \tau_c / 2$.
-\textbf{How do we set k?}
+\textbf{Thibault: Explain the choice of k}
 
@@
 Namely, an \emph{exact} streaming step
-\[
+\begin{equation}
+  \label{eq:lbe}
   f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
+\end{equation}
 during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
@@
 The equilibrium function is given by
-\[
+\begin{equation}
+  \label{eq:lbe_eq}
   f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\]
+    \frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right),
+\end{equation}
 with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice.
@@
 
+\include{appendix_ams}
+
 \EL{

### 2020-02-05 jfm/appendix_perturbation.tex
Add appendix decribing perturbation at branching time

As mentioned in sections~\ref{sec:ams} and~\ref{sec:cloning}, the application of the \ac{ams} and the cloning algorithm to deterministic systems require the the addition of a perturbation to the initial condition of
resampled trajectories.
In the absence of any perturbation, the resampling would not yield new trajectories, but instead exact copies of the trajectories the resampling is based on.

The state of the flow at any point in time is given by the velocity field and the pressure $(\mathbf{u}, p)$, where the dependence on time of both variables is implicit.
A general strategy for perturbing the state of the flow is to construct a perturbative divergence free field $\delta \mathbf{u}$ and corresponding $\delta p$ solution of the
Poisson equation.
In the context of the \ac{lbm} however, the state of the system is described at a mesoscopic level by the ensemble the populations $\{f_k(x_j, y_j)\}_{0\le k \le 8, 0 \le i <N_x, 0 \le j < N_y}$.
A consequence is that, from the knowledge of a macroscopic state ($(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$, it is difficult to obtain the mesoscopic state from with the \ac{lbm} model can be initialised.

In this work we work around this issue by designing the perturbative directly at the level of the mesoscopic populations.
At each lattice site $\mathbf{x}$, the populations ${f_i}_{0\le i \le 8}$ are replaced by
\begin{equation}
  f_{i}(\mathbf{x}) \longrightarrow f_i(\mathbf{x}) + \epsilon \sum_{n=1}^{N} \alpha_{n}f_{i}^{n}(\mathbf{x}), \,\,\, 1\leq i \leq 9,
  \label{eq:perturb_pop}
\end{equation}
with the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the amplitude of the perturbation. The perturbed populations are then rescaled so that no mass is added in the system.
For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}
In other words obtained as a linear combination of $N$ solutions $\{f_i^{(n)}\}_{1\le n \le N}(\mathbf{x})$ of the numerical model~\eqref{eq:lbe}.
Note that, the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions can be assimilated as a solution of the numerical model.
However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq}, and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
  \caption[Illustration of the perturbation of the initial condition]{The blue line shows the evolution of the drag acting on the obstacle over time for a simulation initialised on a given initial condition. The second line represents the evolution of the drag for a simulation initialised on the same initial condition, however slightly perturbed as described in this appendix. $N=10$, $\epsilon=0.002$}
  \label{fig:effect_of_perturbation}
\end{figure}

In this work, we chose $N = 10$ and $\epsilon = 0.002$.
Figure~\ref{fig:effect_of_perturbation} illustrate the effect of applying a small perturbation on the evolution of the drag acting on the square obstacle.
The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with $\epsilon = 0.002$ with a period $\tau = \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
Figure~\ref{fig:pdf_drag_with_perturbation} shows that the correct statistics are obtained even in presence of this slight perturbation.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
  \caption{Estimation of the \ac{pdf} with and without periodic perturbations}
  \label{fig:pdf_drag_with_perturbation}
\end{figure}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End:

### 2020-02-05 jfm/introduction.tex
Add introduction in separate tex file

--- 2020-01-28 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-
-\title{\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\ZZ{This study investigates by means of direct numerical simulation, extreme drag events exerted by a  turbulent flow on a bluff body and examine the relevance of statistical algorithms to sample efficiently these events.
-Two representative algorithms are examined: AMS and GKTL.
-Briefly mention differences and complementarity.
-What do we learn?
-The take-home message is...}
-\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems.
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
 \section{Introduction}
@@
 %
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures.
+A characteristic feature of turbulent flows is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
+By ``extreme'' we refer to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
+In engineering, better understanding of the nature of such extreme dynamical events and their statistics is crucial for the control of excessive mechanical efforts, for instance extreme wind loads on tall structures such as buildings of wind turbines~\cite{kanev2019}.
 %
@@
 From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Otherwise, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
+However, the use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
+Alternatively, numerical simulation offers a convenient approach to gain physical insight into these events, as well as quantifying their intensity and estimating their frequency of occurrence.
+The main issue with this approach is however the very large durations over which the flows must be simulated in order to sample extreme events, since these are rare events.
+The majority of flows relevant to industrial or environmental problems are three-dimensional and involve complex shapes as well as large Reynolds numbers.
+In this context, tremendous computational requirements make the sole accurate simulation of the flow a challenge, and its computation over very long durations is impractical.
+
 % general def on rare-event sampling %
 %
-{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+More generally, the computational cost for sampling a fluctuation of very small probability grows like the inverse of this probability, and simulating rare events through direct numerical simulation is therefore not a scalable approach.
+This issue is addressed by rare-event sampling algorithms, referring to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.
 %
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
+Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
+More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
 
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducted.
+We deliberately focus on a simple two-dimensional flow in order to allow for its simulation over very large durations and sampling of extreme events.
+Furthermore, we consider the application of rare-event sampling algorithms to reduce the computational cost associated to the sampling of these events.
+The motivation for this work is twofold.
+First, we provide a detailed description of the statistics and dynamics of rare events related to extreme
+drag forces acting on a square placed in a two-dimensional channel flow.
+This cannot be achieved based on more complex (and realistic) systems.
+Although our analysis is based on a simplified geometry, it is important to fluid dynamics because it paves the way to a better understanding of this type of events.
+To the best of the authors' knowledge, such analysis has never been reported before.
+Second, we assess the relevance of rare-event sampling approaches for this problem.
 In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
@@
 %
-%
 In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
@@
 %
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
+An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
 %
-%
-\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-}
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-%
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
+Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the \ac{ams} algorithm and the \ac{gktl} algorithm.
 
 % Annonce du plan
-In more detail,  the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
-	drag fluctuations} is investigated based on a direct sampling approach.
+This paper is organised in two main parts.
+Section~\ref{sec:direct_sampling} describes the phenomenology of extreme drag events based on the simulation of the flow over a very long duration.
+Section~\ref{sec:rare_events_algorithms} presents the application of two rare-events algorithms to the same
+problem.
+The flow set-up is introduced in section~\ref{sec:test_flow} and the statistical properties of the drag are discussed.
+In section~\ref{sec:direct_sampling}, the phenomenology of  extreme drag fluctuations is investigated based on a direct sampling approach.
 Both the instantaneous drag and time-averaged drag are considered.
@@
 Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
 Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
@@
 Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
+This latter allows for an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
+As a specific successful application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
 Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}).
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-%
-We believe that this system embeds the characteristic features that makes \sout{the application of rare-event algorithms} \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics.
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square.
-%
-%Finally, the obstacle does not deform or move either.
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
-%
-
-% lattice Boltzmann method
-%
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations.
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition.
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag.
-As a consequence, the drag can eventually be expressed as the difference
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-	\label{fig:typical_drag_signal}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle.
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}
-	\caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
-The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		\textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}.
-One can then argue that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-			One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent.
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that
-\begin{equation}
-\label{eq:return_time}
-\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		As the amplitude of the drag increases, the relative variation of the base pressure also increases.
-\EL{The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
-
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
-%relatively to the forebody pressure.
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag.
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time.
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics.
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base.
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-%
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown.
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, a relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
-\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:IS_GKTL}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
-Indeed, both cases are %equally
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-\bigskip
-
-\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-\bigskip
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-\section{Rare-events algorithms}
-\label{sec:rare_events_algorithms}
-
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
-$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.
-
-In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
-computational cost much lower than their return time.
-Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
-greater complexity than the two-dimensional flow discussed in the previous sections.
-
-We focus on two algorithms, representatives of two popular approaches in rare event simulation.
-The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently, \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
-This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
-Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-More recently it allowed the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
-This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.
-
-The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
-system.
-Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
-The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
-influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
-Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
-committor function. However, with the exception of rare analytical cases, this function is unknown.
-See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.
-
-Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
-However, resampling steps are performed along the simulation of the ensemble.
-Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
-according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
-This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
-Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation_main}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
-See appendix~\ref{app:gktl_description} for a more detailed description of the cloning algorithm.
-
-Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
-rules to bias the sampling towards extreme events.
-In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
-events of interest.
-Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.
-
-Most previous applications of the \ac{ams} and cloning algorithm target stochastic dynamics.
-We stress that this is not the case in this work, as the target flow model is fully deterministic.
-Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
-The amplitude of this perturbation is chosen very low, so as not to impact the statistics of the dynamics.
-The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
-For further details about the addition of this perturbation, see appendix~\ref{sec:app_perturbation}.
-
-\subsection{The \acl{ams} algorithm}
-\label{sec:ams}
-We applied the \ac{ams} algorithm, described in appendix\ref{sec:app_ams}, to the sampling of extreme fluctuations of the drag applied on the square obstacle immersed in the flow described in section~\ref{sec:test_flow}.
-We chose to set the score function as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
-Let us stress that we do not expect this choice to be close to the optimal "committor" function.
-Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
-into the physics that lead to the extreme events of interest.
-Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
-It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
-score function, a choice that is often the only available one.
-
-In addition to the score function, the parameters of the \ac{ams} are the number of trajectories $N$ and the duration of trajectories $T_a$.
-As detailed in appendix~\ref{app_ams}, the size of the ensemble $N$ governs the statistical error affecting the estimators based on the sampled set of trajectories.
-Furthermore, the number of resamplings required to reach a final ensemble of trajectories of probability of the order of $10^{-\beta}$ can be shown to scale like $\beta N$ \citep{lestang_computing_2018}.
-In this work, the choice of $N$ was governed by the available computational resources.
-We performed two experiments with $N = 32$ and $N = 256$, having in mind that higher numbers are unlikely to be accessible in the case of more complex or three-dimensional flows.
-The duration of the trajectories was set to $T_a$ = $5\tau_c$, a value that is larger than the timescale of extreme drag fluctuations whilst maintaining an affordable computational cost.
-As pointed out in appendix~\ref{app_ams}, setting $T_a \gg \tau_c$ does not lead to a better sampling.
-
-\begin{figure}
-  \centering
-  \subfloat[\label{fig:AMS_drag_trajectories_a} Final ensemble of trajectories sampled by the \ac{ams}]{\includegraphics[width=.4\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}}
-  \subfloat[Drag maximum for resampled trajectories]{\label{fig:AMS_drag_trajectories_b}\includegraphics[width=.4\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}}
-  \caption{(a) Ensemble of $N = 32$ trajectories after $181$ \ac{ams} resamplings.
-    (b) Maximum of the instantaneous drag throughout the resampled trajectories as a function of the corresponding computational cost $C_{AMS}$, \textit{i.e.} as a function of the resamplings.
-The maximum over resampled trajectories overlap until saturating at the level of the largest maximum in the initial ensemble.
-  }
-\end{figure}
-
-Figure~\ref{fig:AMS_drag_trajectories_a} illustrates that the splitting introduced by the \ac{ams} does not enhance the sampling of extremes but rather has catastrophic consequences on the sampling of trajectories.
-Indeed, trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.
-This result is confirmed by increasing the number of initial trajectories from $N=32$ to $N=256$, as shown
-in figure~\ref{fig:AMS_drag_trajectories_b}.
-Recalling section~\ref{sec:instantaneous_drag}, extreme drag fluctuations have a lifetime of $\tau_c$, the timescale over which a vortex is blocked near the base of the obstacle.
-After $\tau_c$, the vortex is freed and swept downstream, and further large drag fluctuations can only result from the interaction of the obstacle with the incoming flow, beyond the correlation time of the drag fluctuations.
-Moreover, it takes a certain time $\tau_L$ (known as the Lyapunov timescale) before a trajectory resampled by the \ac{ams} separates from its parent under the effect of the initial perturbation.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories_a} that this separation timescale is larger than the duration of extreme drag fluctuations $\tau_c$, leading to an overlap of trajectories.
-
-We stress that this phenomenology is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble can only increase the amplitude of the the global maximum attained initially, but does not solve the issue of overlapping trajectories.
-Increasing $T_a$ allows further time for resampled trajectories to develop rare fluctuations, however after a duration $\Delta T \geq \tau_L > \tau_c$, and therefore decorrelated from the resampling.
-
-Our results indicate that the \ac{ams} cannot be straightforwardly applied to convective flows past obstacles,
-in which the timescale of extreme events (here a drag fluctuation) is smaller than the duration required
-for trajectories to separate under the influence of turbulence.
-The the author's knowledge, this result is the first evidence of the limitations of splitting approaches such as the \ac{ams} for complex systems, for which the only available choice for the score function displays a memory.
-\textbf{Do we agree on this last paragraph?}
-
-\section{The cloning algorithm}
-\label{sec:cloning}
-We applied the cloning algorithm described, in appendix\ref{sec:app_gktl}, to the two-dimensional flow described in section~\ref{sec:test_flow}.
-Unlike the case of the \ac{ams} in section~\ref{sec:ams}, the algorithm is here used to sampled trajectories that exhibit extremes values of
-the \emph{average} drag.
-
-The cloning algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The cloning strength is the multiplicative factor appearing the the bias term of equation~\eqref{eq:Biased_Path_Approximation_main}.
-The higher $k$, the larger the typical average drag for the trajectories in the ensemble sampled by the algorithm.
-There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.
-Similarly to the \ac{ams}, $N$ governs the error affecting estimates of probabilities and estimators based on the biased ensemble.
-In the cloning algorithm, the population of trajectories is periodically resampled as the trajectories are simulated from $t=0$ to
-$t=T_a$.
-The cloning period determines how often this resampling step is performed.
-For deterministic dynamics, a common rule of thumb is to chose $\tau \approx \tau_c$ \citep{giardina_direct_2006}.
-For further details about the cloning algorithm, see appendix~\ref{sec:app_gktl}.
-
-In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
-Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
-Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
-Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
-Finally, we set $\tau = \tau_c / 2$.
-\textbf{How do we set k?}
-
-Figure~\ref{sec:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
-It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
-The higher $k$, the stronger the bias.
-The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
-We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:pdf_avg}).
-Figure~\ref{sec:IS_GKTL} also illustrates that the sampling by the algorithm is very concentrated on some fluctuations values, and that this effect increases as $k$ increases.
-This results from the overlap of trajectories as a consequence of the cloning and pruning.
-Indeed, if a trajectory is cloned at step $i$, then the trajectory ${\mathbf{x}(t)}$ over the interval $[0;i\tau]$ will be the same for all the clones.
-For a fixed ensemble size $N$, the overlap of trajectories increases as $k$ increases as fewer
-trajectories are cloned into a higher number of duplicates.
-\textbf{How do we quantify this effect? It is probably application-dependant. What are the strategies to mitigate it?}
-
-Figure~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag timeseries for several trajectories within the ensemble sampled by the cloning algorithm.
-It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-  \caption{Drag timeseries corresponding to the 6 highest fluctuations of the average drag in the ensemble sampled by the cloning algorithm.
-The cloning algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-  \label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-
-In addition, figure~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
-sampled by the cloning algorithm.
-It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-  \caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
-  \label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-Beyond dynamical aspects, the cloning algorithm provides ensemble of trajectories over which statistics of rare events can be computed by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}.
-In this section, we show that the biased sampling performed by the cloning algorithm allows for the computation of return times for events otherwise out-of-reach from a direct sampling approach.
-
-Each trajectory $j$ in the biased ensemble results in a timeseries of the time-averaged drag:
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  ,
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}
-\begin{equation}
-  \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{kT_aF_T^{(j)}}s_j(a),
-% \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d^{(j)}(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise. This amounts to taking the sum of the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-Two independent estimates are shown, obtained using different values of the cloning strength $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the cloning algorithm allows for the computation of statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This result shows that the biased ensemble sampled by cloning algorithm enables the estimation of return times for average drag fluctuations that would be out-of-reach from a direct sampling.
-Alternatively, for a fixed target return time, the use of the cloning algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is obviously a major advantage of this rare-event sampling algorithm.
-
-More generally, the successful computation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the cloning algorithm may be used to compute statistical estimators in a very efficient manner.
-We note, however, that some estimators may be more affected by the quality of the sampling than others.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent.
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\]
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice.
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-\ZZ{
-	\section{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-	\label{app:AMS_on_OU}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-	The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-}
-
-\ZZ{
-\section{The \ac{gktl} algorithm}
-\label{app:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
-refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
-%This is true for stochastic dynamics.
-\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-%
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-}
-
-\EL{
-	\section{Perturbation at branching time}
-	\label{app:perturb_branching_time}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}
 
@@
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "draft_p2_jfm"
 %%% End:

### 2020-02-05 jfm/draft_p2_jfm.tex
Update final paragraph on AMS

@@
 
-Our results indicate that the \ac{ams} cannot be straightforwardly applied to convective flows past obstacles,
-in which the timescale of extreme events (here a drag fluctuation) is smaller than the duration required
-for trajectories to separate under the influence of turbulence.
-The the author's knowledge, this result is the first evidence of the limitations of splitting approaches such as the \ac{ams} for complex systems, for which the only available choice for the score function displays a memory.
-\textbf{Do we agree on this last paragraph?}
+Our results indicate that the \ac{ams} does not work with this simple choice of score function.
+Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.
 

### 2020-02-05 jfm/appendix_ams.tex
Add OU example to appendix on AMS

@@
 \newtheorem{theo}{Theorem}
+
+\section{The \ac{ams} algorithm}
+\label{app_ams}
 
@@
 This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
+
+\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+\label{app:AMS_on_OU}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+  \caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
+  \label{fig:comparaison_temps_de_retour}
+\end{figure}
+Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+\begin{equation}
+  \label{eq:ou}
+  \dot{x} = -x + \eta (t),
+\end{equation}
+where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+This basic stochastic process will allow us to highlight differences with fluid dynamics.
+
+The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+
+The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
+In the latter, the typical computational cost is simply the return time $r(a)$.
+One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+
+Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
+
 %%% Local Variables:

### 2020-02-05 jfm/appendix_perturbation.tex
Add section into the appendix on the perturbation

@@
+\section{Perturbation at branching time}
+\label{app:perturb_branching_time}
 As mentioned in sections~\ref{sec:ams} and~\ref{sec:cloning}, the application of the \ac{ams} and the cloning algorithm to deterministic systems require the the addition of a perturbation to the initial condition of
@@
 \end{figure}
+
 %%% Local Variables:

### 2020-02-05 jfm/appendix_lbm.tex
Move all appendices and introduction into separate .tex files

--- 2020-02-05 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-
-\title{\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\ZZ{This study investigates by means of direct numerical simulation, extreme drag events exerted by a  turbulent flow on a bluff body and examine the relevance of statistical algorithms to sample efficiently these events.
-Two representative algorithms are examined: AMS and GKTL.
-Briefly mention differences and complementarity.
-What do we learn?
-The take-home message is...}
-\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems.
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Otherwise, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
-%
-%
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-}
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-%
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-In more detail,  the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
-	drag fluctuations} is investigated based on a direct sampling approach.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}).
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-%
-We believe that this system embeds the characteristic features that makes \sout{the application of rare-event algorithms} \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics.
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square.
-%
-%Finally, the obstacle does not deform or move either.
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
-%
-
-% lattice Boltzmann method
-%
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations.
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition.
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag.
-As a consequence, the drag can eventually be expressed as the difference
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-	\label{fig:typical_drag_signal}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle.
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}
-	\caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
-The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		\textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}.
-One can then argue that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-			One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent.
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that
-\begin{equation}
-\label{eq:return_time}
-\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		As the amplitude of the drag increases, the relative variation of the base pressure also increases.
-\EL{The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
-
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
-%relatively to the forebody pressure.
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag.
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time.
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics.
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base.
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-%
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown.
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, a relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
-\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
-Indeed, both cases are %equally
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-\bigskip
-
-\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-\bigskip
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-\section{Rare-events algorithms}
-\label{sec:rare_events_algorithms}
-
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
-$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.
-
-In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
-computational cost much lower than their return time.
-Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
-greater complexity than the two-dimensional flow discussed in the previous sections.
-
-We focus on two algorithms, representatives of two popular approaches in rare event simulation.
-The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently, \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
-This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
-Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-More recently it allowed the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
-This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.
-
-The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
-system.
-Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
-The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
-influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
-Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
-committor function. However, with the exception of rare analytical cases, this function is unknown.
-See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.
-
-Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
-However, resampling steps are performed along the simulation of the ensemble.
-Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
-according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
-This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
-Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation_main}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
-See appendix~\ref{app:gktl_description} for a more detailed description of the cloning algorithm.
-
-Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
-rules to bias the sampling towards extreme events.
-In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
-events of interest.
-Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.
-
-Most previous applications of the \ac{ams} and cloning algorithm target stochastic dynamics.
-We stress that this is not the case in this work, as the target flow model is fully deterministic.
-Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
-The amplitude of this perturbation is chosen very low, so as not to impact the statistics of the dynamics.
-The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
-For further details about the addition of this perturbation, see appendix~\ref{sec:app_perturbation}.
-
-\subsection{The \acl{ams} algorithm}
-\label{sec:ams}
-We applied the \ac{ams} algorithm, described in appendix\ref{sec:app_ams}, to the sampling of extreme fluctuations of the drag applied on the square obstacle immersed in the flow described in section~\ref{sec:test_flow}.
-We chose to set the score function as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
-Let us stress that we do not expect this choice to be close to the optimal "committor" function.
-Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
-into the physics that lead to the extreme events of interest.
-Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
-It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
-score function, a choice that is often the only available one.
-
-In addition to the score function, the parameters of the \ac{ams} are the number of trajectories $N$ and the duration of trajectories $T_a$.
-As detailed in appendix~\ref{app_ams}, the size of the ensemble $N$ governs the statistical error affecting the estimators based on the sampled set of trajectories.
-Furthermore, the number of resamplings required to reach a final ensemble of trajectories of probability of the order of $10^{-\beta}$ can be shown to scale like $\beta N$ \citep{lestang_computing_2018}.
-In this work, the choice of $N$ was governed by the available computational resources.
-We performed two experiments with $N = 32$ and $N = 256$, having in mind that higher numbers are unlikely to be accessible in the case of more complex or three-dimensional flows.
-The duration of the trajectories was set to $T_a$ = $5\tau_c$, a value that is larger than the timescale of extreme drag fluctuations whilst maintaining an affordable computational cost.
-As pointed out in appendix~\ref{app_ams}, setting $T_a \gg \tau_c$ does not lead to a better sampling.
-
-\begin{figure}
-  \centering
-  \subfloat[\label{fig:AMS_drag_trajectories_a} Final ensemble of trajectories sampled by the \ac{ams}]{\includegraphics[width=.4\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}}
-  \subfloat[Drag maximum for resampled trajectories]{\label{fig:AMS_drag_trajectories_b}\includegraphics[width=.4\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}}
-  \caption{(a) Ensemble of $N = 32$ trajectories after $181$ \ac{ams} resamplings.
-    (b) Maximum of the instantaneous drag throughout the resampled trajectories as a function of the corresponding computational cost $C_{AMS}$, \textit{i.e.} as a function of the resamplings.
-The maximum over resampled trajectories overlap until saturating at the level of the largest maximum in the initial ensemble.
-  }
-\end{figure}
-
-Figure~\ref{fig:AMS_drag_trajectories_a} illustrates that the splitting introduced by the \ac{ams} does not enhance the sampling of extremes but rather has catastrophic consequences on the sampling of trajectories.
-Indeed, trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.
-This result is confirmed by increasing the number of initial trajectories from $N=32$ to $N=256$, as shown
-in figure~\ref{fig:AMS_drag_trajectories_b}.
-Recalling section~\ref{sec:instantaneous_drag}, extreme drag fluctuations have a lifetime of $\tau_c$, the timescale over which a vortex is blocked near the base of the obstacle.
-After $\tau_c$, the vortex is freed and swept downstream, and further large drag fluctuations can only result from the interaction of the obstacle with the incoming flow, beyond the correlation time of the drag fluctuations.
-Moreover, it takes a certain time $\tau_L$ (known as the Lyapunov timescale) before a trajectory resampled by the \ac{ams} separates from its parent under the effect of the initial perturbation.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories_a} that this separation timescale is larger than the duration of extreme drag fluctuations $\tau_c$, leading to an overlap of trajectories.
-
-We stress that this phenomenology is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble can only increase the amplitude of the the global maximum attained initially, but does not solve the issue of overlapping trajectories.
-Increasing $T_a$ allows further time for resampled trajectories to develop rare fluctuations, however after a duration $\Delta T \geq \tau_L > \tau_c$, and therefore decorrelated from the resampling.
-
-Our results indicate that the \ac{ams} does not work with this simple choice of score function.
-Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.
-
-\section{The cloning algorithm}
-\label{sec:cloning}
-We applied the cloning algorithm described, in appendix\ref{sec:app_gktl}, to the two-dimensional flow described in section~\ref{sec:test_flow}.
-Unlike the case of the \ac{ams} in section~\ref{sec:ams}, the algorithm is here used to sampled trajectories that exhibit extremes values of
-the \emph{average} drag.
-
-The cloning algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The cloning strength is the multiplicative factor appearing the the bias term of equation~\eqref{eq:Biased_Path_Approximation_main}.
-The higher $k$, the larger the typical average drag for the trajectories in the ensemble sampled by the algorithm.
-There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.
-Similarly to the \ac{ams}, $N$ governs the error affecting estimates of probabilities and estimators based on the biased ensemble.
-In the cloning algorithm, the population of trajectories is periodically resampled as the trajectories are simulated from $t=0$ to
-$t=T_a$.
-The cloning period determines how often this resampling step is performed.
-For deterministic dynamics, a common rule of thumb is to chose $\tau \approx \tau_c$ \citep{giardina_direct_2006}.
-For further details about the cloning algorithm, see appendix~\ref{sec:app_gktl}.
-
-In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
-Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
-Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
-Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
-Finally, we set $\tau = \tau_c / 2$.
-\textbf{Thibault: Explain the choice of k}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{IS_GKTL/IS_GKTL_2}
-  \caption{Histogram of the values of the time-averaged drag $F_{T_a}$, with $T_a = 10\tau_c$, in the ensemble sampled by the cloning algorithm, for three values of the cloning strength $k$.
-In each case, the algorithm was applied with $N=1024$, $\tau = \tau_c / 2$.
-The grey histogram represents the average histogram obtained through a direct sampling, \textit{i.e.} with $k=0$.
-It was computed using a Gaussian approximation for the \ac{pdf} of the time-averaged drag.}
-  \label{fig:IS_GKTL}
-\end{figure}
-Figure~\ref{fig:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
-It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
-The higher $k$, the stronger the bias.
-The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
-We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:pdf_avg}).
-Figure~\ref{sec:IS_GKTL} also illustrates that the sampling by the algorithm is very concentrated on some fluctuations values, and that this effect increases as $k$ increases.
-This results from the overlap of trajectories as a consequence of the cloning and pruning.
-Indeed, if a trajectory is cloned at step $i$, then the trajectory ${\mathbf{x}(t)}$ over the interval $[0;i\tau]$ will be the same for all the clones.
-For a fixed ensemble size $N$, the overlap of trajectories increases as $k$ increases as fewer
-trajectories are cloned into a higher number of duplicates.
-\textbf{How do we quantify this effect? It is probably application-dependant. What are the strategies to mitigate it?}
-
-Figure~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag timeseries for several trajectories within the ensemble sampled by the cloning algorithm.
-It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-  \caption{Drag timeseries corresponding to the 6 highest fluctuations of the average drag in the ensemble sampled by the cloning algorithm.
-The cloning algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-  \label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-
-In addition, figure~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
-sampled by the cloning algorithm.
-It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-  \caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
-  \label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-Beyond dynamical aspects, the cloning algorithm provides ensemble of trajectories over which statistics of rare events can be computed by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}.
-In this section, we show that the biased sampling performed by the cloning algorithm allows for the computation of return times for events otherwise out-of-reach from a direct sampling approach.
-
-Each trajectory $j$ in the biased ensemble results in a timeseries of the time-averaged drag:
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  ,
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}
-\begin{equation}
-  \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{kT_aF_T^{(j)}}s_j(a),
-% \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d^{(j)}(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise. This amounts to taking the sum of the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-Two independent estimates are shown, obtained using different values of the cloning strength $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the cloning algorithm allows for the computation of statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This result shows that the biased ensemble sampled by cloning algorithm enables the estimation of return times for average drag fluctuations that would be out-of-reach from a direct sampling.
-Alternatively, for a fixed target return time, the use of the cloning algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is obviously a major advantage of this rare-event sampling algorithm.
-
-More generally, the successful computation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the cloning algorithm may be used to compute statistical estimators in a very efficient manner.
-We note, however, that some estimators may be more affected by the quality of the sampling than others.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent.
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
 \section{The \acl{lbm}}
@@
 
-\ZZ{
-	\section{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-	\label{app:AMS_on_OU}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-	The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-}
-
-\ZZ{
-\section{The \ac{gktl} algorithm}
-\label{app:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
-refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
-%This is true for stochastic dynamics.
-\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-%
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-}
-
-\include{appendix_ams}
-
-\EL{
-	\section{Perturbation at branching time}
-	\label{app:perturb_branching_time}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}
-
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "draft_p2_jfm"
 %%% End:

### 2020-02-05 jfm/draft_p2_jfm.tex
Move all appendices and introduction into separate .tex files

@@
 	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
 }
@@
 
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Otherwise, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
-%
-%
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-}
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-%
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-In more detail,  the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
-	drag fluctuations} is investigated based on a direct sampling approach.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
+\include{introduction}
 
@@
 Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
-The amplitude of this perturbation is chosen very low, so as not to impact the statistics of the dynamics.
+The amplitude of this perturbation is chosen very low, and it was tested that it does not impact the statistics of the sampled trajectories.
 The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
@@
 Finally, we set $\tau = \tau_c / 2$.
-\textbf{Thibault: Explain the choice of k}
 
@@
 \appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
 
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\begin{equation}
-  \label{eq:lbe}
-  f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\end{equation}
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\begin{equation}
-  \label{eq:lbe_eq}
-  f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-    \frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right),
-\end{equation}
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice.
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-\ZZ{
-	\section{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-	\label{app:AMS_on_OU}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-	The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-}
-
-\ZZ{
-\section{The \ac{gktl} algorithm}
-\label{app:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$
-refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.}
-%This is true for stochastic dynamics.
-\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-%
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-}
-
+\include{appendix_lbm}
 \include{appendix_ams}
-
-\EL{
-	\section{Perturbation at branching time}
-	\label{app:perturb_branching_time}
-}
+\include{appendix_gktl}
+\include{appendix_perturbation}
 

### 2020-02-05 jfm/appendix_perturbation.tex
Proofread appendix on perturbation

@@
 \label{app:perturb_branching_time}
-As mentioned in sections~\ref{sec:ams} and~\ref{sec:cloning}, the application of the \ac{ams} and the cloning algorithm to deterministic systems require the the addition of a perturbation to the initial condition of
+As mentioned in sections~\ref{sec:ams} and~\ref{sec:cloning}, the application of the \ac{ams} and the cloning algorithm to deterministic systems requires the addition of a perturbation to the initial condition of
 resampled trajectories.
@@
 
-The state of the flow at any point in time is given by the velocity field and the pressure $(\mathbf{u}, p)$, where the dependence on time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence free field $\delta \mathbf{u}$ and corresponding $\delta p$ solution of the
+The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
+A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the
 Poisson equation.
-In the context of the \ac{lbm} however, the state of the system is described at a mesoscopic level by the ensemble the populations $\{f_k(x_j, y_j)\}_{0\le k \le 8, 0 \le i <N_x, 0 \le j < N_y}$.
-A consequence is that, from the knowledge of a macroscopic state ($(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$, it is difficult to obtain the mesoscopic state from with the \ac{lbm} model can be initialised.
+In the context of the \ac{lbm} , the state of the system is described at a mesoscopic level by the ensemble the populations $\{f_k(x_j, y_j)\}_{0\le k \le 8, 0 \le i <N_x, 0 \le j < N_y}$ (see appendix~\ref{app:lbm}).
+As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
 
-In this work we work around this issue by designing the perturbative directly at the level of the mesoscopic populations.
-At each lattice site $\mathbf{x}$, the populations ${f_i}_{0\le i \le 8}$ are replaced by
+In this work, we work around this issue by designing the perturbation directly at the level of the mesoscopic populations.
+At each lattice site $\mathbf{x}$, the populations $\{f_i\}_{0\le i \le 8}$ are replaced by
 \begin{equation}
-  f_{i}(\mathbf{x}) \longrightarrow f_i(\mathbf{x}) + \epsilon \sum_{n=1}^{N} \alpha_{n}f_{i}^{n}(\mathbf{x}), \,\,\, 1\leq i \leq 9,
+  f_{i}(\mathbf{x}) \longrightarrow f_i(\mathbf{x}) + \epsilon \sum_{n=1}^{N} \alpha_{n}f_{i}^{n}(\mathbf{x}), \,\,\, 0\leq i \leq 8,
   \label{eq:perturb_pop}
 \end{equation}
-with the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the amplitude of the perturbation.
+with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the amplitude of the perturbation.
 The perturbed populations are then rescaled so that no mass is added in the system.
-For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}
-In other words obtained as a linear combination of $N$ solutions $\{f_i^{(n)}\}_{1\le n \le N}(\mathbf{x})$ of the numerical model~\eqref{eq:lbe}.
-Note that, the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions can be assimilated as a solution of the numerical model.
-However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq}, and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
+For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
+In other words, the perturbation is computed as a linear combination of $N$ solutions $\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
+Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
+However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
 
@@
 The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with $\epsilon = 0.002$ with a period $\tau = \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
+In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $\tau = \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
 Figure~\ref{fig:pdf_drag_with_perturbation} shows that the correct statistics are obtained even in presence of this slight perturbation.

### 2020-02-06 figures/autocorr_drag/plot_correlation.py
Update figure autocorrelation

@@ -1,28 +1,0 @@
-import numpy as np
-import matplotlib.style
-import matplotlib.pyplot as plt
-
-plt.style.use('ggplot')
-
-buf = np.fromfile("correlationData.dat", float, -1, "");
-
-n=int(len(buf)/2)
-tDomain = buf[0:n]
-c = buf[n:]
-
-fig,ax = plt.subplots(1,1, figsize=(12,8), constrained_layout=True)
-ax.set_xlim(0, 10)
-line0, =ax.plot(np.array([0, 10]), np.array([0, 0]))
-lineC, = ax.plot(tDomain/500,c,color=line0.get_color())
-line0.set_linestyle('--')
-line0.set_linewidth(1.5)
-
-ax.set_xlabel(r'$\tau / \tau_c$',fontsize=22)
-plt.ylabel(r'$C(\tau)$',fontsize=22)
-plt.xticks(fontsize=16)
-plt.yticks(fontsize=16)
-
-fname = 'autocorr_drag.png'
-plt.savefig(fname)
-
-plt.show()

### 2020-02-06 figures/autocorrelation_drag/autocorrelation_drag.py

@@ -1,0 +1,36 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+import sys
+from  scipy import integrate
+from os.path import splitext
+
+R = 16 # Size of the obstable in lattice unit
+U = 0.036 # Mean flow velocity in lattice unit 
+dt = U/R
+
+# -------* Autocorrelation function for drag process *---------
+path_autocorr_drag = '../../data/autocorr_drag.csv'
+try:
+    series = pd.read_csv(path_autocorr_drag, comment='#', index_col=0)
+except FileNotFoundError:
+    sys.exit('Error: Could not find find file {}'.format(path_autocorr_drag))
+
+# Compute integral of autocorrelation to estimate
+# autocorrelation time
+# Note that the autocorrelation is already normalized
+# by variance.
+I = integrate.trapz(series.values, axis=0)
+# Make the autocorrelation time the unit of time
+dt = 1/I
+
+C = series.values
+plt.plot(np.arange(0,C.size)*dt, C,lw=2)
+
+plt.xlabel('$\\tau / T_0$', fontsize=18)
+plt.ylabel('$C(\\tau)$', fontsize=18)
+
+fname = splitext(__file__)[0]+'.png'
+plt.savefig(fname)
+
+plt.show()

### 2020-02-06 figures/PDF_drag/PDF_drag.py
Add gaussian estimate to legend for PDF_drag

@@ -6,7 +6,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 
 
-fig, ax = plt.subplots(constrained_layout=True, figsize=(8, 5))
+fig, ax = plt.subplots()
 
 series = pd.read_csv("../../data/PDF_drag.csv", comment="#", index_col=0)
 series.plot(ax=ax, lw=2, legend=False)
@@ -16,20 +16,17 @@ series.plot(ax=ax, lw=2, legend=False)
 
 # Plot gaussian distribution
 xx = np.linspace(-4, 4, 50)
-ax.plot(
+plt.plot(
     xx,
     (lambda x: np.exp(-x * x * 0.5) / np.sqrt(2 * np.pi))(xx),
     linestyle="--",
     color="k",
     linewidth=1.5,
-    label="Gaussian statistics",
 )
-ax.set_xlabel("$(f_d - \\mu)/\\sigma$", fontsize=18)
-ax.set_ylabel("$\\mathcal{P}$", fontsize=18)
-ax.legend(fontsize=18)
-
-ax.legend(["With obstacle", "Without obstacle"])
-ax.set_yscale("log")
+plt.xlabel("$(f_d - \\mu)/\\sigma$", fontsize=18)
+plt.ylabel("$\\mathcal{P}$", fontsize=18)
+plt.legend(["With obstacle", "Without obstacle", "Gaussian estimate"], fontsize=18)
+plt.yscale("log")
 fname = splitext(__file__)[0] + ".png"
 plt.savefig(fname)
 

### 2020-02-06 jfm/draft_p2_jfm.tex
Implement Emmanuel's modifications for section 2 and 3. Move all 3 sections into separate .tex files

@@
 
-\section{Description of the numerical case study}
-\label{sec:test_flow}
+\include{section_description_case_study}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
+\include{section_extremes_direct_sampling}
 
-% introduce the flow
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}).
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-%
-We believe that this system embeds the characteristic features that makes \sout{the application of rare-event algorithms} \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics.
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square.
-%
-%Finally, the obstacle does not deform or move either.
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
-%
-
-% lattice Boltzmann method
-%
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations.
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition.
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag.
-As a consequence, the drag can eventually be expressed as the difference
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-	\label{fig:typical_drag_signal}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle.
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}
-	\caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
-The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		\textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}.
-One can then argue that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-			One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent.
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that
-\begin{equation}
-\label{eq:return_time}
-\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		As the amplitude of the drag increases, the relative variation of the base pressure also increases.
-\EL{The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
-
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
-%relatively to the forebody pressure.
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag.
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time.
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics.
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base.
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-%
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown.
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, a relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
-\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
-Indeed, both cases are %equally
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-\bigskip
-
-\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-\bigskip
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-\section{Rare-events algorithms}
-\label{sec:rare_events_algorithms}
-
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
-$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.
-
-In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
-computational cost much lower than their return time.
-Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
-greater complexity than the two-dimensional flow discussed in the previous sections.
-
-We focus on two algorithms, representatives of two popular approaches in rare event simulation.
-The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently, \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
-This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
-Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-More recently it allowed the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
-This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.
-
-The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
-system.
-Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
-The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
-influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
-Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
-committor function. However, with the exception of rare analytical cases, this function is unknown.
-See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.
-
-Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
-However, resampling steps are performed along the simulation of the ensemble.
-Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
-according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
-This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
-Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation_main}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
-See appendix~\ref{app:gktl_description} for a more detailed description of the cloning algorithm.
-
-Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
-rules to bias the sampling towards extreme events.
-In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
-events of interest.
-Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.
-
-Most previous applications of the \ac{ams} and cloning algorithm target stochastic dynamics.
-We stress that this is not the case in this work, as the target flow model is fully deterministic.
-Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
-The amplitude of this perturbation is chosen very low, and it was tested that it does not impact the statistics of the sampled trajectories.
-The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
-For further details about the addition of this perturbation, see appendix~\ref{sec:app_perturbation}.
-
-\subsection{The \acl{ams} algorithm}
-\label{sec:ams}
-We applied the \ac{ams} algorithm, described in appendix\ref{sec:app_ams}, to the sampling of extreme fluctuations of the drag applied on the square obstacle immersed in the flow described in section~\ref{sec:test_flow}.
-We chose to set the score function as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
-Let us stress that we do not expect this choice to be close to the optimal "committor" function.
-Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
-into the physics that lead to the extreme events of interest.
-Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
-It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
-score function, a choice that is often the only available one.
-
-In addition to the score function, the parameters of the \ac{ams} are the number of trajectories $N$ and the duration of trajectories $T_a$.
-As detailed in appendix~\ref{app_ams}, the size of the ensemble $N$ governs the statistical error affecting the estimators based on the sampled set of trajectories.
-Furthermore, the number of resamplings required to reach a final ensemble of trajectories of probability of the order of $10^{-\beta}$ can be shown to scale like $\beta N$ \citep{lestang_computing_2018}.
-In this work, the choice of $N$ was governed by the available computational resources.
-We performed two experiments with $N = 32$ and $N = 256$, having in mind that higher numbers are unlikely to be accessible in the case of more complex or three-dimensional flows.
-The duration of the trajectories was set to $T_a$ = $5\tau_c$, a value that is larger than the timescale of extreme drag fluctuations whilst maintaining an affordable computational cost.
-As pointed out in appendix~\ref{app_ams}, setting $T_a \gg \tau_c$ does not lead to a better sampling.
-
-\begin{figure}
-  \centering
-  \subfloat[\label{fig:AMS_drag_trajectories_a} Final ensemble of trajectories sampled by the \ac{ams}]{\includegraphics[width=.4\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}}
-  \subfloat[Drag maximum for resampled trajectories]{\label{fig:AMS_drag_trajectories_b}\includegraphics[width=.4\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}}
-  \caption{(a) Ensemble of $N = 32$ trajectories after $181$ \ac{ams} resamplings.
-    (b) Maximum of the instantaneous drag throughout the resampled trajectories as a function of the corresponding computational cost $C_{AMS}$, \textit{i.e.} as a function of the resamplings.
-The maximum over resampled trajectories overlap until saturating at the level of the largest maximum in the initial ensemble.
-  }
-\end{figure}
-
-Figure~\ref{fig:AMS_drag_trajectories_a} illustrates that the splitting introduced by the \ac{ams} does not enhance the sampling of extremes but rather has catastrophic consequences on the sampling of trajectories.
-Indeed, trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.
-This result is confirmed by increasing the number of initial trajectories from $N=32$ to $N=256$, as shown
-in figure~\ref{fig:AMS_drag_trajectories_b}.
-Recalling section~\ref{sec:instantaneous_drag}, extreme drag fluctuations have a lifetime of $\tau_c$, the timescale over which a vortex is blocked near the base of the obstacle.
-After $\tau_c$, the vortex is freed and swept downstream, and further large drag fluctuations can only result from the interaction of the obstacle with the incoming flow, beyond the correlation time of the drag fluctuations.
-Moreover, it takes a certain time $\tau_L$ (known as the Lyapunov timescale) before a trajectory resampled by the \ac{ams} separates from its parent under the effect of the initial perturbation.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories_a} that this separation timescale is larger than the duration of extreme drag fluctuations $\tau_c$, leading to an overlap of trajectories.
-
-We stress that this phenomenology is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble can only increase the amplitude of the the global maximum attained initially, but does not solve the issue of overlapping trajectories.
-Increasing $T_a$ allows further time for resampled trajectories to develop rare fluctuations, however after a duration $\Delta T \geq \tau_L > \tau_c$, and therefore decorrelated from the resampling.
-
-Our results indicate that the \ac{ams} does not work with this simple choice of score function.
-Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.
-
-\section{The cloning algorithm}
-\label{sec:cloning}
-We applied the cloning algorithm described, in appendix\ref{sec:app_gktl}, to the two-dimensional flow described in section~\ref{sec:test_flow}.
-Unlike the case of the \ac{ams} in section~\ref{sec:ams}, the algorithm is here used to sampled trajectories that exhibit extremes values of
-the \emph{average} drag.
-
-The cloning algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The cloning strength is the multiplicative factor appearing the the bias term of equation~\eqref{eq:Biased_Path_Approximation_main}.
-The higher $k$, the larger the typical average drag for the trajectories in the ensemble sampled by the algorithm.
-There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.
-Similarly to the \ac{ams}, $N$ governs the error affecting estimates of probabilities and estimators based on the biased ensemble.
-In the cloning algorithm, the population of trajectories is periodically resampled as the trajectories are simulated from $t=0$ to
-$t=T_a$.
-The cloning period determines how often this resampling step is performed.
-For deterministic dynamics, a common rule of thumb is to chose $\tau \approx \tau_c$ \citep{giardina_direct_2006}.
-For further details about the cloning algorithm, see appendix~\ref{sec:app_gktl}.
-
-In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
-Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
-Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
-Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
-Finally, we set $\tau = \tau_c / 2$.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{IS_GKTL/IS_GKTL_2}
-  \caption{Histogram of the values of the time-averaged drag $F_{T_a}$, with $T_a = 10\tau_c$, in the ensemble sampled by the cloning algorithm, for three values of the cloning strength $k$.
-In each case, the algorithm was applied with $N=1024$, $\tau = \tau_c / 2$.
-The grey histogram represents the average histogram obtained through a direct sampling, \textit{i.e.} with $k=0$.
-It was computed using a Gaussian approximation for the \ac{pdf} of the time-averaged drag.}
-  \label{fig:IS_GKTL}
-\end{figure}
-Figure~\ref{fig:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
-It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
-The higher $k$, the stronger the bias.
-The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
-We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:pdf_avg}).
-Figure~\ref{sec:IS_GKTL} also illustrates that the sampling by the algorithm is very concentrated on some fluctuations values, and that this effect increases as $k$ increases.
-This results from the overlap of trajectories as a consequence of the cloning and pruning.
-Indeed, if a trajectory is cloned at step $i$, then the trajectory ${\mathbf{x}(t)}$ over the interval $[0;i\tau]$ will be the same for all the clones.
-For a fixed ensemble size $N$, the overlap of trajectories increases as $k$ increases as fewer
-trajectories are cloned into a higher number of duplicates.
-\textbf{How do we quantify this effect? It is probably application-dependant. What are the strategies to mitigate it?}
-
-Figure~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag timeseries for several trajectories within the ensemble sampled by the cloning algorithm.
-It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-  \caption{Drag timeseries corresponding to the 6 highest fluctuations of the average drag in the ensemble sampled by the cloning algorithm.
-The cloning algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-  \label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-
-In addition, figure~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
-sampled by the cloning algorithm.
-It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-  \caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
-  \label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-Beyond dynamical aspects, the cloning algorithm provides ensemble of trajectories over which statistics of rare events can be computed by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}.
-In this section, we show that the biased sampling performed by the cloning algorithm allows for the computation of return times for events otherwise out-of-reach from a direct sampling approach.
-
-Each trajectory $j$ in the biased ensemble results in a timeseries of the time-averaged drag:
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  ,
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}
-\begin{equation}
-  \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{kT_aF_T^{(j)}}s_j(a),
-% \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d^{(j)}(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise. This amounts to taking the sum of the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-Two independent estimates are shown, obtained using different values of the cloning strength $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the cloning algorithm allows for the computation of statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This result shows that the biased ensemble sampled by cloning algorithm enables the estimation of return times for average drag fluctuations that would be out-of-reach from a direct sampling.
-Alternatively, for a fixed target return time, the use of the cloning algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is obviously a major advantage of this rare-event sampling algorithm.
-
-More generally, the successful computation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the cloning algorithm may be used to compute statistical estimators in a very efficient manner.
-We note, however, that some estimators may be more affected by the quality of the sampling than others.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
+\include{rare_events_algorithms}
 

### 2020-02-06 jfm/section_description_case_study.tex
Implement Emmanuel's modifications for section 2 and 3. Move all 3 sections into separate .tex files

--- 2019-10-19 draft_p2.tex
@@
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage,preprint]{revtex4-1}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-\usepackage{caption}
-\graphicspath{{figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\begin{document}
-
-\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-
-\author{Thibault Lestang}
-\email{thibault.lestang@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\author{Freddy Bouchet}
-\email{freddy.bouchet@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\author{Emmanuel Lévêque}
-\email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-
-\begin{abstract}
-This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems. 
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Such anomalous constraints can threaten the structural integrity of embedded structures.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
-%
-A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-
-%
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
-%
-%
-An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-% 
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
-More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
-Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
-This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
-drag fluctuations is investigated based on a direct sampling approach.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
 \section{Description of the numerical case study}
@@
 %
-We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
+We believe that this system embeds the characteristic features that makes this study challenging and relevant for fluid-structure-interaction problems.
 % rapid description of the flow
@@
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
+Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
 %This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
 Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix and references therein.
+Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
 In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
@@
 %
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
+No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.
 %
@@
 	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
-	The vorticity is given in lattice units.}
+	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $4T_0$. The vorticity is expressed in unit of $T_0$.}
 	\label{fig:typical_vorticity}
@@
 \end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
+where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
+Here, the viscous stress makes a negligible contribution to the drag.
+The latter therefore results mostly from pressure forces.
 %
@@
 between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity fiel.
+Pressure fluctuations are related to the dynamics of the vorticity field.
 Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
 
 % Definition of turnover time
-%
 The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
@@
 \label{eq:turnover_time}
-\tau_0 = \frac{R}{U},
+T_0 = \frac{R}{U},
 \end{equation}
@@
 Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
+Vorticity is generated along the forebody and eventually carried away by the flow.
+Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.
 \subsection{The drag as a random process
@@
 	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
+The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
 	\label{fig:typical_drag_signal}
@@
 % pdf
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$.
+Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~T_0$.
 This long simulation will be referred to as the \textit{control run} in the following.
@@
 The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.} ${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$.
 %
@@
 	{\label{fig:pdf_drag_b}
-    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
+		\includegraphics[width=.45\linewidth]{./autocorrelation_drag/autocorrelation_drag.png}}
 	\caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
 The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
+The estimate with obstacle was computed over a timeseries spanning $4\times 10^6T_0$.
+The estimate without square was computed over a timeseries spanning $4\times 10^5 T_0$.
           \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is defined by $C(\tau_c)=0$.
-  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-}
+The correlation time $\tau_c\simeq 4 T_0$ is defined by $C(\tau_c)=0$.}
 	\label{fig:pdf_drag}
@@
 Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4T_0$, illustrating that the drag loses its memory over a time scale corresponding to the sweeping of a few eddies past the obstacle.
 %
@@
 In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
+The ratio $T_0 / \tau_c$ may be viewed as a Strouhal number.
+The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.
 
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-	One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a function the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
- \label{eq:return_time}
-r(a) \underset{a\to\infty}{\propto} \exp(-la)
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{Instantaneous drag}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-	As the amplitude of the drag increases, the relative variation of the base pressure also increases. 
-	%\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} 
-}
-\end{figure}
-
-% teasing
-%
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
-
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-  \tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure. 
-%relatively to the forebody pressure.
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-This is reminiscent of time-irreversibility in turbulent dynamics. 
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-  \caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-  }
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-  \caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-  \caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-  \label{eq:avg_shear_def}
-  \overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, a relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-  \label{eq:def_time_averaged_drag}
-  F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-It is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are equally featured in Fig.~\ref{fig:extreme_avg}.
-Very large value of the averaged-drag results from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
-%
-% theoretical explanation
-%
-This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-\begin{equation}
-  \label{eq:indep}
-  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-\end{equation}
-If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-\begin{equation}
-  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-  \label{eq:ratio_exp}
-\end{equation}
-The observation is therefore well supported theoretically.
-See~\cite{lestang2019JSTAT} for further details.
-
-\section{Rare-event algorithms}
-\label{sec:rare_events_algorithms}
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like $r(a) \propto e^{l a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than their related return times.
-%
-For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
-%
-%
-Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
-In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-%In a nutshell
-Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
-These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables. 
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
-Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-\end{figure}
-
-% Introduction of AMS
-%
-The \acl{ams} algorithm~\cite{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \cite{KahnHarris1951}.
-The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into 
-%a sequence of events that can be sampled with a higher probability  
-a sequence of shorter excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
-The operating principle is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function during the time interval $0\leq t \leq T_a$.
-Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
-{Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-  It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}.
-  The following results are not limited to this particular variant.}
-
-\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-\label{sec:ams_ou}
-% Example: OU process
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-  \label{fig:comparaison_temps_de_retour}
-\end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-\begin{equation}
-  \label{eq:ou}
-  \dot{x} = -x + \eta (t),
-\end{equation}
-where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%
-The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-In the latter, the typical computational cost is simply the return time $r(a)$.
-One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features do no \textit{a priori} persist in the case of fluid dynamics.
-
-\subsubsection{The \ac{ams} for extreme drag fluctuations}
-\label{sec:ams_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
-The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-Our aim is now to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
-In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.
-%
-{Let us also stress that the dynamics is here deterministic.
-Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-%
-Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
-% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
-The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
-
-% application
-%
-The \ac{ams} algorithm is applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%
-Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost, similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-%(see Fig.~\ref{fig:comparaison_temps_de_retour}).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-  \caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
-In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
-  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{ams} re-sampling procedure. 
-%
-Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  
-%
-Interestingly, all trajectories ultimately originate from one unique initial trajectory. 
-%
-This feature may be explained as follows.
-It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations, $\tau_c$, is much shorter than the Lyapunov's timescale $\tau_L$.
-As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and therefore the probability of observing a new extreme fluctuation is also very low.
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
-%
-As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L \gtrsim \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-A possible line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
-%
-
-\subsection{Importance sampling of extreme time-averaged drag forces with the \acl{gktl} algorithm}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
-
-\subsubsection{The \ac{gktl} algorithm}
-\label{sec:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-  \label{eq:mean_field}
-  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\cite{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics.
-All trajectories in the resampled ensemble are solutions of the original dynamical.
-%This is true for stochastic dynamics.
-{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to allow clones of a same trajectory to separate. 
-As for the \ac{ams} algorithm, it is assumed that this perturbation does not significantly affect the statistics of the sampled trajectories.}
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-	The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is now considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-The purpose is to sample trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.
-{Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.}
-
-% Choice of parameters and perturbation
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-In the following experiments, $\tau = \tau_c /2$. 
-%
-
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-\begin{equation}
-  \label{eq:estimate_biased_measure}
-  \rho_k(F) 
-  %= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-\end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-%
-% limitations
-%
-Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
-%
-Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-		\hline
-		${N}_{brute-force}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
-Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independant estimates, obtained using different values of the bias parameter $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to yield an large value of the average.
-Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent. 
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix*
-\section{The \acl{lbm}}
-\label{sec:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \cite{De_Rosis_2016}.
-
-\bibliographystyle{unsrt}
-\bibliography{biblio}
-
-\end{document}
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "draft_p2_jfm"
 %%% End:

### 2020-02-06 jfm/section_extremes_direct_sampling.tex
Implement Emmanuel's modifications for section 2 and 3. Move all 3 sections into separate .tex files

--- 2019-10-19 draft_p2.tex
@@
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage,preprint]{revtex4-1}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-\usepackage{caption}
-\graphicspath{{figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\begin{document}
-
-\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-
-\author{Thibault Lestang}
-\email{thibault.lestang@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\author{Freddy Bouchet}
-\email{freddy.bouchet@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\author{Emmanuel Lévêque}
-\email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-
-\begin{abstract}
-This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems. 
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Such anomalous constraints can threaten the structural integrity of embedded structures.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
-%
-A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-
-%
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
-%
-%
-An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-% 
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
-More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
-Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
-This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
-drag fluctuations is investigated based on a direct sampling approach.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
-	The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity fiel.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\tau_0 = \frac{R}{U},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
-  \label{fig:typical_drag_signal}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-  \centering
-    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
-    {\label{fig:pdf_drag_a}
-    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[Autocorrelation of drag fluctuations]
-    {\label{fig:pdf_drag_b}
-    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
-      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
-The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
-  \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is defined by $C(\tau_c)=0$.
-  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-}
-  \label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
-
 \section{Extreme fluctuations of the drag by means of direct sampling}
@@
 	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a function the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
+	\includegraphics[width=.7\linewidth]{return_time/return_time.png}
+	\caption{Amplitude of drag fluctuations as a function of the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
 	}
@@
 The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
+The mixing time $\tau_m$ is the time needed for the dynamics to lose the memory of its initial condition.
 As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
@@
 
-How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
+How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
 We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
@@
 %
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
 If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
 one might expect that
@@
   \label{eq:return_time}
-r(a) \underset{a\to\infty}{\propto} \exp(-la)
+  r(a) \underset{a\to\infty}{\propto} \exp(la),
 \end{equation}
 where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
 Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
@@
 
-\subsection{Instantaneous drag}
+\subsection{Extreme fluctuations of the instantaneous drag}
 \label{sec:instantaneous_drag}
@@
 	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-	As the amplitude of the drag increases, the relative variation of the base pressure also increases. 
-	%\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} 
-}
+An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).}
 \end{figure}
@@
 %
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
+In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the vorticity around the obstacle.
+In the following we show that the situation os different in the case of extreme drag fluctuations.
 
@@
 Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure. 
-%relatively to the forebody pressure.
+$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the drag fluctuation $\tilde f_d^{\star}$.
+It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from the drop of the base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
+On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).
 
@@
 Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-This is reminiscent of time-irreversibility in turbulent dynamics. 
-%Moreover, such very high drag levels do not persist over time.
+%This is reminiscent of time-irreversibility in turbulent dynamics.
+This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.
 To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
@@
 %
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
 Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
@@
 where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
+In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation of the instantaneous drag.
+In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:IS_GKTL}).
 
@@
 %
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
 By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
+Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
+As a rule of thumb, the threshold has been set so that around one hundred events are selected.
 
@@
 Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-It is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are equally featured in Fig.~\ref{fig:extreme_avg}.
-Very large value of the averaged-drag results from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
-%
-% theoretical explanation
-%
-This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-\begin{equation}
-  \label{eq:indep}
-  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-\end{equation}
-If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-\begin{equation}
-  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-  \label{eq:ratio_exp}
-\end{equation}
-The observation is therefore well supported theoretically.
-See~\cite{lestang2019JSTAT} for further details.
+Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
+Indeed, both cases are %equally
+featured in Fig.~\ref{fig:extreme_avg};
+very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
 
-\section{Rare-event algorithms}
-\label{sec:rare_events_algorithms}
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like $r(a) \propto e^{l a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than their related return times.
-%
-For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
-%
-%
-Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
-In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-%In a nutshell
-Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
-These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables. 
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
-Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-\end{figure}
-
-% Introduction of AMS
-%
-The \acl{ams} algorithm~\cite{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \cite{KahnHarris1951}.
-The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into 
-%a sequence of events that can be sampled with a higher probability  
-a sequence of shorter excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
-The operating principle is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function during the time interval $0\leq t \leq T_a$.
-Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
-{Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-  It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}.
-  The following results are not limited to this particular variant.}
-
-\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-\label{sec:ams_ou}
-% Example: OU process
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-  \label{fig:comparaison_temps_de_retour}
-\end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-\begin{equation}
-  \label{eq:ou}
-  \dot{x} = -x + \eta (t),
-\end{equation}
-where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%
-The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-In the latter, the typical computational cost is simply the return time $r(a)$.
-One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features do no \textit{a priori} persist in the case of fluid dynamics.
-
-\subsubsection{The \ac{ams} for extreme drag fluctuations}
-\label{sec:ams_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
-The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-Our aim is now to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
-In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.
-%
-{Let us also stress that the dynamics is here deterministic.
-Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-%
-Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
-% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
-The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
-
-% application
-%
-The \ac{ams} algorithm is applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%
-Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost, similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-%(see Fig.~\ref{fig:comparaison_temps_de_retour}).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-  \caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
-In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
-  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{ams} re-sampling procedure. 
-%
-Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  
-%
-Interestingly, all trajectories ultimately originate from one unique initial trajectory. 
-%
-This feature may be explained as follows.
-It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations, $\tau_c$, is much shorter than the Lyapunov's timescale $\tau_L$.
-As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and therefore the probability of observing a new extreme fluctuation is also very low.
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
-%
-As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L \gtrsim \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-A possible line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
-%
-
-\subsection{Importance sampling of extreme time-averaged drag forces with the \acl{gktl} algorithm}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
-
-\subsubsection{The \ac{gktl} algorithm}
-\label{sec:gktl_description}
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-  \label{eq:mean_field}
-  Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\cite{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
-A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics.
-All trajectories in the resampled ensemble are solutions of the original dynamical.
-%This is true for stochastic dynamics.
-{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to allow clones of a same trajectory to separate. 
-As for the \ac{ams} algorithm, it is assumed that this perturbation does not significantly affect the statistics of the sampled trajectories.}
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-	The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is now considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-The purpose is to sample trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.
-{Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.}
-
-% Choice of parameters and perturbation
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-In the following experiments, $\tau = \tau_c /2$. 
-%
-
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-\begin{equation}
-  \label{eq:estimate_biased_measure}
-  \rho_k(F) 
-  %= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-\end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-%
-% limitations
-%
-Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
-%
-Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-		\hline
-		${N}_{brute-force}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
-Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independant estimates, obtained using different values of the bias parameter $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to yield an large value of the average.
-Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent. 
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix*
-\section{The \acl{lbm}}
-\label{sec:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \cite{De_Rosis_2016}.
-
-\bibliographystyle{unsrt}
-\bibliography{biblio}
-
-\end{document}
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "draft_p2_jfm"
 %%% End:

### 2020-02-07 jfm/draft_p2_jfm.tex
Update conclusion

@@
 
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
+In a first part, we showed that the phenomenology describing extreme drag fluctuations is very different from typical events.
+By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
+The lifetime of extreme drag fluctuations of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
+Furthermore, this lifetime is of the order of the correlation timescale of the drag and therefore each successive extreme drag events are decorrelated for one another.
+From a timeseries spanning four million turn-over times we also illustrated that the phenomenology of extremes of the time-averaged drag cannot be reduced to a well defined scenario.
+Indeed, extreme average values result both from rare successions of typical fluctuations, of the occurrence a fewer number of rare extreme fluctuations of the instantaneous drag.
 
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
+In order to extend our study, and consider its replication on more complex flows, we considered the application of two popular rare-event sampling algorithms.
+Because of the short lifetime of extreme drag fluctuations, the \ac{ams} does not lead to a better sampling
+when using the drag itself as a score function.
+The choice of the score function is difficult for complex dynamics, and optimal score function is problem-dependant.
+To enable the application of the \ac{ams} to a wide range of complex dynamical systems, an important question is the systematic estimation of the optimal score function, for instance using machine learning approaches.
 
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent.
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
+We obtained more satisfactory results for the averaged drag using the \ac{gktl} algorithm.
+Indeed, using the algorithm we could sample trajectories corresponding to values of the time-averaged drag unreachable from a direct sampling of similar computational cost.
+Qualitatively, we found the phenomenology of the sampled extreme events to be consistent with the findings of the part of the study.
+However, a more quantitative analysis is required to assess the ability of the \ac{gktl} algorithm to equally sample the different dynamics leading to a given extreme value of the average drag.
+Nonetheless, our study shows that the sampled ensemble is useful to compute estimators over extreme events, such as return times of extreme fluctuations of the average drag.
 
@@
 The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
+Finally, the authors would like to thank the three anonymous referees for their detailed review and insightfull comments and suggestions.
 

### 2020-02-07 jfm/draft_p2_jfm.tex
Update abstract following Emmanuel's comments

@@
 
-\title{\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
+\title{
+  Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
 }
@@
 \begin{abstract}
-	\ZZ{This study investigates by means of direct numerical simulation, extreme drag events exerted by a  turbulent flow on a bluff body and examine the relevance of statistical algorithms to sample efficiently these events.
-Two representative algorithms are examined: AMS and GKTL.
-Briefly mention differences and complementarity.
-What do we learn?
-The take-home message is...}
-\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems.
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
+  This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body, and examines the relevance of statistical algorithms to efficiently sample these events.
+  The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
+  Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
+  Two different algorithms are considered to speed up the sampling of such flow scenarii related to extreme values of the drag, namely the \ac{ams} and the \ac{gktl} algorithms.
+  The idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned to sample large-amplitude events more frequently.
+  These two algorithms are representative of the two main class of algorithms applicable to non-equilibrium and deterministic dynamics.
+  Whilst the \ac{ams} is general, the \ac{gktl} algorithm is particularly suited to sample trajectories based on the average value of a given observable.
+  These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
+  Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong
+  influence on the efficiency of the selection-and-cloning rules of these two algorithms. While
+  the \ac{ams} algorithm does not yield significant run-time savings as compared to direct
+  sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of
+  the time-averaged drag and estimate related statistics such as return times.
 \end{abstract}

### 2020-02-07 figures/pressure_ratio/plot_pressure_ratio_revision.py
Add extreme events to figure pressure_ratio

@@ -21,6 +21,9 @@ flucts = [_tuple for i, _tuple in enumerate(flucts_reader) if os.path.isfile("..
 csv_file_handle.seek(0)
 event_dirs = ["../../data/event_{}".format(i) for i, _tuple in enumerate(flucts_reader) if os.path.isfile("../../data/event_{}/dragFile.dat".format(i))]
 csv_file_handle.close()
+NbExtremeEvents = 104
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+extreme_event_dirs = [os.path.join(prefix, "event_{}_feb".format(i+1)) for i in range(NbExtremeEvents)]
 
 maxDrag = []
 frontContrib = []
@@ -39,12 +42,25 @@ for counter, fluct in enumerate(flucts):
     maxDrag.append(drag[t_star])
     frontContrib.append((pFront[t_star]-meanPf)/(maxDrag[counter]-meanDrag))
     rearContrib.append((meanPr+pRear[t_star])/(maxDrag[counter]-meanDrag))
+for dirname in extreme_event_dirs:
+    drag = np.fromfile(os.path.join(dirname,'data_force.datout'), float, -1, "")
+    pFront = np.fromfile(os.path.join(dirname,'pFront.dat'), float, -1, "")
+    pRear = np.fromfile(os.path.join(dirname,'pRear.dat'), float, -1, "")
+    vTop = np.fromfile(os.path.join(dirname,'viscousTop.dat'), float, -1, "")
+    vBot = np.fromfile(os.path.join(dirname,'viscousBot.dat'), float, -1, "")
+    vFront = np.fromfile(os.path.join(dirname,'viscousFront.dat'), float, -1, "")
+    vRear = np.fromfile(os.path.join(dirname,'viscousRear.dat'), float, -1, "")
 
+    _maxDrag = np.amax(drag)
+    maxDrag.append(_maxDrag)
+    I = np.argmax(drag)
+    frontContrib.append((pFront[I]-meanPf)/(_maxDrag-meanDrag));
+    rearContrib.append((meanPr+pRear[I])/(_maxDrag-meanDrag));
+    
 maxDrag = np.array(maxDrag)
 maxDrag = (maxDrag-mu) / sig
     
-fig,ax = plt.subplots(figsize=(9,5), constrained_layout=True)
-
[…]

### 2020-02-08 jfm/appendix_ams.tex
Minor fixes of typos and references

@@
 \section{The \ac{ams} algorithm}
-\label{app_ams}
+\label{app:ams}
 
@@
 One might expect intuitively that $\ell_j=1$.
-However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{Brehier2016a}.
+However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
 The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
@@
 \subsection{How to compute the probability}
-Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,Brehier2016a}:
+Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
 \begin{equation}
@@
 of iterations $N$.
-We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{ams_threshold_def}.
+We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
 The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
@@
 In the following we denote this estimator $\hat{q}_N(a)$.
-One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{Brehier2016a} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
+One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
 \begin{theo}
@@
 \end{theo}
-Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{Brehier2016a,Rolland2015} for extensive numerical simulations concerning the role of the score function.
+Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
 

### 2020-02-08 jfm/appendix_perturbation.tex
Minor fixes of typos and references

@@
 \label{app:perturb_branching_time}
-As mentioned in sections~\ref{sec:ams} and~\ref{sec:cloning}, the application of the \ac{ams} and the cloning algorithm to deterministic systems requires the addition of a perturbation to the initial condition of
+As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl}, the application of the \ac{ams} and the cloning algorithm to deterministic systems requires the addition of a perturbation to the initial condition of
 resampled trajectories.

### 2020-02-08 jfm/section_description_case_study.tex
Minor fixes of typos and references

@@
 The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
+The initial distributions are imposed at equilibrium (see appendix~\ref{app:lbm}).
 In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
+	\includegraphics[width=.9\linewidth]{illustrate_return_time/illustrate_return_time}
 	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
+The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
+          One observes that $r(a) \gg \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.
+          The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.}
 	\label{fig:typical_drag_signal}

### 2020-02-08 jfm/section_extremes_direct_sampling.tex
Minor fixes of typos and references

@@
 \label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-            One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
 
@@
 
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
+The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:typical_drag_signal}.
 The mixing time $\tau_m$ is the time needed for the dynamics to lose the memory of its initial condition.
@@
 In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:IS_GKTL}).
+In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+  \caption{Estimate of the \ac{pdf} for the time-averaged drag $F_T$ with $T = 10\tau_c$.
+This estimate is based on a rolling average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.}
+    \label{fig:PDF_AVG}
+\end{figure}
 

### 2020-02-11 figures/effect_of_perturbation/effect_of_perturbation.py
Track file for section 3 and figures of appendix on perturbation

@@ -1,0 +1,26 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+tau_c = 1999
+N = 2
+labels = ["Trajectory {}".format(i) for i in range(N)]
+labels[0] = "Reference trajectory"
+
+fig, ax = plt.subplots(1, 1, figsize=(10, 4), constrained_layout=True)
+
+for i in range(N):
+    dirname = "simulation_{}".format(i)
+    f = np.fromfile(os.path.join(dirname, "dragFile.dat"), float, -1, "")
+    tspan = np.linspace(0, (len(f) - 1) / tau_c, len(f))
+    plt.plot(tspan, f)
+
+plt.legend(labels)
+
+plt.xlabel(r"$t / \tau_c$", fontsize=22)
+plt.ylabel("$f_d$", fontsize=22)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+figname = os.path.splitext(__file__)[0] + ".png"
+plt.savefig(figname)
+plt.show()

### 2020-02-11 figures/pdf_drag_with_perturbation/compute_pdf_drag.py

@@ -1,0 +1,32 @@
+import sys
+import datetime
+from os.path import realpath, join
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+list_timeseries = ["../../data/cr_perturb/all_the_timeseries.bin",
+              "../../data/control_1_2017_10_31_14_07_55/data_force.datout"]
+output_files = ["PDF_drag_perturb.csv", "PDF_drag_one_file.csv"]
+
+var = 0.00173
+mu = iter([0.0, 0.0252])
+for timeseries, output_file in zip(list_timeseries, output_files):
+    # Note that the mean is already substracted
+    xtot = np.fromfile(timeseries, float, -1, "") - next(mu)
+
+    print("Computing histogram...")
+    xtot = xtot / np.sqrt(var)
+    hist, bins = np.histogram(xtot, bins=100, density=True)
+    bin_centers = [0.5 * (b[1] + b[0]) for b in zip(bins[:-1], bins[1:])]
+
+    series = pd.Series(data=hist, index=bin_centers)
+
+    output_csv = join("../../data/", output_file)
+    with open(output_csv, "w") as f:
+        f.write("# Generated with {}\n".format(sys.argv[0]))
+        f.write("# {}\n".format(str(datetime.datetime.now())))
+        f.write("# mu = 0 var = 1")
+        f.write("# {}".format(realpath(timeseries)))
+        f.write("#\n")
+        series.to_csv(f, header=True)

### 2020-02-11 figures/pdf_drag_with_perturbation/pdf_drag_with_perturbation.py

@@ -1,0 +1,24 @@
+import sys
+from os.path import splitext
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+
+fig, ax = plt.subplots(constrained_layout=True, figsize=(8, 5))
+
+series = pd.read_csv("../../data/PDF_drag_one_file.csv", comment="#", index_col=0)
+series.plot(ax=ax, lw=2, legend=False)
+series_perturb = pd.read_csv("../../data/PDF_drag_with_perturbation.csv", comment="#", index_col=0)
+series_perturb.plot(ax=ax, lw=2, legend=False)
+
+ax.set_xlabel("$(f_d - \\mu)/\\sigma$", fontsize=18)
+ax.set_ylabel("$\\mathcal{P}$", fontsize=18)
+ax.legend(["No perturbation", "$\\epsilon = 0.002$"], fontsize=18)
+
+ax.set_yscale("log")
+fname = splitext(__file__)[0] + ".png"
+plt.savefig(fname)
+
+plt.show()

### 2020-02-11 jfm/appendix_gktl.tex
Add missing appendix on GKTL and section on rare events algorithms

--- 2019-10-19 draft_p2.tex
@@
-\documentclass[pre,aps,floatfix,10pt,superscriptaddress, notitlepage,preprint]{revtex4-1}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-\usepackage{caption}
-\graphicspath{{figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\begin{document}
-
-\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-
-\author{Thibault Lestang}
-\email{thibault.lestang@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\author{Freddy Bouchet}
-\email{freddy.bouchet@ens-lyon.fr}
-\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
-\author{Emmanuel Lévêque}
-\email{emmanuel.leveque@ec-lyon.fr}
-\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-
-\begin{abstract}
-This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems. 
-%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
-Such anomalous constraints can threaten the structural integrity of embedded structures.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
-The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
-Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
-%
-A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-
-%
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
-%
-%
-An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
-% 
-Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
-%
-
-The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}.
-More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
-Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
-%
-{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
-This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
-%
-
-{The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
-%
-Motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
-drag fluctuations is investigated based on a direct sampling approach.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
-This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
-As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
-Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
-	The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity fiel.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\tau_0 = \frac{R}{U},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
-Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-  \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
-  \label{fig:typical_drag_signal}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
-The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~\tau_0$.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-  \centering
-    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
-    {\label{fig:pdf_drag_a}
-    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
-    \subfloat[Autocorrelation of drag fluctuations]
-    {\label{fig:pdf_drag_b}
-    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
-      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value.
-The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle.
-%Note that the amplitudes have not been normalized.
-  \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time $\tau_c\simeq 4 \tau_0$ is defined by $C(\tau_c)=0$.
-  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-}
-  \label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$.
-One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-	One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a function the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to loose the memory of its initial conditions.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
- \label{eq:return_time}
-r(a) \underset{a\to\infty}{\propto} \exp(-la)
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{Instantaneous drag}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-	As the amplitude of the drag increases, the relative variation of the base pressure also increases. 
-	%\TL{in Figure, $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$} 
-}
-\end{figure}
-
-% teasing
-%
-In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.
-
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-  \tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
-%
-It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure. 
-%relatively to the forebody pressure.
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
-This is reminiscent of time-irreversibility in turbulent dynamics. 
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-  \caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-  }
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-  \caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-  \caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-  \label{eq:avg_shear_def}
-  \overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, a relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-  \label{eq:def_time_averaged_drag}
-  F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-Relevant values of $T$ are application dependant.
-We shall consider $T=10\tau_c$ in the following.
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-It is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are equally featured in Fig.~\ref{fig:extreme_avg}.
-Very large value of the averaged-drag results from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.
-%
-% theoretical explanation
-%
-This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-\begin{equation}
-  \label{eq:indep}
-  p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-\end{equation}
-If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-\begin{equation}
-  \frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-  \label{eq:ratio_exp}
-\end{equation}
-The observation is therefore well supported theoretically.
-See~\cite{lestang2019JSTAT} for further details.
-
-\section{Rare-event algorithms}
-\label{sec:rare_events_algorithms}
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows in industrial or environmental domains, the computational cost of  direct sampling becomes prohibitive.
-According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like $r(a) \propto e^{l a}$.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The motivation behind rare-events algorithms is to sample extreme fluctuations for a computational cost much lower than their related return times.
-%
-For this purpose, the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms are considered in the following.
-%
-%
-Both algorithms rely on an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory (of duration $T_a$) of the system in phase space.
-In our case, the dynamical system under investigation is the flow described in section~\ref{sec:test_flow}.
-%In a nutshell
-Trajectories are replicated or discarded in the ensemble according to some \emph{selection rules} designed to sample rare events of some observable, \textit{e.g.} the drag, with a higher frequency.
-These selection rules are such that the introduced statistical bias is known at each iteration of the algorithm.
-The generated trajectories are associated with a weight, from which one is able to compute the probability and expectation values of observables. 
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-  \caption{\label{fig:illustr_AMS} Selection-mutation procedure of the \ac{ams} algorithm with $N=3$.
-Lines 1, 2 and 3 represent the evolution of the score function for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-\end{figure}
-
-% Introduction of AMS
-%
-The \acl{ams} algorithm~\cite{cerou_adaptive_2007} builds on previous ideas of {splitting algorithms} \cite{KahnHarris1951}.
-The sampling of a rare event is made easier by splitting the dynamical path of the system (in phase space) into 
-%a sequence of events that can be sampled with a higher probability  
-a sequence of shorter excursions \cite{glasserman_multilevel_1999,rolland_statistical_2015}.
-The related selection-mutation step of this algorithm {relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t),t)$}.
-The operating principle is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-Iterations of the algorithm consist in discarding the trajectories with the lowest maxima of the score function during the time interval $0\leq t \leq T_a$.
-Discarded trajectories are re-sampled according to the remaining trajectories.
-See~\cite{cerou2019adaptive} for a review of the \ac{ams} algorithm, its history and applications.
-For recent applications of the \ac{ams} in the context of fluid turbulence, see \cite{bouchet2019rare,rolland_extremely_2018}.
-{Let us mention that we use a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-  It can however be formulated in terms of the ``classical \ac{ams}''~\cite{lestang_computing_2018}.
-  The following results are not limited to this particular variant.}
-
-\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-\label{sec:ams_ou}
-% Example: OU process
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-    \caption{Efficiency of the \ac{tams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \cite{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{TAMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-  \label{fig:comparaison_temps_de_retour}
-\end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-\begin{equation}
-  \label{eq:ou}
-  \dot{x} = -x + \eta (t),
-\end{equation}
-where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%
-The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-In the latter, the typical computational cost is simply the return time $r(a)$.
-One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features do no \textit{a priori} persist in the case of fluid dynamics.
-
-\subsubsection{The \ac{ams} for extreme drag fluctuations}
-\label{sec:ams_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maximum of the instantaneous drag throughout the re-sampled trajectories as a function of the corresponding computational cost $C_{TAMS}$.
-The \ac{ams} is unable to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-Our aim is now to use the \ac{ams} algorithm to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d$ acting on the square obstacle.
-In contrast with the \acl{ou} process, the phase space is here highly-dimensional with intrinsically chaotic dynamics.
-In that situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ is no longer straightforward.
-However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.
-%
-{Let us also stress that the dynamics is here deterministic.
-Therefore, randomness must be artificially introduced for the re-sampling procedure to be effective, \textit{i.e.} to generate new trajectories associated to larger drag fluctuations.
-In practice, a small perturbation in the state of the flow has been introduced at the branching points (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-%
-Due to the chaotic behavior of the flow, re-sampled trajectories then separate from their parents.
-% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-We assume that this perturbation at branching times has a negligible impact on the statistical properties of the sampled rare events.
-The validity of this assumption has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-It was checked that the obtained statistics of the drag were consistent with the statistics computed from the (unperturbed) control simulation described in section~\ref{sec:pdfs}.}
-%\EL{Freddy: provide reference for Lyapunov timescale. \#12}
-
-% application
-%
-The \ac{ams} algorithm is applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%
-Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost, similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in figure~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-%(see Fig.~\ref{fig:comparaison_temps_de_retour}).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-  \caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the selection-mutation procedure.
-In this experiment, the \ac{ams} algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ (correlation time of the drag).
-  The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of trajectories after many iterations of the \ac{ams} re-sampling procedure. 
-%
-Let us mention that for the sake of simplicity, the trajectory of the system is here abusively associated with the time evolution of the observed drag.  
-%
-Interestingly, all trajectories ultimately originate from one unique initial trajectory. 
-%
-This feature may be explained as follows.
-It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
-%
-We observe in Fig.~\ref{fig:AMS_drag_trajectories} that the duration of extreme drag fluctuations, $\tau_c$, is much shorter than the Lyapunov's timescale $\tau_L$.
-As a consequence, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and therefore the probability of observing a new extreme fluctuation is also very low.
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
-%
-As discussed previously, the duration of extreme fluctuations of the drag are closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L \gtrsim \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-A possible line of improvement might be to explicitly account for ``the correlation time of the observable being smaller than the Lyapunov time of the system'' by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
-%
-
-\subsection{Importance sampling of extreme time-averaged drag forces with the \acl{gktl} algorithm}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \cite{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-Similarly to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \cite{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\cite{grassberger1998perm}, out of equilibrium statistical physics~\cite{PhysRevLett.118.115702}, computer science~\cite{aldous1994go}, dynamical systems~\cite{tailleur_probing_2007}, quantum mechanics~\cite{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006 \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \cite{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \cite{giardina_direct_2006,tailleur_probing_2007}.
-
-\subsubsection{The \ac{gktl} algorithm}
-\label{sec:gktl_description}
+\section{The \ac{gktl} algorithm}
+\label{app:gktl}
 The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
@@
 A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics.
-All trajectories in the resampled ensemble are solutions of the original dynamical.
-%This is true for stochastic dynamics.
-{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to allow clones of a same trajectory to separate. 
-As for the \ac{ams} algorithm, it is assumed that this perturbation does not significantly affect the statistics of the sampled trajectories.}
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
+All resampled trajectories are solutions of the original dynamical system.
+Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+It was checked that this perturbation does not affect the statistics of the sampled trajectories, see appendix~\ref{app:perturb_branching_time}.
+
 Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
 
-\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-	The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is now considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-The purpose is to sample trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.
-{Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.}
-
-% Choice of parameters and perturbation
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-In the following experiments, $\tau = \tau_c /2$. 
-%
-
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-\begin{equation}
-  \label{eq:estimate_biased_measure}
-  \rho_k(F) 
-  %= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-\end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
-%
-% limitations
-%
-Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
-%
-Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \cite{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-		\hline
-		${N}_{brute-force}$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		${N}_{brute-force}$ is the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is now assessed with respect to direct sampling. 
-Table~\ref{tbl:nbTraj_tbl} indicates the number of trajectories that correspond to a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independant estimates, obtained using different values of the bias parameter $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to yield an large value of the average.
-Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, which lifetime is shorter than the timescale over which re-sampled trajectories separate from their parent. 
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix*
-\section{The \acl{lbm}}
-\label{sec:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice. 
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$.
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \cite{De_Rosis_2016}.
-
-\bibliographystyle{unsrt}
-\bibliography{biblio}
-
-\end{document}
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "draft_p2_jfm"
 %%% End:

### 2020-02-11 figures/illustr_extrms_vorticity_GKTL/resimulate_extremes.py
Trakc missing figure illustr_extrms_vorticity_GKTL.png

@@ -1,0 +1,15 @@
+import csv
+# Files extremes.csv contains the index of the events and the corresponding
+# peak drag amplitude, in units of sigma
+# First row is lowest fluctuation
+idx = []
+F = []
+with open("extremes.csv", mode='r') as extremes_file:
+        extremes_reader = csv.reader(extremes_file, delimiter=',')
+        line_count=0
+        for _idx, _F in extremes_reader:
+            idx.append(_idx)
+            F.append(_F)
+
+print(idx)
+        

### 2020-02-14 jfm_paper/simulate.cpp
Include Python.h instead of python3.7/Python.h in simulate extension

@@ -1,7 +1,7 @@
 #include <iostream>
 #include <fstream>
 #include <algorithm>
-#include <python3.7/Python.h>
+#include <Python.h>
 
 #include "libpipeLBM.h"
 

### 2020-02-14 jfm/appendix_perturbation.tex
Update appendix on perturbation

@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-  \caption[Illustration of the perturbation of the initial condition]{The blue line shows the evolution of the drag acting on the obstacle over time for a simulation initialised on a given initial condition.
-The second line represents the evolution of the drag for a simulation initialised on the same initial condition, however slightly perturbed as described in this appendix.
-$N=10$, $\epsilon=0.002$}
-  \label{fig:effect_of_perturbation}
-\end{figure}
-
 In this work, we chose $N = 10$ and $\epsilon = 0.002$.
-Figure~\ref{fig:effect_of_perturbation} illustrate the effect of applying a small perturbation on the evolution of the drag acting on the square obstacle.
 The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
@@
   \includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-  \caption{Estimation of the \ac{pdf} with and without periodic perturbations}
+  \caption{Estimation of the \ac{pdf} with and without periodic perturbation.
+    Note that the \ac{pdf} with perturbation is based on a timeseries that is ten times shorter ($T_{tot} = 10^5 \tau_c$) than the one used to estimate the blue \ac{pdf} ($T_{tot} = 10^6 \tau_c$), hence the aspect of the tails of the former.}
   \label{fig:pdf_drag_with_perturbation}

### 2020-02-14 figures/pdf_drag_with_perturbation/pdf_drag_with_perturbation.py

@@ -8,14 +8,16 @@ import matplotlib.pyplot as plt
 
 fig, ax = plt.subplots(constrained_layout=True, figsize=(8, 5))
 
-series = pd.read_csv("../../data/PDF_drag_one_file.csv", comment="#", index_col=0)
-series.plot(ax=ax, lw=2, legend=False)
+series = pd.read_csv("../../data/PDF_drag.csv", comment="#", index_col=0)
+trunc_series = series.iloc[np.where((series.index.values > -6.5) & (series.index.values < 8.5))] 
+trunc_series.plot(ax=ax, lw=2, legend=False)
 series_perturb = pd.read_csv("../../data/PDF_drag_with_perturbation.csv", comment="#", index_col=0)
 series_perturb.plot(ax=ax, lw=2, legend=False)
 
-ax.set_xlabel("$(f_d - \\mu)/\\sigma$", fontsize=18)
-ax.set_ylabel("$\\mathcal{P}$", fontsize=18)
+ax.set_xlabel("$(f_d - \\mu)/\\sigma$", fontsize=22)
+ax.set_ylabel("$\\mathcal{P}$", fontsize=22)
 ax.legend(["No perturbation", "$\\epsilon = 0.002$"], fontsize=18)
+ax.tick_params(axis='both', which='major', labelsize=18)
 
 ax.set_yscale("log")
 fname = splitext(__file__)[0] + ".png"

### 2020-02-26 jfm/appendix_lbm.tex

@@
 % details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice (see \cite{kruger_lattice_2017} for a comprehensive introduction).
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
@@
 \]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \citep{BGK}.
 %
@@
 This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
+$\Delta x \to 0$ \citep{succi_book}.
 

### 2020-02-26 jfm/draft_p2_jfm.tex
Minor fixes of typos and references

--- 2020-02-06 jfm/draft_p2_jfm.tex
@@
 \usepackage{subfig}
-\usepackage{caption}
+%\usepackage{caption}
 
@@
 	long=Adaptive Multilevel Splitting,
+}
+\DeclareAcronym{tams}{
+	short=TAMS,
+	long=Trajectory Adaptive Multilevel Splitting,
 }
@@
 \newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
+\newcommand{\QQ}[1]{{\color{green}{#1}}}
 
-\title{\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
+\title{
+	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
+\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
 }
@@
 \begin{abstract}
-	\ZZ{This study investigates by means of direct numerical simulation, extreme drag events exerted by a  turbulent flow on a bluff body and examine the relevance of statistical algorithms to sample efficiently these events.
-Two representative algorithms are examined: AMS and GKTL.
-Briefly mention differences and complementarity.
-What do we learn?
-The take-home message is...}
-\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
+	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
 	%
-Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-This work is the first application to fluid-structure interaction problems.
+	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
 	%
-The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
+	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
+This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
 	%
-Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
+	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
+	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
+	%
+
+		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
+		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
+Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
+
+	%
+	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
+	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
+	%
+	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
+	% 
+	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
+	%
+	\QQ{phrase de conclusion ?}
+}
+%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
+%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
+%%
+%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
+%This work is the first application to fluid-structure interaction problems. 
+%%
+%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
+%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
+%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
+%%
+%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
 }
@@
 
-\include{introduction}
+\section{Introduction}
 
-\include{section_description_case_study}
+% general comments on physical problem %
+%
+Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
+Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
+``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
+In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
+%
 
-\include{section_extremes_direct_sampling}
+From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
+The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
+Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
+However, this requires very long simulations since these events are rare.
+% general def on rare-event sampling %
+%
+{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+%
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
+%
 
-\include{rare_events_algorithms}
+% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
+%
+In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
+In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
+%
+An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
+Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
+Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
+An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
+%
+However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
+%
+%
+%
+In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
+It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
+%    
+%
+Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
+More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
+%
+%
+\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+%
+Two algorithms a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
+%
+
+%
+% MOVED TO SECTION 4
+%
+%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
+%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
+The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
+More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+%
+%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
+%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
+%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
+%%
+%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
+%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
+Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
+%
+
+The paper is organized in two parts.
+The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
+This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
+%
+Motivation for this study is twofold.
+Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
+This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
+Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
+This assessment is developed in the second part of the paper.
+
+% Annonce du plan
+In more details,  the flow set-up is introduced and the dynamics related to typical
+drag fluctuations is described in section~\ref{sec:test_flow}.
+The statistical properties of the drag are then discussed.
+In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
+	drag fluctuations} is investigated based on a direct sampling approach.
+Both the instantaneous drag and time-averaged drag are considered.
+It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
+Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
+%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
+Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
+In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
+This difficulty is put in perspective with the phenomenology developed in the previous sections.
+Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
+This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
+As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
+Perspectives and conclusion end this work.
+
+\section{Description of the numerical case study}
+\label{sec:test_flow}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
+	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
+The flow is artificially damped near the end of the channel.
+In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
+The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
+	\label{fig:illustr_ecoulement}
+\end{figure}
+
+% introduce the flow 
+%
+The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
+% why this flow?
+%
+Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
+% 
+We believe that this system embeds the characteristic features that makes 
+%\sout{the application of rare-event algorithms}
+ \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
+%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
+%
+% rapid description of the flow
+Turbulent eddies generated in the near-wake of the grid are carried downstream.
+They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
+The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
+%
+%Finally, the obstacle does not deform or move either. 
+%
+%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
+%
+
+% lattice Boltzmann method 
+%
+The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
+While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
+Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
+%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
+Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
+Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
+In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
+
+% geometry
+%
+The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
+The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
+The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
+% boundary conditions
+%
+No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
+%
+Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
+The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
+In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
+The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
+Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
+Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+
+\subsection{The drag force}
+\label{sec:drag_force}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
+	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
+		The vorticity is given in lattice units.}
+	\label{fig:typical_vorticity}
+\end{figure}
+
+%% LB parameters  %
+
+% drag signal
+%
+The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
+The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
+\begin{equation}
+\label{eq:drag_definition}
+f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
+\end{equation}
+where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
+Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
+% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
+%
+Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
+As a consequence, the drag can eventually be expressed as the difference 
+\begin{equation}
+\label{eq:drag_approx}
+f_d(t) = p_{fb}(t) - p_{base}(t)
+\end{equation}
+between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
+Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
+Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
+%
+% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
+%
+
+% Definition of turnover time
+%
+The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
+\begin{equation}
+\label{eq:turnover_time}
+\EL{T_0 = \frac{R}{U}},
+\end{equation}
+where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
+%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
+%
+% typical fluctuation of the drag
+%
+Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
+%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
+\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
+Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
+
+\subsection{The drag as a random process
+	%: Probability Density Function and correlation in time
+}
+\label{sec:pdfs}
+
+%\ZZ{As requested by referee 3, Fig.3 has been removed}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
+%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
+The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
+%	\label{fig:typical_drag_signal}
+%\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
+	\caption{
+	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
+The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
+If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
+The selected peak fluctuations are therefore well separated.}
+		\label{fig:illustrate_return_time}
+\end{figure}
+
+% drag signal
+%
+% Description du signal de trainee typique
+In Fig.~\ref{fig:illustrate_return_time},
+the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
+Therefore, it is natural to model the drag as a (scalar) random process.
+
+% drag statistics
+% pdf
+Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
+This long simulation will be referred to as the \textit{control run} in the following.
+It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
+The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
+%
+Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
+%
+In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
+This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
+\begin{figure}
+	\centering
+	\subfloat[\ac{pdf} of normalised drag fluctuations]
+	{\label{fig:pdf_drag_a}
+		\includegraphics[width=.45\linewidth]{PDF_drag/PDF_drag.png}}
+	\subfloat[Autocorrelation of drag fluctuations]
+	{\label{fig:pdf_drag_b}
+		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
+	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
+The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
+%Note that the amplitudes have not been normalized.
+		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
+		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
+The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
+		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
+	}
+	\label{fig:pdf_drag}
+\end{figure}
+
+% drag statistics
+% correlation time
+Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
+It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+%
+This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
+%
+In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
+The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
+The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
+
+\section{Extreme fluctuations of the drag by means of direct sampling}
+\label{sec:direct_sampling}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
+	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
+	}
+	\label{fig:return_time_instant}
+\end{figure}
+
+% return time given from direct sampling
+%
+The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
+Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
+It will provide a trustworthy baseline for the validation of rare-events algorithms.
+
+The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
+The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
+As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
+For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
+
+How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
+We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
+%
+%
+%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
+If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
+% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
+one might expect that 
+\begin{equation}
+\label{eq:return_time}
+\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
+\end{equation}
+where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
+Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
+Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
+Let us point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
+
+% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
+
+\subsection{Extracting extreme drag fluctuations from a very long timeseries}
+\label{sec:extreme_extraction}
+
+% intro
+%
+We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
+This set will be considered as representative of \emph{extreme events} in the upcoming study.
+The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
+%
+According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma \equiv\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
+Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
+Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
+%
+In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
+
+\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
+\label{sec:instantaneous_drag}
+
+\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
+\label{sec:forebody_and_base_contribution}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
+	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
+An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
+		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
+The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
+	}
+\end{figure}
+
+% teasing
+%
+%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
+We shall see that the situation is different in the case of {extreme} drag fluctuations.}
+\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
+We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
+}
+
+% relative contribution of forebody and base pressure
+%
+Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
+The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
+\begin{equation}
+\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
+\end{equation}
+where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
+%
+Fig.~\ref{fig:pressure_ratio} displays the relative contributions
+$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
+	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from the drop of the base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
+On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
+
+%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
+Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
+Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
+
+\subsubsection{Fluid dynamics related to extreme drag fluctuations}
+\label{sec:dynamical_aspects}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
+	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
+The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
+Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
+The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
+\end{figure}
+
+% mean profile around burst
+%
+The focus is now on the flow scenarii that yield extreme values of the drag. 
+Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
+A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
+This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
+%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
+Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
+%This is reminiscent of time-irreversibility in turbulent dynamics. 
+\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
+%Moreover, such very high drag levels do not persist over time.
+To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
+	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
+	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
+In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
+	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
+\end{figure}
+% flow scenario
+%
+Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
+%
+In each case, an intense vortical structure is visible near the base of the obstacle.
+The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
+%+
+The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
+%
+This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
+In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
+
+The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
+As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
+% 
+This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
+%
+Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
+This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
+As the blocking vortex is in turn advected downstream, the vortex against the base is released.
+%
+Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
+This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
+This scenario is generic and has been observed for most extreme events sampled in the control run.
+
+%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
+	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
+Each trajectory corresponds to a single event.
+The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
+\end{figure}
+
+Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
+\begin{equation}
+\label{eq:avg_shear_def}
+\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
+\end{equation}
+where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
+%
+% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
+% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
+%
+Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
+Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
+On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
+%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
+These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
+This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
+In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
+
+\subsection{Extreme fluctuations of the time-averaged drag }
+\label{sec:time_avg}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
+	\label{fig:extreme_avg}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
+		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
+				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
+	}
+	\label{fig:PDF_AVG}
+\end{figure}
+
+% motivation
+%
+%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
+In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
+
+%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
+We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
+%
+In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
+% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
+%
+% definition
+%
+Therefore, another relevant observable is the \textit{time-averaged} drag defined as
+\begin{equation}
+\label{eq:def_time_averaged_drag}
+F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
+\end{equation}
+where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
+%\sout{Relevant values of $T$ are application dependant.}
+\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation of the instantaneous drag.
+In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
+
+%
+During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
+% phenomenology
+%
+% What is the phenomenology leading to extreme values of $F_T(t)$ ?
+It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
+%
+%
+%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
+In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
+By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
+\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
+As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
+
+% neither case (1) nor case (2)
+%
+Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
+\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
+Indeed, both cases are %equally 
+featured in Fig.~\ref{fig:extreme_avg};
+very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
+
+%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
+
+% theoretical explanation is removed because it is speculative
+
+%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
+%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
+%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
+%\begin{equation}
+%\label{eq:indep}
+%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
+%\end{equation}
+%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
+%\begin{equation}
+%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
+%\label{eq:ratio_exp}
+%\end{equation}
+%The observation is therefore well supported theoretically.
+%See~\cite{lestang2019JSTAT} for further details.
+
+%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
+
+\section{Rare-event sampling algorithms}
+\label{sec:rare_events_algorithms}
+
+% limitation of direct sampling
+%
+In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
+Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
+As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
+%% objective
+% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
+The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.
+
+\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
+\medskip
+\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
+To what extent they are representative of rare-event sampling algorithms?
+Complementarity of the two algorithms?
+Make the link between fluid mechanics and statistical algorithms ?}
+\medskip
+
+\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
+Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common.
+The algorithm tracks the value of the ratio of the probability of the new ensemble with the probability of the previous one, in order to be able to estimate the probability of the extreme event in the model.
+Different algorithms use different selection rules.
+The success of the algorithm depends on the quality of the selection rule in order to select the precursors of the extreme event.
+Optimal selection rules are called committor functions, see for instance~\citep{lestang_computing_2018} and references therein.
+In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
+They both proved efficient for various dynamics and adopt opposite strategies:  the \acl{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \acl{gktl} uses as a selection rule based on the increment of a score function.} \ZZ{A complete description of these two algorithms and their operating details is given in appendixes.}
+
+%
+%\ZZ{paragraph moved from introduction}
+The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
+In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
+or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
+%
+% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
+ More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+\ZZ{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
+%
+
+During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
+Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
+Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for \ZZ{estimating the probability of observables which are of time integrals over a very long time \citep{giardina_simulating_2011,Laffargue_2013}.
+%
+Recently the \ac{gktl} algorithm has proven extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
+This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
+
+\ZZ{The aim of this part of the work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which a completely new field of application.}
+
+% These two algorithms will be detailed in the following.
+% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
+% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
+% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
+
+\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
+\label{sec:ams}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
+%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
+On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
+Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
+It is then simulated from the branching point to the final time $T_a$.
+In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
+Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
+%\end{figure}
+
+% Introduction of AMS
+%
+%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
+\ZZ{In the version of the \ac{ams} algorithm we use in this work, we consider %a sequence of events that can be sampled with a higher probability  
+trajectories with a fixed duration $T_a$~\citep{lestang_computing_2018}.%
+%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
+%
+% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
+%
+% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
+%
+% Quick description of TAMS
+%
+ At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
+%The discarded trajectories are then re-sampled according to the remaining trajectories.
+These trajectories are re-sampled by branching them from the remaining trajectories.
+%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
+\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}
+
+%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
+However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
+Our results are not limited to this particular variant.
+
+%
+% SUBSECTION IS MOVED TO APPENDIX %
+%
+%\subsubsection{}
+%\label{sec:ams_ou}
+%% Example: OU process
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
+%	\label{fig:comparaison_temps_de_retour}
+%\end{figure}
+%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+%\begin{equation}
+%\label{eq:ou}
+%\dot{x} = -x + \eta (t),
+%\end{equation}
+%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+%This basic stochastic process will allow us to highlight differences with fluid dynamics.
+%
+%
+%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+%%
+%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+%%
+%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
+%In the latter, the typical computational cost is simply the return time $r(a)$.
+%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+%
+%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
+
+%\subsubsection{The \ac{ams} for extreme drag fluctuations}
+%\label{sec:ams_drag}
+
+%
+% Application to instantaneous drag
+%
+% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
+% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
+%
+%
+%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
+
+The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
+%\sout{In contrast with the \acl{ou} process, } 
+%The phase space is highly-dimensional with deterministic chaotic dynamics.
+%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
+%\sout{is no longer straightforward} 
+%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
+%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
+We use the observable itself as the score function : $\xi (\mathbf{x}(t),t) = f_d(t)$.  
+%
+Let us also stress that the Navier--Stokes dynamics is deterministic.
+\ZZ{Then, small random perturbation are artificially introduced at branching points. This procedure is detailed in Appendix \ref{app:perturb_branching_time}}.
+% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
+%In practice, a small perturbation in the state of the flow has been introduced 
+%at branching points.
+% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
+	%
+Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov time.}
+	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
+\ZZ{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistical properties of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
+It was checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation.
+%\EL{Freddy: provide reference for Lyapunov timescale.
+\#12}
+
+% application
+%
+%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
+%%
+%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
+%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
+%%
+%%
+%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
+%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
+%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
+%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
+%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
+
+\EL{In addition to the score function, two important parameters are the number $N$ and the duration $T_a$ of the trajectories.
+The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
+Therefore, $N$ should be taken as large as possible to reduce these errors with a practical limit given by the available computational resources.
+In this work, we have performed two numerical experiments with $N=32$ and $N=256$.} \ZZ{ The value of $T_a$ should chosen much larger than $\tau_c$~\citep{lestang_computing_2018}, but be kept small enough in order to minimize the computation cost.
+The duration of the trajectories was set to $T_a=5\tau_c$ in both cases.
+We have checked that note that large values of $T_a$ do not affect the result.}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
+	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
+In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
+Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
+		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
+	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
+The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.
+\QQ{Change TAMS to AMS in horizontal axis}}
+\end{figure}
+
+%pitfall
+%
+\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm.} 
+%
+Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
+%
+The algorithm therefore fails to enhance the sampling of extreme events. 
+This result is also confirmed by increasing the number of initial trajectories to $N=256$.
+%
+The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
+%
+% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
+%
+%
+the distribution of the maximal drag for the {initial} trajectories is also shown.
+We observed that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
+%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
+However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
+%
+
+It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
+In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
+%
+\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
+After $\tau_c$, the vortex is freed and carried away by the flow, and further large fluctuations of the drag can only result from the trapping of other vortices.
+%
+Fig.~\ref{fig:AMS_drag_trajectories} indicates that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
+%
+%
+Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
+For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, therefore, the probability of observing a new extreme fluctuation is also very low.
+%
+%
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
+%
+As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
+On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
+The ratio $u_{rms}/U$, called turbulence rate, is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L > \tau_c$.
+
+%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
+Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
+
+% perspectives
+%
+In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
+%
+%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
+%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
+%
+%
+%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
+%
+%Furthermore, such ``tuning'' is problem dependent.
+%
+\EL{The phenomenology described before is independent of the choice of $N$ and $T_a$.
+Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories.
+}
+Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
+That way, new events associated to higher fluctuations may have the opportunity to develop.
+However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
+%
+
+\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
+\label{sec:gktl}
+The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
+The \ac{ams} algorithm could be used in the same way as before
+%in section~\ref{sec:ams_drag} \textit{i.e.} 
+by taking the time-averaged observable itself as the score function.
+However, this would lead to similar unsatisfactory results.
+% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
+%
+For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
+%
+\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
+% In this case, however, the trajectories interact dynamically. 
+%
+At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
+%
+The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
+%
+The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
+{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
+The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
+Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
+\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
+
+%\bigskip
+%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
+%\bigskip
+
+%\subsubsection{The \ac{gktl} algorithm}
+%\label{sec:gktl_description}
+%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%%
+%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
+%%
+%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
+%%
+%\begin{equation}
+%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+%\label{eq:Weight}
+%\end{equation}
+%so that $\sum_{n=1}^N W_n^i = N$.
+%%
+%%
+%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
+The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
+%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+%%
+%The sampled biased distribution writes
+%%
+%\begin{align}
+%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
+%\label{eq:Biased_Path_Approximation}
+%\end{align}
+%where
+%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
+%refers formally to the probability of observing the trajectory 
+%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
+%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+%%
+%One can mention that
+%\begin{equation}
+%\label{eq:mean_field}
+%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
+%\end{equation}
+%with $\mathbb{E}_{0}$ being the expectation value with respect to the
+%distribution $\mathbb{P}_{0}$.
+%This result relies on the \textit{mean-field approximation}
+%\begin{equation}
+%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
+%\label{eq:Mean_Field_Approximation}
+%\end{equation}
+%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
+%%
+%Rejected trajectories are discarded from the statistics.
+%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+%
+%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
+%%This is true for stochastic dynamics.
+%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+%%
+%% 
+%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
+%%
+%%
+%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+%%
+
+%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
+%\label{sec:gktl_drag}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
+	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
+The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
+		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
+	}
+\end{figure}
+
+The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
+\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.}
+%Referring to the notations used in Appendix \ref{app:gktl_description}, 
+Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.    
+%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
+%
+\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period.
+Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details).
+This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
+The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
+}
+
+%
+% Choice of parameters and perturbation
+
+\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
+The higher $k$, the larger is the bias involved in the statistical resampling, as illustrated in Fig.~\ref{fig:IS_GKTL}.
+As for the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
+Finally, the cloning period $\tau$ determines how often the resampling is performed.
+The choice of the cloning period $\tau$ is important in practice.
+A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
+As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$. In the following experiments, $\tau = \tau_c /2$.}
+
+\QQ{Thibault, pour la suite, nous sommes restés sur une version proche du papier initial, par simplicité.
+Si tu penses que certains points que tu avais ajoutés sont vraiment nécessaire pour répondre au referee, dis le nous.}
+
+The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+%The choice of the cloning period $\tau$ is important in practice.
+%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
+%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
+%In the following experiments, $\tau = \tau_c /2$. 
+%
+%
+Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
+%
+%
+%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
+%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
+%
+%
+Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
+\begin{equation}
+\label{eq:estimate_biased_measure}
+\rho_k(F) 
+%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
+\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
+\end{equation}
+in addition to the unbiased \ac{pdf} based on the control run.
+As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
+%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
+%
+% limitations
+%
+Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
+%
+For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
+%
+Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
+%
+An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
+This issue is addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
+%
+% how to assess the value of k_max ?
+In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
+This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
+
+\begin{table}
+	\centering
+	\begin{tabular}{|l|c|c|c|c|c|}
+		\hline
+		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+		\hline
+		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
+		\hline
+		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
+		\hline
+		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
+		\hline
+		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
+		\hline
+	\end{tabular}
+	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
+		Note that trajectories that overlap for more than half of their duration are counted as only one.
+		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
+	\label{tbl:nbTraj_tbl}
+\end{table}
+
+% Comparisons with direct sampling
+%
+The efficiency of the \ac{gktl} algorithm is  assessed in  
+Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
+Note that trajectories that overlap for more than half of their duration are counted as one.
+%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
+Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
+%
+%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
+%
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+%  \caption{PDF of average drag over $10\tau_c$}
+%  \label{fig:PDF_AVG}
+%\end{figure}
+%
+% Conclusion
+%
+We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
+
+%
+% PARAGRAPHE SUIVANT INUTILE ?
+%
+%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
+By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
+However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
+Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
+
+\ZZ{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for extreme trajectories resampled by the \ac{gktl} algorithm.
+%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
+Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
+% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
+	\caption{Drag timeseries corresponding to the highest fluctuations of the average drag in the ensemble of resampled trajectories.
+The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
+}
+	\label{fig:timeseries_extrms_AVG_GKTL}
+\end{figure}
+In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
+sampled by the cloning algorithm.
+It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
+Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
+	\caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
+	\label{fig:illustr_extrms_vorticity_GKTL}
+\end{figure}
+
+%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
+%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
+%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
+%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
+%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
+}
+
+\subsubsection{Computation of return times}
+\label{sec:return_times}
+
+Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
+events unreachable by a direct approach.
+%(see Fig.~\ref{fig:illustrate_return_time})
+
+In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
+Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
+\begin{equation}
+\label{eq:time_averaged}
+F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
+\end{equation}
+and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
+\begin{equation}
+r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
+\end{equation}
+
+The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
+\begin{equation}
+\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
+\end{equation}
+with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
+
+Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
+The figure shows two independent estimates, obtained using different values of the bias parameter $k$.
+Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
+In addition, \EL{Fig.~\ref{fig:return_times_gktl}} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
+Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+This is  obviously a major advantage of this rare-event sampling algorithm.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
+	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
+$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
+The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
+The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
+Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
+Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
+\end{figure}
 
@@
 \appendix
+\section{The \acl{lbm}}
+\label{app:lbm}
 
-\include{appendix_lbm}
-\include{appendix_ams}
-\include{appendix_gktl}
-\include{appendix_perturbation}
+% details about LBM
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
+In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
+Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
+\[
+\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
+\]
+for the mass density and the fluid momentum respectively.
+The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
+
+% algo
+%
+The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
+The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
+Namely, an \emph{exact} streaming step
+\[
+f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
+\]
+during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
+\[
+f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
+\]
+which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
+The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
+\[
+\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
+\]
+This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
+%
+The equilibrium function is given by
+\[
+f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
+\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
+\] 
+with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
+This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
+$\Delta x \to 0$ \EL{\citep{succi_book}}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
+	\caption{Sketch of the D2Q9 lattice.
+Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
+By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
+	\label{fig:D2Q9}
+\end{figure}
+
+As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
+\[
+\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\]
+so that the total stress expresses as
+\begin{equation}\label{eq:def_stress}
+\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
+\end{equation}
+Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
+
+\newtheorem{theo}{Theorem}
+\section{The \ac{ams} algorithm -- nouvelle version Thibault}
+\label{app:ams}
+
+\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
+Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
+Les referees n'ont pas demandé de détailler les algos.}
+
+This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
+mathematical results and discusses its use in practice.
+
+\subsection{Description of the algorithm}
+An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
+%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
+%
+From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
+
+To start off with, each of these trajectories is associated a weight $w_0=1$.
+Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
+\begin{equation}
+\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
+\end{equation}
+Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
+\begin{equation}
+\label{eq:ams_threshold_def}
+\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
+\end{equation}
+One might expect intuitively that $\ell_j=1$.
+However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
+The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
+Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
+For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
+For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
+The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
+The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
+	\caption{Illustration of one iteration the \ac{ams} algorithm.
+Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
+Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
+Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
+	\label{fig:AMS_schema}
+\end{figure}
+
+After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
+$\tilde{J} = \sum_{j=1}^J \ell_j$.
+Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
+\subsection{How to compute the probability}
+Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
+\begin{equation}
+w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
+\end{equation}
+This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
+Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
+Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
+Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
+\subsection{Summary of mathematical results}
+We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
+of iterations $N$.
+We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
+The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
+\begin{equation}
+\label{eq:estimator_ams}
+q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
+\end{equation}
+In the following we denote this estimator $\hat{q}_N(a)$.
+One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
+\begin{theo}
+	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
+	\begin{equation}
+	\mathbb{E}[\hat{q}_N]=q.
+	\end{equation}
+\end{theo}
+Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
+
+More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
+\begin{equation}
+\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
+\end{equation}
+with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
+The minimal variance $-q^2\ln q$ is obtained when choosing
+\begin{equation}
+\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
+\label{eq:time_dependent_committor}
+\end{equation}
+for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
+
+\subsection{How to use it in practice}
+In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
+Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
+Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
+
+Another parameter of the algorithm is the duration $T_a$.
+This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
+Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
+This can be explained as follows.
+Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
+after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
+As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
+This amounts to direct sampling.
+
+Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
+Rather, iterations were carried out until all the trajectories in the ensemble overlap.
+Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
+\begin{equation}
+\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
+\end{equation}
+and the iterations stop.
+This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
+
+\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+\label{app:AMS_on_OU}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
+	\label{fig:comparaison_temps_de_retour}
+\end{figure}
+Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+\begin{equation}
+\label{eq:ou}
+\dot{x} = -x + \eta (t),
+\end{equation}
+where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+This basic stochastic process will allow us to highlight differences with fluid dynamics.
+
+The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+
+The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
+In the latter, the typical computational cost is simply the return time $r(a)$.
+One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+
+Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
+
+\newpage
+
+\section{The \ac{ams} algorithm -- version initiale}
+\label{app:ams-short}
+
+\subsection{The operating principle}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
+	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
+$\xi(x(t))$ is a prescribed score function defined along each trajectory.
+	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
+\end{figure}
+
+The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
+%
+The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
+On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
+Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
+It is then simulated from this branching point to the final time $T_a$.
+In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
+This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
+
+\subsection{Application to a simple case: the \acl{ou} process}
+
+	\begin{figure}
+		\centering
+		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+		\caption{\QQ{Change TAMS TO AMS} Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
+The blue line is the analytical solution for the return time of amplitude $a$.}
+		\label{fig:comparaison_temps_de_retour}
+	\end{figure}
+	A one-dimensional \acl{ou} process is considered:
+	\begin{equation}
+	\label{eq:ou}
+	\dot{x} = -x + \eta (t),
+	\end{equation}
+	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
+
+	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
+	%
+	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
+	%
+	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
+	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
+	In the latter, the typical computational cost is simply the return time $r(a)$.
+	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+
+	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
+	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
+	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+Such favourable features \EL{do not} persist in the case of fluid dynamics.
+
+\newpage
+
+\QQ{Appendix sur GKTL à valider : Freddy ?}
+
+\section{The \acl{gktl} algorithm -- version initiale}
+\label{app:gktl_description}
+The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
+%
+Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
+%
+At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
+%
+\begin{equation}
+W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
+\label{eq:Weight}
+\end{equation}
+so that $\sum_{n=1}^N W_n^i = N$.
+%
+%
+{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
+The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
+For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
+The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
+%
+The sampled biased distribution writes
+%
+\begin{align}
+\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
+\label{eq:Biased_Path_Approximation}
+\end{align}
+where
+$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
+refers formally to the probability of observing the trajectory 
+$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
+The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
+%
+One can mention that
+\begin{equation}
+\label{eq:mean_field}
+Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
+\end{equation}
+with $\mathbb{E}_{0}$ being the expectation value with respect to the
+distribution $\mathbb{P}_{0}$.
+This result relies on the \textit{mean-field approximation}
+\begin{equation}
+R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
+\label{eq:Mean_Field_Approximation}
+\end{equation}
+where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
+The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
+%
+Rejected trajectories are discarded from the statistics.
+Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
+
+ \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
+%This is true for stochastic dynamics.
+Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
+	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+%
+% 
+%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
+We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
+%
+%
+Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
+%
+
+\newpage
+
+\EL{
+\section{Perturbation at branching time}
+\label{app:perturb_branching_time}
+%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
+The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
+resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
+
+%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
+A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
+In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
+%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
+%
+Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
+\begin{equation}
+f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
+\label{eq:perturb_pop}
+\end{equation}
+where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
+%
+%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
+The perturbed populations are then rescaled so that no mass is added in the system.
+% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
+In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
+Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
+%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
+%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
+%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
+The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
+%	\label{fig:effect_of_perturbation}
+%\end{figure}
+%
+In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.  \QQ{is it correct?}.
+%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
+We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
+%
+In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ have been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
+%
+%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
+In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
+Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
+
+\begin{figure}
+  \centering
+\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
+\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.} \QQ{indicate $x=(f_d-\mu)/\sigma$ and ${\cal P}(x)$  as before}}
+\label{fig:pdf_drag_with_perturbation}
+\end{figure}
+}
 
@@
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2020-02-26 jfm/introduction.tex

@@
 By ``extreme'' we refer to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, better understanding of the nature of such extreme dynamical events and their statistics is crucial for the control of excessive mechanical efforts, for instance extreme wind loads on tall structures such as buildings of wind turbines~\cite{kanev2019}.
+In engineering, better understanding of the nature of such extreme dynamical events and their statistics is crucial for the control of excessive mechanical efforts, for instance extreme wind loads on tall structures such as buildings of wind turbines~\cite{kanev2010}.
 %
@@
 %
-More generally, the computational cost for sampling a fluctuation of very small probability grows like the inverse of this probability, and simulating rare events through direct numerical simulation is therefore not a scalable approach.
+More generally, the computational cost for sampling a fluctuation of very small probability grows like the inverse of this probability~\cite{wouters2016rare}, and simulating rare events through direct numerical simulation is therefore not a scalable approach.
 This issue is addressed by rare-event sampling algorithms, referring to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.

### 2020-02-26 jfm/rare_events_algorithms.tex

\section{Rare-events algorithms}
\label{sec:rare_events_algorithms}

According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.

In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
computational cost much lower than their return time.
Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
greater complexity than the two-dimensional flow discussed in the previous sections.

We focus on two algorithms, representatives of two popular approaches in rare event simulation.
The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. More recently, it has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
More recently it allowed the numerical simulation of extreme heat waves in a simplified model of the atmosphere \citep{ragone_computation_2018}.
This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
Along the same line, rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow.

The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
system.
Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
committor function. However, with the exception of rare analytical cases, this function is unknown.
See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.

Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
However, resampling steps are performed along the simulation of the ensemble.
Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
\begin{align}
\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
\label{eq:Biased_Path_Approximation_main}
\end{align}
where
$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory
$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
See appendix~\ref{app:gktl} for a more detailed description of the cloning algorithm.

Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
rules to bias the sampling towards extreme events.
In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
events of interest.
Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.

Most previous applications of the \ac{ams} and cloning algorithm target stochastic dynamics.
We stress that this is not the case in this work, as the target flow model is fully deterministic.
Because in both algorithms the initial condition of resampled trajectories is taken as the flow state of another, it is therefore required to add a random perturbation to the initial conditions of resampled trajectories, so that they separate from the trajectory they are resampled from.
The amplitude of this perturbation is chosen very low, and it was tested that it does not impact the statistics of the sampled trajectories.
The separation of trajectories under this slight perturbation is guaranteed by the chaoticty of the flow.
For further details about the addition of this perturbation, see appendix~\ref{app:perturb_branching_time}.

\subsection{The \acl{ams} algorithm}
\label{sec:ams}
We applied the \ac{ams} algorithm, described in appendix\ref{app:ams}, to the sampling of extreme fluctuations of the drag applied on the square obstacle immersed in the flow described in section~\ref{sec:test_flow}.
We chose to set the score function as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
Let us stress that we do not expect this choice to be close to the optimal "committor" function.
Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
into the physics that lead to the extreme events of interest.
Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
score function, a choice that is often the only available one.

In addition to the score function, the parameters of the \ac{ams} are the number of trajectories $N$ and the duration of trajectories $T_a$.
As detailed in appendix~\ref{app:ams}, the size of the ensemble $N$ governs the statistical error affecting the estimators based on the sampled set of trajectories.
Furthermore, the number of resamplings required to reach a final ensemble of trajectories of probability of the order of $10^{-\beta}$ can be shown to scale like $\beta N$ \citep{lestang_computing_2018}.
In this work, the choice of $N$ was governed by the available computational resources.
We performed two experiments with $N = 32$ and $N = 256$, having in mind that higher numbers are unlikely to be accessible in the case of more complex or three-dimensional flows.
The duration of the trajectories was set to $T_a$ = $5\tau_c$, a value that is larger than the timescale of extreme drag fluctuations whilst maintaining an affordable computational cost.
As pointed out in appendix~\ref{app:ams}, setting $T_a \gg \tau_c$ does not lead to a better sampling.

\begin{figure}
  \centering
  \subfloat[\label{fig:AMS_drag_trajectories_a} Final ensemble of trajectories sampled by the \ac{ams}]{\includegraphics[width=.4\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}}
  \subfloat[Drag maximum for resampled trajectories]{\label{fig:AMS_drag_trajectories_b}\includegraphics[width=.4\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}}
  \caption{(a) Ensemble of $N = 32$ trajectories after $181$ \ac{ams} resamplings.
    (b) Maximum of the instantaneous drag throughout the resampled trajectories as a function of the corresponding computational cost $C_{AMS}$, \textit{i.e.} as a function of the resamplings. The maximum over resampled trajectories overlap until saturating at the level of the largest maximum in the initial ensemble.
  }
\end{figure}

Figure~\ref{fig:AMS_drag_trajectories_a} illustrates that the splitting introduced by the \ac{ams} does not enhance the sampling of extremes but rather has catastrophic consequences on the sampling of trajectories.
Indeed, trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.
This result is confirmed by increasing the number of initial trajectories from $N=32$ to $N=256$, as shown
in figure~\ref{fig:AMS_drag_trajectories_b}.
Recalling section~\ref{sec:instantaneous_drag}, extreme drag fluctuations have a lifetime of $\tau_c$, the timescale over which a vortex is blocked near the base of the obstacle.
After $\tau_c$, the vortex is freed and swept downstream, and further large drag fluctuations can only result from the interaction of the obstacle with the incoming flow, beyond the correlation time of the drag fluctuations.
Moreover, it takes a certain time $\tau_L$ (known as the Lyapunov timescale) before a trajectory resampled by the \ac{ams} separates from its parent under the effect of the initial perturbation.
In our situation, this ``memory effect'' originates from the fact that the score function is of dimension much smaller (here 1) than the dimension of the phase space (very large for this fluid dynamics problem).
%
We observe in Fig.~\ref{fig:AMS_drag_trajectories_a} that this separation timescale is larger than the duration of extreme drag fluctuations $\tau_c$, leading to an overlap of trajectories.

We stress that this phenomenology is independent of the choice of $N$ and $T_a$.
Increasing the size of the initial ensemble can only increase the amplitude of the the global maximum attained initially, but does not solve the issue of overlapping trajectories.
Increasing $T_a$ allows further time for resampled trajectories to develop rare fluctuations, however after a duration $\Delta T \geq \tau_L > \tau_c$, and therefore decorrelated from the resampling.

Our results indicate that the \ac{ams} does not work with this simple choice of score function.
Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.

\subsection{The \ac{gktl} algorithm}
\label{sec:gktl}
We applied the cloning algorithm described, in appendix\ref{app:gktl}, to the two-dimensional flow described in section~\ref{sec:test_flow}.
Unlike the case of the \ac{ams} in section~\ref{sec:ams}, the algorithm is here used to sampled trajectories that exhibit extremes values of
the \emph{average} drag.

The cloning algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
The cloning strength is the multiplicative factor appearing the the bias term of equation~\eqref{eq:Biased_Path_Approximation_main}.
The higher $k$, the larger the typical average drag for the trajectories in the ensemble sampled by the algorithm.
There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.
Similarly to the \ac{ams}, $N$ governs the error affecting estimates of probabilities and estimators based on the biased ensemble.
In the cloning algorithm, the population of trajectories is periodically resampled as the trajectories are simulated from $t=0$ to
$t=T_a$.
The cloning period determines how often this resampling step is performed.
For deterministic dynamics, a common rule of thumb is to chose $\tau \approx \tau_c$ \citep{giardina_direct_2006}.
For further details about the cloning algorithm, see appendix~\ref{app:gktl}.

In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
Finally, we set $\tau = \tau_c / 2$.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{IS_GKTL/IS_GKTL_2}
  \caption{Histogram of the values of the time-averaged drag $F_{T_a}$, with $T_a = 10\tau_c$, in the ensemble sampled by the cloning algorithm, for three values of the cloning strength $k$. In each case, the algorithm was applied with $N=1024$, $\tau = \tau_c / 2$. The grey histogram represents the average histogram obtained through a direct sampling, \textit{i.e.} with $k=0$. It was computed using a Gaussian approximation for the \ac{pdf} of the time-averaged drag.}
  \label{fig:IS_GKTL}
\end{figure}
Figure~\ref{fig:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.
It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
The higher $k$, the stronger the bias.
The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:PDF_AVG}).
Figure~\ref{fig:IS_GKTL} also illustrates that the sampling by the algorithm is very concentrated on some fluctuations values, and that this effect increases as $k$ increases.
This results from the overlap of trajectories as a consequence of the cloning and pruning.
Indeed, if a trajectory is cloned at step $i$, then the trajectory ${\mathbf{x}(t)}$ over the interval $[0;i\tau]$ will be the same for all the clones.
For a fixed ensemble size $N$, the overlap of trajectories increases as $k$ increases as fewer
trajectories are cloned into a higher number of duplicates.
\textbf{How do we quantify this effect? It is probably application-dependant. What are the strategies to mitigate it?}

Figure~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag timeseries for several trajectories within the ensemble sampled by the cloning algorithm.
It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
  \caption{Drag timeseries corresponding to the 6 highest fluctuations of the average drag in the ensemble sampled by the cloning algorithm. The cloning algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$. }
  \label{fig:timeseries_extrms_AVG_GKTL}
\end{figure}

In addition, figure~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
sampled by the cloning algorithm.
It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
  \caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
  \label{fig:illustr_extrms_vorticity_GKTL}
\end{figure}

An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.

\subsubsection{Computation of return times}
\label{sec:return_times}

Beyond dynamical aspects, the cloning algorithm provides ensemble of trajectories over which statistics of rare events can be computed by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}.
In this section, we show that the biased sampling performed by the cloning algorithm allows for the computation of return times for events otherwise out-of-reach from a direct sampling approach.

Each trajectory $j$ in the biased ensemble results in a timeseries of the time-averaged drag:
\begin{equation}
\label{eq:time_averaged}
F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  ,
\end{equation}
and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
\begin{equation}
r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
\end{equation}

The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation_main}
\begin{equation}
  \mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{kT_aF_T^{(j)}}s_j(a),
\end{equation}
with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise. This amounts to taking the sum of the weights of the timeseries which maximum is larger than $a$.

Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
Two independent estimates are shown, obtained using different values of the cloning strength $k$.
Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
In addition, figure~\ref{fig:return_times_gktl} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the cloning algorithm allows for the computation of statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
This result shows that the biased ensemble sampled by cloning algorithm enables the estimation of return times for average drag fluctuations that would be out-of-reach from a direct sampling.
Alternatively, for a fixed target return time, the use of the cloning algorithm can reduce the computational cost of the estimation by several orders of magnitude.
This is obviously a major advantage of this rare-event sampling algorithm.

More generally, the successful computation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the cloning algorithm may be used to compute statistical estimators in a very efficient manner.
We note, however, that some estimators may be more affected by the quality of the sampling than others.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
  \caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle. $\tilde{F}_T$ denotes the time-averaged drag with zero mean. The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$. The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments. Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments. Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End:

### 2020-02-26 jfm/draft_p2_jfm.tex
Add Thibault's comments

@@
 Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
+\textbf{However, this requires very long simulations, as the computational cost for sampling a fluctuation of very small probability grows like the inverse of this probability~\cite{wouters2016rare}.}
+
 % general def on rare-event sampling %
@@
 %Note that the amplitudes have not been normalized.
-		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
+		\QQ{x-axis is now $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
 		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
@@
 As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
+\textbf{The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.}
+
 %% objective
 % In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.
+\sout{The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.}
+\textbf{In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
+computational cost much lower than their return time.
+Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
+greater complexity than the two-dimensional flow discussed in the previous sections.}
 
@@
 
-\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
+\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, \textbf{where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
+system.}
 Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common.
@@
 %
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
+% The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
+
+% introduce score function and optimal score function
+\textbf{In the \ac{ams} algorithm, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
+The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
+influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
+Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
+committor function. However, with the exception of rare analytical cases, this function is unknown.
+See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.}
+
+% Quick description of TAMS
+%
+ At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
+These trajectories are re-sampled by branching them from the remaining trajectories.
+\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}
+
+The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
+
+% Discuss naive choice for the score function, in line
+% with the objectives of the paper
+\textbf{The score function is set as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
+Let us stress that we do not expect this choice to be close to the optimal "committor" function.
+Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
+into the physics that lead to the extreme events of interest.
+Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
+It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
+score function, a choice that is often the only available one.}
+
 \ZZ{In the version of the \ac{ams} algorithm we use in this work, we consider %a sequence of events that can be sampled with a higher probability
-trajectories with a fixed duration $T_a$~\citep{lestang_computing_2018}.%
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
- At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
+
 %\sout{In contrast with the \acl{ou} process, } 
@@
 %\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-We use the observable itself as the score function : $\xi (\mathbf{x}(t),t) = f_d(t)$.  
 %
@@
 	%
-Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov time.}
+Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov \textbf{not introduced} time.}
 	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
@@
 Therefore, $N$ should be taken as large as possible to reduce these errors with a practical limit given by the available computational resources.
-In this work, we have performed two numerical experiments with $N=32$ and $N=256$.} \ZZ{ The value of $T_a$ should chosen much larger than $\tau_c$~\citep{lestang_computing_2018}, but be kept small enough in order to minimize the computation cost.
+In this work, we have performed two numerical experiments with $N=32$ and $N=256$.}
+\ZZ{The value of $T_a$ should be chosen larger than $\tau_c$~\citep{lestang_computing_2018}, but kept small enough in order to minimize the computational cost.
   The duration of the trajectories was set to $T_a=5\tau_c$ in both cases.
-We have checked that note that large values of $T_a$ do not affect the result.}
+  We checked that larger values of $T_a$ do not affect the result.}
 
@@
 The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.
-\QQ{Change TAMS to AMS in horizontal axis}}
+\QQ{Changed TAMS to AMS in horizontal axis}}
 \end{figure}
@@
 %
-\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-%
+\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after 181 iterations of the \ac{ams} algorithm.} 
+%
+\textbf{It illustrates that trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.}
+%
+This result is also confirmed by increasing the number of initial trajectories to $N=256$, see figure~\ref{fig:AMS_drag_resampling}.
+(\textbf{Leger remaniement ici})
+Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
+Although the trajectories with the lowest maxima of the score function are discarded after a few iterations and new trajectories with higher maxima are re-sampled, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
 The algorithm therefore fails to enhance the sampling of extreme events.
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observed that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%
 
@@
 }
-Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
+(\textbf{Now we know the following is not effective})
+\sout{Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
 That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
-%
+However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.}
+%
+\textbf{Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.}
 
@@
 %
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
+\sout{At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
+%
+The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.}
+
+\textbf{\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period.
+Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details).
+This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
+The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.}}
+%
+\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
+
+(\textbf{Is the following necessary? Could go in appendix. We already introduced provious applications of GKTL in the introduction of this section.})
+\textit{The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
 {Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
 The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
+Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.}
+
+The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
+\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.} 
+The observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
+
+\textbf{(The algorithm is already described above. I replaced the above (striked-out) by the following.)}
+\textit{\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period.
+Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details).
+This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
+The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
+}}
+
+%
+% Choice of parameters and perturbation
+
+\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
+The higher $k$, the larger is the bias involved in the statistical resampling, as illustrated in Fig.~\ref{fig:IS_GKTL}.}
+\textbf{There is however no general relation between this typical value and the value of $k$, and this value must be set empirically.}
+\EL{As for the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
+Finally, the cloning period $\tau$ determines how often the resampling is performed.
+The choice of the cloning period $\tau$ is important in practice.
+\textbf{Too short a cloning period} \sout{A cloning period too-short} can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
+As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$. \sout{In the following experiments, $\tau = \tau_c /2$.} (\textbf{Written below)}}
+
+\QQ{Thibault, pour la suite, nous sommes restés sur une version proche du papier initial, par simplicité.
+Si tu penses que certains points que tu avais ajoutés sont vraiment nécessaire pour répondre au referee, dis le nous.}
+
+(\textbf{Il faut mieux discuter le choix des parametres de l'algorithme}
+)
+\textbf{In practice, we started by fixing the computational cost associated with a single run of the cloning algorithm, \textit{i.e.} by fixing $N \times T_a$.
+Consistently with section~\ref{sec:time_avg}, we aim at sampling trajectories with a duration $T = 10\tau_c$.
+Consequently, we fixed $T_a = 15\tau_c$ to allow for an initial transient regime of the cloning algorithm.
+Then, we chose $N = 1024$ to obtain a computational cost of order $\mathcal{O}(10^4 \tau_c)$, \textit{i.e.} roughly one hundredth of the computational cost associated with the direct numerical simulation used for direct sampling in section~\ref{sec:direct_sampling}.
+Finally, we set $\tau = \tau_c / 2$.}
+\sout{The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.} (\textbf{Out-of-date values})
 
@@
 \end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period.
-Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details).
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling, as illustrated in Fig.~\ref{fig:IS_GKTL}.
-As for the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-Finally, the cloning period $\tau$ determines how often the resampling is performed.
-The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$. In the following experiments, $\tau = \tau_c /2$.}
-
-\QQ{Thibault, pour la suite, nous sommes restés sur une version proche du papier initial, par simplicité.
-Si tu penses que certains points que tu avais ajoutés sont vraiment nécessaire pour répondre au referee, dis le nous.}
-
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
+%
+
+(\textbf{The disscusion of figure~\ref{fig:IS_GKTL} must be updated (referees 2 \& 3})
+\sout{Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
+%
+Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}}
 \begin{equation}
@@
 \end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
+\sout{
+in addition to the unbiased \ac{pdf} based on the control run.}
+
+\textbf{Figure~\ref{fig:IS_GKTL} shows the histogram of the values of the average drag $F_{T_a} = \frac{1}{T_a}\int_{0}^{T_a}f_d(t)\mathrm{d}t$ for three ensemble of trajectories sampled with three different values of the cloning strength $k$.}
+It illustrates that the algorithm effectively samples preferentially trajectories with a higher value of the average.
+The higher $k$, the stronger the bias.
+\textbf{The grey histogram represents the average histogram that would be obtained using a direct approach -- \textit{i.e.} with $k=0$ -- for the same computational cost.
+We computed this histogram by approximating the \ac{pdf} of the drag averaged over $10\tau_c$ as a Gaussian \ac{pdf} (see figure~\ref{fig:PDF_AVG}).}
+
 %Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
@@
 
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-		\hline
-		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is  assessed in  
-Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\ZZ{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for extreme trajectories resampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
+\ZZ{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for extreme trajectories resampled by the \ac{gktl} algorithm.}
+
 \begin{figure}
@@
 
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
+(\textbf{Le paragraphe suivant n'est il pas important en reponse au premier commentaire du referee \#1 ??})
+
+\textbf{An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
+Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
+dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
+In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
+A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.}
 

### 2020-02-26 figures/AMS_drag_resampling/make_plot_256.py
Modify labels of PDFs

@@ -11,9 +11,9 @@ color_list = plt.rcParams['axes.prop_cycle'].by_key()['color']
 # transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/courbe_efficacite_instant
 # Original MATLAB script is make_plot_256.m
 
-prefix = '/home/tlestang/transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/' \
-        +'data/instant_256_20/'
-
+# prefix = '/home/tlestang/transition_hdd/thibault/These/AMS/libTAMS/tests/experience_lbm/' \
+#         +'data/instant_256_20/'
+prefix = "../../data/AMS_instant_256_20/"
 # Defines average and std of drag process, compute over the control run
 m = 0.0252
 sig = 0.0412
@@ -58,7 +58,7 @@ resamp.set_markeredgecolor(resamp.get_markerfacecolor())
 resamp.set_label('Maxima over resampled trajectories')
 
 ax.set_xscale('log')
-ax.set_xlabel(r'$C_{TAMS} / \tau_c$',fontsize=22)
+ax.set_xlabel(r'$C_{AMS} / \tau_c$',fontsize=22)
 ax.set_ylabel(r'$f_d / \sigma$',fontsize=22)
 ax.tick_params(axis="x", labelsize=18)
 ax.tick_params(axis="y", labelsize=18)

### 2020-02-26 figures/PDF_drag/PDF_drag.py

@@ -6,7 +6,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 
 
-fig, ax = plt.subplots()
+fig, ax = plt.subplots(constrained_layout=True)
 
 series = pd.read_csv("../../data/PDF_drag.csv", comment="#", index_col=0)
 series.plot(ax=ax, lw=2, legend=False)
@@ -23,8 +23,8 @@ plt.plot(
     color="k",
     linewidth=1.5,
 )
-plt.xlabel("$(f_d - \\mu)/\\sigma$", fontsize=18)
-plt.ylabel("$\\mathcal{P}$", fontsize=18)
+plt.xlabel("$x=(f_d - \\mu)/\\sigma$", fontsize=20)
+plt.ylabel("$P(x)$", fontsize=20)
 plt.legend(["With obstacle", "Without obstacle", "Gaussian estimate"], fontsize=18)
 plt.yscale("log")
 fname = splitext(__file__)[0] + ".png"

### 2020-02-26 figures/pdf_drag_with_perturbation/pdf_drag_with_perturbation.py

@@ -14,8 +14,8 @@ trunc_series.plot(ax=ax, lw=2, legend=False)
 series_perturb = pd.read_csv("../../data/PDF_drag_with_perturbation.csv", comment="#", index_col=0)
 series_perturb.plot(ax=ax, lw=2, legend=False)
 
-ax.set_xlabel("$(f_d - \\mu)/\\sigma$", fontsize=22)
-ax.set_ylabel("$\\mathcal{P}$", fontsize=22)
+ax.set_xlabel("$x=(f_d - \\mu)/\\sigma$", fontsize=22)
+ax.set_ylabel("$P(x)$", fontsize=22)
 ax.legend(["No perturbation", "$\\epsilon = 0.002$"], fontsize=18)
 ax.tick_params(axis='both', which='major', labelsize=18)
 

### 2020-02-27 jfm/draft_p2_jfm.tex
Iteration d'Emmanuel

--- 2020-02-26 jfm/draft_p2_jfm.tex
@@
 From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
+The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
 Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these events are rare.
+However, this requires very long simulations since these extreme events are rare.
+\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
+%
+%However, this requires very long simulations since these events are rare.
 % general def on rare-event sampling %
 %
-{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+%
+{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+%
+%	
+% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
 %
 In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
-%
-
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
+
 %
@@
 %
-In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
+Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
+It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
+%, and is designed to sample rare events of some observable with a higher frequency.
 %    
@@
 Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+More importantly, these types of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
+%
+%
+\EL{An original contribution of our study is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
 %
@@
 %
-Motivation for this study is twofold.
+The motivation for this study is twofold.
 Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
 This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them.
+Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
 This assessment is developed in the second part of the paper.
@@
 The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
-	drag fluctuations} is investigated based on a direct sampling approach.
+In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
 Both the instantaneous drag and time-averaged drag are considered.
-It is found that sampled extreme events for the instantaneous drag share very similar dynamics.
-Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
+It is found in particular that sampled extreme events for the instantaneous drag result from very similar dynamics.
+%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
 %\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
-In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology developed in the previous sections.
-Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
+Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
+In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
+This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
+Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
 This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
@@
 %
-The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
+The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{book_lbm}}. 
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
@@
 Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational manageability and efficiency.
+In our context, this numerical method has been chosen principally for its computational efficiency.
 
@@
 The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
+Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
 Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
@@
 	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.45\linewidth]{PDF_drag/PDF_drag.png}}
+		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
 	\subfloat[Autocorrelation of drag fluctuations]
@@
 Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.
+Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
 
@@
 This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma \equiv\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
+The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
+%
+According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
 Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
@@
 This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.
+In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
 
@@
 % In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.
-
-\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-\medskip
-\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
+The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
+%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
+%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
 To what extent they are representative of rare-event sampling algorithms?
@@
 Make the link between fluid mechanics and statistical algorithms ?}
-\medskip
-
-\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
-Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common.
-The algorithm tracks the value of the ratio of the probability of the new ensemble with the probability of the previous one, in order to be able to estimate the probability of the extreme event in the model.
-Different algorithms use different selection rules.
-The success of the algorithm depends on the quality of the selection rule in order to select the precursors of the extreme event.
-Optimal selection rules are called committor functions, see for instance~\citep{lestang_computing_2018} and references therein.
+%
+\EL{Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
+At each step of the algorithm, certain trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
+The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, which makes it possible to estimate the statistical bias and, therefore, to infer the statistics of extreme events.
+
+Different algorithms use different \emph{selection rules}. 
+The success of the algorithm essentially depends on the quality of the selection rule to select the precursors of the extreme event \citep{rolland_statistical_2015}. 
+However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
 In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved efficient for various dynamics and adopt opposite strategies:  the \acl{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \acl{gktl} uses as a selection rule based on the increment of a score function.} \ZZ{A complete description of these two algorithms and their operating details is given in appendixes.}
+They both proved to be efficient for various dynamics but adopt opposite strategies:  The \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating details are provided in the appendixes \ref{app:ams-short} and \ref{app:gktl_description}.}
 
@@
  More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\ZZ{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
+\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
 %
@@
 Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for \ZZ{estimating the probability of observables which are of time integrals over a very long time \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-
-\ZZ{The aim of this part of the work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which a completely new field of application.}
+Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
+%
+Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
+This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
+\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
 
@@
 %The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\ZZ{In the version of the \ac{ams} algorithm we use in this work, we consider %a sequence of events that can be sampled with a higher probability  
-trajectories with a fixed duration $T_a$~\citep{lestang_computing_2018}.%
+\EL{In the version of the \ac{ams} algorithm used in this work,  %a sequence of events that can be sampled with a higher probability  
+trajectories have a fixed duration $T_a$~\citep{lestang_computing_2018}.
 %The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
@@
 %
- At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
+ At each iteration, the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
 %The discarded trajectories are then re-sampled according to the remaining trajectories.
@@
 %The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}
+\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
 
@@
 
-The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
+The objective here is to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
 %\sout{In contrast with the \acl{ou} process, } 
@@
 %\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-We use the observable itself as the score function : $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Let us also stress that the Navier--Stokes dynamics is deterministic.
-\ZZ{Then, small random perturbation are artificially introduced at branching points. This procedure is detailed in Appendix \ref{app:perturb_branching_time}}.
+The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
+%
+Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
 % for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
@@
 	%
-Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov time.}
+The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval usually referred to as the Lyapunov timescale.}
 	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\ZZ{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistical properties of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-It was checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation.
+\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
+We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
 %\EL{Freddy: provide reference for Lyapunov timescale.
@@
 
-\EL{In addition to the score function, two important parameters are the number $N$ and the duration $T_a$ of the trajectories.
+\EL{In addition to the score function, two important parameters are the number, $N$, and the duration, $T_a$, of trajectories.
 The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with a practical limit given by the available computational resources.
-In this work, we have performed two numerical experiments with $N=32$ and $N=256$.} \ZZ{ The value of $T_a$ should chosen much larger than $\tau_c$~\citep{lestang_computing_2018}, but be kept small enough in order to minimize the computation cost.
-The duration of the trajectories was set to $T_a=5\tau_c$ in both cases.
-We have checked that note that large values of $T_a$ do not affect the result.}
+Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
+In this work, we have performed two numerical simulations with $N=32$ and $N=256$.} \EL{ Concerning $T_a$, it should be chosen much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
+In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
 
@@
 %
-\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-%
-The algorithm therefore fails to enhance the sampling of extreme events. 
+\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
+%
+Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
+%
+The algorithm thus fails to enhance the sampling of extreme events. 
 This result is also confirmed by increasing the number of initial trajectories to $N=256$.
@@
 the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observed that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
+We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
 %Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
 However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%
-
-It takes a certain time $\tau_L$ (called Lyapunov timescale) before a re-sampled trajectory separates from its parent.
+\EL{A phenomenological explanation is developed in the next paragraph.}
+%
+
+It takes a certain time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
 In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
 %
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is freed and carried away by the flow, and further large fluctuations of the drag can only result from the trapping of other vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} indicates that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
+\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ corresponding to the timescale over which a vortex remains trapped against the base of the obstacle.
+After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
+%
+Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
 %
@@
 Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, therefore, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the \emph{turbulence rate} of the flow.
+For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
+%
+%
+
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
 %
@@
 On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$, called turbulence rate, is much lower than one in our case of grid-generated turbulence, which implies that $\tau_L > \tau_c$.
+The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
 
@@
 %
-\EL{The phenomenology described before is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories.
-}
-Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
+\EL{This behaviour is independent of the choice of $N$ and $T_a$.
+Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories.}
+%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
 That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
+However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
 %
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
+	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
 	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=16384$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
+The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
 		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
@@
 %
-\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period.
-Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details).
+\EL{In brief, trajectories are first evolved in time  independently up to $t=\tau $ with $\tau < T_a$ referring to a cloning period.
+Then, the selection-and-cloning rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
 This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
+As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
+Further details are provided in the Appendix \ref{app:gktl_description}.
 }
@@
 \EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling, as illustrated in Fig.~\ref{fig:IS_GKTL}.
+The higher $k$, the larger is the bias involved in the statistical resampling.
+%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
 As for the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
@@
 A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$. In the following experiments, $\tau = \tau_c /2$.}
-
-\QQ{Thibault, pour la suite, nous sommes restés sur une version proche du papier initial, par simplicité.
-Si tu penses que certains points que tu avais ajoutés sont vraiment nécessaire pour répondre au referee, dis le nous.}
-
-The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-We have taken $N=16384$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps to allow for efficient importance sampling.
+As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.}
+
+\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
+We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
+%
+%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
+%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
+%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
+%
 %The choice of the cloning period $\tau$ is important in practice.
@@
 %
-Three numerical experiments corresponding to three different values of the parameter $k$ have been carried out.
+Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
 %
@@
 %
-Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-\begin{equation}
-\label{eq:estimate_biased_measure}
-\rho_k(F) 
-%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-\end{equation}
-in addition to the unbiased \ac{pdf} based on the control run.
-As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.
+%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
+%\begin{equation}
+%\label{eq:estimate_biased_measure}
+%\rho_k(F) 
+%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
+%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
+%\end{equation}
+%in addition to the unbiased \ac{pdf} based on the control run.
+%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
+%%
+\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the biased \ac{pdf}s of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased \ac{pdf} based on the control run (with the same computational cost).
+As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
+%
 %
@@
 %
-Nevertheless, let us mention that for a fixed number of trajectories, there is necessarily an upper limit $k_{max}$ over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
+Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
 %
 For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ is artificially peaked.
-%
-Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$.
-This issue is addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ becomes artificially peaked.
+%
+%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
+%
+%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
+%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
 %
 % how to assess the value of k_max ?
-In the present work, the order of magnitude of $k_{max}$ has been estimated \emph{empirically} by evaluating the diversity of the trajectories in the biased ensemble.
-This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-\begin{table}
-	\centering
-	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline
-		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-		\hline
-		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-		\hline
-		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-		\hline
-		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-		\hline
-		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-		\hline
-	\end{tabular}
-	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-		Note that trajectories that overlap for more than half of their duration are counted as only one.
-		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-	\label{tbl:nbTraj_tbl}
-\end{table}
-
-% Comparisons with direct sampling
-%
-The efficiency of the \ac{gktl} algorithm is  assessed in  
-Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-Note that trajectories that overlap for more than half of their duration are counted as one.
-%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%
-%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%
-%\begin{figure}
+\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
+%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
+
+%\begin{table}
 %	\centering
-%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%  \caption{PDF of average drag over $10\tau_c$}
-%  \label{fig:PDF_AVG}
-%\end{figure}
-%
-% Conclusion
-%
-We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
+%	\begin{tabular}{|l|c|c|c|c|c|}
+%		\hline
+%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+%		\hline
+%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
+%		\hline
+%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
+%		\hline
+%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
+%		\hline
+%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
+%		\hline
+%	\end{tabular}
+%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
+%		Note that trajectories that overlap for more than half of their duration are counted as only one.
+%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
+%	\label{tbl:nbTraj_tbl}
+%\end{table}
+%
+%% Comparisons with direct sampling
+%%
+%The efficiency of the \ac{gktl} algorithm is  assessed in  
+%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
+%Note that trajectories that overlap for more than half of their duration are counted as one.
+%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
+%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
+%%
+%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
+%%
+%%\begin{figure}
+%%  \centering
+%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
+%%  \caption{PDF of average drag over $10\tau_c$}
+%%  \label{fig:PDF_AVG}
+%%\end{figure}
+%%
+%% Conclusion
+%%
+%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
 
@@
 
-\ZZ{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for extreme trajectories resampled by the \ac{gktl} algorithm.
+\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
 %It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
@@
 	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the average drag in the ensemble of resampled trajectories.
+	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
 The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
@@
 \end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field at the time of maximum drag over a subset of trajectories
-sampled by the cloning algorithm.
-It shows that the sampled events are consistent with the picture of strong vorticity levels being concentrated in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Moreover, we stress that the computational cost for the sampling of the events shown in figures~\ref{fig:timeseries_extrms_AVG_GKTL} and~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly a hundredth of computational cost of the direct sampling required to sample events shown in figures~\ref{fig:extreme_avg} and~\ref{fig:top_4_events_vorticity}.
+In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
+One observes that extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
+Nevertheless, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
 
@@
 	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field corresponding the maximum of the drag $f_d$ over trajectories sampled by the cloning algorithm.}
+	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
 	\label{fig:illustr_extrms_vorticity_GKTL}
@@
 
-Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-events unreachable by a direct approach.
+\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
+In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
+Analytical details about the estimate of return times are provided in Appendix \ref{sec:return-time-gktl}.}
+
+%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
+% events unreachable by a direct approach.
 %(see Fig.~\ref{fig:illustrate_return_time})
 
-In the following, the \ac{gktl} algorithm is used with $N=1024$ trajectories of duration $T_a > T= 10\tau_c$.
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T(t) = \int_{t-T}^{t}f_d(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\cite{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k \int_{0}^{T_a}f_d(t)dt}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag acting on the square obstacle.
-The figure shows two independent estimates, obtained using different values of the bias parameter $k$.
-Note that both estimates have been computed with the same computational cost $T_{tot}=N\times T_a$.
-In addition, \EL{Fig.~\ref{fig:return_times_gktl}} shows an estimate computed through direct sampling with the same computational cost, \textit{i.e.} using a timeseries of duration $T_{tot}$.
-Whilst a direct approach cannot access events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us for the computation of statistics for drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
+%\begin{equation}
+%\label{eq:time_averaged}
+%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
+%\end{equation}
+%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
+%\begin{equation}
+%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
+%\end{equation}
+%%
+%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
+%\begin{equation}
+%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
+%\end{equation}
+%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
+
+%
+
+Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
+%
+Both estimates have been obtained with the same computational cost $N\times T_a$.
+An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$, is also shown. 
+Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
 This is  obviously a major advantage of this rare-event sampling algorithm.
+%
+%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
 
@@
 $\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$ and $T_a=30\tau_c$.
+The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
 The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
@@
 
-In a first part, we investigated the dynamics and statistics of extreme fluctuations of the drag acting on a square mounted in a turbulent channel flow, in two dimensions.
+The dynamics and statistics of extreme fluctuations of the drag acting on a square obstacle placed in a turbulent two-dimensional channel flow are investigated numerically. 
 By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
@@
 The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
-Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-This property is linked to the phenomenology of extremes of the time-averaged drag.
-Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
+%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
+%This property is linked to the phenomenology of extremes of the time-averaged drag.
+%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
+Concerning the time-averaged drag, extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
 %Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
 
-The application of rare event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of the score function which has to be well suited to the phenomenology of the extreme dynamics.
-For complex dynamics including turbulent flows, this choice is  difficult because of the lack of characterisation of rare events.
-These latter are expected to depend on the dynamical regime (Reynolds number), the stirring mechanism and boundary conditions of the flow.
-In this context, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-
-On the basis of the same two-dimensional test flow, we then applied the \ac{ams} algorithm choosing the drag itself as a score function.
+The second part is dedicated to the application of two rare-event algorithms, namely the \ac{ams} and \ac{gktl} algorithms, to possibly accelerate the sampling of extreme fluctuations of the drag. 
+%While such algorithms have already proved to be efficient for a wide range of dynamical systems, its applicability in the context of turbulent flows remains open.
+
+Importantly, rare-event algorithms relies on the definition of a score function for the selection of trajectories.
+The efficiency of the algorithm depends on the choice of this score function, which is expected to select the precursors of the extreme events.
+%For complex dynamics such as turbulent flows, this choice is difficult because of the lack of characterisation of rare events. 
+%These latter are expected to depend on the dynamical regime, the stirring mechanism and boundary conditions of the flow. 
+In the present case, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
+%
+Therefore, we applied the \ac{ams} algorithm choosing the drag itself as a score function.
 In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
@@
 
-\newtheorem{theo}{Theorem}
-\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-\label{app:ams}
-
-\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
+%
+%
+%\newtheorem{theo}{Theorem}
+%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
+%\label{app:ams}
+%
+%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
 Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
 Les referees n'ont pas demandé de détailler les algos.}
-
-This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-mathematical results and discusses its use in practice.
-
-\subsection{Description of the algorithm}
-An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%
-From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-
-To start off with, each of these trajectories is associated a weight $w_0=1$.
-Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-\begin{equation}
-\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-\end{equation}
-Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-\begin{equation}
-\label{eq:ams_threshold_def}
-\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-\end{equation}
-One might expect intuitively that $\ell_j=1$.
-However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{Illustration of one iteration the \ac{ams} algorithm.
+%
+%
+%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
+%mathematical results and discusses its use in practice.
+%
+%\subsection{Description of the algorithm}
+%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
+%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
+%%
+%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
+%
+%To start off with, each of these trajectories is associated a weight $w_0=1$.
+%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
+%\begin{equation}
+%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
+%\end{equation}
+%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
+%\begin{equation}
+%\label{eq:ams_threshold_def}
+%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
+%\end{equation}
+%One might expect intuitively that $\ell_j=1$.
+%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
+%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
+%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
+%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
+%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
+%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
+%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
+%	\caption{Illustration of one iteration the \ac{ams} algorithm.
 Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
@@
 Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-	\label{fig:AMS_schema}
-\end{figure}
-
-After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-$\tilde{J} = \sum_{j=1}^J \ell_j$.
-Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-\subsection{How to compute the probability}
-Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-\begin{equation}
-w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-\end{equation}
-This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-\subsection{Summary of mathematical results}
-We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-of iterations $N$.
-We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-\begin{equation}
-\label{eq:estimator_ams}
-q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-\end{equation}
-In the following we denote this estimator $\hat{q}_N(a)$.
-One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-\begin{theo}
-	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-	\begin{equation}
-	\mathbb{E}[\hat{q}_N]=q.
-	\end{equation}
-\end{theo}
-Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-
-More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-\begin{equation}
-\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-\end{equation}
-with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-The minimal variance $-q^2\ln q$ is obtained when choosing
-\begin{equation}
-\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-\label{eq:time_dependent_committor}
-\end{equation}
-for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-
-\subsection{How to use it in practice}
-In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-
-Another parameter of the algorithm is the duration $T_a$.
-This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-This can be explained as follows.
-Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-This amounts to direct sampling.
-
-Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-\begin{equation}
-\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-\end{equation}
-and the iterations stop.
-This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-
-\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-\label{app:AMS_on_OU}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+%	\label{fig:AMS_schema}
+%\end{figure}
+%
+%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
+%$\tilde{J} = \sum_{j=1}^J \ell_j$.
+%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
+%\subsection{How to compute the probability}
+%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
+%\begin{equation}
+%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
+%\end{equation}
+%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
+%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
+%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
+%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
+%\subsection{Summary of mathematical results}
+%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
+%of iterations $N$.
+%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
+%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
+%\begin{equation}
+%\label{eq:estimator_ams}
+%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
+%\end{equation}
+%In the following we denote this estimator $\hat{q}_N(a)$.
+%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
+%\begin{theo}
+%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
+%	\begin{equation}
+%	\mathbb{E}[\hat{q}_N]=q.
+%	\end{equation}
+%\end{theo}
+%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
+%
+%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
+%\begin{equation}
+%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
+%\end{equation}
+%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
+%The minimal variance $-q^2\ln q$ is obtained when choosing
+%\begin{equation}
+%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
+%\label{eq:time_dependent_committor}
+%\end{equation}
+%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
+%
+%
+%\subsection{How to use it in practice}
+%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
+%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
+%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
+%
+%Another parameter of the algorithm is the duration $T_a$.
+%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
+%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
+%This can be explained as follows.
+%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
+%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
+%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
+%This amounts to direct sampling.
+%
+%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
+%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
+%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
+%\begin{equation}
+%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
+%\end{equation}
+%and the iterations stop.
+%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
+%
+%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
+%\label{app:AMS_on_OU}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
 The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
 The blue line is the analytical solution for the return time of amplitude $a$.}
-	\label{fig:comparaison_temps_de_retour}
-\end{figure}
-Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-\begin{equation}
-\label{eq:ou}
-\dot{x} = -x + \eta (t),
-\end{equation}
-where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-
-The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-In the latter, the typical computational cost is simply the return time $r(a)$.
-One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
+%	\label{fig:comparaison_temps_de_retour}
+%\end{figure}
+%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
+%\begin{equation}
+%\label{eq:ou}
+%\dot{x} = -x + \eta (t),
+%\end{equation}
+%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
+%This basic stochastic process will allow us to highlight differences with fluid dynamics.
+%
+%
+%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
+%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
+%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
+%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
+%
+%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
+%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
+%In the latter, the typical computational cost is simply the return time $r(a)$.
+%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
+%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
+%
+%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
+%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
+%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
 Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-\newpage
-
-\section{The \ac{ams} algorithm -- version initiale}
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%\newpage
+
+\section{The \ac{ams} algorithm}
 \label{app:ams-short}
@@
 
-\newpage
-
-\QQ{Appendix sur GKTL à valider : Freddy ?}
-
-\section{The \acl{gktl} algorithm -- version initiale}
+s
+
+\section{The \acl{gktl} algorithm}
 \label{app:gktl_description}
+
+\subsection{The operating principle}
+\label{sec:gktl-operating-principle}
+
 The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
@@
 
+\subsection{Computation of return times}
+\label{sec:return-time-gktl}
+
+Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
+\begin{equation}
+\label{eq:time_averaged}
+F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
+\end{equation}
+and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
+\begin{equation}
+r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
+\end{equation}
+%
+The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
+\begin{equation}
+\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
+\end{equation}
+with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
+
 \newpage

### 2020-02-27 figures/ecoulement_typique/plot_ecoulement_typique.py
Revert "Update units plot_ecoulement_typique"
This reverts commit 782227546de6fcdbc8b73ffe7c03f772d32bf8fe.

Fix units of time

@@ -6,16 +6,13 @@ import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
 
-T0 = 500 # Turn-over time in lattice units
-
 Dx = 513
 Dy = 129
 N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
 
-prefix = '/home/tlestang/transition_hdd/thibault/These/' \
-         + 'articles/data/ecoulement_typique/'
+prefix = '../../data/ecoulement_typique/'
 
 fig = plt.figure(figsize=(16,7), constrained_layout=True)
 # Declares a 2x5 grid.
@@ -57,16 +54,14 @@ for i in range(0,nplots-1):
         # The sign of the vorticity is wrong if using a right-handed coord. system.
         # The velocity field should probably be flipped upside-down
         vort = - vort
-        # Express vorticity in unit of T0 and R
-        vort = vort*T0
 
         # Print time as title
-        timeFromBeg=0.4+i*0.8
-        tit='$t={:2.1f}T_0$'.format(timeFromBeg)
+        timeFromBeg=0.1+i*0.2
+        tit='$t={:2.1f}\\tau_c$'.format(timeFromBeg)
 
         ax_list[i].set_title(tit,fontsize=20,bbox=bbox)
         # Print vorticity field
-        im=ax_list[i].imshow(vort, interpolation='spline36', cmap='BrBG', vmin=-35, vmax=35)
+        im=ax_list[i].imshow(vort, interpolation='spline36', cmap='BrBG', vmin=-0.09, vmax=0.09)
         # Adds a filled Rectangle to mark the square obstacle
         rect = ax_list[i].add_artist(Rectangle((21,25), 14, 14))
         rect.set_color('#545454')
@@ -98,7 +93,7 @@ ax_list[-1].plot(tspan,fd)
 for t in range(0,2000,400):
         ax_list[-1].plot(tspan[t+200],fd[t+200],marker='o',markersize=10,linestyle='None',color='k')
 ax_list[-1].set_xlim((0,4))
-#ax_list[-1].set_xticklabels(['$0$','$0.2\\tau_c$','$0.4\\tau_c$','$0.6\\tau_c$','$0.8\\tau_c$','$1.0\\tau_c$'], fontsize=22)
[…]

### 2020-02-27 jfm/draft_p2_jfm.tex
Relecture apres Emmanuel

@@
 \begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
+	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
 	%
@@
 	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-	%
-
-		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
+	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
+	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
+These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
+        The present study is the first application to a fluid-structure interaction problem. 
+	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
+	%
+	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
+	% 
+	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
+%
+%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
+%	%
+%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
+%	%
+%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
+This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
+%	%
+%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
+%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
+%	%
+%	
+%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
+%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
 Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-
-	%
-	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-	%
-	\QQ{phrase de conclusion ?}
-}
+%
+%	%
+%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
+%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
+%	%
+%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
+%	% 
+%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
+%	%
+%	\QQ{phrase de conclusion ?}
+%
 %\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
@@
 %
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
 
@@
 %
-Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, these types of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{An original contribution of our study is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
+Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
+More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
+%
+%
+\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+%
+Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
 %
@@
 % Annonce du plan
-In more details,  the flow set-up is introduced and the dynamics related to typical
+The flow set-up is introduced and the dynamics related to typical
 drag fluctuations is described in section~\ref{sec:test_flow}.
@@
 Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme events for the instantaneous drag result from very similar dynamics.
+It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
 %Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
@@
 %
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{book_lbm}}. 
+The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
@@
 The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel.
+The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
 The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
@@
 Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.
+The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
+For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
 
@@
 This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
+It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
 The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
@@
 	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
+		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
 	\subfloat[Autocorrelation of drag fluctuations]
@@
 %Note that the amplitudes have not been normalized.
-		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
 		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
@@
   \label{eq:return_time}
-\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
+  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
 \end{equation}
@@
 %
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
-Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified.
+According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
+Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
 Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
@@
 }
-
+%
 % relative contribution of forebody and base pressure
@@
 $\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from the drop of the base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
+	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
 On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
@@
 %Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
+Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
 %This is reminiscent of time-irreversibility in turbulent dynamics. 
@@
 		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
+%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
 	}
@@
 %\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation of the instantaneous drag.
+\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
 In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
@@
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected.
+Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
 \EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that around one hundred events are selected.}
+As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
 
@@
 In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
+Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
+As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
 %% objective
 % In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
+\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
 %\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
@@
 %
-\EL{Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, certain trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, which makes it possible to estimate the statistical bias and, therefore, to infer the statistics of extreme events.
+Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
+At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
+The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
 
 Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to select the precursors of the extreme event \citep{rolland_statistical_2015}. 
+The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
 However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
 In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies:  The \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating details are provided in the appendixes \ref{app:ams-short} and \ref{app:gktl_description}.}
+They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
 
@@
 During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
+Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
 Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
@@
 %The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{In the version of the \ac{ams} algorithm used in this work,  %a sequence of events that can be sampled with a higher probability  
-trajectories have a fixed duration $T_a$~\citep{lestang_computing_2018}.
+\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
+trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
 %The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
@@
 %
- At each iteration, the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
+At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
 %The discarded trajectories are then re-sampled according to the remaining trajectories.
@@
 
-The objective here is to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
+The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
 %\sout{In contrast with the \acl{ou} process, } 
@@
 	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval usually referred to as the Lyapunov timescale.}
+The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
 	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
@@
 
-\EL{In addition to the score function, two important parameters are the number, $N$, and the duration, $T_a$, of trajectories.
+\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
   The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
 Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.} \EL{ Concerning $T_a$, it should be chosen much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
+In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
+\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
 In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
@@
 	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.
-\QQ{Change TAMS to AMS in horizontal axis}}
+The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
 \end{figure}
@@
 
-It takes a certain time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
+It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
 In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
 %
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ corresponding to the timescale over which a vortex remains trapped against the base of the obstacle.
+\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
 After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
@@
 \EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories.}
+Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
 %Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
@@
 %
-The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
+The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
 {Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
+The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
 Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
@@
 The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.}
+\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
 %Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.    
+Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
 %Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
 %
-\EL{In brief, trajectories are first evolved in time  independently up to $t=\tau $ with $\tau < T_a$ referring to a cloning period.
-Then, the selection-and-cloning rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
+\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
+Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
 This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
 As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in the Appendix \ref{app:gktl_description}.
+Further details are provided in Appendix~\ref{app:gktl_description}.
 }
@@
 %, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-As for the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-Finally, the cloning period $\tau$ determines how often the resampling is performed.
-The choice of the cloning period $\tau$ is important in practice.
-A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps to allow for efficient importance sampling.
-As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.}
+Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
+The cloning period $\tau$ determines how often the resampling is performed.
+A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
+On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
+As a result, a rule of thumb is $\tau \approx \tau_c$.}
 
@@
 %%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the biased \ac{pdf}s of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased \ac{pdf} based on the control run (with the same computational cost).
+\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
 As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
@@
 For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the biased \ac{pdf} corresponding to $k=0.03$ becomes artificially peaked.
+This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
 %
@@
 % how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
+\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
 %This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
@@
 In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Nevertheless, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
+One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
+Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
+However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
 
@@
 In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-Analytical details about the estimate of return times are provided in Appendix \ref{sec:return-time-gktl}.}
+The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
 
@@
 Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$, is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
+An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
+Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
 Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
@@
 
-The dynamics and statistics of extreme fluctuations of the drag acting on a square obstacle placed in a turbulent two-dimensional channel flow are investigated numerically. 
-By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
-Extreme drag amplitudes do not persist over time since the mean flow eventually sweeps away the fluid structures responsible for this situation.
-%
-The lifetime of extreme drag fluctuations is therefore of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
+%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
+%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
+%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
+Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
+%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
+%%This property is linked to the phenomenology of extremes of the time-averaged drag.
+%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
+%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
+%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
+%
+%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
+These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
+%%
+%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
+%%
+%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
+Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
+%%
+%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
+However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
+The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
+%
+%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
+%%
+%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
+%Here, the observable itself was chosen as score function, as a safe fallback solution.
+%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
+%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
+
+\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
+By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
+This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
+Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
 %Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
@@
 %Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-Concerning the time-averaged drag, extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
+In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
+% but can be either formed by one of these two cases or by a combination of both.
 %Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
 
-The second part is dedicated to the application of two rare-event algorithms, namely the \ac{ams} and \ac{gktl} algorithms, to possibly accelerate the sampling of extreme fluctuations of the drag. 
-%While such algorithms have already proved to be efficient for a wide range of dynamical systems, its applicability in the context of turbulent flows remains open.
-
-Importantly, rare-event algorithms relies on the definition of a score function for the selection of trajectories.
-The efficiency of the algorithm depends on the choice of this score function, which is expected to select the precursors of the extreme events.
-%For complex dynamics such as turbulent flows, this choice is difficult because of the lack of characterisation of rare events. 
-%These latter are expected to depend on the dynamical regime, the stirring mechanism and boundary conditions of the flow. 
-In the present case, the choice of the observable itself as score function is certainly a safe fall-back option, nevertheless not optimal.
-%
-Therefore, we applied the \ac{ams} algorithm choosing the drag itself as a score function.
-In this case, our results illustrate that the selection-mutation procedure is unable to generate rare trajectories at a better rate than a direct sampling.
-This can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which re-sampled trajectories separate from their parent. 
-%following the addition of a small perturbation in the initial conditions.
-The \ac{gktl} algorithm has been applied to the sampling of trajectories displaying extreme fluctuations of the time-averaged drag.
-In this case, we showed that using \ac{gktl} leads to a significant gain with respect to a direct approach, allowing for the simulation and computation of statistics for (very) rare trajectories.
-The sampling of extreme time-averages is aided by the selection of trajectories displaying an exceptional succession of drag fluctuations, resulting in an extreme value of the average.
-It is however unclear whether the \ac{gktl} is suited to sample trajectories for which the extreme fluctuations of the time-averaged drag results from a unique, exceptionally large drag fluctuation.
-This issue is postponed for future investigations.
-
-Importantly, successful usage of the \ac{ams}, \ac{gktl} or similar algorithms for complex flows relevant to industrial or environmental situations will require coping with the fact that optimal score functions
-are difficult to identify, if even possible.
-A promising direction explored in current research is to take advantage of recent advances in \emph{learning methods} to optimise score functions beforehand.
+A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
+These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
+%
+On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
+This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
+Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
+%
+On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
+The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
+The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
+This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
+
+%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
+%
+Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
+In this study, the observable itself (the drag) has been chosen as the score function.
+Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
+Our study shows this quite clearly.
+A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
 
@@
 % details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \cite{book_lbm} for a comprehensive introduction)}.
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
@@
 		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{\QQ{Change TAMS TO AMS} Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
 The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
@@
 
-s
-
 \section{The \acl{gktl} algorithm}
@@
 
-\newpage
-
 \EL{
@@
 %
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.  \QQ{is it correct?}.
+In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
 %Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
@@
 %
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ have been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
+In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
 %
@@
 \includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.} \QQ{indicate $x=(f_d-\mu)/\sigma$ and ${\cal P}(x)$  as before}}
+\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
 \label{fig:pdf_drag_with_perturbation}

### 2020-02-28 figures/PDF_AVG/plot_pdf_AVG.py
Update plot for PDF for average drag

@@ -8,9 +8,9 @@ def ret_gaussian(x):
     # Returns zero-mean, unit variance at point
     return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
 
-musigma = np.fromfile("musigma_AVG_10.dat", float, -1, "");
+musigma = np.fromfile("../../data/musigma_AVG_10.dat", float, -1, "");
 
-buf = np.fromfile("pdfData_AVG.dat", float, -1, "");
+buf = np.fromfile("../../data/pdfData_AVG.dat", float, -1, "");
 n=int(len(buf)/2)
 f = buf[0:n]*musigma[1]
 xi = (buf[n:]-musigma[0])/musigma[1]
@@ -26,10 +26,9 @@ p = np.poly1d(c) # p is the polynomial and can be evaluated
 fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
 # plots the PDF estimate
 line, = ax.plot(xi,f)
-line.set_label('PDF of $F_T$')
 
 # Plots the gaussian fit based on mu and sigma
-lineGauss, = ax.plot(xi,ret_gaussian(xi), label='Gaussian PDF')
+lineGauss, = ax.plot(xi,ret_gaussian(xi), label='Gaussian estimate')
 
 # Plots the local least-square fit for the positive tail
 #lineFit, = ax.plot(xiTrunc,np.exp(p(xiTrunc)), linewidth=2, \
@@ -40,8 +39,8 @@ ax.set_yscale('log')
 ax.set_xticks([-5,-3,-1,0,1,3,5])
 ax.set_xticklabels(['$-5\sigma_T$','$-3\sigma_T$','$-\sigma_T$',
                     '$0$','$\sigma_T$','$3\sigma_T$','$5\sigma_T$'], fontsize=16)
-ax.set_xlabel('$F_T$',fontsize=22)
-plt.ylabel('$\log(PDF)$',fontsize=22)
+ax.set_xlabel('$x=F_T$',fontsize=22)
+plt.ylabel('$P(x)$',fontsize=22)
 ax.legend(loc='best',fontsize=22)
 
 # Prints figure

### 2020-02-28 jfm/draft_p2_jfm.tex
Add references suggested by referees

@@
 %
-
+\EL{Another problem related to rare events is the estimation of parameters or statistics from a limited number of samples.
+Several authors have proposed original strategies  \citep{Mohamad-ws,Blonigan-ws} such as sequential sampling.
+These approaches and rare event algorithms could also be considered for applications in wave-structure problems, ship motion and load acting on offshore platforms \citep{belenky-ws}.}
 %
@@
 This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
+\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is an unexplored area of application.}
 

### 2020-02-28 jfm/draft_p2_jfm.tex
Add link to github repo

@@
 \usepackage{subfig}
+\usepackage{hyperref}
 %\usepackage{caption}
@@
 	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
+      This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
 %
@@
 The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
+For more details about the implementation of the flow, see the software repository attached to this paper (\url{https://github.com/tlestang/paper_extreme_drag_fluctuations}).
 

### 2020-04-03 figures/PDF_drag/PDF_drag.py
fix paths and relocate data for fig 1-6

@@ -1,17 +1,20 @@
 import sys
 from os.path import splitext
-
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
-
+from os.path import abspath, dirname, join, basename, splitext
 
 fig, ax = plt.subplots(constrained_layout=True)
 
-series = pd.read_csv("../../data/PDF_drag.csv", comment="#", index_col=0)
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
+series = pd.read_csv(join(prefix, "PDF_drag.csv"), comment="#", index_col=0)
 series.plot(ax=ax, lw=2, legend=False)
 
-series = pd.read_csv("../../data/PDF_drag_nosq.csv", comment="#", index_col=0)
+series = pd.read_csv(join(prefix, "PDF_drag_nosq.csv"), comment="#", index_col=0)
 series.plot(ax=ax, lw=2, legend=False)
 
 # Plot gaussian distribution
@@ -27,7 +30,9 @@ plt.xlabel("$x=(f_d - \\mu)/\\sigma$", fontsize=20)
 plt.ylabel("$P(x)$", fontsize=20)
 plt.legend(["With obstacle", "Without obstacle", "Gaussian estimate"], fontsize=18)
 plt.yscale("log")
-fname = splitext(__file__)[0] + ".eps"
-plt.savefig(fname)
 
-plt.show()
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
+
+plt.savefig(fname)

### 2020-04-03 figures/autocorrelation_drag/autocorrelation_drag.py

@@ -4,13 +4,18 @@ import pandas as pd
 import sys
 from  scipy import integrate
 from os.path import splitext
+from os.path import abspath, dirname, join, basename
 
 R = 16 # Size of the obstable in lattice unit
 U = 0.036 # Mean flow velocity in lattice unit 
 dt = U/R
 
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
 # -------* Autocorrelation function for drag process *---------
-path_autocorr_drag = '../../data/autocorr_drag.csv'
+path_autocorr_drag = join(prefix, 'autocorr_drag.csv')
 try:
     series = pd.read_csv(path_autocorr_drag, comment='#', index_col=0)
 except FileNotFoundError:
@@ -30,7 +35,7 @@ plt.plot(np.arange(0,C.size)*dt, C,lw=2)
 plt.xlabel('$\\tau / T_0$', fontsize=18)
 plt.ylabel('$C(\\tau)$', fontsize=18)
 
-fname = splitext(__file__)[0]+'.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0])+'.eps'
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/ecoulement_typique/ecoulement_typique.py

@@ -5,6 +5,7 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join
 
 Dx = 513
 Dy = 129
@@ -12,8 +13,9 @@ N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
 
-prefix = "./data/"
-print(prefix)
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 fig = plt.figure(figsize=(16,7), constrained_layout=True)
 # Declares a 2x5 grid.
@@ -28,8 +30,8 @@ plt.subplot(grid[1,:])
 ax_list = fig.axes
 nplots = len(ax_list)
 
-path_to_file_ux = prefix+'ux.dat'
-path_to_file_uy = prefix+'uy.dat'
+path_to_file_ux = join(prefix, 'ux.dat')
+path_to_file_uy = join(prefix, "uy.dat")
 fx = open(path_to_file_ux, "rb")  # reopen the file
 fy = open(path_to_file_uy, "rb")  # reopen the file
 bbox = dict(boxstyle="round", fc="0.9")
@@ -87,7 +89,7 @@ cbar.ax.tick_params(labelsize=16)
 sig=0.0412
 m=0.0252
 Nt = 2000;
-fd = (np.fromfile(prefix+'data_force.datout',float,Nt,"")-m)/sig
+fd = (np.fromfile(join(prefix,"data_force.datout"),float,Nt,"")-m)/sig
 tspan = np.linspace(0,4,Nt)
 ax_list[-1].plot(tspan,fd)
 # And plots nplots-1 dots for the nplots-1 snapshots
@@ -102,8 +104,7 @@ ax_list[-1].set_ylabel('$f_d(t)$',fontsize=22)
 ax_list[-1].set_xlabel('$t/T_0$',fontsize=22)
 
 
-fname = 'ecoulement_typique.eps'
-plt.savefig(fname)
-
-plt.show()
-            
+fname = join(
+    abspath(dirname(__file__)), "ecoulement_typique.eps"
+    )
+plt.savefig(fname)            

### 2020-04-03 figures/illustr_ecoulement/illustr_ecoulement.py

@@ -2,6 +2,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
 
 Dx = 513
 Dy = 129
@@ -72,8 +73,8 @@ ax.set_ylabel(r'$y/R$',fontsize=24)
 #cbar=fig.colorbar(im, ax=ax,location='top')
 #cbar.ax.tick_params(labelsize=16)
 
-fname = 'illustr_ecoulement.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()
             

### 2020-04-03 figures/illustrate_return_time/illustrate_return_time.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join
 
 def bundle_nearest_points(I, dist):
     """
@@ -86,8 +87,8 @@ plt.ylabel('$f_d(t)$',fontsize=22)
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)
 
+fname = join(
+    abspath(dirname(__file__)), "illustrate_return_time.eps"
+    )
 
-fname= 'illustrate_return_time.eps'
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/pressure_ratio/pressure_ratio.py

@@ -3,8 +3,13 @@ import os
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 meanPf = 5.68;
 meanPr = 5.65;
@@ -13,13 +18,13 @@ M = (meanPf + meanPr)/2;
 sig = 0.0412;
 mu = 0.0252;
 
-path_csv = "../../utilities/peaks.csv"
+path_csv = join(prefix, "peaks.csv")
 csv_file_handle = open(path_csv, "r")
 
 flucts_reader = csv.reader(csv_file_handle, delimiter=",")
-flucts = [_tuple for i, _tuple in enumerate(flucts_reader) if os.path.isfile("../../data/event_{}/dragFile.dat".format(i))]
+flucts = [_tuple for i, _tuple in enumerate(flucts_reader) if os.path.isfile(join(prefix, "event_{}/dragFile.dat".format(i)))]
 csv_file_handle.seek(0)
-event_dirs = ["../../data/event_{}".format(i) for i, _tuple in enumerate(flucts_reader) if os.path.isfile("../../data/event_{}/dragFile.dat".format(i))]
+event_dirs = [join(prefix, "event_{}".format(i)) for i, _tuple in enumerate(flucts_reader) if os.path.isfile(join(prefix, "event_{}/dragFile.dat".format(i)))]
 csv_file_handle.close()
 NbExtremeEvents = 104
 prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
@@ -29,7 +34,7 @@ maxDrag = []
 frontContrib = []
 rearContrib = []
 for counter, fluct in enumerate(flucts):
-    dirname, init_file, nb_timesteps, peak, value = fluct
+    dir_name, init_file, nb_timesteps, peak, value = fluct
     drag = np.fromfile(os.path.join(event_dirs[counter],'dragFile.dat'), float, -1, "")
     pFront = np.fromfile(os.path.join(event_dirs[counter],'pFront.dat'), float, -1, "")
     pRear = np.fromfile(os.path.join(event_dirs[counter],'pRear.dat'), float, -1, "")
[…]

### 2020-04-03 figures/return_time/return_time.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join
 
 plt.style.use('ggplot')
 
@@ -31,7 +32,8 @@ ax.set_yticklabels((r'$3\sigma$',r'$5\sigma$',r'$7\sigma$',r'$9\sigma$'))
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)
 
-fname = 'return_time.eps'
-plt.savefig(fname)
+fname = join(
+    abspath(dirname(__file__)), splitext(__file__)[0] + ".eps"
+    )
 
-plt.show()
+plt.savefig(fname)

### 2020-04-03 figures/PDF_AVG/PDF_AVG.py
fix paths and relocate data for fig 7-13

@@ -1,16 +1,21 @@
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 def ret_gaussian(x):
     # Returns zero-mean, unit variance at point
     return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
 
-musigma = np.fromfile("../../data/musigma_AVG_10.dat", float, -1, "");
+musigma = np.fromfile(join(prefix, "musigma_AVG_10.dat"), float, -1, "");
 
-buf = np.fromfile("../../data/pdfData_AVG.dat", float, -1, "");
+buf = np.fromfile(join(prefix, "pdfData_AVG.dat"), float, -1, "");
 n=int(len(buf)/2)
 f = buf[0:n]*musigma[1]
 xi = (buf[n:]-musigma[0])/musigma[1]
@@ -44,7 +49,7 @@ plt.ylabel('$P(x)$',fontsize=22)
 ax.legend(loc='best',fontsize=22)
 
 # Prints figure
-fname = 'PDF_AVG.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/dynamics_extremes/dynamics_extremes.py

@@ -5,13 +5,15 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
 
 Dx = 513
 Dy = 129
 N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
-prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
+
+prefix = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/"
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is lowest fluctuation
@@ -23,29 +25,29 @@ with open(prefix+'extremes.csv', mode='r') as extremes_file:
             F[line_count]=row[1]
             line_count += 1
 
-fig = plt.figure(figsize=(16,6),constrained_layout=True)
+fig = plt.figure(figsize=(16,7), constrained_layout=True)
 # Declares a 2x5 grid.
-# the last axes spans 4 cells
+
 grid = plt.GridSpec(2, 5, figure=fig)
-for i in range(0,4):
+for i in range(0,5):
     plt.subplot(grid[0,i])
     plt.subplot(grid[1,i])
-plt.subplot(grid[0:,3:])
 
-fig = plt.gcf()
 ax_list = fig.axes
-Nevents = len(ax_list)
+nplots = len(ax_list)
 
-for i in range(0,Nevents):
-        path_to_file_ux = prefix+'event_{:d}_feb/ux.dat'.format(int(idx[Nevents-i-1]+1))
-        path_to_file_uy = prefix+'event_{:d}_feb/uy.dat'.format(int(idx[Nevents-i-1]+1))
-
-        # Opens the files and jump to peak drag
-        # (middle of timeseries)
-        fx = open(path_to_file_ux, "rb")  # reopen the file
-        fy = open(path_to_file_uy, "rb")  # reopen the file
-        fx.seek(49*N*8, 0)  
-        fy.seek(49*N*8, 0)  
+path_to_file_ux = prefix+'event_{:d}_feb/ux.dat'.format(int(idx[1]+1))
+path_to_file_uy = prefix+'event_{:d}_feb/uy.dat'.format(int(idx[1]+1))
+fx = open(path_to_file_ux, "rb")  # reopen the file
[…]

### 2020-04-03 figures/illustr_density_streamlines/illustr_density_streamlines.py

@@ -5,6 +5,7 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
 
 Dx = 513
 Dy = 129
@@ -109,8 +110,7 @@ ax_list.set_xlabel(r'$x/R$',fontsize=22)
 ax_list.set_ylabel(r'$y/R$',fontsize=22)
 
 
-fname = 'illustr_density_streamlines.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()
-            

### 2020-04-03 figures/illustr_extrms_vorticity/illustr_extrms_vorticity.py

@@ -5,14 +5,14 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
 
 Dx = 513
 Dy = 129
 N = Dx*Dy
 F = np.zeros(104)
 idx = np.zeros(104)
-
-prefix = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/"
+prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is lowest fluctuation
@@ -24,29 +24,29 @@ with open(prefix+'extremes.csv', mode='r') as extremes_file:
             F[line_count]=row[1]
             line_count += 1
 
-fig = plt.figure(figsize=(16,7), constrained_layout=True)
+fig = plt.figure(figsize=(16,6),constrained_layout=True)
 # Declares a 2x5 grid.
-
+# the last axes spans 4 cells
 grid = plt.GridSpec(2, 5, figure=fig)
-for i in range(0,5):
+for i in range(0,4):
     plt.subplot(grid[0,i])
     plt.subplot(grid[1,i])
+plt.subplot(grid[0:,3:])
 
+fig = plt.gcf()
 ax_list = fig.axes
-nplots = len(ax_list)
+Nevents = len(ax_list)
 
-path_to_file_ux = prefix+'event_{:d}_feb/ux.dat'.format(int(idx[1]+1))
-path_to_file_uy = prefix+'event_{:d}_feb/uy.dat'.format(int(idx[1]+1))
-fx = open(path_to_file_ux, "rb")  # reopen the file
-fy = open(path_to_file_uy, "rb")  # reopen the file
-bbox = dict(boxstyle="round", fc="0.9")
-for i in range(0,nplots):
-        # trick to advance in time with row major ordering
-        I = 2*(i%5) + i//5
-        
-        offset=(40+2*i)-1
-        fx.seek(offset*N*8, 0)  
-        fy.seek(offset*N*8, 0)  
+for i in range(0,Nevents):
+        path_to_file_ux = prefix+'event_{:d}_feb/ux.dat'.format(int(idx[Nevents-i-1]+1))
[…]

### 2020-04-03 figures/return_time/return_time.py

@@ -1,7 +1,7 @@
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
-from os.path import abspath, dirname, join
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
 
@@ -33,7 +33,7 @@ plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)
 
 fname = join(
-    abspath(dirname(__file__)), splitext(__file__)[0] + ".eps"
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
     )
 
 plt.savefig(fname)

### 2020-04-03 figures/timeseries_extremes/timeseries_extremes.py

@@ -1,6 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
-
+from os.path import abspath, dirname, join, basename, splitext
 
 prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/instant_events/'
 
@@ -37,7 +37,7 @@ ax.set_xlabel('$t$',fontsize=22)
 plt.xticks(fontsize=18)
 plt.yticks(fontsize=18)
 
-fname = 'timeseries_extremes.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/timeseries_extrms_AVG/sort_extremes.py

@@ -29,5 +29,5 @@ for i in range(0,Nevents):
     f = np.fromfile(fileName, float, -1, "")
     F[i] = np.mean(f)
 
-np.savetxt('list.txt', np.argsort(-F), '%0.2d')
+np.savetxt('data/list.txt', np.argsort(-F), '%0.2d')
 

### 2020-04-03 figures/timeseries_extrms_AVG/timeseries_extrms_AVG.py

@@ -1,20 +1,25 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
 
-prefix = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/' \
+path_events = '/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/etude_dynamique/' \
          + 'averaged_drag/'
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 # list.txt is generated with ./sort_extremes.py
 # contains the sorted list of extremes according to the
 # time-averaged drag (highest value first)
-sortedI = np.loadtxt('list.txt', dtype=int)
+sortedI = np.loadtxt(join(prefix, "list.txt"), dtype=int)
 
 sig=0.0412
 m=0.0252
 # Gets the value of mu and sigma for time-averaged drag
-musigma = np.fromfile("../../data/musigma_AVG_10.dat", float, -1, "");
+musigma = np.fromfile(join(prefix, "musigma_AVG_10.dat"), float, -1, "");
 
 T = 10 # Physical duration of instant. trajectories
 
@@ -23,7 +28,7 @@ minArray=np.zeros(Nsubplots) # containers for max and min over instant. drag
 maxArray=np.zeros(Nsubplots) # timeseries
 fig, axes = plt.subplots(nrows=2,ncols=3, figsize=(18,6), constrained_layout=True)
 for i in range(0,Nsubplots):
-    fileName=prefix+'event_{:d}_AVG/'.format(sortedI[i]+1)+'data_force.datout'
+    fileName=path_events+'event_{:d}_AVG/'.format(sortedI[i]+1)+'data_force.datout'
     f = np.fromfile(fileName, float, -1, "")
     # amplitude of fluctuation for time-average
     F_avg = np.sum(f)/(len(f)-1)
@@ -62,7 +67,7 @@ for i in range(0,6):
                                    for x in axes[row,col].get_xticks()])
     axes[row,col].tick_params(axis='both', which='major', labelsize=22)
 
-fname = 'timeseries_extrms_AVG.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/AMS_drag_resampling/AMS_drag_resampling.py
Fix paths and relocate data Gif 14-19

@@ -1,5 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
 color_list = plt.rcParams['axes.prop_cycle'].by_key()['color']
@@ -65,7 +66,7 @@ ax.tick_params(axis="y", labelsize=18)
 
 ax.legend(loc='best', fontsize=18)
 
-fname = 'AMS_drag_resampling.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/AMS_drag_trajectories/AMS_drag_trajectories.py

@@ -1,5 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
 
@@ -58,7 +59,8 @@ ax.tick_params(axis="y", labelsize=18)
 ax.legend(loc='best', fontsize=18)
 ax.set_xlim((0,5))
 
-fname = 'AMS_drag_trajectories.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
 
-plt.show()

### 2020-04-03 figures/IS_GKTL/IS_GKTL.py

@@ -1,49 +1,55 @@
+import os
 import numpy as np
-import seaborn as sns
 import matplotlib.pyplot as plt
+from scipy.special import erf
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
+
+from jfm_paper import utils
+
 
 def ret_gaussian(x):
     """ Returns zero-mean, unit variance at point
     """
-    return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
+    return (1.0 / (np.sqrt(2.0 * np.pi))) * np.exp(-0.5 * x * x)
 
-prefix = '/home/tlestang/transition_hdd/thibault/These/tailleur_lecomte/lbm/2_sq/' \
-         + 'grid/illustration_IS/'
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
 
-directories = ['k_0.02_2018_03_26_10_01_46/', 'k_0.025_2018_03_25_20_13_40/',
-           'k_0.03_2018_03_25_20_13_40/'];
-k_value = [0.02, 0.015, 0.03]
+musigma = np.fromfile(join(str(main_data_path), "musigma_AVG_10.dat"), float, -1, "")
 
-sig = 0.0087714;
+gktl_dirs = ["er_lannic_1", "er_lannic_0", "er_lannic"]
+gktl_dirs = [os.path.join(str(main_data_path), gktl_dir) for gktl_dir in gktl_dirs]
 
-fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
+fig, ax = plt.subplots(figsize=(8, 5), constrained_layout=True)
 
-i=0
-for directory in directories:
-    A = np.fromfile(prefix+directory+'A.datout', float, -1, "")
-    A = A/sig # A is already zero-mean
+for gktl_dir in gktl_dirs:
+    gktl_params = utils.get_gktl_parameters(gktl_dir)
+    df = utils.get_gktl_drag_trajectories(gktl_dir)
+    traj = df.values
+    gktl_params = utils.get_gktl_parameters(gktl_dir)
+    A = [np.mean(traj[:, i]) for i in range(gktl_params["nc"])]
 
-    # https://seaborn.pydata.org/generated/seaborn.kdeplot.html
[…]

### 2020-04-03 figures/IS_GKTL/IS_GKTL_2.py

@@ -1,53 +1,0 @@
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-from scipy.special import erf
-
-from jfm_paper import utils
-
-
-def ret_gaussian(x):
-    """ Returns zero-mean, unit variance at point
-    """
-    return (1.0 / (np.sqrt(2.0 * np.pi))) * np.exp(-0.5 * x * x)
-
-
-musimga = np.fromfile("../..//data/musigma_AVG_10.dat", float, -1, "")
-
-prefix = "../../data/"
-gktl_dirs = ["er_lannic_1", "er_lannic_0", "er_lannic"]
-gktl_dirs = [os.path.join(prefix, gktl_dir) for gktl_dir in gktl_dirs]
-
-fig, ax = plt.subplots(figsize=(8, 5), constrained_layout=True)
-
-for gktl_dir in gktl_dirs:
-    gktl_params = utils.get_gktl_parameters(gktl_dir)
-    df = utils.get_gktl_drag_trajectories(gktl_dir)
-    traj = df.values
-    gktl_params = utils.get_gktl_parameters(gktl_dir)
-    A = [np.mean(traj[:, i]) for i in range(gktl_params["nc"])]
-
-    k = gktl_params["k"]
-    label = "$k = {:1.2f}$".format(k)
-    n, bins, patches = plt.hist(A / musimga[1], 50, alpha=0.5, label=label)
-
-binsize = bins[1] - bins[0]
-nc = utils.get_gktl_parameters(gktl_dirs[0])["nc"]
-A = np.sqrt(np.pi) * 0.5
-mesh = np.arange(-5, 5, binsize)
-hist = A * nc * (erf(mesh[1:]) - erf(mesh[:-1]))
-midpoints = [0.5 * (b + a) for a, b in zip(mesh[:-1], mesh[1:])]
-plt.bar(
-    x=midpoints, height=hist, width=binsize, alpha=0.5, color="#808080", label="$k=0$"
-)
-
-plt.xlabel("$F_{T_a} / \\sigma_{T_a}$", fontsize=22)
-plt.ylabel("#trajectories", fontsize=22)
-ax.tick_params(axis="both", which="major", labelsize=16)
-ax.legend(loc="best", fontsize=16)
-plt.xlim((-2, 6))
-
-fname = os.path.splitext(__file__)[0] + ".eps"
-plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/illustr_extrms_vorticity/illustr_extrms_vorticity.py

@@ -93,6 +93,4 @@ fname = join(
     abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
     )
 plt.savefig(fname)
-
-plt.show()
             

### 2020-04-03 figures/illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL.py

@@ -6,17 +6,24 @@ import numpy as np
 import csv
 import matplotlib.pyplot as plt
 from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
+
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 Dx = 513
 Dy = 129
 N = Dx * Dy
 F = []
 idx = []
-gktl_dir = "../../er_lannic"
+gktl_dir = join(str(main_data_path), "er_lannic")
 # Files extremes.csv contains the index of the events and the corresponding
 # peak drag amplitude, in units of sigma
 # First row is highest fluctuation
-with open("extremes.csv", mode="r") as extremes_file:
+with open(join(prefix, "extremes.csv"), mode="r") as extremes_file:
     extremes_reader = csv.reader(extremes_file, delimiter=",")
     for _idx, _F, t_star in extremes_reader:
         idx.append(int(_idx))
@@ -92,7 +99,7 @@ for i in range(0, Nevents):
 cbar = fig.colorbar(im, ax=ax_list)
 cbar.ax.tick_params(labelsize=16)
 
-fname = "illustr_extrms_vorticity.eps"
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/illustr_extrms_vorticity_GKTL/sort_extremes.py

@@ -2,6 +2,8 @@ import os
 import numpy as np
 import argparse
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
 
 from jfm_paper import utils
 from jfm_paper.resimulate_extreme_GKTL import resimulate
@@ -21,8 +23,13 @@ parser.add_argument(
 
 args = parser.parse_args()
 
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
 musimga = np.fromfile(
-    "/home/tlestang/articles/jstat_paper/data/musigma_AVG_10.dat", float, -1, ""
+    join(str(main_data_path), "musigma_AVG_10.dat"), float, -1, ""
 )
 sig = 0.0412  # Standard deviation of instant. drag computed over CR
 m = 0.0252  # Average instant. drag computed over CR
@@ -53,7 +60,7 @@ for counter, common_ancestor_idx in enumerate(dict_of_parents):
     traj_idx.append(j)
     drag_maxima.append(np.amax(drag_maxima_ingroup))
 
-with open("extremes.csv", "w") as f:
+with open(join(prefix, "extremes.csv"), "w") as f:
     for i, idx in enumerate(np.argsort(-np.array(drag_maxima))):
         fileName = os.path.join(
             args.directory, "recon_rep_0_clone_{}.traj".format(traj_idx[idx])

### 2020-04-03 figures/return_times_GKTL/return_times_GKTL.py

@@ -3,6 +3,8 @@ import pandas as pd
 import matplotlib.pyplot as plt
 from scipy import interpolate
 import returntimes_GKTL as rtt
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
 
 def get_interpolated_return_times(directories, Ta, T, k, N, nb_points=10):
     nb_directories = len(directories)
@@ -30,14 +32,19 @@ nbfiles = 10
 tau_c = 2000
 T = 10
 
-mu_and_sigma_file = '../../data/return_times_GKTL/musigma_AVG_10.dat'
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
+mu_and_sigma_file = join(str(main_data_path), "musigma_AVG_10.dat")
 musigma = np.fromfile(mu_and_sigma_file, float, -1, "")
 
 fig, ax = plt.subplots(figsize=(8,5),constrained_layout=True)
 ax.set_xscale('log')
 
 for k in [0.005, 0.01]:
-    GKTL_dir_template = '../../data/return_times_GKTL/gktl_test_{}_k_{}_T_{}'.format(N,k,Ta)
+    GKTL_dir_template = join(prefix, "gktl_test_{}_k_{}_T_{}".format(N,k,Ta))
     directories = ['{}_{}'.format(GKTL_dir_template, i) for i in range(0,nbfiles)]
     ra_interp, df = get_interpolated_return_times(directories, Ta, T,
                                                   k, N, nb_points=30)
@@ -49,7 +56,7 @@ for k in [0.005, 0.01]:
     plt.fill_between(ra_interp, mean-error, mean+error, color=line.get_color(), alpha=0.2)
     
 
-direct_sampling_file='../../data/return_times_GKTL/rt_Nc1024_Ta30_T10.csv'
+direct_sampling_file=join(prefix, "rt_Nc1024_Ta30_T10.csv")
 # Direct sampling file contains M rows (blocks) and K columns (repetitions).
 # Numbers are the block maxima
 df = pd.read_csv(direct_sampling_file, comment='#', index_col=0)
@@ -77,8 +84,7 @@ ax.plot([np.amax(ra), np.amax(ra)], [0, 6], linestyle='--', linewidth=0.7, color
[…]

### 2020-04-03 figures/timeseries_extrms_AVG_GKTL/sort_extremes.py

@@ -11,14 +11,19 @@
 from os.path import join
 import numpy as np
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
 
 from jfm_paper import utils
 
-gktl_dir = "../../data/er_lannic"
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
-musimga = np.fromfile(
-    "../../data/musigma_AVG_10.dat", float, -1, ""
-)
+gktl_dir = join(str(main_data_path), "er_lannic")
+
+musigma = np.fromfile(join(str(main_data_path), "musigma_AVG_10.dat"), float, -1, "");
 sig=0.0412 # Standard deviation of instant. drag computed over CR
 m=0.0252 # Average instant. drag computed over CR
 
@@ -50,5 +55,5 @@ for counter, common_ancestor_idx in enumerate(dict_of_parents):
     f = np.fromfile(fileName, float, -1, "")
     F.append(np.mean(f))
 
-np.savetxt('list.txt', [traj_idx[ii] for ii in np.argsort(-np.array(F))], '%0.2d')
+np.savetxt(join(prefix, "list.txt"), [traj_idx[ii] for ii in np.argsort(-np.array(F))], '%0.2d')
 

### 2020-04-03 figures/timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL.py

@@ -1,20 +1,27 @@
 from os.path import join
 import numpy as np
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
+from pathlib import PurePath
 
 plt.style.use('ggplot')
 
-gktl_dir = "../../data/er_lannic"
+main_data_path = PurePath(abspath(dirname(__file__))).parent.parent / "data"
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
+gktl_dir = join(str(main_data_path), "er_lannic")
 
 # list.txt is generated with ./sort_extremes.py
 # contains the sorted list of extremes according to the
 # time-averaged drag (highest value first)
-sortedI = np.loadtxt('list.txt', dtype=int)
+sortedI = np.loadtxt(join(prefix, "list.txt"), dtype=int)
 
 sig=0.0412
 m=0.0252
 # Gets the value of mu and sigma for time-averaged drag
-musigma = np.fromfile("../../data/musigma_AVG_10.dat", float, -1, "");
+musigma = np.fromfile(join(str(main_data_path), "musigma_AVG_10.dat"), float, -1, "");
 
 T = 10 # Physical duration of instant. trajectories
 
@@ -62,8 +69,7 @@ for i in range(0,6):
                                    for x in axes[row,col].get_xticks()])
     axes[row,col].tick_params(axis='both', which='major', labelsize=22)
 
-fname = 'timeseries_extrms_AVG.eps'
-fname = 'timeseries_extrms_AVG.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/AMS_OU/AMS_OU.py
Fix paths and relocate data Gif 21-23

@@ -1,20 +1,25 @@
 import numpy as np
 import matplotlib.style
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
 plt.style.use('ggplot')
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 Ta = 5;
 Nc = 32;
 sig = 1/np.sqrt(2);
 
-compTime = np.fromfile("./comp.dat", float, -1, "")
-amax = np.fromfile("./amax.dat", float, -1, "")
-Z = np.fromfile("./Z.dat", float, -1, "")
+compTime = np.fromfile(join(prefix, "comp.dat"), float, -1, "")
+amax = np.fromfile(join(prefix, "amax.dat"), float, -1, "")
+Z = np.fromfile(join(prefix, "Z.dat"), float, -1, "")
 
 xx = Ta*Nc + np.cumsum(compTime);
 
-buf = np.fromfile("./return_time_analytical.dat", float, -1, "")
+buf = np.fromfile(join(prefix, "return_time_analytical.dat"), float, -1, "")
 n = int(len(buf)/2)
 avec = buf[0:n]
 rtheo = buf[n:]
@@ -36,7 +41,7 @@ ax.legend(loc='best', fontsize=18)
 plt.xticks(fontsize=18)
 plt.yticks(fontsize=18)
 
-fname = 'AMS_OU.eps'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-
-plt.show()

### 2020-04-03 figures/AMS_OU/compute_analytical_rt.py

@@ -1,6 +1,11 @@
 import numpy as np
 import scipy.integrate as integrate
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 def g(z):
     return np.exp(z*z)
@@ -23,7 +28,7 @@ for a in avec:
     rstat[k] = np.sqrt(prefact) * integrate.quad(fstat,-20,a)[0]
     k = k+1;
 
-fo = open("return_time_analytical.dat", "wb")
+fo = open(join(prefix, "return_time_analytical.dat"), "wb")
 avec.tofile(fo, "")
 rstat.tofile(fo, "")
 fo.close()

### 2020-04-03 figures/illustr_AMS/illustr_AMS.py

@@ -1,8 +1,8 @@
 import os
-
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas as pd
+from os.path import abspath, dirname, join, basename, splitext
 
 prefix = '/home/tlestang/transition_hdd/thibault/These/AMS/AMS_OU_AVG/run_for_figure_AMS/'
 with open(os.path.join(prefix, 'input.in'), 'r') as f:
@@ -73,6 +73,7 @@ plt.xticks(fontsize=28)
 ax.set_xlim((0,9))
 ax.set_xlabel(r'$t$', fontsize=30)
 
-fname = 'illustr_AMS.pdf'
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
-plt.show()

### 2020-04-03 figures/pdf_drag_with_perturbation/pdf_drag_with_perturbation.py

@@ -1,17 +1,19 @@
 import sys
-from os.path import splitext
-
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
 
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
 
 fig, ax = plt.subplots(constrained_layout=True, figsize=(8, 5))
 
-series = pd.read_csv("../../data/PDF_drag.csv", comment="#", index_col=0)
+series = pd.read_csv(join(prefix, "PDF_drag.csv"), comment="#", index_col=0)
 trunc_series = series.iloc[np.where((series.index.values > -6.5) & (series.index.values < 8.5))] 
 trunc_series.plot(ax=ax, lw=2, legend=False)
-series_perturb = pd.read_csv("../../data/PDF_drag_with_perturbation.csv", comment="#", index_col=0)
+series_perturb = pd.read_csv(join(prefix, "PDF_drag_with_perturbation.csv"), comment="#", index_col=0)
 series_perturb.plot(ax=ax, lw=2, legend=False)
 
 ax.set_xlabel("$x=(f_d - \\mu)/\\sigma$", fontsize=22)
@@ -20,7 +22,9 @@ ax.legend(["No perturbation", "$\\epsilon = 0.002$"], fontsize=18)
 ax.tick_params(axis='both', which='major', labelsize=18)
 
 ax.set_yscale("log")
-fname = splitext(__file__)[0] + ".eps"
+
+fname = join(
+    abspath(dirname(__file__)), basename(splitext(__file__)[0]) + ".eps"
+    )
 plt.savefig(fname)
 
-plt.show()

### 2020-04-03 figures/ecoulement_typique/ecoulement_typique.py
Change name of dirs, figure scripts to fig1, fig2, fig3 ...

@@ -1,110 +1,0 @@
-# Produces a figure illustrating the vorticity field at the moment of peak drag for 7 events.
-# The figure consists of 7 axes. One axes is four times bigger as the others, and contains 
-# annotations
-import numpy as np
-import csv
-import matplotlib.pyplot as plt
-from matplotlib.patches import Rectangle
-from os.path import abspath, dirname, join
-
-Dx = 513
-Dy = 129
-N = Dx*Dy
-F = np.zeros(104)
-idx = np.zeros(104)
-
-prefix = join(
-    abspath(dirname(__file__)), "data"
-    )
-
-fig = plt.figure(figsize=(16,7), constrained_layout=True)
-# Declares a 2x5 grid.
-
-grid = plt.GridSpec(2, 5, figure=fig)
-# the first row is divided into five snapshots
-for i in range(0,5):
-    plt.subplot(grid[0,i])
-#the second row is for the drag signal
-plt.subplot(grid[1,:])
-
-ax_list = fig.axes
-nplots = len(ax_list)
-
-path_to_file_ux = join(prefix, 'ux.dat')
-path_to_file_uy = join(prefix, "uy.dat")
-fx = open(path_to_file_ux, "rb")  # reopen the file
-fy = open(path_to_file_uy, "rb")  # reopen the file
-bbox = dict(boxstyle="round", fc="0.9")
-
-for i in range(0,nplots-1):
-        # trick to advance in time with row major ordering
-        
-        fx.seek(i*N*8, 0)  
-        fy.seek(i*N*8, 0)  
-
-        # Get velocity fields
-        ux = np.fromfile(fx, float, N, "")
-        ux = np.reshape(ux, (129,513), 'F')
-        uy = np.fromfile(fy, float, N, "")
-        uy = np.reshape(uy, (129,513), 'F')
-    
-        ymin=32;ymax=96
-        xmin=236;xmax=310
-        # Compute vorticity using finite-differences
-        dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
-        dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
-        vort = dxy - dyx
-        # The sign of the vorticity is wrong if using a right-handed coord. system.
[…]

### 2020-04-03 figures/fig19/returntimes_avg_blocks.py

@@ -1,0 +1,48 @@
+import os
+import sys
+import datetime
+import numpy as np
+import pandas as pd
+
+tau_c = 2000
+Nc = 1024
+Ta = 30
+T = 10
+blockLength = Nc*Ta*tau_c
+DT = 10*T*tau_c
+
+nb_files = 10
+files = ["/home/tlestang/grid/control_long_{:d}/"
+         "data_force.datout_AVG_20000".format(i+1) for i in range(0,nb_files)]
+
+for filename in files:
+    if not(os.path.isfile(filename)):
+        raise FileNotFoundError('File {} not found.'.format(filename))
+
+n_windows_in_block = int(np.floor(blockLength/DT))
+Ntot = int(os.path.getsize(files[0])/8.)
+Nblocks = int(np.floor(Ntot / blockLength))
+
+amaxtilde = np.zeros((n_windows_in_block,nb_files*Nblocks))
+for i,filename in enumerate(files):
+    print(filename)
+    if Ntot < blockLength:
+        raise valueError("Length of timeseries for file {} is smaller"
+                        "than block lenght".format(filename))
+    with open(filename, 'rb') as f:
+        for block in range(0,Nblocks):
+            f.seek(block*blockLength,0)
+            list_of_max = [np.max(np.fromfile(f, float, DT, '')) for n in range(0,n_windows_in_block)]
+            amaxtilde[:,i*Nblocks + block] = np.sort(list_of_max)[::-1]
+
+df = pd.DataFrame(amaxtilde)
+
+with open('rt_Nc1024_Ta30_T10.csv', 'w') as f:
+    f.write('# Generated with {}\n'.format(sys.argv[0]))
+    f.write('# {}\n'.format(str(datetime.datetime.now())))
+    f.write('# NBlocks = {}\n'.format(Nblocks))
+    f.write('# DT = {}\n'.format(DT))
+    f.write('# using {} files'.format(nb_files))
+    f.write('#\n')
+    df.to_csv(f, header=True)
+

### 2020-04-03 figures/fig2/fig2.py

@@ -1,0 +1,109 @@
+# Produces a figure illustrating the vorticity field at the moment of peak drag for 7 events.
+# The figure consists of 7 axes. One axes is four times bigger as the others, and contains 
+# annotations
+import numpy as np
+import csv
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
+from os.path import abspath, dirname, join, basename, splitext
+
+Dx = 513
+Dy = 129
+N = Dx*Dy
+F = np.zeros(104)
+idx = np.zeros(104)
+
+prefix = join(
+    abspath(dirname(__file__)), "data"
+    )
+
+fig = plt.figure(figsize=(16,7), constrained_layout=True)
+# Declares a 2x5 grid.
+
+grid = plt.GridSpec(2, 5, figure=fig)
+# the first row is divided into five snapshots
+for i in range(0,5):
+    plt.subplot(grid[0,i])
+#the second row is for the drag signal
+plt.subplot(grid[1,:])
+
+ax_list = fig.axes
+nplots = len(ax_list)
+
+path_to_file_ux = join(prefix, 'ux.dat')
+path_to_file_uy = join(prefix, "uy.dat")
+fx = open(path_to_file_ux, "rb")  # reopen the file
+fy = open(path_to_file_uy, "rb")  # reopen the file
+bbox = dict(boxstyle="round", fc="0.9")
+
+for i in range(0,nplots-1):
+        # trick to advance in time with row major ordering
+        
+        fx.seek(i*N*8, 0)  
+        fy.seek(i*N*8, 0)  
+
+        # Get velocity fields
+        ux = np.fromfile(fx, float, N, "")
+        ux = np.reshape(ux, (129,513), 'F')
+        uy = np.fromfile(fy, float, N, "")
+        uy = np.reshape(uy, (129,513), 'F')
+    
+        ymin=32;ymax=96
+        xmin=236;xmax=310
+        # Compute vorticity using finite-differences
+        dxy = (uy[ymin:ymax, xmin+1:xmax+1] - uy[ymin:ymax, xmin-1:xmax-1])*0.5
+        dyx = (ux[ymin+1:ymax+1, xmin:xmax] - ux[ymin-1:ymax-1, xmin:xmax])*0.5
+        vort = dxy - dyx
[…]

### 2020-04-03 figures/fig3/fig3.py

@@ -1,0 +1,93 @@
+import numpy as np
+import matplotlib.style
+import matplotlib.pyplot as plt
+from os.path import abspath, dirname, join, basename, splitext
+
+def bundle_nearest_points(I, dist):
+    """
+       Given a list of number I (int or floats), return a list of lists, each sublist containing
+       subsequent elements distant from less than dist of the next one.
+
+       Parameters
+       -----------
+       I: list
+       dist: int or float
+
+       Returns
+       --------
+       list of lists
+    """
+    bundle = []
+    list_of_bundles = []
+    bundle.append(I[0])
+    for tup in zip(I,I[1:]):     
+        if (tup[1]-tup[0]) < dist:
+            bundle.append(tup[1])
+        else:
+            list_of_bundles.append(bundle.copy())
+            bundle.clear()
+            bundle.append(tup[1])
+            
+    list_of_bundles.append(bundle.copy())
+    return list_of_bundles
+
+
+plt.style.use('ggplot')
+
+path_to_file = "/home/tlestang/articles/jfm_paper/data/control_2_2017_10_31_14_07_55" \
+               + "/data_force.datout"
+tau_0 = 500 # Turnover time in units of LBM timesteps
+             # The correlation time tau_c is 2000 timesteps, and the
+             # tau_c is 4tau_0
+T = 1000
+f = np.fromfile(path_to_file, float, T*tau_0, "");
+tspan = np.linspace(0,T,len(f))
+
+a = 0.17 # threshold
+I = np.argwhere(f>a) # find all indices where f>a
+
+tau_c = 4*tau_0
+list_of_bundles = bundle_nearest_points(I, tau_c)
+
+point_array = np.zeros(len(list_of_bundles))
+value_array = np.zeros(len(list_of_bundles))                    
+for i, bundle in enumerate(list_of_bundles):
+    # argmax return the index relative to f[bundle]
+    # must add offset bundle[0]
+    point_array[i] = tspan[np.argmax(f[bundle])+bundle[0]]
+    value_array[i] = np.amax(f[bundle])
+
+fig,ax = plt.subplots(1,1,figsize=(12,4), constrained_layout=True)
+ax.set_xlim(0, T)
+line, = ax.plot(tspan, f)
[…]

### 2020-04-03 figures/fig4b/autocorr.py

@@ -1,0 +1,28 @@
+import sys
+import datetime
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+from jfm_paper.utils import get_autocorr
+
+data = "/home/tlestang/grid/control_long_1/data_force.datout"
+
+nreps = 15000
+tau_0 = 450  # Turn over in units of timesteps
+T = 11 * tau_0
+
+
+var = 0.00173
+mu = 0.0253
+
+# In the following the lag window is prescribed in units of timesteps (dt=1)
+series = get_autocorr(data, nreps, T, mu=mu, var=var, dtype=float, verbose=True)
+
+with open("autocorr_drag.csv", "w") as f:
+    f.write("# Generated with {}\n".format(sys.argv[0]))
+    f.write("# {}\n".format(str(datetime.datetime.now())))
+    f.write("# nreps = {}\n".format(nreps))
+    f.write("# using {} \n".format(filename))
+    f.write("#\n")
+    series.to_csv(f, header=True)

### 2020-04-03 figures/illustrate_return_time/illustrate_return_time.py

@@ -1,94 +1,0 @@
-import numpy as np
-import matplotlib.style
-import matplotlib.pyplot as plt
-from os.path import abspath, dirname, join
-
-def bundle_nearest_points(I, dist):
-    """
-       Given a list of number I (int or floats), return a list of lists, each sublist containing
-       subsequent elements distant from less than dist of the next one.
-
-       Parameters
-       -----------
-       I: list
-       dist: int or float
-
-       Returns
-       --------
-       list of lists
-    """
-    bundle = []
-    list_of_bundles = []
-    bundle.append(I[0])
-    for tup in zip(I,I[1:]):     
-        if (tup[1]-tup[0]) < dist:
-            bundle.append(tup[1])
-        else:
-            list_of_bundles.append(bundle.copy())
-            bundle.clear()
-            bundle.append(tup[1])
-            
-    list_of_bundles.append(bundle.copy())
-    return list_of_bundles
-
-
-plt.style.use('ggplot')
-
-path_to_file = "/home/tlestang/articles/jfm_paper/data/control_2_2017_10_31_14_07_55" \
-               + "/data_force.datout"
-tau_0 = 500 # Turnover time in units of LBM timesteps
-             # The correlation time tau_c is 2000 timesteps, and the
-             # tau_c is 4tau_0
-T = 1000
-f = np.fromfile(path_to_file, float, T*tau_0, "");
-tspan = np.linspace(0,T,len(f))
-
-a = 0.17 # threshold
-I = np.argwhere(f>a) # find all indices where f>a
-
-tau_c = 4*tau_0
-list_of_bundles = bundle_nearest_points(I, tau_c)
-
-point_array = np.zeros(len(list_of_bundles))
-value_array = np.zeros(len(list_of_bundles))                    
-for i, bundle in enumerate(list_of_bundles):
-    # argmax return the index relative to f[bundle]
-    # must add offset bundle[0]
-    point_array[i] = tspan[np.argmax(f[bundle])+bundle[0]]
-    value_array[i] = np.amax(f[bundle])
-
-fig,ax = plt.subplots(1,1,figsize=(12,4), constrained_layout=True)
-ax.set_xlim(0, T)
-line, = ax.plot(tspan, f)
-line2, = ax.plot(point_array, value_array, 'o', linestyle='None', markersize=4)
[…]

### 2020-04-03 appendixA.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
 \section{The \acl{lbm}}
@@
 	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
+	\includegraphics[width=0.3\linewidth]{fig20/fig20}
 	\caption{Sketch of the D2Q9 lattice.
@@
 Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 appendixC.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
 \section{The \acl{gktl} algorithm}
@@
 with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 appendixD.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
 \section{Perturbation at branching time}
 \label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
 The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
 resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
 In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
 %
@@
 %
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
+In other words, the perturbation is a  random linear combination of $N_s$ snapshots.
 Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
 %
 In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
 %
@@
 %
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
 Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
@@
   \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
+\includegraphics[width=.7\linewidth]{fig23/fig23}
 \caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
@@
 \end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 draft_p2_jfm.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
 \usepackage{subfig}
-%\usepackage{caption}
+\usepackage{hyperref}
 
@@
 
-\graphicspath{{../figures/}}
+\graphicspath{{./figures/}}
 
@@
 \title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
 \EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
@@
 	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
+      This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
 \end{abstract}
 
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
+\include{section1}
+\include{section2}
+\include{section3}
+\include{section4}
+\include{section5}
 
@@
 \appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
+\include{appendixA}
+\include{appendixB}
+\include{appendixC}
+\include{appendixD}
 
 \bibliographystyle{jfm}
-\bibliography{../biblio}
+\bibliography{biblio}
 

### 2020-04-03 section2.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
 \section{Description of the numerical case study}
@@
 	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
+	\includegraphics[width=\linewidth]{fig1/fig1}
 	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
@@
 We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
 \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
+
 % rapid description of the flow
@@
 The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
 % lattice Boltzmann method 
@@
 	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
+	\includegraphics[width=\linewidth]{fig2/fig2}
 	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
@@
 Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
 %
@@
 %
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
 
@@
 where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
 %
@@
 
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
 \begin{figure}
 	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
+	\includegraphics[width=\linewidth]{fig3/fig3}
 	\caption{
@@
 	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
+		\includegraphics[width=.4\linewidth]{fig4a/fig4a}}
 	\subfloat[Autocorrelation of drag fluctuations]
 	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
+		\includegraphics[width=.45\linewidth]{fig4b/fig4b}}  
 	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
@@
 The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 section3.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
 \section{Extreme fluctuations of the drag by means of direct sampling}
@@
 	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
+	\includegraphics[width=.6\linewidth]{fig5/fig5}
 	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
@@
 We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
+
 If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
 one might expect that 
@@
 Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
 
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
+	\includegraphics[width=.8\linewidth]{fig6/fig6}
 	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
+An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).}
 \end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
 \EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
@@
 
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
 \subsubsection{Fluid dynamics related to extreme drag fluctuations}
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
+	\includegraphics[width=.7\linewidth]{fig7/fig7}
 	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
@@
 This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
 Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
 \EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
 To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
@@
 	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
+	\includegraphics[width=\linewidth]{fig8/fig8}
 	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
@@
 	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
+	\includegraphics[width=.5\linewidth]{fig9/fig9}
 	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
+	\includegraphics[width=.8\linewidth]{fig10/fig10}
 	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
@@
 
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
 \begin{figure}
 	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
+	\includegraphics[width=.7\linewidth]{fig11/fig11}
 	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
@@
 On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
 These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
@@
 	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
+	\includegraphics[width=\linewidth]{fig12/fig12}
 	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
+	\includegraphics[width=.7\linewidth]{fig13/fig13}
+	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. }
 	\label{fig:PDF_AVG}
@@
 
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
 We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
@@
 In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
 %
@@
 %
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
 It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
@@
 %
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
 In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
@@
 very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 section4.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
 \section{Rare-event sampling algorithms}
@@
 %% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
 \EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
 %
@@
 %
-%\ZZ{paragraph moved from introduction}
 The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
@@
 %
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
  More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
@@
 This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
+\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is an unexplored area of application.}
 
@@
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
+\EL{The \ac{ams} algorithm is here used with, trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
 %
@@
 At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
 These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
 \EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
 The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
 The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
@@
 Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
 The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
 \EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
 We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
 
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
+	\includegraphics[width=.7\linewidth]{fig14/fig14}
 	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
 In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
+Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
+	\includegraphics[width=.7\linewidth]{fig15/fig15}
 	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
@@
 %
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
 %
@@
 We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
 However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
@@
 
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
 % perspectives
@@
 %
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
 \EL{This behaviour is independent of the choice of $N$ and $T_a$.
 Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
 
@@
 The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
+The \ac{ams} algorithm could be used in the same way as before by taking the time-averaged observable itself as the score function.
 However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
 %
@@
 \EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
 %
@@
 
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
 \begin{figure}
 	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
+	\includegraphics[width=0.7\linewidth]{fig16/fig16}
 	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
@@
 \EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
 Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
 %
@@
 % Choice of parameters and perturbation
-
 \EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
 The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
 Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
@@
 %
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
 %
@@
 %
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
 \EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
@@
 %
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
 \EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
 \EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
 \begin{figure}
 	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
+	\includegraphics[width=\linewidth]{fig17/fig17}
 	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
@@
 	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
+	\includegraphics[width=\linewidth]{fig18/fig18}
 	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
 	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
+\end{figure}}
 
@@
 The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
 %
@@
 %
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
 \begin{figure}
 	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
+	\includegraphics[width=.7\linewidth]{fig19/fig19}
 	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
@@
 \end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 section5.tex
Split master latex file into several files for each section

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
 \section{Conclusion}
 \label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
 \EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
@@
 Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
 In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
 A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
@@
 
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
 %
@@
 A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-03 section5.tex
Fix typos spotted by referee 1

@@
 In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
+Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, for example how these events build up.
 Our study shows this quite clearly.

### 2020-04-04 texfiles/appendixB.tex
Move body of paper into texfiles/ and rename main document to paper.tex

--- 2020-02-27 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
+\section{The \ac{ams} algorithm}
+\label{app:ams-short}
 
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
-\section{Introduction}
-
-% general comments on physical problem %
-%
-Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
-Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
-``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
-%
-
-From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
-The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
-Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
-However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
-%
-%
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
-%
-%	
-% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
-%
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
-
-%
-In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}.
-In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
-%
-An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
-Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise).
-Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
-An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
-%
-However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
-%
-%
-%
-Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
-It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
-%
-Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
-More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
-%
-%
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
-%
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
-%
-
-%
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
-The paper is organized in two parts.
-The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow.
-This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
-%
-The motivation for this study is twofold.
-Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations.
-This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
-Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them.
-This assessment is developed in the second part of the paper.
-
-% Annonce du plan
-The flow set-up is introduced and the dynamics related to typical
-drag fluctuations is described in section~\ref{sec:test_flow}.
-The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
-Both the instantaneous drag and time-averaged drag are considered.
-It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics.
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
-Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
-In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly.
-This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
-Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
-This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
-Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
+\subsection{Operating principle}
 
@@
 	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is a completely new field of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
+	\includegraphics[width=\linewidth]{fig21/fig21}
 	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
@@
 		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
+		\includegraphics[width=.7\linewidth]{fig22/fig22}
 		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
@@
 	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
 
@@
 	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
 	%
@@
 Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-04 texfiles/section1.tex
Move body of paper into texfiles/ and rename main document to paper.tex

--- 2020-02-28 jfm/draft_p2_jfm.tex
@@
-\documentclass{jfm}
-\pdfoutput=1
-\usepackage[english]{babel}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb}
-\usepackage[svgnames]{xcolor}
-\usepackage{graphicx}
-\usepackage{acro}
-\usepackage{subfig}
-%\usepackage{caption}
-
-\usepackage[normalem]{ulem}
-\usepackage{soul}
-
-\graphicspath{{../figures/}}
-
-\captionsetup[figure]{justification=raggedright}
-
-\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}
-
-\DeclareAcronym{pdf}{
-	short=PDF,
-	long=Probability Density Function,
-}
-\DeclareAcronym{iid}{
-	short=i.i.d.,
-	long=Independent Identically Distributed
-}
-\DeclareAcronym{ou}{
-	short=OU,
-	long=Ornstein-Uhlenbeck,
-}
-\DeclareAcronym{gktl}{
-	short=GKTL,
-	long=Giardina-Kurchan-Tailleur-Lecomte,
-}
-\DeclareAcronym{ams}{
-	short=AMS,
-	long=Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{tams}{
-	short=TAMS,
-	long=Trajectory Adaptive Multilevel Splitting,
-}
-\DeclareAcronym{scgf}{
-	short=SCGF,
-	long=Scaled Cumulant Generating Function,
-}
-\DeclareAcronym{lbm}{
-	short=LBM,
-	long=Lattice Boltzmann Method,
-}
-\DeclareAcronym{lbe}{
-	short=LBE,
-	long=Lattice Boltzmann Equation,
-}
-\DeclareAcronym{lgca}{
-	short=LGCA,
-	long=Lattice Gas Cellular Automata,
-}
-\DeclareAcronym{lbgk}{
-	short=LBGK,
-	long=Lattice Bhatnagar-Gross-Krook
-}
-\DeclareAcronym{OU}{
-	short=O-U,
-	long=Ornstein--Ulhenbeck
-}
-\DeclareAcronym{dns}{
-	short=DNS,
-	long=Direct Numerical Simulation,
-}
-\DeclareAcronym{md}{
-	short=MD,
-	long=Molecular Dynamics,
-}
-\DeclareAcronym{cfd}{
-	short=CFD,
-	long=Computational Fluid Dynamics,
-}
-
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
-\title{
-	%\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
-}
-
-\author{Thibault Lestang\aff{1}\aff{2}
-  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
-  Freddy Bouchet\aff{1}
-  \and Emmanuel L\'evêque\aff{2}}
-\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
-\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
-	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-	%
-	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-	%
-	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
-	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.
-These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
-        The present study is the first application to a fluid-structure interaction problem. 
-	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-	%
-	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
-	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
-%
-%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events.
-%	%
-%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
-%	%
-%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
-This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag.
-%	%
-%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
-%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.
-%	%
-%	
-%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
-%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics.
-Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}
-%
-%	%
-%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
-%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
-%	%
-%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms.
-%	% 
-%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%	%
-%	\QQ{phrase de conclusion ?}
-%
-%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
-%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently.
-%%
-%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
-%This work is the first application to fluid-structure interaction problems. 
-%%
-%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
-%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
-%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-%%
-%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body.
-}
-\end{abstract}
-
 \section{Introduction}
@@
 \EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
-%
-%However, this requires very long simulations since these events are rare.
-% general def on rare-event sampling %
 %
@@
 It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
-%, and is designed to sample rare events of some observable with a higher frequency.
-%    
 %
@@
 %
-% MOVED TO SECTION 4
-%
-%The \acl{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.
-%In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}.
-The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
-More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-%
-%During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-%Alongside numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-%Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
-%%
-%{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}. 
-%	This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems.
-Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.}
-%
-
 The paper is organized in two parts.
@@
 It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics. 
-%Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
-%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
 Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
@@
 Perspectives and conclusion end this work.
-
-\section{Description of the numerical case study}
-\label{sec:test_flow}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
-	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions.
-The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
-The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
-	\label{fig:illustr_ecoulement}
-\end{figure}
-
-% introduce the flow 
-%
-The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
-% why this flow?
-%
-Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
-% 
-We believe that this system embeds the characteristic features that makes 
-%\sout{the application of rare-event algorithms}
- \EL{this study} both relevant and challenging for fluid-structure-interaction problems. 
-%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
-%
-% rapid description of the flow
-Turbulent eddies generated in the near-wake of the grid are carried downstream.
-They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
-The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
-%
-%Finally, the obstacle does not deform or move either. 
-%
-%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
-%
-
-% lattice Boltzmann method 
-%
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
-While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
-%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
-Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
-Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
-In our context, this numerical method has been chosen principally for its computational efficiency.
-
-% geometry
-%
-The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points.
-The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel.
-The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
-% boundary conditions
-%
-No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
-%
-Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
-The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
-In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$.
-The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
-Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
-Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
-The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}.
-For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.
-
-\subsection{The drag force}
-\label{sec:drag_force}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
-		The vorticity is given in lattice units.}
-	\label{fig:typical_vorticity}
-\end{figure}
-
-%% LB parameters  %
-
-% drag signal
-%
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
-The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
-\begin{equation}
-\label{eq:drag_definition}
-f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
-\end{equation}
-where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm}). 
-Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
-% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
-%
-Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
-As a consequence, the drag can eventually be expressed as the difference 
-\begin{equation}
-\label{eq:drag_approx}
-f_d(t) = p_{fb}(t) - p_{base}(t)
-\end{equation}
-between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
-Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
-%
-% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
-%
-
-% Definition of turnover time
-%
-The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
-\begin{equation}
-\label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
-\end{equation}
-where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
-%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
-%
-% typical fluctuation of the drag
-%
-Fig.~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle for typical drag fluctuations.
-%\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
-
-\subsection{The drag as a random process
-	%: Probability Density Function and correlation in time
-}
-\label{sec:pdfs}
-
-%\ZZ{As requested by referee 3, Fig.3 has been removed}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
-%	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
-%	\label{fig:typical_drag_signal}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
-	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
-If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
-The selected peak fluctuations are therefore well separated.}
-		\label{fig:illustrate_return_time}
-\end{figure}
-
-% drag signal
-%
-% Description du signal de trainee typique
-In Fig.~\ref{fig:illustrate_return_time},
-the time signal of the drag acting on the square, $f_d(t)$, appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
-Therefore, it is natural to model the drag as a (scalar) random process.
-
-% drag statistics
-% pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
-This long simulation will be referred to as the \textit{control run} in the following.
-It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
-The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
-%
-Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
-%
-In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails.
-This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
-\begin{figure}
-	\centering
-	\subfloat[\ac{pdf} of normalised drag fluctuations]
-	{\label{fig:pdf_drag_a}
-		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
-	\subfloat[Autocorrelation of drag fluctuations]
-	{\label{fig:pdf_drag_b}
-		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
-	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation.
-The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle.
-%Note that the amplitudes have not been normalized.
-		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
-		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
-	}
-	\label{fig:pdf_drag}
-\end{figure}
-
-% drag statistics
-% correlation time
-Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
-%
-This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
-%
-In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
-
-\section{Extreme fluctuations of the drag by means of direct sampling}
-\label{sec:direct_sampling}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-	}
-	\label{fig:return_time_instant}
-\end{figure}
-
-% return time given from direct sampling
-%
-The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
-Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
-It will provide a trustworthy baseline for the validation of rare-events algorithms.
-
-The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
-As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
-For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.
-
-How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
-We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
-%
-%
-%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
-If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
-% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
-one might expect that 
-\begin{equation}
-  \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
-\end{equation}
-where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
-Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
-Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$.
-Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.
-
-% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.
-
-\subsection{Extracting extreme drag fluctuations from a very long timeseries}
-\label{sec:extreme_extraction}
-
-% intro
-%
-We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
-This set will be considered as representative of \emph{extreme events} in the upcoming study.
-The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
-%
-According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
-Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified.
-Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
-%
-In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
-
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
-\label{sec:instantaneous_drag}
-
-\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
-\label{sec:forebody_and_base_contribution}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
-	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag fluctuation $\tilde f_d$.
-An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
-		%\sout{As the amplitude of the drag increases, the relative variation of the base pressure also increases slightly.
-The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
-	}
-\end{figure}
-
-% teasing
-%
-%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
-We shall see that the situation is different in the case of {extreme} drag fluctuations.}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
-We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
-%
-% relative contribution of forebody and base pressure
-%
-Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
-The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
-\begin{equation}
-\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
-\end{equation}
-where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
-%
-Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
-	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
-
-%\sout{It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
-Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
-Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.}
-
-\subsubsection{Fluid dynamics related to extreme drag fluctuations}
-\label{sec:dynamical_aspects}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
-	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$.
-The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation.
-Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$.
-The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
-\end{figure}
-
-% mean profile around burst
-%
-The focus is now on the flow scenarii that yield extreme values of the drag. 
-Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events.
-A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$.
-This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
-%Starting from typical values, extreme drags are typically reached in less than a correlation time. 
-Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-%This is reminiscent of time-irreversibility in turbulent dynamics. 
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
-%Moreover, such very high drag levels do not persist over time.
-To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
-	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
-	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$.
-In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
-	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
-\end{figure}
-% flow scenario
-%
-Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
-%
-In each case, an intense vortical structure is visible near the base of the obstacle.
-The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
-%+
-The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle. 
-%
-This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
-In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.
-
-The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base. 
-As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
-% 
-This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown. 
-%
-Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
-As the blocking vortex is in turn advected downstream, the vortex against the base is released.
-%
-Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
-This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
-This scenario is generic and has been observed for most extreme events sampled in the control run.
-
-%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
-Each trajectory corresponds to a single event.
-The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
-\end{figure}
-
-Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
-\begin{equation}
-\label{eq:avg_shear_def}
-\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
-\end{equation}
-where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
-%
-% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
-% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
-%
-Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
-Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
-On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
-%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
-These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
-This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
-
-\subsection{Extreme fluctuations of the time-averaged drag }
-\label{sec:time_avg}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
-	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and the averaged drag, respectively.}
-	\label{fig:extreme_avg}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-	\caption{\ac{pdf} of the time-averaged drag $F_T$ with $T = 10\tau_c$. 
-		%This estimate is based on a moving average over a timeseries of the instantaneous drag $f_d$ of duration $4 \times 10^6 T_0$.
-%				\QQ{x-axis should be $x = F_T$ and y-axis $P(x)$ in agreement with Fig. 4(a). Gaussian PDF should be replaced by Gaussian estimate for the same reason. No need to mention PDF of $F_T$.}
-	}
-	\label{fig:PDF_AVG}
-\end{figure}
-
-% motivation
-%
-%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
-In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.
-
-%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
-We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
-%
-In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
-% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
-%
-% definition
-%
-Therefore, another relevant observable is the \textit{time-averaged} drag defined as
-\begin{equation}
-\label{eq:def_time_averaged_drag}
-F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
-\end{equation}
-where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
-%\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
-
-%
-During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.  
-% phenomenology
-%
-% What is the phenomenology leading to extreme values of $F_T(t)$ ?
-It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
-%
-%
-%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$. 
-In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
-Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
-
-% neither case (1) nor case (2)
-%
-Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
-
-%\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}
-
-% theoretical explanation is removed because it is speculative
-
-%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations. 
-%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
-%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes   
-%\begin{equation}
-%\label{eq:indep}
-%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)  
-%\end{equation}
-%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
-%\begin{equation}
-%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
-%\label{eq:ratio_exp}
-%\end{equation}
-%The observation is therefore well supported theoretically.
-%See~\cite{lestang2019JSTAT} for further details.
-
-%\ZZ{ The next section has been strongly reduced by keeping only key results accessible to broad JFM audience. Technical details are either removed or moved to Appendix}
-
-\section{Rare-event sampling algorithms}
-\label{sec:rare_events_algorithms}
-
-% limitation of direct sampling
-%
-In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive.
-Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{\ell a}~(\ell>0)$ according to Eq.~\eqref{eq:return_time}.
-As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{\ell a}$.
-%% objective
-% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
-%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
-%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study.
-To what extent they are representative of rare-event sampling algorithms?
-Complementarity of the two algorithms?
-Make the link between fluid mechanics and statistical algorithms ?}
-%
-Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
-At each step of the algorithm, some trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations.
-The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, allowing the estimation of the statistical bias and therefore the inference of the statistics of extreme events.
-
-Different algorithms use different \emph{selection rules}. 
-The success of the algorithm essentially depends on the quality of the selection rule to detect the precursors of the extreme event \citep{rolland_statistical_2015}. 
-However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
-In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
-
-%
-%\ZZ{paragraph moved from introduction}
-The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
-In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
-or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
-%
-% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}.
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
-%
-
-During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \citep{touchette_large_2009}.
-Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
-%
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
-This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is an unexplored area of application.}
-
-% These two algorithms will be detailed in the following.
-% In section~\ref{sec:ams}, we highlight that sampling  extreme fluctuations of the instantaneous drag with the \ac{ams} algorithm is a difficult task.
-% In section~\ref{sec:gktl}, we apply an importance sampling algorithm, the \ac{gktl} algorithm, and efficiently sample rare trajectories corresponding to extreme fluctuations of the time-averaged drag.
-% As an application, we use the \ac{gktl} algorithm to compute return times of very rare drag fluctuations.
-
-\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
-\label{sec:ams}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the selection-mutation procedure in the \ac{ams} algorithm with $N=3$.} Lines 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from the branching point to the final time $T_a$.
-In case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-Depending on the application, this re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.}
-%\end{figure}
-
-% Introduction of AMS
-%
-%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
-\EL{The \ac{ams} algorithm is here used with,  %a sequence of events that can be sampled with a higher probability  
-trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
-%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
-%
-% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
-%
-% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
-%
-% Quick description of TAMS
-%
-At each iteration, trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
-%The discarded trajectories are then re-sampled according to the remaining trajectories.
-These trajectories are re-sampled by branching them from the remaining trajectories.
-%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
-
-%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$.
-However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}.
-Our results are not limited to this particular variant.
-
-%
-% SUBSECTION IS MOVED TO APPENDIX %
-%
-%\subsubsection{}
-%\label{sec:ams_ou}
-%% Example: OU process
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%%
-%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-%%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-
-%\subsubsection{The \ac{ams} for extreme drag fluctuations}
-%\label{sec:ams_drag}
-
-%
-% Application to instantaneous drag
-%
-% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
-% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
-%
-%
-%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}
-
-The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
-%\sout{In contrast with the \acl{ou} process, } 
-%The phase space is highly-dimensional with deterministic chaotic dynamics.
-%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
-%\sout{is no longer straightforward} 
-%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
-%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
-The observable itself is used as the score function, \emph{i.e.} $\xi (\mathbf{x}(t),t) = f_d(t)$.  
-%
-Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
-%In practice, a small perturbation in the state of the flow has been introduced 
-%at branching points.
-% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
-	%
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
-We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
-%\EL{Freddy: provide reference for Lyapunov timescale.
-\#12}
-
-% application
-%
-%\EL{In the following,} the \ac{ams} algorithm has been applied to $N=256$ trajectories of duration $T_a = 20\tau_c$.
-%%
-%Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
-%% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%%
-%%
-%%The distribution of the maximal drag for the \textit{initial} trajectories is also shown.
-%%After a few iterations, the trajectories with the lowest maxima of the score function are discarded and new trajectories with higher maxima are re-sampled.
-%%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-%The re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-%%Note that the red points in figure~\ref{fig:AMS_drag_resampling} are analogous to the red line in Fig.~\ref{fig:comparaison_temps_de_retour} for the \ac{ou} process.
-
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
-  The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
-Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_trajectories/AMS_drag_trajectories.png}
-	\caption{\label{fig:AMS_drag_trajectories} Ensemble of $N = 32$ trajectories after $181$ iterations of the \ac{ams} algorithm.
-In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.
-		%The \ac{ams} fails to efficiently sample rare trajectories. All trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
-The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
-\end{figure}
-
-%pitfall
-%
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
-%
-Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble, and overlap over most of their duration.
-%
-The algorithm thus fails to enhance the sampling of extreme events. 
-This result is also confirmed by increasing the number of initial trajectories to $N=256$.
-%
-The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
-%
-% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
-%
-%
-the distribution of the maximal drag for the {initial} trajectories is also shown.
-We observe that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
-%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
-However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
-%
-
-It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
-%
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
-After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
-%
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
-%
-%
-Therefore, the re-sampling of a trajectory branched close to $t=t^{\star}$ (when the maximum drag occurs) cannot lead to larger values at $t^{\star} \leq t \leq t^{\star}+\tau_L$.
-For $t - t^{\star} >\tau_L$, the drag process has lost the memory of the drag fluctuations on which the re-sampling was based and, thus, the probability of observing a new extreme fluctuation is also very low.
-%
-%
-
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
-%
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
-The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.
-
-%As discussed previously, extreme fluctuations of the drag $f_d$ have a lifetime of length roughly $\tau_c$.
-Fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$, shorter than the time it takes for trajectories to separate, \textit{i.e} the Lyapunov timescale.
-
-% perspectives
-%
-In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
-%
-%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
-%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
-%
-%
-%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
-%
-%Furthermore, such ``tuning'' is problem dependent.
-%
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
-%Nevertheless, a possible line of improvement could  be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached.
-That way, new events associated to higher fluctuations may have the opportunity to develop.
-However, such modification of the procedure should be implemented with care especially to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.
-%
-
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
-\label{sec:gktl}
-The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
-The \ac{ams} algorithm could be used in the same way as before
-%in section~\ref{sec:ams_drag} \textit{i.e.} 
-by taking the time-averaged observable itself as the score function.
-However, this would lead to similar unsatisfactory results.
-% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
-%
-For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
-%
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
-% In this case, however, the trajectories interact dynamically. 
-%
-At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
-%
-The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
-%
-The \ac{gktl} algorithm belongs to a family of algorithms known as ``go-with-the-winners'' \citep{aldous1994go,grassberger2002go}.
-{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
-The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \citep{giardina_direct_2006}.
-Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
-
-%\bigskip
-%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
-%\bigskip
-
-%\subsubsection{The \ac{gktl} algorithm}
-%\label{sec:gktl_description}
-%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%%
-%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%%
-%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%%
-%\begin{equation}
-%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-%\label{eq:Weight}
-%\end{equation}
-%so that $\sum_{n=1}^N W_n^i = N$.
-%%
-%%
-%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.
-%The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%%
-%The sampled biased distribution writes
-%%
-%\begin{align}
-%\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-%\label{eq:Biased_Path_Approximation}
-%\end{align}
-%where
-%$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-%refers formally to the probability of observing the trajectory 
-%$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-%The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%%
-%One can mention that
-%\begin{equation}
-%\label{eq:mean_field}
-%Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-%\end{equation}
-%with $\mathbb{E}_{0}$ being the expectation value with respect to the
-%distribution $\mathbb{P}_{0}$.
-%This result relies on the \textit{mean-field approximation}
-%\begin{equation}
-%R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-%\label{eq:Mean_Field_Approximation}
-%\end{equation}
-%where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-%The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%%
-%Rejected trajectories are discarded from the statistics.
-%Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-%
-%A key feature of the \ac{gktl} algorithm is that the sampling procedure does not involve any alteration of the dynamics. \EL{All resampled trajectories are solutions of the original dynamical.} 
-%%This is true for stochastic dynamics.
-%\EL{Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-%	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%%
-%% 
-%%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%%
-%%
-%Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%%
-
-%\subsubsection{The \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
-%\label{sec:gktl_drag}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL_2}
-	\caption{\label{fig:IS_GKTL} Rare-event sampling of the (zero-mean) time-averaged drag $\tilde F_T = F_T - \overline{F_T}$ with $T=10\tau_c$; $\tau_c$ is the correlation time of the instantaneous drag.
-The shaded \ac{pdf}s are estimated from the biased ensemble resulting from the \ac{gktl} algorithm applied to $N=1024$ trajectories of duration $T_a=T$ with the cloning period $\tau=\tau_c/2$.
-		The dashed line refers to the unbiased \ac{pdf} of $\tilde F_T$, \textit{i.e.} obtained from direct sampling or with $k=0$ (no bias) in the \ac{gktl} algorithm.
-	}
-\end{figure}
-
-The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
-%Referring to the notations used in Appendix \ref{app:gktl_description}, 
-Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
-%Referring to the previous notations, $A(\mathbf{x}(t))=f_d(t)$ is our observable of interest and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
-%
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
-Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
-This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$.
-As a result, the resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.
-Further details are provided in Appendix~\ref{app:gktl_description}.
-}
-
-%
-% Choice of parameters and perturbation
-
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
-The higher $k$, the larger is the bias involved in the statistical resampling.
-%, as illustrated in Fig.~\ref{fig:IS_GKTL}. 
-Similar to the \ac{ams} algorithm, $N$ governs the error affecting the averaged quantities evaluated from the biased ensemble of trajectories, and should be chosen as high as possible.
-The cloning period $\tau$ determines how often the resampling is performed.
-A small cloning period can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
-
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
-%
-%In the following experiments, we fixed $\tau = \tau_c /2$ as a satisfactory trade-off \citep{lestang:tel-01974316}.}
-%The computational cost $C_{{gktl}}$ depends on both the duration of each trajectory and the number of trajectories: $C_{gktl} = N \times T_a$.
-%We have taken $N=1024$ and $T_a = 10\tau_c$ which yields $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
-%
-%The choice of the cloning period $\tau$ is important in practice.
-%A  cloning period too-short can result in a loss of information if the clones do not separate from their parents between two cloning steps.
-On the contrary, choosing $\tau \gg \tau_c$ may result in insufficient cloning steps to allow for efficient importance sampling.
-%As a result, a safe rule of thumb is to take $\tau$ of the order of $\tau_c$.
-%In the following experiments, $\tau = \tau_c /2$. 
-%
-%
-Three numerical experiments corresponding to three different values of the strength parameter $k$ have been carried out.
-%
-%
-%Let us remark that the value of $k$ should be chosen empirically unless a \textit{large deviation regime} $T_a \to \infty$ is verified  \cite{touchette_large_2009}.
-%In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem \cite{touchette_large_2009}.
-%
-%
-%Fig.~\ref{fig:IS_GKTL} shows the estimate of the biased \ac{pdf}s of the (zero-mean) time-averaged drag $\tilde F_T \equiv F_T - \overline{F_T}$, \emph{i.e}
-%\begin{equation}
-%\label{eq:estimate_biased_measure}
-%\rho_k(F) 
-%%= \mathbb{E}_{\mathbb{P}_k}[\delta(\tilde F_T(\{\mathbf{x}\}_{0\leq t \leq T_a})-F)]
-%\approx \frac{1}{N}\sum_{j=1}^{N}\delta(\tilde F_{T}(\{\mathbf{x}_j\}_{0\leq t \leq T_a}) - F)
-%\end{equation}
-%in addition to the unbiased \ac{pdf} based on the control run.
-%As expected, the biased PDFs obtained by applying the \ac{gktl} algorithm are centered around increasing fluctuations with the parameter $k$.
-%%
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
-%
-%
-% limitations
-%
-Nevertheless, it should be mentioned that for a given number $N$ of trajectories, there is necessarily an upper limit $k_{max}$ of the strength parameter over which the finite number of trajectories becomes detrimental to the efficiency and accuracy of the selection procedure.
-%
-For $k \gtrsim k_{max}$, the re-sampling relies only on a small number of ``independent trajectories'' and most of the trajectories in the biased ensemble overlap. 
-This effect is highlighted in Fig.~\ref{fig:IS_GKTL} where the histogram corresponding to $k=0.03$ becomes artificially peaked.
-%
-%Such a finite-size effect is expected to gradually increase as the bias amplitude $k$ is increased.
-%
-%An important question concerning the application of the \ac{gktl} algorithm is therefore that of the dependence of $k_{max}$ on the ensemble size $N$. 
-%Let us mention that this issue has been addressed in recent studies \citep{nemoto2017finite,Guevara_Hidalgo_2018}.
-%
-% how to assess the value of k_max ?
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-%This diversity has been monitored at each selection step of the algorithm by calculating the proportion of trajectories sharing the same ancestor trajectory.
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{|l|c|c|c|c|c|}
-%		\hline
-%		& $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		\hline
-%		$k=0.02$ & ~1594~ & ~799~ & ~155~ & 22 & 0  \\ 
-%		\hline
-%		$k=0.025~$ & 1019 & 834 & 521 & 198 & 27  \\ 
-%		\hline
-%		$k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
-%		\hline
-%		$k=0$ & 2599 & 37 & 22 & ~0.5~ & ~0.005~ \\
-%		\hline
-%	\end{tabular}
-%	\caption{Number of fluctuations $F_{T} \geq a$ with a=$\sigma,~2\sigma,$ etc. captured in the biased ensemble for $k=0.02,~0.025$ and $0.03$.
-%		Note that trajectories that overlap for more than half of their duration are counted as only one.
-%		$k=0$ yields the number of fluctuations expected from the direct sampling of $N$ independent realizations of $F_{T}$.}
-%	\label{tbl:nbTraj_tbl}
-%\end{table}
-%
-%% Comparisons with direct sampling
-%%
-%The efficiency of the \ac{gktl} algorithm is  assessed in  
-%Table~\ref{tbl:nbTraj_tbl} by indicating the number of trajectories that yield a time-averaged drag $\tilde{F}_{T} \geq a$ for different values of the bias parameter $k$.
-%Note that trajectories that overlap for more than half of their duration are counted as one.
-%%\footnote{Trajectories that overlap over an interval $[O;t]$ with $t\geq T_a / 2$.}
-%Table~\ref{tbl:nbTraj_tbl} also indicates the expected number of fluctuations obtained by simulating $N$ independent trajectories $\{F_T(t)\}_{0\leq t \leq T_a}$ without resampling.
-%%
-%%This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed. This hypothesis is motivated by Fig.~\ref{fig:PDF_AVG}.
-%%
-%%\begin{figure}
-%%  \centering
-%%  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
-%%  \caption{PDF of average drag over $10\tau_c$}
-%%  \label{fig:PDF_AVG}
-%%\end{figure}
-%%
-%% Conclusion
-%%
-%We obtain that the \ac{gktl} algorithm is able to sample drag fluctuations that are far from reach by direct sampling with an equivalent computational cost.
-
-%
-% PARAGRAPHE SUIVANT INUTILE ?
-%
-%In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
-By selecting trajectories based on the previous average value of the drag over a duration $\tau \approx \tau_c$, the \ac{gktl} algorithm is well suited for the preferential selection of trajectories of the former kind.
-However, such selection strategy does not appear equally as favourable to trajectories of the latter type, for which all of the drag density is located in one or two very large fluctuations of the instantaneous drag.
-Further work must now investigate the nature of the sampled trajectories, and determine the amount of discrepancy concerning their respective phenomenology.
-
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
-%It illustrates that the cloning algorithm samples trajectories that exhibit successive positive fluctuations, resulting in a very large value of the average drag.
-Note that during the post-processing of the sampled trajectories, trajectories overlapping over more than half of their duration ($5 \tau_c$) have been considered duplicates, and treated as effectively the same trajectory.
-% This procedure resulted in an ensemble of ?? "distinct" trajectories (out of the original $N = 1024$).
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{timeseries_extrms_AVG_GKTL/timeseries_extrms_AVG_GKTL}
-	\caption{Drag timeseries corresponding to the highest fluctuations of the averaged drag in the ensemble of trajectories sampled  by the \ac{gktl} algorithm.
-The algorithm was applied with $N = 1024$, $\tau = \tau_c / 2$ and $k = 0.03$.
-}
-	\label{fig:timeseries_extrms_AVG_GKTL}
-\end{figure}
-In addition, Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} displays the vorticity field related to the maximum of the drag in sampled trajectories.
-One observes that the extreme events are consistent with the picture of strong vorticity being trapped in the vicinity of the base of the obstacle, as pointed out in section~\ref{sec:instantaneous_drag}.
-Qualitatively, the related drag signals are also found very similar to those obtained from direct sampling.
-However, it should be stressed that the computational cost for sampling the events shown in Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} and Fig.~\ref{fig:illustr_extrms_vorticity_GKTL} is roughly one hundred time lower than the computational cost required to capture by direct sampling the events displayed in  Fig.~\ref{fig:top_4_events_vorticity} and Fig.~\ref{fig:extreme_avg}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_extrms_vorticity_GKTL/illustr_extrms_vorticity_GKTL}
-	\caption{Vorticity field associated with the maximum of the instantaneous drag $f_d$ in  trajectories sampled by the \ac{gktl} algorithm.}
-	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}
-
-%An important remaining question is whether or not the sampled ensemble of trajectories are representative of the full range of dynamics that can lead to extreme values of the average drag.
-%Indeed, it was observed in section~\ref{sec:time_avg} that large values of the average drag over $10\tau_c$ can result from a variety of different
-%dynamics, including the succession of independent positive fluctuations, but also the occurrence of fewer and larger fluctuations separated by periods of typical drag.
-%In particular, it is not clear from the results of the present work whether or not the cloning algorithm is suited to sample trajectories of the latter form.
-%A possible continuation of this study is to go beyond the qualitative description of the events sampled by the cloning algorithm, a task that is left to future work.
-}
-
-\subsubsection{Computation of return times}
-\label{sec:return_times}
-
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
-In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
-
-%Statistics of extreme events can be computed from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}. For instance, it is possible to compute return times for
-% events unreachable by a direct approach.
-%(see Fig.~\ref{fig:illustrate_return_time})
-
-%Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-%\begin{equation}
-%\label{eq:time_averaged}
-%F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-%\end{equation}
-%and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-%\begin{equation}
-%r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-%\end{equation}
-%%
-%The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-%\begin{equation}
-%\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-%\end{equation}
-%with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-%
-
-Fig.~\ref{fig:return_times_gktl} displays the return times for extreme fluctuations of the time-averaged drag obtained by using the \ac{gktl} algorithm with two different values of the strength parameter $k$.
-%
-Both estimates have been obtained with the same computational cost $N\times T_a$.
-An estimate given by direct sampling with a time-series of the same effective duration $T_{tot}=N\times T_a$ is also shown. 
-Whilst a direct approach cannot access (by definition) events with a return time greater than $T_{tot}$, the \ac{gktl} algorithm allows us to estimate the statistics of time-averaged drag fluctuations having a return time several orders of magnitude above $T_{tot}$.
-Alternatively, for a fixed target return time, the use of the \ac{gktl} algorithm can reduce the computational cost of the estimation by several orders of magnitude.
-This is  obviously a major advantage of this rare-event sampling algorithm.
-%
-%\EL{More generally, the successful estimation of return times indicates that, despite the complex dynamics and the overlapping trajectories, the biased ensemble generated by the \ac{gktl} algorithm may be used to compute statistical estimators in a very efficient manner.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.7\linewidth]{return_times_GKTL/return_times_GKTL}
-	\caption{\label{fig:return_times_gktl} Return times for the time-averaged drag acting on the square obstacle.
-$\tilde{F}_T$ denotes the time-averaged drag with zero mean.
-The blue and red lines are obtained from the biased ensemble of trajectories generated by the \ac{gktl} algorithm with $N=1024$, $T_a=30\tau_c$ and a cloning period $\tau=\tau_c/2$.
-The green line is the return times obtained from a single timeseries of duration equal to the computational cost of both \ac{gktl} experiments.
-Uncertainty ranges for the \ac{gktl} estimates are computed as the standard deviation over a set of 10 independent experiments.
-Uncertainty ranges for the direct estimation are computed as the standard deviation over a ensemble of direct estimates resulting from 60 independent timeseries.}
-\end{figure}
-
-\section{Conclusion}
-\label{conlusion}
-
-%In this study, the application of two rare-event sampling algorithms has been assessed for the simulation of extreme mechanical efforts on structures immersed in a turbulent flow; a situation relevant to many industrial applications.
-
-%The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-%By means of direct sampling throughout a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle.
-%This arrangement does not persist over time since the main stream eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around  extreme fluctuations.
-%%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%Concerning the time-averaged drag, it was found that extreme fluctuations do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-%%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-%
-%A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly accelerate the sampling of extreme fluctuations.
-%%
-%Interestingly, the \ac{ams} and \ac{gktl} algorithms rely on different selection rules and have different fortunes.
-%%
-%On the one hand, the \ac{ams} algorithm fails to generate trajectories exhibiting extreme events at a better rate than a direct sampling, which can be related to the phenomenology of extreme drag fluctuations, \EL{whose lifetime} is shorter than the timescale over which the replicated trajectories manage to separate from their parent.
-Therefore, the algorithm is unable to capitalize on the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories.
-%%
-%On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-However, this latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as in the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations, which can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-%
-%%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%%
-%Selection rules entering in rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-%Here, the observable itself was chosen as score function, as a safe fallback solution.
-%Optimizing the choice of this score function is certainly desirable but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-%Our study shows this quite clearly. A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.
-
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
-By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
-This arrangement does not persist over time, however, since the main flow eventually sweeps away the surrounding fluid structures responsible for this trapping.
-Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-%Our long drag timeseries also reveal that the tails of the \ac{pdf} for the drag are well described by an exponential \ac{pdf}.
-%This property is linked to the phenomenology of extremes of the time-averaged drag.
-%Especially, we observed that such extreme values for the average do not preferentially result from a small number of very large fluctuations, or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
-% but can be either formed by one of these two cases or by a combination of both.
-%Such a phenomenology is observed for stochastic processes with an exponential \ac{pdf}.
-
-A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
-These algorithms rely on selection rules that determine how an initial ensemble of trajectories is evolved to possibly enhance the sampling of extreme fluctuations.
-%
-On the one hand, the \ac{ams} algorithm fails (for this specific application) to generate trajectories exhibiting extreme events at a better rate than a direct sampling.
-This result can be related to the phenomenology of extreme drag fluctuations, whose lifetime is shorter than the timescale over which the replicated trajectories manage to separate from their duplicate.
-Therefore, the algorithm is unable to benefit from the precursors of extreme fluctuations to enhance the realization of new rare-event trajectories. 
-%
-On the other hand, the \ac{gktl} algorithm leads to a computational gain of several orders of magnitude.
-The latter is based on the cumulative evolution of the system rather than its instantaneous behaviour, as opposed to the \ac{ams} algorithm.
-The counterpart is that the algorithm provides only  statistics of extreme values of the time-averaged fluctuations.
-This can nevertheless be of great interest for computing return times of extreme (time-averaged) fluctuations, for instance.
-
-%Our conclusions on the applicability of rare-event algorithms to turbulent flows and fluid-structure interaction are mitigate, and obviously calls for further investigation and deeper understanding.
-%
-Selection rules at the heart of rare-event algorithms rely heavily on the choice of a score function that drives the selection and replication of trajectories. 
-In this study, the observable itself (the drag) has been chosen as the score function.
-Optimizing the choice of this score function is desirable (especially for the \ac{ams} algorithm) but difficult in practice, since it should account for the phenomenology of the extreme events themselves, how these events build up.
-Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
-
-\section{Acknowledgements}
-The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
-T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
-Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
-The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
-
-\appendix
-\section{The \acl{lbm}}
-\label{app:lbm}
-
-% details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
-In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
-Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
-\[
-\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
-\]
-for the mass density and the fluid momentum respectively.
-The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.
-
-% algo
-%
-The complexity of the flow emerges from the repeated application of simple rules of streaming and collision.
-The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure.
-Namely, an \emph{exact} streaming step
-\[
-f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
-\]
-during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
-\[
-f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
-\]
-which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level).
-The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by
-\[
-\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
-\]
-This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
-%
-The equilibrium function is given by
-\[
-f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
-\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
-\] 
-with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
-This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e.
-$\Delta x \to 0$ \EL{\citep{succi_book}}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\linewidth]{D2Q9/D2Q9}
-	\caption{Sketch of the D2Q9 lattice.
-Particles move exactly from a lattice node towards one of its nine neighbours (including the node itself) during one time step.
-By definition, the lattice spacing is related to the time step by $\Delta x/ \Delta t = \sqrt{3} c_s$ where $c_s$ is interpreted as a speed of sound.}
-	\label{fig:D2Q9}
-\end{figure}
-
-As mentioned before, the pressure is directly accessible from the mass density: $p = \rho c_s^2$. The viscous stress is also obtained easily from the densities of particles by
-\[
-\tau^\mathrm{visc.}_{\alpha \beta} = -\frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\]
-so that the total stress expresses as
-\begin{equation}\label{eq:def_stress}
-\tau_{\alpha \beta} = -  c_s^2 \sum_i f_i ~ \delta_{\alpha\beta}  - \frac{\nu}{\tau_\nu ~ c_s^2 \Delta t} \sum_i  {c_i}_\alpha {c_i}_\beta (f_i - f_i^\mathrm{eq})
-\end{equation}
-Finally, let us mention that in the present context of turbulent flows, the single-relaxation-time BGK collision has been replaced by a multi-relaxation-time procedure based on central moments with an improved stability \citep{De_Rosis_2016}.
-
-%
-%
-%\newtheorem{theo}{Theorem}
-%\section{The \ac{ams} algorithm -- nouvelle version Thibault}
-%\label{app:ams}
-%
-%\QQ{Pour moi, le texte de Thibault ci-dessous est beaucoup trop détaillé (théorème, etc.) et absolument inutile pour les lecteurs de JFM !
-Je proposerais de ne garder que le schéma de principe et l'application à Ornstein-Ulhenbeck (comme dans la version déjà soumise).
-Les referees n'ont pas demandé de détailler les algos.}
-%
-%
-%This appendix provides a brief overview of the \ac{ams} algorithm, a brief summary of main
-%mathematical results and discusses its use in practice.
-%
-%\subsection{Description of the algorithm}
-%An ensemble of $N$ independent trajectories (with fixed duration $T_a$) $\{x_n^{(0)}(t)\}_{1\leq n \leq N}$ is considered.
-%%These trajectories could be generated by any continuous time Markov model: a stochastic process, for instance a diffusion, or a chaotic deterministic dynamical system.
-%%
-%From this initial ensemble, the \ac{ams} consists in a series of iterations in which one or more trajectories are discarded, and re-simulated from a given point in time $t \in [0, T_a]$ to the final time $t = T_a$.
-%
-%To start off with, each of these trajectories is associated a weight $w_0=1$.
-%Then, at iteration $j\geq 1$, we evaluate the score of all trajectories $\{x_n^{(j-1)}(t)\}_{1 \leq n \leq N}$ at iteration $j-1$, measured by the maximum of the score function $\xi$ over the whole trajectory:
-%\begin{equation}
-%\mathcal{Q}_n^{(j)} = \sup_{0 \leq t \leq T_a} \xi(t,x_n^{(j-1)}(t)).
-%\end{equation}
-%Then, the trajectories corresponding to the lowest $\mathcal{Q}_n^{(j)}$ are selected: we denote $\mathcal{Q}_j^\star= \min_{1\leq n \leq N} \mathcal{Q}_n^{(j)}$ and $n_{j,1}^\star,\ldots,n_{j,\ell_j}^\star$ the indices of trajectories such that:
-%\begin{equation}
-%\label{eq:ams_threshold_def}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star.
-%\end{equation}
-%One might expect intuitively that $\ell_j=1$.
-%However, because of the discretization of the dynamical equations in the numerical model, two or more trajectories may yield the same level $\mathcal{Q}_n^{(j)}$, see~\cite{brehier:hal-01142704}.
-%The next step is known as the \textit{mutation step}: for each trajectory $x_{n_{j,\ell}^\star}^{(j-1)}$ ($1 \leq \ell \leq \ell_j$), we choose a trajectory $x_{n_\ell}^{(j-1)}$ ($n_{\ell} \neq n_{j,1},\ldots n_{j,\ell_j}$) randomly among the $N-\ell_j$ remaining trajectories, and define the time $t_{j,\ell}$ defined as the smallest time $t$ such that $\xi(t,x_{n_\ell}^{(j-1)}(t))>\mathcal{Q}_j^\star$.
-%Finally, we define the new replica $x_{n_{j,\ell}^\star}^{(j)}$ by copying the trajectory $x_{n_\ell}^{(j-1)}$ from $t_0$ to $t_{j,\ell}$, and simulating the rest of the trajectory, from $t_{j,\ell}$ to $T_a$.
-%For a Markov process, for instance a diffusion, a new realisation of the noise is used in order to simulate the new trajectory from $t_j$ to $T_a$.
-%For a chaotic deterministic system, a small amplitude noise is added to the initial condition at time $t_j$.
-%The other trajectories are not modified: $x_n^{(j)}=x_n^{(j-1)}$ for $n \neq n_{j,1}^\star,\ldots,n_{j,\ell}^\star$.
-%The selection-mutation process is illustrated on Fig.~\ref{fig:AMS_schema}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{illustr_AMS/illustr_AMS}
-%	\caption{Illustration of one iteration the \ac{ams} algorithm.
-Starting from an initial ensemble of 3 trajectories, trajectory 1 is selected as it corresponds to the lowest maximum of the score function $Q_1 = \mathcal{Q}_{j}^\star$.
-Out of the 2 other trajectories, trajectory 3 is chosen randomly (with probability 1/2) to serve as the base of the resampling of trajectory 1.
-Trajectory 1 is first made a copy of trajectory 3 from $t=0$ to $t_j$, then simulated from $t_j$ to $T_a$.}
-%	\label{fig:AMS_schema}
-%\end{figure}
-%
-%After $N$ iterations of this selection-mutation procedure, the number of resampled trajectories is
-%$\tilde{J} = \sum_{j=1}^J \ell_j$.
-%Ultimately, the algorithm generates $M=N+\tilde{J}$ trajectories, given explicitly by the set $\{x_n^{(0)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$, or equivalently, the set $\{x_n^{(J)}\}_{1 \leq n \leq N} \cup \{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$.
-%\subsection{How to compute the probability}
-%Trajectories $x_n^{(j)}$ forming the ensemble at step $j$ are attributed a weight $w_j$ given by~\cite{Cerou2007,Cerou2011,brehier:hal-01142704}:
-%\begin{equation}
-%w_j = \prod_{i=1}^j \left( 1 - \frac{\ell_i}{N}\right)=\left( 1 - \frac{\ell_j}{N}\right)w_{j-1}.
-%\end{equation}
-%This means that each trajectory in the sampled ensemble has an associated weight, given by the iteration until which it was a member of the ensemble: $w_J$ for the final trajectories $\{x_n^{(J)}\}_{1 \leq n \leq N}$, and $w_{j-1}$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j-1)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Alternatively, $w_0 = 1$ for the initial trajectories, and $w_j$ for the trajectories $\{ x_{n_{j,\ell}^\star}^{(j)}\}_{1 \leq \ell \leq \ell_j, 1 \leq j \leq J}$ resampled at iteration $1 \leq j \leq J$.
-%Let us relabel these trajectories and their associated weights as $\{(x_m,w_m)\}_{1 \leq m \leq M}$.
-%Normalising the weights with $W=\sum_{m=1}^M w_m$, we obtain the probabilities $p_m=w_m/W$ associated with the trajectories.
-%\subsection{Summary of mathematical results}
-%We now consider the final ensemble of trajectories sampled by the \ac{ams} algorithm, after a given number
-%of iterations $N$.
-%We call $a$ the lowest maximum among the trajectories in the ensemble, \textit{i.e.} the final value of the threshold $\mathcal{Q}^{\star}$ defined in Eq.~\eqref{eq:ams_threshold_def}.
-%The \ac{ams} provides an estimator of the probability that a the maximum of the score function over a trajectory of duration $T_a$ is above $a$, that is:
-%\begin{equation}
-%\label{eq:estimator_ams}
-%q(a) = \mathbb{P}\left[ \underset{0\leq t \leq T_a}{\max} \xi(\mathbf{x}, t) > a \right]
-%\end{equation}
-%In the following we denote this estimator $\hat{q}_N(a)$.
-%One of the main properties of the \ac{ams} algorithm is the following unbiasedness result, see~\cite{brehier:hal-01142704} for more general statements, and discussion on the influence of the time discretization of the Markov dynamics.
-%\begin{theo}
-%	For every $N$, for every score function $\xi$, $\hat{q}_N$ is an unbiased estimator of $q$:
-%	\begin{equation}
-%	\mathbb{E}[\hat{q}_N]=q.
-%	\end{equation}
-%\end{theo}
-%Thus only the statistical error $\text{Var}(\hat{q}_N)$ depends on the choice of $N$, and, more importantly, on the score function $\xi$; see~\cite{brehier:hal-01142704,rolland_statistical_2015} for extensive numerical simulations concerning the role of the score function.
-%
-%More precisely, it is proved in~\cite{Cerou2016}, that the estimator $\hat{q}_N$ satisfies a Central Limit Theorem,
-%\begin{equation}
-%\sqrt{N}\bigl(\hat{q}_N-q\bigr)\underset{N\to \infty}\to\mathcal{N}(0,\sigma^2(\xi,q)),
-%\end{equation}
-%with an asymptotic variance $\sigma^2(\xi,q)\in [-q^2\ln q,2q(1-q)]$.
-%The minimal variance $-q^2\ln q$ is obtained when choosing
-%\begin{equation}
-%\bar{\xi}(t,x;T_a,a)=\mathbb{P}_{x,t}\left\lbrack\underset{t\le s\le T_a}\max O[X,s] > a\right\rbrack,
-%\label{eq:time_dependent_committor}
-%\end{equation}
-%for all $(t,x)\in[0,T_a]\times \mathbb{R}^d$, where we denote $\mathbb{P}_{x,t}$ the probability over the process initialised at position $x$ at time $t$, and the threshold $a$ and trajectory duration $T_a$ are fixed parameters.
-%
-%
-%\subsection{How to use it in practice}
-%In practice, the optimal score function $\overline{\xi}$, also referred to as the \emph{committor}, is of course not known: it is the output of algorithm.
-%Indeed, $q(a) = \bar{\xi}(0, x_0;T_a,a)$, where $x_0$ denotes an initial condition drawn according to the stationary measure of the dynamics.
-%Therefore, a crucial point to implement the \ac{ams} is to choose a score function that provides a good approximation of the committor.
-%
-%Another parameter of the algorithm is the duration $T_a$.
-%This duration must naturally be chosen larger than the timescale of the rare events of interest, and in this work we use $T_a = 5\tau_c$, where $\tau_c$ is the typical duration of extreme drag fluctuations, see section~\ref{sec:instantaneous_drag}.
-%Furthermore, we (empirically) observed that the sampling does not benefit form much longer trajectories.
-%This can be explained as follows.
-%Considering an arbitrary iteration $j$ and calling $t_j$ is the resampling time (the time at which the parent trajectory reaches $\mathcal{Q}^{\star}$, the resampled trajectory will only be decorrelated from its parent
-%after a time $\tau_L > \tau_c$ (see section~\ref{sec:ams} and appendix~\ref{app:perturb_branching_time}).
-%As a result, the probability that it exhibits a fluctuation $\xi > \mathcal{Q}^{\star}$ is $\mathbb{P}_{x_0}[\underset{t_j+\tau_L \le t \le T_a}\max \xi(\mathbf{x},t) > a]$, where $\mathbb{P}_{x_0}$ denotes the probability given a random initial condition drawn according to the stationary measure of the dynamics.
-%This amounts to direct sampling.
-%
-%Finally, the \ac{ams} experiments presented in this work do not specify a specific number of iterations $J$, neither a target drag threshold $\mathcal{Q}^{\star}$.
-%Rather, iterations were carried out until all the trajectories in the ensemble overlap.
-%Ultimately, because of the discretization induced by the numerical model, all trajectories reach the exact same maximum, \textit{i.e.}
-%\begin{equation}
-%\mathcal{Q}_{n_{j,1}^\star}^{(j)} = \cdots = \mathcal{Q}_{n_{j,\ell_j}^\star}^{(j)} = \mathcal{Q}_j^\star. \quad \text{with} \quad l_j = N
-%\end{equation}
-%and the iterations stop.
-%This phenomenon is sometimes referred as the \textit{extinction} of the algorithm.
-%
-%\subsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
-%\label{app:AMS_on_OU}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-%	\label{fig:comparaison_temps_de_retour}
-%\end{figure}
-%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
-%\begin{equation}
-%\label{eq:ou}
-%\dot{x} = -x + \eta (t),
-%\end{equation}
-%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-%
-%
-%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
-%
-%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
-%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling.
-%In the latter, the typical computational cost is simply the return time $r(a)$.
-%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-%
-%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
-%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
-%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%\newpage
-
-\section{The \ac{ams} algorithm}
-\label{app:ams-short}
-
-\subsection{The operating principle}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{illustr_AMS/illustr_AMS}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
-\end{figure}
-
-The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
-%
-The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories.
-On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line).
-Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$.
-It is then simulated from this branching point to the final time $T_a$.
-In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
-
-\subsection{Application to a simple case: the \acl{ou} process}
-
-	\begin{figure}
-		\centering
-		\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
-The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
-The blue line is the analytical solution for the return time of amplitude $a$.}
-		\label{fig:comparaison_temps_de_retour}
-	\end{figure}
-	A one-dimensional \acl{ou} process is considered:
-	\begin{equation}
-	\label{eq:ou}
-	\dot{x} = -x + \eta (t),
-	\end{equation}
-	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
-	%This basic stochastic process will allow us to highlight differences with fluid dynamics.
-
-	The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
-	Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
-	Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
-	The score function is simply $x(t)$ and only one trajectory is re-sampled at each iteration.
-	%
-	%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
-	%
-	The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the number of re-sampled trajectories.
-	Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
-	In the latter, the typical computational cost is simply the return time $r(a)$.
-	One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
-	For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
-
-	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
-	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
-	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
-
-\section{The \acl{gktl} algorithm}
-\label{app:gktl_description}
-
-\subsection{The operating principle}
-\label{sec:gktl-operating-principle}
-
-The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
-%
-Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
-%
-At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
-%
-\begin{equation}
-W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
-\label{eq:Weight}
-\end{equation}
-so that $\sum_{n=1}^N W_n^i = N$.
-%
-%
-{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$.
-The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}).
-For more information about the practical implementation of the algorithm, the interested reader can refer to~\citep{brewer2018efficient, lestang:tel-01974316}}.
-The application of this re-sampling at each step $t_i$ eventually leads to a biased sampling in the trajectory space; the trajectories corresponding to extreme values of $\int_{0}^{T_a}A(\mathbf{x}_{n}(t))dt$ have a higher probability.
-%
-The sampled biased distribution writes
-%
-\begin{align}
-\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
-\label{eq:Biased_Path_Approximation}
-\end{align}
-where
-$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
-refers formally to the probability of observing the trajectory 
-$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$.
-The normalisation factor is given by $Z(k,T_a)=\prod_{i=1}^{T_a/\tau}R_i$.
-%
-One can mention that
-\begin{equation}
-\label{eq:mean_field}
-Z(k,T_a) \underset{N\to \infty}{\sim} \mathbb{E}_0\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right],
-\end{equation}
-with $\mathbb{E}_{0}$ being the expectation value with respect to the
-distribution $\mathbb{P}_{0}$.
-This result relies on the \textit{mean-field approximation}
-\begin{equation}
-R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_i)= \mathbb{E}_{i}\left[e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{X}(t))dt}\right],
-\label{eq:Mean_Field_Approximation}
-\end{equation}
-where $\mathbb{E}_{i}[.]$ denotes the expectation value with respect to the biased distribution $\mathbb{P}_k^{(i)}$ obtained after $i$ cloning steps.
-The typical relative error related to this approximation can be shown to be of order $1/\sqrt{N}$ for a family of rare-event algorithms including the \ac{gktl} algorithm~\citep{DelMoralBook,DelMoral2013}.
-%
-Rejected trajectories are discarded from the statistics.
-Eventually, an effective ensemble of $N$ trajectories of duration $T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
-
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
-%This is true for stochastic dynamics.
-Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
-%
-% 
-%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones.
-We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
-%
-%
-Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
-%
-
-\subsection{Computation of return times}
-\label{sec:return-time-gktl}
-
-Each trajectory in the biased ensemble results in a timeseries of the time-averaged drag
-\begin{equation}
-\label{eq:time_averaged}
-F_T^{(j)}(t) = \int_{t-T}^{t}f_d^{(j)}(\tau)d\tau, \quad t\in [T,T_a]  
-\end{equation}
-and the return time of a fluctuation $F_T \geq a$ is given by~\citep{lestang_computing_2018}
-\begin{equation}
-r(a) = - \frac{T_a - T}{\ln (1-\mathbb{P}(F_T \geq a))}.
-\end{equation}
-%
-The probability $\mathbb{P}(F_T \geq a)$ can be estimated from the biased ensemble by inverting Eq.~\eqref{eq:Biased_Path_Approximation}
-\begin{equation}
-\mathbb{P}(f_d \geq a) \approx \frac{1}{N}\sum_{j=1}^{N}e^{T_a \lambda(k)}e^{k T_a  F_T^{(j)}}s_j(a)
-\end{equation}
-with $s_j(a) = 1$ if $\max_{T\leq t \leq T_a}[F_T^{(j)}] \geq a$ and $s_j(a) = 0$ otherwise, \emph{i.e.} by summing the weights of the timeseries which maximum is larger than $a$.
-
-\EL{
-\section{Perturbation at branching time}
-\label{app:perturb_branching_time}
-%As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl},
-The application  of the \ac{ams} and \ac{gktl} algorithms to a \emph{deterministic} dynamical system requires a perturbation at initial time of the
-resampled trajectories. In the absence of such perturbation, the resampling would not yield new trajectories but exact copies of the original trajectories.
-
-%The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
-A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the Poisson equation.
-In the framework of Lattice Boltzmann simulations, the state of the system is described at a mesoscopic level by the particle densities $\{f_i(\mathbf{x},t)\}_{i=0\cdots8}$ (see appendix~\ref{app:lbm}).
-%As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.
-%
-Therefore, the perturbation of the solution at time $t_0$ applies directly to the $f_i$'s with 
-\begin{equation}
-f_{i}(\mathbf{x},t_0) \longrightarrow f_i(\mathbf{x},t_0) + \epsilon \sum_{n=1}^{N_s} \alpha_{n}f_{i}^{(n)}(\mathbf{x}),
-\label{eq:perturb_pop}
-\end{equation}
-where the $\alpha_n$'s are random numbers uniformly picked in the interval $[0,1]$, $\epsilon$ is the relative amplitude of the perturbation and $\{f_i^{(n)}\}_{n=1 \cdots N_s}$ is a set of arbitrary snapshots of the flow.
-%
-%with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the relative amplitude of the perturbation.
-The perturbed populations are then rescaled so that no mass is added in the system.
-% For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
-In other words, the perturbation is a  random linear combination of $N_s$ snapshots. %$\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
-Eventually, the perturbed densities are rescaled so that the mass of the system is preserved.
-%Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
-%However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=.7\linewidth]{effect_of_perturbation/effect_of_perturbation}
-%	\caption[Illustration of the perturbation of the initial condition]{\EL{Time signal of the drag acting on the obstacle for a reference simulation (blue) and the same simulation perturbed at time $t_0=0$.
-The time variable is normalized by the correlation time of the drag ($\tau_c$).}}
-%	\label{fig:effect_of_perturbation}
-%\end{figure}
-%
-In practice, we chose $\epsilon = 0.002$ and $N_s = 10$ to ensure that the perturbations remain sufficiently small, random and independent.
-%Fig.~\ref{fig:effect_of_perturbation} illustrates the effect of applying such perturbation to the drag acting on the square obstacle.
-We observe that the trajectory separates from its ancestor after a few $\tau_c$ (correlation time of the drag).
-%
-In order to check that this perturbation does not impact the statistics of drag fluctuations, a long simulation of duration $T_{tot} = 10^5 \tau_c$ has been performed with a periodic perturbation (with period $\tau_c / 2$) mimicking the perturbation of the clones in the algorithm (see section \ref{sec:rare_events_algorithms}).
-%
-%The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
-In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $ \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
-Fig.~\ref{fig:pdf_drag_with_perturbation} shows that the statistics of the perturbed and unperturbed simulations are equivalent.
-
-\begin{figure}
-  \centering
-\includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
-\label{fig:pdf_drag_with_perturbation}
-\end{figure}
-}
-
-\bibliographystyle{jfm}
-\bibliography{../biblio}
-
-\end{document}

### 2020-04-04 texfiles/section2.tex
Move body of paper into texfiles/ and rename main document to paper.tex

--- 2020-04-03 section2.tex
@@
 % 
-We believe that this system embeds the characteristic features that makes 
+We believe that this system embeds the characteristic features that make
 \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
@@
 	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
+The time has been normalised by the turnover time related to the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
 If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.

### 2020-04-04 paper.tex
Move body of paper into texfiles/ and rename main document to paper.tex
Remove \EL{} macros

--- 2020-04-03 draft_p2_jfm.tex
@@
 
-\definecolor{myred}{HTML}{b30000}
-\definecolor{mygreen}{HTML}{008000}
-\newcommand{\EL}[1]{{\color{myred}{#1}}}
-\newcommand{\TL}[1]{{\color{mygreen}{#1}}}
-
-\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
-\newcommand{\QQ}[1]{{\color{green}{#1}}}
-
 \title{
-\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
+Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
 }
@@
 \begin{abstract}
-	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
+	This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
 	%
@@
 	% 
-	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
+	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
       This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
@@
 
-\include{section1}
-\include{section2}
-\include{section3}
-\include{section4}
-\include{section5}
+\include{./texfiles/section1}
+\include{./texfiles/section2}
+\include{./texfiles/section3}
+\include{./texfiles/section4}
+\include{./texfiles/section5}
 
@@
 \appendix
-\include{appendixA}
-\include{appendixB}
-\include{appendixC}
-\include{appendixD}
+\include{./texfiles/appendixA}
+\include{./texfiles/appendixB}
+\include{./texfiles/appendixC}
+\include{./texfiles/appendixD}
 

### 2020-04-04 texfiles/appendixA.tex
Remove \EL{} macros

--- 2020-04-03 appendixA.tex
@@
 % details about LBM
-In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}.
+In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice (see \citep{kruger_lattice_2017} for a comprehensive introduction).
 In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).

### 2020-04-04 texfiles/appendixB.tex
Remove \EL{} macros

@@
 	\includegraphics[width=\linewidth]{fig21/fig21}
-	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$.
-$\xi(x(t))$ is a prescribed score function defined along each trajectory.
-	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
+	\caption{\label{fig:illustr_AMS} Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$. $\xi(x(t))$ is a prescribed score function defined along each trajectory. 
+	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$. }
 \end{figure}
@@
 In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories.
-This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.
+This re-sampling procedure can be iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.
 
@@
 		\includegraphics[width=.7\linewidth]{fig22/fig22}
-		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
+		\caption{Efficiency of the \ac{ams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}.
 The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$.
@@
 	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents.
-Such favourable features \EL{do not} persist in the case of fluid dynamics.
+Such favourable features do not persist in the case of fluid dynamics.

### 2020-04-04 texfiles/appendixC.tex
Remove \EL{} macros

--- 2020-04-03 appendixC.tex
@@
 
- \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamical. 
+ A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamics. 
 %This is true for stochastic dynamics.
 Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.
-	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
+	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.
 %

### 2020-04-04 texfiles/appendixD.tex
Remove \EL{} macros

--- 2020-04-03 appendixD.tex
@@
 \includegraphics[width=.7\linewidth]{fig23/fig23}
-\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
+\caption{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}
 \label{fig:pdf_drag_with_perturbation}

### 2020-04-04 texfiles/section1.tex
Remove \EL{} macros

@@
+buffer-name
 \section{Introduction}
@@
 ``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
+In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \citep{kanev2010}.
 %
@@
 However, this requires very long simulations since these extreme events are rare.
-\EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
+The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.
 %
 %
-{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
+{Rare-event sampling refers to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
 %
@@
 %
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.
+In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducted using both very long time-series (direct sampling) and rare-event sampling techniques.
 
@@
 %
-\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
+Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom.
 %
-Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
+Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.
 %
-\EL{Another problem related to rare events is the estimation of parameters or statistics from a limited number of samples.
+Another problem related to rare events is the estimation of parameters or statistics from a limited number of samples.
 Several authors have proposed original strategies  \citep{Mohamad-ws,Blonigan-ws} such as sequential sampling.
-These approaches and rare event algorithms could also be considered for applications in wave-structure problems, ship motion and load acting on offshore platforms \citep{belenky-ws}.}
+These approaches and rare event algorithms could also be considered for applications in wave-structure problems, ship motion and load acting on offshore platforms \citep{belenky-ws}.
 %
@@
 The statistical properties of the drag are then discussed.
-In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
+In section~\ref{sec:direct_sampling}, the phenomenology of  extreme drag fluctuations is investigated based on direct sampling.
 Both the instantaneous drag and time-averaged drag are considered.
@@
 This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
-As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
+As a specific successful application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
 Perspectives and conclusion end this work.

### 2020-04-04 texfiles/section2.tex
Remove \EL{} macros

--- 2020-04-03 section2.tex
@@
 % 
-We believe that this system embeds the characteristic features that makes 
-\EL{this study} both relevant and challenging for fluid-structure-interaction problems.
+We believe that this system embeds the characteristic features that make
+this study both relevant and challenging for fluid-structure-interaction problems.
 
@@
 %
-The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
+The flow dynamics is integrated numerically by the \ac{lbm} \citep{kruger_lattice_2017}. 
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
@@
 	\includegraphics[width=\linewidth]{fig2/fig2}
-	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
+	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4T_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
 		The vorticity is given in lattice units.}
@@
 between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
-Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
+Pressure fluctuations are related to the dynamics of the vorticity field.
 Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
@@
 \label{eq:turnover_time}
-\EL{T_0 = \frac{R}{U}},
+T_0 = \frac{R}{U},
 \end{equation}
@@
 %\sout{Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.}
-\EL{Vorticity is generated along the forebody and eventually carried away by the flow.
-Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.}
+Vorticity is generated along the forebody and eventually carried away by the flow.
+Typical fluctuations of the drag (within one standard deviation) do not result from some preferred arrangement of the vorticity around the obstacle.
 
@@
 	\caption{
-	\EL{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
-The time has been normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.} The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
+	Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow.
+The time has been normalised by the turnover time related to the mean-flow velocity and the size of the obstacle, \emph{i.e.} $T_0=R/U$.
+The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
 If the threshold $a$ is sufficiently high, one observes that $r(a)$ is much larger than the correlation time of the signal.
@@
 % pdf
-Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
+Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4\times 10^6~T_0$.
 This long simulation will be referred to as the \textit{control run} in the following.
@@
 The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
-It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
+It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.} ${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$.
 %
@@
 		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$.
-The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
+The correlation time $\tau_c\simeq 4 T_0$ is given by $C(\tau_c)=0$.
 		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
@@
 Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}.
-It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}, illustrating that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
+It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4T_0$, illustrating that the drag loses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
 %
@@
 In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
-The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}.
-The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \EL{\citep{rodi1998}}.
+The ratio $T_0 / \tau_c$ may be viewed as a {Strouhal number}.
+The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.

### 2020-04-04 texfiles/section3.tex
Remove \EL{} macros

--- 2020-04-03 section3.tex
@@
 	\includegraphics[width=.6\linewidth]{fig5/fig5}
-	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
+	\caption{Amplitude of drag fluctuations as a function of the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
 	}
@@
 The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
-The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
+The mixing time $\tau_m$ is the time needed for the dynamics to lose the memory of its initial condition.
 As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \citep{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
@@
   \label{eq:return_time}
-  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
+  r(a) \underset{a\to\infty}{\propto} \exp(\ell a)
 \end{equation}
@@
 
-\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
+\subsection{Extreme fluctuations of the instantaneous drag}
 \label{sec:instantaneous_drag}
@@
 \end{figure}
-\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
+In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle.
 We shall see in the following that the situation is different in the case of {extreme} drag fluctuations.
-}
 %
@@
 Fig.~\ref{fig:pressure_ratio} displays the relative contributions
-$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
+$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the drag fluctuation $\tilde f_d^{\star}$.
 	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$.
-On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}
+On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).
 
@@
 Interestingly, the profile is also slightly skewed, indicating that the step up of the drag is slower than the return to typical values past the peak value.
-\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
+This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.
 To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.
@@
 Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square.
-This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
+This positive vorticity  prevents the shedding of negative vorticity and enforces the development of an intense vortex against the base of the square.
 As the blocking vortex is in turn advected downstream, the vortex against the base is released.
@@
 This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
-In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.
+In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}.
 
@@
 %\sout{Relevant values of $T$ are application dependant.}
-\EL{In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
-In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).}
+In the following, we shall consider $T=10\tau_c$, where $\tau_c$ is the correlation time of the instantaneous drag.
+In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig:PDF_AVG}).
 
@@
 Setting $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events were sampled.
-\EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
-As a rule of thumb, the threshold has been set so that about one hundred events are sampled.}
+Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics.
+As a rule of thumb, the threshold has been set so that about one hundred events are sampled.
 
@@
 Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
-\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
-Indeed, both cases are %equally 
-featured in Fig.~\ref{fig:extreme_avg};
-very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}
+Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2). 
+Indeed, both cases are featured in Fig.~\ref{fig:extreme_avg};
+very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.

### 2020-04-04 texfiles/section4.tex
Move body of paper into texfiles/ and rename main document to paper.tex
Remove \EL{} macros

--- 2020-04-03 section4.tex
@@
 %% objective
-\EL{The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
+The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
 %
@@
 In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms.
-They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.}
+They both proved to be efficient for various dynamics but adopt opposite strategies: the \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, whereas the \ac{gktl} uses a selection rule based on the increment of a score function.
+A complete description of these two algorithms and their operating principles are provided in appendices \ref{app:ams-short} and \ref{app:gktl_description}.
 
@@
 %
- More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
-\EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}.
+ More recently it has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
+A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}.
 %
@@
 Alongside, numerical methods have been developed to sample rare events \citep{DelMoralBook}. 
-Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for \EL{estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
+Among them the \ac{gktl} algorithm \citep{giardina_direct_2006} is particularly suitable for estimating the probability of observables which are temporal integrals over a very long period \citep{giardina_simulating_2011,Laffargue_2013}.
 %
-Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.}
+Recently the \ac{gktl} algorithm has proven to be extremely efficient to simulate extreme heat waves in a comprehensive climate model \citep{ragone_computation_2018}, with a gain of a factor 100 to 1000 for the computation time.
 This achievement already represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. 
-\EL{The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is an unexplored area of application.}
+The aim of this part of our work is to test rare-event algorithms for fluid-structure interaction in a turbulent flow, which is an unexplored area of application.
 
@@
 
-\EL{The \ac{ams} algorithm is here used with, trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
+The \ac{ams} algorithm is here used with trajectories of fixed duration $T_a$~\citep{lestang_computing_2018}.
 %
@@
 These trajectories are re-sampled by branching them from the remaining trajectories.
-\EL{The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.}
+The operating principle of the \ac{ams} algorithm is detailed and sketched in Appendix \ref{app:ams-short}.
 The objective is here to sample flow evolutions which exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
@@
 Since the Navier--Stokes dynamics is deterministic, small random perturbations are artificially introduced at branching points; this procedure is detailed in Appendix \ref{app:perturb_branching_time}.
-The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.}
-\EL{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
+The chaotic dynamics then ensures that re-sampled trajectories separate from their parents over a time interval $\tau_L$, usually referred to as the Lyapunov timescale.
+Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistics of the sampled rare events.
+This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
 We checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation (see Appendix \ref{app:perturb_branching_time}).
 
-\EL{In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
+In addition to the score function, two important parameters are the number of trajectories $N$ and their duration $T_a$.
   The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories.
 Therefore, $N$ should be taken as large as possible to reduce these errors with, nevertheless, a practical limit given by the available computational resources.
-In this work, we have performed two numerical simulations with $N=32$ and $N=256$.}
-\EL{The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
-In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.}
+In this work, we have performed two numerical simulations with $N=32$ and $N=256$.
+The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
+In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.
 
@@
 In this experiment, the algorithm is used with the instantaneous drag $f_d(t)$ as score function.
-Each trajectory has a duration $T_a = 5\tau_c$ \EL{ where $\tau_c$ is the correlation time of the instantaneous drag}.}
+Each trajectory has a duration $T_a = 5\tau_c$  where $\tau_c$ is the correlation time of the instantaneous drag.}
 \end{figure}
@@
 	\includegraphics[width=.7\linewidth]{fig15/fig15}
-	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit.
+	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding computational cost $C_{AMS}$ in correlation-time unit.
 The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble.}
@@
 %
-\EL{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$.} 
+Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm with $N=32$. 
 %
@@
 However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
-\EL{A phenomenological explanation is developed in the next paragraph.}
+A phenomenological explanation is developed in the next paragraph.
 %
@@
 It takes a time $\tau_L$ (Lyapunov timescale) before a re-sampled trajectory separates from its parent.
-In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space\EL{ or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.}
+In our situation, this ``memory effect'' is related to the fact that the score function is of dimension much smaller (one) than the dimension of the phase space or, in other words, that the score function results from the contribution of a very large number of degrees of freedom.
 %
-\EL{As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
+As shown in section \ref{sec:dynamical_aspects}, extreme drag fluctuations have a lifetime $\tau_c$ related to the timescale over which a vortex remains trapped against the base of the obstacle.
 After $\tau_c$, the vortex is swept away by the flow and further large fluctuations of the drag can only result from the trapping of new vortices.
 %
-Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. }
+Fig.~\ref{fig:AMS_drag_trajectories} shows that $\tau_c$ is shorter than the Lyapunov's timescale $\tau_L$. 
 %
@@
 
-The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \EL{\citep{frisch_book}}.
+The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_L$ may be heuristically associated with the so-called \emph{turbulence rate} \citep{frisch_book}.
 %
-As discussed previously, the duration of extreme fluctuations of the drag \EL{is closely} related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
+As discussed previously, the duration of extreme fluctuations of the drag is closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
 On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
@@
 %
-\EL{This behaviour is independent of the choice of $N$ and $T_a$.
-Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.}
+This behaviour is independent of the choice of $N$ and $T_a$.
+Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially but does not solve the issue of overlapping trajectories.
 
-\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
+\subsection{Extreme time-averaged drag forces with the \acl{gktl} algorithm}
 \label{sec:gktl}
@@
 %
-\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
+Similar to the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
 %
@@
 Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
-\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}
+The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.
 
@@
 The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
-\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.}
+The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged drag}, $F_T$.
 Our observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging, \emph{i.e.} $T_a=T$.    
 %
-\EL{In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
+In a nutshell, trajectories are first evolved in time  independently up to $t=\tau $, with $\tau < T_a$ referring to a cloning period.
 Then, the selection (and cloning) rules apply according to the average of the observable of interest, here $f_d(t)$, over the interval $[0,\tau]$.
@@
 Further details are provided in Appendix~\ref{app:gktl_description}.
-}
 
@@
 % Choice of parameters and perturbation
-\EL{The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
+The algorithm depends on three parameters: the cloning strength $k$, the number of trajectories $N$ and the resampling period $\tau$.
 The higher $k$, the larger is the bias involved in the statistical resampling.
@@
 On the contrary, choosing $\tau$ much higher than  $\tau_c$  results in insufficient cloning steps.
-As a result, a rule of thumb is $\tau \approx \tau_c$.}
+As a result, a rule of thumb is $\tau \approx \tau_c$.
 
-\EL{In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
-We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.}
+In the following experiments, $N=1024$ and $T_a = 10\tau_c$, which yields a computational cost $C_{gktl} = N \times T_a \approx 10^4 \tau_c$.
+We fixed $\tau = \tau_c /2$ as a satisfactory compromise in order to ensure an efficient sampling \citep{lestang:tel-01974316}.
 %
@@
 %
-\EL{Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
-As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.}
+Fig.~\ref{fig:IS_GKTL} shows the histograms of the (zero-mean) time-averaged drag for the three ensembles of trajectories, in addition to the unbiased histogram based on a gaussian approximation of the \ac{pdf} of the time-averaged drag.
+As expected, the algorithm samples preferentially trajectories with a higher value of the averaged drag. Furthermore, the higher $k$, the stronger the bias.
 %
@@
 %
-\EL{In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.}
-\EL{Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
+In the present simulations with $N=1024$ trajectories, one can empirically estimate that  $k$ should be kept smaller than $k_{max} \approx 0.03$ to ensure the independence of the trajectories in the biased ensemble.
+Fig.~\ref{fig:timeseries_extrms_AVG_GKTL} displays the drag signal for the extreme trajectories sampled by the \ac{gktl} algorithm.
 \begin{figure}
@@
 	\label{fig:illustr_extrms_vorticity_GKTL}
-\end{figure}}
+\end{figure}
 
@@
 
-\EL{Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
+Importantly, the \ac{gktl} algorithm provides an ensemble of trajectories over which statistics of rare events can be evaluated.
 In this section, we show in particular that this ensemble allows for the estimate of return times for fluctuation amplitudes that would be unreachable by direct sampling (with the same computational cost).
-The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.}
+The method for estimating return times is described in Appendix~\ref{sec:return-time-gktl}.
 %

### 2020-04-04 texfiles/section5.tex
Remove \EL{} macros

--- 2020-04-03 section5.tex
@@
 \label{conlusion}
-\EL{The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
+The dynamics and statistics of extreme fluctuations of the drag acting on a squared obstacle mounted in a turbulent channel flow (in two dimensions) have been investigated numerically through both direct and rare-event sampling methods.
 By means of direct sampling based on a very long simulation, it was first observed that such extreme events are generically related to the trapping of a strong vortex against the base of the obstacle by a streamline cap.
@@
 Therefore, the lifetime of extreme drag fluctuations is found of the order of the sweeping time of the flow past the obstacle, and the corresponding drag signal is very peaked around extreme fluctuations.
-In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to \EL{yield a large value} of the average; both configurations are observed.
+In addition, it was found that extreme fluctuations of the time-averaged drag do not preferentially result from a small number of very large fluctuations or an exceptional succession of moderate fluctuations that pile up to yield a large value of the average; both configurations are observed.
 A second part of this study has been dedicated to the application of two representative rare-event algorithms, namely the \acl{ams} and the \acl{gktl} algorithms, to our fluid-mechanical problem.
@@
 Our study shows this quite clearly.
-A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.}
+A possible direction would then be to take advantage of recent advances in \emph{learning methods} to dynamically optimise the score function.

### 2020-04-04 figures/effect_of_perturbation/effect_of_perturbation.py
Delete obsolete figure dirs

@@ -1,26 +1,0 @@
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-
-tau_c = 1999
-N = 2
-labels = ["Trajectory {}".format(i) for i in range(N)]
-labels[0] = "Reference trajectory"
-
-fig, ax = plt.subplots(1, 1, figsize=(10, 4), constrained_layout=True)
-
-for i in range(N):
-    dirname = "simulation_{}".format(i)
-    f = np.fromfile(os.path.join(dirname, "dragFile.dat"), float, -1, "")
-    tspan = np.linspace(0, (len(f) - 1) / tau_c, len(f))
-    plt.plot(tspan, f)
-
-plt.legend(labels)
-
-plt.xlabel(r"$t / \tau_c$", fontsize=22)
-plt.ylabel("$f_d$", fontsize=22)
-plt.xticks(fontsize=16)
-plt.yticks(fontsize=16)
-figname = os.path.splitext(__file__)[0] + ".png"
-plt.savefig(figname)
-plt.show()

### 2020-04-04 figures/table_nb_traj/table_nb_traj_fit.py

@@ -1,122 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-from scipy import special,integrate
-
-def ret_gaussian(x):
-    """ Returns zero-mean, unit variance at point
-    """
-    return (1./(np.sqrt(2.*np.pi)))*np.exp(-0.5*x*x)
-
-def integrate_composite_pdf(borneInf, borneSup, pol, norm_factor=1):
-    """Return value of integral of the piecewise function
-         f(x) = f_gauss(x) for x <=3
-                f_fit(x)   for x > 3
-       where f_gauss is the zero-mean, unit variance Gaussian PDF and 
-       f_fit is the exponential of a polynomial pol
-
-    Parameters
-    ----------
-    borneInf : float
-    borneSup : float
-    pol : poly1d
-    norm_factor : float, optional
-
-    Returns
-    -------
-    integral : float
-    """
-    xvec = np.linspace(borneInf,borneSup,1000)
-    dx = xvec[1]-xvec[0]
-    gaussian_part = ret_gaussian(xvec[xvec<=3])
-    non_gaussian_part = np.exp(p(xvec[xvec>3]))
-    full_pdf = np.append(gaussian_part,non_gaussian_part)
-
-    return integrate.trapz(full_pdf,xvec,dx)/norm_factor
-
-Nc = 16384 # Nb of total traj
-NN = 20 # Nb of cloning steps
-sig = 0.0087714 # Std dev of time-averaged drag (T=10\tau_c)
-
-# List containing the names of the result directories
-directories = ['k_0.02_2018_03_26_10_01_46/', 'k_0.025_2018_03_25_20_13_40/',
-	   'k_0.03_2018_03_25_20_13_40/']
-prefix = '/home/tlestang/transition_hdd/thibault/These/tailleur_lecomte/lbm/2_sq/' \
-         + 'grid/illustration_IS/'
-
-# Open a text file
-fo = open("foo.txt", "w")
-for directory in directories:
-    # Read labels and average drag
-    # labels is a Nc*NN vector containing the index of ancestors
-    # for the next evolution step
-    # Ex: Ancestor for traj. 1 after third step is
-    # labels[2*Nc + 1]
-    #          |
-    #    Ancestors for first 2 steps run from 0 to 2*Nc-1   
-    full_path = prefix+directory+'labels.datout'
-    labels = np.fromfile(full_path, np.int32, -1, "")
-    full_path = prefix+directory+'A.datout'
[…]

### 2020-04-04 figures/typical_drag_signal/plot_drag_timeseries.py

@@ -1,31 +1,0 @@
-import numpy as np
-import matplotlib.style
-import matplotlib.pyplot as plt
-
-plt.style.use('ggplot')
-
-path_to_file = "/home/tlestang/transition_hdd/thibault/These/lbm_code/seq/draft/control_run/control_1" \
-               + "/data_force.datout"
-tau_0 = 500 # Turnover time in units of LBM timesteps
-             # The correlation time tau_c is 2000 timesteps, and the
-             # tau_c is 4tau_0
-T = 1000
-f = np.fromfile(path_to_file, float, T*tau_0, "");
-tspan = np.linspace(0,T,len(f))
-
-fig,ax = plt.subplots(1,1,figsize=(12,3), constrained_layout=True)
-ax.set_xlim(0, T)
-line, = ax.plot(tspan, f)
-line.set_linewidth(.7)
-
-ax.set_xlabel(r'$t / \tau_0$',fontsize=22)
-plt.ylabel('$f_d(t)$',fontsize=22)
-#ax.set_yticks(np.arange(-0.10,0.30,0.10))
-plt.xticks(fontsize=16)
-plt.yticks(fontsize=16)
-
-
-fname= 'typical_drag_signal.png'
-plt.savefig(fname)
-
-plt.show()

### 2020-04-04 paper.tex
Modify sentence about github repo in abstract and add graphics extension priorities

@@
 
+\DeclareGraphicsExtensions{.eps,.png}
+
 \title{
@@
 	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
-      This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
+      Software used for simulations and data processing is available at \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
 \end{abstract}

### 2020-05-20 paper.tex
Efforts ---> force

@@
 }
+\DeclareAcronym{pde}{
+	short=PDE,
+	long=Partial Differential Equation,
+}
 
@@
 \title{
-Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
+Numerical study of extreme mechanical force exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
 }
@@
 \begin{abstract}
-	This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
+	This study investigates, by means of numerical simulations, extreme mechanical force exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events.
 	%
-	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
+	The drag experienced by a square obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
 	%

### 2020-05-20 texfiles/section1.tex
Remove spurious emacs lisp in intro
Efforts ---> force

@@
-buffer-name
 \section{Introduction}
@@
 ``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
-In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \citep{kanev2010}.
+In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive force, which can threaten the structural integrity of embedded structures \citep{kanev2010}.
 %
@@
 %
-In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducted using both very long time-series (direct sampling) and rare-event sampling techniques.
+In the present work, a computational study of extreme mechanical force acting on an immersed bluff body is conducted using both very long time-series (direct sampling) and rare-event sampling techniques.
 

### 2020-05-20 texfiles/section2.tex
Efforts ---> force

@@
 The flow is artificially damped near the end of the channel.
-In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square.
+In the developed flow, turbulent eddies have typically the size of the square, which results in strong load fluctuations.
 The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
@@
 While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
-Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
+Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving nonlinear \acp{pde}. 
 %This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level.
@@
 Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
-The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
+The centreline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively.
 The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}).
@@
 %
-The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
+The incoming turbulent flow exerts a fluctuating force onto the squared obstacle.
 The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 

### 2020-05-20 texfiles/section3.tex
Efforts ---> force

@@
 	\includegraphics[width=.7\linewidth]{fig11/fig11}
-	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$.
+	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}+2\tau_c$.
 Each trajectory corresponds to a single event.

### 2020-05-20 texfiles/section4.tex
Efforts ---> force

@@
 In this work, we have performed two numerical simulations with $N=32$ and $N=256$.
-The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational effort.
+The duration $T_a$ should be much larger than $\tau_c$~\citep{lestang_computing_2018} but, again, kept small enough to limit the computational cost.
 In practice, the duration of the trajectories was eventually set to $T_a=5\tau_c$ in both simulations; we checked in particular that larger values of $T_a$ did not improve the results.
@@
 As discussed previously, the duration of extreme fluctuations of the drag is closely related to the sweeping time of the flow past the obstacle, and consequently to the mean-flow velocity $U$.
-On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the fluctuating velocity $u_{rms}$.
+On the contrary, the Lyapunov timescale is rather associated with the intrinsic evolution of turbulent fluctuations in the reference frame of the mean flow, \textit{i.e.} with the root mean square velocity $u_{rms}$.
 The ratio $u_{rms}/U$ (turbulence rate) is much lower than one in our case of grid-generated turbulence, which thus justifies that $\tau_L > \tau_c$.

### 2020-05-20 figures/fig22/fig22.py
Remove reference to TAMS in figure 22

@@ -26,12 +26,12 @@ rtheo = buf[n:]
 
 fig,ax = plt.subplots(figsize=(8,5),constrained_layout=True)
 #ax.plot(xx,amax/sig,marker='o',markersize=1,linestyle='None');
-lineLvl, = ax.plot(xx,Z/sig,linestyle='-',label='TAMS');
+lineLvl, = ax.plot(xx,Z/sig,linestyle='-',label='AMS');
 ax.plot(rtheo,avec/sig, linestyle='-',label='Direct sampling (return time)')
 plt.annotate('', xy=(12584,10), xytext=(1.7e19,10), arrowprops=dict(arrowstyle='<->',lw=2,color=lineLvl.get_color()))
 plt.text(1e7, 10.2, r'$C(a) = 10^{-17}r(a), a=10\sigma$', fontsize=18)
 ax.set_xscale('log')
-ax.set_xlabel(r'$r(x \geq a), C_{TAMS}(x \geq a)$',fontsize=22)
+ax.set_xlabel(r'$r(x \geq a), C_{AMS}(x \geq a)$',fontsize=22)
 ax.set_ylabel('$a$',fontsize=22)
 ax.set_yticks([0,2,4,6,8,10])
 ax.set_xticks([1e2,1e8,1e14,1e20])
